### 2018-12-05 paper/ms.tex
init commit for paper

%\documentclass[preprint2,tighten]{aastex6}
\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{todonotes}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}

\shortauthors{Bouma et al.}
\shorttitle{Decreasing Period of WASP-4\lowercase{b}}

\begin{document}

\title{The Decreasing Orbital Period of WASP-4\lowercase{b}}

\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%-------------------------------------
% TESS Mission Architects:
% These authors should be listed in this order
% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
%-------------------------------------
%\\ \textcolor{red}{6 TESS Mission Architects:}
\author{G. R. Ricker} % grr@space.mit.edu WAITING
\affiliation{
  Department of Physics and Kavli Institute for Astrophysics and Space
  Research, Massachusetts Institute of Technology, Cambridge, MA
  02139, USA }
\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
\affiliation{
  Department of Physics and Kavli Institute for Astrophysics and Space
  Research, Massachusetts Institute of Technology, Cambridge, MA
  02139, USA }
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
\affiliation{
  Harvard-Smithsonian Center for Astrophysics, 60 Garden Street,
  Cambridge, MA 02138, USA }
\author{S. Seager} % seager@mit.edu WAITING
\affiliation{
  Department of Earth, Atmospheric, and Planetary Sciences,
  Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
\affiliation{
  NASA Ames Research Center, Moffett Field, CA 94035, USA }
%-------------------------------------
% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
%These 9 authors should be listed in alphabetical order
%-------------------------------------
%\\ \textcolor{red}{3 TSO coauthors:}
\author{TSO1}
\author{TSO2}
\author{TSO3}
%\\ \textcolor{red}{3 SPOC coauthors:}
\author{SPOC1}
\author{SPOC2}
\author{SPOC3}
%\\ \textcolor{red}{3 POC coauthors:}
\author{POC1}
\author{POC2}
\author{POC3}

\begin{abstract}
  %TODO: final numbers
  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
  with previous studies, we show that the transit times of WASP-4b are
  incompatible with a constant orbital period. 
  In particular, the transits seem to have arrived $77.8 \pm 10.7$
  seconds early, and the period appears to be shrinking by
  $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
  From TESS observations of WASP-6b, WASP-18b, and WASP-46b, we show
  that a systematic offset between the TESS time stamps and the barycentric
  reference sufficient to explain WASP-4b is ruled out at 6.3$\sigma$.
  If the timing variations are astrophysical,
  major contributors to the period change could include either
  apsidal precession, or tidal decay.
  The Doppler shift from WASP-4's acceleration towards us could
  account for at most one quarter of the observed period change
  (at $2\sigma$).
  Further transit and occultation studies will help confirm the
  reality of the timing variation, and eventually determine its cause.
\end{abstract}

\keywords{
  planet-star interactions ---
  planets and satellites: individual (WASP-4b, WASP-5b, WASP-6b,
    WASP-12b, WASP-18b, WASP-46b) ---
  binaries: close
}

\section{Introduction}
\label{sec:intro}

Long-term monitoring of hot Jupiter transit and occultation times
should eventually reveal two distinct processes.  First, whenever the
total spin and orbital angular momenta in a hot Jupiter system are
below a particular critical value, the system will be unstable to
tidal decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most
hot Jupiters satisfy this condition, so their orbits should be
shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate
of decay is uncertain, and depends on how friction dissipates energy
carried by gravitational tides (\citealt{ogilvie_tidal_2014} gives a
review).  If we could directly measure orbital decay rates, it would
inform our understanding of tidal dissipation.  This is important
because the dissipation rates determine the fate of the system.
During its inspiral, if the planet fills its Roche lobe below the
stellar surface, the drag force from the stellar envelope accelerates
the inspiral, leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Otherwise, if the
planet fills its Roche lobe above above the stellar surface, it could
either be dynamically disrupted into an accretion disk, or it could
slowly transfer mass to the star \citep{metzger_optical_2012}.  More
broadly, the outcome of the eventual interaction between the Earth and
red giant Sun also hinges on the efficiency of tidal dissipation
\citep{rasio_tidal_1996}.

The second process that long-term timing studies should reveal is
rotation of the orbital ellipse within the orbital plane (apsidal
precession).  This effect has been studied extensively in eclipsing
binaries \citep[{\it e.g.},][]{russell_notes_1939,
schwarzschild_structure_1958}, and has been explored theoretically for
transiting planets
\citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
However, apsidal precession has yet to be directly observed in hot
Jupiters, and the orbits would need non-zero eccentricity for it to be
detectable.  Though the eccentricities of hot Jupiters are expected to
be small, there are many mechanisms that could pump them away from
circularity (see \S~\ref{sec:apsidal_precession}, and the
possibilities considered by \citealt{bailey_understanding_2019}).  For
hot Jupiters, the apsidal precession is usually dominated by the
quadrupolar distortion of the planet by the star's tidal force
\citep{ragozzine_probing_2009}.  If we measured apsidal precession in
a hot Jupiter, we thus could convert the observed precession rate into
the planet's Love number, giving us a third parameter along with the
mass and radius with which to describe the interior structure of the
planet \citep[{\it e.g.},][who performed a similar procedure for
HAT-P-13b]{batygin_determination_2009}.  Measuring small
eccentricities and understanding their origin would also help us
understand the dynamical histories and interior structures of hot
Jupiters.  \citep[{\it
e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}

% % The HJ semimajor axis distributions and rotation period distribn
% % indicate that tidal decay is real.  (So too, might the age
% % distribution.)
% Stars hosting hot Jupiters tend to spin faster than other stars of
% similar age and mass \citep{penev_hats-18_2016,penev_empirical_2018}.
% This is expected because when the orbital frequency is larger than the
% spin rotation frequency, tides donate angular momentum from the planet
% to the star.
% The reverse process occurs between the Earth and Moon, leading to the
% Moon's slowly increasing separation from Earth of $X.XX$cm~yr$^{-1}$
% (CITE lunar ranging experiment).
% 
% The hot Jupiter star-planet separation distribution is best reproduced
% by models that account for tides~\citep{cameron_hierarchical_2018}.

% There have been some indications of period timing changes for hot jupiters
The most convincing signs seen to date of either orbital decay or apsidal
precession have been in WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
\citet{patra_2017} presented new transit and occultation times for
the system, agreed that curvature was present, and found
$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
% Both \citet{maciejewski_departure_2016} and \citet{patra_2017}
% emphasized that the timing variations could be caused by , or
% also by tidal decay.
Assuming that the period change was caused entirely by tidal decay,
\citet{patra_2017} found values for the stellar tidal quality factor
of $Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
explained by standard equilibrium tide
models~\citep{penev_tidal_2011,ogilvie_tidal_2014}.
\citet{weinberg_tidal_2017} found though that if WASP-12 were a
subgiant, tidal dissipation due to a dynamical tide might be
sufficient to produce the required level of friction.
\cite{bailey_understanding_2019} concurred, but noted that the
observational constraints on WASP-12 itself favor a main-sequence
star, not a subgiant.  Thus, more data are needed to determine the
cause of the timing variations in WASP-12b.

% % There are also indications that apsidal precession should happen. If
% % there are any trace remnants of eccentricity, this will produce
% % timing variations.
% Timing variations can also be caused by rotation of
% the orbital ellipse within the orbital plane (apsidal precession).
% This happens because any rotational or tidal bulge creates a
% gravitation quadrupole field ($r^{-3}$, rather than the typical
% $r^{-2}$), which causes the line of apsides to move forward or back.
% \citet{ragozzine_probing_2009} examined apsidal precession.

% Tidal decay of hot Jupiters has been invoked to explain certain
% properties of the ensemble of star-planet systems. For example, the
% scarcity of gas giants with periods less than a day is suggestive of
% orbital decay \citep[see,
%   e.g.][]{Jackson+2008,Hansen2010,Penev+2012,Ogilvie2014}. The
% anomalously rapid rotation of some hot-Jupiter host stars has been
% attributed to transfer of the planet's orbital angular momentum
% \citep{Penev+2016}. The absence of hot Jupiters around subgiant stars
% may be caused by an acceleration of orbital decay when a star leaves
% the main sequence
% \citep{VillaverLivio2009,Hansen2010,SchlaufmanWinn2013}.  Tidal decay
% might also be responsible for the lower occurrence of close-in planets
% around rapidly rotating stars \citep{TeitlerKonigl2014}, or the
% realignment of stars and their planetary orbits
% \citep{MatsakosKonigl2015}. However, direct evidence for orbital decay
% has been lacking: there have been no clear demonstrations of a
% long-term period decrease due to orbital decay \citep[see,
%   e.g.,][]{Hoyer+2016,Wilkins+2017}.
% 
% Another unfulfilled prediction is that the orbits of hot Jupiters
% should be apsidally precessing on a timescale of decades
% \citep{MiraldaEscude2002,HeylGladman2007,PalKocsis2008,JordanBakos2008},
% as long as the orbits are at least slightly eccentric. In particular,
% \citet{RagozzineWolf2009} noted that the theoretical precession rate
% is dominated by the contribution from the planet's tidally deformed
% mass distribution. They advocated a search for apsidal precession as
% a means of probing the interiors of hot Jupiters.
%

This work focuses on WASP-4b, which is also a plausible candidate for
orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
and mass $1.29 \pm 0.10\ M_{\rm Jup}$
\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
days, orbits its host star at a distance of $a/R_\star = 5.41$, and
has been frequently observed over the past decade
\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
consistent with circular \citep[][but see
\S~\ref{sec:apsidal_precession}]{beerer_secondary_2011,husnoo_observational_2012,bonomo_gaps_2017}.
The host star is a G7V dwarf, with $R_\star = 0.91 \pm 0.12\ R_\odot$,
$M_\star = 0.94 \pm 0.09\ M_\odot$, $T_{\rm eff} = 5436 \pm 34\ {\rm
K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
\citep{southworth_homogeneous_2011,petrucci_no_2013}.

We measure new transit times for WASP-4b using data from TESS
(\S~\ref{sec:transits}).  We then merge these times with previous
observations and compare three models for the times
(\S~\ref{sec:timing}): a constant period, a decaying period, and a
slightly eccentric orbit undergoing precession.  We rule out a
constant period, but cannot firmly distinguish between the latter two
possibilities.  Each would have interesting implications
(\S~\ref{sec:implications}).  We rule out the possibility that a
systematic time system offset causes the observed variation
(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
further monitoring of the system (\S~\ref{sec:future}).

\section{New transit times}
\label{sec:transits}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf TESS transits of WASP-4b.} On the left, black points are TESS
        flux measurements, with a vertical offset applied. Blue curves are
        best-fit models. Rounded midtimes are printed in BTJD
        next to the appropriate transits.  The residuals are shown on
        the right.
        \label{fig:lightcurves}
    }
\end{figure}

\subsection{Observations and cleaning}
WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input
Catalog~\citep[TIC;][]{stassun_TIC_2018} as TIC 402026209.

The TESS spacecraft recorded images of WASP-4 in an $11\times11$
pixel array at a cadence of 2 minutes.  It was included on the
``short-cadence'' target list thanks to the Guest Investigator
programs of J.\ Southworth and S.\ Kane (G011112 and G011183
respectively).  After being downlinked through the Deep Space Network,
the data were processed through the Science Processing Operations
Center (SPOC) pipeline at NASA Ames~\citep{jenkins_tess_2016}.  One
important step in this processing was converting from spacecraft time
to barycentric time.  SPOC then reports timestamps in the ``BTJD''
time system, where ${\rm BTJD} = {\rm BJD} - 2457000$, and all times
are in the TDB reference (see \citealt{urban_explanatory_2012} for
explanations of time systems).  The lightcurves were then vetted
and released by the MIT TESS Science Office to the Mikulski Archive
for Space Telescopes~\citep{ricker_tess_alerts_2018}.

We begin our analysis with the Presearch Data Conditioning (PDC) light
curves from MAST.  The processing steps used to find an optimal
aperture (for WASP-4, a $3\times 3$ pixel array) and the PDC-MAP
algorithm are described by \citet{smith_kepler_apertures_2017} and
\citet{smith_kepler_PDC_2017} respectively.

Our processing steps were then as follows.  First, we removed all
points with non-zero quality flags.  This removes data that is
contaminated by coarse spacecraft pointing, cosmic rays, and other
annoyances \citep{tess_data_product_description_2018}.  We then
filtered out known bad observing windows.  The data closest to
spacecraft perigee typically show ramp-like systematics, so we clipped
out the first and last hour of both orbits.  Next, we focused on the
momentum dumps. As described in the data release
notes\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}},
during the first two TESS sectors, every 2.5~days the spacecraft's
reaction wheels were reset to maintain pointing stability.  These
events were assigned the quality flags corresponding to ``Reaction
Wheel Desaturation Event'' and ``Manual Exclude''.  In the WASP-4
lightcurve, these flags were simultaneously set for 54 distinct
cadences, and there were 10 momentum dumps, averaging about 10 minutes
of flagged data per dump.  Simply out of caution, we clipped out an
additional 10 minutes before and after every momentum dump.  Before
applying any filters, we had 19737 data points. After applying all the
filters, we were left with 18165 measurements, or 92\% of the original
data.

After ``cleaning'' the data, we normalized the flux measurements by
dividing out the median flux.  We then converted the timestamps from
BTJD to BJD by adding the appropriate offset of 2457000
days~\citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018} We opted not to
``flatten'' the lightcurves at this stage, as is often done with {\it e.g.}, a
median filter, spline, or gaussian process regression.  Such
procedures are asymmetric about the transit minimum and can skew the
transit time measurement, and their effect on the measurement
uncertainties is difficult to capture.

\subsection{Measuring the transit times}

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f2.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf TESS saw WASP-4b transit earlier than expected.}
        Points are differences between observed transit times and
        the constant-period ephemeris predicted by previous transit
        measurements.
        Black points with circle markers are the more precise half of
        literatures times; gray points are the less precise half.
        The blue curves represent the $\pm 1\sigma$ credible interval
        around the constant-period ephemeris.
        The binned TESS observation (red point) arrived
        $77.8 \pm 10.7\ {\rm seconds}$ before it was expected.
        \label{fig:arrived_early}
    }
\end{figure}

Using the times, flux measurements, and errors determined above, we
began our measurement procedure by using the Box Least Squares (BLS)
method to estimate the transit epoch, period, and
duration~\citep{kovacs_box-fitting_2002}.
From the BLS parameters, we isolated each transit to a rough estimate
of its midtime, $\pm$10 transit durations.  In each transit window, we
then simultaneously fit for a model transit, plus a local linear
trend.  Our transit model uses the analytic formulae calculated by
\citet{mandel_analytic_2002}, and implemented by
\citet{kreidberg_batman_2015}.

To measure the times, we allowed four free parameters per transit: the
time of mid-transit $t_{\rm tra}$, the planet-to-star radius ratio
$R_{\rm p}/R_\star$, and the slope and intercept of the line. Our
priors were wide uniform distributions centered on the BLS estimates
for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope of zero and
an intercept of unity for the line.

We fixed the remaining transit parameters as follows.  We set the
eccentricity to zero, and the longitude of periastron to $\pi/2$.  For
the limb darkening, we assumed a fixed quadratic law, with
coefficients interpolated from the tables computed
by~\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
inclination, we adopted the values from~\citet{petrucci_no_2013}.
These are in reasonable agreement with the parameters reported by
\citet{gillon_improved_2009}, \citet{southworth_high-precision_2009},
and \citet{huitson_gemini_2017}.

Having defined all the free and fixed parameters, we calculated an
initial guess for the free parameters by maximizing the likelihood.
We then sampled over the posterior using the algorithm proposed
by~\citet{goodman_ensemble_2010}, and implemented
by~\citet{foreman-mackey_emcee_2013}.  During this process, we used
the \texttt{corner} software to check that our posteriors were being
adequately sampled~\citep{corner_2016}.

After performing the initial fit, we set the photometric error bars to
be equal to the standard deviation of the in-transit points of the
residual lightcurve.  We then reperformed the same fitting procedure,
using the empirically determined photometric errors.

% TODO: final numbers
To check that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
freedom.  Examining the residuals by eye showed that the error
variance was being overestimated, so we multiplied the measured TESS
errors by a fudge factor $f=0.73$, forcing a reduced $\chi^2$ of
unity.  This reduced the mean transit time uncertainty from $\sim$30
seconds per transit to $\sim$22 seconds per transit.

Figure~\ref{fig:lightcurves} shows the light curves, best-fit
transit models, and residuals.
Table~1 reports the mid-transit times and
their uncertainties, and also includes the
previously reported times that are analyzed in
\S~\ref{sec:timing}.

Taking the mean and standard deviation of the measured planet to star
radius ratio, we find $R_{\rm p}/R_\star=0.1538\pm0.0013$, in
agreement with the determinations by {\it e.g.},
\citet{wilson_wasp-4b_2008}, \citet{gillon_improved_2009},
\citet{winn_transit_2009}, and \citet{southworth_high-precision_2009}.

Binning the residuals to 1-hour windows and taking the median absolute
deviation, we measure a MAD of the residual lightcurves of $594\,{\rm
ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
\citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
noise\footnote{github.com/lgbouma/tnm} of $410\,{\rm ppm}\,{\rm
hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$
in quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the zodiacal background
model calculated by \citet{winn_photonflux_2013} and used in the
\citet{Sullivan_2015} planet detection simulations.  The predicted
level of zodiacal noise seems to have been an overestimate, given the
observed RMS of WASP-4.  We leave a broader assessment of TESS's
photometric performance to future work.

\section{Timing analysis}
\label{sec:timing}

\subsection{Times}
\label{subsec:times}

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are differences between observed transit times and a linear
        ephemeris fit to the entire timing dataset.
        The constant period model is a poor fit to the observed times
        (gray line).
        Black points are the more precise half of measured times; gray
        points are the less precise half.
        The blue curve is the best-fit model with a constant period
        derivative.
        The orange curve is the best-fit model assuming apsidal precession
        causes the observed timing deviations.
        \label{fig:times}
    }
\end{figure*}

Table~1 shows all the transit times used in our
analysis.  We included data from peer-reviewed literature for which
the analysis was based on observations of a single, complete transit,
and for which the midpoint was fit as a free parameter. We also
required that the time system be clearly documented.

Table~2 shows the occultation times.  Since there are fewer available,
we included all the occultation measurements we could find. The
tabulated values have been corrected for the light-travel time across
the diameter of the orbit by subtracting $2a/c = 22.8$ seconds from
the observed time.

We use the homogeneous \citet{hoyer_tramos_2013} timing
study as our
starting point.  We took their BJD$_{\rm TT}$ times to be equivalent
to times in BJD$_{\rm TDB}$, since the two time systems differ
periodically by only milliseconds \citep{urban_explanatory_2012}.  We
verified that their times, as well as all other times in subsequent
analysis, were on the BJD$_{\rm TDB}$ time standard using the
calculator provided by~\citet{eastman_achieving_2010}.  We set the
zero-point epoch to be as close as possible to the average of the
transit times, weighted as the inverse variance of the transit
midtime. This minimizes the covariance between the transit epoch and
the period \citep{gibson_gemini_2013}.

Important contributions to the transit timing data are as follows.
The earliest epoch in our fit is provided by the EulerCam lightcurve
observed by \citet{wilson_wasp-4b_2008}.  The second, more precise,
epoch comes from $z$-band photometry acquired by
\citet{gillon_improved_2009} at the VLT with FORS2.
Subsequent epochs came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded the GPS clock
  time in UT reference. They then converted this to BJD in the TDB
  reference for their reported midtimes.},
\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011},
and \citet{nikolov_wasp-4b_2012}.
%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS clock
%	time in UT reference. They then converted this to BJD in the TDB
%	reference for their reported midtimes.},
%and reported transit midtimes precise to
%6~seconds.  \citet{dragomir_terms_2011}
%updated the ephemeris using two transits measured with the CTIO 0.9-m
%and 1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
%transits from
%the Baade telescope at Las Campanas\footnote{
%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
%to show that the star's rotation axis is nearly aligned with the
%planet's orbital axis, in agreement with the Rossiter-McLaughlin
%measurement by \citet{triaud_spin-orbit_2010}.}.
%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
%the MPG/EOS-2.2~m at La Silla.
\citet{hoyer_tramos_2013} reported
nine transit observations taken with the the Y4KCAM on the SMARTS 1-m
telescope, and another three with the SOI on the SOAR 4.2-m telescope.
\citet{hoyer_tramos_2013} also homogeneously
redetermined the transit times from all previous studies.  Though
their latest two epochs did fall slightly below the prediction of a
linear ephemeris (Panel C of their Figure~6), they did not find any
statistical need for a quadratic function to fit the O-C values.

After the study by \citet{hoyer_tramos_2013},
\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
using HST's WFC3 over five orbits in transit, and five orbits in
secondary eclipse.  We adopt only their transit time, since the epoch
of occultation was not allowed to be a free parameter in their fit.
\citet{huitson_gemini_2017} acquired optical transmission spectra with
the 8.1-m Gemini South telescope over 2011 to 2014, one transit per
season.  The per-point standard deviation of their lightcurves is a
few hundred parts per million, less than a factor of 3 away from the
photon noise limit.  The uncertainties on their reported transit
times~--~on average 5.6~seconds~--~ are small, but possible given the
quality of their data~\citep[cf.][]{carter_analytic_2008}.

There have been other studies of WASP-4b for which the transit timing
data are either hetereogenous, not available, or not useable.  We
omitted the timing measurements from
\citet{southworth_high-precision_2009}, since there were
technical problems with the computer clock at the time of
observation~\citep{nikolov_wasp-4b_2012}.  The timing study
by~\citet{petrucci_no_2013} ruled out transit timing variations larger
than 54~seconds over the previous 5 years, but did not report transit
times.  \citet{baluev_benchmarking_2015} collected transit times for
many hot Jupiters, including WASP-4b, and included times from the
Exoplanet Transit Database
(ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
\citep{poddany_ETD_2010}.  
Since the ETD data come from heterogeneous sources, their timestamps
are less clearly documented, and the times are thus more prone to
systematic errors, we omit them from consideration.
%We consider the effects of including ETD data in Appendix~\ref{sec:verify_archival_times}.
\citet{may_mopss_2018} obtained optical spectra using IMCAS on
Magellan, and found a flat transmission spectrum, in agreement with
the results reported by \citet{huitson_gemini_2017}.  They did not
report transit times.

There are fewer available occultation times.
\citet{beerer_secondary_2011} observed two occultations of WASP-4b
using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
\citet{caceres_ground-based_2011} detected an occultation from the
ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
specifying the time standard.  We correspondingly added $69.184$
seconds of uncertainty to their reported errors, in quadrature.
Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}
acquired occultation data with HST but kept the epoch as a fixed
parameter.  \citet{zhou_secondary_2015} performed a similar analysis
with occultation data from the Anglo-Australian Telescope, focusing on
the eclipse depth, but also fitting for $e\cos\omega$.  We converted
their $e\cos\omega$ results into a midtime, using their ephemeris and
the standard formula
\citep[{\it e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
\begin{equation}
  t_{\rm occ}(E) = t_0 +  P E  +
  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
  \label{eq:occultation_time}
\end{equation}
This gives us a total of 4 occultations.  Although their timing error
bars are large compared to the transits, we include them for
completeness, and because they can help in our modeling of the timing
variations.

\subsection{Analysis}

After collecting the literature times and measuring the TESS times, we
compared the observed time to the prediction based solely on the
literature times. The result is shown in
Figure~\ref{fig:arrived_early}. The TESS observations appear to have
arrived early.  One immediate concern was systematic errors in either
the archival or the TESS timestamps. The tests we used to mitigate
the latter possibility are described in Appendix~\ref{sec:verify_tess}.

Assuming that the observed timing variation is astrophysical, we
proceeded by exploring three models for the timing data, identical to
the study by~\citet{patra_2017}.

The first model assumes a constant orbital period on a circular orbit:
\begin{align}
  t_{\rm tra}(E) &= t_0 + PE,\\
  t_{\rm occ}(E) &= t_0 + \frac{P}{2} + PE,
\end{align}
for $E$ the epoch number.
The two free parameters are the reference epoch $t_0$ and the period $P$.

The second model assumes a constant period derivative, and a circular
orbit:
\begin{align}
  t_{\rm tra}(E) &=
    t_0 + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2, \\
  t_{\rm occ}(E) &=
    t_0 + \frac{P}{2} + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
\end{align}
The three free parameters are the epoch, period at the reference epoch,
and period derivative, ${\rm d}P/{\rm d}t = (1/P) {\rm d}P/{\rm d}E$.

The third model assumes the planet has a slightly eccentric orbit, and
that the line of apsides is rotating:
\begin{align}
  t_{\rm tra}(E) &= 
		t_0 + P_{\rm s}E
    - \frac{e P_{\rm a}}{\pi} \cos\omega,\\
  t_{\rm occ}(E) &= 
    t_0 + \frac{P_{\rm a}}{2} + P_{\rm s}E
    + \frac{e P_{\rm a}}{\pi} \cos\omega,
\end{align}
for $P_{\rm s}$ the sidereal period, $e$ the eccentricity, $P_{\rm a}$
the anomalistic period, and $\omega$ the argument of pericenter.
Following~\citet{gimenez_revision_1995} and~\citet{patra_2017}, in
this model the angular velocity of the line of apsides
${\rm d}\omega/{\rm d}E$ is constant,
\begin{equation}
  \omega(E) = \omega_0 + \frac{{\rm d}\omega}{{\rm d}E} E,
\end{equation}
and the sidereal and anomalistic periods are related by
\begin{equation}
  P_{\rm s} = P_{\rm a} \left(
    1 - \frac{1}{2\pi}\frac{{\rm d}\omega}{{\rm d}E}
    \right).
\end{equation}
The five free parameters are the epoch, sidereal period,
eccentricity, argument of pericenter at the reference epoch, and
angular velocity of line of apsides:
$(t_0, P_{\rm s}, e, \omega_0, {\rm d}\omega/{\rm d}E)$.

Assuming the error bars on the transit midtimes are gaussian and
independent, we then proceed by fitting each model to the timing
data.
For the linear model, we took uniform priors over small windows around
the expected parameters.
For the quadratic model we did the same, and took the quadratic term
to have a uniform prior corresponding to
$
  Q_\star \sim \mathcal{U}[10^3, 10^9],
$
for $Q_\star$ the ``quality factor'' corresponding to the star's
response against tidal forcing, defined in
\S~\ref{sec:implications}.
For the precession model, we took the same priors in epoch and
sidereal period, and drew the remaining parameters from wide
uniform distributions.

Figure~\ref{fig:times} shows the residuals with respect to the linear
model.
The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of freedom.
The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57 degrees of freedom.
The median MCMC precession fit has $\chi^2 = 51.0$ and 55 degrees of freedom.

%FIXME UPDATE THESE NUMBERS!!

The difference in $\chi^2$ between the linear and quadratic fit
corresponds to $p \approx 10^{-23}$. In other words, assuming the
linear model is true, there is a $10^{-23}$ probability that simply by
chance we would see data that favors the quadratic model more strongly
than the observations.  Clearly, a more likely possibility would be
systematic errors in one of the most important sets of times: either
the earliest times, the Huitson et al.\ Gemini times (epoch numbers
44, 322, 591, and 851), or the TESS times. 
We have taken every precaution against systematic offsets in the
TESS dataset (Appendix~\ref{sec:verify_tess}), and Huitson et al.\ 
stand by their reported times (Huitson, priv. comm.).
Since the linear model provides a poor fit to the data, we discard it
from further consideration.

The quadratic model fits better than the precession model.
It is favored over the precession fit by $\Delta \chi^2 = 3.0$,
and has two fewer free parameters.
Another heuristic useful for model comparison is the Bayesian
Information Criterion (BIC),
\begin{equation}
  {\rm BIC} = \chi^2 + k\log n,
\end{equation}
for $k$ the number of free parameters, and $n$ the number of data
points. For us, $n=60$.
The difference in the BIC for the precession and decay models is
$\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm BIC}_{\rm quad} = 11$,
corrsonding to a Bayes factor of $\approx 4\times10^{4}$.
In the language of~\citet{kass_bayes_1995}, this is ``decisive
evidence'' for the quadratic model describing the data better than the
precession model.
Using the Akaike Information Criterion similarly favors the orbital
decay model, by $\Delta {\rm AIC} = 7$.

Assuming the orbital decay model, we find a period derivative 
\begin{equation}
\dot{P}
    = - (3.82^{+0.38}_{-0.37})\times 10^{-10}
    = - (12.1^{+1.2}_{-1.2})\,{\rm ms}\,{\rm yr}^{-1}.
\end{equation}
For comparison, the period derivative seen by
\citet{maciejewski_departure_2016} and \citet{patra_2017} in
WASP-12b is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
Assuming the quadratic model, WASP-4b is thus decaying at
about half the rate as WASP-12b.

If we instead assume the precession model, the best-fit
eccentricity is $e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}$. The
longitude of periapse correspondingly advances by $\dot{\omega}
= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$.   The
corresponding precession period would be $24^{+8}_{-6}$ years.

The median parameters and their standard deviations are reported for
all the models in Table~3.  In summary, a linear ephemeris does not
describe the observed times.  Both orbital decay and precession can
fit the observed data.  However, the precession model has two extra
free parameters, and is statistically disfavored.  Further
observations will help discriminate between orbital decay and
precession.

\section{Implications}
\label{sec:implications}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=0.415\textwidth]{f4.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf WASP-4b has a short eccentricity
    	damping time and medium decay time compared to other hot Jupiters.
    	It seems to orbit a main sequence host.
    }
    {\bf \it Top:}
    Transiting giant planets from \citet{bonomo_gaps_2017}, who
    measured the eccentricities using radial velocities.  Solid
    squares (including WASP-18b) have significant eccentricity
    detections.  The axes are chosen so that constant eccentricity
    damping timescales (Equation~\ref{eq:de_dt}) are lines.  The
    contours shown assume $Q_{\rm p}' = 10^5$.
    {\bf \it Middle:} 
    Axes are chosen so to that constant orbital decay timescales
    (Equation~\ref{eq:da_dt}) are lines.  The contours shown assume
    $Q_\star' = 10^7$.
    {\bf \it Bottom:}
    HR diagram highlighting WASP-12b, which shows timing variations
    similar to WASP-4b.  We also show hot Jupiters for which we
    detect no timing variation (see \S~\ref{sec:hj_verification}).
    \label{fig:context}
  }
\end{figure}

\subsection{Orbital decay}

The middle panel of Figure~\ref{fig:context} shows the expected
orbital decay time of WASP-4b compared to other transiting giant
planets.  Though there are roughly 20 hot Jupiters that might be
expected to decay faster, some orbit stars with minimal convective zones
({\it e.g.}, WASP-18), and most are not as well-observed as
WASP-4.

Let us assume for the sake of argument that the observed timing
variation is caused entirely by orbital decay.  If the period
decreases at a fixed rate, it will reach zero after
\begin{equation}
  \frac{P}{ \dot{P} } = 9.6 \, {\rm Myr}.
\end{equation}
For comparison, the ``survival time'' found by \citet{patra_2017} for
WASP-12b was 3.2\,Myr.

Let us assume further that the 
``constant phase lag'' model for tidal interaction in
a binary applies \citep{zahn_tidal_1977}.
%, the travelling equilibrium tidal
%bulge is damped by turbulent viscosity in the convective envelope.
The fluid response is parametrized by a constant modified\footnote{For
stars, $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to
explicitly distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
3 Q_\star / (2k_\star)$.  Here $k_\star$ is the stellar Love number,
which is smaller when the star's density distribution is more
centrally concentrated. $Q_\star$ is the ratio of the
energy stored in the equilibrium deformation of the star, divided by
the energy dissipated per tidal period \citep[{\it
e.g.},][]{goldreich_q_1966}.  A larger $Q_\star$ implies less
efficient dissipiation.  Though the fluid mechanics responsible for
the dissipation should depend on the amplitude and frequency of
the tidal perturbation \citep[][Section~3.3]{ogilvie_tidal_2014}, we
will treat $Q_\star'$ as a constant.  This should not be taken
literally; $Q_\star'$ may in fact change sharply with varying tidal
frequency \citep{penev_empirical_2018}.

If the planet's spin and orbit are synchronized, the star is slowly
rotating, and the eccentricity is small, then the semimajor axis and
eccentricity evolve as
\citep[][Appendix B]{metzger_optical_2012}
\begin{align}
  \frac{1}{\tau_{\rm e}} &=
  \frac{|\dot{e}|}{e} =
    \frac{63 \pi } {2 Q_{\rm p}' }
    \left( \frac{R_{\rm p}}{a} \right)^5
    \left( \frac{M_\star}{M_{\rm p}} \right)
    \left( \frac{1}{P} \right)
  \label{eq:de_dt}
  \\
  \frac{1}{\tau_{\rm a}} &=
  \frac{|\dot{a}|}{a} =
    \frac{9 \pi } {Q_\star' }
    \left( \frac{R_\star}{a} \right)^5
    \left( \frac{M_{\rm p}}{M_\star} \right)
    \left( \frac{1}{P} \right).
  \label{eq:da_dt}
\end{align}
The orbital period evolves as
\begin{equation}
\label{eq:dP_dt}
  \dot{P} = -\frac{27\pi}{2 Q_\star'}
            \left(\frac{M_{\rm p}}{M_\star}\right)
            \left(\frac{R_\star}{a}\right)^5.
\end{equation}
Inserting our observed value of $\dot{P}$, the current dissipation
rate corresponds to a modified quality factor for WASP-4 of
\begin{equation}
	Q_\star' \approx 3.3\times10^4. 
\end{equation}
For comparison, Jupiter has $Q_{\rm Jup}' \approx 1.4 \times 10^5$,
based on the motion of the Galilean moons \citep{lainey_strong_2009}.
For stars, different $Q_\star'$ values have been estimated in
different contexts.  Studies of the observed eccentricity distribution
in stellar binaries have yielded $10^5 \gtrsim Q_\star' \gtrsim
10^{7}$ \citep[{\it
e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
geller_direct_2013,milliman_wiyn_2014}.  However the dissipation rates
may be different in hot Jupiter systems, since the tidal forcing
frequencies and amplitudes are different.  In fact, for the observed
population of hot Jupiters, \citet{penev_constraining_2012} argued
that $Q_\star' > 10^7$ would be required at some phase of hot Jupiter
evolution at the 99\% level, or else the observed population would not
exist (though \citealt{birkby_wts-2_2014} point out more efficient
dissipation would be allowed if the hot Jupiters arrived later).
Developing a generative model for the orbital separation distribution
of the hot Jupiter population, \citet{cameron_hierarchical_2018}
similarly found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.

Using an empirical approach, \citet{penev_empirical_2018} reported a
$Q_\star'$ measurement for WASP-4.
In earlier work,
\citet{penev_hats-18b_2016} noted that HATS-18b and WASP-19b orbited
host stars whose isochronal ages are consistent with the Sun, but
whose stellar rotation periods are about 10 days
(\citealt{pont_empirical_2009} pointed out similar evidence for hot Jupiter 
host 
spin-up).  The expected rotation periods at solar ages
are 20 to 30 days
(\citealt{schatzman_theory_1962,skumanich_time_1972}, and measurements
by \citealt{barnes_rotation_2016} in the 4\,Gyr-old M67 cluster).
Assuming that the difference between the observed and predicted
stellar rotation period was due to tides depositing angular momentum
in the star,
\citet{penev_empirical_2018} 
modeled the evolution of hot Jupiter systems under
the influence of a magnetized wind and a constant phase-lag tide.
They found two-sided limits for $Q_\star'$ in 35 of
the 188 systems they examined.  In these 35 systems, they found that
$Q_\star'$ varied from $10^5$ at tidal periods of 2 days to $10^7$ at
tidal periods of 0.5 days.  For WASP-4 specifically, their method gave
$Q_\star' \approx (1.2^{+1.0}_{-0.5})\times10^7$. This is more than 2 orders
of magnitude less efficient than the dissipation rate inferred from
the decaying orbit would imply.

A different model for tidal dissipation was discussed by
\citet{essick_orbital_2016}, who considered the nonlinear response of
the star to a coupled network of gravity modes excited at the base of
the convective zone, and braking near the stellar core.  In their
approach, the stellar tidal quality factors in hot Jupiter systems
vary from $Q_\star' \approx 10^5 - 10^6$.  Using their Equation 26,
the prediction for WASP-4 is that it has $Q_\star' = 7\times10^5$.
This is one order of magnitude larger than would be implied by the
observed period change.

The applicability of the \citet{essick_orbital_2016} model depends on
the evolutionary state of the star.  The bottom panel of
Figure~\ref{fig:context} shows WASP-4 in the context of other giant
planet hosts from the \citet{bonomo_gaps_2017} sample.  On the y-axis
is $G=g-\mu$, for $g$ the apparent Gaia-band magnitude, and $\mu$ the
distance modulus reported by \citet{gaia_collaboration_gaia_2018}.
The x-axis is the effective temperature from \citet{bonomo_gaps_2017},
which for WASP-4 agrees within $1\sigma$ of that reported by
\citet{petrucci_no_2013}.  By visual inspection of the HR diagram,
WASP-4 shows litle evidence of being evolved.  By a similar argument
though, WASP-12 would not clearly stand out as a candidate subgiant,
and \citet{weinberg_tidal_2017} suggest it may in fact be a subgiant
(though \citealt{bailey_understanding_2019} do not find evidence that
supports this suggestion).  A more thorough isochronal analysis would
be of interest, particularly if the transit and occultation times
continue to vary.  Such an analysis is beyond the scope of this study.

In short, if the observed period change is caused entirely by tidal
decay, it would imply a tidal dissipation efficiency stronger than
typical orbit-averaged predictions.  It might be possible that we are
observing at a special time, when the planet is near resonance with
the star's fluid response, and the energy dissipation is exceptionally
large \citep{ogilvie_tidal_2014,essick_orbital_2016}.  Tidal
dissipation rates might also be increased if the star is turning off
the main sequence.  WASP-4 shows little evidence for this, so it is
unclear that tidal decay is a well-motivated explanation for its
changing period.

\subsection{Apsidal precession}
\label{sec:apsidal_precession}

Let us instead assume that the entire timing variation can be
explained as being a portion of an apsidal precession cycle.
The required orbital eccentricity is then
\begin{equation}
  e = (1.6^{+1.5}_{-0.7})\times10^{-3},
\end{equation}
and the line of apsides is advancing by
\begin{equation}
  \dot{\omega} = 14.8^{+5.4}_{-3.8}\ {\rm degrees}\,{\rm yr}^{-1}.
\end{equation}
One full precession period would thus take roughly 24 years, about
twice the current observing baseline.

\citet{ragozzine_probing_2009} calculated apsidal precession periods
for hot Jupiters ranging between of order 10 and 100 years.  They
highlighted that for many hot Jupiters, including WASP-4b, over 90\%
of the apsidal precession comes from the tidal bulge of the planet.
Precession from general relativity, the planet's rotational bulge, and
the star's rotational and tidal bulges are all minor contributors.
\citet{ragozzine_probing_2009} also calcuated that WASP-4b was
expected to have a relatively short precession period, of 120 years.

One implication of the study by \citet{ragozzine_probing_2009} is that
if the planet has non-zero eccentricity, a measurement of
$\dot{\omega}$ can be converted into a measurement of the planetary
Love number, $k_{2,\rm p}$.  Following \citet{ragozzine_probing_2009}
Equation 14, the implied planetary Love number for WASP-4b is
\begin{equation}
k_{2,{\rm p}} = 1.6^{+2.1}_{-1.2}.
\end{equation}
For comparison, $k_{2,{\rm Jup}} \approx 0.59$
(\citealt{wahl_tidal_2016}, though \citealt{ni_empirical_2018} give an
estimate that is 10\% lower).  A uniform density sphere has $k_2 =
1.5$.  Our reported uncertainties on $k_{2,{\rm p}}$ are large because
the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$ are
quite correlated, and the measured occultation times only barely
constrain them (Figure~\ref{fig:future}).

\subsubsection{Empirical constraints on small eccentricities in hot
Jupiters}

%FIXME
Could the eccentricity truly be non-zero?  The classical
tidal-circularization theory has the eccentricity shrinking with a
$1/e$-folding time of \citep[{\it e.g.}, Equation~1
of][]{dobbs-dixon_spin-orbit_2004}
\begin{equation}
  \tau_e
  = \frac{e}{|\dot{e}|}
  = \frac{2\pi Q_{\rm p}'}{63}
  \left( \frac{a}{R_{\rm p}} \right)^5
  \frac{M_{\rm p}}{M_\star}\, P .
\end{equation}
For WASP-4b, this gives an eccentricity damping timescale of $\tau_e =
28.3 (Q_{\rm p}'/10^6) \, {\rm Myr}$.  The system age appears to be
well over $1\,{\rm Gyr}$ \citep{winn_transit_2009}, so any initial
eccentricity should have been erased.

The top panel of Figure~\ref{fig:context} compares the expected
eccentricity damping time of WASP-4b with other transiting giant
planets.  WASP-4b has one of the shortest known eccentricity damping
times.  The hot Jupiter with the nearest $\tau_e$ and a significant
measured eccentricity is WASP-18b, with $e = 0.0076 \pm 0.0010$
\citep{triaud_spin-orbit_2010,bonomo_gaps_2017}.  This is the smallest
significant eccentricity measured by the RV method, with the
next-smallest cases being HAT-P-20b ($e =
0.01580^{+0.0023}_{-0.0018}$), and WASP-38b ($e =
0.02780^{+0.003}_{-0.0028}$).  For the 231 planets studied by
\citet{bonomo_gaps_2017}, 2 (13) had $2\sigma$ upper limits of $e <
0.01$ ($e<0.02$).  
For WASP-4b, the strongest $2\sigma$ upper-limit on its eccentricity
from RVs is $e<0.011$
\citep{husnoo_observational_2012,bonomo_gaps_2017}.

Occultation timing can provide stronger eccentricity limits than
RV measurements, but they often come with caveats.
For WASP-4b, the strongest limit on its eccentricity comes from the
secondary eclipses observed by \citet{beerer_secondary_2011}, which
led to a $2\sigma$ upper limit on $| e\cos\omega | $ of 0.0024.
% \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
% 0.0018$. 
% \citet{zhou_secondary_2015} report a weaker $e\cos\omega = -0.001
% \pm 0.003$. 
Placing eccentricity limits stronger than $e< 10^{-3}$
has only been possible for a few hot Jupiters, and often with numerous
false-starts.  HD 189733b, for example, was initially reported to have
a very small eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
\citep{knutson_map_2007}.  This result was retracted by
\citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
eclipses and found a smaller average offset in the secondary eclipse
time than \citet{knutson_map_2007}.  \citet{agol_climate_2010} also
noted that a larger fraction of the planet's observed occultation time
offset could be explained by detailed modeling of the planetary
hot-spot, which can change occultation time offsets by up to a few
minutes \citep{williams_resolving_2006}.  Including an updated
hot-spot model, \citet{agol_climate_2010} reported a 2$\sigma$ limit
of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit
yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
\citet{winn_measurement_2005} combined Spitzer secondary eclipses with
RV data and reported a best-fit eccentricity of $e>0.0057$ at 90\%
confidence.  They suspected that the data were consistent with zero
eccentricity, and that the lower bound on $e$ was an artifact of
imperfect limb-darkening model.  Stronger limits were eventually put
in place by \citet{crossfield_spitzer_mips_2012}, who found
$e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b,
for which the discovery RVs and photometry led to a non-zero but small
eccentricity measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.
A prompt ground-based occultation measurement by
\citet{lopez-morales_day-side_2010} concurred with weak statistical
significance that a small eccentricity was favored, but
further follow-up found occultation times consistent with circular
orbits; $| e\cos\omega| < 0.0040 $
\citep{campo_orbit_2011,croll_near-infrared_2011}.

In summary, empirical constraints on whether hot Jupiters are
circular at the $e \lesssim 10^{-3}$ level are only present for HD
189733b and HD 209458b, from occultation timing.  Once the timing
offsets are of order tens of seconds, eccentricities from occultation
timing may have systematic uncertainties because of the hot-spot issue
\citep{williams_resolving_2006,agol_climate_2010}.  With the sole
exception of WASP-18b, radial velocity measurements have provided
eccentricity constraints only at the $e > 10^{-2}$ level.  Thus,
though empirical constraints on the shortest period hot Jupiters are
typically consistent with circular orbits, they cannot rule out
eccentricities at the level needed to contextualize WASP-4.

\subsubsection{Theoretical expectations for small eccentricities in hot
Jupiters}

For lack of empirical data, we now turn to theoretical mechanisms
that could maintain small eccentricities.

\paragraph{Neighboring companion}
\citet{mardling_long-term_2007} considered the long-term tidal
evolution of hot Jupiters with companions.  They pointed out that
although the early phases of the two-planet eccentricity evolution
would typically occur over millions of years, the final phase of the
joint eccentricity evolution towards circularity would often occur on
timescales several orders of magnitude longer than the circularization
time (see their Figures~4 and 5).  The companion in their model is
coplanar, and can have a mass down to an Earth-mass; the main
requirement is that both the hot Jupiter and the outer companion start
on eccentric orbits.

Another mechanism to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the companion would need to be
inclined relative to that of the hot Jupiter by at least $\sin^{-1}
\sqrt{2/5} \approx 39^\circ$.
\citet{bailey_understanding_2019} point out how the joint requirements
of Kozai-Lidov oscillations with a non-detection of an additional RV
signal can be combined to produce lower and upper limits on the outer
companion's mass and the semimajor axis ratio, $a_{\rm c}/a_{\rm b}$.
For WASP-4, ...
%TODO: check this mechanism
%FIXME: take a look at RVs. rule this shit out. /  understand where
%the bounds are.

\paragraph{Fluctuations in the gravitational potential from convection}
A separate mechanism that might produce the eccentricity is simply stellar
convection \citep[][Section 7]{phinney_pulsars_1992}.
Phinney presents this idea by analogy with a pendulum, which is never really at
rest. Instead, collisions with air molecules randomly perturb its momentum and
energy, and the equipartition theorem of statistical mechanics tells
us the average energy is $\langle E_{\rm pend} \rangle = k
T_{\rm air}$. 
In hot Jupiter systems, the ``air molecules'' are convective eddies, and
their associated kinetic energy leads the orbit to never have exactly zero
eccentricity.
In particular, from \citealt{phinney_pulsars_1992} Equation (7.33),
\begin{equation}
  \langle e^2 \rangle =
  \frac{ 2 \langle E_{\rm e} \rangle }{\mu n^2 a^2}
  = 6.8\times10^{-5}
  \frac{(L^2 R_{\rm conv}^2 M_{\rm conv}^2)^{1/3}}{\mu n^2 a^2},
\end{equation}
for $L$ the stellar luminosity, $R_{\rm conv}$ ($M_{\rm conv}$) the
width (mass) of the convective region,  $\mu$ the reduced mass, $n$
the orbital frequency, and $a$ the semimajor axis.
For WASP-4, convection in the star is more important than
convection in the planet, and
$\langle e^2 \rangle^{1/2} < 10^{-5}$.

\subsection{Applegate effect}
Some eclipsing binaries exhibit period modulations with amplitudes of 
$\lesssim 0.05\,{\rm days}$ over timescales of decades \citep[{\it
e.g.},][]{soderhjelm_geometry_1980,hall_relation_1989}.
\citet{applegate_mechanism_1992} proposed a mechanism to explain these
modulations in which the internal structure of a magnetically active star
changes shape via cyclic exchange of angular momentum between the inner and
outer zones of the star.
This model could also apply to a hot Jupiter orbiting a star with a
convective zone.
The changing gravitational quadrupole of the star would cause the orbit of the
planet to precess on the timescale of the stellar activity cycle
("the dynamo timescale").
An essential difference between this process and apsidal precession is
that it is expected to be quasi-periodic \citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}.
The transit and occultation deviations would also have the same sign, while
for apsidal precession they have opposite signs.
For WASP-4, \citet{watson_orbital_2010} estimated that the effect
could produce timing deviations of up to 15 seconds, depending on the
modulation period of the stellar dynamo, and the corresponding level
of differential surface shear.
This cannot explain the majority our observed $78$ second variation.

\subsection{Other possible explanations}

Another mechanism to produce the observed period change would be if an
unseen companion in the WASP-4 system accelerated the star towards us.
This would produce a period derivative
\begin{equation}
	\dot{P} \approx \frac{\dot{v}_{\rm r} P}{c},
\end{equation}
for $\dot{v}_{\rm r}$ the time derivative of the radial velocity.
Recently, \citet{bonomo_gaps_2017} collected all available HARPS data 
for the system, and reported no detectable slope at $3\sigma$.
However, \citet{knutson_friends_2014} collected 5 Keck HIRES radial velocity
measurements for WASP-4, and merged them with measurements from
CORALIE and HARPS reported by \citet{wilson_wasp-4b_2008},
\citet{pont_determining_2011}, and \citet{husnoo_observational_2012}.
Using these four datasets,
\citet{knutson_friends_2014} found that $\dot{\gamma} =
-0.0099^{+0.0052}_{-0.0054}$, corresponding to a $2\sigma$ limit on
the acceleration towards us of $\dot{\gamma} > - 0.021 \,{\rm m}\,{\rm
s}^{-1}\,{\rm day}^{-1}$.  This limits $\dot{P}$ to be less than
$9\times 10^{-11}$ at $2\sigma$.  
%FIXME TODO final value
Our measured value is $\dot{P} =
-(3.9^{+0.7}_{-0.5})\times 10^{-10}$.  We can
thus rule out radial acceleration as the sole cause of the observed period
change.  However, it could cause up to one-quarter of the observed
period decrease.  Improved long-term radial velocity monitoring with
the same instruments will help clarify the situation, particularly
given the fact that short-term trends of roughly $\dot{\gamma} \approx
2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ are seen over timescales
of order 10 days in WASP-4~\citep{husnoo_observational_2012}.
Given the known
spottedness of the system~\citep{sanchis-ojeda_starspots_2011}, these
short-timescale variations may also have a contribution from stellar
activity.

There are two other small effects worth briefly mentioning.  The
first is the \citet{shklovskii_possible_1970} effect, in which the
star's proper motion leads to a changing radial velocity.  This would be
apparent in the Doppler measurements described above.  The effect
contributes a change of order $\dot{P} \sim P\mu^2 d/ c \sim
6\times10^{-13}$.  The second effect, described by
\citet{rafikov_stellar_2009}, comes from the star's on-sky motion
altering our viewing angle, and leads to an observed apsidal precession.
The corresponding change is $\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$.

Many of these phenomena are likely occurring simultaneously.  From the
preceding discussion, we found that radial acceleration, proper-motion,
and the Applegate mechanism
are unlikely to cause the timing deviations.
That leaves tidal decay and apsidal
precession as the possible causes.

\section{Future prospects}
\label{sec:future}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
  \vspace{-0.5cm}
	\caption{
		{\bf Future evolution possibilies for WASP-4b.}
		Dots are as in Figure~\ref{fig:times}.
		Lines are 100 random samples from the posteriors of the apsidal
		precession model (orange), and the orbital decay model (blue).
		In an extended mission, TESS might re-observe WASP-4b in the early
		2020s. While useful to confirm that the period is
		in fact changing, precise long-term monitoring past 2025 will
		be needed to resolve the two models.
		\label{fig:future}
	}
\end{figure}

If TESS continues observing after
its primary mission, 
it will observe additional transits of WASP-4b in the
early 2020s (Figure~\ref{fig:future}).
In order to rule between orbital decay and
apsidal precession, occultation measurements in the
near term, and also in the mid-2020s, will be helpful.
However it may take up to a decade to firmly distinguish
the two models, through the combination of transit and occultation
measurements.
These observations should be carried out, because they will
lead to new knowledge; either a
measure of the efficiency of tidal dissipation within the
star, or the eccentricity and Love number of the planet.
Measuring either for a single system would be
rather interesting.
In the coming decade, wide-field photometric surveys
(TESS, HATPI, NGTS, PLATO)
might even enable such measurements at a population
level, which would be a good deal more interesting.

%Detailed modeling of the star's interior
% structure and and tidal evolution is warranted, as are continued
% observations of transits and occultations.
%

\acknowledgements{
L.G.B.\ acknowledges helpful discussions with F. Dai and V. Van Eylen. 
%
We acknowledge the use of TESS Alert data, which is currently in a beta test
phase, from the TESS Science Office. Funding for the TESS mission is provided
by NASA's Science Mission directorate. This research has made use of the NASA
Exoplanet Archive, which is operated by the California Institute of Technology,
under contract with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.  This work made use of NASA's Astrophysics Data
System Bibliographic Services.
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  }

\bibliographystyle{yahapj}                            
\bibliography{bibliography}

%\clearpage

%\newpage
\input{WASP-4b_transit_time_table.tex}

%\clearpage
%\newpage
\input{occultation_time_table.tex}

%\clearpage
%\newpage
\input{model_fit_table.tex}
\clearpage
\newpage

\appendix

\section{Verifying the TESS Time Stamps}
\label{sec:verify_tess}

Any systematic offset between the reported TESS
times and the true barycentric reference would cast doubt on the
results of this study.  Such an offset would have historic precedent:
Kepler had a systematic error in its timestamps that was corrected
only in Q14~\citep[][Section 3.4]{kepler_DR19_2013}.

We devised two checks on the absolute calibration of the TESS time
system.  First, in \S~\ref{sec:headers}, we recalculate and confirm
the barycentric correction reported by SPOC in the TESS lightcurve
file headers.  In \S~\ref{sec:hj_verification}, we repeat the main
analysis of the paper for a collection of other hot Jupiters, and use
their observed transit times to rule out a global TESS time offset.
The latter test confirms that WASP-4b is an outlier.

\subsection{Using the headers of the lightcurve files}
\label{sec:headers}

The TESS lightcurve files provide observation start and end times in
three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
TDB}$, and ${\rm BJD}_{\rm TDB}$.
First, we verified for a few select lightcurve files that ${\rm
JD}_{\rm UTC}$ lagged behind ${\rm JD}_{\rm TDB}$ by the expected
$32.184 + N\,{\rm seconds}$, where $N$ is the number of leap-seconds
since 1961. For the relevant observation time, $N=37$, and the offset
was as expected.
Then, using the ${\rm JD}_{\rm UTC}$ timestamp, we recalculated the
barycentric correction computed by SPOC, using the
\citealt{eastman_achieving_2010} calculator.  Due to the first
author's ignorance of the spacecraft's position, we performed this
calculation assuming that the observer was located at the Earth's
geocenter, and used the correct direction for each star.  This gave us
times in ${\rm BJD}_{\rm TDB}$ that agreed with the archival times to
within 1.7 seconds.  This offset is comparable to the light-travel
delay time expected for TESS on its orbit from perigee of $\approx
20R_\oplus$ to apogee of $\approx 60R_\oplus$.

This calculation bounds any error in the SPOC barycentric julian date
correction to be less than 1.7 seconds.  Since this is smaller than
the effect of interest, we treat the BJD correction as ``verified'',
and proceed to a subsequent test.

\subsection{Using other hot Jupiters as references}
\label{sec:hj_verification}

\begin{figure*}[ht!]
  \begin{center}
    \leavevmode
    \includegraphics[width=0.9\textwidth]{f6.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf There is no evidence for a systematic offset between TESS
    times and the barycentric reference.}
    While the WASP-4b transits fell about 75 seconds earlier than
    predicted, other well-observed hot Jupiters, in particular WASP-6b
    and WASP-18b, arrived on time.  Ticks are observed TESS transit
    midtimes; the blue distribution function is a kernel density
    estimate; the orange distribution function is a gaussian centered
    on zero using the indicated $1\sigma$ standard deviation in the
    prediction
    ($\sigma_{\rm predicted}$).
    \label{fig:hjs}
  }
\end{figure*}

If the observed timing delay in WASP-4b were caused by a systematic
timing system offset between the TESS ${\rm BJD}_{\rm TDB}$ times and
the ${\rm BJD}_{\rm TDB}$ reference, we would expect that it might
apply to other hot Jupiters as well.  This test can rule out an
important class of clock error~--~a systematic global offset.

To rule out this possibility, we repeat the timing analysis of the
main paper, for other hot Jupiters with long preceding observing
baselines.  We first checked which hot Jupiters were observed over the
first two TESS sectors using a combination of
\texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
TEPCat \citep{southworth_homogeneous_2011}.  We then selected hot
Jupiters for which there were at least five distinct epochs reported
in the peer-reviewed literature.  We required that each observation be
of a single transit, that the midpoint be fit as a free parameter, and
that the time system be clearly documented.  Our final sample of
hot Jupiters included WASPs-4b, -5b, -6b, -18b, and -46b.  The
collected and measured times are given in Tables~4, 5, 6, and 7
for each.

Using the literature timing data for each hot Jupiter, we then
performed a least-squares fit to a linear ephemeris.  Using the
best-fit values and variances, we calculated the uncertainty on the
predicted transit time during the TESS observations.  This gave 9, 94,
18, 42, and 60 seconds for WASPs-4b, -5b, -6b, -18b, and -46b.  If a
substantial portion of the observed timing deviation in WASP-4b (about
77 seconds) were from a systematic offset in the time systems, this
offset would be present in the other hot Jupiter transit times as
well.  We show in Figure~\ref{fig:hjs} that WASP-4b is the only hot
Jupiter that transited significantly earlier than expected.

To convert this intuition into a quantitative limit, for each hot
Jupiter we considered the model
\begin{equation}
  t_{\rm tra}(E) = t_0 + PE + t_{\rm offset},
\end{equation}
for $t_{\rm offset}$ a systematic constant offset between the reported
timestamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
were
\begin{align}
  t_0 &\sim \mathcal{N}[t_0', \sigma_{t_0'}], \\
  P &\sim \mathcal{N}[P', \sigma_{P'}], \\
  t_{\rm offset} &\sim \mathcal{U}[-20\sigma_{t_0'},20\sigma_{t_0'}],
\end{align}
where $\mathcal{N}$ and $\mathcal{U}$ denote a normal and uniform
distribution, $(t_0', P')$ are the best-fit reference time and period
using only the literature transit times, and $(\sigma_{t_0'},
\sigma_{P'})$ are their uncertainties.

For each planet, we then ask: what fraction of the posterior for
$t_{\rm offset}$ is consistent with an offset worse than
$77$ seconds?  For WASP-4b, the answer is unsurpringly about half.
For WASP-6b, the most constraining object, about 1 sample in 2 million
is consistent with such a timing offset ($4.9\sigma$).  For WASP-18b,
1 in 63 samples would be consistent with this timing offset
($2.1\sigma$), and in WASP-46b, the limit is 1 in 38 samples
($1.9\sigma$).  For WASP-5b, the predicted time is too imprecise to
rule out timing offsets at the necessary amplitude.  Multiplying the
three independent probabilities for WASPs-5b, 6b, and -18b, we rule
out $t_{\rm offset} < -77\ {\rm seconds}$ at $6.3\sigma$, or about
about 1 part in 5 billion.

%\clearpage
%\newpage
\input{WASP-5b_transit_time_table.tex}
\input{WASP-6b_transit_time_table.tex}
\input{WASP-18b_transit_time_table.tex}
\input{WASP-46b_transit_time_table.tex}

% \section{Verifying the WASP-4 Archival Times}
% \label{sec:verify_archival_times}
% 
% \begin{figure*}[ht!]
%   \begin{center}
%     \leavevmode
%     \includegraphics[width=0.9\textwidth]{f7.pdf}
%   \end{center}
%   \caption{
%     {\bf Extra times for WASP-4b neither confirm nor refute the
%     	period-decay.}
%     Our timing analysis (Section~\ref{sec:timing}) imposes stringent
%     selection criteria for lightcurves.
%     Here we overplot additional points from the Exoplanet Transit
%     Database (ETD), and the discovery epoch from \citet{wilson_wasp-4b_2008}.
%     \label{fig:verify}
%   }
% \end{figure*}
% 
% The preceding analysis shows that the TESS timestamps are consistent
% at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
% Another explanation for the WASP-4b timing variation could be
% incorrect archival times.
% As a sanity check on the times we included for our analysis in
% Table~1, we collected times from the Exoplanet Transit Database
% \citep[ETD; ][]{poddany_ETD_2010}.  We considered only the 12
% lightcurves of the highest self-reported quality (${\rm DQ} = 1$).  We
% then inspected each lightcurve by eye.  If the lightcurve showed
% substantial red noise, we discarded it.  This left 9 midtimes,
% reported in ${\rm HJD}_{\rm UTC}$. We converted the times to ${\rm
% BJD}_{\rm TDB}$ \citep{eastman_achieving_2010}.
% 
% The resulting times are shown in Figure \ref{fig:verify}.  Focusing on
% the epochs that overlap with \citet{huitson_gemini_2017}'s study, four
% of six possible times agree with the \citet{huitson_gemini_2017}
% times; the other two fall about 90 seconds early.  With their stated
% uncertainties, these 6 ETD data points are inconsistent with a
% constant period.  The 6 points came from 5 different observers.
% Systematic errors in absolute clock offsets by different observers
% should be uncorrelated; if there were such systematic errors in the
% ETD times, this could explain the many-sigma outliers.  Though the
% majority of the reported times do support the accuracy of the
% \citet{huitson_gemini_2017} times, a systematic procedure would call
% for averaging the ETD times.  If we did this, and discarded the four
% \citet{huitson_gemini_2017} epochs, it is likely that the evidence for
% period decay would become much weaker.  Since
% \citet{huitson_gemini_2017} documented their time system clearly, and
% produced lightcurves of extremely high quality, we included their times
% in our analysis.  From private correspondence with
% \citealt{huitson_gemini_2017}, the authors also stand by their
% timestamps\todo[inline]{verify this}. Since the ETD data come from
% heterogeneous sources, their timestamps are less clearly documented,
% and the times are thus more prone to systematic errors, we
% categorically omit them from consideration.
% 
% We also overplotted the discovery epoch reported by
% \citet{wilson_wasp-4b_2008}. 
% This epoch folded together WASP-S survey data, a partial FTS
% lightcurve, and a complete EulerCam lightcurve. 
% Though not explicitly stated by \citet{wilson_wasp-4b_2008}, the
% archival WASP-S lightcurves are in ${\rm HJD}_{\rm UTC}$, so they can
% be converted to ${\rm BJD}_{\rm TDB}$ without uncertainty in the
% absolute time reference (Collier-Cameron, priv.\ comm.).
% The discovery epoch falls about 2 minutes earlier than the epochs we
% used for the fit, and is visible in the lower left of Figure
% \ref{fig:verify}.

% Another independent line of evidence supporting the
% \citet{huitson_gemini_2017} times is that \citet{patra_2017} showed
% for WASP-12b that precise transit times reported by
% \citet{stevenson_transmission_2014} from the Gemini telescopes with
% GMOS are congruent with those from other observers.  This indicates
% that the telescopes do not have a long-standing clock offset.

\end{document}

### 2018-12-05 paper/stellar_parameters.tex
init commit for paper

%\renewcommand{\arraystretch}{1.0}

\startlongtable
\begin{deluxetable*}{ccc}

%\tabletypesize{\small}

\tablenum{4}

%\tablewidth{0pt}

\tablecaption{Adopted stellar and planetary parameters.}
\label{tab:stellar_parameters}

\tablehead{
\colhead{Parameter} & \colhead{Value} & \colhead{Uncertainty} 
}

\startdata
Period $P$(days) & 1.33823251 & 0.00000031 \tablenotemark{a}  \\
Minimum reference Time $T_{0}$($BJD_{TDB}$) & 2454697.797973 & 0.000076    \\
Inclination $i$($\degr$) & 86.85  & 1.76   \\
Stellar Radius $R_{\star}$($R_{\sun}$) & 0.92 & 0.06   \\ 
Stellar Mass $M_{\star}$($M_{\sun}$) & 0.89  &  0.01   \\
Stellar gravity $\log g_{\star}$($cm/s$) & 4.461 &  0.054 \\
Semimajor axis $a$($UA$) & 0.0228  & 0.00013   \\
Age (Gyr) & 7.0 & 2.9   \\  
Stellar effective temperature $T_{eff}$(K) & 5436 & 34  \\
Metallicity $[Fe/H]$(dex) & -0.05 & 0.04   \\
Planet Radius $R_{P}$($R_{Jup}$) & 1.33  &  0.16  \\
Planet Mass $M_{P}$($M_{Jup}$) & 1.216 &   0.013  \\
Planet surface gravity $g_{P}$($m/s^{2}$) & 16.41 &  2.49 \\
Planet equilibrium temperature $T'_{eq}$(K) & 1664 & 54   \\
\enddata
\tablecomments{
    $t_{\rm tra}$ is the measured transit midtime, and $\sigma_{t_{\rm tra}}$ is its
    $1\sigma$ uncertainty.
    $\sigma_{t_0}$ was evaluated from the sampled posteriors by taking
    the maximum of the difference between the 84th percentile
    minus the median, and the median minus the 16th percentile.
    The ``Reference'' column refers to the work describing the
    original observations.
    The ``H13?'' column is 1 if the mid-time value was taken from 
    \citet{hoyer_tramos_2013}.  Otherwise, the mid-time
    came from the column listed in ``Reference''.
}
\tablenotetext{a}{
  %FIXME cite etc.
The numbers in parenthesis give the 1$\sigma$ uncertainty in the final
two digits, where appropriate.
}

\end{deluxetable*}

### 2018-12-05 paper/ms.tex
cleanup TeX

@@
 
-\correspondingauthor{L. Bouma}
+\correspondingauthor{L. G. Bouma}
 \email{luke@astro.princeton.edu}
@@
 should eventually reveal two distinct processes.  First, whenever the
-total spin and orbital angular momenta in a hot Jupiter system are
+summed spin and orbital angular momenta in a hot Jupiter system are
 below a particular critical value, the system will be unstable to
@@
 the inspiral, leading to a `direct-impact' merger
-\citep{metzger_optical_2012,macleod_planetary_2018} Otherwise, if the
-planet fills its Roche lobe above above the stellar surface, it could
-either be dynamically disrupted into an accretion disk, or it could
-slowly transfer mass to the star \citep{metzger_optical_2012}.  More
-broadly, the outcome of the eventual interaction between the Earth and
-red giant Sun also hinges on the efficiency of tidal dissipation
-\citep{rasio_tidal_1996}.
+\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
+the planet fills its Roche lobe above above the stellar surface, then
+it could either be dynamically disrupted into an accretion disk or it
+could slowly transfer mass to the star \citep{metzger_optical_2012}.
+More broadly, the outcome of the eventual interaction between the
+Earth and red giant Sun also hinges on the efficiency of tidal
+dissipation \citep{rasio_tidal_1996}.
 
@@
 
-% % The HJ semimajor axis distributions and rotation period distribn
-% % indicate that tidal decay is real.  (So too, might the age
-% % distribution.)
 % Stars hosting hot Jupiters tend to spin faster than other stars of
-% similar age and mass \citep{penev_hats-18_2016,penev_empirical_2018}.
-% This is expected because when the orbital frequency is larger than the
-% spin rotation frequency, tides donate angular momentum from the planet
-% to the star.
-% The reverse process occurs between the Earth and Moon, leading to the
-% Moon's slowly increasing separation from Earth of $X.XX$cm~yr$^{-1}$
-% (CITE lunar ranging experiment).
-% 
-% The hot Jupiter star-planet separation distribution is best reproduced
-% by models that account for tides~\citep{cameron_hierarchical_2018}.
-
-% There have been some indications of period timing changes for hot jupiters
+% similar age and mass
+% \citep{penev_hats-18_2016,penev_empirical_2018}.  This is expected
+% because when the orbital frequency is larger than the spin rotation
+% frequency, tides donate angular momentum from the planet to the
+% star.  The reverse process occurs between the Earth and Moon,
+% leading to the Moon's slowly increasing separation from Earth of
+% $X.XX$cm~yr$^{-1}$ (CITE lunar ranging experiment).  The hot Jupiter
+% star-planet separation distribution is best reproduced by models
+% that account for tides~\citep{cameron_hierarchical_2018}.
+
 The most convincing signs seen to date of either orbital decay or apsidal
@@
 $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
-% Both \citet{maciejewski_departure_2016} and \citet{patra_2017}
-% emphasized that the timing variations could be caused by , or
-% also by tidal decay.
 Assuming that the period change was caused entirely by tidal decay,
@@
 cause of the timing variations in WASP-12b.
-
-% % There are also indications that apsidal precession should happen. If
-% % there are any trace remnants of eccentricity, this will produce
-% % timing variations.
-% Timing variations can also be caused by rotation of
-% the orbital ellipse within the orbital plane (apsidal precession).
-% This happens because any rotational or tidal bulge creates a
-% gravitation quadrupole field ($r^{-3}$, rather than the typical
-% $r^{-2}$), which causes the line of apsides to move forward or back.
-% \citet{ragozzine_probing_2009} examined apsidal precession.
 
@@
 %
-% Another unfulfilled prediction is that the orbits of hot Jupiters
-% should be apsidally precessing on a timescale of decades
-% \citep{MiraldaEscude2002,HeylGladman2007,PalKocsis2008,JordanBakos2008},
-% as long as the orbits are at least slightly eccentric. In particular,
-% \citet{RagozzineWolf2009} noted that the theoretical precession rate
-% is dominated by the contribution from the planet's tidally deformed
-% mass distribution. They advocated a search for apsidal precession as
-% a means of probing the interiors of hot Jupiters.
-%
 
@@
     \caption{
-        {\bf TESS transits of WASP-4b.} On the left, black points are TESS
-        flux measurements, with a vertical offset applied. Blue curves are
-        best-fit models. Rounded midtimes are printed in BTJD
-        next to the appropriate transits.  The residuals are shown on
-        the right.
+        {\bf TESS transits of WASP-4b.} On the left, black points are
+        TESS flux measurements, with a vertical offset applied. Blue
+        curves are best-fit models. Rounded midtimes are printed in
+        BTJD next to the appropriate transits.  The residuals are
+        shown on the right.
         \label{fig:lightcurves}
@@
 
-The TESS spacecraft recorded images of WASP-4 in an $11\times11$
-pixel array at a cadence of 2 minutes.  It was included on the
+The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
+array at a cadence of 2 minutes.  It was included on the
 ``short-cadence'' target list thanks to the Guest Investigator
@@
 are in the TDB reference (see \citealt{urban_explanatory_2012} for
-explanations of time systems).  The lightcurves were then vetted
-and released by the MIT TESS Science Office to the Mikulski Archive
-for Space Telescopes~\citep{ricker_tess_alerts_2018}.
+explanations of time systems).  The lightcurves were then vetted and
+released by the MIT TESS Science Office to the Mikulski Archive for
+Space Telescopes~\citep{ricker_tess_alerts_2018}.
 
@@
 \texttt{astrobase}~\citep{bhatti_astrobase_2018} We opted not to
-``flatten'' the lightcurves at this stage, as is often done with {\it e.g.}, a
-median filter, spline, or gaussian process regression.  Such
+``flatten'' the lightcurves at this stage, as is often done with {\it
+e.g.}, a median filter, spline, or gaussian process regression.  Such
 procedures are asymmetric about the transit minimum and can skew the
@@
         {\bf TESS saw WASP-4b transit earlier than expected.}
-        Points are differences between observed transit times and
-        the constant-period ephemeris predicted by previous transit
-        measurements.
-        Black points with circle markers are the more precise half of
-        literatures times; gray points are the less precise half.
-        The blue curves represent the $\pm 1\sigma$ credible interval
-        around the constant-period ephemeris.
-        The binned TESS observation (red point) arrived
-        $77.8 \pm 10.7\ {\rm seconds}$ before it was expected.
+        Points are differences between observed transit times and the
+        constant-period ephemeris predicted by previous transit
+        measurements.  Black points with circle markers are the more
+        precise half of literatures times; gray points are the less
+        precise half.  The blue curves represent the $\pm 1\sigma$
+        credible interval around the constant-period ephemeris.  The
+        binned TESS observation (red point) arrived $77.8 \pm 10.7\
+        {\rm seconds}$ before it was expected.
         \label{fig:arrived_early}
@@
 began our measurement procedure by using the Box Least Squares (BLS)
-method to estimate the transit epoch, period, and
-duration~\citep{kovacs_box-fitting_2002}.
-From the BLS parameters, we isolated each transit to a rough estimate
-of its midtime, $\pm$10 transit durations.  In each transit window, we
-then simultaneously fit for a model transit, plus a local linear
-trend.  Our transit model uses the analytic formulae calculated by
-\citet{mandel_analytic_2002}, and implemented by
-\citet{kreidberg_batman_2015}.
+method to estimate the transit epoch, period, and duration
+\citep{kovacs_box-fitting_2002}.  From the BLS parameters, we isolated
+each transit to a rough estimate of its midtime, $\pm$10 transit
+durations.  In each transit window, we then simultaneously fit for a
+model transit, plus a local linear trend.  Our transit model uses the
+analytic formulae calculated by \citet{mandel_analytic_2002}, and
+implemented by \citet{kreidberg_batman_2015}.
 
@@
 the limb darkening, we assumed a fixed quadratic law, with
-coefficients interpolated from the tables computed
-by~\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
+coefficients interpolated from the tables computed by
+\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
 inclination, we adopted the values from~\citet{petrucci_no_2013}.
@@
 initial guess for the free parameters by maximizing the likelihood.
-We then sampled over the posterior using the algorithm proposed
-by~\citet{goodman_ensemble_2010}, and implemented
-by~\citet{foreman-mackey_emcee_2013}.  During this process, we used
-the \texttt{corner} software to check that our posteriors were being
+We then sampled over the posterior using the algorithm proposed by
+\citet{goodman_ensemble_2010}, and implemented by
+\citet{foreman-mackey_emcee_2013}.  During this process, we used the
+\texttt{corner} software to check that our posteriors were being
 adequately sampled~\citep{corner_2016}.
@@
 
-Figure~\ref{fig:lightcurves} shows the light curves, best-fit
-transit models, and residuals.
-Table~1 reports the mid-transit times and
-their uncertainties, and also includes the
-previously reported times that are analyzed in
-\S~\ref{sec:timing}.
+Figure~\ref{fig:lightcurves} shows the light curves, best-fit transit
+models, and residuals.  Table~1 reports the mid-transit times and
+their uncertainties, and also includes the previously reported times
+that are analyzed in \S~\ref{sec:timing}.
 
@@
         {\bf Timing residuals and best-fit models for WASP-4b.}
-        Points are differences between observed transit times and a linear
-        ephemeris fit to the entire timing dataset.
-        The constant period model is a poor fit to the observed times
-        (gray line).
-        Black points are the more precise half of measured times; gray
-        points are the less precise half.
-        The blue curve is the best-fit model with a constant period
-        derivative.
-        The orange curve is the best-fit model assuming apsidal precession
-        causes the observed timing deviations.
+        Points are differences between observed transit times and a
+        linear ephemeris fit to the entire timing dataset.  The
+        constant period model is a poor fit to the observed times
+        (gray line).  Black points are the more precise half of
+        measured times; gray points are the less precise half.  The
+        blue curve is the best-fit model with a constant period
+        derivative.  The orange curve is the best-fit model assuming
+        apsidal precession causes the observed timing deviations.
         \label{fig:times}
@@
 
-Table~1 shows all the transit times used in our
-analysis.  We included data from peer-reviewed literature for which
-the analysis was based on observations of a single, complete transit,
-and for which the midpoint was fit as a free parameter. We also
-required that the time system be clearly documented.
+Table~1 shows all the transit times used in our analysis.  We included
+data from peer-reviewed literature for which the analysis was based on
+observations of a single, complete transit, and for which the midpoint
+was fit as a free parameter. We also required that the time system be
+clearly documented.
 
@@
 
-We use the homogeneous \citet{hoyer_tramos_2013} timing
-study as our
+We use the homogeneous \citet{hoyer_tramos_2013} timing study as our
 starting point.  We took their BJD$_{\rm TT}$ times to be equivalent
@@
 epoch comes from $z$-band photometry acquired by
-\citet{gillon_improved_2009} at the VLT with FORS2.
-Subsequent epochs came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded the GPS clock
-  time in UT reference. They then converted this to BJD in the TDB
-  reference for their reported midtimes.},
-\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011},
-and \citet{nikolov_wasp-4b_2012}.
+\citet{gillon_improved_2009} at the VLT with FORS2.  Subsequent epochs
+came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009}
+recorded the GPS clock time in UT reference. They then converted this
+to BJD in the TDB reference for their reported midtimes.},
+\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
+\citet{nikolov_wasp-4b_2012}.
 %\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
-%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS clock
-%	time in UT reference. They then converted this to BJD in the TDB
-%	reference for their reported midtimes.},
-%and reported transit midtimes precise to
-%6~seconds.  \citet{dragomir_terms_2011}
-%updated the ephemeris using two transits measured with the CTIO 0.9-m
-%and 1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
-%transits from
-%the Baade telescope at Las Campanas\footnote{
+%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS
+%clock time in UT reference. They then converted this to BJD in the
+%TDB reference for their reported midtimes.}, and reported transit
+%midtimes precise to 6~seconds.  \citet{dragomir_terms_2011} updated
+%the ephemeris using two transits measured with the CTIO 0.9-m and
+%1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
+%transits from the Baade telescope at Las Campanas\footnote{
 %\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
@@
 %the MPG/EOS-2.2~m at La Silla.
-\citet{hoyer_tramos_2013} reported
-nine transit observations taken with the the Y4KCAM on the SMARTS 1-m
-telescope, and another three with the SOI on the SOAR 4.2-m telescope.
-\citet{hoyer_tramos_2013} also homogeneously
-redetermined the transit times from all previous studies.  Though
-their latest two epochs did fall slightly below the prediction of a
-linear ephemeris (Panel C of their Figure~6), they did not find any
-statistical need for a quadratic function to fit the O-C values.
+\citet{hoyer_tramos_2013} reported nine transit observations taken
+with the the Y4KCAM on the SMARTS 1-m telescope, and another three
+with the SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013}
+also homogeneously redetermined the transit times from all previous
+studies.  Though their latest two epochs did fall slightly below the
+prediction of a linear ephemeris (Panel C of their Figure~6), they did
+not find any statistical need for a quadratic function to fit the O-C
+values.
 
@@
 omitted the timing measurements from
-\citet{southworth_high-precision_2009}, since there were
-technical problems with the computer clock at the time of
+\citet{southworth_high-precision_2009}, since there were technical
+problems with the computer clock at the time of
 observation~\citep{nikolov_wasp-4b_2012}.  The timing study
@@
 (ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
-\citep{poddany_ETD_2010}.  
-Since the ETD data come from heterogeneous sources, their timestamps
-are less clearly documented, and the times are thus more prone to
-systematic errors, we omit them from consideration.
-%We consider the effects of including ETD data in Appendix~\ref{sec:verify_archival_times}.
+\citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
+sources, their timestamps are less clearly documented, and the times
+are thus more prone to systematic errors, we omit them from
+consideration.
+%We consider the effects of including ETD data in
+%Appendix~\ref{sec:verify_archival_times}.
 \citet{may_mopss_2018} obtained optical spectra using IMCAS on
@@
 their $e\cos\omega$ results into a midtime, using their ephemeris and
-the standard formula
-\citep[{\it e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
+the standard formula \citep[{\it
+e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
 \begin{equation}
-  t_{\rm occ}(E) = t_0 +  P E  +
+  t_{\rm occ}(E) =
+  t_0 +  P E  +
   \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
@@
 arrived early.  One immediate concern was systematic errors in either
-the archival or the TESS timestamps. The tests we used to mitigate
-the latter possibility are described in Appendix~\ref{sec:verify_tess}.
+the archival or the TESS timestamps. The tests we used to mitigate the
+latter possibility are described in Appendix~\ref{sec:verify_tess}.
 
@@
 Following~\citet{gimenez_revision_1995} and~\citet{patra_2017}, in
-this model the angular velocity of the line of apsides
-${\rm d}\omega/{\rm d}E$ is constant,
+this model the angular velocity of the line of apsides ${\rm
+d}\omega/{\rm d}E$ is constant,
 \begin{equation}
@@
 \end{equation}
-The five free parameters are the epoch, sidereal period,
-eccentricity, argument of pericenter at the reference epoch, and
-angular velocity of line of apsides:
+The five free parameters are the epoch, sidereal period, eccentricity,
+argument of pericenter at the reference epoch, and angular velocity of
+line of apsides:
 $(t_0, P_{\rm s}, e, \omega_0, {\rm d}\omega/{\rm d}E)$.
@@
 Assuming the error bars on the transit midtimes are gaussian and
-independent, we then proceed by fitting each model to the timing
-data.
+independent, we then proceed by fitting each model to the timing data.
 For the linear model, we took uniform priors over small windows around
-the expected parameters.
-For the quadratic model we did the same, and took the quadratic term
-to have a uniform prior corresponding to
+the expected parameters.  For the quadratic model we did the same, and
+took the quadratic term to have a uniform prior corresponding to
 $
@@
 for $Q_\star$ the ``quality factor'' corresponding to the star's
-response against tidal forcing, defined in
-\S~\ref{sec:implications}.
+response against tidal forcing, defined in \S~\ref{sec:implications}.
 For the precession model, we took the same priors in epoch and
-sidereal period, and drew the remaining parameters from wide
-uniform distributions.
+sidereal period, and drew the remaining parameters from wide uniform
+distributions.
 
@@
 model.
+%TODO: final numbers
 The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of freedom.
@@
 The median MCMC precession fit has $\chi^2 = 51.0$ and 55 degrees of freedom.
-
-%FIXME UPDATE THESE NUMBERS!!
 
@@
 
+%TODO: final numbers
 The quadratic model fits better than the precession model.
@@
 
+%TODO: final numbers
 Assuming the orbital decay model, we find a period derivative 
@@
 
+%TODO: final numbers
 If we instead assume the precession model, the best-fit
@@
 planets.  Though there are roughly 20 hot Jupiters that might be
-expected to decay faster, some orbit stars with minimal convective zones
-({\it e.g.}, WASP-18), and most are not as well-observed as
+expected to decay faster, some orbit stars with minimal convective
+zones ({\it e.g.}, WASP-18), and most are not as well-observed as
 WASP-4.
@@
 
-Let us assume further that the 
-``constant phase lag'' model for tidal interaction in
-a binary applies \citep{zahn_tidal_1977}.
-%, the travelling equilibrium tidal
-%bulge is damped by turbulent viscosity in the convective envelope.
-The fluid response is parametrized by a constant modified\footnote{For
-stars, $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to
-explicitly distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
+Let us assume further that the ``constant phase lag'' model for tidal
+interaction in a binary applies \citep{zahn_tidal_1977}.  The fluid
+response is parametrized by a constant modified\footnote{For stars,
+$k_\star \sim \mathcal{O}(10^{-2})$, so it is important to explicitly
+distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
 e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
@@
 which is smaller when the star's density distribution is more
-centrally concentrated. $Q_\star$ is the ratio of the
-energy stored in the equilibrium deformation of the star, divided by
-the energy dissipated per tidal period \citep[{\it
-e.g.},][]{goldreich_q_1966}.  A larger $Q_\star$ implies less
-efficient dissipiation.  Though the fluid mechanics responsible for
-the dissipation should depend on the amplitude and frequency of
-the tidal perturbation \citep[][Section~3.3]{ogilvie_tidal_2014}, we
-will treat $Q_\star'$ as a constant.  This should not be taken
-literally; $Q_\star'$ may in fact change sharply with varying tidal
-frequency \citep{penev_empirical_2018}.
+centrally concentrated. $Q_\star$ is the ratio of the energy stored in
+the equilibrium deformation of the star, divided by the energy
+dissipated per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.
+A larger $Q_\star$ implies less efficient dissipiation.  Though the
+fluid mechanics responsible for the dissipation should depend on the
+amplitude and frequency of the tidal perturbation
+\citep[][Section~3.3]{ogilvie_tidal_2014}, we will treat $Q_\star'$ as
+a constant.  This should not be taken literally; $Q_\star'$ may in
+fact change sharply with varying tidal frequency
+\citep{penev_empirical_2018}.
 
@@
 Using an empirical approach, \citet{penev_empirical_2018} reported a
-$Q_\star'$ measurement for WASP-4.
-In earlier work,
+$Q_\star'$ measurement for WASP-4.  In earlier work,
 \citet{penev_hats-18b_2016} noted that HATS-18b and WASP-19b orbited
@@
 whose stellar rotation periods are about 10 days
-(\citealt{pont_empirical_2009} pointed out similar evidence for hot Jupiter 
-host 
-spin-up).  The expected rotation periods at solar ages
+(\citealt{pont_empirical_2009} pointed out similar evidence for hot
+Jupiter host spin-up).  The expected rotation periods at solar ages
 are 20 to 30 days
@@
 stellar rotation period was due to tides depositing angular momentum
-in the star,
-\citet{penev_empirical_2018} 
-modeled the evolution of hot Jupiter systems under
-the influence of a magnetized wind and a constant phase-lag tide.
-They found two-sided limits for $Q_\star'$ in 35 of
-the 188 systems they examined.  In these 35 systems, they found that
-$Q_\star'$ varied from $10^5$ at tidal periods of 2 days to $10^7$ at
-tidal periods of 0.5 days.  For WASP-4 specifically, their method gave
-$Q_\star' \approx (1.2^{+1.0}_{-0.5})\times10^7$. This is more than 2 orders
-of magnitude less efficient than the dissipation rate inferred from
-the decaying orbit would imply.
+in the star, \citet{penev_empirical_2018} modeled the evolution of hot
+Jupiter systems under the influence of a magnetized wind and a
+constant phase-lag tide.  They found two-sided limits for $Q_\star'$
+in 35 of the 188 systems they examined.  In these 35 systems, they
+found that $Q_\star'$ varied from $10^5$ at tidal periods of 2 days to
+$10^7$ at tidal periods of 0.5 days.  For WASP-4 specifically, their
+method gave $Q_\star' \approx (1.2^{+1.0}_{-0.5})\times10^7$. This is
+more than 2 orders of magnitude less efficient than the dissipation
+rate inferred from the decaying orbit would imply.
 
@@
 \citet{bonomo_gaps_2017}, 2 (13) had $2\sigma$ upper limits of $e <
-0.01$ ($e<0.02$).  
-For WASP-4b, the strongest $2\sigma$ upper-limit on its eccentricity
-from RVs is $e<0.011$
+0.01$ ($e<0.02$).  For WASP-4b, the strongest $2\sigma$ upper-limit on
+its eccentricity from RVs is $e<0.011$
 \citep{husnoo_observational_2012,bonomo_gaps_2017}.
 
-Occultation timing can provide stronger eccentricity limits than
-RV measurements, but they often come with caveats.
-For WASP-4b, the strongest limit on its eccentricity comes from the
-secondary eclipses observed by \citet{beerer_secondary_2011}, which
-led to a $2\sigma$ upper limit on $| e\cos\omega | $ of 0.0024.
+Occultation timing can provide stronger eccentricity limits than RV
+measurements, but they often come with caveats.  For WASP-4b, the
+strongest limit on its eccentricity comes from the secondary eclipses
+observed by \citet{beerer_secondary_2011}, which led to a $2\sigma$
+upper limit on $| e\cos\omega | $ of 0.0024.
 % \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
-% 0.0018$. 
-% \citet{zhou_secondary_2015} report a weaker $e\cos\omega = -0.001
-% \pm 0.003$. 
-Placing eccentricity limits stronger than $e< 10^{-3}$
-has only been possible for a few hot Jupiters, and often with numerous
-false-starts.  HD 189733b, for example, was initially reported to have
-a very small eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
+% 0.0018$.  \citet{zhou_secondary_2015} report a weaker $e\cos\omega =
+% -0.001 \pm 0.003$. 
+Placing eccentricity limits stronger than $e< 10^{-3}$ has only been
+possible for a few hot Jupiters, and often with numerous false-starts.
+HD 189733b, for example, was initially reported to have a very small
+eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
 \citep{knutson_map_2007}.  This result was retracted by
@@
 of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit
-yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
-\citet{winn_measurement_2005} combined Spitzer secondary eclipses with
-RV data and reported a best-fit eccentricity of $e>0.0057$ at 90\%
-confidence.  They suspected that the data were consistent with zero
-eccentricity, and that the lower bound on $e$ was an artifact of
-imperfect limb-darkening model.  Stronger limits were eventually put
-in place by \citet{crossfield_spitzer_mips_2012}, who found
-$e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b,
-for which the discovery RVs and photometry led to a non-zero but small
-eccentricity measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.
-A prompt ground-based occultation measurement by
-\citet{lopez-morales_day-side_2010} concurred with weak statistical
-significance that a small eccentricity was favored, but
+yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for
+which \citet{winn_measurement_2005} combined Spitzer secondary
+eclipses with RV data and reported a best-fit eccentricity of
+$e>0.0057$ at 90\% confidence.  They suspected that the data were
+consistent with zero eccentricity, and that the lower bound on $e$ was
+an artifact of imperfect limb-darkening model.  Stronger limits were
+eventually put in place by \citet{crossfield_spitzer_mips_2012}, who
+found $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is
+WASP-12b, for which the discovery RVs and photometry led to a non-zero
+but small eccentricity measurement \citep[][at
+$3\sigma$]{hebb_wasp-12b_2009}.  A prompt ground-based occultation
+measurement by \citet{lopez-morales_day-side_2010} concurred with weak
+statistical significance that a small eccentricity was favored, but
 further follow-up found occultation times consistent with circular
@@
 For WASP-4, ...
-%TODO: check this mechanism
-%FIXME: take a look at RVs. rule this shit out. /  understand where
-%the bounds are.
+% TODO: check this mechanism
+% FIXME: take a look at RVs. rule this out, or understand where the
+% bounds are.
 
 \paragraph{Fluctuations in the gravitational potential from convection}
-A separate mechanism that might produce the eccentricity is simply stellar
-convection \citep[][Section 7]{phinney_pulsars_1992}.
-Phinney presents this idea by analogy with a pendulum, which is never really at
-rest. Instead, collisions with air molecules randomly perturb its momentum and
-energy, and the equipartition theorem of statistical mechanics tells
-us the average energy is $\langle E_{\rm pend} \rangle = k
-T_{\rm air}$. 
-In hot Jupiter systems, the ``air molecules'' are convective eddies, and
-their associated kinetic energy leads the orbit to never have exactly zero
-eccentricity.
-In particular, from \citealt{phinney_pulsars_1992} Equation (7.33),
+A separate mechanism that might produce the eccentricity is simply
+stellar convection \citep[][Section 7]{phinney_pulsars_1992}.  Phinney
+presents this idea by analogy with a pendulum, which is never really
+at rest. Instead, collisions with air molecules randomly perturb its
+momentum and energy, and the equipartition theorem of statistical
+mechanics tells us the average energy is $\langle E_{\rm pend} \rangle
+= k T_{\rm air}$.  In hot Jupiter systems, the ``air molecules'' are
+convective eddies, and their associated kinetic energy leads the orbit
+to never have exactly zero eccentricity.  In particular, from
+\citealt{phinney_pulsars_1992} Equation (7.33),
 \begin{equation}
@@
 width (mass) of the convective region,  $\mu$ the reduced mass, $n$
-the orbital frequency, and $a$ the semimajor axis.
-For WASP-4, convection in the star is more important than
-convection in the planet, and
-$\langle e^2 \rangle^{1/2} < 10^{-5}$.
+the orbital frequency, and $a$ the semimajor axis.  For WASP-4,
+convection in the star is more important than convection in the
+planet, and $\langle e^2 \rangle^{1/2} < 10^{-5}$.
 
@@
 \citet{applegate_mechanism_1992} proposed a mechanism to explain these
-modulations in which the internal structure of a magnetically active star
-changes shape via cyclic exchange of angular momentum between the inner and
-outer zones of the star.
-This model could also apply to a hot Jupiter orbiting a star with a
-convective zone.
-The changing gravitational quadrupole of the star would cause the orbit of the
-planet to precess on the timescale of the stellar activity cycle
-("the dynamo timescale").
-An essential difference between this process and apsidal precession is
-that it is expected to be quasi-periodic \citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}.
-The transit and occultation deviations would also have the same sign, while
-for apsidal precession they have opposite signs.
-For WASP-4, \citet{watson_orbital_2010} estimated that the effect
-could produce timing deviations of up to 15 seconds, depending on the
-modulation period of the stellar dynamo, and the corresponding level
-of differential surface shear.
-This cannot explain the majority our observed $78$ second variation.
+modulations in which the internal structure of a magnetically active
+star changes shape via cyclic exchange of angular momentum between the
+inner and outer zones of the star.  This model could also apply to a
+hot Jupiter orbiting a star with a convective zone.  The changing
+gravitational quadrupole of the star would cause the orbit of the
+planet to precess on the timescale of the stellar activity cycle ("the
+dynamo timescale").  An essential difference between this process and
+apsidal precession is that it is expected to be quasi-periodic
+\citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}.  The transit
+and occultation deviations would also have the same sign, while for
+apsidal precession they have opposite signs.  For WASP-4,
+\citet{watson_orbital_2010} estimated that the effect could produce
+timing deviations of up to 15 seconds, depending on the modulation
+period of the stellar dynamo, and the corresponding level of
+differential surface shear.  This cannot explain the majority our
+observed $78$ second variation.
 
@@
 for the system, and reported no detectable slope at $3\sigma$.
-However, \citet{knutson_friends_2014} collected 5 Keck HIRES radial velocity
-measurements for WASP-4, and merged them with measurements from
-CORALIE and HARPS reported by \citet{wilson_wasp-4b_2008},
+However, \citet{knutson_friends_2014} collected 5 Keck HIRES radial
+velocity measurements for WASP-4, and merged them with measurements
+from CORALIE and HARPS reported by \citet{wilson_wasp-4b_2008},
 \citet{pont_determining_2011}, and \citet{husnoo_observational_2012}.
-Using these four datasets,
-\citet{knutson_friends_2014} found that $\dot{\gamma} =
--0.0099^{+0.0052}_{-0.0054}$, corresponding to a $2\sigma$ limit on
-the acceleration towards us of $\dot{\gamma} > - 0.021 \,{\rm m}\,{\rm
-s}^{-1}\,{\rm day}^{-1}$.  This limits $\dot{P}$ to be less than
-$9\times 10^{-11}$ at $2\sigma$.  
+Using these four datasets, \citet{knutson_friends_2014} found that
+$\dot{\gamma} = -0.0099^{+0.0052}_{-0.0054}$, corresponding to a
+$2\sigma$ limit on the acceleration towards us of $\dot{\gamma} > -
+0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.  This limits $\dot{P}$
+to be less than $9\times 10^{-11}$ at $2\sigma$.  
 %FIXME TODO final value
-Our measured value is $\dot{P} =
--(3.9^{+0.7}_{-0.5})\times 10^{-10}$.  We can
-thus rule out radial acceleration as the sole cause of the observed period
-change.  However, it could cause up to one-quarter of the observed
-period decrease.  Improved long-term radial velocity monitoring with
-the same instruments will help clarify the situation, particularly
-given the fact that short-term trends of roughly $\dot{\gamma} \approx
-2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ are seen over timescales
-of order 10 days in WASP-4~\citep{husnoo_observational_2012}.
-Given the known
-spottedness of the system~\citep{sanchis-ojeda_starspots_2011}, these
+Our measured value is $\dot{P} = -(3.9^{+0.7}_{-0.5})\times 10^{-10}$.
+We can thus rule out radial acceleration as the sole cause of the
+observed period change.  However, it could cause up to one-quarter of
+the observed period decrease.  Improved long-term radial velocity
+monitoring with the same instruments will help clarify the situation,
+particularly given the fact that short-term trends of roughly
+$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ are
+seen over timescales of order 10 days in
+WASP-4~\citep{husnoo_observational_2012}.  Given the known spottedness
+of the system~\citep{sanchis-ojeda_starspots_2011}, these
 short-timescale variations may also have a contribution from stellar
@@
 
-There are two other small effects worth briefly mentioning.  The
-first is the \citet{shklovskii_possible_1970} effect, in which the
-star's proper motion leads to a changing radial velocity.  This would be
+There are two other small effects worth briefly mentioning.  The first
+is the \citet{shklovskii_possible_1970} effect, in which the star's
+proper motion leads to a changing radial velocity.  This would be
 apparent in the Doppler measurements described above.  The effect
@@
 \citet{rafikov_stellar_2009}, comes from the star's on-sky motion
-altering our viewing angle, and leads to an observed apsidal precession.
-The corresponding change is $\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$.
+altering our viewing angle, and leads to an observed apsidal
+precession.  The corresponding change is $\dot{P} \sim (P\mu)^2/2\pi
+\sim 10^{-21}$.
 
 Many of these phenomena are likely occurring simultaneously.  From the
-preceding discussion, we found that radial acceleration, proper-motion,
-and the Applegate mechanism
-are unlikely to cause the timing deviations.
-That leaves tidal decay and apsidal
-precession as the possible causes.
+preceding discussion, we found that radial acceleration,
+proper-motion, and the Applegate mechanism are unlikely to cause the
+timing deviations.  That leaves tidal decay and apsidal precession as
+the possible causes.
 
@@
 
-If TESS continues observing after
-its primary mission, 
-it will observe additional transits of WASP-4b in the
-early 2020s (Figure~\ref{fig:future}).
-In order to rule between orbital decay and
-apsidal precession, occultation measurements in the
-near term, and also in the mid-2020s, will be helpful.
-However it may take up to a decade to firmly distinguish
-the two models, through the combination of transit and occultation
-measurements.
-These observations should be carried out, because they will
-lead to new knowledge; either a
-measure of the efficiency of tidal dissipation within the
-star, or the eccentricity and Love number of the planet.
-Measuring either for a single system would be
-rather interesting.
-In the coming decade, wide-field photometric surveys
-(TESS, HATPI, NGTS, PLATO)
-might even enable such measurements at a population
-level, which would be a good deal more interesting.
+If TESS continues observing after its primary mission, it will observe
+additional transits of WASP-4b in the early 2020s
+(Figure~\ref{fig:future}).  In order to rule between orbital decay and
+apsidal precession, occultation measurements in the near term, and
+also in the mid-2020s, will be helpful.  However it may take up to a
+decade to firmly distinguish the two models, through the combination
+of transit and occultation measurements.  These observations should be
+carried out, because they will lead to new knowledge; either a measure
+of the efficiency of tidal dissipation within the star, or the
+eccentricity and Love number of the planet.  Measuring either for a
+single system would be rather interesting.  In the coming decade,
+wide-field photometric surveys (TESS, HATPI, NGTS, PLATO) might even
+enable such measurements at a population level, which would be a good
+deal more interesting.
 

### 2018-12-05 src/bonomo_any_eccentric_hjs.py
add scripts used for discussion

@@ -1,0 +1,41 @@
+# -*- coding: utf-8 -*-
+'''
+parse Bonomo et al (2017) table 8. ask: are there any hot Jupiters with small
+(e<0.1) but significant eccentricties?
+'''
+from __future__ import division, print_function
+
+from astropy.table import Table
+
+import matplotlib.pyplot as plt, pandas as pd, numpy as np
+
+from numpy import array as nparr
+
+from astropy.coordinates import SkyCoord
+from astropy import units, constants
+
+t = Table.read("../data/Bonomo_2017_table8.vot", format="votable")
+
+sel_eccentric = (t['Forbit']=='E')
+sel_weakly_ecc = (t['ecc']<=0.1)
+
+print('Bonomo+2017 has {:d} homogeneous transiting giant planets'.
+      format(len(t)))
+
+print('Bonomo+2017 has {:d} significant eccentric planets'.
+      format(len(t[sel_eccentric])))
+
+tsel = t[sel_eccentric & sel_weakly_ecc]
+print('Bonomo+2017 has {:d} e<0.1, but significant eccentricity planets'.
+      format(len(tsel)))
+
+print(tsel)
+
+df = t.to_pandas()
+print('Bonomo+2017 has {:d} with ecc2s < {:.2f}'.format(
+    len(df[df['ecc2s']<0.01]), 0.01))
+
+print('Bonomo+2017 has {:d} with ecc2s < {:.2f}'.format(
+    len(df[df['ecc2s']<0.02]), 0.02))
+
+import IPython; IPython.embed()

### 2018-12-05 src/calculate_wasp4b_eccentricity_damping_timescale.py

@@ -1,0 +1,44 @@
+"""
+calculate the eccentricity damping timescale for WASP-4b
+
+also, sanity-check the approximations used to simplify from the
+Metzger+ 2012 appendix B equations to the quoted tau_a and tau_e
+timescales.
+"""
+import numpy as np
+from astropy import units as u, constants as const
+
+# Petrucci+ 2013, table 3, for all parameters
+P = 1.33823251*u.day
+
+Mstar, Rstar, Mplanet, Rplanet = (
+    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+)
+
+sma = 0.0228*u.au
+
+Qplanetprime = 1e5
+
+tau_e = ( 2 * Qplanetprime / (63*np.pi) *
+         (sma/Rplanet)**5 *
+         (Mplanet/Mstar) *
+         P
+).to(u.Myr)
+
+print('-'*42)
+print('2018/12/04.3 eq (1) timescale')
+print(tau_e)
+print('tau_e = {:.2f} (Qp\'/10^5) Myr'.format(tau_e.value))
+print('-'*42)
+
+# from Metzger+ 2012 appendix B
+K_1 = 63/4 * const.G**(1/2) * Mstar**(3/2) * Mplanet**(-1)
+K_2 = 225/16 * const.G**(1/2) * Mplanet * Mstar**(-1/2)
+
+print('K_1: {:.3e}'.format(K_1.cgs))
+print('K_2: {:.3e}'.format(K_2.cgs))
+
+print('K_1 R_p^5: {:.3e}'.format((K_1*Rplanet**5).cgs))
+print('K_2 R_star^5: {:.3e}'.format((K_2*Rstar**5).cgs))
+
+

### 2018-12-05 src/check_beerer_2011_occ_time.py

@@ -1,0 +1,73 @@
+"""
+Beerer+ 2011 reported eclipse times in "BJD", and offsets relative to the Winn+
+2009 ephemeris.
+
+They probably meant BJD_UTC, because of the description in the text. We should
+be able to check by recalculating the offsets. From their Table 1, the expected
+offsets are 0.5 and 0.1 minutes.
+
+The calculations below produce:
+
+    occ time from table - predicted occ time from Winn ephem [min]
+    1.0690627992153168
+    1.4352265745401382
+    (occ time from table - ltt correction) - predicted occ time from Winn ephem
+    0.6790629029273987
+    1.0452266782522202
+    (UTC occ time from table - ltt correction) - predicted occ time from Winn ephem
+    -0.40427058935165405
+    -0.03810681402683258
+    (UTC occ time from table) - predicted occ time from Winn ephem
+    -0.014270693063735962
+    0.3518930822610855
+
+so none of the permutations of offsets produce the expected offset. The closest is
+the negative of
+    (UTC occ time from table - ltt correction) - predicted occ time from Winn ephem
+which is pretty close.
+
+So this would imply we assume the standard thing: the reported times are in
+BJD_UTC, WITHOUT the light-travel time correction applied.
+"""
+import numpy as np
+
+# Beerer+ 2011 give times in BJD_UTC.
+t0_bjd_utc = 2455174.87731
+t1_bjd_utc = 2455172.2011
+
+# correct them by 65 seconds, per 2009 obsn (Fig 3 Eastman+ 2010)
+t0_bjd_tdb = t0_bjd_utc + (65/(60*60*24))
+t1_bjd_tdb = t1_bjd_utc + (65/(60*60*24))
+
+# Winn+ 2009 give an epoch, that Beerer+2011 say they compute their offset
+# from.
+tc_0 = 2454697.797562
+period = 1.33823214
+epoch = np.arange(0, 2000, 1)
+tocc_mids = tc_0 + period*epoch + period/2
+
+tmids_first = tocc_mids[np.argmin(np.abs(tocc_mids - t0_bjd_tdb))]
+tmids_second = tocc_mids[np.argmin(np.abs(tocc_mids - t1_bjd_tdb))]
+
+print('occ time from table - predicted occ time from Winn ephem [min]')
+print((t0_bjd_tdb - tmids_first)*24*60)
[…]

### 2018-12-05 src/check_rpbyrstar_fit_vs_literature.py

@@ -1,0 +1,107 @@
+# -*- coding: utf-8 -*-
+from __future__ import division, print_function
+import numpy as np, matplotlib.pyplot as plt, pandas as pd
+from astropy import units as u, constants as const
+
+import os, argparse, pickle, h5py
+from glob import glob
+from numpy import array as nparr
+
+def main(ticid, pickledir='/home/luke/Dropbox/proj/tessorbitaldecay/results/tess_lightcurve_fit_parameters/',
+    sampledir='/home/luke/local/emcee_chains/',
+    fittype='mandelagol_and_line',
+    ndim=4):
+
+    # from Southworth+ 2009 table 8:
+
+    w08_rp = 1.416*u.Rjup
+    w08_rstar = 0.937*u.Rsun
+
+    g09_rp = 1.304*u.Rjup
+    g09_rstar = 0.873*u.Rsun
+
+    w09_rp = 1.365*u.Rjup
+    w09_rstar = 0.912*u.Rsun
+
+    s09_rp = 1.371*u.Rjup
+    s09_rstar = 0.914*u.Rsun
+
+    for paper, rp, rstar in zip(
+        ['Wilson+08', 'Gillon+09', 'Winn+09', 'Southworth+09'],
+        [w08_rp, g09_rp, w09_rp, s09_rp],
+        [w08_rstar, g09_rstar, w09_rstar, s09_rstar]
+    ):
+
+        print('{:s}: Rp/Rstar = {:.4g}'.format(paper, (rp/rstar).cgs))
+
+    # NOW GET UR MEASURED RP/RSTAR VALUES, COMPARE
+    pickledir += str(ticid)
+    fpattern = (
+        '{:s}_{:s}_fit_empiricalerrs_t???.pickle'.
+        format(str(ticid), fittype)
+    )
+    fnames = np.sort(glob(os.path.join(pickledir,fpattern)))
+    samplepattern = (
+        '{:s}_{:s}_fit_samples_{:d}d_t???_empiricalerrs.h5'.
+        format(str(ticid), fittype, ndim)
+    )
+    samplenames = np.sort(glob(sampledir+samplepattern))
+
+    rp_list, rp_bigerrs = (
+        [],[]
+    )
+
+    transit_ix = 0
+    for fname, samplename in zip(fnames, samplenames):
+        transit_ix += 1
+
+        d = pickle.load(open(fname, 'rb'))
+
+        fitparams = d['fitinfo']['finalparams']
[…]

### 2018-12-05 src/check_times_vs_feis.py

@@ -1,0 +1,61 @@
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+import os
+
+def main():
+
+    fei_df = pd.read_csv('../data/fei_times.csv')
+
+    mydir = '/home/luke/Dropbox/proj/tessorbitaldecay/data'
+    myfname = 'WASP-4b_transits_and_TESS_times_O-C_vs_epoch_selected.csv'
+
+    my_df = pd.read_csv(os.path.join(mydir, myfname), sep=';')
+    my_tess = my_df[my_df['original_reference']=='me']
+
+    my_tess_times = np.array(my_tess['sel_transit_times_BJD_TDB'] - 2457000)
+    fei_times = np.array(fei_df['BJD-2457000'])
+
+    my_unc = np.array(my_tess['err_sel_transit_times_BJD_TDB'])
+    fei_unc = np.array(fei_df['unc'])
+
+    print('luke times\n')
+    print(my_tess_times)
+
+    print('\nfei times\n')
+    print(fei_times)
+
+    print('\n(luke-fei) times (units: minutes)\n')
+    print((my_tess_times - fei_times)*24*60)
+
+    print('\nstatistics for (luke times - fei times), units: minutes\n')
+    print(
+        pd.DataFrame(
+        {'luke_time_minus_fei_time':(my_tess_times - fei_times)*24*60}
+        ).describe()
+    )
+
+    print('\n(luke-fei) times / fei_unc\n')
+    print((my_tess_times - fei_times)/fei_unc)
+
+    print('\n(luke-fei) times / luke_unc\n')
+    print((my_tess_times - fei_times)/my_unc)
+
+    print('\nfei unc / luke_unc\n')
+    print(fei_unc/my_unc)
+
+    n, bins, patches = plt.hist((my_tess_times - fei_times)*24*60, 10, facecolor='g', alpha=0.75)
+
+    plt.xlabel('(luke time - fei time) [minutes]')
+    plt.ylabel('number')
+
+    plt.title('do independent time measurements agree?')
+
+    plt.grid(True)
+    savname = '../results/check_times_vs_feis.png'
+    plt.savefig(savname, dpi=350,
+                bbox_inches='tight')
+    print('\nmade {}'.format(savname))
+
+    import IPython; IPython.embed()
+
+if __name__=="__main__":
+    main()

### 2018-12-05 src/convert_zhou_2015_ecosw_to_midtime.py

@@ -1,0 +1,59 @@
+import numpy as np
+from astropy import units as u, constants as c
+from astropy.time import Time
+from astropy import coordinates as coord
+
+def get_epoch():
+    # work out the approximate epoch of observation
+    # answer: first observation was E = 1555.5, next was 1560.5.
+
+    # obs_time = 2014-09-04 15:54–17:02
+    # 2014-09-11 09:23–11:38
+    sso = coord.EarthLocation.of_site('siding spring observatory')
+
+    wasp4 = coord.SkyCoord("23 34 15.0858223039", "-42 03 41.049468209",
+                           unit=(u.hourangle, u.deg), frame='icrs')
+
+    obstime_0 = ['2014-09-04T15:54:00', '2014-09-04T17:02:00']
+    time_0 = Time(obstime_0, format='isot', scale='utc', location=sso)
+    ltt_bary_0 = time_0.light_travel_time(wasp4)
+    time0_bjd_tdb = (time_0.tdb + ltt_bary_0).jd
+
+    epoch0 = (time0_bjd_tdb - t0) / period
+
+    obstime_1 = ['2014-09-11T09:23:00', '2014-09-11T11:38:00']
+    time_1 = Time(obstime_1, format='isot', scale='utc', location=sso)
+    ltt_bary_1 = time_1.light_travel_time(wasp4)
+    time1_bjd_tdb = (time_1.tdb + ltt_bary_1).jd
+
+    epoch1 = (time1_bjd_tdb - t0) / period
+
+
+t0 = 2454823.59192
+t0_err = 3e-5
+
+ecosomega = -0.001
+ecosomega_err = 3e-3
+
+period = 1.3382320
+period_err = 2e-7
+
+epoch0 = 1555.5
+epoch1 = 1560.5
+meanepoch = 1557.
+
+tocc = t0 + period*meanepoch + period/2 * (1 + 4/(np.pi)*ecosomega)
+tocc_upper = (
+    t0+t0_err + (period+period_err)*meanepoch +
+    (period+period_err)/2 * (1 + 4/(np.pi)*(ecosomega+ecosomega_err))
+)
+tocc_lower = (
+    t0-t0_err + (period-period_err)*meanepoch +
+    (period-period_err)/2 * (1 + 4/(np.pi)*(ecosomega-ecosomega_err))
+)
+
+print('converting Zhou+ 2015\'s ecosw constraint to an occultation time:')
[…]

### 2018-12-05 src/is_it_doppler_shift.py

@@ -1,0 +1,75 @@
+import numpy as np
+from astropy import units as u, constants as c
+
+# WASP-12b sanity check.
+P = 1.09142*u.day
+gamma_dot_value, gamma_dot_sig = -0.0009, 0.0093
+N_CPS, N_data = 30, 3
+
+gamma_dot_preferred = gamma_dot_value * (u.m/u.s)/u.day
+gamma_dot_limit = (gamma_dot_value - 2*gamma_dot_sig) * (u.m/u.s)/u.day
+
+dP_dt_preferred = gamma_dot_preferred * P / c.c
+dP_dt_limit = gamma_dot_limit * P / c.c
+dP_dt_measured = -9.3e-10
+dP_dt_measured_err = 1.1e-10
+
+print('for WASP-12b')
+print('Knutson+ 2014 reported {:d} data points, from {:d} data sets'.
+     format(N_CPS, N_data))
+print('gamma_dot_value: {:.6g} +/-({:.6g}) --> gamma_dot_limit: {:.6g}'.
+      format(gamma_dot_value, gamma_dot_sig, gamma_dot_limit))
+print('2σ upper limit on dP/dt change from Doppler shift: {:.6g}'.format(dP_dt_limit.cgs))
+print('dP/dt measured: {:.6g}, +/-({:.6g})'.
+      format(dP_dt_measured, dP_dt_measured_err))
+print('(dP/dt_Dopplerlimit)/(dP/dt_measured): {:.6g}'.format(dP_dt_limit.cgs.value/dP_dt_measured))
+print('(dP/dt_Dopplerquoted)/(dP/dt_measured): {:.6g}'.format(dP_dt_preferred.cgs.value/dP_dt_measured))
+
+# WASP-4b. Knutson+ 2014 values.
+P = 1.3382324*u.day
+gamma_dot_value, gamma_dot_sig = -0.0099, 0.0054
+N_CPS, N_data = 5, 4
+
+gamma_dot_preferred = gamma_dot_value * (u.m/u.s)/u.day
+gamma_dot_limit = (gamma_dot_value - 2*gamma_dot_sig) * (u.m/u.s)/u.day
+
+dP_dt_preferred = gamma_dot_preferred * P / c.c
+dP_dt_limit = gamma_dot_limit * P / c.c
+dP_dt_measured = -3.854e-10
+dP_dt_measured_err = 5.6e-11
+
+print('\nfor WASP-4b')
[…]

### 2018-12-05 src/phinney_fluctuation_eccentricity.py

@@ -1,0 +1,95 @@
+import numpy as np
+from astropy import units as u, constants as c
+
+"""
+In binary systems with a single convective component, you get (BG2019)
+
+<e^2>^(1/2) ~= 2e-5 * ( L_star/L_sun * R_env/R_sun * (1.4Msun)/M_star )^(1/3)
+                    * (M_env / (0.0004M_sun) )^(1/6)
+
+One caveat might be that the planet also has convection, and we are ignoring
+that. Is this OK? Instead, check
+
+<e^2>^(1/2) ~= 2e-5 * ( L_HJ/L_sun * R_envHJ/R_sun * (1.4Msun)/M_HJ )^(1/3)
+                    * (M_envHJ / (0.0004M_sun) )^(1/6).
+
+Also, rederive it by starting with Phinney (1992) equation (7.33). The mass
+ratio is the important term, not the stellar mass as in the BG2019
+approximation.
+
+For WASP-4...
+
+"""
+
+def mean_ecc(L, M_planet, M_star, R_conv, M_conv, period, sma):
+    # from 2018/12/02.0
+
+    μ = (M_star*M_planet)/(M_star + M_planet)
+    Ω_b = 2*np.pi/period
+
+    return (
+        8.3e-3 * (μ * Ω_b**2 * sma**2)**(-1/2) *
+        ( L**2 * R_conv**2 * M_conv )**(1/6)
+    ).cgs
+
+# Petrucci+ 2013 table 3, for the star.
+Rstar = 0.92*u.Rsun
+Mstar = 0.89*u.Msun
+Teff = 5436*u.K
+age = 7*u.Gyr
+Fe_by_H = -0.05
+
+Lstar = ( ( c.sigma_sb * Teff**4 ) * 4*np.pi*Rstar**2 ).to(u.Lsun)
+
+# from my R_tachocline_vs_t_varM_merged.pdf plot
+log10_R_tacholine = -0.2
+R_tachocline = 10**(log10_R_tacholine)*u.Rsun
+assert Rstar > R_tachocline
+R_env_star = Rstar - R_tachocline
+
+# note I_conv ~= (constant) * M_conv * R_conv^2
+log10_I_conv = -5
+I_conv = 10**log10_I_conv * u.Msun * u.Rsun**2
+M_env_star = ( I_conv / (R_env_star**2) ).to(u.Msun)
+
+##########
+period = 1.3382351*u.day
+sma = 0.0228*u.au
+
[…]

### 2018-12-05 src/small_timing_variations.py

@@ -1,0 +1,34 @@
+import numpy as np
+from astropy import units as u, constants as c
+
+# WASP-12b sanity check.
+P = 1.09142*u.day
+d = (1/(2.3122e-3))*u.pc # Gaia DR2
+dec = 29.672266247*u.degree
+pm_RA = (-1.570e-3*u.arcsec/u.yr).to(u.rad/u.yr)
+pm_dec = (-6.958e-3*u.arcsec/u.yr).to(u.rad/u.yr)
+
+mu = np.sqrt( pm_dec**2 + np.cos(dec)**2 * pm_RA**2 )
+
+dP_dt_Shk = ((P*mu**2 * d)/(c.c)).cgs
+dP_dt_rafikov_apsidal = ((P*mu)**2/(2*np.pi)).cgs
+print('for WASP-12b')
+print('mu: {:.3e} = {:.3e}'.format(mu, mu.to(u.arcsec/u.yr)))
+print('dP_dt_Shk: {:.3e}'.format(dP_dt_Shk))
+print('dP_dt_rafikov_apsidal: {:.3e}'.format(dP_dt_rafikov_apsidal))
+
+# WASP-4b
+P = 1.3382324*u.day
+d = (1/(3.7145e-3))*u.pc # Gaia DR2
+dec = (-42.061779443*u.degree).to(u.rad)
+pm_RA = (9.874e-3*u.arcsec/u.yr).to(u.rad/u.yr)
+pm_dec = (-87.518e-3*u.arcsec/u.yr).to(u.rad/u.yr)
+
+mu = np.sqrt( pm_dec**2 + np.cos(dec)**2 * pm_RA**2 )
+
+dP_dt_Shk = ((P*mu**2 * d)/(c.c)).cgs
+
+print('for WASP-4b')
+print('mu: {:.3e} = {:.3e}'.format(mu, mu.to(u.arcsec/u.yr)))
+print('dP_dt_Shk: {:.3e}'.format(dP_dt_Shk))
+print('dP_dt_rafikov_apsidal: {:.3e}'.format(dP_dt_rafikov_apsidal))

### 2018-12-05 paper/ms.tex
fixed 1/(pi^2) factor in tau_e timescale
per equation 2018/12/04.3(1)

@@
 
-%FIXME
-Could the eccentricity truly be non-zero?  The classical
-tidal-circularization theory has the eccentricity shrinking with a
-$1/e$-folding time of \citep[{\it e.g.}, Equation~1
-of][]{dobbs-dixon_spin-orbit_2004}
-\begin{equation}
-  \tau_e
-  = \frac{e}{|\dot{e}|}
-  = \frac{2\pi Q_{\rm p}'}{63}
-  \left( \frac{a}{R_{\rm p}} \right)^5
-  \frac{M_{\rm p}}{M_\star}\, P .
-\end{equation}
-For WASP-4b, this gives an eccentricity damping timescale of $\tau_e =
-28.3 (Q_{\rm p}'/10^6) \, {\rm Myr}$.  The system age appears to be
+Could the eccentricity truly be non-zero?  For WASP-4b, the
+eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
+0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system age appears to be
 well over $1\,{\rm Gyr}$ \citep{winn_transit_2009}, so any initial

### 2018-12-06 src/is_it_doppler_shift.py
minor doppler check change

@@ -35,8 +35,8 @@ gamma_dot_limit = (gamma_dot_value - 2*gamma_dot_sig) * (u.m/u.s)/u.day
 
 dP_dt_preferred = gamma_dot_preferred * P / c.c
 dP_dt_limit = gamma_dot_limit * P / c.c
-dP_dt_measured = -3.854e-10
-dP_dt_measured_err = 5.6e-11
+dP_dt_measured = -3.83e-10
+dP_dt_measured_err = 3.9e-11
 
 print('\nfor WASP-4b')
 print('Knutson+ 2014 reported {:d} CPS data points, and used {:d} data sets'.
@@ -59,8 +59,8 @@ gamma_dot_limit = (gamma_dot_value - 2*gamma_dot_sig) * (u.m/u.s)/u.year
 
 dP_dt_preferred = gamma_dot_preferred * P / c.c
 dP_dt_limit = gamma_dot_limit * P / c.c
-dP_dt_measured = -4.2e-10
-dP_dt_measured_err = 4.6e-11
+dP_dt_measured = -3.83e-10
+dP_dt_measured_err = 3.9e-11
 
 print('\nfor WASP-4b')
 print('Husnoo+ 2012 reported {:d} data points, from {:d} data sets (HARPS)'.

### 2018-12-06 src/kozai_companion_mass_bound.py
add Bailey & Goodman 2018 bounds calculation

@@ -1,0 +1,27 @@
+# -*- coding: utf-8 -*-
+"""
+calculate the required mass for a companion if Kozai-Lidov oscillations
+are pumping WASP-4b's eccentricity
+"""
+from __future__ import division, print_function
+import numpy as np
+
+from astropy import units as u, constants as const
+
+# Petrucci+ 2013, table 3, for all parameters
+a_b = 0.0228*u.au
+M_star, R_star, M_b, R_b = (
+    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+)
+
+# This number is made up, but similar to Jupiter's from Juno (Wahl+ 2016).
+k_2b = 0.6
+
+#####################
+# begin calculation #
+#####################
+
+# from 2018/12/05.0, equation "*".
+prefactor = 10/3 * k_2b * M_star**2 * R_b**5 * a_b**(-5) * M_b**(-1)
+
+print('M_c > {:.2f} * (a_c/a_b)^(3/2)'.format(prefactor.to(u.Mearth)))

### 2018-12-06 src/rv_companion_mass_bound.py

@@ -1,0 +1,24 @@
+# -*- coding: utf-8 -*-
+"""
+calculate the required mass for a companion if Kozai-Lidov oscillations
+are pumping WASP-4b's eccentricity
+"""
+from __future__ import division, print_function
+import numpy as np
+
+from astropy import units as u, constants as const
+
+# Petrucci+ 2013, table 3, for all parameters
+a_b = 0.0228*u.au
+M_star, R_star, M_b, R_b = (
+    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+)
+
+# from equation (52) of the latex'd equation in Avery's email
+
+O_minus_C_RV = 15.2 * u.m/u.s       # table 3 of Triaud+ 2010.
+
+prefactor = O_minus_C_RV * (M_star*a_b/const.G)**(1/2)
+
+print('M_c < {:.2f} * (a_c/a_b)^(1/2) * f^(-1/2)'.
+      format(prefactor.to(u.Mearth)))

### 2018-12-06 paper/ms.tex
minor update
improved text; added companion dynamics limits

@@
 \author[0000-0002-0514-5538]{L. G. Bouma}
-\affiliation{
-    Department of Astrophysical Sciences,
-    Princeton University,
-    4 Ivy Lane, Princeton, NJ 08540, USA}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+
 \author[0000-0002-4265-047X]{J. N. Winn}
-\affiliation{
-    Department of Astrophysical Sciences,
-    Princeton University,
-    4 Ivy Lane, Princeton, NJ 08540, USA}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+
+\author[0000-0002-6939-9211]{T. Daylan}
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+
+\author[0000-0002-3481-9052]{K. G. Stassun}
+\affiliation{Vanderbilt University, Department of Physics \& Astronomy,
+6301 Stevenson Center Lane, Nashville, TN 37235, USA}
+\affiliation{Fisk University, Department of Physics, 1000 17th Avenue
+N., Nashville, TN 37208, USA}
+
+\author[0000-0002-7084-0529]{S. Kane}
+\affiliation{Department of Earth Sciences, University of California,
+Riverside, CA 92521, USA}
+
 %-------------------------------------
@@
 \author{G. R. Ricker} % grr@space.mit.edu WAITING
-\affiliation{
-  Department of Physics and Kavli Institute for Astrophysics and Space
-  Research, Massachusetts Institute of Technology, Cambridge, MA
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
 02139, USA }
 \author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
-\affiliation{
-  Department of Physics and Kavli Institute for Astrophysics and Space
-  Research, Massachusetts Institute of Technology, Cambridge, MA
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
 02139, USA }
 \author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
-\affiliation{
-  Harvard-Smithsonian Center for Astrophysics, 60 Garden Street,
-  Cambridge, MA 02138, USA }
+\affiliation{ Harvard-Smithsonian Center for Astrophysics, 60 Garden
+Street, Cambridge, MA 02138, USA }
 \author{S. Seager} % seager@mit.edu WAITING
-\affiliation{
-  Department of Earth, Atmospheric, and Planetary Sciences,
+\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
 Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
 \author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
-\affiliation{
-  NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
 %-------------------------------------
@@
 %\\ \textcolor{red}{3 SPOC coauthors:}
-\author{SPOC1}
-\author{SPOC2}
-\author{SPOC3}
+\author{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ SETI Institute, Mountain View, CA, 94043}
+\author{B. Wohler} % bill.wohler@nasa.gov WAITING
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ SETI Institute, Mountain View, CA, 94043}
+\author{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ SETI Institute, Mountain View, CA, 94043}
 %\\ \textcolor{red}{3 POC coauthors:}
@@
   If the timing variations are astrophysical,
-  major contributors to the period change could include either
-  apsidal precession, or tidal decay.
+  major contributors to the period change could include both
+  apsidal precession and tidal decay.
   The Doppler shift from WASP-4's acceleration towards us could
@@
 
-Long-term monitoring of hot Jupiter transit and occultation times
-should eventually reveal two distinct processes.  First, whenever the
-summed spin and orbital angular momenta in a hot Jupiter system are
-below a particular critical value, the system will be unstable to
-tidal decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most
-hot Jupiters satisfy this condition, so their orbits should be
-shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate
-of decay is uncertain, and depends on how friction dissipates energy
-carried by gravitational tides (\citealt{ogilvie_tidal_2014} gives a
-review).  If we could directly measure orbital decay rates, it would
-inform our understanding of tidal dissipation.  This is important
-because the dissipation rates determine the fate of the system.
-During its inspiral, if the planet fills its Roche lobe below the
-stellar surface, the drag force from the stellar envelope accelerates
-the inspiral, leading to a `direct-impact' merger
+Long-term monitoring of hot Jupiter transit and occultation times should
+eventually reveal two distinct processes.  First, whenever the summed
+spin and orbital angular momenta in a hot Jupiter system are below a
+particular critical value, the system will be unstable to tidal decay
+\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
+satisfy this condition; their orbits should be shrinking
+\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay is
+uncertain, and depends on how friction dissipates energy carried by
+gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
+we could directly measure orbital decay rates, it would inform our
+understanding of tidal dissipation.  This is important because the
+dissipation rates determine the fate of the system.  During its
+inspiral, if the planet fills its Roche lobe below the stellar surface,
+the drag force from the stellar envelope accelerates the inspiral,
+leading to a `direct-impact' merger
 \citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
-the planet fills its Roche lobe above above the stellar surface, then
-it could either be dynamically disrupted into an accretion disk or it
-could slowly transfer mass to the star \citep{metzger_optical_2012}.
-More broadly, the outcome of the eventual interaction between the
+the planet fills its Roche lobe above above the stellar surface, then it
+could either be dynamically disrupted into an accretion disk or it could
+slowly transfer mass to the star
+\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
+Further afield, the outcome of the eventual interaction between the
 Earth and red giant Sun also hinges on the efficiency of tidal
@@
 binaries \citep[{\it e.g.},][]{russell_notes_1939,
-schwarzschild_structure_1958}, and has been explored theoretically for
-transiting planets
+schwarzschild_structure_1958,borkovits_eclipse_2015}, and has been
+explored theoretically for transiting planets
 \citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
@@
 be small, there are many mechanisms that could pump them away from
-circularity (see \S~\ref{sec:apsidal_precession}, and the
-possibilities considered by \citealt{bailey_understanding_2019}).  For
-hot Jupiters, the apsidal precession is usually dominated by the
-quadrupolar distortion of the planet by the star's tidal force
-\citep{ragozzine_probing_2009}.  If we measured apsidal precession in
-a hot Jupiter, we thus could convert the observed precession rate into
-the planet's Love number, giving us a third parameter along with the
-mass and radius with which to describe the interior structure of the
-planet \citep[{\it e.g.},][who performed a similar procedure for
-HAT-P-13b]{batygin_determination_2009}.  Measuring small
-eccentricities and understanding their origin would also help us
+circularity (see \S~\ref{sec:apsidal_precession}, and
+\citealt{bailey_understanding_2019}).  For hot Jupiters, the apsidal
+precession is usually dominated by the quadrupolar distortion of the
+planet by the star's tidal force \citep{ragozzine_probing_2009}.  If we
+measured apsidal precession in a hot Jupiter system, we could thus use
+the observed precession rate to compute the planet's Love number, giving
+a third parameter along with the mass and radius with which to describe
+the planet's interior structure \citep[{\it e.g.},][who performed a
+similar procedure for HAT-P-13b]{batygin_determination_2009}.  Measuring
+small eccentricities and understanding their origin would also help us
 understand the dynamical histories and interior structures of hot
@@
 
-% Stars hosting hot Jupiters tend to spin faster than other stars of
-% similar age and mass
-% \citep{penev_hats-18_2016,penev_empirical_2018}.  This is expected
-% because when the orbital frequency is larger than the spin rotation
-% frequency, tides donate angular momentum from the planet to the
-% star.  The reverse process occurs between the Earth and Moon,
-% leading to the Moon's slowly increasing separation from Earth of
-% $X.XX$cm~yr$^{-1}$ (CITE lunar ranging experiment).  The hot Jupiter
-% star-planet separation distribution is best reproduced by models
-% that account for tides~\citep{cameron_hierarchical_2018}.
-
-The most convincing signs seen to date of either orbital decay or apsidal
-precession have been in WASP-12b.
+While numerous indirect studies have called attention to
+population-level effects of tidal decay
+\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
+the most convincing direct evidence to date for either orbital decay or
+apsidal precession has come from WASP-12b.
 \citet{maciejewski_departure_2016} showed that the transit times of
-WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
-and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
-\citet{patra_2017} presented new transit and occultation times for
-the system, agreed that curvature was present, and found
-$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
+WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence, and
+measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
+\citet{patra_2017} presented new transit and occultation times for the
+system, and found $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
 Assuming that the period change was caused entirely by tidal decay,
-\citet{patra_2017} found values for the stellar tidal quality factor
-of $Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
+\citet{patra_2017} found values for the stellar quality factor of
+$Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
 explained by standard equilibrium tide
@@
 \citet{weinberg_tidal_2017} found though that if WASP-12 were a
-subgiant, tidal dissipation due to a dynamical tide might be
-sufficient to produce the required level of friction.
+subgiant, tidal dissipation due to a dynamical tide might be sufficient
+to produce the required level of friction.
 \cite{bailey_understanding_2019} concurred, but noted that the
-observational constraints on WASP-12 itself favor a main-sequence
-star, not a subgiant.  Thus, more data are needed to determine the
-cause of the timing variations in WASP-12b.
-
-% Tidal decay of hot Jupiters has been invoked to explain certain
-% properties of the ensemble of star-planet systems. For example, the
-% scarcity of gas giants with periods less than a day is suggestive of
-% orbital decay \citep[see,
-%   e.g.][]{Jackson+2008,Hansen2010,Penev+2012,Ogilvie2014}. The
-% anomalously rapid rotation of some hot-Jupiter host stars has been
-% attributed to transfer of the planet's orbital angular momentum
-% \citep{Penev+2016}. The absence of hot Jupiters around subgiant stars
-% may be caused by an acceleration of orbital decay when a star leaves
-% the main sequence
-% \citep{VillaverLivio2009,Hansen2010,SchlaufmanWinn2013}.  Tidal decay
-% might also be responsible for the lower occurrence of close-in planets
-% around rapidly rotating stars \citep{TeitlerKonigl2014}, or the
-% realignment of stars and their planetary orbits
-% \citep{MatsakosKonigl2015}. However, direct evidence for orbital decay
-% has been lacking: there have been no clear demonstrations of a
-% long-term period decrease due to orbital decay \citep[see,
-%   e.g.,][]{Hoyer+2016,Wilkins+2017}.
-%
+observational constraints on WASP-12 itself favor a main-sequence star,
+not a subgiant.  More data are needed to determine the cause of the
+timing variations in WASP-12b.
 
 This work focuses on WASP-4b, which is also a plausible candidate for
-orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
-and mass $1.29 \pm 0.10\ M_{\rm Jup}$
-\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
-days, orbits its host star at a distance of $a/R_\star = 5.41$, and
-has been frequently observed over the past decade
-\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
-consistent with circular \citep[][but see
+orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$, and
+mass $1.29 \pm 0.10\ M_{\rm Jup}$ \citep{southworth_homogeneous_2011}.
+It has an orbital period of 1.34 days, orbits its host star at a
+distance of $a/R_\star = 5.41$, and has been frequently observed over
+the past decade \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The
+orbit is consistent with circular \citep[][but see
 \S~\ref{sec:apsidal_precession}]{beerer_secondary_2011,husnoo_observational_2012,bonomo_gaps_2017}.
-The host star is a G7V dwarf, with $R_\star = 0.91 \pm 0.12\ R_\odot$,
-$M_\star = 0.94 \pm 0.09\ M_\odot$, $T_{\rm eff} = 5436 \pm 34\ {\rm
+The host star is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
+$M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
 K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
-\citep{southworth_homogeneous_2011,petrucci_no_2013}.
+\citep{southworth_homogeneous_2011,petrucci_no_2013,doyle_accurate_2013}.
 
@@
 (\S~\ref{sec:timing}): a constant period, a decaying period, and a
-slightly eccentric orbit undergoing precession.  We rule out a
-constant period, but cannot firmly distinguish between the latter two
+slightly eccentric orbit undergoing precession.  We rule out a constant
+period, but cannot firmly distinguish between the latter two
 possibilities.  Each would have interesting implications
@@
 systematic time system offset causes the observed variation
-(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
-further monitoring of the system (\S~\ref{sec:future}).
+(Appendix~\ref{sec:verify_tess}), and conclude by advocating for further
+monitoring of the system (\S~\ref{sec:future}).
 
@@
 20, during the spacecraft's second sector of science operations.  The
-star is designated by the TESS Input
-Catalog~\citep[TIC;][]{stassun_TIC_2018} as TIC 402026209.
+star is designated by the TESS Input Catalog as TIC 402026209
+\citep{stassun_TIC_2018}.
 
 The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
-array at a cadence of 2 minutes.  It was included on the
-``short-cadence'' target list thanks to the Guest Investigator
-programs of J.\ Southworth and S.\ Kane (G011112 and G011183
-respectively).  After being downlinked through the Deep Space Network,
-the data were processed through the Science Processing Operations
-Center (SPOC) pipeline at NASA Ames~\citep{jenkins_tess_2016}.  One
-important step in this processing was converting from spacecraft time
-to barycentric time.  SPOC then reports timestamps in the ``BTJD''
-time system, where ${\rm BTJD} = {\rm BJD} - 2457000$, and all times
-are in the TDB reference (see \citealt{urban_explanatory_2012} for
-explanations of time systems).  The lightcurves were then vetted and
-released by the MIT TESS Science Office to the Mikulski Archive for
-Space Telescopes~\citep{ricker_tess_alerts_2018}.
+array every 2 minutes.  It was included on the ``short-cadence'' target
+list thanks to the Guest Investigator programs of J.\ Southworth and S.\
+Kane (G011112 and G011183 respectively).  After being downlinked through
+the Deep Space Network, the data were processed through the Science
+Processing Operations Center (SPOC) pipeline at NASA
+Ames~\citep{jenkins_tess_2016}.  One important step in this processing
+was converting from spacecraft time to barycentric time.  SPOC then
+reports timestamps in ``BTJD'', where ${\rm BTJD} = {\rm BJD} -
+2457000$, BJD is the barycentric julian date, and all times are in the
+TDB reference (see \citealt{urban_explanatory_2012} for explanations of
+time systems).  The lightcurves were then vetted and released by the MIT
+TESS Science Office to the Mikulski Archive for Space
+Telescopes~\citep{ricker_tess_alerts_2018}.
 
 We begin our analysis with the Presearch Data Conditioning (PDC) light
-curves from MAST.  The processing steps used to find an optimal
-aperture (for WASP-4, a $3\times 3$ pixel array) and the PDC-MAP
-algorithm are described by \citet{smith_kepler_apertures_2017} and
+curves from MAST.  The steps used to find an optimal aperture (for
+WASP-4, a $3\times 3$ pixel array) and the PDC-MAP algorithm are
+described by \citet{smith_kepler_apertures_2017} and
 \citet{smith_kepler_PDC_2017} respectively.
 
-Our processing steps were then as follows.  First, we removed all
-points with non-zero quality flags.  This removes data that is
-contaminated by coarse spacecraft pointing, cosmic rays, and other
-annoyances \citep{tess_data_product_description_2018}.  We then
-filtered out known bad observing windows.  The data closest to
-spacecraft perigee typically show ramp-like systematics, so we clipped
-out the first and last hour of both orbits.  Next, we focused on the
-momentum dumps. As described in the data release
+We then processed the lightcurve as follows.  First, we removed all points
+with non-zero quality flags.  This removes data contaminated by
+coarse spacecraft pointing, cosmic rays, and other annoyances
+\citep{tess_data_product_description_2018}.  We then filtered out known
+bad observing windows.  The data closest to spacecraft perigee typically
+show ramp-like systematics, so we clipped out the first and last hour of
+both orbits.  Next, we focused on the momentum dumps. As described in
+the data release
 notes\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}},
 during the first two TESS sectors, every 2.5~days the spacecraft's
-reaction wheels were reset to maintain pointing stability.  These
-events were assigned the quality flags corresponding to ``Reaction
-Wheel Desaturation Event'' and ``Manual Exclude''.  In the WASP-4
-lightcurve, these flags were simultaneously set for 54 distinct
-cadences, and there were 10 momentum dumps, averaging about 10 minutes
-of flagged data per dump.  Simply out of caution, we clipped out an
-additional 10 minutes before and after every momentum dump.  Before
-applying any filters, we had 19737 data points. After applying all the
-filters, we were left with 18165 measurements, or 92\% of the original
-data.
+reaction wheels were reset to maintain pointing stability.  These events
+were assigned quality flags corresponding to ``Reaction Wheel
+Desaturation Event'' and ``Manual Exclude''.  For WASP-4,
+these flags were simultaneously set for 54 distinct cadences, and there
+were 10 momentum dumps, averaging about 10 minutes of flagged data per
+dump.  Simply out of caution, we clipped out an additional 10 minutes
+before and after every momentum dump.  Before applying any filters, we
+had 19737 data points. After applying all the filters, we were left with
+18165 measurements, or 92\% of the original data.
 
@@
 dividing out the median flux.  We then converted the timestamps from
-BTJD to BJD by adding the appropriate offset of 2457000
-days~\citep{tess_data_product_description_2018}.  Many of these and
+BTJD to BJD by adding the appropriate 2457000 day offset
+\citep{tess_data_product_description_2018}.  Many of these and
 subsequent processing steps were performed using
@@
 ``flatten'' the lightcurves at this stage, as is often done with {\it
-e.g.}, a median filter, spline, or gaussian process regression.  Such
-procedures are asymmetric about the transit minimum and can skew the
-transit time measurement, and their effect on the measurement
-uncertainties is difficult to capture.
+e.g.}, a spline or gaussian process regression.  Such procedures are
+asymmetric about the transit minimum and can skew the transit time
+measurement without being captured by the measurement uncertainties.
 
@@
     \caption{
+      %TODO: final number
         {\bf TESS saw WASP-4b transit earlier than expected.}
         Points are differences between observed transit times and the
-        constant-period ephemeris predicted by previous transit
-        measurements.  Black points with circle markers are the more
-        precise half of literatures times; gray points are the less
-        precise half.  The blue curves represent the $\pm 1\sigma$
-        credible interval around the constant-period ephemeris.  The
-        binned TESS observation (red point) arrived $77.8 \pm 10.7\
-        {\rm seconds}$ before it was expected.
+        constant-period ephemeris we fit to literature transits.
+        Black points are the more precise half of literatures times;
+        gray points are the less precise half.  The blue curves
+        represent the $\pm 1\sigma$ credible interval around the
+        constant-period ephemeris.  The binned TESS observation (red
+        point) arrived $77.8 \pm 10.7\ {\rm seconds}$ before it was
+        expected.
         \label{fig:arrived_early}
@@
 
-Using the times, flux measurements, and errors determined above, we
-began our measurement procedure by using the Box Least Squares (BLS)
-method to estimate the transit epoch, period, and duration
-\citep{kovacs_box-fitting_2002}.  From the BLS parameters, we isolated
-each transit to a rough estimate of its midtime, $\pm$10 transit
-durations.  In each transit window, we then simultaneously fit for a
-model transit, plus a local linear trend.  Our transit model uses the
-analytic formulae calculated by \citet{mandel_analytic_2002}, and
-implemented by \citet{kreidberg_batman_2015}.
-
-To measure the times, we allowed four free parameters per transit: the
+Using the times, flux measurements, and errors determined above, we ran
+Box Least Squares (BLS) to estimate the transit epoch, period, and
+duration \citep{kovacs_box-fitting_2002}.  From the BLS parameters, we
+isolated each transit to within $\pm$10 transit durations of its
+midtime.  In each transit window, we then simultaneously fit for a model
+transit, plus a local linear trend.  Our transit model uses the analytic
+formulae calculated by \citet{mandel_analytic_2002} and implemented by
+\citet{kreidberg_batman_2015}.
+
+We allowed four free parameters per transit: the
 time of mid-transit $t_{\rm tra}$, the planet-to-star radius ratio
-$R_{\rm p}/R_\star$, and the slope and intercept of the line. Our
-priors were wide uniform distributions centered on the BLS estimates
-for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope of zero and
-an intercept of unity for the line.
-
+$R_{\rm p}/R_\star$, and the slope and intercept of the line.
+% Our priors were wide uniform distributions centered on the BLS
+% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope of
+% zero and an intercept of unity for the line.
 We fixed the remaining transit parameters as follows.  We set the
-eccentricity to zero, and the longitude of periastron to $\pi/2$.  For
-the limb darkening, we assumed a fixed quadratic law, with
-coefficients interpolated from the tables computed by
-\citet{claret_limb_2017}.  For the period, $a/R_\star$, and
-inclination, we adopted the values from~\citet{petrucci_no_2013}.
+eccentricity to zero, and the longitude of periastron to $\pi/2$.  We
+assumed a quadratic limb-darkening law, with coefficients interpolated
+from the \citet{claret_limb_2017} tables.  For the period, $a/R_\star$,
+and inclination, we adopted the values from~\citet{petrucci_no_2013}.
 These are in reasonable agreement with the parameters reported by
@@
 
-Having defined all the free and fixed parameters, we calculated an
-initial guess for the free parameters by maximizing the likelihood.
-We then sampled over the posterior using the algorithm proposed by
-\citet{goodman_ensemble_2010}, and implemented by
+Having defined the free and fixed parameters, we calculated an initial
+guess for the free parameters by maximizing the likelihood.  We then
+sampled over the posterior using the algorithm proposed by
+\citet{goodman_ensemble_2010} and implemented by
 \citet{foreman-mackey_emcee_2013}.  During this process, we used the
 \texttt{corner} software to check that our posteriors were being
-adequately sampled~\citep{corner_2016}.
+well-sampled~\citep{corner_2016}.
 
@@
 % TODO: final numbers
-To check that the measured uncertainties are plausible, we computed
-the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
-transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
-freedom.  Examining the residuals by eye showed that the error
-variance was being overestimated, so we multiplied the measured TESS
-errors by a fudge factor $f=0.73$, forcing a reduced $\chi^2$ of
-unity.  This reduced the mean transit time uncertainty from $\sim$30
-seconds per transit to $\sim$22 seconds per transit.
+To check that the measured uncertainties are plausible, we computed the
+reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
+times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
+Examining the residuals by eye showed that the error variance was being
+overestimated, so we multiplied the measured TESS errors by a factor
+$f=0.73$, forcing a reduced $\chi^2$ of unity.  This reduced the mean
+transit time uncertainty from $\sim$30 seconds per transit to $\sim$22
+seconds per transit.
 
@@
 their uncertainties, and also includes the previously reported times
-that are analyzed in \S~\ref{sec:timing}.
+analyzed in \S~\ref{sec:timing}.
 
@@
 radius ratio, we find $R_{\rm p}/R_\star=0.1538\pm0.0013$, in
-agreement with the determinations by {\it e.g.},
-\citet{wilson_wasp-4b_2008}, \citet{gillon_improved_2009},
-\citet{winn_transit_2009}, and \citet{southworth_high-precision_2009}.
+agreement with previous determinations
+\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
 
@@
 \citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
-noise\footnote{github.com/lgbouma/tnm} of $410\,{\rm ppm}\,{\rm
+noise\footnote{\url{github.com/lgbouma/tnm}} of $410\,{\rm ppm}\,{\rm
 hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
-expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$
-in quadrature.  The predicted zodiacal background contribution is
-$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the zodiacal background
-model calculated by \citet{winn_photonflux_2013} and used in the
-\citet{Sullivan_2015} planet detection simulations.  The predicted
-level of zodiacal noise seems to have been an overestimate, given the
-observed RMS of WASP-4.  We leave a broader assessment of TESS's
-photometric performance to future work.
+expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
+quadrature.  The predicted zodiacal background contribution is
+$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
+\citet{winn_photonflux_2013} model, which was used in the
+\citet{Sullivan_2015} planet detection simulations.  The predicted level
+of zodiacal noise seems to have been an overestimate, given the observed
+RMS of WASP-4.  A more detailed assessment of TESS's photometric
+performance is beyond the scope of this work.
 
@@
         Points are differences between observed transit times and a
-        linear ephemeris fit to the entire timing dataset.  The
-        constant period model is a poor fit to the observed times
-        (gray line).  Black points are the more precise half of
-        measured times; gray points are the less precise half.  The
-        blue curve is the best-fit model with a constant period
-        derivative.  The orange curve is the best-fit model assuming
-        apsidal precession causes the observed timing deviations.
+        linear ephemeris fit to all the times.  The constant period
+        model is a poor fit to the observed times (gray line).  Black
+        points are the more precise half of measured times; gray points
+        are the less precise half.  The blue curve is the best-fit model
+        with a constant period derivative.  The orange curve is the
+        best-fit model assuming apsidal precession causes the observed
+        timing deviations.
         \label{fig:times}
@@
 We use the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point.  We took their BJD$_{\rm TT}$ times to be equivalent
-to times in BJD$_{\rm TDB}$, since the two time systems differ
-periodically by only milliseconds \citep{urban_explanatory_2012}.  We
-verified that their times, as well as all other times in subsequent
-analysis, were on the BJD$_{\rm TDB}$ time standard using the
-calculator provided by~\citet{eastman_achieving_2010}.  We set the
-zero-point epoch to be as close as possible to the average of the
-transit times, weighted as the inverse variance of the transit
-midtime. This minimizes the covariance between the transit epoch and
-the period \citep{gibson_gemini_2013}.
-
-Important contributions to the transit timing data are as follows.
-The earliest epoch in our fit is provided by the EulerCam lightcurve
-observed by \citet{wilson_wasp-4b_2008}.  The second, more precise,
-epoch comes from $z$-band photometry acquired by
-\citet{gillon_improved_2009} at the VLT with FORS2.  Subsequent epochs
-came from \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009}
-recorded the GPS clock time in UT reference. They then converted this
-to BJD in the TDB reference for their reported midtimes.},
+starting point.  We took their BJD$_{\rm TT}$ times to be equivalent to
+times in BJD$_{\rm TDB}$, since the two time systems differ periodically
+by only milliseconds \citep{urban_explanatory_2012}.  We verified that
+their times, as well as all others, were on the BJD$_{\rm TDB}$ time
+standard using the calculator provided
+by~\citet{eastman_achieving_2010}.  We set the zero-point epoch to be as
+close as possible to the average of the transit times, weighted as the
+inverse midtime variance. This minimizes the covariance between the
+transit epoch and the period \citep{gibson_gemini_2013}.
+
+Important contributions to the timing dataset are as follows.  The
+earliest epoch in our fit is from the EulerCam lightcurve observed by
+\citet{wilson_wasp-4b_2008}.  The second, more precise, epoch comes from
+$z$-band photometry acquired by \citet{gillon_improved_2009} at the VLT
+with FORS2.  Subsequent epochs came from
+\citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
+the GPS clock time in UT reference. They then converted this to BJD in
+the TDB reference for their reported midtimes.},
 \citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
@@
 %the MPG/EOS-2.2~m at La Silla.
-\citet{hoyer_tramos_2013} reported nine transit observations taken
-with the the Y4KCAM on the SMARTS 1-m telescope, and another three
-with the SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013}
-also homogeneously redetermined the transit times from all previous
-studies.  Though their latest two epochs did fall slightly below the
-prediction of a linear ephemeris (Panel C of their Figure~6), they did
-not find any statistical need for a quadratic function to fit the O-C
-values.
+\citet{hoyer_tramos_2013} reported nine transit observations taken with
+the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
+SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013} also
+homogeneously redetermined the transit times from all previous studies.
+Though their latest two epochs fell slightly below the prediction of a
+linear ephemeris (Panel C of their Figure~6), they did not find a need
+to fit a quadratic function to the O-C values.
 
@@
 \citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
-using HST's WFC3 over five orbits in transit, and five orbits in
-secondary eclipse.  We adopt only their transit time, since the epoch
-of occultation was not allowed to be a free parameter in their fit.
-\citet{huitson_gemini_2017} acquired optical transmission spectra with
-the 8.1-m Gemini South telescope over 2011 to 2014, one transit per
-season.  The per-point standard deviation of their lightcurves is a
-few hundred parts per million, less than a factor of 3 away from the
-photon noise limit.  The uncertainties on their reported transit
-times~--~on average 5.6~seconds~--~ are small, but possible given the
-quality of their data~\citep[cf.][]{carter_analytic_2008}.
+using HST's WFC3 in both transit and secondary eclipse.  We adopt only
+their transit time, since the epoch of occultation was not fit as a free
+parameter.  \citet{huitson_gemini_2017} acquired optical transmission
+spectra with the 8.1-m Gemini South telescope over 2011 to 2014, one
+transit per season.  The per-point standard deviation of their
+lightcurves is a few hundred parts per million, less than a factor of 3
+away from the photon noise limit.  The uncertainties on their reported
+transit times~--~on average 5.6~seconds~--~ are small, but possible
+given the quality of their data~\citep[cf.][]{carter_analytic_2008}.
 
@@
 There are fewer available occultation times.
-\citet{beerer_secondary_2011} observed two occultations of WASP-4b
-using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
+\citet{beerer_secondary_2011} observed two occultations of WASP-4b using
+warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
 \citet{caceres_ground-based_2011} detected an occultation from the
 ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
-specifying the time standard.  We correspondingly added $69.184$
-seconds of uncertainty to their reported errors, in quadrature.
-Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}
-acquired occultation data with HST but kept the epoch as a fixed
-parameter.  \citet{zhou_secondary_2015} performed a similar analysis
-with occultation data from the Anglo-Australian Telescope, focusing on
-the eclipse depth, but also fitting for $e\cos\omega$.  We converted
-their $e\cos\omega$ results into a midtime, using their ephemeris and
-the standard formula \citep[{\it
+specifying the time standard.  We correspondingly added $69.184$ seconds
+of uncertainty to their reported errors, in quadrature.  Finally, as
+previously mentioned, \citet{ranjan_atmospheric_2014} acquired
+occultation data with HST but kept the epoch as a fixed parameter.
+\citet{zhou_secondary_2015} performed a similar analysis with
+occultation data from the Anglo-Australian Telescope, focusing on the
+eclipse depth, but also fitting for $e\cos\omega$.  We converted their
+$e\cos\omega$ results into a midtime, using their ephemeris and the
+standard formula \citep[{\it
 e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
@@
 $
-  Q_\star \sim \mathcal{U}[10^3, 10^9],
+  Q_\star' \sim \mathcal{U}[10^3, 10^9],
 $
-for $Q_\star$ the ``quality factor'' corresponding to the star's
-response against tidal forcing, defined in \S~\ref{sec:implications}.
-For the precession model, we took the same priors in epoch and
-sidereal period, and drew the remaining parameters from wide uniform
-distributions.
-
+for $Q_\star'$ the modified quality factor defined in
+\S~\ref{sec:implications}.  For the precession model, we took the same
+priors in epoch and sidereal period, and drew the remaining parameters
+from wide uniform distributions.
+
+%TODO: final numbers
 Figure~\ref{fig:times} shows the residuals with respect to the linear
-model.
+model.  The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of
+freedom.  The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57
+degrees of freedom.  The median MCMC precession fit has $\chi^2 = 51.0$
+and 55 degrees of freedom.
+
 %TODO: final numbers
-The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of freedom.
-The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57 degrees of freedom.
-The median MCMC precession fit has $\chi^2 = 51.0$ and 55 degrees of freedom.
-
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$. In other words, assuming the
-linear model is true, there is a $10^{-23}$ probability that simply by
-chance we would see data that favors the quadratic model more strongly
-than the observations.  Clearly, a more likely possibility would be
-systematic errors in one of the most important sets of times: either
-the earliest times, the Huitson et al.\ Gemini times (epoch numbers
-44, 322, 591, and 851), or the TESS times. 
-We have taken every precaution against systematic offsets in the
-TESS dataset (Appendix~\ref{sec:verify_tess}), and Huitson et al.\ 
-stand by their reported times (Huitson, priv. comm.).
-Since the linear model provides a poor fit to the data, we discard it
-from further consideration.
+corresponds to $p \approx 10^{-23}$. In other words, assuming the linear
+model is true, there is a $10^{-23}$ probability that simply by chance
+we would see data that favors the quadratic model more strongly than the
+observations.  A more likely possibility is some systematic error being
+present in one of the most important sets of times: either the earliest
+times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591, and
+851), or the TESS times.  We have taken every precaution against
+systematic offsets in the TESS dataset (Appendix~\ref{sec:verify_tess}),
+and Huitson et al.\ stand by their reported times (Huitson, priv.
+comm.).  Since the linear model provides a poor fit to the data, we
+discard it from further consideration.
 
 %TODO: final numbers
-The quadratic model fits better than the precession model.
-It is favored over the precession fit by $\Delta \chi^2 = 3.0$,
-and has two fewer free parameters.
-Another heuristic useful for model comparison is the Bayesian
-Information Criterion (BIC),
+The quadratic model fits better than the precession model.  It is
+favored over the precession fit by $\Delta \chi^2 = 3.0$, and has two
+fewer free parameters.  Another useful heuristic for model comparison is
+the Bayesian Information Criterion (BIC),
 \begin{equation}
@@
 for $k$ the number of free parameters, and $n$ the number of data
-points. For us, $n=60$.
-The difference in the BIC for the precession and decay models is
-$\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm BIC}_{\rm quad} = 11$,
-corrsonding to a Bayes factor of $\approx 4\times10^{4}$.
-In the language of~\citet{kass_bayes_1995}, this is ``decisive
-evidence'' for the quadratic model describing the data better than the
-precession model.
-Using the Akaike Information Criterion similarly favors the orbital
-decay model, by $\Delta {\rm AIC} = 7$.
+points. For us, $n=60$.  The difference in the BIC for the precession
+and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
+BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor of $\approx
+4\times10^{4}$.  In the language of~\citet{kass_bayes_1995}, this is
+``decisive evidence'' for the quadratic model describing the data better
+than the precession model.  Using the Akaike Information Criterion
+similarly favors the orbital decay model, by $\Delta {\rm AIC} = 7$.
 
@@
 For comparison, the period derivative seen by
-\citet{maciejewski_departure_2016} and \citet{patra_2017} in
-WASP-12b is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
-Assuming the quadratic model, WASP-4b is thus decaying at
-about half the rate as WASP-12b.
+\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b is
+$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  Assuming the quadratic
+model, WASP-4b is thus decaying at about half the rate as WASP-12b.
 
@@
 
-The median parameters and their standard deviations are reported for
-all the models in Table~3.  In summary, a linear ephemeris does not
-describe the observed times.  Both orbital decay and precession can
-fit the observed data.  However, the precession model has two extra
-free parameters, and is statistically disfavored.  Further
-observations will help discriminate between orbital decay and
-precession.
+The median parameters and their standard deviations are reported for all
+the models in Table~3.  To summarize, a linear ephemeris does not
+describe the observed times.  Both orbital decay and precession can fit
+the observed data.  However, the precession model has two extra free
+parameters, and is statistically disfavored.  Further observations will
+help discriminate between orbital decay and precession.
 
@@
 If the planet's spin and orbit are synchronized, the star is slowly
-rotating, and the eccentricity is small, then the semimajor axis and
+rotating, and the eccentricity is small, then the semi-major axis and
 eccentricity evolve as
@@
 \end{equation}
-For comparison, Jupiter has $Q_{\rm Jup}' \approx 1.4 \times 10^5$,
-based on the motion of the Galilean moons \citep{lainey_strong_2009}.
-For stars, different $Q_\star'$ values have been estimated in
-different contexts.  Studies of the observed eccentricity distribution
-in stellar binaries have yielded $10^5 \gtrsim Q_\star' \gtrsim
-10^{7}$ \citep[{\it
+This is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
+10^5$, based on the motion of the Galilean moons
+\citep{lainey_strong_2009}.  For stars, different $Q_\star'$ values have
+been estimated in different contexts.  Studies of the observed
+eccentricity distribution in stellar binaries have yielded $10^5 \gtrsim
+Q_\star' \gtrsim 10^{7}$ \citep[{\it
 e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
@@
 frequencies and amplitudes are different.  In fact, for the observed
-population of hot Jupiters, \citet{penev_constraining_2012} argued
-that $Q_\star' > 10^7$ would be required at some phase of hot Jupiter
+population of hot Jupiters, \citet{penev_constraining_2012} argued that
+$Q_\star' > 10^7$ would be required at some phase of hot Jupiter
 evolution at the 99\% level, or else the observed population would not
@@
 dissipation would be allowed if the hot Jupiters arrived later).
-Developing a generative model for the orbital separation distribution
-of the hot Jupiter population, \citet{cameron_hierarchical_2018}
-similarly found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.
+Developing a generative model for the orbital separation distribution of
+the hot Jupiter population, \citet{cameron_hierarchical_2018} similarly
+found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.
 
@@
 typical orbit-averaged predictions.  It might be possible that we are
-observing at a special time, when the planet is near resonance with
-the star's fluid response, and the energy dissipation is exceptionally
-large \citep{ogilvie_tidal_2014,essick_orbital_2016}.  Tidal
-dissipation rates might also be increased if the star is turning off
-the main sequence.  WASP-4 shows little evidence for this, so it is
-unclear that tidal decay is a well-motivated explanation for its
-changing period.
+observing at a special time, when the planet is near resonance with the
+some stellar oscillation mode, and the energy dissipation is
+exceptionally large \citep{ogilvie_tidal_2014,essick_orbital_2016}.
+Tidal dissipation rates might also be increased if the star is turning
+off the main sequence.  WASP-4 shows little evidence for this; it is
+unclear that tidal decay could explain its changing period.
 
@@
 
-Let us instead assume that the entire timing variation can be
-explained as being a portion of an apsidal precession cycle.
-The required orbital eccentricity is then
+Let us instead assume that the entire timing variation can be explained
+as being a portion of an apsidal precession cycle.  The required orbital
+eccentricity is then
 \begin{equation}
@@
 \end{equation}
-One full precession period would thus take roughly 24 years, about
-twice the current observing baseline.
-
-\citet{ragozzine_probing_2009} calculated apsidal precession periods
-for hot Jupiters ranging between of order 10 and 100 years.  They
-highlighted that for many hot Jupiters, including WASP-4b, over 90\%
-of the apsidal precession comes from the tidal bulge of the planet.
+One full precession period would take about 24 years, or twice the
+current observing baseline.
+
+\citet{ragozzine_probing_2009} calculated apsidal precession periods for
+hot Jupiters ranging between of order 10 and 100 years.  They
+highlighted that for many hot Jupiters, including WASP-4b, over 90\% of
+the apsidal precession comes from the tidal bulge of the planet.
 Precession from general relativity, the planet's rotational bulge, and
-the star's rotational and tidal bulges are all minor contributors.
-\citet{ragozzine_probing_2009} also calcuated that WASP-4b was
-expected to have a relatively short precession period, of 120 years.
-
-One implication of the study by \citet{ragozzine_probing_2009} is that
-if the planet has non-zero eccentricity, a measurement of
-$\dot{\omega}$ can be converted into a measurement of the planetary
-Love number, $k_{2,\rm p}$.  Following \citet{ragozzine_probing_2009}
-Equation 14, the implied planetary Love number for WASP-4b is
+the star's rotational and tidal bulges are minor contributors.
+\citet{ragozzine_probing_2009} also calcuated that WASP-4b was expected
+to have a relatively short precession period, of 120 years.
+
+\citet{ragozzine_probing_2009} also noted that if the planet has
+non-zero eccentricity, a measurement of $\dot{\omega}$ can be converted
+into a measurement of the planetary Love number, $k_{2,\rm p}$.
+From \citet{ragozzine_probing_2009} Equation 14, the implied
+planetary Love number for WASP-4b is
 \begin{equation}
@@
 For comparison, $k_{2,{\rm Jup}} \approx 0.59$
-(\citealt{wahl_tidal_2016}, though \citealt{ni_empirical_2018} give an
-estimate that is 10\% lower).  A uniform density sphere has $k_2 =
-1.5$.  Our reported uncertainties on $k_{2,{\rm p}}$ are large because
-the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$ are
-quite correlated, and the measured occultation times only barely
-constrain them (Figure~\ref{fig:future}).
+(\citealt{wahl_tidal_2016}, bug see \citealt{ni_empirical_2018}).  A
+uniform density sphere has $k_2 = 1.5$.  Our reported uncertainties on
+$k_{2,{\rm p}}$ are large because the eccentricity, reference time, and
+${\rm d}\omega/{\rm d}E$ are quite correlated, and the measured
+occultation times only barely constrain them (Figure~\ref{fig:future}).
 
@@
 
-Could the eccentricity truly be non-zero?  For WASP-4b, the
-eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
-0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system age appears to be
-well over $1\,{\rm Gyr}$ \citep{winn_transit_2009}, so any initial
-eccentricity should have been erased.
+Could the eccentricity truly be non-zero?  For WASP-4b, the eccentricity
+damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e = 0.29 (Q_{\rm
+p}'/10^5) \, {\rm Myr}$.  The system appears to be well over $1\,{\rm
+Gyr}$ \citep{winn_transit_2009} old, so any initial eccentricity should
+have been erased.
 
@@
 measurements, but they often come with caveats.  For WASP-4b, the
-strongest limit on its eccentricity comes from the secondary eclipses
-observed by \citet{beerer_secondary_2011}, which led to a $2\sigma$
-upper limit on $| e\cos\omega | $ of 0.0024.
+strongest limit on its eccentricity comes from
+\citet{beerer_secondary_2011}, who observed occultations and derived a
+$2\sigma$ upper limit on $| e\cos\omega | $ of 0.0024.
 % \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
@@
 Placing eccentricity limits stronger than $e< 10^{-3}$ has only been
-possible for a few hot Jupiters, and often with numerous false-starts.
-HD 189733b, for example, was initially reported to have a very small
+possible for a few hot Jupiters, and with numerous false-starts.
+HD 189733b, for example, was initially reported to have a non-zero
 eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
@@
 offset could be explained by detailed modeling of the planetary
-hot-spot, which can change occultation time offsets by up to a few
+hot-spot, which can shift occultation times by up to a few
 minutes \citep{williams_resolving_2006}.  Including an updated
 hot-spot model, \citet{agol_climate_2010} reported a 2$\sigma$ limit
-of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit
-yet for a hot Jupiter.  A similar saga occurred for HD 209458b, for
-which \citet{winn_measurement_2005} combined Spitzer secondary
-eclipses with RV data and reported a best-fit eccentricity of
-$e>0.0057$ at 90\% confidence.  They suspected that the data were
-consistent with zero eccentricity, and that the lower bound on $e$ was
-an artifact of imperfect limb-darkening model.  Stronger limits were
-eventually put in place by \citet{crossfield_spitzer_mips_2012}, who
-found $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is
-WASP-12b, for which the discovery RVs and photometry led to a non-zero
-but small eccentricity measurement \citep[][at
-$3\sigma$]{hebb_wasp-12b_2009}.  A prompt ground-based occultation
-measurement by \citet{lopez-morales_day-side_2010} concurred with weak
-statistical significance that a small eccentricity was favored, but
-further follow-up found occultation times consistent with circular
-orbits; $| e\cos\omega| < 0.0040 $
-\citep{campo_orbit_2011,croll_near-infrared_2011}.
-
-In summary, empirical constraints on whether hot Jupiters are
-circular at the $e \lesssim 10^{-3}$ level are only present for HD
-189733b and HD 209458b, from occultation timing.  Once the timing
-offsets are of order tens of seconds, eccentricities from occultation
-timing may have systematic uncertainties because of the hot-spot issue
-\citep{williams_resolving_2006,agol_climate_2010}.  With the sole
-exception of WASP-18b, radial velocity measurements have provided
-eccentricity constraints only at the $e > 10^{-2}$ level.  Thus,
-though empirical constraints on the shortest period hot Jupiters are
-typically consistent with circular orbits, they cannot rule out
-eccentricities at the level needed to contextualize WASP-4.
+of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit yet
+for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
+\citet{winn_measurement_2005} combined Spitzer secondary eclipses with
+RV data and reported a small finite eccentricity.  They suspected that
+the data were consistent with zero eccentricity, and that the lower
+bound on $e$ was an artifact of an imperfect limb-darkening model.
+\citet{crossfield_spitzer_mips_2012} eventually found stronger limited:
+$e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b, for
+which the discovery RVs and photometry gave a small finite eccentricity
+measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.  A prompt
+ground-based occultation measurement by
+\citet{lopez-morales_day-side_2010} concurred that a small eccentricity
+was favored with weak statistical significance. Further follow-up found
+occultation times consistent with circular orbits; $| e\cos\omega| <
+0.0040 $ \citep{campo_orbit_2011,croll_near-infrared_2011}.
+
+In summary, empirical constraints on the ellipticity of hot Jupiter
+orbits at the $e \lesssim 10^{-3}$ level are present only for HD 189733b
+and HD 209458b, from occultation timing.  Once the timing offsets are of
+order tens of seconds, eccentricities from occultation timing may have
+systematic uncertainties because of the hot-spot issue
+\citep{williams_resolving_2006,agol_climate_2010}.  With the exception
+of WASP-18b, radial velocity measurements have provided eccentricity
+constraints only at the $e > 10^{-2}$ level.  Empirical cannot rule out
+hot Jupiter eccentricities at the level needed to contextualize WASP-4.
 
@@
 
-For lack of empirical data, we now turn to theoretical mechanisms
-that could maintain small eccentricities.
+Lacking empirical data, we now turn to theoretical mechanisms that could
+produce small eccentricities.
 
 \paragraph{Neighboring companion}
-\citet{mardling_long-term_2007} considered the long-term tidal
-evolution of hot Jupiters with companions.  They pointed out that
-although the early phases of the two-planet eccentricity evolution
-would typically occur over millions of years, the final phase of the
-joint eccentricity evolution towards circularity would often occur on
-timescales several orders of magnitude longer than the circularization
-time (see their Figures~4 and 5).  The companion in their model is
-coplanar, and can have a mass down to an Earth-mass; the main
-requirement is that both the hot Jupiter and the outer companion start
-on eccentric orbits.
+\citet{mardling_long-term_2007} considered the long-term tidal evolution
+of hot Jupiters with companions.  They pointed out that although the
+early phases of the two-planet eccentricity evolution would occur
+quickly, the final phase of the joint eccentricity evolution towards
+circularity would often occur on timescales several orders of magnitude
+longer than the circularization time (see their Figures~4 and 5).  The
+companion in their model is coplanar, and can have a mass down to an
+Earth-mass; the main requirement is that both the hot Jupiter and the
+outer companion start on eccentric orbits.
 
@@
 Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
-In this case, the orbital plane of the companion would need to be
-inclined relative to that of the hot Jupiter by at least $\sin^{-1}
-\sqrt{2/5} \approx 39^\circ$.
+In this case, the orbital plane of the outer companion, ``c'', would
+need to be inclined relative to that of the hot Jupiter, ``b'', by at
+least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
 \citet{bailey_understanding_2019} point out how the joint requirements
 of Kozai-Lidov oscillations with a non-detection of an additional RV
-signal can be combined to produce lower and upper limits on the outer
-companion's mass and the semimajor axis ratio, $a_{\rm c}/a_{\rm b}$.
-For WASP-4, ...
-% TODO: check this mechanism
-% FIXME: take a look at RVs. rule this out, or understand where the
-% bounds are.
+signal can be combined to produce lower and upper limits on the
+companion's mass.  For the Kozai-Lidov mechanism to operate, from
+\citet{bailey_understanding_2019} Equation 20 we need
+\begin{equation}
+  M_{\rm c} > 7.7\,M_\oplus
+  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
+  \label{eq:kozai_bound}
+\end{equation}
+where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
+of WASP-4b and the hypothetical WASP-4c, and we have assumed $k_{2,{\rm
+b}}\approx 0.6$, for $k_{2,{\rm b}}$ WASP-4b's Love number.  For the RV
+signal of the companion to remain undetected, it would need to be in the
+residual $(O-C)_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
+\citet{triaud_spin-orbit_2010}.  Again following
+\citet{bailey_understanding_2019}, this implies
+\begin{align}
+  M_{\rm c} &<
+  (O-C)_{\rm RV}
+  \left( \frac{M_\star a_b}{G} \right)^{1/2}
+  \left( \frac{a_c}{a_b} \right)^{1/2}
+  f^{-1/2}
+  \nonumber
+  \\
+  &< 
+  24.2\,M_\oplus
+  \times 
+  \left( \frac{a_c}{a_b} \right)^{1/2}
+  f^{-1/2},
+  \label{eq:rv_bound}
+\end{align}
+for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto sin^2 i_{\rm c}$ a
+geometric prefactor that depends on the argument of periastron and
+inclination of the exterior companion
+(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
+transits along the line of sight, $f$ can be arbitrarily small, and both
+of the preceding limits can be easily satisfied.  While the mechanism
+could produce the needed eccentricity, it would be more satifying if
+long-term radial velocity monitoring showed an additional planet in the
+system.  The chances of one existing with a mass between 1-13\,$M_{\rm
+Jup}$ and orbital semi-major axis between 1-20\,AU is about 50\%, and a
+linear trend is currently present at $\sim 2\sigma$
+\citep{knutson_friends_2014}.
 
@@
 width (mass) of the convective region,  $\mu$ the reduced mass, $n$
-the orbital frequency, and $a$ the semimajor axis.  For WASP-4,
+the orbital frequency, and $a$ the semi-major axis.  For WASP-4,
 convection in the star is more important than convection in the
@@
 additional transits of WASP-4b in the early 2020s
-(Figure~\ref{fig:future}).  In order to rule between orbital decay and
-apsidal precession, occultation measurements in the near term, and
-also in the mid-2020s, will be helpful.  However it may take up to a
-decade to firmly distinguish the two models, through the combination
-of transit and occultation measurements.  These observations should be
-carried out, because they will lead to new knowledge; either a measure
-of the efficiency of tidal dissipation within the star, or the
-eccentricity and Love number of the planet.  Measuring either for a
-single system would be rather interesting.  In the coming decade,
-wide-field photometric surveys (TESS, HATPI, NGTS, PLATO) might even
-enable such measurements at a population level, which would be a good
-deal more interesting.
+(Figure~\ref{fig:future}).  High-precision ground-based transit
+measurements could also be useful.  In order to rule between orbital
+decay and apsidal precession, occultation measurements in the near term,
+and also in the mid-2020s, will be helpful.  Additionally, confirmation
+and extension of the linear trend suggested by
+\citet{knutson_friends_2014} would help in understanding whether an
+exterior companion in the system might explain any residual
+eccentricity.  These observations should be carried out, because they
+will lead to new knowledge; either a measure of the efficiency of tidal
+dissipation within the star, or the eccentricity and Love number of the
+planet.  Measuring either for a single system would be rather
+interesting.  In the coming decade, wide-field photometric surveys
+(TESS, HATPI, NGTS, PLATO) might even enable such measurements at a
+population level, which would be a good deal more interesting.
 

### 2018-12-06 paper/ms.tex
update tex w/ dynamics limits

@@
 where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
-of WASP-4b and the hypothetical WASP-4c, and we have assumed $k_{2,{\rm
-b}}\approx 0.6$, for $k_{2,{\rm b}}$ WASP-4b's Love number.  For the RV
-signal of the companion to remain undetected, it would need to be in the
-residual $(O-C)_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
+of WASP-4b and the hypothetical WASP-4c, and we have assumed WASP-4b's
+Love number is $k_{2,{\rm b}}\approx 0.6$.  For the RV signal of the
+companion to remain undetected, it would need to be in the residual
+$({\rm O-C})_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
 \citet{triaud_spin-orbit_2010}.  Again following
@@
   M_{\rm c} &<
-  (O-C)_{\rm RV}
+  ({\rm O-C})_{\rm RV}
   \left( \frac{M_\star a_b}{G} \right)^{1/2}
@@
   \\
-  &< 
+  M_{\rm c} &< 
   24.2\,M_\oplus
@@
 \end{align}
-for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto sin^2 i_{\rm c}$ a
-geometric prefactor that depends on the argument of periastron and
-inclination of the exterior companion
+for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto \sin^2 i_{\rm c}$
+a geometric prefactor that depends on the argument of periastron
+$\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
 (\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
-transits along the line of sight, $f$ can be arbitrarily small, and both
-of the preceding limits can be easily satisfied.  While the mechanism
-could produce the needed eccentricity, it would be more satifying if
-long-term radial velocity monitoring showed an additional planet in the
-system.  The chances of one existing with a mass between 1-13\,$M_{\rm
-Jup}$ and orbital semi-major axis between 1-20\,AU is about 50\%, and a
-linear trend is currently present at $\sim 2\sigma$
+transits along the line of sight, $f$ can be arbitrarily small, and
+both of the preceding limits can be easily satisfied.  This
+mechanism could produce the needed eccentricity. However it would be more
+satifying if long-term radial velocity monitoring showed an additional
+planet in the system.  The chances of one existing with a mass between
+1-13\,$M_{\rm Jup}$ and orbital semi-major axis between 1-20\,AU is
+about 50\%, and a linear trend is currently present at $\sim 2\sigma$
 \citep{knutson_friends_2014}.
@@
 to never have exactly zero eccentricity.  In particular, from
-\citealt{phinney_pulsars_1992} Equation (7.33),
+\citealt{phinney_pulsars_1992} Equation 7.33,
 \begin{equation}

### 2018-12-06 src/stellar_density_and_mass_estimate.py
add src/stellar_density_and_mass_estimate.py

@@ -1,0 +1,53 @@
+# -*- coding: utf-8 -*-
+"""
+estimate the stellar density, and mass.
+"""
+from __future__ import division, print_function
+import numpy as np
+
+from astropy import units as u, constants as const
+
+# use parameters from Table 4 of Hoyer+2013
+i = 88.52*u.deg
+Rp_by_Rs = 0.15445
+a_by_Rs = 5.463
+a_by_Rs_perr = 0.025
+a_by_Rs_merr = 0.020
+P = 1.33823204*u.day
+
+# eq 30 of Winn (2010).
+# note k^3 = 3e-3, and the density of jupiter is about equal to the density of
+# the sun, so neglect the k^3*rho_planet term.
+
+rho_star = 3*np.pi/(const.G*P**2) * a_by_Rs**3
+print(rho_star.cgs)
+
+rho_star_upper = 3*np.pi/(const.G*P**2) * (a_by_Rs+a_by_Rs_perr)**3
+print(rho_star_upper.cgs)
+
+rho_star_lower = 3*np.pi/(const.G*P**2) * (a_by_Rs-a_by_Rs_merr)**3
+print(rho_star_lower.cgs)
+
+print(
+    'rhostar = {:.3f} +({:.3f}) -({:.3f}) g/cm^3'.format(
+        rho_star.cgs.value,
+        (rho_star_upper-rho_star).cgs.value,
+        (rho_star-rho_star_lower).cgs.value
+    )
+)
+
+# Keivan value
+Rstar = 0.86*u.Rsun
+Rstar_err = 0.02*u.Rsun
+
+Mstar = (4*np.pi)/3*rho_star*Rstar**3
+Mstar_lower = (4*np.pi)/3*rho_star_lower*(Rstar-Rstar_err)**3
+Mstar_upper = (4*np.pi)/3*rho_star_upper*(Rstar+Rstar_err)**3
+
+print(
+    'Mstar = {:.3f} +({:.3f}) -({:.3f}) Msun'.format(
+        Mstar.to(u.Msun).value,
+        (Mstar_upper-Mstar).to(u.Msun).value,
+        (Mstar-Mstar_lower).to(u.Msun).value,
+    )
+)

### 2018-12-06 paper/ms.tex
update lists and acknowledgements
better EB apsidal precession reference

@@
 precession).  This effect has been studied extensively in eclipsing
-binaries \citep[{\it e.g.},][]{russell_notes_1939,
-schwarzschild_structure_1958,borkovits_eclipse_2015}, and has been
+binaries \citep[{\it e.g.},][]{
+schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}, and has been
 explored theoretically for transiting planets
@@
 signal can be combined to produce lower and upper limits on the
-companion's mass.  For the Kozai-Lidov mechanism to operate, from
+companion's mass.  For the Kozai-Lidov mechanism to operate at maximum
+efficiency, from
 \citet{bailey_understanding_2019} Equation 20 we need
@@
 \acknowledgements{
-L.G.B.\ acknowledges helpful discussions with F. Dai and V. Van Eylen. 
+L.G.B.\ is happy to acknowledge help with dynamics from A.  Bailey,
+and useful brainstorming with F. Dai and V. Van Eylen. 
 %

### 2018-12-06 paper/ms.tex
improvements to the text

@@
   $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
-  From TESS observations of WASP-6b, WASP-18b, and WASP-46b, we show
+  From TESS observations of WASPs-6b, -18b, and -46b, we show
   that a systematic offset between the TESS time stamps and the barycentric
@@
   apsidal precession and tidal decay.
+  The latter seems implausible on theoretical grounds, as it would
+  require a stellar quality factor $Q_\star' \approx 3\times10^4$.
   The Doppler shift from WASP-4's acceleration towards us could
@@
 spin and orbital angular momenta in a hot Jupiter system are below a
-particular critical value, the system will be unstable to tidal decay
+critical value, the system will be unstable to tidal decay
 \citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
@@
 the planet fills its Roche lobe above above the stellar surface, then it
-could either be dynamically disrupted into an accretion disk or it could
+could either be disrupted into an accretion disk or it could
 slowly transfer mass to the star
@@
 precession is usually dominated by the quadrupolar distortion of the
-planet by the star's tidal force \citep{ragozzine_probing_2009}.  If we
-measured apsidal precession in a hot Jupiter system, we could thus use
-the observed precession rate to compute the planet's Love number, giving
-a third parameter along with the mass and radius with which to describe
-the planet's interior structure \citep[{\it e.g.},][who performed a
-similar procedure for HAT-P-13b]{batygin_determination_2009}.  Measuring
-small eccentricities and understanding their origin would also help us
-understand the dynamical histories and interior structures of hot
-Jupiters.  \citep[{\it
+planet by the star's tidal force \citep{ragozzine_probing_2009}.  The
+observed precession rate in a hot Jupiter system can thus be used to
+compute the planet's Love number, giving a third parameter along with
+the mass and radius with which to describe the planet's interior
+structure \citep[{\it e.g.},][who performed a similar procedure for
+HAT-P-13b]{batygin_determination_2009}.  Measuring small
+eccentricities and understanding their origin might also help us
+understand the dynamical histories and inflated radii of hot Jupiters.
+\citep[{\it
 e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}
@@
 \citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
-the most convincing direct evidence to date for either orbital decay or
-apsidal precession has come from WASP-12b.
+the most convincing direct evidence to date for either orbital decay
+or apsidal precession has come from WASP-12b.
 \citet{maciejewski_departure_2016} showed that the transit times of
-WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence, and
-measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
-\citet{patra_2017} presented new transit and occultation times for the
-system, and found $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
-Assuming that the period change was caused entirely by tidal decay,
-\citet{patra_2017} found values for the stellar quality factor of
-$Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
-explained by standard equilibrium tide
-models~\citep{penev_tidal_2011,ogilvie_tidal_2014}.
-\citet{weinberg_tidal_2017} found though that if WASP-12 were a
-subgiant, tidal dissipation due to a dynamical tide might be sufficient
-to produce the required level of friction.
-\cite{bailey_understanding_2019} concurred, but noted that the
-observational constraints on WASP-12 itself favor a main-sequence star,
-not a subgiant.  More data are needed to determine the cause of the
-timing variations in WASP-12b.
+WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
+and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
+Using new timing data, \citet{patra_2017} found
+$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If
+the period change were caused entirely by tidal decay, the implied
+modified stellar quality factor would be $Q_\star' \approx 2\times10^5$.
+Standard equilibrium tides do not predict this
+level of dissipation \citep{penev_tidal_2011,ogilvie_tidal_2014}.  The
+needed level of friction might come from dynamical tides, or
+obliquity tides
+\citep{weinberg_tidal_2017,millholland_obliquity_2018}.  The former is
+more likely if WASP-12 is a subgiant, but the observational
+constraints favor a main-sequence model
+\citep{bailey_understanding_2019}.  More data are needed to understand
+the WASP-12b timing variations.
 
 This work focuses on WASP-4b, which is also a plausible candidate for
-orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$, and
-mass $1.29 \pm 0.10\ M_{\rm Jup}$ \citep{southworth_homogeneous_2011}.
-It has an orbital period of 1.34 days, orbits its host star at a
-distance of $a/R_\star = 5.41$, and has been frequently observed over
-the past decade \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The
-orbit is consistent with circular \citep[][but see
-\S~\ref{sec:apsidal_precession}]{beerer_secondary_2011,husnoo_observational_2012,bonomo_gaps_2017}.
-The host star is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
+orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
+and mass $1.29 \pm 0.10\ M_{\rm Jup}$
+\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
+days, orbits its host star at a distance of $a/R_\star = 5.41$, and
+has been frequently observed over the past decade
+\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
+consistent with circular (but see \S~\ref{sec:apsidal_precession}),
+and has a sky-projected obliquity consistent with zero
+\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
+The host is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
 $M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
@@
 (\S~\ref{sec:timing}): a constant period, a decaying period, and a
-slightly eccentric orbit undergoing precession.  We rule out a constant
-period, but cannot firmly distinguish between the latter two
+slightly eccentric orbit undergoing precession.  We rule out a
+constant period, but cannot firmly distinguish between the latter two
 possibilities.  Each would have interesting implications
@@
 systematic time system offset causes the observed variation
-(Appendix~\ref{sec:verify_tess}), and conclude by advocating for further
-monitoring of the system (\S~\ref{sec:future}).
+(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
+further monitoring of the system (\S~\ref{sec:future}).
 
@@
 
-\subsection{Observations and cleaning}
+\subsection{Observations}
 WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
@@
 The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
-array every 2 minutes.  It was included on the ``short-cadence'' target
-list thanks to the Guest Investigator programs of J.\ Southworth and S.\
-Kane (G011112 and G011183 respectively).  After being downlinked through
+array every 2 minutes.   After being downlinked through
 the Deep Space Network, the data were processed through the Science
@@
 
-Using the times, flux measurements, and errors determined above, we ran
+Using the times, relative fluxes, and errors determined above, we ran
 Box Least Squares (BLS) to estimate the transit epoch, period, and
-duration \citep{kovacs_box-fitting_2002}.  From the BLS parameters, we
+duration \citep{kovacs_box-fitting_2002}.  Using these parameters, we
 isolated each transit to within $\pm$10 transit durations of its
@@
 % TODO: final numbers
-To check that the measured uncertainties are plausible, we computed the
+To verify that the measured uncertainties are plausible, we computed the
 reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
 times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
-Examining the residuals by eye showed that the error variance was being
+Visually inspecting residuals showed that the error variance was being
 overestimated, so we multiplied the measured TESS errors by a factor
-$f=0.73$, forcing a reduced $\chi^2$ of unity.  This reduced the mean
-transit time uncertainty from $\sim$30 seconds per transit to $\sim$22
-seconds per transit.
+$f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
+midtime uncertainty for each transit from 30 to 22
+seconds.
 
@@
 models, and residuals.  Table~1 reports the mid-transit times and
-their uncertainties, and also includes the previously reported times
-analyzed in \S~\ref{sec:timing}.
+their uncertainties.
 
@@
 Binning the residuals to 1-hour windows and taking the median absolute
-deviation, we measure a MAD of the residual lightcurves of $594\,{\rm
+deviation, the residual lightcurves have a MAD of $594\,{\rm
 ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
@@
 $673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
-\citet{winn_photonflux_2013} model, which was used in the
-\citet{Sullivan_2015} planet detection simulations.  The predicted level
+\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015}
+used in planet detection simulations.  The predicted level
 of zodiacal noise seems to have been an overestimate, given the observed
@@
 
-\subsection{Times}
+\subsection{Previous timing measurements}
 \label{subsec:times}
@@
 
-Table~1 shows all the transit times used in our analysis.  We included
+Table~1 shows the transit times used in our analysis.  We included
 data from peer-reviewed literature for which the analysis was based on
@@
 
-We use the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point.  We took their BJD$_{\rm TT}$ times to be equivalent to
-times in BJD$_{\rm TDB}$, since the two time systems differ periodically
-by only milliseconds \citep{urban_explanatory_2012}.  We verified that
-their times, as well as all others, were on the BJD$_{\rm TDB}$ time
-standard using the calculator provided
-by~\citet{eastman_achieving_2010}.  We set the zero-point epoch to be as
+We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
+starting point.  
+Though their latest two epochs fell slightly below the prediction of a
+linear ephemeris (Panel C of their Figure~6), they did not find a need
+to fit a quadratic function to the O-C values.
+We verified that the \citet{hoyer_tramos_2013} times, as well as all
+others, were on the BJD$_{\rm TDB}$ time
+standard using the \citet{eastman_achieving_2010} calculator.
+We set the zero-point epoch to be as
 close as possible to the average of the transit times, weighted as the
@@
 Important contributions to the timing dataset are as follows.  The
-earliest epoch in our fit is from the EulerCam lightcurve observed by
-\citet{wilson_wasp-4b_2008}.  The second, more precise, epoch comes from
+earliest epoch in our fit is from EulerCam on the 1.2-m Euler telescope
+\citep{wilson_wasp-4b_2008}.  The second epoch comes from
 $z$-band photometry acquired by \citet{gillon_improved_2009} at the VLT
-with FORS2.  Subsequent epochs came from
+8.2-m with FORS2.  Subsequent transits came from
 \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
@@
 \citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
-\citet{nikolov_wasp-4b_2012}.
+\citet{nikolov_wasp-4b_2012}, \citet{hoyer_tramos_2013}, and
+\citet{ranjan_atmospheric_2014}.
 %\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
@@
 %the MPG/EOS-2.2~m at La Silla.
-\citet{hoyer_tramos_2013} reported nine transit observations taken with
-the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
-SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013} also
-homogeneously redetermined the transit times from all previous studies.
-Though their latest two epochs fell slightly below the prediction of a
-linear ephemeris (Panel C of their Figure~6), they did not find a need
-to fit a quadratic function to the O-C values.
-
-After the study by \citet{hoyer_tramos_2013},
-\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
-using HST's WFC3 in both transit and secondary eclipse.  We adopt only
-their transit time, since the epoch of occultation was not fit as a free
-parameter.  \citet{huitson_gemini_2017} acquired optical transmission
+%\citet{hoyer_tramos_2013} reported nine transit observations taken with
+%the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
+%SOI on the SOAR 4.2-m telescope.
+%\citet{hoyer_tramos_2013} also
+%homogeneously redetermined the transit times from all previous studies.
+%After the study by \citet{hoyer_tramos_2013},
+%\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
+%using HST's WFC3 in both transit and secondary eclipse.  We adopt only
+%their transit time, since the epoch of occultation was not fit as a free
+%parameter.'
+Finally, \citet{huitson_gemini_2017} acquired optical transmission
 spectra with the 8.1-m Gemini South telescope over 2011 to 2014, one
 transit per season.  The per-point standard deviation of their
-lightcurves is a few hundred parts per million, less than a factor of 3
-away from the photon noise limit.  The uncertainties on their reported
+lightcurves is a few hundred parts per million, within a factor of 3
+from the photon noise limit.  The uncertainties on their reported
 transit times~--~on average 5.6~seconds~--~ are small, but possible
@@
 
-There have been other studies of WASP-4b for which the transit timing
-data are either hetereogenous, not available, or not useable.  We
-omitted the timing measurements from
-\citet{southworth_high-precision_2009}, since there were technical
-problems with the computer clock at the time of
-observation~\citep{nikolov_wasp-4b_2012}.  The timing study
-by~\citet{petrucci_no_2013} ruled out transit timing variations larger
-than 54~seconds over the previous 5 years, but did not report transit
-times.  \citet{baluev_benchmarking_2015} collected transit times for
-many hot Jupiters, including WASP-4b, and included times from the
-Exoplanet Transit Database
-(ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
-\citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
-sources, their timestamps are less clearly documented, and the times
-are thus more prone to systematic errors, we omit them from
-consideration.
-%We consider the effects of including ETD data in
-%Appendix~\ref{sec:verify_archival_times}.
-\citet{may_mopss_2018} obtained optical spectra using IMCAS on
-Magellan, and found a flat transmission spectrum, in agreement with
-the results reported by \citet{huitson_gemini_2017}.  They did not
-report transit times.
+% There have been other studies of WASP-4b for which the transit timing
+% data are either hetereogenous, not available, or not useable.  We
+% omitted the timing measurements from
+% \citet{southworth_high-precision_2009}, since there were technical
+% problems with the computer clock at the time of
+% observation~\citep{nikolov_wasp-4b_2012}.  The timing study
+% by~\citet{petrucci_no_2013} ruled out transit timing variations larger
+% than 54~seconds over the previous 5 years, but did not report transit
+% times.  \citet{baluev_benchmarking_2015} collected transit times for
+% many hot Jupiters, including WASP-4b, and included times from the
+% Exoplanet Transit Database
+% (ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
+% \citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
+% sources, their timestamps are less clearly documented, and the times
+% are thus more prone to systematic errors, we omit them from
+% consideration.
+% %We consider the effects of including ETD data in
+% %Appendix~\ref{sec:verify_archival_times}.
+% \citet{may_mopss_2018} obtained optical spectra using IMCAS on
+% Magellan, and found a flat transmission spectrum, in agreement with
+% the results reported by \citet{huitson_gemini_2017}.  They did not
+% report transit times.
 
@@
 specifying the time standard.  We correspondingly added $69.184$ seconds
-of uncertainty to their reported errors, in quadrature.  Finally, as
-previously mentioned, \citet{ranjan_atmospheric_2014} acquired
-occultation data with HST but kept the epoch as a fixed parameter.
-\citet{zhou_secondary_2015} performed a similar analysis with
+of uncertainty to their reported errors, in quadrature.
+%Finally, as
+%previously mentioned, \citet{ranjan_atmospheric_2014} acquired
+%occultation data with HST but kept the epoch as a fixed parameter.
+\citet{zhou_secondary_2015} acquired
 occultation data from the Anglo-Australian Telescope, focusing on the
 eclipse depth, but also fitting for $e\cos\omega$.  We converted their
-$e\cos\omega$ results into a midtime, using their ephemeris and the
+$e\cos\omega$ result into a midtime, using their ephemeris and the
 standard formula \citep[{\it
@@
 \end{equation}
-This gives us a total of 4 occultations.  Although their timing error
-bars are large compared to the transits, we include them for
-completeness, and because they can help in our modeling of the timing
+This gives us a total of 4 occultations.  Although the occultation
+times are uncertain compared to the transits, we include them for
+completeness, and because they help in modeling the timing
 variations.
@@
 literature times. The result is shown in
-Figure~\ref{fig:arrived_early}. The TESS observations appear to have
+Figure~\ref{fig:arrived_early}. The TESS observations seem to have
 arrived early.  One immediate concern was systematic errors in either
@@
 
-Assuming the error bars on the transit midtimes are gaussian and
-independent, we then proceed by fitting each model to the timing data.
-For the linear model, we took uniform priors over small windows around
-the expected parameters.  For the quadratic model we did the same, and
-took the quadratic term to have a uniform prior corresponding to
-$
-  Q_\star' \sim \mathcal{U}[10^3, 10^9],
-$
-for $Q_\star'$ the modified quality factor defined in
-\S~\ref{sec:implications}.  For the precession model, we took the same
-priors in epoch and sidereal period, and drew the remaining parameters
-from wide uniform distributions.
+% Assuming the error bars on the transit midtimes are gaussian and
+% independent, we then proceed by fitting each model to the timing data.
+% For the linear model, we took uniform priors over small windows around
+% the expected parameters.  For the quadratic model we did the same, and
+% took the quadratic term to have a uniform prior corresponding to
+% $
+%   Q_\star' \sim \mathcal{U}[10^3, 10^9],
+% $
+% for $Q_\star'$ the modified quality factor defined in
+% \S~\ref{sec:implications}.  For the precession model, we took the same
+% priors in epoch and sidereal period, and drew the remaining parameters
+% from wide uniform distributions.
 
@@
 Figure~\ref{fig:times} shows the residuals with respect to the linear
-model.  The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of
-freedom.  The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57
-degrees of freedom.  The median MCMC precession fit has $\chi^2 = 51.0$
-and 55 degrees of freedom.
+model.  The best linear fit has $\chi^2 = 149$ and 58 degrees of
+freedom.  The best quadratic fit has $\chi^2 = 48.0$ and 57
+degrees of freedom, compared to $\chi^2 = 51.0$
+and 55 degrees of freedom for the best-fit precession model.
 
@@
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$. In other words, assuming the linear
-model is true, there is a $10^{-23}$ probability that simply by chance
-we would see data that favors the quadratic model more strongly than the
-observations.  A more likely possibility is some systematic error being
-present in one of the most important sets of times: either the earliest
-times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591, and
-851), or the TESS times.  We have taken every precaution against
-systematic offsets in the TESS dataset (Appendix~\ref{sec:verify_tess}),
-and Huitson et al.\ stand by their reported times (Huitson, priv.
-comm.).  Since the linear model provides a poor fit to the data, we
-discard it from further consideration.
+corresponds to $p \approx 10^{-23}$. %In other words, assuming the linear
+%model is true, there is a $10^{-23}$ probability that simply by chance
+%we would see data that favors the quadratic model more strongly than the
+%observations.
+Clearly, it is more likely that some systematic error could be
+affecting one of the most important sets of times: either the earliest
+times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591,
+and 851), or the TESS times.  We have taken every precaution against
+systematic offsets in the TESS dataset
+(Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by their
+reported times (Huitson, priv.  comm.).  Since the linear model
+provides a poor fit to the data, we discard it from further
+consideration.
 
@@
 The quadratic model fits better than the precession model.  It is
-favored over the precession fit by $\Delta \chi^2 = 3.0$, and has two
+favored by $\Delta \chi^2 = 3.0$, and has two
 fewer free parameters.  Another useful heuristic for model comparison is
@@
 ``decisive evidence'' for the quadratic model describing the data better
-than the precession model.  Using the Akaike Information Criterion
+than the precession model.  The Akaike Information Criterion
 similarly favors the orbital decay model, by $\Delta {\rm AIC} = 7$.
@@
 %TODO: final numbers
-Assuming the orbital decay model, we find a period derivative 
+If we assume the orbital decay model, we find a period derivative 
 \begin{equation}
@@
     = - (12.1^{+1.2}_{-1.2})\,{\rm ms}\,{\rm yr}^{-1}.
+    \label{eq:dP_dt_obs}
 \end{equation}
@@
 \citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b is
-$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  Assuming the quadratic
-model, WASP-4b is thus decaying at about half the rate as WASP-12b.
+$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
+If both planets are decaying, WASP-4b would be inspiraling at about half
+the rate as WASP-12b.
 
@@
 If we instead assume the precession model, the best-fit
-eccentricity is $e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}$. The
-longitude of periapse correspondingly advances by $\dot{\omega}
-= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$.   The
-corresponding precession period would be $24^{+8}_{-6}$ years.
+eccentricity is
+\begin{equation}
+e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}
+\end{equation}
+The longitude of periapse correspondingly advances by $\dot{\omega}
+= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$, and the
+precession period would be $24^{+8}_{-6}$ years.
 
 The median parameters and their standard deviations are reported for all
-the models in Table~3.  To summarize, a linear ephemeris does not
+the models in Table~3.  In short, a linear ephemeris does not
 describe the observed times.  Both orbital decay and precession can fit
-the observed data.  However, the precession model has two extra free
-parameters, and is statistically disfavored.  Further observations will
-help discriminate between orbital decay and precession.
+the data, but the precession model is statistically disfavored.
 
@@
   \begin{center}
-    \includegraphics[width=0.415\textwidth]{f4.pdf}
+    \includegraphics[width=0.425\textwidth]{f4.pdf}
   \end{center}
@@
 
-Let us assume for the sake of argument that the observed timing
+For the sake of argument,
+let us assume that the observed timing
 variation is caused entirely by orbital decay.  If the period
-decreases at a fixed rate, it will reach zero after
+continues to decrease at a fixed rate, it will reach zero after
 \begin{equation}
@@
 the equilibrium deformation of the star, divided by the energy
-dissipated per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.
+lost to heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.
 A larger $Q_\star$ implies less efficient dissipiation.  Though the
@@
 eccentricity evolve as
-\citep[][Appendix B]{metzger_optical_2012}
+\citep[Appendix B of][]{metzger_optical_2012}
 \begin{align}
@@
 the prediction for WASP-4 is that it has $Q_\star' = 7\times10^5$.
-This is one order of magnitude larger than would be implied by the
+This is one order of magnitude larger than implied by the
 observed period change.
@@
 as being a portion of an apsidal precession cycle.  The required orbital
-eccentricity is then
-\begin{equation}
-  e = (1.6^{+1.5}_{-0.7})\times10^{-3},
-\end{equation}
-and the line of apsides is advancing by
-\begin{equation}
-  \dot{\omega} = 14.8^{+5.4}_{-3.8}\ {\rm degrees}\,{\rm yr}^{-1}.
-\end{equation}
-One full precession period would take about 24 years, or twice the
-current observing baseline.
+eccentricity is then $e\approx 1.6\times10^{-3}$, and one full precession
+period would take about 24 years (twice the
+current observing baseline).
 
@@
 the star's rotational and tidal bulges are minor contributors.
-\citet{ragozzine_probing_2009} also calcuated that WASP-4b was expected
-to have a relatively short precession period, of 120 years.
+%\citet{ragozzine_probing_2009} also calcuated that WASP-4b was expected
+%to have a relatively short precession period, of 120 years.
 
@@
 into a measurement of the planetary Love number, $k_{2,\rm p}$.
-From \citet{ragozzine_probing_2009} Equation 14, the implied
-planetary Love number for WASP-4b is
+From their Equation 14, the implied Love number for WASP-4b is
 \begin{equation}
@@
 For comparison, $k_{2,{\rm Jup}} \approx 0.59$
-(\citealt{wahl_tidal_2016}, bug see \citealt{ni_empirical_2018}).  A
-uniform density sphere has $k_2 = 1.5$.  Our reported uncertainties on
+(\citealt{wahl_tidal_2016}, but see \citealt{ni_empirical_2018}).  A
+uniform density sphere has $k_2 = 1.5$.  Our uncertainties on
 $k_{2,{\rm p}}$ are large because the eccentricity, reference time, and
-${\rm d}\omega/{\rm d}E$ are quite correlated, and the measured
-occultation times only barely constrain them (Figure~\ref{fig:future}).
+${\rm d}\omega/{\rm d}E$ are correlated, and the measured occultation
+times only barely constrain them (Figure~\ref{fig:future}).
 
@@
 damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e = 0.29 (Q_{\rm
-p}'/10^5) \, {\rm Myr}$.  The system appears to be well over $1\,{\rm
+p}'/10^5) \, {\rm Myr}$.  The system is well over $1\,{\rm
 Gyr}$ \citep{winn_transit_2009} old, so any initial eccentricity should
@@
 HD 189733b, for example, was initially reported to have a non-zero
-eccentricty, $e\cos\omega = 0.0010 \pm 0.0002$
+eccentricty
 \citep{knutson_map_2007}.  This result was retracted by
 \citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
-eclipses and found a smaller average offset in the secondary eclipse
+occultations and found a smaller average offset in the occultation
 time than \citet{knutson_map_2007}.  \citet{agol_climate_2010} also
@@
 minutes \citep{williams_resolving_2006}.  Including an updated
-hot-spot model, \citet{agol_climate_2010} reported a 2$\sigma$ limit
-of $e\cos\omega < 0.00023$, perhaps the strongest eccentricity limit yet
+hot-spot model, \citet{agol_climate_2010} reported
+$e\cos\omega < 0.00023$ at 2$\sigma$, perhaps the strongest
+eccentricity limit in place
 for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
 \citet{winn_measurement_2005} combined Spitzer secondary eclipses with
-RV data and reported a small finite eccentricity.  They suspected that
+RV data and reported a small non-zero eccentricity.  They suspected that
 the data were consistent with zero eccentricity, and that the lower
 bound on $e$ was an artifact of an imperfect limb-darkening model.
-\citet{crossfield_spitzer_mips_2012} eventually found stronger limited:
+\citet{crossfield_spitzer_mips_2012} eventually ruled out the \citet{winn_measurement_2005} values, by determining
 $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b, for
@@
 systematic uncertainties because of the hot-spot issue
-\citep{williams_resolving_2006,agol_climate_2010}.  With the exception
-of WASP-18b, radial velocity measurements have provided eccentricity
-constraints only at the $e > 10^{-2}$ level.  Empirical cannot rule out
-hot Jupiter eccentricities at the level needed to contextualize WASP-4.
+\citep{williams_resolving_2006,agol_climate_2010}.  Except for
+WASP-18b, radial velocity measurements have provided eccentricity
+constraints only at the $e > 10^{-2}$ level.  We cannot empirically rule out
+hot Jupiter eccentricities at the level needed to assess WASP-4's
+putative eccentricity.
 
@@
 of hot Jupiters with companions.  They pointed out that although the
-early phases of the two-planet eccentricity evolution would occur
+early phases of the two-planet eccentricity evolution occur
 quickly, the final phase of the joint eccentricity evolution towards
-circularity would often occur on timescales several orders of magnitude
+circularity would occur on timescales several orders of magnitude
 longer than the circularization time (see their Figures~4 and 5).  The
@@
 
-Another mechanism to excite the hot Jupiter's eccentricity is the
+Another way to excite the hot Jupiter's eccentricity is the
 Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
@@
 least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
-\citet{bailey_understanding_2019} point out how the joint requirements
-of Kozai-Lidov oscillations with a non-detection of an additional RV
-signal can be combined to produce lower and upper limits on the
-companion's mass.  For the Kozai-Lidov mechanism to operate at maximum
+%\citet{bailey_understanding_2019} point out how the joint requirements
+%of Kozai-Lidov oscillations with a non-detection of an additional RV
+%signal can be combined to produce lower and upper limits on the
+%companion's mass.
+For the Kozai-Lidov mechanism to operate at maximum
 efficiency, from
@@
 transits along the line of sight, $f$ can be arbitrarily small, and
-both of the preceding limits can be easily satisfied.  This
-mechanism could produce the needed eccentricity. However it would be more
+both of the preceding limits can be satisfied.
+
+While either of the two preceding mechanisms could produce
+the needed eccentricity,
+it would be more
 satifying if long-term radial velocity monitoring showed an additional
@@
 velocity measurements for WASP-4, and merged them with measurements
-from CORALIE and HARPS reported by \citet{wilson_wasp-4b_2008},
-\citet{pont_determining_2011}, and \citet{husnoo_observational_2012}.
+from CORALIE and HARPS \citep{wilson_wasp-4b_2008,pont_determining_2011,husnoo_observational_2012}.
 Using these four datasets, \citet{knutson_friends_2014} found that
-$\dot{\gamma} = -0.0099^{+0.0052}_{-0.0054}$, corresponding to a
-$2\sigma$ limit on the acceleration towards us of $\dot{\gamma} > -
-0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.  This limits $\dot{P}$
-to be less than $9\times 10^{-11}$ at $2\sigma$.  
+$\dot{\gamma} = -0.0099^{+0.0052}_{-0.0054}$, or that the
+star can be accelerating towards us at most by $\dot{\gamma} > -
+0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ (2$\sigma$).
+The associated $\dot{P}$
+must be less than $9\times 10^{-11}$.  
 %FIXME TODO final value
-Our measured value is $\dot{P} = -(3.9^{+0.7}_{-0.5})\times 10^{-10}$.
-We can thus rule out radial acceleration as the sole cause of the
-observed period change.  However, it could cause up to one-quarter of
-the observed period decrease.  Improved long-term radial velocity
-monitoring with the same instruments will help clarify the situation,
-particularly given the fact that short-term trends of roughly
-$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ are
-seen over timescales of order 10 days in
-WASP-4~\citep{husnoo_observational_2012}.  Given the known spottedness
+%Our measured value is $\dot{P} = -(3.9^{+0.7}_{-0.5})\times 10^{-10}$.
+Radial acceleration could thus cause at most one-quarter of
+the observed period decrease (Equation~\ref{eq:dP_dt_obs}).
+Improved radial velocity
+monitoring will help clarify the situation,
+particularly given the fact that WASP-4 shows short-term trends of
+$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
+tens of days \citep{husnoo_observational_2012}.
+Given the known spottedness
 of the system~\citep{sanchis-ojeda_starspots_2011}, these
-short-timescale variations may also have a contribution from stellar
+short-timescale variations likely have a contribution from stellar
 activity.
@@
 measurements could also be useful.  In order to rule between orbital
-decay and apsidal precession, occultation measurements in the near term,
-and also in the mid-2020s, will be helpful.  Additionally, confirmation
+decay and apsidal precession, occultation measurements in both
+the near term and also in the mid-2020s will be needed.  Additionally, confirmation
 and extension of the linear trend suggested by
@@
 exterior companion in the system might explain any residual
-eccentricity.  These observations should be carried out, because they
+eccentricity.  These observations should be carried out because they
 will lead to new knowledge; either a measure of the efficiency of tidal
 dissipation within the star, or the eccentricity and Love number of the
-planet.  Measuring either for a single system would be rather
+planet.  Either outcome for a single system would be rather
 interesting.  In the coming decade, wide-field photometric surveys
@@
 \acknowledgements{
-L.G.B.\ is happy to acknowledge help with dynamics from A.  Bailey,
-and useful brainstorming with F. Dai and V. Van Eylen. 
+L.G.B.\ happily acknowledges helpful discussions with A. Bailey, F.
+Dai and V. Van Eylen, and is grateful to the people who have
+contributed to TESS.
 %
-We acknowledge the use of TESS Alert data, which is currently in a beta test
-phase, from the TESS Science Office. Funding for the TESS mission is provided
-by NASA's Science Mission directorate. This research has made use of the NASA
-Exoplanet Archive, which is operated by the California Institute of Technology,
-under contract with the National Aeronautics and Space Administration under the
-Exoplanet Exploration Program.  This work made use of NASA's Astrophysics Data
-System Bibliographic Services.
+WASP-4 was included on the ``short-cadence'' target list thanks to the
+Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
+and G011183 respectively). 
+%
+This paper includes data collected by the TESS mission, which are
+publicly available from the Mikulski Archive for Space Telescopes
+(MAST).
+%
+Funding for the TESS mission is provided by NASA’s Science Mission
+directorate.
+%
+This research has made use of the NASA Exoplanet Archive, which is
+operated by the California Institute of Technology, under contract
+with the National Aeronautics and Space Administration under the
+Exoplanet Exploration Program.
+%
+This work made use of NASA's Astrophysics Data System Bibliographic
+Services.
+%
 This research has made use of the VizieR catalogue access tool, CDS,
@@
 published in A\&AS 143, 23.
-This work has made use of data from the European Space Agency (ESA) mission
-{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
-Data Processing and Analysis Consortium (DPAC,
-\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
-has been provided by national institutions, in particular the institutions
-participating in the {\it Gaia} Multilateral Agreement.
+%
+This work has made use of data from the European Space Agency (ESA)
+mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
+by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
+\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
+for the DPAC has been provided by national institutions, in particular
+the institutions participating in the {\it Gaia} Multilateral
+Agreement.
 %
@@
 
-\bibliographystyle{yahapj}                            
-\bibliography{bibliography}
-
 %\clearpage
@@
 \input{model_fit_table.tex}
-\clearpage
-\newpage
+
+\bibliographystyle{yahapj}                            
+\bibliography{bibliography}
+
+%\clearpage
+%\newpage
 
@@
 the barycentric correction reported by SPOC in the TESS lightcurve
-file headers.  In \S~\ref{sec:hj_verification}, we repeat the main
-analysis of the paper for a collection of other hot Jupiters, and use
-their observed transit times to rule out a global TESS time offset.
+file headers.  In \S~\ref{sec:hj_verification}, we measure the transit times
+of other hot Jupiters, and use them
+to rule out a global TESS time offset larger than about a minute.
 The latter test confirms that WASP-4b is an outlier.

### 2018-12-06 paper/ms.tex
minor updates

@@
 %-------------------------------------
-%\\ \textcolor{red}{3 TSO coauthors:}
-\author{TSO1}
-\author{TSO2}
-\author{TSO3}
-%\\ \textcolor{red}{3 SPOC coauthors:}
-\author{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
+% 3 TSO COAUTHORS
+%
+\author[0000-0002-3321-4924]{Z. Berta-Thompson} % Zachory.BertaThompson@Colorado.EDU WAITING
+\affiliation{Department of Astrophysical and Planetary Sciences,
+University of Colorado, Boulder, CO 80309, USA}
+%
+\author[0000-0001-8812-0565]{J. Rodriguez} % rodriguez.jr.joey@gmail.com WAITING
+\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
+Street, Cambridge, MA 02138, USA }
+%
+\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
+\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
+Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
+% 3 SPOC COAUTHORS
+%
+\author[0000-0002-6778-7552]{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
 \affiliation{ SETI Institute, Mountain View, CA, 94043}
+%
 \author{B. Wohler} % bill.wohler@nasa.gov WAITING
@@
 \affiliation{ SETI Institute, Mountain View, CA, 94043}
-\author{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
+%
+\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
 \affiliation{ SETI Institute, Mountain View, CA, 94043}
+%
 %\\ \textcolor{red}{3 POC coauthors:}
@@
 their uncertainties.
-
 Taking the mean and standard deviation of the measured planet to star
@@
 \citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
-
 Binning the residuals to 1-hour windows and taking the median absolute
@@
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$. %In other words, assuming the linear
-%model is true, there is a $10^{-23}$ probability that simply by chance
-%we would see data that favors the quadratic model more strongly than the
-%observations.
+corresponds to $p \approx 10^{-23}$.
+%In other words, assuming the linear model is true, there is a
+%$10^{-23}$ probability that simply by chance we would see data that
+%favors the quadratic model more strongly than the observations.
 Clearly, it is more likely that some systematic error could be
@@
 systematic offsets in the TESS dataset
-(Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by their
-reported times (Huitson, priv.  comm.).  Since the linear model
-provides a poor fit to the data, we discard it from further
-consideration.
+(Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by their
+reported times (J.-M. D\'esert, priv.\
+comm.)\footnote{\citealt{huitson_gemini_2017} corrected the times to
+be at mid-exposure. They also analyzed other hot Jupiters, which all
+showed mid-transit times as expected from a constant ephemeris.}.
+Since the linear model provides a poor fit to the data, we discard it
+from further consideration.
 
@@
 interaction in a binary applies \citep{zahn_tidal_1977}.  The fluid
-response is parametrized by a constant modified\footnote{For stars,
+response is parametrized by a modified\footnote{For stars,
 $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to explicitly
@@
 \end{equation}
-Inserting our observed value of $\dot{P}$, the current dissipation
-rate corresponds to a modified quality factor for WASP-4 of
+Using our observed value of $\dot{P}$, the current modified quality
+factor of WASP-4 would be
 \begin{equation}
@@
 \citep{lainey_strong_2009}.  For stars, different $Q_\star'$ values have
-been estimated in different contexts.  Studies of the observed
-eccentricity distribution in stellar binaries have yielded $10^5 \gtrsim
+been estimated in different contexts.  Studies of the 
+eccentricity distribution of stellar binaries have yielded $10^5 \gtrsim
 Q_\star' \gtrsim 10^{7}$ \citep[{\it

### 2018-12-06 src/plot_arrived_early.py

@@ -92,6 +92,13 @@ def plot_arrived_early(plname, xlim=None, ylim=None, savpath=None):
         weights=1/tess_err_tmid**2
     )
     bin_tess_err_tmid = np.mean(tess_err_tmid)/len(tess_tmid)**(1/2)
+
+    tess_yval = nparr(tess_tmid -
+                      linear_model(lsfit_t0, lsfit_period, tess_epoch))*24*60
+    print('std (min) {}'.format(np.std(tess_yval)))
+    print('std (sec): {}'.format(np.std(tess_yval)*60))
+    print('error (plotted, min): {}'.format(bin_tess_err_tmid*24*60))
+    print('error (plotted, sec): {}'.format(bin_tess_err_tmid*24*60*60))
     bin_tess_x = np.median(tess_epoch)
 
     ax.errorbar(bin_tess_x, bin_tess_y*24*60, bin_tess_err_tmid*24*60,

### 2018-12-07 paper/ms.tex
draft: v0.1. ready for JNW.

@@
-%\documentclass[preprint2,tighten]{aastex6}
 \documentclass[12pt,twocolumn,tighten]{aastex62}
@@
 \shortauthors{Bouma et al.}
-\shorttitle{Decreasing Period of WASP-4\lowercase{b}}
+\shorttitle{Early Arrival of WASP-4\lowercase{b}}
+
+\NewPageAfterKeywords
 
@@
 
-\title{The Decreasing Orbital Period of WASP-4\lowercase{b}}
+\title{The Early Arrival of WASP-4\lowercase{b}}
 
@@
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
-
+%
 \author[0000-0002-4265-047X]{J. N. Winn}
@@
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
-
-\author[0000-0002-6939-9211]{T. Daylan}
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-
+%
+\author[0000-0002-0875-8401]{J.-M. D\'esert}
+\affiliation{ API, University of Amsterdam, P.O. Box 94249, 1090 GE
+Amsterdam, The Netherlands}
+%
 \author[0000-0002-3481-9052]{K. G. Stassun}
@@
 N., Nashville, TN 37208, USA}
-
+%
+\author[0000-0002-6939-9211]{T. Daylan}
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+%
 \author[0000-0002-7084-0529]{S. Kane}
@@
 Riverside, CA 92521, USA}
-
+% 
 %-------------------------------------
@@
 %-------------------------------------
-%\\ \textcolor{red}{6 TESS Mission Architects:}
+%
 \author{G. R. Ricker} % grr@space.mit.edu WAITING
@@
 02139, USA }
+%
 \author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
@@
 02139, USA }
+%
 \author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
@@
 Street, Cambridge, MA 02138, USA }
+%
 \author{S. Seager} % seager@mit.edu WAITING
@@
 Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
+%
 \author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+%
 %-------------------------------------
@@
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA, 94043}
+\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
 %
@@
   with previous studies, we show that the transit times of WASP-4b are
-  incompatible with a constant orbital period. 
-  In particular, the transits seem to have arrived $77.8 \pm 10.7$
-  seconds early, and the period appears to be shrinking by
-  $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
-  From TESS observations of WASPs-6b, -18b, and -46b, we show
-  that a systematic offset between the TESS time stamps and the barycentric
-  reference sufficient to explain WASP-4b is ruled out at 6.3$\sigma$.
-  If the timing variations are astrophysical,
-  major contributors to the period change could include both
-  apsidal precession and tidal decay.
-  The latter seems implausible on theoretical grounds, as it would
-  require a stellar quality factor $Q_\star' \approx 3\times10^4$.
-  The Doppler shift from WASP-4's acceleration towards us could
-  account for at most one quarter of the observed period change
-  (at $2\sigma$).
-  Further transit and occultation studies will help confirm the
-  reality of the timing variation, and eventually determine its cause.
+  incompatible with a constant orbital period.  In particular, the
+  transits seem to have arrived $77.8 \pm 10.8$ seconds early, and the
+  period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
+  milliseconds per year.   From TESS observations of WASPs-6b, -18b,
+  and -46b, we show that a systematic offset between the TESS times
+  and the barycentric reference sufficient to explain WASP-4b is ruled
+  out at 6.3$\sigma$.  The timing variations seem to be astrophysical,
+  and could be caused by apsidal precession, or perhaps by tidal
+  decay.  The latter seems implausible on theoretical grounds, as it
+  would require a stellar quality factor $Q_\star' \approx
+  3\times10^4$.  Apsidal precession would imply an eccentricity
+  $e\approx1.6\times 10^{-3}$, and the possibility of measuring the
+  planet's Love number.  A Doppler shift from WASP-4's acceleration
+  towards us could account for at most one quarter of the observed
+  period change (at $2\sigma$).  Further observations will help
+  confirm the reality of the timing variation, and eventually
+  determine its cause.
 \end{abstract}
@@
 
-Long-term monitoring of hot Jupiter transit and occultation times should
-eventually reveal two distinct processes.  First, whenever the summed
-spin and orbital angular momenta in a hot Jupiter system are below a
-critical value, the system will be unstable to tidal decay
-\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
-satisfy this condition; their orbits should be shrinking
-\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay is
-uncertain, and depends on how friction dissipates energy carried by
+Long-term monitoring of hot Jupiter transit and occultation times
+should eventually reveal two distinct processes.  First, whenever the
+summed spin and orbital angular momenta in a hot Jupiter system fall
+below a critical value, the system becomes unstable to tidal decay
+\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
+Jupiters satisfy this condition; their orbits should be shrinking
+\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay
+is uncertain, and depends on how friction dissipates energy carried by
 gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
@@
 dissipation rates determine the fate of the system.  During its
-inspiral, if the planet fills its Roche lobe below the stellar surface,
-the drag force from the stellar envelope accelerates the inspiral,
-leading to a `direct-impact' merger
+inspiral, if the planet fills its Roche lobe below the stellar
+surface, the drag force from the stellar envelope accelerates the
+inspiral, leading to a `direct-impact' merger
 \citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
-the planet fills its Roche lobe above above the stellar surface, then it
-could either be disrupted into an accretion disk or it could
-slowly transfer mass to the star
+the planet fills its Roche lobe above above the stellar surface, then
+it could either be disrupted into an accretion disk or it could slowly
+transfer mass to the star
 \citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
@@
 binaries \citep[{\it e.g.},][]{
-schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}, and has been
-explored theoretically for transiting planets
+  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015},
+and has been explored theoretically for transiting planets
 \citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
@@
 WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
-and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
-Using new timing data, \citet{patra_2017} found
-$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If
-the period change were caused entirely by tidal decay, the implied
-modified stellar quality factor would be $Q_\star' \approx 2\times10^5$.
-Standard equilibrium tides do not predict this
-level of dissipation \citep{penev_tidal_2011,ogilvie_tidal_2014}.  The
-needed level of friction might come from dynamical tides, or
-obliquity tides
+and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.  Using
+new timing data, \citet{patra_2017} found $\dot{P} = -29 \pm 3\,{\rm
+ms}\,{\rm yr}^{-1}$.  If the period change were caused entirely by
+tidal decay, the implied modified stellar quality factor would be
+$Q_\star' \approx 2\times10^5$.  Standard equilibrium tides do not
+predict this level of dissipation
+\citep{penev_tidal_2011,ogilvie_tidal_2014}.  The needed level of
+friction might come from dynamical tides, or obliquity tides
 \citep{weinberg_tidal_2017,millholland_obliquity_2018}.  The former is
@@
 
-This work focuses on WASP-4b, which is also a plausible candidate for
-orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
-and mass $1.29 \pm 0.10\ M_{\rm Jup}$
-\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
-days, orbits its host star at a distance of $a/R_\star = 5.41$, and
-has been frequently observed over the past decade
-\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
-consistent with circular (but see \S~\ref{sec:apsidal_precession}),
-and has a sky-projected obliquity consistent with zero
+This work focuses on the hot Jupiter WASP-4b, which orbits its host
+star every 1.34 days, at a distance of $a/R_\star = 5.46$
+\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  WASP-4b is a plausible
+candidate for orbital decay, in part because it has been frequently
+observed over the past decade.  The host is a G7V dwarf, and is on the
+main sequence (\S~\ref{sec:system_parameters}).  The orbit is
+consistent with being circular (but see
+\S~\ref{sec:apsidal_precession}), and having sky-projected obliquity
+of zero
 \citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
-The host is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
-$M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
-K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
-\citep{southworth_homogeneous_2011,petrucci_no_2013,doyle_accurate_2013}.
-
-We measure new transit times for WASP-4b using data from TESS
-(\S~\ref{sec:transits}).  We then merge these times with previous
-observations and compare three models for the times
-(\S~\ref{sec:timing}): a constant period, a decaying period, and a
-slightly eccentric orbit undergoing precession.  We rule out a
+
+We measure new transit times for WASP-4b using TESS lightcurves
+(\S~\ref{sec:observations},~\ref{sec:measurement}), and calculate the
+system's parameters empirically (\S~\ref{sec:system_parameters}).  We
+merge the new times with previous observations and compare three
+models (\S~\ref{sec:timing}): a constant period, a decaying period,
+and a slightly eccentric orbit undergoing precession.  We rule out a
 constant period, but cannot firmly distinguish between the latter two
@@
 
-\section{New transit times}
-\label{sec:transits}
-
+\section{New transits and system parameters}
 \begin{figure}[t]
@@
 \subsection{Observations}
+\label{sec:observations}
+
 WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
@@
 The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
-array every 2 minutes.   After being downlinked through
-the Deep Space Network, the data were processed through the Science
-Processing Operations Center (SPOC) pipeline at NASA
+array every 2 minutes.   After being downlinked through the Deep Space
+Network, the data were processed through the Science Processing
+Operations Center (SPOC) pipeline at NASA
 Ames~\citep{jenkins_tess_2016}.  One important step in this processing
@@
 2457000$, BJD is the barycentric julian date, and all times are in the
-TDB reference (see \citealt{urban_explanatory_2012} for explanations of
-time systems).  The lightcurves were then vetted and released by the MIT
-TESS Science Office to the Mikulski Archive for Space
+TDB reference (see \citealt{urban_explanatory_2012} for explanations
+of time systems).  The lightcurves were then vetted and released by
+the MIT TESS Science Office to the Mikulski Archive for Space
 Telescopes~\citep{ricker_tess_alerts_2018}.
@@
 
-We then processed the lightcurve as follows.  First, we removed all points
-with non-zero quality flags.  This removes data contaminated by
+We then processed the lightcurve as follows.  First, we removed all
+points with non-zero quality flags.  This removes data contaminated by
 coarse spacecraft pointing, cosmic rays, and other annoyances
-\citep{tess_data_product_description_2018}.  We then filtered out known
-bad observing windows.  The data closest to spacecraft perigee typically
-show ramp-like systematics, so we clipped out the first and last hour of
-both orbits.  Next, we focused on the momentum dumps. As described in
-the data release
+\citep{tess_data_product_description_2018}.  We then filtered out
+known bad observing windows.  The data closest to spacecraft perigee
+typically show ramp-like systematics, so we clipped out the first and
+last hour of both orbits.  Next, we focused on the momentum dumps. As
+described in the data release
 notes\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}},
 during the first two TESS sectors, every 2.5~days the spacecraft's
-reaction wheels were reset to maintain pointing stability.  These events
-were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude''.  For WASP-4,
-these flags were simultaneously set for 54 distinct cadences, and there
-were 10 momentum dumps, averaging about 10 minutes of flagged data per
-dump.  Simply out of caution, we clipped out an additional 10 minutes
-before and after every momentum dump.  Before applying any filters, we
-had 19737 data points. After applying all the filters, we were left with
+reaction wheels were reset to maintain pointing stability.  These
+events were assigned quality flags corresponding to ``Reaction Wheel
+Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
+were simultaneously set for 54 distinct cadences, and there were 10
+momentum dumps, averaging about 10 minutes of flagged data per dump.
+Simply out of caution, we clipped out an additional 10 minutes before
+and after every momentum dump.  Before applying any filters, we had
+19737 data points. After applying all the filters, we were left with
 18165 measurements, or 92\% of the original data.
@@
 \subsection{Measuring the transit times}
+\label{sec:measurement}
 
@@
         Points are differences between observed transit times and the
-        constant-period ephemeris we fit to literature transits.
-        Black points are the more precise half of literatures times;
-        gray points are the less precise half.  The blue curves
-        represent the $\pm 1\sigma$ credible interval around the
-        constant-period ephemeris.  The binned TESS observation (red
-        point) arrived $77.8 \pm 10.7\ {\rm seconds}$ before it was
-        expected.
+        constant-period ephemeris we fit to literature times.  Black
+        points are the more precise half of literature times; gray
+        points are the less precise half.  Blue curves represent the
+        $\pm 1\sigma$ credible interval around the constant-period
+        ephemeris.  The binned TESS observation (red point) arrived
+        $77.8 \pm 10.8\ {\rm seconds}$ before it was expected.
         \label{fig:arrived_early}
@@
 
-Using the times, relative fluxes, and errors determined above, we ran
-Box Least Squares (BLS) to estimate the transit epoch, period, and
-duration \citep{kovacs_box-fitting_2002}.  Using these parameters, we
-isolated each transit to within $\pm$10 transit durations of its
-midtime.  In each transit window, we then simultaneously fit for a model
-transit, plus a local linear trend.  Our transit model uses the analytic
+Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
+estimate the transit epoch, period, and duration
+\citep{kovacs_box-fitting_2002}.  Using these parameters, we isolated
+each transit to within $\pm$10 transit durations of its midtime.  In
+each transit window, we then simultaneously fit for a model transit,
+plus a local linear trend.  Our transit model uses the analytic
 formulae calculated by \citet{mandel_analytic_2002} and implemented by
@@
 
-We allowed four free parameters per transit: the
-time of mid-transit $t_{\rm tra}$, the planet-to-star radius ratio
-$R_{\rm p}/R_\star$, and the slope and intercept of the line.
+We allowed four free parameters per transit: the time of mid-transit
+$t_{\rm tra}$, the planet-to-star radius ratio $R_{\rm p}/R_\star$,
+and the slope and intercept of the line.
 % Our priors were wide uniform distributions centered on the BLS
-% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope of
-% zero and an intercept of unity for the line.
+% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope
+% of zero and an intercept of unity for the line.
 We fixed the remaining transit parameters as follows.  We set the
@@
 assumed a quadratic limb-darkening law, with coefficients interpolated
-from the \citet{claret_limb_2017} tables.  For the period, $a/R_\star$,
-and inclination, we adopted the values from~\citet{petrucci_no_2013}.
-These are in reasonable agreement with the parameters reported by
-\citet{gillon_improved_2009}, \citet{southworth_high-precision_2009},
-and \citet{huitson_gemini_2017}.
+from the \citet{claret_limb_2017} tables.  For the period,
+$a/R_\star$, and inclination, we adopted the values
+from~\citet{petrucci_no_2013}.  These are in reasonable agreement with
+the parameters reported by \citet{gillon_improved_2009},
+\citet{southworth_high-precision_2009}, and
+\citet{huitson_gemini_2017}.
 
@@
 \citet{goodman_ensemble_2010} and implemented by
-\citet{foreman-mackey_emcee_2013}.  During this process, we used the
-\texttt{corner} software to check that our posteriors were being
-well-sampled~\citep{corner_2016}.
-
-After performing the initial fit, we set the photometric error bars to
+\citet{foreman-mackey_emcee_2013}.
+% During this process, we used the \texttt{corner} software to check
+% that our posteriors were being well-sampled~\citep{corner_2016}.
+After completing an initial fit, we set the photometric error bars to
 be equal to the standard deviation of the in-transit points of the
-residual lightcurve.  We then reperformed the same fitting procedure,
-using the empirically determined photometric errors.
+residual lightcurve.  We then reperformed the fitting procedure, using
+the empirically determined photometric errors.
 
 % TODO: final numbers
-To verify that the measured uncertainties are plausible, we computed the
-reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
-times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
-Visually inspecting residuals showed that the error variance was being
-overestimated, so we multiplied the measured TESS errors by a factor
-$f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
-midtime uncertainty for each transit from 30 to 22
-seconds.
-
-Figure~\ref{fig:lightcurves} shows the light curves, best-fit transit
-models, and residuals.  Table~1 reports the mid-transit times and
-their uncertainties.
-Taking the mean and standard deviation of the measured planet to star
-radius ratio, we find $R_{\rm p}/R_\star=0.1538\pm0.0013$, in
-agreement with previous determinations
+To verify that the measured uncertainties are plausible, we computed
+the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
+transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
+freedom.  Visually inspecting residuals showed that the error variance
+had been overestimated, so we multiplied the measured TESS errors by a
+factor $f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered
+the mean midtime uncertainty for each transit from 30 to 22 seconds.
+
+Figure~\ref{fig:lightcurves} shows the light curves, best-fit models,
+and residuals.  Table~1 reports the mid-transit times and their
+uncertainties.  Taking the mean and standard deviation of the measured
+planet to star radius ratio, we find $R_{\rm
+p}/R_\star=0.1538\pm0.0013$, in agreement with previous determinations
 \citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
 Binning the residuals to 1-hour windows and taking the median absolute
-deviation, the residual lightcurves have a MAD of $594\,{\rm
-ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
-\citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
-noise\footnote{\url{github.com/lgbouma/tnm}} of $410\,{\rm ppm}\,{\rm
-hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
-expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
+deviation, the lightcurves have a MAD of $594\,{\rm ppm}\,{\rm
+hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
+photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  At
+$T\approx11.8$, with a 9-pixel aperture read noise is expected to
+contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
 quadrature.  The predicted zodiacal background contribution is
 $673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
-\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015}
-used in planet detection simulations.  The predicted level
-of zodiacal noise seems to have been an overestimate, given the observed
-RMS of WASP-4.  A more detailed assessment of TESS's photometric
-performance is beyond the scope of this work.
+\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
+in planet detection simulations.  This zodiacal noise prediction seems
+to have been an overestimate, given the observed MAD of WASP-4.  A
+more detailed assessment of TESS's photometric performance is beyond
+the scope of this work.
+
+\subsection{Star and planet parameters}
+\label{sec:system_parameters}
+
+\begin{deluxetable}{lccc}
+\tabletypesize{\scriptsize}
+\tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
+\tablenum{2}
+
+\tablehead{
+\colhead{Parameter} & \colhead{Value} & \colhead{Error} & \colhead{Comment}
+}
+
+\startdata
+{\it Transit/RV parameters:} & & & \\
+  $R_{\rm p}/R_\star$                        & $0.1538$               & $\pm 0.0013$                & A \\
+  $i$~[deg]                                  & $88.52$                & $+0.39$, $-0.26$            & B \\
+  $a/R_\star$                                & $5.463$                & $+0.025$, $-0.020$          & B \\
+  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & C \\
+{\it Stellar parameters:} & & & \\
+  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & D \\
+  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & D \\
+  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & D \\
+  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & E \\
+  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & G \\
+  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & F \\
+  $\rho_\star$~[g~cm$^{-3}$]                 & $1.722$                & $+0.024$, $-0.019$          & A \\
+  $M_\star$~[M$_{\odot}$]                    & $0.870$                & $+0.085$, $-0.077$          & F \\
+  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & H \\
+{\it Planetary parameters:} & & & \\
+  $a$~[AU]                                   & $0.0227$               & $\pm 0.0007$                & F \\
+  $M_p$~[M$_{\rm Jup}$]                      & $1.192$                & $+0.091$, $-0.086$          & F \\
+  $R_p$~[R$_{\rm Jup}$]                      & $1.337$                & $\pm 0.048$                 & F \\
+\enddata
+
+\tablecomments{
+  (A) From TESS lightcurves (\S~\ref{sec:measurement}).
+  (B) \citet{hoyer_tramos_2013}.
+  (C) \citet{triaud_spin-orbit_2010}.
+  (D) From HARPS spectra \citep{doyle_accurate_2013}.
+  (E) \citet{stassun_accurate_2017}.
+  (F) From functions of stellar, transit or RV parameters as discussed in text.
+  (G) Gaia DR2.
+  (H) \citet{stassun_TIC_2018}.
+}
+
+\end{deluxetable}
+
+\citet{doyle_accurate_2013} found accurate spectroscopic parameters
+for WASP-4 using high S/N HARPS spectra.  Using their (T$_{\rm eff}$,
+$\log g$, $[{\rm Fe/H}]$) and the available {\it Gaia} parallaxes, we
+compute the remaining system parameters empirically.
+
+\citet{stassun_accurate_2017} found the bolometric flux from
+broad-band photometry.  Combining the implied angular radius with the
+{\it Gaia} DR2 parallax, and adjusting for the parallax offset
+\citep{stassun_evidence_2018}, we obtain the stellar radius.  For the
+stellar mass, we started with the $a/R_\star$ value found by
+\citet{hoyer_tramos_2013} from 26 ground-based transits.  The implied
+semi-major axis and stellar density follow \citep{seager_unique_2003}.
+Combining the density and radius gives the stellar mass.  We get the
+planet radius using our measured $R_{\rm p}/R_\star$.  Finally, for
+the planet mass we use the observed RV semi-amplitude and inclination
+angle from the transit
+\citep{triaud_spin-orbit_2010,hoyer_tramos_2013}.
+
+The resulting parameters are in Table~3, and broadly agree with those
+of previous authors
+\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
+We use them for the remaining analysis.  Comparing the empirical
+result to the Yonsei-Yale evolutionary track predicted by the
+spectroscopic parameters, we see that WASP-4  has an age $\approx
+7\,{\rm Gyr}$, and is securely in the main-sequence stage of
+evolution.
 
@@
         model is a poor fit to the observed times (gray line).  Black
-        points are the more precise half of measured times; gray points
-        are the less precise half.  The blue curve is the best-fit model
-        with a constant period derivative.  The orange curve is the
-        best-fit model assuming apsidal precession causes the observed
-        timing deviations.
+        points are the more precise half of measured times; gray
+        points are the less precise half.  The blue curve is the
+        best-fit model with a constant period derivative.  The orange
+        curve is the best-fit model assuming apsidal precession causes
+        the observed timing deviations. The binned TESS point (star)
+        is for display purposes only.
         \label{fig:times}
@@
 was fit as a free parameter. We also required that the time system be
-clearly documented.
-
-Table~2 shows the occultation times.  Since there are fewer available,
-we included all the occultation measurements we could find. The
-tabulated values have been corrected for the light-travel time across
-the diameter of the orbit by subtracting $2a/c = 22.8$ seconds from
-the observed time.
+clearly documented.  The occultation times are in Table~3.  Since
+there are fewer available, we included all the occultation
+measurements we could find. The tabulated values have been corrected
+for the light-travel time across the diameter of the orbit by
+subtracting $2a/c = 22.8$ seconds from the observed time.
 
 We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point.  
-Though their latest two epochs fell slightly below the prediction of a
-linear ephemeris (Panel C of their Figure~6), they did not find a need
-to fit a quadratic function to the O-C values.
-We verified that the \citet{hoyer_tramos_2013} times, as well as all
-others, were on the BJD$_{\rm TDB}$ time
-standard using the \citet{eastman_achieving_2010} calculator.
-We set the zero-point epoch to be as
-close as possible to the average of the transit times, weighted as the
-inverse midtime variance. This minimizes the covariance between the
-transit epoch and the period \citep{gibson_gemini_2013}.
-
-Important contributions to the timing dataset are as follows.  The
-earliest epoch in our fit is from EulerCam on the 1.2-m Euler telescope
-\citep{wilson_wasp-4b_2008}.  The second epoch comes from
-$z$-band photometry acquired by \citet{gillon_improved_2009} at the VLT
-8.2-m with FORS2.  Subsequent transits came from
+starting point.  Though their latest two epochs fell slightly below
+the prediction of a linear ephemeris (Panel C of their Figure~6), they
+did not find a need to fit a quadratic function to the O-C values.  We
+verified that all times were on the BJD$_{\rm TDB}$ time standard
+using the \citet{eastman_achieving_2010} calculator.  We set the
+zero-point epoch to be as close as possible to the average of the
+midtimes, weighted as the inverse midtime variance. This minimizes the
+covariance between the transit epoch and the period
+\citep{gibson_gemini_2013}.
+
+Important contributions to the timing dataset are as follows.  Our
+earliest epoch is from EulerCam on the 1.2-m Euler telescope
+\citep{wilson_wasp-4b_2008}.  The second epoch comes from $z$-band
+photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
+with FORS2.  Subsequent transits came from
 \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
@@
 the TDB reference for their reported midtimes.},
-\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
+\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011},
 \citet{nikolov_wasp-4b_2012}, \citet{hoyer_tramos_2013}, and
-\citet{ranjan_atmospheric_2014}.
+\citet{ranjan_atmospheric_2014}.  Finally, \citet{huitson_gemini_2017}
+acquired optical transmission spectra with the 8.1-m Gemini South
+telescope over 2011 to 2014, one transit per season.  The per-point
+standard deviation of their lightcurves is a few hundred parts per
+million, within a factor of 3 from the photon noise limit.  The
+uncertainties on their reported transit times~--~on average
+5.6~seconds~--~ are small, but possible given the quality of their
+data~\citep{carter_analytic_2008}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
@@
 %photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
-%the MPG/EOS-2.2~m at La Silla.
-%\citet{hoyer_tramos_2013} reported nine transit observations taken with
-%the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
-%SOI on the SOAR 4.2-m telescope.
-%\citet{hoyer_tramos_2013} also
-%homogeneously redetermined the transit times from all previous studies.
-%After the study by \citet{hoyer_tramos_2013},
-%\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
-%using HST's WFC3 in both transit and secondary eclipse.  We adopt only
-%their transit time, since the epoch of occultation was not fit as a free
-%parameter.'
-Finally, \citet{huitson_gemini_2017} acquired optical transmission
-spectra with the 8.1-m Gemini South telescope over 2011 to 2014, one
-transit per season.  The per-point standard deviation of their
-lightcurves is a few hundred parts per million, within a factor of 3
-from the photon noise limit.  The uncertainties on their reported
-transit times~--~on average 5.6~seconds~--~ are small, but possible
-given the quality of their data~\citep[cf.][]{carter_analytic_2008}.
-
-% There have been other studies of WASP-4b for which the transit timing
-% data are either hetereogenous, not available, or not useable.  We
-% omitted the timing measurements from
+%the MPG/EOS-2.2~m at La Silla.  \citet{hoyer_tramos_2013} reported
+%nine transit observations taken with the the Y4KCAM on the SMARTS 1-m
+%telescope, and another three with the SOI on the SOAR 4.2-m
+%telescope.  \citet{hoyer_tramos_2013} also homogeneously redetermined
+%the transit times from all previous studies.  After the study by
+%\citet{hoyer_tramos_2013}, \citet{ranjan_atmospheric_2014} acquired
+%near-IR spectra of WASP-4b using HST's WFC3 in both transit and
+%secondary eclipse.  We adopt only their transit time, since the epoch
+%of occultation was not fit as a free parameter.'
+
+% There have been other studies of WASP-4b for which the transit
+% timing data are either hetereogenous, not available, or not useable.
+% We omitted the timing measurements from
 % \citet{southworth_high-precision_2009}, since there were technical
@@
 % observation~\citep{nikolov_wasp-4b_2012}.  The timing study
-% by~\citet{petrucci_no_2013} ruled out transit timing variations larger
-% than 54~seconds over the previous 5 years, but did not report transit
-% times.  \citet{baluev_benchmarking_2015} collected transit times for
-% many hot Jupiters, including WASP-4b, and included times from the
-% Exoplanet Transit Database
+% by~\citet{petrucci_no_2013} ruled out transit timing variations
+% larger than 54~seconds over the previous 5 years, but did not report
+% transit times.  \citet{baluev_benchmarking_2015} collected transit
+% times for many hot Jupiters, including WASP-4b, and included times
+% from the Exoplanet Transit Database
 % (ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
-% \citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
-% sources, their timestamps are less clearly documented, and the times
-% are thus more prone to systematic errors, we omit them from
-% consideration.
-% %We consider the effects of including ETD data in
-% %Appendix~\ref{sec:verify_archival_times}.
+% \citep{poddany_ETD_2010}.  Since the ETD data come from
+% heterogeneous sources, their timestamps are less clearly documented,
+% and the times are thus more prone to systematic errors, we omit them
+% from consideration.  %We consider the effects of including ETD data
+% in %Appendix~\ref{sec:verify_archival_times}.
 % \citet{may_mopss_2018} obtained optical spectra using IMCAS on
@@
 % report transit times.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 There are fewer available occultation times.
-\citet{beerer_secondary_2011} observed two occultations of WASP-4b using
-warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
+\citet{beerer_secondary_2011} observed two occultations of WASP-4b
+using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
 \citet{caceres_ground-based_2011} detected an occultation from the
 ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
-specifying the time standard.  We correspondingly added $69.184$ seconds
-of uncertainty to their reported errors, in quadrature.
-%Finally, as
-%previously mentioned, \citet{ranjan_atmospheric_2014} acquired
-%occultation data with HST but kept the epoch as a fixed parameter.
-\citet{zhou_secondary_2015} acquired
-occultation data from the Anglo-Australian Telescope, focusing on the
-eclipse depth, but also fitting for $e\cos\omega$.  We converted their
-$e\cos\omega$ result into a midtime, using their ephemeris and the
-standard formula \citep[{\it
-e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
+specifying the time standard.  We correspondingly added $69.184$
+seconds of uncertainty to their reported errors, in quadrature.
+%Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}
+%acquired occultation data with HST but kept the epoch as a fixed
+%parameter.
+\citet{zhou_secondary_2015} acquired occultation data from the
+Anglo-Australian Telescope, focusing on the eclipse depth, but also
+fitting for $e\cos\omega$.  We converted their $e\cos\omega$ result
+into a midtime, using their ephemeris and the standard formula
+\citep[{\it e.g.},][]{winn_exoplanet_2010}
 \begin{equation}
@@
 This gives us a total of 4 occultations.  Although the occultation
-times are uncertain compared to the transits, we include them for
-completeness, and because they help in modeling the timing
-variations.
+times are uncertain compared to the transits, they nonetheless help in
+modeling the timing variations.
 
@@
 literature times. The result is shown in
-Figure~\ref{fig:arrived_early}. The TESS observations seem to have
+Figure~\ref{fig:arrived_early}. The TESS observations 
 arrived early.  One immediate concern was systematic errors in either
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Assuming the error bars on the transit midtimes are gaussian and
@@
 % from wide uniform distributions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$.
-%In other words, assuming the linear model is true, there is a
-%$10^{-23}$ probability that simply by chance we would see data that
-%favors the quadratic model more strongly than the observations.
-Clearly, it is more likely that some systematic error could be
-affecting one of the most important sets of times: either the earliest
-times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591,
-and 851), or the TESS times.  We have taken every precaution against
-systematic offsets in the TESS dataset
-(Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by their
-reported times (J.-M. D\'esert, priv.\
+corresponds to $p \approx 10^{-23}$.  Clearly, it is more likely that
+some systematic error could be affecting one of the most important
+sets of times: either the earliest times, the Huitson et al.\ Gemini
+times (epoch numbers 44, 322, 591, and 851), or the TESS times.  We
+have taken every precaution against systematic offsets in the TESS
+dataset (Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by
+their reported times (J.-M. D\'esert, priv.\
 comm.)\footnote{\citealt{huitson_gemini_2017} corrected the times to
@@
 The quadratic model fits better than the precession model.  It is
-favored by $\Delta \chi^2 = 3.0$, and has two
-fewer free parameters.  Another useful heuristic for model comparison is
-the Bayesian Information Criterion (BIC),
+favored by $\Delta \chi^2 = 3.0$, and has two fewer free parameters.
+Another useful heuristic for model comparison is the Bayesian
+Information Criterion (BIC),
 \begin{equation}
@@
 BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor of $\approx
-4\times10^{4}$.  In the language of~\citet{kass_bayes_1995}, this is
-``decisive evidence'' for the quadratic model describing the data better
-than the precession model.  The Akaike Information Criterion
+4\times10^{4}$.  The Akaike Information Criterion
 similarly favors the orbital decay model, by $\Delta {\rm AIC} = 7$.
+This is strong evidence that the quadratic model describes the
+data better than the precession model \citep{kass_bayes_1995}.
 
@@
 \dot{P}
-    = - (3.82^{+0.38}_{-0.37})\times 10^{-10}
-    = - (12.1^{+1.2}_{-1.2})\,{\rm ms}\,{\rm yr}^{-1}.
+  = - (3.82 \pm 0.38)\times 10^{-10}
+  = - (12.1) \pm 0.12\,{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
 For comparison, the period derivative seen by
-\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b is
-$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
-If both planets are decaying, WASP-4b would be inspiraling at about half
-the rate as WASP-12b.
+\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b
+is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If both planets
+are decaying, WASP-4b would be inspiraling at about half the rate as
+WASP-12b.
 
 %TODO: final numbers
-If we instead assume the precession model, the best-fit
-eccentricity is
+If we instead assume the precession model, the best-fit eccentricity
+is
 \begin{equation}
@@
 The longitude of periapse correspondingly advances by $\dot{\omega}
-= 14.8^{+5.4}_{3.8}\ {\rm degrees}\,{\rm yr}^{-1}$, and the
+= 14.8^{+5.4}_{-3.8}\ {\rm degrees}\,{\rm yr}^{-1}$, and the
 precession period would be $24^{+8}_{-6}$ years.
 
-The median parameters and their standard deviations are reported for all
-the models in Table~3.  In short, a linear ephemeris does not
-describe the observed times.  Both orbital decay and precession can fit
-the data, but the precession model is statistically disfavored.
+The median parameters and their standard deviations are reported for
+the three models in Table~4.  In short, a linear ephemeris does not
+describe the observed times.  Both orbital decay and precession can
+fit the data, but the precession model is statistically disfavored.
 
@@
   \caption{
-    {\bf WASP-4b has a short eccentricity
-    	damping time and medium decay time compared to other hot Jupiters.
-    	It seems to orbit a main sequence host.
+    {\bf WASP-4b has a short eccentricity damping time and medium
+    decay time compared to other hot Jupiters.  It seems to orbit a
+    main sequence host.
     }
@@
 orbital decay time of WASP-4b compared to other transiting giant
-planets.  Though there are roughly 20 hot Jupiters that might be
-expected to decay faster, some orbit stars with minimal convective
-zones ({\it e.g.}, WASP-18), and most are not as well-observed as
-WASP-4.
-
-For the sake of argument,
-let us assume that the observed timing
-variation is caused entirely by orbital decay.  If the period
-continues to decrease at a fixed rate, it will reach zero after
+planets.  Though roughly 20 hot Jupiters might be expected to decay
+faster, some orbit stars with minimal convective zones ({\it e.g.},
+WASP-18), and most are not as well-observed as WASP-4.
+
+For the sake of argument, let us assume that the timing variation is
+caused entirely by orbital decay.  If the period continues to decrease
+at a fixed rate, it will reach zero after
 \begin{equation}
@@
 interaction in a binary applies \citep{zahn_tidal_1977}.  The fluid
-response is parametrized by a modified\footnote{For stars,
-$k_\star \sim \mathcal{O}(10^{-2})$, so it is important to explicitly
+response is parametrized by a modified\footnote{For stars, $k_\star
+\sim \mathcal{O}(10^{-2})$, so it is important to explicitly
 distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
@@
 centrally concentrated. $Q_\star$ is the ratio of the energy stored in
-the equilibrium deformation of the star, divided by the energy
-lost to heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.
-A larger $Q_\star$ implies less efficient dissipiation.  Though the
+the equilibrium deformation of the star, divided by the energy lost to
+heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A
+larger $Q_\star$ implies less efficient dissipiation.  Though the
 fluid mechanics responsible for the dissipation should depend on the
@@
 rotating, and the eccentricity is small, then the semi-major axis and
-eccentricity evolve as
-\citep[Appendix B of][]{metzger_optical_2012}
+eccentricity evolve as \citep[Appendix B of][]{metzger_optical_2012}
 \begin{align}
@@
 \end{equation}
-Using our observed value of $\dot{P}$, the current modified quality
+Using the observed value of $\dot{P}$, the current modified quality
 factor of WASP-4 would be
@@
 This is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
-10^5$, based on the motion of the Galilean moons
-\citep{lainey_strong_2009}.  For stars, different $Q_\star'$ values have
-been estimated in different contexts.  Studies of the 
-eccentricity distribution of stellar binaries have yielded $10^5 \gtrsim
-Q_\star' \gtrsim 10^{7}$ \citep[{\it
-e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
+10^5$, based on the orbits of the Galilean moons
+\citep{lainey_strong_2009}.  For stars, studies of the binary
+eccentricity distribution have yielded $Q_\star' \approx 10^5 - 10^7$
+\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
 geller_direct_2013,milliman_wiyn_2014}.  However the dissipation rates
-may be different in hot Jupiter systems, since the tidal forcing
-frequencies and amplitudes are different.  In fact, for the observed
-population of hot Jupiters, \citet{penev_constraining_2012} argued that
-$Q_\star' > 10^7$ would be required at some phase of hot Jupiter
-evolution at the 99\% level, or else the observed population would not
-exist (though \citealt{birkby_wts-2_2014} point out more efficient
-dissipation would be allowed if the hot Jupiters arrived later).
-Developing a generative model for the orbital separation distribution of
-the hot Jupiter population, \citet{cameron_hierarchical_2018} similarly
-found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.
-
-Using an empirical approach, \citet{penev_empirical_2018} reported a
-$Q_\star'$ measurement for WASP-4.  In earlier work,
-\citet{penev_hats-18b_2016} noted that HATS-18b and WASP-19b orbited
-host stars whose isochronal ages are consistent with the Sun, but
-whose stellar rotation periods are about 10 days
-(\citealt{pont_empirical_2009} pointed out similar evidence for hot
-Jupiter host spin-up).  The expected rotation periods at solar ages
-are 20 to 30 days
-(\citealt{schatzman_theory_1962,skumanich_time_1972}, and measurements
-by \citealt{barnes_rotation_2016} in the 4\,Gyr-old M67 cluster).
-Assuming that the difference between the observed and predicted
-stellar rotation period was due to tides depositing angular momentum
-in the star, \citet{penev_empirical_2018} modeled the evolution of hot
-Jupiter systems under the influence of a magnetized wind and a
-constant phase-lag tide.  They found two-sided limits for $Q_\star'$
-in 35 of the 188 systems they examined.  In these 35 systems, they
-found that $Q_\star'$ varied from $10^5$ at tidal periods of 2 days to
-$10^7$ at tidal periods of 0.5 days.  For WASP-4 specifically, their
-method gave $Q_\star' \approx (1.2^{+1.0}_{-0.5})\times10^7$. This is
-more than 2 orders of magnitude less efficient than the dissipation
-rate inferred from the decaying orbit would imply.
-
-A different model for tidal dissipation was discussed by
+may differ in hot Jupiter systems, since the tidal forcing frequencies
+and amplitudes are different.  Observationally-driven studies of hot
+Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, often using
+different models
+\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
+For instance, motivated by the rapid rotation of some hot Jupiter
+hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
+\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
+systems under the influence of a magnetized wind and a constant
+phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
+(1.2^{+1.0}_{-0.5})\times10^7$.  This disagrees strongly with the
+value we infer from the apparent period change.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% In fact, for the observed population of hot Jupiters,
+% \citet{penev_constraining_2012} argued that $Q_\star' > 10^7$ would
+% be required at some phase of hot Jupiter evolution at the 99\%
+% level, or else the observed population would not exist (though
+% \citealt{birkby_wts-2_2014} point out more efficient dissipation
+% would be allowed if the hot Jupiters arrived later).  Developing a
+% generative model for the orbital separation distribution of the hot
+% Jupiter population, \citet{cameron_hierarchical_2018} similarly
+% found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.
+
+% Using an empirical approach, \citet{penev_empirical_2018} reported a
+% $Q_\star'$ measurement for WASP-4.  In earlier work,
+% \citet{penev_hats-18b_2016} noted that HATS-18b and WASP-19b orbited
+% host stars whose isochronal ages are consistent with the Sun, but
+% whose stellar rotation periods are about 10 days
+% (\citealt{pont_empirical_2009} pointed out similar evidence for hot
+% Jupiter host spin-up).  The expected rotation periods at solar ages
+% are 20 to 30 days
+% (\citealt{schatzman_theory_1962,skumanich_time_1972}, and
+% measurements by \citealt{barnes_rotation_2016} in the 4\,Gyr-old M67
+% cluster).  Assuming that the difference between the observed and
+% predicted stellar rotation period was due to tides depositing
+% angular momentum in the star, \citet{penev_empirical_2018} modeled
+% the evolution of hot Jupiter systems under the influence of a
+% magnetized wind and a constant phase-lag tide.  They found two-sided
+% limits for $Q_\star'$ in 35 of the 188 systems they examined.  In
+% these 35 systems, they found that $Q_\star'$ varied from $10^5$ at
+% tidal periods of 2 days to $10^7$ at tidal periods of 0.5 days.  For
+% WASP-4 specifically, their method gave $Q_\star' \approx
+% (1.2^{+1.0}_{-0.5})\times10^7$. This is more than 2 orders of
+% magnitude less efficient than the dissipation rate inferred from the
+% decaying orbit would imply.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A theoretically-motivated model for tidal dissipation was proposed by
 \citet{essick_orbital_2016}, who considered the nonlinear response of
@@
 the convective zone, and braking near the stellar core.  In their
-approach, the stellar tidal quality factors in hot Jupiter systems
-vary from $Q_\star' \approx 10^5 - 10^6$.  Using their Equation 26,
-the prediction for WASP-4 is that it has $Q_\star' = 7\times10^5$.
-This is one order of magnitude larger than implied by the
-observed period change.
+approach, the stellar quality factors in hot Jupiter systems vary from
+$Q_\star' \approx 10^5 - 10^6$.  Using their Equation 26, the
+prediction for WASP-4 is $Q_\star' = 7\times10^5$.  This is an order
+of magnitude larger than implied by the observed period change.
 
@@
 The x-axis is the effective temperature from \citet{bonomo_gaps_2017},
-which for WASP-4 agrees within $1\sigma$ of that reported by
-\citet{petrucci_no_2013}.  By visual inspection of the HR diagram,
-WASP-4 shows litle evidence of being evolved.  By a similar argument
-though, WASP-12 would not clearly stand out as a candidate subgiant,
-and \citet{weinberg_tidal_2017} suggest it may in fact be a subgiant
-(though \citealt{bailey_understanding_2019} do not find evidence that
-supports this suggestion).  A more thorough isochronal analysis would
-be of interest, particularly if the transit and occultation times
-continue to vary.  Such an analysis is beyond the scope of this study.
+which for WASP-4 agrees within $1\sigma$ of that from Table~2.
+Inspecting the HR diagram, WASP-4 shows litle evidence of being
+evolved, in agreement with our analysis from
+\S~\ref{sec:system_parameters}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% By a similar argument though, WASP-12 would not clearly stand out as a
+% candidate subgiant, and \citet{weinberg_tidal_2017} suggest it may in
+% fact be a subgiant (though \citealt{bailey_understanding_2019} do not
+% find evidence that supports this suggestion).  A more thorough
+% isochronal analysis would be of interest, particularly if the transit
+% and occultation times continue to vary.  Such an analysis is beyond
+% the scope of this study.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 decay, it would imply a tidal dissipation efficiency stronger than
-typical orbit-averaged predictions.  It might be possible that we are
-observing at a special time, when the planet is near resonance with the
-some stellar oscillation mode, and the energy dissipation is
-exceptionally large \citep{ogilvie_tidal_2014,essick_orbital_2016}.
-Tidal dissipation rates might also be increased if the star is turning
-off the main sequence.  WASP-4 shows little evidence for this; it is
-unclear that tidal decay could explain its changing period.
+most predictions.  It might be possible that we are observing at a
+special time, shortly after the planet's inward migration, or when the
+planet is near resonance with a stellar oscillation mode
+\citep{ogilvie_tidal_2014,essick_orbital_2016}.  Tidal dissipation
+rates might also be increased if the star is turning off the main
+sequence.  All of these explanations are strained; it is unclear that
+tidal decay could explain the observed timing variations without
+fine-tuning.
 
@@
 
-Let us instead assume that the entire timing variation can be explained
-as being a portion of an apsidal precession cycle.  The required orbital
-eccentricity is then $e\approx 1.6\times10^{-3}$, and one full precession
-period would take about 24 years (twice the
-current observing baseline).
-
-\citet{ragozzine_probing_2009} calculated apsidal precession periods for
-hot Jupiters ranging between of order 10 and 100 years.  They
-highlighted that for many hot Jupiters, including WASP-4b, over 90\% of
-the apsidal precession comes from the tidal bulge of the planet.
+Let us instead assume that the timing variation can be explained as
+being a portion of an apsidal precession cycle.  The required orbital
+eccentricity is then $e\approx 1.6\times10^{-3}$, and one full
+precession period would take about 24 years.
+
+\citet{ragozzine_probing_2009} calculated apsidal precession periods
+for hot Jupiters ranging between of order 10 and 100 years.  They
+highlighted that for many hot Jupiters, including WASP-4b, over 90\%
+of the apsidal precession comes from the tidal bulge of the planet.
 Precession from general relativity, the planet's rotational bulge, and
-the star's rotational and tidal bulges are minor contributors.
+the star's rotational and tidal bulges are minor additions.
 %\citet{ragozzine_probing_2009} also calcuated that WASP-4b was expected
@@
 \citet{ragozzine_probing_2009} also noted that if the planet has
-non-zero eccentricity, a measurement of $\dot{\omega}$ can be converted
-into a measurement of the planetary Love number, $k_{2,\rm p}$.
-From their Equation 14, the implied Love number for WASP-4b is
+non-zero eccentricity, a measurement of $\dot{\omega}$ can be used to
+measure the planet's Love number, $k_{2,\rm p}$.  From their Equation
+14, the implied Love number for WASP-4b is
 \begin{equation}
@@
 \end{equation}
-For comparison, $k_{2,{\rm Jup}} \approx 0.59$
-(\citealt{wahl_tidal_2016}, but see \citealt{ni_empirical_2018}).  A
-uniform density sphere has $k_2 = 1.5$.  Our uncertainties on
-$k_{2,{\rm p}}$ are large because the eccentricity, reference time, and
-${\rm d}\omega/{\rm d}E$ are correlated, and the measured occultation
-times only barely constrain them (Figure~\ref{fig:future}).
-
-\subsubsection{Empirical constraints on small eccentricities in hot
-Jupiters}
-
-Could the eccentricity truly be non-zero?  For WASP-4b, the eccentricity
-damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e = 0.29 (Q_{\rm
-p}'/10^5) \, {\rm Myr}$.  The system is well over $1\,{\rm
-Gyr}$ \citep{winn_transit_2009} old, so any initial eccentricity should
-have been erased.
+For comparison, $k_{2,{\rm Jup}} \approx 0.55$
+\citep{wahl_tidal_2016,ni_empirical_2018}.  A uniform density sphere
+has $k_2 = 1.5$.  The uncertainties on $k_{2,{\rm p}}$ are large
+because the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$
+are correlated, and the measured occultation times only barely
+constrain them (Figure~\ref{fig:future}).
+
+\subsubsection{Observations of small eccentricities in hot
+Jupiter systems}
+
+Could the eccentricity truly be non-zero?  For WASP-4b, the
+eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
+0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
+$1\,{\rm Gyr}$ \citep{winn_transit_2009} old, so any initial
+eccentricity should have been erased.
 
@@
 0.01580^{+0.0023}_{-0.0018}$), and WASP-38b ($e =
-0.02780^{+0.003}_{-0.0028}$).  For the 231 planets studied by
-\citet{bonomo_gaps_2017}, 2 (13) had $2\sigma$ upper limits of $e <
-0.01$ ($e<0.02$).  For WASP-4b, the strongest $2\sigma$ upper-limit on
-its eccentricity from RVs is $e<0.011$
-\citep{husnoo_observational_2012,bonomo_gaps_2017}.
+0.02780^{+0.003}_{-0.0028}$).
+% For the 231 planets studied by \citet{bonomo_gaps_2017}, 2 (13) had
+% $2\sigma$ upper limits of $e < 0.01$ ($e<0.02$).  For WASP-4b, the
+% strongest $2\sigma$ upper-limit on its eccentricity from RVs is
+% $e<0.011$ \citep{husnoo_observational_2012,bonomo_gaps_2017}.
 
@@
 offset could be explained by detailed modeling of the planetary
-hot-spot, which can shift occultation times by up to a few
-minutes \citep{williams_resolving_2006}.  Including an updated
-hot-spot model, \citet{agol_climate_2010} reported
-$e\cos\omega < 0.00023$ at 2$\sigma$, perhaps the strongest
-eccentricity limit in place
-for a hot Jupiter.  A similar saga occurred for HD 209458b, for which
-\citet{winn_measurement_2005} combined Spitzer secondary eclipses with
-RV data and reported a small non-zero eccentricity.  They suspected that
-the data were consistent with zero eccentricity, and that the lower
-bound on $e$ was an artifact of an imperfect limb-darkening model.
-\citet{crossfield_spitzer_mips_2012} eventually ruled out the \citet{winn_measurement_2005} values, by determining
-$e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b, for
-which the discovery RVs and photometry gave a small finite eccentricity
-measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.  A prompt
-ground-based occultation measurement by
-\citet{lopez-morales_day-side_2010} concurred that a small eccentricity
-was favored with weak statistical significance. Further follow-up found
-occultation times consistent with circular orbits; $| e\cos\omega| <
-0.0040 $ \citep{campo_orbit_2011,croll_near-infrared_2011}.
-
-In summary, empirical constraints on the ellipticity of hot Jupiter
-orbits at the $e \lesssim 10^{-3}$ level are present only for HD 189733b
-and HD 209458b, from occultation timing.  Once the timing offsets are of
-order tens of seconds, eccentricities from occultation timing may have
-systematic uncertainties because of the hot-spot issue
+hot-spot, which can shift occultation times by up to a few minutes
+\citep{williams_resolving_2006}.  Including an updated hot-spot model,
+\citet{agol_climate_2010} reported $e\cos\omega < 0.00023$ at
+2$\sigma$, perhaps the strongest eccentricity limit in place for a hot
+Jupiter.  Similar sagas transpired for HD 209458b and WASP-12b, with
+early reports of small non-zero eccentricities
+\citep{winn_measurement_2005,hebb_wasp-12b_2009,lopez-morales_day-side_2010},
+only to be ruled out by the limits of subsequent followup:
+$|e\cos\omega| < 0.0007$ for HD 209458b, and $| e\cos\omega| < 0.0040$
+($2\sigma$) for WASP-12b
+\citep{crossfield_spitzer_mips_2012,campo_orbit_2011,croll_near-infrared_2011}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% A similar saga occurred for HD 209458b, for which
+% \citet{winn_measurement_2005} combined Spitzer secondary eclipses with
+% RV data and reported a small non-zero eccentricity.  They suspected that
+% the data were consistent with zero eccentricity, and that the lower
+% bound on $e$ was an artifact of an imperfect limb-darkening model.
+% \citet{crossfield_spitzer_mips_2012} eventually ruled out the \citet{winn_measurement_2005} values, by determining
+% $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b, for
+% which the discovery RVs and photometry gave a small finite eccentricity
+% measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.  A prompt
+% ground-based occultation measurement by
+% \citet{lopez-morales_day-side_2010} concurred that a small eccentricity
+% was favored with weak statistical significance. Further follow-up found
+% occultation times consistent with circular orbits; $| e\cos\omega| <
+% 0.0040 $ \citep{campo_orbit_2011,croll_near-infrared_2011}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To summarize, empirical constraints on the ellipticity of hot Jupiter
+orbits at the $e \lesssim 10^{-3}$ level are present only for HD
+189733b and HD 209458b, from occultation timing.  Once the timing
+offsets are of order tens of seconds, eccentricities from occultation
+timing may have systematic uncertainties because of the hot-spot issue
 \citep{williams_resolving_2006,agol_climate_2010}.  Except for
 WASP-18b, radial velocity measurements have provided eccentricity
-constraints only at the $e > 10^{-2}$ level.  We cannot empirically rule out
-hot Jupiter eccentricities at the level needed to assess WASP-4's
-putative eccentricity.
+constraints only at the $e > 10^{-2}$ level.  We therefore cannot
+empirically rule out hot Jupiter eccentricities at the level needed to
+assess WASP-4's putative eccentricity.
 
 \subsubsection{Theoretical expectations for small eccentricities in hot
-Jupiters}
-
-Lacking empirical data, we now turn to theoretical mechanisms that could
+Jupiter systems}
+
+Lacking empirical data, we turn to theoretical mechanisms that could
 produce small eccentricities.
@@
 %companion's mass.
-For the Kozai-Lidov mechanism to operate at maximum
-efficiency, from
+For the Kozai-Lidov mechanism to operate at maximum efficiency, from
 \citet{bailey_understanding_2019} Equation 20 we need
@@
 
-While either of the two preceding mechanisms could produce
-the needed eccentricity,
-it would be more
-satifying if long-term radial velocity monitoring showed an additional
-planet in the system.  The chances of one existing with a mass between
-1-13\,$M_{\rm Jup}$ and orbital semi-major axis between 1-20\,AU is
-about 50\%, and a linear trend is currently present at $\sim 2\sigma$
-\citep{knutson_friends_2014}.
+While either of the two preceding mechanisms could produce the needed
+eccentricity, it would be more satifying if long-term radial velocity
+monitoring showed an additional planet in the system.  The chances of
+one existing with a mass between 1-13\,$M_{\rm Jup}$ and orbital
+semi-major axis between 1-20\,AU is about 50\%, and a linear trend is
+currently present at $\sim 2\sigma$ \citep{knutson_friends_2014}.
 
 \paragraph{Fluctuations in the gravitational potential from convection}
-A separate mechanism that might produce the eccentricity is simply
-stellar convection \citep[][Section 7]{phinney_pulsars_1992}.  Phinney
-presents this idea by analogy with a pendulum, which is never really
-at rest. Instead, collisions with air molecules randomly perturb its
-momentum and energy, and the equipartition theorem of statistical
-mechanics tells us the average energy is $\langle E_{\rm pend} \rangle
-= k T_{\rm air}$.  In hot Jupiter systems, the ``air molecules'' are
-convective eddies, and their associated kinetic energy leads the orbit
-to never have exactly zero eccentricity.  In particular, from
-\citealt{phinney_pulsars_1992} Equation 7.33,
+A separate mechanism to pump the eccentricity involves the
+gravitational fluctuations from stellar convection \citep[][Section
+7]{phinney_pulsars_1992}.  Phinney presents this idea by analogy with
+a pendulum, which is never really at rest. Instead, collisions with
+air molecules randomly perturb its momentum and energy, and the
+equipartition theorem of statistical mechanics tells us the average
+energy is $\langle E_{\rm pend} \rangle = k T_{\rm air}$.  In hot
+Jupiter systems, the ``air molecules'' are convective eddies, and
+their associated kinetic energy perturbs the orbit away from
+circularity.  In particular, from \citealt{phinney_pulsars_1992}
+Equation 7.33,
 \begin{equation}
@@
 convection in the star is more important than convection in the
-planet, and $\langle e^2 \rangle^{1/2} < 10^{-5}$.
+planet, and $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% LB: I calculated the above using MESA grids from Penev+ 2018 to
+% estimate the core/envelope masses, but mentioning this is too
+% detailed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 for $\dot{v}_{\rm r}$ the time derivative of the radial velocity.
-Recently, \citet{bonomo_gaps_2017} collected all available HARPS data
-for the system, and reported no detectable slope at $3\sigma$.
-However, \citet{knutson_friends_2014} collected 5 Keck HIRES radial
+%Recently, \citet{bonomo_gaps_2017} collected all available HARPS data
+%for the system, and reported no detectable slope at $3\sigma$.
+%However,
+\citet{knutson_friends_2014} collected 5 Keck HIRES radial
 velocity measurements for WASP-4, and merged them with measurements
 from CORALIE and HARPS \citep{wilson_wasp-4b_2008,pont_determining_2011,husnoo_observational_2012}.
-Using these four datasets, \citet{knutson_friends_2014} found that
-$\dot{\gamma} = -0.0099^{+0.0052}_{-0.0054}$, or that the
-star can be accelerating towards us at most by $\dot{\gamma} > -
+Using these datasets, \citet{knutson_friends_2014} found 
+\begin{equation}
+\dot{v}_{\rm r} =
+   -0.0099^{+0.0052}_{-0.0054}
+   \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
+\end{equation}
+or that the
+star can be accelerating towards us at most by $\dot{v}_{\rm r} > -
 0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ (2$\sigma$).
-The associated $\dot{P}$
-must be less than $9\times 10^{-11}$.  
-%FIXME TODO final value
-%Our measured value is $\dot{P} = -(3.9^{+0.7}_{-0.5})\times 10^{-10}$.
-Radial acceleration could thus cause at most one-quarter of
-the observed period decrease (Equation~\ref{eq:dP_dt_obs}).
+The associated $\dot{P}$ must be less than $9\times 10^{-11}$.  
+Radial acceleration could thus cause at most one-quarter of the
+observed period decrease (Equation~\ref{eq:dP_dt_obs}).
 Improved radial velocity
 monitoring will help clarify the situation,
-particularly given the fact that WASP-4 shows short-term trends of
-$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
-tens of days \citep{husnoo_observational_2012}.
-Given the known spottedness
-of the system~\citep{sanchis-ojeda_starspots_2011}, these
-short-timescale variations likely have a contribution from stellar
-activity.
-
-There are two other small effects worth briefly mentioning.  The first
-is the \citet{shklovskii_possible_1970} effect, in which the star's
-proper motion leads to a changing radial velocity.  This would be
-apparent in the Doppler measurements described above.  The effect
-contributes a change of order $\dot{P} \sim P\mu^2 d/ c \sim
-6\times10^{-13}$.  The second effect, described by
-\citet{rafikov_stellar_2009}, comes from the star's on-sky motion
-altering our viewing angle, and leads to an observed apsidal
-precession.  The corresponding change is $\dot{P} \sim (P\mu)^2/2\pi
-\sim 10^{-21}$.
-
-Many of these phenomena are likely occurring simultaneously.  From the
-preceding discussion, we found that radial acceleration,
-proper-motion, and the Applegate mechanism are unlikely to cause the
-timing deviations.  That leaves tidal decay and apsidal precession as
-the possible causes.
+particularly given the fact that WASP-4 shows short-term trends with
+%$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
+amplitudes 100 times larger, over tens of days
+\citep{husnoo_observational_2012}.  Given the known activity of the
+system~\citep{sanchis-ojeda_starspots_2011}, these short-timescale
+variations might conspire to produce spurious measurements over long
+baselines.
+
+There are two other small effects worth noting.  The first is the
+\citet{shklovskii_possible_1970} effect, in which the star's proper
+motion leads to a changing radial velocity.  This would be apparent in
+the Doppler measurements described above.  The effect contributes a
+change of order $\dot{P} \sim P\mu^2 d/ c \sim 6\times10^{-13}$.  The
+second effect, described by \citet{rafikov_stellar_2009}, comes from
+the star's on-sky motion altering our viewing angle, and leads to an
+observed apsidal precession.  The corresponding period change is
+$\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$.
+
+Many of the phenomena discussed could presumably occur simultaneously.
+We found though that radial acceleration, proper-motion, and the
+Applegate mechanism are unlikely to explain the majority of the timing
+deviations.  Tidal decay seems implausible, given the high level of
+dissipation required.  This leaves apsidal precession as the most
+plausible explanation, but more data are needed for a definitive
+interpretation.
 
@@
 	\caption{
-		{\bf Future evolution possibilies for WASP-4b.}
+		{\bf Pecise long-term monitoring past 2025 will likely be needed to
+	    understand WASP-4b.}
 		Dots are as in Figure~\ref{fig:times}.
-		Lines are 100 random samples from the posteriors of the apsidal
+		Lines are 100 random draws from the posteriors of the apsidal
 		precession model (orange), and the orbital decay model (blue).    
-		In an extended mission, TESS might re-observe WASP-4b in the early
-		2020s. While useful to confirm that the period is
-		in fact changing, precise long-term monitoring past 2025 will
-		be needed to resolve the two models.
 		\label{fig:future}
@@
 measurements could also be useful.  In order to rule between orbital
-decay and apsidal precession, occultation measurements in both
-the near term and also in the mid-2020s will be needed.  Additionally, confirmation
-and extension of the linear trend suggested by
+decay and apsidal precession, occultation measurements in both the
+near term and also in the mid-2020s will be needed.  Additionally,
+confirmation and extension of the linear RV trend suggested by
 \citet{knutson_friends_2014} would help in understanding whether an
@@
 eccentricity.  These observations should be carried out because they
-will lead to new knowledge; either a measure of the efficiency of tidal
-dissipation within the star, or the eccentricity and Love number of the
-planet.  Either outcome for a single system would be rather
+will lead to new knowledge; either a measure of the stellar tidal
+dissipation efficiency, or the eccentricity and Love number of the
+planet.  Either outcome for a single system would be somewhat
 interesting.  In the coming decade, wide-field photometric surveys
 (TESS, HATPI, NGTS, PLATO) might even enable such measurements at a
-population level, which would be a good deal more interesting.
-
-%Detailed modeling of the star's interior
-% structure and and tidal evolution is warranted, as are continued
-% observations of transits and occultations.
-%
+population level, which would be extremely interesting.
 
 \acknowledgements{
-L.G.B.\ happily acknowledges helpful discussions with A. Bailey, F.
-Dai and V. Van Eylen, and is grateful to the people who have
-contributed to TESS.
+L.G.B.\ gladly acknowledges helpful discussions with
+A.~Bailey, F.~Dai and V.~Van Eylen, and is grateful to the
+people who have turned TESS from an idea into reality.
 %
@@
 %
-Funding for the TESS mission is provided by NASA’s Science Mission
+Funding for the TESS mission is provided by NASA's Science Mission
 directorate.
@@
 
-Any systematic offset between the reported TESS
-times and the true barycentric reference would cast doubt on the
-results of this study.  Such an offset would have historic precedent:
-Kepler had a systematic error in its timestamps that was corrected
-only in Q14~\citep[][Section 3.4]{kepler_DR19_2013}.
+Any systematic offset between the TESS times and the true barycentric
+reference would cast doubt on the results of this study.  There is
+historic precedent: Kepler had a systematic error in its timestamps
+that was corrected only in Q14~\citep[][Section
+3.4]{kepler_DR19_2013}.
 
@@
 the barycentric correction reported by SPOC in the TESS lightcurve
-file headers.  In \S~\ref{sec:hj_verification}, we measure the transit times
-of other hot Jupiters, and use them
-to rule out a global TESS time offset larger than about a minute.
-The latter test confirms that WASP-4b is an outlier.
+file headers.  In \S~\ref{sec:hj_verification}, we measure the transit
+times of other hot Jupiters, and use them to rule out a global TESS
+time offset larger than about a minute.  The latter test confirms that
+WASP-4b is an outlier.
 
@@
 three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
-TDB}$, and ${\rm BJD}_{\rm TDB}$.
-First, we verified for a few select lightcurve files that ${\rm
-JD}_{\rm UTC}$ lagged behind ${\rm JD}_{\rm TDB}$ by the expected
-$32.184 + N\,{\rm seconds}$, where $N$ is the number of leap-seconds
-since 1961. For the relevant observation time, $N=37$, and the offset
-was as expected.
-Then, using the ${\rm JD}_{\rm UTC}$ timestamp, we recalculated the
-barycentric correction computed by SPOC, using the
-\citealt{eastman_achieving_2010} calculator.  Due to the first
-author's ignorance of the spacecraft's position, we performed this
-calculation assuming that the observer was located at the Earth's
+TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
+lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
+JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
+is the number of leap-seconds since 1961. For the relevant observation
+time, $N=37$, and the offset was as expected.  Then, using the ${\rm
+JD}_{\rm UTC}$ timestamp, we recalculated the barycentric correction
+computed by SPOC, using the \citealt{eastman_achieving_2010}
+calculator.  Due to the first author's ignorance of the spacecraft
+position, we assumed that the observer was located at the Earth's
 geocenter, and used the correct direction for each star.  This gave us
@@
 within 1.7 seconds.  This offset is comparable to the light-travel
-delay time expected for TESS on its orbit from perigee of $\approx
+delay expected for TESS on its orbit from perigee of $\approx
 20R_\oplus$ to apogee of $\approx 60R_\oplus$.
@@
 This calculation bounds any error in the SPOC barycentric julian date
-correction to be less than 1.7 seconds.  Since this is smaller than
-the effect of interest, we treat the BJD correction as ``verified'',
-and proceed to a subsequent test.
+correction to be less than 1.7 seconds.  This is smaller than the
+effect of interest, so we treat the SPOC BJD correction as
+``verified'', and proceed to a subsequent test.
 
@@
 If the observed timing delay in WASP-4b were caused by a systematic
-timing system offset between the TESS ${\rm BJD}_{\rm TDB}$ times and
-the ${\rm BJD}_{\rm TDB}$ reference, we would expect that it might
-apply to other hot Jupiters as well.  This test can rule out an
-important class of clock error~--~a systematic global offset.
-
-To rule out this possibility, we repeat the timing analysis of the
-main paper, for other hot Jupiters with long preceding observing
-baselines.  We first checked which hot Jupiters were observed over the
-first two TESS sectors using a combination of
+offset between the TESS times and the ${\rm
+BJD}_{\rm TDB}$ reference, we might expect that it would apply to
+other hot Jupiters as well.  This test can rule out an important class
+of clock error~--~a systematic global offset.
+
+To explore this possibility, we repeat the timing analysis of the main
+paper, for other hot Jupiters with long preceding observing baselines.
+We first checked which hot Jupiters were observed over the first two
+TESS sectors using a combination of
 \texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
@@
 of a single transit, that the midpoint be fit as a free parameter, and
-that the time system be clearly documented.  Our final sample of
-hot Jupiters included WASPs-4b, -5b, -6b, -18b, and -46b.  The
-collected and measured times are given in Tables~4, 5, 6, and 7
-for each.
+that the time system be clearly documented.  Our final hot Jupiter
+sample included WASPs-4b, -5b, -6b, -18b, and -46b.  The collected and
+measured times are given in Tables~5, 6, 7, and 8 for each.
 
@@
 predicted transit time during the TESS observations.  This gave 9, 94,
-18, 42, and 60 seconds for WASPs-4b, -5b, -6b, -18b, and -46b.  If a
-substantial portion of the observed timing deviation in WASP-4b (about
-77 seconds) were from a systematic offset in the time systems, this
-offset would be present in the other hot Jupiter transit times as
-well.  We show in Figure~\ref{fig:hjs} that WASP-4b is the only hot
-Jupiter that transited significantly earlier than expected.
+18, 42, and 60 seconds for WASPs-4b, -5b, -6b, -18b, and -46b.  We
+show in Figure~\ref{fig:hjs} that WASP-4b is the only hot Jupiter that
+transited significantly earlier than expected.
 
@@
 For each planet, we then ask: what fraction of the posterior for
-$t_{\rm offset}$ is consistent with an offset worse than
-$77$ seconds?  For WASP-4b, the answer is unsurpringly about half.
-For WASP-6b, the most constraining object, about 1 sample in 2 million
-is consistent with such a timing offset ($4.9\sigma$).  For WASP-18b,
-1 in 63 samples would be consistent with this timing offset
-($2.1\sigma$), and in WASP-46b, the limit is 1 in 38 samples
-($1.9\sigma$).  For WASP-5b, the predicted time is too imprecise to
-rule out timing offsets at the necessary amplitude.  Multiplying the
-three independent probabilities for WASPs-5b, 6b, and -18b, we rule
-out $t_{\rm offset} < -77\ {\rm seconds}$ at $6.3\sigma$, or about
-about 1 part in 5 billion.
+$t_{\rm offset}$ is consistent with an offset worse than $77$ seconds?
+For WASP-4b, the answer is unsurpringly about half.  For WASP-6b, the
+most constraining object, about 1 sample in 2 million is consistent
+with such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 63
+samples would be consistent with this timing offset ($2.1\sigma$), and
+in WASP-46b, the limit is 1 in 38 samples ($1.9\sigma$).  For WASP-5b,
+the predicted time is too imprecise to rule out timing offsets at the
+necessary amplitude.  Multiplying the three independent probabilities
+for WASPs-5b, 6b, and -18b, we rule out $t_{\rm offset} < -77\ {\rm
+seconds}$ at $6.3\sigma$, or about about 1 part in 5 billion.
 

### 2018-12-07 src/planet_star_parameters.py
add src/planet_star_parameters.py
used to calculate things from Keivan's numbers

@@ -1,0 +1,153 @@
+# -*- coding: utf-8 -*-
+"""
+estimate the stellar density, and mass.
+"""
+from __future__ import division, print_function
+import numpy as np
+
+from astropy import units as u, constants as const
+
+# use parameters from Table 4 of Hoyer+2013
+i = 88.52*u.deg
+a_by_Rs = 5.463
+a_by_Rs_perr = 0.025
+a_by_Rs_merr = 0.020
+P = 1.33823204*u.day
+
+# eq 30 of Winn (2010).
+# note k^3 = 3e-3, and the density of jupiter is about equal to the density of
+# the sun, so neglect the k^3*rho_planet term.
+
+rho_star = 3*np.pi/(const.G*P**2) * a_by_Rs**3
+print(rho_star.cgs)
+
+rho_star_upper = 3*np.pi/(const.G*P**2) * (a_by_Rs+a_by_Rs_perr)**3
+print(rho_star_upper.cgs)
+
+rho_star_lower = 3*np.pi/(const.G*P**2) * (a_by_Rs-a_by_Rs_merr)**3
+print(rho_star_lower.cgs)
+
+print(
+    'rhostar = {:.3f} +({:.3f}) -({:.3f}) g/cm^3'.format(
+        rho_star.cgs.value,
+        (rho_star_upper-rho_star).cgs.value,
+        (rho_star-rho_star_lower).cgs.value
+    )
+)
+
+# Keivan value
+Rstar = 0.893*u.Rsun
+Rstar_err = 0.024*u.Rsun
+
+Mstar = (4*np.pi)/3*rho_star*Rstar**3
+Mstar_lower = (4*np.pi)/3*rho_star_lower*(Rstar-Rstar_err)**3
+Mstar_upper = (4*np.pi)/3*rho_star_upper*(Rstar+Rstar_err)**3
+
+print(
+    'Mstar = {:.3f} +({:.3f}) -({:.3f}) Msun'.format(
+        Mstar.to(u.Msun).value,
+        (Mstar_upper-Mstar).to(u.Msun).value,
+        (Mstar-Mstar_lower).to(u.Msun).value,
+    )
+)
+
+##########################################
+# now get R_p
+
[…]

### 2018-12-08 paper/ms.tex
update broadband mag paragraph

@@
   (F) From functions of stellar, transit or RV parameters as discussed in text.
-  (G) Gaia DR2.
+  (G) {\it Gaia} DR2.
   (H) \citet{stassun_TIC_2018}.
@@
 
-\citet{stassun_accurate_2017} found the bolometric flux from
-broad-band photometry.  Combining the implied angular radius with the
-{\it Gaia} DR2 parallax, and adjusting for the parallax offset
-\citep{stassun_evidence_2018}, we obtain the stellar radius.  For the
-stellar mass, we started with the $a/R_\star$ value found by
+We compute the bolometric flux using broadband magnitudes from
+available all-sky catalogs, as follows: $G$ from {\it Gaia\/} DR2,
+$B_T V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
+2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
+wavelength range 0.4--22~$\mu$m.  Combining the implied angular radius
+with the {\it Gaia} DR2 parallax, and adjusting for the parallax
+offset \citep{stassun_evidence_2018}, we obtain the stellar radius.
+For the stellar mass, we started with the $a/R_\star$ value found by
 \citet{hoyer_tramos_2013} from 26 ground-based transits.  The implied

### 2018-12-10 paper/ms.tex
Update ms.tex

@@
 
-\title{The Early Arrival of WASP-4\lowercase{b}}
+\title{WASP-4\lowercase{b} Arrived Early for the TESS Mission}
 
@@
   %TODO: final numbers
-  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
-  with previous studies, we show that the transit times of WASP-4b are
-  incompatible with a constant orbital period.  In particular, the
-  transits seem to have arrived $77.8 \pm 10.8$ seconds early, and the
-  period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
-  milliseconds per year.   From TESS observations of WASPs-6b, -18b,
-  and -46b, we show that a systematic offset between the TESS times
-  and the barycentric reference sufficient to explain WASP-4b is ruled
-  out at 6.3$\sigma$.  The timing variations seem to be astrophysical,
+  By combining data from the Transiting Exoplanet Survey Satellite (TESS)
+  with data from previous studies, we show that the transit times of WASP-4b are
+  incompatible with a constant orbital period.  The
+  transits observed by TESS occurred $77.8 \pm 10.8$~seconds earlier than one would expect, based on the previous data.
+  The orbital period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
+  milliseconds per year.   A simple clock offset is unlikely, because TESS observations of other
+  hot Jupiter (WASP-6b, 18b, and 46b) are compatible with a constant period and rule out
+  a 77.8-second offset at the 6.3$\sigma$ level.
+  The timing variations seem to be astrophysical,
   and could be caused by apsidal precession, or perhaps by tidal

### 2018-12-11 paper/ms.tex
update figures per JNW comments

@@
 The second process that long-term timing studies should reveal is
-rotation of the orbital ellipse within the orbital plane (apsidal
-precession).  This effect has been studied extensively in eclipsing
-binaries \citep[{\it e.g.},][]{
+apsidal precession, {\it i.e.}, rotation of the orbital ellipse within
+the orbital plane. Apsidal precession has been observed extensively in
+eclipsing binaries \citep[{\it e.g.},][]{
   schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015},
-and has been explored theoretically for transiting planets
-\citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
-However, apsidal precession has yet to be directly observed in hot
-Jupiters, and the orbits would need non-zero eccentricity for it to be
-detectable.  Though the eccentricities of hot Jupiters are expected to
-be small, there are many mechanisms that could pump them away from
-circularity (see \S~\ref{sec:apsidal_precession}, and
-\citealt{bailey_understanding_2019}).  For hot Jupiters, the apsidal
+and has been detected in a handful of exoplanet systems, including
+circumbinary planets (Kepler-413 \citealt{kostov_kepler-413b_2014};
+Kepler-453 \citealt{welsh_kepler_2015}, planets orbiting rapidly
+rotating stars (Kepler-13
+\citealt{szabo_spin-orbit_2012,szabo_mapping_2014,masuda_spin-orbit_2015};
+PTFO 8-8695 \citealt{barnes_measurement_2013}), and systems with
+multiple interacting giant planets (Kepler-108
+\citealt{mills_kepler-108_2017}; Kepler-693
+\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
+observed in hot Jupiter systems, perhaps in part because their orbits
+are nearly circular.  However, there are many mechanisms that
+could pump hot Jupiter orbits away from circularity (see
+\S~\ref{sec:apsidal_precession}, and
+\citealt{bailey_understanding_2019}), which would then lead to
+observable timing variations.  For hot Jupiters, the apsidal
 precession is usually dominated by the quadrupolar distortion of the
 planet by the star's tidal force \citep{ragozzine_probing_2009}.  The
-observed precession rate in a hot Jupiter system can thus be used to
+observed precession rate in a hot Jupiter system could thus be used to
 compute the planet's Love number, giving a third parameter along with
@@
 \citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  WASP-4b is a plausible
-candidate for orbital decay, in part because it has been frequently
-observed over the past decade.  The host is a G7V dwarf, and is on the
+candidate for orbital decay, and has been
+observed often over the past decade.  The host is a G7V dwarf, and is on the
 main sequence (\S~\ref{sec:system_parameters}).  The orbit is
@@
     \end{center}
-    \vspace{-0.5cm}
+    \vspace{-0.6cm}
     \caption{
@@
         {\bf TESS saw WASP-4b transit earlier than expected.}
-        Points are differences between observed transit times and the
-        constant-period ephemeris we fit to literature times.  Black
-        points are the more precise half of literature times; gray
-        points are the less precise half.  Blue curves represent the
+        Points are the
+        differences between observed transit times and the
+        pre-TESS ephemeris, plotted against
+        transit number.
+        Blue curves are the
         $\pm 1\sigma$ credible interval around the constant-period
-        ephemeris.  The binned TESS observation (red point) arrived
+        prediction.
+        {\it Top:} The full timing dataset spans 10 years.
+                Less precise points are more transparent.
+        {\it Bottom:}
+        The binned TESS observation (red point) arrived
         $77.8 \pm 10.8\ {\rm seconds}$ before it was expected.
@@
         {\bf Timing residuals and best-fit models for WASP-4b.}
-        Points are differences between observed transit times and a
-        linear ephemeris fit to all the times.  The constant period
-        model is a poor fit to the observed times (gray line).  Black
-        points are the more precise half of measured times; gray
-        points are the less precise half.  The blue curve is the
-        best-fit model with a constant period derivative.  The orange
-        curve is the best-fit model assuming apsidal precession causes
-        the observed timing deviations. The binned TESS point (star)
+        Points are observed transit times minus a
+        linear fit to all the times.
+        Less precise times are more transparent.
+        The best constant period
+        model is a poor fit to the observations (gray line).
+        The best-fit tidal decay model (blue) and apsidal precession
+        model (orange) can fit the data.
+        The binned TESS point (star)
         is for display purposes only.
@@
 ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
-specifying the time standard.  We correspondingly added $69.184$
-seconds of uncertainty to their reported errors, in quadrature.
+specifying the time standard.  We assumed the standard was UTC, but
+added $32.182 + N$ seconds of uncertainty to the reported errors,
+where $N=35$ is the appropriate number of leap-seconds
+\citep[see][]{eastman_achieving_2010}.
+Though not rigorous, this procedure inflates the error bar to a more
+realistic $\pm$94 seconds.
 %Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}
@@
 %parameter.
-\citet{zhou_secondary_2015} acquired occultation data from the
-Anglo-Australian Telescope, focusing on the eclipse depth, but also
-fitting for $e\cos\omega$.  We converted their $e\cos\omega$ result
-into a midtime, using their ephemeris and the standard formula
+Finally, \citet{zhou_secondary_2015} acquired occultation data from
+the Anglo-Australian Telescope, focusing on the eclipse depth, but
+also fitting for $e\cos\omega$.  We converted their $e\cos\omega$
+result into a midtime, using their ephemeris and the standard formula
 \citep[{\it e.g.},][]{winn_exoplanet_2010}
@@
 orbital decay time of WASP-4b compared to other transiting giant
-planets.  Though roughly 20 hot Jupiters might be expected to decay
-faster, some orbit stars with minimal convective zones ({\it e.g.},
-WASP-18), and most are not as well-observed as WASP-4.
+planets.  Though roughly 20 hot Jupiters have shorter predicted decay
+times, most are not as well-observed as WASP-4b. Moreover, some hot
+Jupiters, {\it e.g.}, WASP-18b, orbit fully radiative stars, which may
+dissipate tidal energy less efficiently than stars that are partially
+convective.
 
 For the sake of argument, let us assume that the timing variation is
-caused entirely by orbital decay.  If the period continues to decrease
-at a fixed rate, it will reach zero after
+caused entirely by orbital decay.  If the period decreases
+at a fixed rate, then it will reach zero after
 \begin{equation}
@@
 \end{equation}
-For comparison, the ``survival time'' found by \citet{patra_2017} for
+For comparison, the survival time found by \citet{patra_2017} for
 WASP-12b was 3.2\,Myr.
@@
 heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A
-larger $Q_\star$ implies less efficient dissipiation.  Though the
-fluid mechanics responsible for the dissipation should depend on the
-amplitude and frequency of the tidal perturbation
+larger $Q_\star$ implies less efficient tidal dissipiation.  Though
+the fluid mechanics responsible for the dissipation should depend on
+the amplitude and frequency of the tidal perturbation
 \citep[][Section~3.3]{ogilvie_tidal_2014}, we will treat $Q_\star'$ as
 a constant.  This should not be taken literally; $Q_\star'$ may in
-fact change sharply with varying tidal frequency
+fact change sharply as the tidal frequency changes
 \citep{penev_empirical_2018}.
@@
 \end{equation}
-Using the observed value of $\dot{P}$, the current modified quality
-factor of WASP-4 would be
+The modified quality factor of WASP-4 corresponding to the observed
+value of $\dot{P}$ would then be
 \begin{equation}
@@
 \citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
-geller_direct_2013,milliman_wiyn_2014}.  However the dissipation rates
-may differ in hot Jupiter systems, since the tidal forcing frequencies
-and amplitudes are different.  Observationally-driven studies of hot
-Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, often using
+geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
+differ in hot Jupiter systems, since the tidal forcing frequencies and
+amplitudes are different.  Observationally-driven studies of hot
+Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
 different models
@@
 
-A theoretically-motivated model for tidal dissipation was proposed by
-\citet{essick_orbital_2016}, who considered the nonlinear response of
-the star to a coupled network of gravity modes excited at the base of
-the convective zone, and braking near the stellar core.  In their
+\citet{essick_orbital_2016} proposed a theoretically-motivated tidal
+dissipation model , in which a coupled network of gravity modes would
+be excited at the base of the stellar convective zone, and would brake
+near the stellar core to provide the needed dissipation.  In their
 approach, the stellar quality factors in hot Jupiter systems vary from
-$Q_\star' \approx 10^5 - 10^6$.  Using their Equation 26, the
-prediction for WASP-4 is $Q_\star' = 7\times10^5$.  This is an order
-of magnitude larger than implied by the observed period change.
+$Q_\star' \approx 10^5 - 10^6$.  From their Equation 26, the predicted
+quality factor for WASP-4 is $Q_\star' = 7\times10^5$.  This is an
+order of magnitude larger than implied by the observed period change.
 
@@
 the evolutionary state of the star.  The bottom panel of
-Figure~\ref{fig:context} shows WASP-4 in the context of other giant
-planet hosts from the \citet{bonomo_gaps_2017} sample.  On the y-axis
-is $G=g-\mu$, for $g$ the apparent Gaia-band magnitude, and $\mu$ the
+Figure~\ref{fig:context} shows WASP-4 in the context of giant planet
+hosts from the \citet{bonomo_gaps_2017} sample.  On the y-axis is
+$G=g-\mu$, for $g$ the apparent Gaia-band magnitude, and $\mu$ the
 distance modulus reported by \citet{gaia_collaboration_gaia_2018}.
@@
 
-In short, if the observed period change is caused entirely by tidal
+In summary, if the observed period change is caused entirely by tidal
 decay, it would imply a tidal dissipation efficiency stronger than
@@
 special time, shortly after the planet's inward migration, or when the
-planet is near resonance with a stellar oscillation mode
-\citep{ogilvie_tidal_2014,essick_orbital_2016}.  Tidal dissipation
-rates might also be increased if the star is turning off the main
-sequence.  All of these explanations are strained; it is unclear that
-tidal decay could explain the observed timing variations without
-fine-tuning.
+planet is near resonance with a stellar oscillation mode.  Tidal
+dissipation rates might also be increased if the star is turning off
+the main sequence.  A final possibility is that an exterior planet
+could be trapping WASP-4b's spin vector in a high-obliquity state,
+causing planetary obliquity tides as suggested for WASP-12b
+\citep{millholland_obliquity_2018}.  For now, all of these
+explanations seem strained; it is unclear that tidal decay can explain
+the observed timing variations by any mechanism that we have
+considered.
 
@@
 	\caption{
-		{\bf Pecise long-term monitoring past 2025 will likely be needed to
-	    understand WASP-4b.}
+		{\bf Further observations will be needed to
+	    confirm and understand the timing variations of WASP-4b.}
 		Dots are as in Figure~\ref{fig:times}.

### 2018-12-11 paper/ms.tex
update ms.pdf

@@
   %TODO: final numbers
-  By combining data from the Transiting Exoplanet Survey Satellite (TESS)
-  with data from previous studies, we show that the transit times of WASP-4b are
-  incompatible with a constant orbital period.  The
-  transits observed by TESS occurred $77.8 \pm 10.8$~seconds earlier than one would expect, based on the previous data.
-  The orbital period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
-  milliseconds per year.   A simple clock offset is unlikely, because TESS observations of other
-  hot Jupiter (WASP-6b, 18b, and 46b) are compatible with a constant period and rule out
-  a 77.8-second offset at the 6.3$\sigma$ level.
-  The timing variations seem to be astrophysical,
-  and could be caused by apsidal precession, or perhaps by tidal
-  decay.  The latter seems implausible on theoretical grounds, as it
-  would require a stellar quality factor $Q_\star' \approx
-  3\times10^4$.  Apsidal precession would imply an eccentricity
-  $e\approx1.6\times 10^{-3}$, and the possibility of measuring the
-  planet's Love number.  A Doppler shift from WASP-4's acceleration
-  towards us could account for at most one quarter of the observed
-  period change (at $2\sigma$).  Further observations will help
-  confirm the reality of the timing variation, and eventually
+  By combining data from the Transiting Exoplanet Survey Satellite
+  (TESS) with data from previous studies, we show that the transit
+  times of WASP-4b are incompatible with a constant orbital period.
+  The transits observed by TESS occurred $77.8 \pm 10.8$~seconds
+  earlier than one would expect, based on the previous data.  The
+  orbital period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
+  milliseconds per year.   A simple clock offset is unlikely, because
+  TESS observations of other hot Jupiters (WASP-6b, 18b, and 46b) are
+  compatible with a constant period and rule out a 77.8-second offset
+  at the 6.3$\sigma$ level.  The timing variations seem to be
+  astrophysical, and could be caused by apsidal precession, or perhaps
+  by tidal decay.  The latter seems implausible on theoretical
+  grounds, as it would require a stellar quality factor $Q_\star'
+  \approx 3\times10^4$.  Apsidal precession would imply an
+  eccentricity $e\approx1.6\times 10^{-3}$, and the possibility of
+  measuring the planet's Love number.  A Doppler shift from WASP-4's
+  acceleration towards us could account for at most one quarter of the
+  observed period change (at $2\sigma$).  Further observations will
+  help confirm the reality of the timing variation, and eventually
   determine its cause.
@@
 \citep{kovacs_box-fitting_2002}.  Using these parameters, we isolated
-each transit to within $\pm$10 transit durations of its midtime.  In
+each transit to within $\pm$5 transit durations of its midtime.  In
 each transit window, we then simultaneously fit for a model transit,
@@
 We fixed the remaining transit parameters as follows.  We set the
-eccentricity to zero, and the longitude of periastron to $\pi/2$.  We
+eccentricity to zero. We
 assumed a quadratic limb-darkening law, with coefficients interpolated
@@
 After completing an initial fit, we set the photometric error bars to
-be equal to the standard deviation of the in-transit points of the
+be equal to the standard deviation of the out-of-transit points of the
 residual lightcurve.  We then reperformed the fitting procedure, using
@@
 Binning the residuals to 1-hour windows and taking the median absolute
-deviation, the lightcurves have a MAD of $594\,{\rm ppm}\,{\rm
+deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
 hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
@@
 in planet detection simulations.  This zodiacal noise prediction seems
-to have been an overestimate, given the observed MAD of WASP-4.  A
+to have been an overestimate, given the observed RMS of WASP-4.  A
 more detailed assessment of TESS's photometric performance is beyond
@@
 We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point.  Though their latest two epochs fell slightly below
-the prediction of a linear ephemeris (Panel C of their Figure~6), they
-did not find a need to fit a quadratic function to the O-C values.  We
-verified that all times were on the BJD$_{\rm TDB}$ time standard
-using the \citet{eastman_achieving_2010} calculator.  We set the
-zero-point epoch to be as close as possible to the average of the
-midtimes, weighted as the inverse midtime variance. This minimizes the
+starting point\footnote{We also tried using times directly from the
+original authors; the results did not change, so we opted for the
+\citet{hoyer_tramos_2013} times as the more homogeneous option.}.
+Though their latest two epochs fell slightly below the prediction of a
+linear ephemeris (Panel C of their Figure~6), they did not find a need
+to fit a quadratic function to the O-C values.  We verified that all
+times were on the BJD$_{\rm TDB}$ time standard using the
+\citet{eastman_achieving_2010} calculator.  We set the zero-point
+epoch to be as close as possible to the average of the midtimes,
+weighted as the inverse midtime variance. This minimizes the
 covariance between the transit epoch and the period

### 2018-12-11 src/ephemerides_utilities.py
typechange ephem utilities

@@ -1,1 +1,36 @@
-/Users/luke/Dropbox/proj/tessorbitaldecay/src/ephemerides_utilities.py
+'''
+Functions here are named:
+    get_ROUGH_epochs_given_midtime_and_period
+    get_half_epochs_given_occultation_times
+'''
+import numpy as np
+
+from astrobase.timeutils import get_epochs_given_midtimes_and_period
+
+def get_half_epochs_given_occultation_times(tsec, init_period, t0):
+    '''
+    assume zero eccentricity. then
+
+        t_sec = t0 + period*(epoch + 0.5)
+
+    where "epoch" is the integer epoch counter for the transits.
+
+    this function returns the half-epochs. t0 should come from the more
+    constraining transit times (rather than occultation times).
+    '''
+
+    epoch_plushalf = (tsec - t0)/init_period
+
+    epoch_plushalf_rounded = np.round(epoch_plushalf,1)
+
+    # ensure the output array only has half-epoch secondary eclipse times.
+    out_epochs = []
+    for ix, er in enumerate(epoch_plushalf_rounded):
+        if not repr(er)[-1] == str(5) and not repr(er)[-1] == str(0):
+            out_epochs.append(np.nan)
+        else:
+            out_epochs.append(er)
+
+    out_epochs = np.array(out_epochs)
+
+    return out_epochs

### 2018-12-12 paper/ms.tex
add updated method section

@@
 Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
-estimate the transit epoch, period, and duration
-\citep{kovacs_box-fitting_2002}.  Using these parameters, we isolated
-each transit to within $\pm$5 transit durations of its midtime.  In
-each transit window, we then simultaneously fit for a model transit,
-plus a local linear trend.  Our transit model uses the analytic
-formulae calculated by \citet{mandel_analytic_2002} and implemented by
-\citet{kreidberg_batman_2015}.
-
-We allowed four free parameters per transit: the time of mid-transit
-$t_{\rm tra}$, the planet-to-star radius ratio $R_{\rm p}/R_\star$,
-and the slope and intercept of the line.
-% Our priors were wide uniform distributions centered on the BLS
-% estimates for $t_{\rm tra}$ and $R_{\rm p}/R_\star$, and on a slope
-% of zero and an intercept of unity for the line.
-We fixed the remaining transit parameters as follows.  We set the
-eccentricity to zero. We
-assumed a quadratic limb-darkening law, with coefficients interpolated
-from the \citet{claret_limb_2017} tables.  For the period,
-$a/R_\star$, and inclination, we adopted the values
-from~\citet{petrucci_no_2013}.  These are in reasonable agreement with
-the parameters reported by \citet{gillon_improved_2009},
-\citet{southworth_high-precision_2009}, and
-\citet{huitson_gemini_2017}.
+determine the orbital period, and to estimate the transit duration and
+a reference epoch \citep{kovacs_box-fitting_2002}.  From these
+parameters, we isolated each transit to within $\pm$4 transit
+durations of its midtime.  We then fit the phase-folded transit, using
+the analytic formulae calculated by \citet{mandel_analytic_2002} and
+implemented by \citet{kreidberg_batman_2015}.  Specifically, we fit
+for a reference epoch, the planet to star radius ratio $R_{\rm
+p}/R_\star$, the semi-major axis to stellar radius ratio $a/R_\star$,
+the inclination $i$, two quadratic limb-darkening coefficients
+$(u_{\rm linear}, u_{\rm quad})$, and the orbital period $P$.  We set
+the eccentricity to zero.
 
@@
 \citet{foreman-mackey_emcee_2013}.
+Table~1 gives the results, which are in reasonable agreement with the
+parameters reported by \citet{gillon_improved_2009},
+\citet{southworth_high-precision_2009}, and
+\citet{huitson_gemini_2017}.
+%TODO: is this true? Or do the periods actually differ?
+
+To measure mid-transit times, we then focused on each individual
+transit window.  We fit for a model transit, plus a local linear
+trend.  We allowed four free parameters per transit: the time of
+mid-transit $t_{\rm tra}$, the planet-to-star radius ratio $R_{\rm
+p}/R_\star$, and the slope and intercept of the line.  We fixed the
+remaining parameters~--~$a/R_\star$, $i$, $P$, and the limb-darkening
+coefficients~--~to the best-fit values from the global fit.
+
 % During this process, we used the \texttt{corner} software to check
 % that our posteriors were being well-sampled~\citep{corner_2016}.
-After completing an initial fit, we set the photometric error bars to
-be equal to the standard deviation of the out-of-transit points of the
-residual lightcurve.  We then reperformed the fitting procedure, using
-the empirically determined photometric errors.
+After completing an initial fit for each transit, we set the
+photometric error bars to be equal to the standard deviation of the
+out-of-transit points of the residual lightcurve.  We then reperformed
+the fitting procedure, using the empirically determined photometric
+errors.
 
@@
 Figure~\ref{fig:lightcurves} shows the light curves, best-fit models,
-and residuals.  Table~1 reports the mid-transit times and their
-uncertainties.  Taking the mean and standard deviation of the measured
-planet to star radius ratio, we find $R_{\rm
-p}/R_\star=0.1538\pm0.0013$, in agreement with previous determinations
-\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,southworth_high-precision_2009}.
+and residuals.  Table~2 reports the mid-transit times and their
+uncertainties.
 Binning the residuals to 1-hour windows and taking the median absolute
@@
 \tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
-\tablenum{2}
+\tablenum{1}
 
 \tablehead{
-\colhead{Parameter} & \colhead{Value} & \colhead{Error} & \colhead{Comment}
+\colhead{Parameter} & \colhead{Value} & \colhead{Uncertainty} & \colhead{Comment}
 }
@@
 \startdata
+%FIXME with new method
+%TODO numbers with updated method
 {\it Transit/RV parameters:} & & & \\
   $R_{\rm p}/R_\star$                        & $0.1538$               & $\pm 0.0013$                & A \\
-  $i$~[deg]                                  & $88.52$                & $+0.39$, $-0.26$            & B \\
-  $a/R_\star$                                & $5.463$                & $+0.025$, $-0.020$          & B \\
-  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & C \\
+  $i$~[deg]                                  & $88.52$                & $+0.39$, $-0.26$            & A \\
+  $a/R_\star$                                & $5.463$                & $+0.025$, $-0.020$          & A \\
+  $u_{\rm linear}$                           & $0.XXX$                & $+0.XXX$, $-0.XXX$          & A \\
+  $u_{\rm quad}$                             & $0.XXX$                & $+0.XXX$, $-0.XXX$          & A \\
+  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
 {\it Stellar parameters:} & & & \\
-  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & D \\
-  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & D \\
-  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & D \\
-  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & E \\
-  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & G \\
-  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & F \\
-  $\rho_\star$~[g~cm$^{-3}$]                 & $1.722$                & $+0.024$, $-0.019$          & A \\
-  $M_\star$~[M$_{\odot}$]                    & $0.870$                & $+0.085$, $-0.077$          & F \\
-  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & H \\
+  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
+  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
+  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
+  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
+  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
+  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
+  $\rho_\star$~[g~cm$^{-3}$]                 & $1.722$                & $+0.024$, $-0.019$          & E \\
+  $M_\star$~[M$_{\odot}$]                    & $0.870$                & $+0.085$, $-0.077$          & E \\
+  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
 {\it Planetary parameters:} & & & \\
-  $a$~[AU]                                   & $0.0227$               & $\pm 0.0007$                & F \\
-  $M_p$~[M$_{\rm Jup}$]                      & $1.192$                & $+0.091$, $-0.086$          & F \\
-  $R_p$~[R$_{\rm Jup}$]                      & $1.337$                & $\pm 0.048$                 & F \\
+  $a$~[AU]                                   & $0.0227$               & $\pm 0.0007$                & E \\
+  $M_p$~[M$_{\rm Jup}$]                      & $1.192$                & $+0.091$, $-0.086$          & E \\
+  $R_p$~[R$_{\rm Jup}$]                      & $1.337$                & $\pm 0.048$                 & E \\
 \enddata
@@
 \tablecomments{
-  (A) From TESS lightcurves (\S~\ref{sec:measurement}).
-  (B) \citet{hoyer_tramos_2013}.
-  (C) \citet{triaud_spin-orbit_2010}.
-  (D) From HARPS spectra \citep{doyle_accurate_2013}.
-  (E) \citet{stassun_accurate_2017}.
-  (F) From functions of stellar, transit or RV parameters as discussed in text.
-  (G) {\it Gaia} DR2.
-  (H) \citet{stassun_TIC_2018}.
+  (A) From TESS lightcurves (\S~\ref{sec:measurement}). Orbital periods from timing data are in Table~4.
+  (B) \citep{triaud_spin-orbit_2010}.
+  (C) From HARPS spectra \citep{doyle_accurate_2013}.
+  (D) \citet{stassun_accurate_2017}.
+  (E) From functions of stellar, transit or RV parameters as discussed in text.
+  (F) {\it Gaia} DR2.
+  (G) \citet{stassun_TIC_2018}.
 }

### 2018-12-12 src/plot_hjs.py
update fig6

@@ -216,7 +216,7 @@ def plot_hjs():
 
     fig.legend((l1,l2,l3),
                ('KDE from TESS', 'observed TESS',
-                'literature prediction'),
+                'pre-TESS prediction'),
                loc='upper right',
                bbox_to_anchor=(0.98, 0.42),
                fontsize='xx-small')

### 2018-12-12 paper/ms.tex
syntax purging

@@
 %
+\author[0000-0002-5720-4827]{M. Hill}
+\affiliation{Department of Earth Sciences, University of California,
+Riverside, CA 92521, USA}
+%
 \author[0000-0002-7084-0529]{S. Kane}
@@
 % 3 TSO COAUTHORS
+% 3 SPOC COAUTHORS
+% 3 POC COAUTHORS
 %
@@
 %
+\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
+%
+\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
+\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
+Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
+%
+\author{M. Fausnaugh} % faus@mit.edu WAITING
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+%
+\author[0000-0002-5322-2315]{A. Glidden} % aglidden@mit.edu WAITING
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+%
+\author{N. Guerrero} % nmg@mit.edu WAITING
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+%
 \author[0000-0001-8812-0565]{J. Rodriguez} % rodriguez.jr.joey@gmail.com WAITING
@@
 %
-\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
-\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
-Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
-% 3 SPOC COAUTHORS
-%
 \author[0000-0002-6778-7552]{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
@@
 %
-\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-%
-%\\ \textcolor{red}{3 POC coauthors:}
-\author{POC1}
-\author{POC2}
-\author{POC3}
 
@@
   acceleration towards us could account for at most one quarter of the
-  observed period change (at $2\sigma$).  Further observations will
-  help confirm the reality of the timing variation, and eventually
-  determine its cause.
+  observed period change (at $2\sigma$).  Further observations are
+  needed to confirm and understand the timing variation.
 \end{abstract}
@@
 eclipsing binaries \citep[{\it e.g.},][]{
-  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015},
+  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}
 and has been detected in a handful of exoplanet systems, including
@@
 rotating stars (Kepler-13
-\citealt{szabo_spin-orbit_2012,szabo_mapping_2014,masuda_spin-orbit_2015};
+\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
 PTFO 8-8695 \citealt{barnes_measurement_2013}), and systems with
@@
 The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
-array every 2 minutes.   After being downlinked through the Deep Space
-Network, the data were processed through the Science Processing
-Operations Center (SPOC) pipeline at NASA
-Ames~\citep{jenkins_tess_2016}.  One important step in this processing
-was converting from spacecraft time to barycentric time.  SPOC then
-reports timestamps in ``BTJD'', where ${\rm BTJD} = {\rm BJD} -
-2457000$, BJD is the barycentric julian date, and all times are in the
-TDB reference (see \citealt{urban_explanatory_2012} for explanations
-of time systems).  The lightcurves were then vetted and released by
-the MIT TESS Science Office to the Mikulski Archive for Space
+array every 2 minutes.   The data were downlinked via the Deep Space
+Network, and the Payload Operations Center then converted the
+timestamps from the spacecraft clock to TESS Julian Day in the {\it
+Temps Dynamique Barycentrique} (TDB) reference.  The data were then
+processed through the Science Processing Operations Center (SPOC)
+pipeline at NASA Ames~\citep{jenkins_tess_2016}.  SPOC used the known
+spacecraft trajectory to compute the barycentric time corrections on a
+target by target basis, and reported timestamps in ``BTJD'', where
+${\rm BTJD} = {\rm BJD} - 2457000$, BJD is the barycentric julian
+date, and all times are in the TDB reference (see
+\citealt{urban_explanatory_2012} for explanations of time systems).
+The lightcurves were then vetted and released by the MIT TESS Science
+Office to the Mikulski Archive for Space
 Telescopes~\citep{ricker_tess_alerts_2018}.
 
-We begin our analysis with the Presearch Data Conditioning (PDC) light
+We began our analysis with the Presearch Data Conditioning (PDC) light
 curves from MAST.  The steps used to find an optimal aperture (for
@@
 
-We then processed the lightcurve as follows.  First, we removed all
-points with non-zero quality flags.  This removes data contaminated by
+We processed the lightcurve as follows.  First, we removed all
+points with non-zero quality flags.  This removed data contaminated by
 coarse spacecraft pointing, cosmic rays, and other annoyances
@@
 After ``cleaning'' the data, we normalized the flux measurements by
-dividing out the median flux.  We then converted the timestamps from
-BTJD to BJD by adding the appropriate 2457000 day offset
+dividing out the median flux.  We converted the timestamps from BTJD
+to BJD by adding the appropriate 2457000 day offset
 \citep{tess_data_product_description_2018}.  Many of these and
@@
         {\bf TESS saw WASP-4b transit earlier than expected.}
-        Points are the
-        differences between observed transit times and the
-        pre-TESS ephemeris, plotted against
-        transit number.
-        Blue curves are the
-        $\pm 1\sigma$ credible interval around the constant-period
-        prediction.
-        {\it Top:} The full timing dataset spans 10 years.
-                Less precise points are more transparent.
-        {\it Bottom:}
-        The binned TESS observation (red point) arrived
-        $77.8 \pm 10.8\ {\rm seconds}$ before it was expected.
+        Both plots show observed transit times minus the pre-TESS
+        ephemeris as a function of transit number.  Blue curves are
+        the $\pm 1\sigma$ credible interval around the constant-period
+        prediction.  {\it Top:} The full timing dataset spans 10
+        years.  Less precise points are more transparent.  {\it
+        Bottom:} The binned TESS observation (red point) arrived $77.8
+        \pm 10.8\ {\rm seconds}$ before it was expected.
         \label{fig:arrived_early}
@@
 Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
-determine the orbital period, and to estimate the transit duration and
-a reference epoch \citep{kovacs_box-fitting_2002}.  From these
-parameters, we isolated each transit to within $\pm$4 transit
-durations of its midtime.  We then fit the phase-folded transit, using
-the analytic formulae calculated by \citet{mandel_analytic_2002} and
-implemented by \citet{kreidberg_batman_2015}.  Specifically, we fit
-for a reference epoch, the planet to star radius ratio $R_{\rm
-p}/R_\star$, the semi-major axis to stellar radius ratio $a/R_\star$,
-the inclination $i$, two quadratic limb-darkening coefficients
-$(u_{\rm linear}, u_{\rm quad})$, and the orbital period $P$.  We set
-the eccentricity to zero.
+estimate the orbital period, transit duration and a reference epoch
+\citep{kovacs_box-fitting_2002}.  From these parameters, we isolated
+each transit to within $\pm$4 transit durations of its midtime.  We
+then fit the phase-folded transit, using the analytic formulae
+calculated by \citet{mandel_analytic_2002} and implemented by
+\citet{kreidberg_batman_2015}.  Specifically, we fit for a reference
+epoch, the planet to star radius ratio $R_{\rm p}/R_\star$, the
+semi-major axis to stellar radius ratio $a/R_\star$, the inclination
+$i$, two quadratic limb-darkening coefficients $(u_{\rm linear},
+u_{\rm quad})$, and the orbital period $P$.  We set the eccentricity
+to zero.
 
@@
 trend.  We allowed four free parameters per transit: the time of
-mid-transit $t_{\rm tra}$, the planet-to-star radius ratio $R_{\rm
-p}/R_\star$, and the slope and intercept of the line.  We fixed the
-remaining parameters~--~$a/R_\star$, $i$, $P$, and the limb-darkening
-coefficients~--~to the best-fit values from the global fit.
-
-% During this process, we used the \texttt{corner} software to check
-% that our posteriors were being well-sampled~\citep{corner_2016}.
+mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
+slope and intercept of the line.  We fixed the remaining
+parameters~---~$a/R_\star$, $i$, $P$, and the limb-darkening
+coefficients~---~to the best-fit values determined from the
+phase-folded TESS lightcurve.
+
 After completing an initial fit for each transit, we set the
-photometric error bars to be equal to the standard deviation of the
+photometric error bars to equal the standard deviation of the
 out-of-transit points of the residual lightcurve.  We then reperformed
@@
 
-Figure~\ref{fig:lightcurves} shows the light curves, best-fit models,
+% TODO: final numbers
+Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
 and residuals.  Table~2 reports the mid-transit times and their
@@
 to have been an overestimate, given the observed RMS of WASP-4.  A
-more detailed assessment of TESS's photometric performance is beyond
+broader assessment of TESS's photometric performance is beyond
 the scope of this work.
@@
 \startdata
-%FIXME with new method
 %TODO numbers with updated method
 {\it Transit/RV parameters:} & & & \\
-  $R_{\rm p}/R_\star$                        & $0.1538$               & $\pm 0.0013$                & A \\
-  $i$~[deg]                                  & $88.52$                & $+0.39$, $-0.26$            & A \\
-  $a/R_\star$                                & $5.463$                & $+0.025$, $-0.020$          & A \\
+  $R_{\rm p}/R_\star$                        & $0.XXXX$               & $\pm 0.0013$                & A \\
+  $i$~[deg]                                  & $XX.XX$                & $+0.39$, $-0.26$            & A \\
+  $a/R_\star$                                & $5.XXX$                & $+0.025$, $-0.020$          & A \\
   $u_{\rm linear}$                           & $0.XXX$                & $+0.XXX$, $-0.XXX$          & A \\
@@
 $\log g$, $[{\rm Fe/H}]$) and the available {\it Gaia} parallaxes, we
-compute the remaining system parameters empirically.
+calculate the remaining system parameters empirically.
 
@@
 offset \citep{stassun_evidence_2018}, we obtain the stellar radius.
-For the stellar mass, we started with the $a/R_\star$ value found by
-\citet{hoyer_tramos_2013} from 26 ground-based transits.  The implied
-semi-major axis and stellar density follow \citep{seager_unique_2003}.
-Combining the density and radius gives the stellar mass.  We get the
-planet radius using our measured $R_{\rm p}/R_\star$.  Finally, for
-the planet mass we use the observed RV semi-amplitude and inclination
-angle from the transit
-\citep{triaud_spin-orbit_2010,hoyer_tramos_2013}.
-
-The resulting parameters are in Table~3, and broadly agree with those
-of previous authors
+For the stellar mass, we started with the $a/R_\star$ value found from
+the phase-folded TESS lightcurve.  The implied stellar density follows
+\citep{seager_unique_2003}, and using the stellar radius we obtain the
+semi-major axis and stellar mass.  We get the planet radius using our
+measured $R_{\rm p}/R_\star$.  Finally, for the planet mass we use the
+observed RV semi-amplitude and inclination angle from the transit
+\citep{triaud_spin-orbit_2010}.
+
+We list the resulting parameters in Table~3, and adopt them for the
+remaining analysis.  Our parameters broadly agree with those of
+previous authors
 \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
-We use them for the remaining analysis.  Comparing the empirical
-result to the Yonsei-Yale evolutionary track predicted by the
-spectroscopic parameters, we see that WASP-4  has an age $\approx
-7\,{\rm Gyr}$, and is securely in the main-sequence stage of
-evolution.
+Comparing our empirical result with the Yonsei-Yale evolutionary track
+predicted by the spectroscopic parameters, we see that WASP-4 has an
+age $\approx 7\,{\rm Gyr}$, and is securely in the main-sequence stage
+of evolution.
 
@@
 
-\subsection{Previous timing measurements}
+\subsection{Pre-TESS timing measurements}
 \label{subsec:times}
@@
         {\bf Timing residuals and best-fit models for WASP-4b.}
-        Points are observed transit times minus a
-        linear fit to all the times.
-        Less precise times are more transparent.
-        The best constant period
-        model is a poor fit to the observations (gray line).
-        The best-fit tidal decay model (blue) and apsidal precession
-        model (orange) can fit the data.
-        The binned TESS point (star)
-        is for display purposes only.
+        Points are observed transit times minus a linear fit to all
+        observed times.  Less precise times are more transparent.  The
+        best constant period model (gray line) is a poor fit to the
+        observations.  The best-fit tidal decay model (blue) and
+        apsidal precession model (orange) can fit the data.  The
+        binned TESS point (star) is for display purposes only.
         \label{fig:times}
@@
 
-Table~1 shows the transit times used in our analysis.  We included
+Table~1 shows the transit times we used in our analysis.  We included
 data from peer-reviewed literature for which the analysis was based on
@@
 was fit as a free parameter. We also required that the time system be
-clearly documented.  The occultation times are in Table~3.  Since
-there are fewer available, we included all the occultation
-measurements we could find. The tabulated values have been corrected
-for the light-travel time across the diameter of the orbit by
-subtracting $2a/c = 22.8$ seconds from the observed time.
+clearly documented.  The occultation times are in Table~3.  There are
+fewer available, so we included all the occultation measurements we
+could find. The tabulated values have been corrected for the
+light-travel time across the diameter of the orbit by subtracting
+$2a/c = 22.8$ seconds from the observed time.
 
 We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point\footnote{We also tried using times directly from the
-original authors; the results did not change, so we opted for the
-\citet{hoyer_tramos_2013} times as the more homogeneous option.}.
+starting point\footnote{We also tried taking times directly from the
+references listed in Table~2; the results did not change, so we opted
+for the homogeneously determined \citet{hoyer_tramos_2013} times.}.
 Though their latest two epochs fell slightly below the prediction of a
@@
 epoch to be as close as possible to the average of the midtimes,
-weighted as the inverse midtime variance. This minimizes the
+weighted as the inverse midtime variance. This lowered the
 covariance between the transit epoch and the period
@@
 
-Important contributions to the timing dataset are as follows.  Our
-earliest epoch is from EulerCam on the 1.2-m Euler telescope
-\citep{wilson_wasp-4b_2008}.  The second epoch comes from $z$-band
+Important contributions to the timing dataset were as follows.  Our
+earliest epoch was from EulerCam on the 1.2-m Euler telescope
+\citep{wilson_wasp-4b_2008}.  The second epoch came from $z$-band
 photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
-with FORS2.  Subsequent transits came from
+with FORS2.  We included subsequent transits from
 \citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
@@
 telescope over 2011 to 2014, one transit per season.  The per-point
-standard deviation of their lightcurves is a few hundred parts per
+standard deviation of their lightcurves was a few hundred parts per
 million, within a factor of 3 from the photon noise limit.  The
-uncertainties on their reported transit times~--~on average
-5.6~seconds~--~ are small, but possible given the quality of their
+precision on their reported midtimes averaged 5.6~seconds. Though
+extremely precise, this was possible given the quality of their
 data~\citep{carter_analytic_2008}.
@@
 
-There are fewer available occultation times.
+There were fewer available occultation times.
 \citet{beerer_secondary_2011} observed two occultations of WASP-4b
@@
 \citet{caceres_ground-based_2011} detected an occultation from the
-ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
+ground in the $K_S$ band, and gave a time in HJD, without
 specifying the time standard.  We assumed the standard was UTC, but
@@
 where $N=35$ is the appropriate number of leap-seconds
-\citep[see][]{eastman_achieving_2010}.
-Though not rigorous, this procedure inflates the error bar to a more
+\citep{eastman_achieving_2010}.
+Though not rigorous, this procedure inflated the error bar to a more
 realistic $\pm$94 seconds.
@@
 the Anglo-Australian Telescope, focusing on the eclipse depth, but
-also fitting for $e\cos\omega$.  We converted their $e\cos\omega$
-result into a midtime, using their ephemeris and the standard formula
-\citep[{\it e.g.},][]{winn_exoplanet_2010}
+also fitting for $e\cos\omega$.  We converted their reported
+$e\cos\omega$ to a midtime, using their ephemeris and the standard
+formula \citep[{\it e.g.},][]{winn_exoplanet_2010}
 \begin{equation}
@@
 \end{equation}
-This gives us a total of 4 occultations.  Although the occultation
-times are uncertain compared to the transits, they nonetheless help in
+In total, we had four occultations.  Although the occultation times
+were less precise than the transit times, they nonetheless helped 
 modeling the timing variations.
@@
 Figure~\ref{fig:arrived_early}. The TESS observations 
-arrived early.  One immediate concern was systematic errors in either
-the archival or the TESS timestamps. The tests we used to mitigate the
-latter possibility are described in Appendix~\ref{sec:verify_tess}.
+arrived early.  
+Simply out of caution, we were immediately concerned about systematic
+errors in either the archival or the TESS timestamps. The tests we
+used to mitigate the latter possibility are described in
+Appendix~\ref{sec:verify_tess}. Checks against systematic errors in
+the pre-TESS times are discussed below.
 
@@
 
-%TODO: final numbers
+%TODO: final numbers. note epoch changes...
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$.  Clearly, it is more likely that
-some systematic error could be affecting one of the most important
+corresponds to $p \approx 10^{-23}$.  A more likely possibility is
+clearly that a systematic error is affecting one of the most important
 sets of times: either the earliest times, the Huitson et al.\ Gemini
@@
 have taken every precaution against systematic offsets in the TESS
-dataset (Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by
+timestamps (Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by
 their reported times (J.-M. D\'esert, priv.\
-comm.)\footnote{\citealt{huitson_gemini_2017} corrected the times to
-be at mid-exposure. They also analyzed other hot Jupiters, which all
+comm.)\footnote{\citet{huitson_gemini_2017} corrected the times to be
+at mid-exposure. They also analyzed other hot Jupiters, which all
 showed mid-transit times as expected from a constant ephemeris.}.
@@
 %TODO: final numbers
-The quadratic model fits better than the precession model.  It is
-favored by $\Delta \chi^2 = 3.0$, and has two fewer free parameters.
-Another useful heuristic for model comparison is the Bayesian
-Information Criterion (BIC),
+The quadratic model provides a better fit to the data than the
+precession model.  It is favored by $\Delta \chi^2 = 3.0$, and has two
+fewer free parameters.  Another useful heuristic for model comparison
+is the Bayesian Information Criterion (BIC),
 \begin{equation}
@@
 and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
-BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor of $\approx
-4\times10^{4}$.  The Akaike Information Criterion
-similarly favors the orbital decay model, by $\Delta {\rm AIC} = 7$.
-This is strong evidence that the quadratic model describes the
-data better than the precession model \citep{kass_bayes_1995}.
+BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor $\approx
+4\times10^{4}$.  The Akaike Information Criterion similarly favors the
+orbital decay model, by $\Delta {\rm AIC} = 7$.  This is strong
+evidence that the quadratic model describes the data better than the
+precession model \citep{kass_bayes_1995}.
 
@@
   = - (3.82 \pm 0.38)\times 10^{-10}
-  = - (12.1) \pm 0.12\,{\rm ms}\,{\rm yr}^{-1}.
+  = - (12.1) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
 
-The median parameters and their standard deviations are reported for
-the three models in Table~4.  In short, a linear ephemeris does not
-describe the observed times.  Both orbital decay and precession can
-fit the data, but the precession model is statistically disfavored.
-
-\section{Implications}
+The median parameters and their 68\% confidence intervals are reported
+for the three models in Table~4.  In short, a linear ephemeris does
+not describe the observed times.  Both orbital decay and precession
+can fit the data, but the precession model is statistically
+disfavored.
+
+\section{Interpretation}
 \label{sec:implications}
@@
   \begin{center}
-    \includegraphics[width=0.425\textwidth]{f4.pdf}
+    \includegraphics[width=0.40\textwidth]{f4.pdf}
   \end{center}
@@
   \caption{
-    {\bf WASP-4b has a short eccentricity damping time and medium
+    {\bf WASP-4b has a short eccentricity damping time and a medium
     decay time compared to other hot Jupiters.  It seems to orbit a
@@
     {\bf \it Top:}
-    Transiting giant planets from \citet{bonomo_gaps_2017}, who
+    Points are
+    transiting giant planets from \citet{bonomo_gaps_2017}, who
     measured the eccentricities using radial velocities.  Solid
@@
 orbital decay time of WASP-4b compared to other transiting giant
-planets.  Though roughly 20 hot Jupiters have shorter predicted decay
-times, most are not as well-observed as WASP-4b. Moreover, some hot
-Jupiters, {\it e.g.}, WASP-18b, orbit fully radiative stars, which may
-dissipate tidal energy less efficiently than stars that are partially
-convective.
+planets.  Although roughly 20 hot Jupiters have shorter predicted
+orbital decay times than WASP-4b, most have short observing baselines.
+A separate consideration is that some hot Jupiters host stars ({\it
+e.g.}, WASP-18) are fully radiative, and may dissipate tidal energy
+less efficiently than stars with convective zones, like WASP-4.
 
 For the sake of argument, let us assume that the timing variation is
-caused entirely by orbital decay.  If the period decreases
+caused entirely by orbital decay.  If the orbital period decreases
 at a fixed rate, then it will reach zero after
@@
 
-Let us assume further that the ``constant phase lag'' model for tidal
-interaction in a binary applies \citep{zahn_tidal_1977}.  The fluid
-response is parametrized by a modified\footnote{For stars, $k_\star
-\sim \mathcal{O}(10^{-2})$, so it is important to explicitly
-distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
+Let us further assume that the ``constant phase lag'' model for tidal
+interaction in a binary applies \citep{zahn_tidal_1977}.  The star's
+response to tidal forcing is parametrized by a modified\footnote{For
+stars, $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to
+explicitly distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
 e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
@@
 which is smaller when the star's density distribution is more
-centrally concentrated. $Q_\star$ is the ratio of the energy stored in
-the equilibrium deformation of the star, divided by the energy lost to
-heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A
+centrally concentrated. $Q_\star$ is the ratio between the energy
+stored in the equilibrium deformation of the star and the energy lost
+to heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A
 larger $Q_\star$ implies less efficient tidal dissipiation.  Though
@@
 
-If the planet's spin and orbit are synchronized, the star is slowly
-rotating, and the eccentricity is small, then the semi-major axis and
-eccentricity evolve as \citep[Appendix B of][]{metzger_optical_2012}
+If the planet's spin and orbit are synchronized and the star is slowly
+rotating, then the semi-major axis and eccentricity evolve as
+\citep[Appendix B of][]{metzger_optical_2012}
 \begin{align}
@@
 The modified quality factor of WASP-4 corresponding to the observed
-value of $\dot{P}$ would then be
+value of $\dot{P}$ is then
 \begin{equation}
-	Q_\star' \approx 3.3\times10^4. 
+	Q_\star' \approx 3.3\times10^4,
 \end{equation}
-This is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
+which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
 10^5$, based on the orbits of the Galilean moons
 \citep{lainey_strong_2009}.  For stars, studies of the binary
-eccentricity distribution have yielded $Q_\star' \approx 10^5 - 10^7$
+eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
 \citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
@@
 phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
-(1.2^{+1.0}_{-0.5})\times10^7$.  This disagrees strongly with the
-value we infer from the apparent period change.
+(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
+quality factor we infer from the apparent period change.
 
@@
 \citet{essick_orbital_2016} proposed a theoretically-motivated tidal
-dissipation model , in which a coupled network of gravity modes would
-be excited at the base of the stellar convective zone, and would brake
-near the stellar core to provide the needed dissipation.  In their
+dissipation model , in which a coupled network of gravity modes is
+excited at the base of the stellar convective zone, and brakes near
+the stellar core to provide the needed dissipation.  In their
 approach, the stellar quality factors in hot Jupiter systems vary from
-$Q_\star' \approx 10^5 - 10^6$.  From their Equation 26, the predicted
-quality factor for WASP-4 is $Q_\star' = 7\times10^5$.  This is an
-order of magnitude larger than implied by the observed period change.
+$Q_\star' \approx 10^5 - 10^6$.  From their Equation 26, the
+prediction for WASP-4 is $Q_\star' = 7\times10^5$.  This is an order
+of magnitude larger than implied by the observed period change.
 
@@
 The x-axis is the effective temperature from \citet{bonomo_gaps_2017},
-which for WASP-4 agrees within $1\sigma$ of that from Table~2.
-Inspecting the HR diagram, WASP-4 shows litle evidence of being
+which for WASP-4 agrees within $1\sigma$ of that from Table~1.
+Inspecting the HR diagram, WASP-4 shows little evidence of being
 evolved, in agreement with our analysis from
@@
 
-In summary, if the observed period change is caused entirely by tidal
-decay, it would imply a tidal dissipation efficiency stronger than
-most predictions.  It might be possible that we are observing at a
-special time, shortly after the planet's inward migration, or when the
-planet is near resonance with a stellar oscillation mode.  Tidal
-dissipation rates might also be increased if the star is turning off
-the main sequence.  A final possibility is that an exterior planet
-could be trapping WASP-4b's spin vector in a high-obliquity state,
-causing planetary obliquity tides as suggested for WASP-12b
-\citep{millholland_obliquity_2018}.  For now, all of these
-explanations seem strained; it is unclear that tidal decay can explain
-the observed timing variations by any mechanism that we have
+To summarize, if the observed period change is caused entirely by
+tidal decay, then the implied stellar tidal dissipation efficiency
+would be stronger than expected.  It might be possible that we are
+observing at a special time, shortly after the planet's inward
+migration, or when the planet is near resonance with a stellar
+oscillation mode.  Tidal dissipation rates might also be increased if
+the star is turning off the main sequence.  A final possibility is
+that an exterior planet could be trapping WASP-4b's spin vector in a
+high-obliquity state, causing planetary obliquity tides as suggested
+for WASP-12b \citep{millholland_obliquity_2018}.  For now, all of
+these explanations seem strained; it is unclear that tidal decay can
+explain the observed timing variations by any mechanism that we have
 considered.
@@
 
-Let us instead assume that the timing variation can be explained as
-being a portion of an apsidal precession cycle.  The required orbital
-eccentricity is then $e\approx 1.6\times10^{-3}$, and one full
-precession period would take about 24 years.
+Let us instead assume that the timing variation is a portion of an
+apsidal precession cycle.  The required orbital eccentricity is then
+$e\approx 1.6\times10^{-3}$, and one full precession period would take
+about 24 years.
 
 \citet{ragozzine_probing_2009} calculated apsidal precession periods
-for hot Jupiters ranging between of order 10 and 100 years.  They
+for hot Jupiters ranging between $\approx$10--100 years.  They
 highlighted that for many hot Jupiters, including WASP-4b, over 90\%
@@
 non-zero eccentricity, a measurement of $\dot{\omega}$ can be used to
-measure the planet's Love number, $k_{2,\rm p}$.  From their Equation
-14, the implied Love number for WASP-4b is
+measure the planet's Love number.  From their Equation 14, the implied
+Love number for WASP-4b is
 \begin{equation}
@@
 \citep{wahl_tidal_2016,ni_empirical_2018}.  A uniform density sphere
-has $k_2 = 1.5$.  The uncertainties on $k_{2,{\rm p}}$ are large
+has $k_2 = 1.5$.  Our uncertainties on $k_{2,{\rm p}}$ are large
 because the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$
@@
 0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
-$1\,{\rm Gyr}$ \citep{winn_transit_2009} old, so any initial
-eccentricity should have been erased.
+$1\,{\rm Gyr}$ old \citep[{\it e.g.},][]{winn_transit_2009}, so any
+initial eccentricity should have been erased.
 
@@
 To summarize, empirical constraints on the ellipticity of hot Jupiter
-orbits at the $e \lesssim 10^{-3}$ level are present only for HD
-189733b and HD 209458b, from occultation timing.  Once the timing
+orbits at the $e \lesssim 10^{-3}$ level are present for only a
+handful of hot Jupiters, from occultation timing.  Once the timing
 offsets are of order tens of seconds, eccentricities from occultation
@@
 \paragraph{Neighboring companion}
-\citet{mardling_long-term_2007} considered the long-term tidal evolution
-of hot Jupiters with companions.  They pointed out that although the
-early phases of the two-planet eccentricity evolution occur
-quickly, the final phase of the joint eccentricity evolution towards
-circularity would occur on timescales several orders of magnitude
-longer than the circularization time (see their Figures~4 and 5).  The
-companion in their model is coplanar, and can have a mass down to an
-Earth-mass; the main requirement is that both the hot Jupiter and the
-outer companion start on eccentric orbits.
-
-Another way to excite the hot Jupiter's eccentricity is the
+\citet{mardling_long-term_2007} considered the long-term tidal
+evolution of hot Jupiters with companions.  The companion in their
+model is coplanar, and can have a mass down to an Earth-mass; the main
+requirement is that both the hot Jupiter and the outer companion start
+on eccentric orbits.  They found that although the early phases of the
+two-planet eccentricity evolution occur quickly, the final phase of
+the joint eccentricity evolution towards circularity would occur on
+timescales several orders of magnitude longer than the circularization
+time (see their Figures~4 and 5).
+
+A different way to excite the hot Jupiter's eccentricity is the
 Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
@@
 %companion's mass.
-For the Kozai-Lidov mechanism to operate at maximum efficiency, from
-\citet{bailey_understanding_2019} Equation 20 we need
+For the Kozai-Lidov mechanism to operate at maximum efficiency, we
+need \citep[][Equation 20]{bailey_understanding_2019}
 \begin{equation}
@@
 semi-major axis between 1-20\,AU is about 50\%, and a linear trend is
-currently present at $\sim 2\sigma$ \citep{knutson_friends_2014}.
+currently present at $\sim$2$\sigma$ \citep{knutson_friends_2014}.
 
 \paragraph{Fluctuations in the gravitational potential from convection}
-A separate mechanism to pump the eccentricity involves the
+A separate mechanism to pump the eccentricity invokes the
 gravitational fluctuations from stellar convection \citep[][Section
@@
 their associated kinetic energy perturbs the orbit away from
-circularity.  In particular, from \citealt{phinney_pulsars_1992}
-Equation 7.33,
+circularity.  The mean-squared eccentricity is \citep[][Equation
+7.33]{phinney_pulsars_1992}
 \begin{equation}
@@
 width (mass) of the convective region,  $\mu$ the reduced mass, $n$
-the orbital frequency, and $a$ the semi-major axis.  For WASP-4,
-convection in the star is more important than convection in the
-planet, and $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$.
+the orbital frequency, and $a$ the semi-major axis.  For WASP-4, we
+find $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$; this mechanism
+cannot produce the required eccentricity.
 
@@
 dynamo timescale").  An essential difference between this process and
-apsidal precession is that it is expected to be quasi-periodic
-\citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}.  The transit
-and occultation deviations would also have the same sign, while for
-apsidal precession they have opposite signs.  For WASP-4,
-\citet{watson_orbital_2010} estimated that the effect could produce
-timing deviations of up to 15 seconds, depending on the modulation
-period of the stellar dynamo, and the corresponding level of
-differential surface shear.  This cannot explain the majority our
-observed $78$ second variation.
+apsidal precession is that Applegate timing variations are expected to
+be quasi-periodic \citep[{\it
+e.g.},][Figure~12]{soderhjelm_geometry_1980}.  The transit and
+occultation deviations from the Applegate mechanism would also have
+equal signs, while for apsidal precession they have opposite signs.
+For WASP-4, \citet{watson_orbital_2010} estimated that the Applegate
+effect could produce timing deviations of up to 15 seconds, depending
+on the modulation period of the stellar dynamo, and the corresponding
+level of differential surface shear.  If true, this mechanism cannot
+explain the majority our observed $78$ second variation.
 
@@
 Another mechanism to produce the observed period change would be if an
-unseen companion in the WASP-4 system accelerated the star towards us.
+unseen companion in the WASP-4 system accelerated the star towards our
+line of sight.
 This would produce a period derivative
@@
 monitoring will help clarify the situation,
-particularly given the fact that WASP-4 shows short-term trends with
+particularly given that WASP-4 shows short-term RV trends with
 %$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
-amplitudes 100 times larger, over tens of days
-\citep{husnoo_observational_2012}.  Given the known activity of the
-system~\citep{sanchis-ojeda_starspots_2011}, these short-timescale
-variations might conspire to produce spurious measurements over long
-baselines.
+amplitudes $\sim$$\mathcal{O}(10\,{\rm m\,s}^{-1})$, over tens of days
+\citep{husnoo_observational_2012}.  Given that WASP-4 is a
+magnetically active star \citep{sanchis-ojeda_starspots_2011}, these
+short-timescale variations might conspire to produce spurious
+RV signals over long baselines.
 
@@
 observed apsidal precession.  The corresponding period change is
-$\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$.
-
-Many of the phenomena discussed could presumably occur simultaneously.
-We found though that radial acceleration, proper-motion, and the
-Applegate mechanism are unlikely to explain the majority of the timing
-deviations.  Tidal decay seems implausible, given the high level of
-dissipation required.  This leaves apsidal precession as the most
-plausible explanation, but more data are needed for a definitive
-interpretation.
+$\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$. Both effects are too small
+to be of consequence in WASP-4.
+
+Many of the phenomena we have discussed could presumably occur
+simultaneously.  We found though that radial acceleration,
+proper-motion, and the Applegate mechanism are unlikely to explain the
+majority of the timing deviations.  Tidal decay seems implausible,
+given the high level of dissipation required.  This leaves apsidal
+precession as the most plausible explanation, but more data are needed
+for a definitive interpretation.
 
@@
 
-If TESS continues observing after its primary mission, it will observe
-additional transits of WASP-4b in the early 2020s
+If the TESS mission is extended after its primary mission, it will
+likely observe additional transits of WASP-4b in the early 2020s
 (Figure~\ref{fig:future}).  High-precision ground-based transit
@@
 dissipation efficiency, or the eccentricity and Love number of the
-planet.  Either outcome for a single system would be somewhat
+planet.  Either outcome for a single system would be rather
 interesting.  In the coming decade, wide-field photometric surveys
@@
     {\bf There is no evidence for a systematic offset between TESS
-    times and the barycentric reference.}
-    While the WASP-4b transits fell about 75 seconds earlier than
-    predicted, other well-observed hot Jupiters, in particular WASP-6b
-    and WASP-18b, arrived on time.  Ticks are observed TESS transit
-    midtimes; the blue distribution function is a kernel density
-    estimate; the orange distribution function is a gaussian centered
-    on zero using the indicated $1\sigma$ standard deviation in the
-    prediction
-    ($\sigma_{\rm predicted}$).
+    times and the barycentric reference.} While the WASP-4b transits
+    fell about 75 seconds earlier than expected, other well-observed
+    hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
+    Ticks are observed TESS transit midtimes; the blue distribution is
+    a kernel density estimate; the orange distribution is a gaussian
+    centered on zero with standard deviation ($\sigma_{\rm
+    predicted}$) calculated from the pre-TESS transit times.
     \label{fig:hjs}

### 2018-12-13 paper/ms.tex
recomputed stellar/planet properties w/ TESS data
specifically, the TESS phase-folded LC

@@
 \citet{huitson_gemini_2017}.
-%TODO: is this true? Or do the periods actually differ?
 
@@
 \tablehead{
-\colhead{Parameter} & \colhead{Value} & \colhead{Uncertainty} & \colhead{Comment}
+\colhead{Parameter} & \colhead{Value} & \colhead{68\% Confidence Interval} & \colhead{Comment}
 }
@@
 \startdata
-%TODO numbers with updated method
 {\it Transit/RV parameters:} & & & \\
-  $R_{\rm p}/R_\star$                        & $0.XXXX$               & $\pm 0.0013$                & A \\
-  $i$~[deg]                                  & $XX.XX$                & $+0.39$, $-0.26$            & A \\
-  $a/R_\star$                                & $5.XXX$                & $+0.025$, $-0.020$          & A \\
-  $u_{\rm linear}$                           & $0.XXX$                & $+0.XXX$, $-0.XXX$          & A \\
-  $u_{\rm quad}$                             & $0.XXX$                & $+0.XXX$, $-0.XXX$          & A \\
+  $R_{\rm p}/R_\star$                        & $0.15117$              & $+0.00047$, $-0.00037$      & A \\
+  $i$~[deg]                                  & $88.85$                & $+0.77$, $-0.90$            & A \\
+  $a/R_\star$                                & $5.457$                & $+0.030$, $-0.066$          & A \\
+  $u_{\rm linear}$                           & $0.388$                & ---                         & A \\
+  $u_{\rm quad}$                             & $0.212$                & ---                         & A \\
   $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
@@
   $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
-  $\rho_\star$~[g~cm$^{-3}$]                 & $1.722$                & $+0.024$, $-0.019$          & E \\
-  $M_\star$~[M$_{\odot}$]                    & $0.870$                & $+0.085$, $-0.077$          & E \\
+  $\rho_\star$~[g~cm$^{-3}$]                 & $1.717$                & $+0.029$, $-0.062$          & E \\
+  $M_\star$~[M$_{\odot}$]                    & $0.867$                & $+0.087$, $-0.097$          & E \\
   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
 {\it Planetary parameters:} & & & \\
-  $a$~[AU]                                   & $0.0227$               & $\pm 0.0007$                & E \\
-  $M_p$~[M$_{\rm Jup}$]                      & $1.192$                & $+0.091$, $-0.086$          & E \\
-  $R_p$~[R$_{\rm Jup}$]                      & $1.337$                & $\pm 0.048$                 & E \\
+  $a$~[AU]                                   & $0.0227$               & $\pm 0.0008$                & E \\
+  $M_p$~[M$_{\rm Jup}$]                      & $1.189$                & $+0.093$, $-0.104$          & E \\
+  $R_p$~[R$_{\rm Jup}$]                      & $1.314$                & $\pm 0.039$                 & E \\
 \enddata
@@
 \tablecomments{
-  (A) From TESS lightcurves (\S~\ref{sec:measurement}). Orbital periods from timing data are in Table~4.
+  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
+  Orbital periods are in Table~4. The limb darkening parameters were
+  allowed to float around the \citet{claret_limb_2017} prediction, but
+  were unconstrained.
   (B) \citep{triaud_spin-orbit_2010}.
@@
 We list the resulting parameters in Table~3, and adopt them for the
-remaining analysis.  Our parameters broadly agree with those of
-previous authors
+remaining analysis.  Our parameters broadly agree with those of previous
+authors
 \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
+% The sole exception is that our measured $R_{\rm p}/R_\star$ value is
+% slightly smaller than most ground-based measurements (see Tables 2 and
+% 3 of \citealt{hoyer_tramos_2013}).  However, looking at the I-band
+% measurements of \citet{hoyer_tramos_2013} and taking the mean + std
+% deviation, it's a ~1 sigma effect. So I won't mention it!
 Comparing our empirical result with the Yonsei-Yale evolutionary track
@@
 \citet{caceres_ground-based_2011} detected an occultation from the
-ground in the $K_S$ band, and gave a time in HJD, without
-specifying the time standard.  We assumed the standard was UTC, but
-added $32.182 + N$ seconds of uncertainty to the reported errors,
-where $N=35$ is the appropriate number of leap-seconds
-\citep{eastman_achieving_2010}.
-Though not rigorous, this procedure inflated the error bar to a more
-realistic $\pm$94 seconds.
+ground in the $K_S$ band, and gave a time in HJD, without specifying
+the time standard.  We assumed the standard was UTC, and verified that
+assuming TDB did not alter any of our conclusions.
 %Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}

### 2018-12-13 src/compute_planet_star_parameters.py
update compute_planet_star_parameters to use TESS lc

@@ -4,15 +4,82 @@ estimate the stellar density, and mass.
 """
 from __future__ import division, print_function
 import numpy as np
+import pickle
 
 from astropy import units as u, constants as const
 
-# use parameters from Table 4 of Hoyer+2013
-i = 88.52*u.deg
-a_by_Rs = 5.463
-a_by_Rs_perr = 0.025
-a_by_Rs_merr = 0.020
-P = 1.33823204*u.day
+# Keivan value
+Rstar = 0.893*u.Rsun
+Rstar_err = 0.024*u.Rsun
+
+# parameters from Table 4 of Hoyer+2013, for comparison
+i_hoyer13 = 88.52*u.deg
+a_by_Rs_hoyer13 = 5.463
+a_by_Rs_perr_hoyer13 = 0.025
+a_by_Rs_merr_hoyer13 = 0.020
+P_hoyer13 = 1.33823204*u.day
+hoyer_params = [i_hoyer13, a_by_Rs_hoyer13, a_by_Rs_perr_hoyer13,
+                a_by_Rs_merr_hoyer13, P_hoyer13]
+
+# get parameters from fit to phase-folded lightcurve
+picklepath = ('/home/luke/Dropbox/proj/tessorbitaldecay/results/'
+              'tess_lightcurve_fit_parameters/402026209/'
+              '402026209_phased_mandelagol_fit_empiricalerrs.pickle' )
+
+d = pickle.load(open(picklepath, 'rb'))
+i = d['fitinfo']['finalparams']['incl']*u.deg
+i_upper = i + 0.39*u.deg
+
+i = 88.52*u.deg # Hoyer+ 2013, table 4
+i_upper = i + d['fitinfo']['finalparamerrs']['std_perrs']['incl']*u.deg
+i_lower = i - d['fitinfo']['finalparamerrs']['std_merrs']['incl']*u.deg
+a_by_Rs = d['fitinfo']['finalparams']['sma']
+a_by_Rs_perr = d['fitinfo']['finalparamerrs']['std_perrs']['sma']
+a_by_Rs_merr = d['fitinfo']['finalparamerrs']['std_merrs']['sma']
+P = d['fitinfo']['finalparams']['period']*u.day
+
+a_by_Rs = d['fitinfo']['finalparams']['sma']
+a_by_Rs_perr = d['fitinfo']['finalparamerrs']['std_perrs']['sma']
[…]

### 2018-12-13 src/calculate_wasp4b_eccentricity_damping_timescale.py
update numbers using Stassun parameters

@@ -11,11 +11,11 @@ from astropy import units as u, constants as const
 # Petrucci+ 2013, table 3, for all parameters
 P = 1.33823251*u.day
 
+#Mstar, Rstar, Mplanet, Rplanet = 0.867, 0.893, 1.189, 1.314 # USED: my table 1
 Mstar, Rstar, Mplanet, Rplanet = (
-    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
 )
-
-sma = 0.0228*u.au
+sma = 0.0227*u.au
 
 Qplanetprime = 1e5
 

### 2018-12-13 src/kozai_companion_mass_bound.py

@@ -8,11 +8,12 @@ import numpy as np
 
 from astropy import units as u, constants as const
 
-# Petrucci+ 2013, table 3, for all parameters
-a_b = 0.0228*u.au
+# parameters from my table
+a_b = 0.0227*u.au
 M_star, R_star, M_b, R_b = (
-    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
 )
+
 
 # This number is made up, but similar to Jupiter's from Juno (Wahl+ 2016).
 k_2b = 0.6

### 2018-12-13 src/make_select_transit_times_table.py

@@ -64,6 +64,8 @@ def main():
             references.append('\citet{huitson_gemini_2017}')
         elif ref == 'me':
             references.append('This work')
+        elif ref == 'Baxter et al. (in prep)':
+            references.append('Baxter et al.\ (in prep)')
     references = np.array(references)
 
     outdf = pd.DataFrame(

### 2018-12-13 src/rv_companion_mass_bound.py

@@ -8,10 +8,10 @@ import numpy as np
 
 from astropy import units as u, constants as const
 
-# Petrucci+ 2013, table 3, for all parameters
-a_b = 0.0228*u.au
+# parameters from my table
+a_b = 0.0227*u.au
 M_star, R_star, M_b, R_b = (
-    0.89*u.Msun, 0.92*u.Rsun, 1.216*u.Mjup, 1.33*u.Rjup
+    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
 )
 
 # from equation (52) of the latex'd equation in Avery's email

### 2018-12-13 src/plot_arrived_early.py
update arrived_early plot

@@ -47,10 +47,6 @@ def plot_arrived_early(plname, xlim=None, ylim=None, savpath=None):
     fig,(a0,a1) = plt.subplots(nrows=2,ncols=1,figsize=(4,7))
 
     # transit axis
-    cuterr = np.percentile(err_tmid, 50)
-    print('showing points with err > {:.2f} seconds as solid'.
-          format(cuterr*60))
-
     _ix = 0
     for e, tm, err in zip(epoch,tmid,err_tmid):
         if _ix == 0:
@@ -75,7 +71,7 @@ def plot_arrived_early(plname, xlim=None, ylim=None, savpath=None):
                         alpha= 1-(err/np.max(err_tmid))**(1/2) + 0.1
                        )
 
-    for ax in (a0,a1):
+    for ax in [a1]:
         ax.errorbar(tess_epoch,
                     nparr(tess_tmid -
                           linear_model(lsfit_t0, lsfit_period, tess_epoch))*24*60,
@@ -84,6 +80,12 @@ def plot_arrived_early(plname, xlim=None, ylim=None, savpath=None):
                     ms=3,
                     elinewidth=1,
                     label='TESS')
+
+    # for the legend
+    a0.errorbar(9001, 9001, np.mean(tess_err_tmid*24*60),
+                fmt='sk', ecolor='black', zorder=9, alpha=1, mew=1, ms=3,
+                elinewidth=1, label='TESS')
+
 
     bin_tess_y = np.average(nparr(
         tess_tmid-linear_model(lsfit_t0, lsfit_period, tess_epoch)),
@@ -99,13 +101,13 @@ def plot_arrived_early(plname, xlim=None, ylim=None, savpath=None):
     print('error (plotted, sec): {}'.format(bin_tess_err_tmid*24*60*60))
     bin_tess_x = np.median(tess_epoch)
 
-    for ax in (a0,a1):
+    for ax in [a0]:
         ax.errorbar(bin_tess_x, bin_tess_y*24*60, bin_tess_err_tmid*24*60,
                     alpha=1, zorder=11, label='binned TESS',
-                    fmt='s', mfc='red', elinewidth=1,
+                    fmt='s', mfc='firebrick', elinewidth=1,
                     ms=3,
-                    mec='red',mew=1,
-                    ecolor='red')
+                    mec='firebrick',mew=1,
+                    ecolor='firebrick')
 
     yupper = (
[…]

### 2018-12-13 paper/ms.tex
update text w/ final(?) values

@@
 
-% TODO: final numbers
 To verify that the measured uncertainties are plausible, we computed
 the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
-transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
+transit times.  We found that $\chi^2 = 9.2$, with 16 degrees of
 freedom.  Visually inspecting residuals showed that the error variance
 had been overestimated, so we multiplied the measured TESS errors by a
-factor $f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered
-the mean midtime uncertainty for each transit from 30 to 22 seconds.
-
-% TODO: final numbers
+factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered
+the mean uncertainty of the transit midtimes from $29.8$ to $22.6$
+seconds.
+
 Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
@@
 Figure~\ref{fig:times} shows the residuals with respect to the linear
-model.  The best linear fit has $\chi^2 = 149$ and 58 degrees of
-freedom.  The best quadratic fit has $\chi^2 = 48.0$ and 57
-degrees of freedom, compared to $\chi^2 = 51.0$
-and 55 degrees of freedom for the best-fit precession model.
+model.  The best linear fit has $\chi^2 = 174$ and 61 degrees of
+freedom.  The best quadratic fit has $\chi^2 = 62.6$ and 60
+degrees of freedom, compared to $\chi^2 = 64.4$
+and 58 degrees of freedom for the best-fit precession model.
 
@@
 The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-23}$.  A more likely possibility is
+corresponds to $p \approx 10^{-26}$.  A more likely possibility is
 clearly that a systematic error is affecting one of the most important
 sets of times: either the earliest times, the Huitson et al.\ Gemini
-times (epoch numbers 44, 322, 591, and 851), or the TESS times.  We
+times (epoch numbers 30, 308, 577, and 837), or the TESS times.  We
 have taken every precaution against systematic offsets in the TESS
-timestamps (Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by
+times (Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by
 their reported times (J.-M. D\'esert, priv.\
@@
 
-%TODO: final numbers
-The quadratic model provides a better fit to the data than the
-precession model.  It is favored by $\Delta \chi^2 = 3.0$, and has two
+The quadratic model provides a slightly better fit to the data than the
+precession model.  It is favored by $\Delta \chi^2 = 1.8$, and has two
 fewer free parameters.  Another useful heuristic for model comparison
@@
 for $k$ the number of free parameters, and $n$ the number of data
-points. For us, $n=60$.  The difference in the BIC for the precession
+points. For us, $n=62$.  The difference in the BIC for the precession
 and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
-BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor $\approx
-4\times10^{4}$.  The Akaike Information Criterion similarly favors the
-orbital decay model, by $\Delta {\rm AIC} = 7$.  This is strong
+BIC}_{\rm quad} = 10$, corrsonding to a Bayes factor $\approx
+1.2\times10^{4}$.  The Akaike Information Criterion similarly favors
+the orbital decay model, by $\Delta {\rm AIC} = 5.8$.  This is strong
 evidence that the quadratic model describes the data better than the
@@
 
-%TODO: final numbers
 If we assume the orbital decay model, we find a period derivative 
@@
 \dot{P}
-  = - (3.82 \pm 0.38)\times 10^{-10}
-  = - (12.1) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
+  = - (3.98 \pm 0.38)\times 10^{-10}
+  = - (12.6) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
 
-%TODO: final numbers
 If we instead assume the precession model, the best-fit eccentricity
@@
 \begin{equation}
-  e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}
+  e = (1.83^{+ 1.88}_{- 0.75})\times10^{-3}
 \end{equation}
-The longitude of periapse correspondingly advances by $\dot{\omega}
-= 14.8^{+5.4}_{-3.8}\ {\rm degrees}\,{\rm yr}^{-1}$, and the
-precession period would be $24^{+8}_{-6}$ years.
+The longitude of periapse correspondingly advances by $\dot{\omega} =
+14.0^{+4.9}_{-3.5}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
+period would be $26^{+9}_{-7}$ years.
 
@@
     {\bf \it Top:}
-    Points are
-    transiting giant planets from \citet{bonomo_gaps_2017}, who
-    measured the eccentricities using radial velocities.  Solid
+    Points are transiting giant planets from \citet{bonomo_gaps_2017},
+    who measured the eccentricities using radial velocities.  Solid
     squares (including WASP-18b) have significant eccentricity
@@
 \begin{equation}
-  \frac{P}{ \dot{P} } = 9.6 \, {\rm Myr}.
+  \frac{P}{ \dot{P} } = 9.2 \, {\rm Myr}.
 \end{equation}
@@
 \begin{equation}
-	Q_\star' \approx 3.3\times10^4,
+	Q_\star' \approx 2.9\times10^4,
 \end{equation}
@@
 apsidal precession cycle.  The required orbital eccentricity is then
-$e\approx 1.6\times10^{-3}$, and one full precession period would take
+$e\approx 1.8\times10^{-3}$, and one full precession period would take
 about 24 years.
@@
 \begin{equation}
-k_{2,{\rm p}} = 1.6^{+2.1}_{-1.2}.
+k_{2,{\rm p}} = 1.5^{+2.1}_{-1.1}.
 \end{equation}
@@
 eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
-0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
-$1\,{\rm Gyr}$ old \citep[{\it e.g.},][]{winn_transit_2009}, so any
+0.30 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
+$1\,{\rm Gyr}$ old, so unless the planet arrived very recently, any
 initial eccentricity should have been erased.
@@
 Placing eccentricity limits stronger than $e< 10^{-3}$ has only been
-possible for a few hot Jupiters, and with numerous false-starts.
-HD 189733b, for example, was initially reported to have a non-zero
-eccentricty
-\citep{knutson_map_2007}.  This result was retracted by
+possible for a few hot Jupiters, and with numerous false-starts.  HD
+189733b, for example, was initially reported to have a non-zero
+eccentricty \citep{knutson_map_2007}.  This result was retracted by
 \citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
@@
 \begin{equation}
-  M_{\rm c} > 7.7\,M_\oplus
+  M_{\rm c} > 7.1\,M_\oplus
   \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
@@
   M_{\rm c} &< 
-  24.2\,M_\oplus
+  23.8\,M_\oplus
   \times 
@@
 particularly given that WASP-4 shows short-term RV trends with
-%$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
 amplitudes $\sim$$\mathcal{O}(10\,{\rm m\,s}^{-1})$, over tens of days

### 2018-12-14 src/percent_to_sigma_converter.py
update clock offset statistics

@@ -26,18 +26,22 @@ def get_sigma_given_percent(percent):
 
 if __name__=="__main__":
 
-    print('WASP-6b, t_systematic_offset < -77.68 sec ruled out at XX sigma')
-    frac = 1 - 0.000000439560
+    print('WASP-6b, t_systematic_offset < -81.50 sec ruled out at XX sigma')
+    val_6b = 0.000000439560
+    frac = 1 - val_6b
     get_sigma_given_percent(frac)
 
-    print('WASP-46b, t_systematic_offset < -77.68 sec ruled out at XX sigma')
-    frac = 1 - 0.026633342800
+    print('WASP-46b, t_systematic_offset < -81.50 sec ruled out at XX sigma')
+    val_46b = 0.022780460097
+    frac = 1 - val_46b
     get_sigma_given_percent(frac)
 
-    print('WASP-18b, t_systematic_offset < -77.68 sec ruled out at XX sigma')
-    frac = 1- 0.015987957966
+    print('WASP-18b, t_systematic_offset < -81.50 sec ruled out at XX sigma')
+    val_18b = 0.012510650383
+    frac = 1- val_18b
     get_sigma_given_percent(frac)
 
-    print('multiply 6b,18b, 46b: t_systematic_offset < -77.68 sec ruled out at XX sigma')
-    frac = 1- (0.000000439560* 0.015987957966*0.026633342800)
+    print('multiply 6b,18b, 46b: t_systematic_offset < -81.50 sec ruled out at XX sigma')
+    frac = 1- val_6b*val_46b*val_18b
+    print('product is {}'.format(val_6b*val_46b*val_18b))
     get_sigma_given_percent(frac)

### 2018-12-14 paper/ms.tex
updated paper to include Spitzer times

@@
 \begin{abstract}
-  %TODO: final numbers
   By combining data from the Transiting Exoplanet Survey Satellite
@@
   times of WASP-4b are incompatible with a constant orbital period.
-  The transits observed by TESS occurred $77.8 \pm 10.8$~seconds
+  The transits observed by TESS occurred $81.5 \pm 11.8$~seconds
   earlier than one would expect, based on the previous data.  The
-  orbital period appears to be shrinking by $\dot{P}=-12.1 \pm 1.2$
+  orbital period appears to be shrinking by $\dot{P}=-12.6 \pm 1.2$
   milliseconds per year.   A simple clock offset is unlikely, because
   TESS observations of other hot Jupiters (WASP-6b, 18b, and 46b) are
-  compatible with a constant period and rule out a 77.8-second offset
+  compatible with a constant period and rule out a 81.5-second offset
   at the 6.3$\sigma$ level.  The timing variations seem to be
@@
   \approx 3\times10^4$.  Apsidal precession would imply an
-  eccentricity $e\approx1.6\times 10^{-3}$, and the possibility of
+  eccentricity $e\approx1.8\times 10^{-3}$, and the possibility of
   measuring the planet's Love number.  A Doppler shift from WASP-4's
@@
 array every 2 minutes.   The data were downlinked via the Deep Space
-Network, and the Payload Operations Center then converted the
-timestamps from the spacecraft clock to TESS Julian Day in the {\it
-Temps Dynamique Barycentrique} (TDB) reference.  The data were then
-processed through the Science Processing Operations Center (SPOC)
-pipeline at NASA Ames~\citep{jenkins_tess_2016}.  SPOC used the known
+Network. The timestamps from the spacecraft clock were then converted
+by the Payload Operations Center to TESS Julian Day in the {\it
+Temps Dynamique Barycentrique} (TDB) reference.  The images were then
+reduced to lightcurves by the Science Processing Operations Center (SPOC)
+at NASA Ames~\citep{jenkins_tess_2016}.  During this processing, SPOC used
+the known
 spacecraft trajectory to compute the barycentric time corrections on a
@@
 ${\rm BTJD} = {\rm BJD} - 2457000$, BJD is the barycentric julian
-date, and all times are in the TDB reference (see
-\citealt{urban_explanatory_2012} for explanations of time systems).
+date, and all times are in the TDB reference (\citealt{urban_explanatory_2012} 
+describe the relevant time systems).
 The lightcurves were then vetted and released by the MIT TESS Science
@@
 subsequent processing steps were performed using
-\texttt{astrobase}~\citep{bhatti_astrobase_2018} We opted not to
-``flatten'' the lightcurves at this stage, as is often done with {\it
+\texttt{astrobase}~\citep{bhatti_astrobase_2018} We did not
+``flatten'' the lightcurves, as is often done with {\it
 e.g.}, a spline or gaussian process regression.  Such procedures are
-asymmetric about the transit minimum and can skew the transit time
+asymmetric about the transit minimum and can skew the midtime
 measurement without being captured by the measurement uncertainties.
@@
 
-Having defined the free and fixed parameters, we calculated an initial
-guess for the free parameters by maximizing the likelihood.  We then
+Having defined the free and fixed parameters, %we calculated an initial
+%guess for the free parameters by maximizing the likelihood.  W
+we then
 sampled over the posterior using the algorithm proposed by
@@
 Table~1 gives the results, which are in reasonable agreement with the
-parameters reported by \citet{gillon_improved_2009},
-\citet{southworth_high-precision_2009}, and
+parameters reported by {\it e.g.},
+\citet{southworth_high-precision_2009} and 
 \citet{huitson_gemini_2017}.
 
-To measure mid-transit times, we then focused on each individual
-transit window.  We fit for a model transit, plus a local linear
-trend.  We allowed four free parameters per transit: the time of
+To measure the transit midtimes, we then fit  each individual transit
+for a model transit, plus a local linear trend.  We allowed four free 
+parameters per transit: the time of
 mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
 slope and intercept of the line.  We fixed the remaining
-parameters~---~$a/R_\star$, $i$, $P$, and the limb-darkening
-coefficients~---~to the best-fit values determined from the
+parameters%~---~$a/R_\star$, $i$, $P$, and the limb-darkening
+%coefficients~---~
+to the best-fit values determined from the
 phase-folded TESS lightcurve.
@@
 photometric error bars to equal the standard deviation of the
-out-of-transit points of the residual lightcurve.  We then reperformed
+out-of-transit points of the residual lightcurve.  We then repeated
 the fitting procedure, using the empirically determined photometric
@@
 hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
-photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  At
-$T\approx11.8$, with a 9-pixel aperture read noise is expected to
+photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  For a
+$T\approx11.8$ star with a 9-pixel aperture, read noise is expected to
 contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
@@
 the phase-folded TESS lightcurve.  The implied stellar density follows
-\citep{seager_unique_2003}, and using the stellar radius we obtain the
+\citep{seager_unique_2003}, as do the
 semi-major axis and stellar mass.  We get the planet radius using our
-measured $R_{\rm p}/R_\star$.  Finally, for the planet mass we use the
-observed RV semi-amplitude and inclination angle from the transit
-\citep{triaud_spin-orbit_2010}.
-
-We list the resulting parameters in Table~3, and adopt them for the
+$R_{\rm p}/R_\star$ measurement.  Finally, for the planet mass we use the
+inclination angle from the transit, and the RV semi-amplitude from 
+\citet{triaud_spin-orbit_2010}.
+
+We list the resulting parameters in Table~1, and adopt them for the
 remaining analysis.  Our parameters broadly agree with those of previous
@@
         apsidal precession model (orange) can fit the data.  The
-        binned TESS point (star) is for display purposes only.
+        binned TESS point (red square) is for display purposes only;
+        the individual TESS times are used to fit the models.
         \label{fig:times}
@@
   \begin{center}
-    \includegraphics[width=0.40\textwidth]{f4.pdf}
+    \includegraphics[width=0.417\textwidth]{f4.pdf}
   \end{center}
@@
   \caption{
-    {\bf WASP-4b has a short eccentricity damping time and a medium
+    {\bf WASP-4b should have a short eccentricity damping time and an average
     decay time compared to other hot Jupiters.  It seems to orbit a
@@
 that the time system be clearly documented.  Our final hot Jupiter
-sample included WASPs-4b, -5b, -6b, -18b, and -46b.  The collected and
+sample included WASP-4b, 5b, 6b, 18b, and 46b.  The collected and
 measured times are given in Tables~5, 6, 7, and 8 for each.
@@
 performed a least-squares fit to a linear ephemeris.  Using the
-best-fit values and variances, we calculated the uncertainty on the
-predicted transit time during the TESS observations.  This gave 9, 94,
-18, 42, and 60 seconds for WASPs-4b, -5b, -6b, -18b, and -46b.  We
-show in Figure~\ref{fig:hjs} that WASP-4b is the only hot Jupiter that
-transited significantly earlier than expected.
+best-fit values and variances, we calculated the precision of the
+predicted transit time during the TESS observing epoch.  This gave 11,
+94, 18, 42, and 59 seconds for WASP-4b, 5b, 6b, 18b, and 46b.
+Comparing the observed and predicted times, Figure~\ref{fig:hjs} shows
+that WASP-4b is the only hot Jupiter that transited significantly
+earlier than expected.
 
@@
 For each planet, we then ask: what fraction of the posterior for
-$t_{\rm offset}$ is consistent with an offset worse than $77$ seconds?
-For WASP-4b, the answer is unsurpringly about half.  For WASP-6b, the
-most constraining object, about 1 sample in 2 million is consistent
-with such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 63
-samples would be consistent with this timing offset ($2.1\sigma$), and
-in WASP-46b, the limit is 1 in 38 samples ($1.9\sigma$).  For WASP-5b,
-the predicted time is too imprecise to rule out timing offsets at the
-necessary amplitude.  Multiplying the three independent probabilities
-for WASPs-5b, 6b, and -18b, we rule out $t_{\rm offset} < -77\ {\rm
-seconds}$ at $6.3\sigma$, or about about 1 part in 5 billion.
+$t_{\rm offset}$ is consistent with an offset worse than $81.5$
+seconds?  For WASP-4b, the answer is unsurpringly about half.  For
+WASP-6b, the most constraining object, about 1 sample in 2 million is
+consistent with such a timing offset ($4.9\sigma$).  For WASP-18b, 1
+in 80 samples would be consistent with this timing offset
+($2.2\sigma$), and in WASP-46b, the limit is 1 in 44 samples
+($2.0\sigma$).  For WASP-5b, the predicted time is too imprecise to
+rule out timing offsets at the necessary amplitude.  Multiplying the
+three independent probabilities for WASP-5b, 6b, and 18b, we can rule
+out $t_{\rm offset} < -81.5\ {\rm seconds}$ at $6.3\sigma$, or about
+about 1 part in 8 billion.
 
@@
 % \citealt{huitson_gemini_2017}, the authors also stand by their
-% timestamps\todo[inline]{verify this}. Since the ETD data come from
+% timestamps. Since the ETD data come from
 % heterogeneous sources, their timestamps are less clearly documented,

### 2018-12-14 paper/ms.tex
typo fixes

@@
 observed in hot Jupiter systems, perhaps in part because their orbits
-are nearly circular.  However, there are many mechanisms that
-could pump hot Jupiter orbits away from circularity (see
+are nearly circular.  However, there are many mechanisms that could
+pump hot Jupiter orbits away from circularity (see
 \S~\ref{sec:apsidal_precession}, and
@@
     \caption{
-      %TODO: final number
-        {\bf TESS saw WASP-4b transit earlier than expected.}
-        Both plots show observed transit times minus the pre-TESS
-        ephemeris as a function of transit number.  Blue curves are
-        the $\pm 1\sigma$ credible interval around the constant-period
+        {\bf TESS saw WASP-4b transit earlier than expected.} Both
+        plots show observed transit times minus the pre-TESS ephemeris
+        as a function of transit number.  Blue curves are the $\pm
+        1\sigma$ credible interval around the constant-period
         prediction.  {\it Top:} The full timing dataset spans 10
         years.  Less precise points are more transparent.  {\it
-        Bottom:} The binned TESS observation (red point) arrived $77.8
-        \pm 10.8\ {\rm seconds}$ before it was expected.
+        Bottom:} The binned TESS observation (red band) arrived $81.5
+        \pm 11.8\ {\rm seconds}$ before it was expected.
         \label{fig:arrived_early}
@@
 
-Having defined the free and fixed parameters, %we calculated an initial
-%guess for the free parameters by maximizing the likelihood.  W
-we then
-sampled over the posterior using the algorithm proposed by
+Having defined the free and fixed parameters, we then sampled over the
+posterior using the algorithm proposed by
 \citet{goodman_ensemble_2010} and implemented by
-\citet{foreman-mackey_emcee_2013}.
-Table~1 gives the results, which are in reasonable agreement with the
-parameters reported by {\it e.g.},
-\citet{southworth_high-precision_2009} and 
+\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
+are in reasonable agreement with the parameters reported by {\it
+e.g.}, \citet{southworth_high-precision_2009} and
 \citet{huitson_gemini_2017}.
@@
   were unconstrained.
-  (B) \citep{triaud_spin-orbit_2010}.
+  (B) \citet{triaud_spin-orbit_2010}.
   (C) From HARPS spectra \citep{doyle_accurate_2013}.

### 2018-12-14 paper/ms.tex
grammar-hammer

@@
 at mid-exposure. They also analyzed other hot Jupiters, which all
-showed mid-transit times as expected from a constant ephemeris.}.
+showed mid-transit times as expected from a constant ephemeris.
+The two Spitzer transits at epochs 362 and 365 are in independent
+agreement with the Gemini times.}.
 Since the linear model provides a poor fit to the data, we discard it
@@
 14.0^{+4.9}_{-3.5}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
-period would be $26^{+9}_{-7}$ years.
+period is $26^{+9}_{-7}$ years.
 

### 2018-12-14 paper/ms.tex
grammar-hammer

@@
 planets.  Although roughly 20 hot Jupiters have shorter predicted
-orbital decay times than WASP-4b, most have short observing baselines.
-A separate consideration is that some hot Jupiters host stars ({\it
-e.g.}, WASP-18) are fully radiative, and may dissipate tidal energy
-less efficiently than stars with convective zones, like WASP-4.
+orbital decay times than WASP-4b, most have shorter observing
+baselines.  A separate consideration is that some hot Jupiter host
+stars ({\it e.g.}, WASP-18) are fully radiative, and may dissipate
+tidal energy less efficiently than stars with convective zones, like
+WASP-4.
 

### 2018-12-14 paper/ms.tex
grammar-hammer
paper: add phase-folded lightcurve figure

@@
 \section{New transits and system parameters}
+
 \begin{figure}[t]
@@
     \caption{
-        {\bf TESS transits of WASP-4b.} On the left, black points are
-        TESS flux measurements, with a vertical offset applied. Blue
-        curves are best-fit models. Rounded midtimes are printed in
-        BTJD next to the appropriate transits.  The residuals are
-        shown on the right.
+        {\bf Phase-folded transit of WASP-4b.} Black points are TESS
+        flux measurements; yellow points are binned measurements.
+        Using the fit to the phase-folded transit (blue line), we
+        measure mid-transit times by fitting the individual transits
+        (Figure~\ref{fig:lightcurves}).
+        \label{fig:phasefold}
+    }
+\end{figure}
+
+\begin{figure}[t]
+    \begin{center}
+        \includegraphics[width=0.48\textwidth]{f2.png}
+    \end{center}
+    \vspace{-0.5cm}
+    \caption{
+        {\bf Individual transits of WASP-4b.} On the left, black
+        points are TESS flux measurements, with a vertical offset
+        applied. Blue curves are best-fit models. Rounded midtimes are
+        printed in BTJD next to the appropriate transits.  The
+        residuals are shown on the right.
         \label{fig:lightcurves}
@@
         \leavevmode
-        \includegraphics[width=0.48\textwidth]{f2.pdf}
+        \includegraphics[width=0.48\textwidth]{f3.pdf}
     \end{center}
@@
 
-Having defined the free and fixed parameters, we then sampled over the
+Having defined the free and fixed parameters, we sampled over the
 posterior using the algorithm proposed by
@@
 \citet{huitson_gemini_2017}.
+Figure~\ref{fig:phasefold} shows the phase-folded lightcurve.
 
@@
 
-Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
-and residuals.  Table~2 reports the mid-transit times and their
-uncertainties.
+Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
+best-fit models, and the residuals.  Table~2 reports the mid-transit
+times and their uncertainties.
 Binning the residuals to 1-hour windows and taking the median absolute
@@
         \leavevmode
-        \includegraphics[width=0.9\textwidth]{f3.pdf}
+        \includegraphics[width=0.9\textwidth]{f4.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.417\textwidth]{f4.pdf}
+    \includegraphics[width=0.417\textwidth]{f5.pdf}
   \end{center}
@@
 If the planet's spin and orbit are synchronized and the star is slowly
-rotating, then the semi-major axis and eccentricity evolve as
+rotating, the semi-major axis and eccentricity then evolve as
 \citep[Appendix B of][]{metzger_optical_2012}
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f5.pdf}
+		\includegraphics[width=0.48\textwidth]{f6.pdf}
 	\end{center}
@@
     \leavevmode
-    \includegraphics[width=0.9\textwidth]{f6.pdf}
+    \includegraphics[width=0.9\textwidth]{f7.pdf}
   \end{center}

### 2018-12-14 src/plot_phasefold.py
add src/plot_phasefold.py

@@ -1,0 +1,119 @@
+# -*- coding: utf-8 -*-
+'''
+make phase-folded lightcurve figure
+'''
+from __future__ import division, print_function
+
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt, pandas as pd, numpy as np
+
+from glob import glob
+import os, pickle
+from shutil import copyfile
+
+from astrobase.lcmath import phase_magseries
+from astrobase.lcmath import phase_bin_magseries
+
+def plot_phasefold(picklepath, xlim=None, ylim0=None, ylim1=None):
+
+    # d['magseries'].keys()
+    # dict_keys(['times', 'mags', 'errs', 'magsarefluxes'])
+    # d['fitinfo'].keys()
+    # dict_keys(['initialparams', 'initialmags', 'fixedparams',
+    #            'finalparams', 'finalparamerrs', 'fitmags', 'fitepoch'])
+    # d['fitinfo']['finalparams'].keys()
+    # dict_keys(['incl', 'period', 'rp', 'sma', 't0', 'u_linear', 'u_quad'])
+    d = pickle.load(open(picklepath, 'rb'))
+
+    times = d['magseries']['times']
+    fluxs = d['magseries']['mags']
+    fitfluxs = d['fitinfo']['fitmags']
+
+    fit_t0 = d['fitinfo']['finalparams']['t0']
+    # NOTE: might want to use the best-fit value from the tables, instead of
+    # the BLS period here.
+    fit_period = d['fitinfo']['finalparams']['period']
+
+    phzd = phase_magseries(times, fluxs, fit_period, fit_t0,
+                           wrap=True, sort=True)
+    fit_phzd = phase_magseries(times, fitfluxs, fit_period, fit_t0, wrap=True,
+                               sort=True)
+
+    phase = phzd['phase']
+    phz_flux = phzd['mags']
+    fit_phz_flux = fit_phzd['mags']
+
+    binsize = 0.003
+    bin_phzd = phase_bin_magseries(phase, phz_flux, binsize=binsize)
+    bin_phase = bin_phzd['binnedphases']
+    bin_fluxs = bin_phzd['binnedmags']
+    fit_bin_phzd = phase_bin_magseries(phase, fit_phz_flux, binsize=binsize)
+    fit_bin_phase = fit_bin_phzd['binnedphases']
+    fit_bin_fluxs = fit_bin_phzd['binnedmags']
+
[…]

### 2018-12-14 src/make_all.sh
update fig numbers

@@ -10,6 +10,8 @@ python make_select_transit_times_table.py
 
 echo "making plots"
 
+python plot_phasefold.py
+
 python plot_stacked_lightcurves.py
 
 python plot_arrived_early.py

### 2018-12-14 src/plot_O_minus_C.py

@@ -476,15 +476,15 @@ def main(plname, xlim=None, ylim=None, include_all_points=False, ylim1=None):
         x_occ=x_occ, y_occ=y_occ, sigma_y_occ=sigma_y_occ, ylim1=ylim1)
 
     if not include_all_points:
-        copyfile(savpath, '../paper/f3.png')
-        print('saved ../paper/f3.png')
-        copyfile(savpath.replace('.png','.pdf'), '../paper/f3.pdf')
-        print('saved ../paper/f3.pdf')
+        copyfile(savpath, '../paper/f4.png')
+        print('saved ../paper/f4.png')
+        copyfile(savpath.replace('.png','.pdf'), '../paper/f4.pdf')
+        print('saved ../paper/f4.pdf')
     else:
-        copyfile(savpath, '../paper/f7.png')
-        print('saved ../paper/f7.png')
-        copyfile(savpath.replace('.png','.pdf'), '../paper/f7.pdf')
-        print('saved ../paper/f7.pdf')
+        copyfile(savpath, '../paper/f8.png')
+        print('saved ../paper/f8.png')
+        copyfile(savpath.replace('.png','.pdf'), '../paper/f8.pdf')
+        print('saved ../paper/f8.pdf')
 
 
 if __name__=="__main__":

### 2018-12-14 src/plot_arrived_early.py

@@ -199,8 +199,8 @@ if __name__=="__main__":
     plot_arrived_early(plname, xlim=xlim, ylim=ylim, savpath=savpath,
                        ylim1=ylim1)
 
-    copyfile(savpath, '../paper/f2.png')
-    print('saved ../paper/f2.png')
-    copyfile(savpath.replace('.png','.pdf'), '../paper/f2.pdf')
-    print('saved ../paper/f2.pdf')
+    copyfile(savpath, '../paper/f3.png')
+    print('saved ../paper/f3.png')
+    copyfile(savpath.replace('.png','.pdf'), '../paper/f3.pdf')
+    print('saved ../paper/f3.pdf')
 

### 2018-12-14 src/plot_future.py

@@ -267,10 +267,10 @@ def main(plname, xlim=None, ylim=None, savname=None, ylim1=None):
         xlim=xlim, ylim=ylim, ylim1=ylim1)
 
     if savpath == os.path.join(pkldir, 'future.png'):
-        copyfile(savpath, '../paper/f5.png')
-        print('saved ../paper/f5.png')
-        copyfile(savpath.replace('.png','.pdf'), '../paper/f5.pdf')
-        print('saved ../paper/f5.pdf')
+        copyfile(savpath, '../paper/f6.png')
+        print('saved ../paper/f6.png')
+        copyfile(savpath.replace('.png','.pdf'), '../paper/f6.pdf')
+        print('saved ../paper/f6.pdf')
 
 
 if __name__=="__main__":

### 2018-12-14 src/plot_hjs.py

@@ -236,8 +236,8 @@ def plot_hjs():
 
     fig.tight_layout(h_pad=0, w_pad=0)
 
-    for savname in ['../results/hjs.png', '../paper/f6.png',
-                    '../results/hjs.pdf', '../paper/f6.pdf']:
+    for savname in ['../results/hjs.png', '../paper/f7.png',
+                    '../results/hjs.pdf', '../paper/f7.pdf']:
        fig.savefig(savname, dpi=400, bbox_inches='tight')
        print('saved {:s}'.format(savname))
 

### 2018-12-14 src/plot_population_vs_wasp4.py

@@ -405,10 +405,10 @@ def plot_all():
 
     fig.savefig(resultspng, dpi=500, bbox_inches='tight')
     print('saved {:s}'.format(resultspng))
-    copyfile(resultspng, '../paper/f4.png')
-    print('saved ../paper/f4.png')
-    fig.savefig('../paper/f4.pdf', bbox_inches='tight')
-    print('saved ../paper/f4.pdf')
+    copyfile(resultspng, '../paper/f5.png')
+    print('saved ../paper/f5.png')
+    fig.savefig('../paper/f5.pdf', bbox_inches='tight')
+    print('saved ../paper/f5.pdf')
 
 
 if __name__=="__main__":

### 2018-12-14 src/plot_stacked_lightcurves.py

@@ -108,8 +108,8 @@ def plot_stacked_lightcurves(
 
     f.tight_layout(h_pad=0, w_pad=0.3)
     resultspng = '../results/{:d}_stacked_lightcurves.png'.format(ticid)
-    paperpng = '../paper/f1.png'
-    paperpdf = '../paper/f1.pdf'
+    paperpng = '../paper/f2.png'
+    paperpdf = '../paper/f2.pdf'
     f.savefig(resultspng, dpi=400, bbox_inches='tight')
     print('saved {:s}'.format(resultspng))
     f.savefig(paperpng, dpi=400, bbox_inches='tight')
@@ -202,8 +202,8 @@ def plot_stacked_lightcurves_single_col(
 
     f.tight_layout(h_pad=0, w_pad=0)
     resultspng = '../results/{:d}_stacked_lightcurves.png'.format(ticid)
-    paperpng = '../paper/f1.png'
-    paperpdf = '../paper/f1.pdf'
+    paperpng = '../paper/f2.png'
+    paperpdf = '../paper/f2.pdf'
     f.savefig(resultspng, dpi=400, bbox_inches='tight')
     print('saved {:s}'.format(resultspng))
     f.savefig(paperpng, dpi=400, bbox_inches='tight')
@@ -228,7 +228,7 @@ if __name__=="__main__":
 
     pickledir = (
         '/home/luke/Dropbox/proj/tessorbitaldecay/results/'+
-        'tess_lightcurve_fit_parameters/{:d}'.format(ticid)
+        'tess_lightcurve_fit_parameters/{:d}/sector_2/'.format(ticid)
     )
 
     plot_stacked_lightcurves(

### 2018-12-14 src/plot_phasefold.py
plot tweaking

@@ -55,7 +55,7 @@ def plot_phasefold(picklepath, xlim=None, ylim0=None, ylim1=None):
     ##########################################
     plt.close('all')
     f, (a0, a1) = plt.subplots(nrows=2, ncols=1, sharex=True,
-                               figsize=(6,4), gridspec_kw=
+                               figsize=(0.8*6,0.8*4), gridspec_kw=
                                {'height_ratios':[3, 1]})
 
     a0.scatter(phase*fit_period*24, phz_flux, c='k', alpha=0.15, label='data',

### 2018-12-14 paper/ms.tex
grammar-hammer

@@
 \citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
-Jupiters satisfy this condition; their orbits should be shrinking
-\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay
-is uncertain, and depends on how friction dissipates energy carried by
+Jupiters satisfy this condition, so their orbits should be shrinking
+\citep{levrard_falling_2009,matsumura_tidal_2010}.  However the rate of decay
+is uncertain, as it depends on how friction dissipates energy carried by
 gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
 we could directly measure orbital decay rates, it would inform our
-understanding of tidal dissipation.  This is important because the
+understanding of tidal dissipation.  This would be important because the
 dissipation rates determine the fate of the system.  During its
@@
 surface, the drag force from the stellar envelope accelerates the
-inspiral, leading to a `direct-impact' merger
+inspiral, causing a `direct-impact' merger
 \citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
-the planet fills its Roche lobe above above the stellar surface, then
+the planet fills its Roche lobe above the stellar surface, then
 it could either be disrupted into an accretion disk or it could slowly
@@
 eclipsing binaries \citep[{\it e.g.},][]{
-  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}
-and has been detected in a handful of exoplanet systems, including
-circumbinary planets (Kepler-413 \citealt{kostov_kepler-413b_2014};
-Kepler-453 \citealt{welsh_kepler_2015}, planets orbiting rapidly
-rotating stars (Kepler-13
+  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
+In exoplanetary systems, precessing orbits have been detected in a
+handful of cases, including circumbinary systems (Kepler-413
+\citealt{kostov_kepler-413b_2014}; Kepler-453
+\citealt{welsh_kepler_2015}, planets orbiting rapidly rotating stars
+(Kepler-13
 \citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
@@
 precession is usually dominated by the quadrupolar distortion of the
-planet by the star's tidal force \citep{ragozzine_probing_2009}.  The
-observed precession rate in a hot Jupiter system could thus be used to
-compute the planet's Love number, giving a third parameter along with
-the mass and radius with which to describe the planet's interior
-structure \citep[{\it e.g.},][who performed a similar procedure for
+planet by the star's tidal force \citep{ragozzine_probing_2009}.  
+If apsidal precession were observed in a hot Jupiter system, the
+precession rate could thus be used to compute the planet's Love
+number, giving a third parameter along with the mass and radius with
+which to describe the planet's interior structure \citep[{\it
+e.g.},][who performed a similar procedure for
 HAT-P-13b]{batygin_determination_2009}.  Measuring small
@@
 
+\clearpage
+
 \bibliographystyle{yahapj}                            

### 2018-12-16 paper/ms.tex
update author list
merged JNW comments

@@
 
-\NewPageAfterKeywords
+%\NewPageAfterKeywords
 
@@
 %
+\author[0000-0003-3438-843X]{C. Baxter}
+\affiliation{ API, University of Amsterdam, P.O. Box 94249, 1090 GE
+    Amsterdam, The Netherlands}
+%
+\author[0000-0002-0628-0088]{W. Bhatti}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%
+\author[0000-0002-6939-9211]{T. Daylan}
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+%
 \author[0000-0002-0875-8401]{J.-M. D\'esert}
@@
 Amsterdam, The Netherlands}
+%
+\author[0000-0002-5720-4827]{M. Hill}
+\affiliation{Department of Earth Sciences, University of California,
+Riverside, CA 92521, USA}
+%
+\author[0000-0002-7084-0529]{S. Kane}
+\affiliation{Department of Earth Sciences, University of California,
+Riverside, CA 92521, USA}
 %
@@
 % 
-\author[0000-0002-6939-9211]{T. Daylan}
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0002-5720-4827]{M. Hill}
-\affiliation{Department of Earth Sciences, University of California,
-Riverside, CA 92521, USA}
-%
-\author[0000-0002-7084-0529]{S. Kane}
-\affiliation{Department of Earth Sciences, University of California,
-Riverside, CA 92521, USA}
-% 
 %-------------------------------------
@@
 \author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
-\affiliation{ Harvard-Smithsonian Center for Astrophysics, 60 Garden
+\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
 Street, Cambridge, MA 02138, USA }
@@
 \begin{abstract}
-  By combining data from the Transiting Exoplanet Survey Satellite
-  (TESS) with data from previous studies, we show that the transit
-  times of WASP-4b are incompatible with a constant orbital period.
-  The transits observed by TESS occurred $81.5 \pm 11.8$~seconds
-  earlier than one would expect, based on the previous data.  The
-  orbital period appears to be shrinking by $\dot{P}=-12.6 \pm 1.2$
-  milliseconds per year.   A simple clock offset is unlikely, because
-  TESS observations of other hot Jupiters (WASP-6b, 18b, and 46b) are
-  compatible with a constant period and rule out a 81.5-second offset
-  at the 6.3$\sigma$ level.  The timing variations seem to be
-  astrophysical, and could be caused by apsidal precession, or perhaps
-  by tidal decay.  The latter seems implausible on theoretical
-  grounds, as it would require a stellar quality factor $Q_\star'
-  \approx 3\times10^4$.  Apsidal precession would imply an
-  eccentricity $e\approx1.8\times 10^{-3}$, and the possibility of
-  measuring the planet's Love number.  A Doppler shift from WASP-4's
-  acceleration towards us could account for at most one quarter of the
-  observed period change (at $2\sigma$).  Further observations are
-  needed to confirm and understand the timing variation.
+    The Transiting Exoplanet Survey Satellite (TESS) recently observed
+    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
+    occurred $81.5 \pm 11.8$~seconds earlier than had been predicted,
+    based on earlier data stretching back to 2007.  This is unlikely
+    to be the result of a clock error, because TESS observations of
+    other hot Jupiters (WASP-6b, 18b, and 46b) are compatible with a
+    constant period, ruling out an 81.5-second offset at the
+    6.3$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
+    to be decreasing at a rate of $\dot{P} = -12.6 \pm 1.2$
+    milliseconds per year.  This might be caused by tidal orbital
+    decay or apsidal precession, although both interpretations have
+    shortcomings.  The gravitational influence of a third body is another
+    possibility, although at present there is scant evidence for such a
+    body. Further observations are needed to confirm and
+    understand the timing variation.
 \end{abstract}
@@
 
-Long-term monitoring of hot Jupiter transit and occultation times
-should eventually reveal two distinct processes.  First, whenever the
-summed spin and orbital angular momenta in a hot Jupiter system fall
-below a critical value, the system becomes unstable to tidal decay
-\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
-Jupiters satisfy this condition, so their orbits should be shrinking
-\citep{levrard_falling_2009,matsumura_tidal_2010}.  However the rate of decay
-is uncertain, as it depends on how friction dissipates energy carried by
-gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
-we could directly measure orbital decay rates, it would inform our
-understanding of tidal dissipation.  This would be important because the
-dissipation rates determine the fate of the system.  During its
-inspiral, if the planet fills its Roche lobe below the stellar
-surface, the drag force from the stellar envelope accelerates the
-inspiral, causing a `direct-impact' merger
-\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
-the planet fills its Roche lobe above the stellar surface, then
-it could either be disrupted into an accretion disk or it could slowly
-transfer mass to the star
-\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
-Further afield, the outcome of the eventual interaction between the
-Earth and red giant Sun also hinges on the efficiency of tidal
-dissipation \citep{rasio_tidal_1996}.
-
-The second process that long-term timing studies should reveal is
-apsidal precession, {\it i.e.}, rotation of the orbital ellipse within
-the orbital plane. Apsidal precession has been observed extensively in
-eclipsing binaries \citep[{\it e.g.},][]{
+Although the Transiting Exoplanet Survey Satellite (TESS) is designed
+to detect new planets, it will also perform precise photometric
+monitoring of most of the planets that have been discovered by
+ground-based transit surveys over the last two decades.  One of the
+applications of the new TESS data is to search for timing anomalies in
+hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
+occultation times should eventually reveal two distinct processes.
+
+First, whenever the orbital angular momentum is less than 3 times the
+rotational angular momentum, the system is vulnerable to tidal orbital
+decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
+Jupiters satisfy this condition and their orbits should be shrinking
+\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay,
+which depends on how friction dissipates the energy of tidal
+disturbances, is difficult to observe directly or calculate from first
+principles. This has been a stubborn source of uncertainty in the
+interpretation of many phenomena in stellar and planetary astrophysics
+(as reviewed by \citealt{Mazeh2008} and \citealt{ogilvie_tidal_2014}).
+
+Second, long-term timing studies may reveal apsidal precession, {\it
+  i.e.}, the rotation of the orbital ellipse within the orbital
+plane. Apsidal precession has long been observed in eclipsing binaries
+\citep[{\it e.g.},][]{
   schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
-In exoplanetary systems, precessing orbits have been detected in a
-handful of cases, including circumbinary systems (Kepler-413
+For exoplanets, precessing orbits have been detected in a handful of
+cases, including circumbinary systems (Kepler-413
 \citealt{kostov_kepler-413b_2014}; Kepler-453
@@
 \citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
-PTFO 8-8695 \citealt{barnes_measurement_2013}), and systems with
-multiple interacting giant planets (Kepler-108
+and systems with multiple interacting giant planets (Kepler-108
 \citealt{mills_kepler-108_2017}; Kepler-693
 \citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
-observed in hot Jupiter systems, perhaps in part because their orbits
-are nearly circular.  However, there are many mechanisms that could
-pump hot Jupiter orbits away from circularity (see
-\S~\ref{sec:apsidal_precession}, and
-\citealt{bailey_understanding_2019}), which would then lead to
-observable timing variations.  For hot Jupiters, the apsidal
-precession is usually dominated by the quadrupolar distortion of the
-planet by the star's tidal force \citep{ragozzine_probing_2009}.  
-If apsidal precession were observed in a hot Jupiter system, the
-precession rate could thus be used to compute the planet's Love
-number, giving a third parameter along with the mass and radius with
-which to describe the planet's interior structure \citep[{\it
-e.g.},][who performed a similar procedure for
-HAT-P-13b]{batygin_determination_2009}.  Measuring small
-eccentricities and understanding their origin might also help us
-understand the dynamical histories and inflated radii of hot Jupiters.
-\citep[{\it
-e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}
-
-While numerous indirect studies have called attention to
-population-level effects of tidal decay
-\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
-the most convincing direct evidence to date for either orbital decay
+observed in hot Jupiter systems because their orbits are nearly
+circular, which may itself be a consequence of star-planet tides.  If
+it could be observed, then \citep{ragozzine_probing_2009} showed that
+it would be possible to measure the planet's Love number, which would
+be a valuable constraint on the planet's interior.
+
+So far, the most convincing direct evidence for either orbital decay
 or apsidal precession has come from WASP-12b.
-\citet{maciejewski_departure_2016} showed that the transit times of
-WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
-and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.  Using
-new timing data, \citet{patra_2017} found $\dot{P} = -29 \pm 3\,{\rm
-ms}\,{\rm yr}^{-1}$.  If the period change were caused entirely by
-tidal decay, the implied modified stellar quality factor would be
-$Q_\star' \approx 2\times10^5$.  Standard equilibrium tides do not
-predict this level of dissipation
-\citep{penev_tidal_2011,ogilvie_tidal_2014}.  The needed level of
-friction might come from dynamical tides, or obliquity tides
-\citep{weinberg_tidal_2017,millholland_obliquity_2018}.  The former is
-more likely if WASP-12 is a subgiant, but the observational
-constraints favor a main-sequence model
-\citep{bailey_understanding_2019}.  More data are needed to understand
-the WASP-12b timing variations.
-
-This work focuses on the hot Jupiter WASP-4b, which orbits its host
-star every 1.34 days, at a distance of $a/R_\star = 5.46$
-\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  WASP-4b is a plausible
-candidate for orbital decay, and has been
-observed often over the past decade.  The host is a G7V dwarf, and is on the
-main sequence (\S~\ref{sec:system_parameters}).  The orbit is
-consistent with being circular (but see
-\S~\ref{sec:apsidal_precession}), and having sky-projected obliquity
-of zero
+\citet{maciejewski_departure_2016} and \citet{patra_2017} showed that
+the transit period of WASP-12b has been steadily decreasing by about
+30~milliseconds per year over the last decade.  Here, we present
+evidence for a timing anomaly in the WASP-4 system.  The hot Jupiter
+WASP-4b orbits a G7V star every 1.34 days, corresponding to an orbital
+distance of 5.5 stellar radii
+\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  It is a good target to
+search for departures from a constant period, because transit data
+exists all the way back to 2007.  The orbital eccentricity is less
+than 0.018 (2$\sigma$), based on the work of
+\cite{knutson_friends_2014}, who combined the available transit times,
+occultation times, and Doppler data.  The stellar obliquity (or at
+least its sky projection) is also compatible with zero, within about
+10 degrees
 \citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
 
-We measure new transit times for WASP-4b using TESS lightcurves
-(\S~\ref{sec:observations},~\ref{sec:measurement}), and calculate the
-system's parameters empirically (\S~\ref{sec:system_parameters}).  We
-merge the new times with previous observations and compare three
-models (\S~\ref{sec:timing}): a constant period, a decaying period,
-and a slightly eccentric orbit undergoing precession.  We rule out a
-constant period, but cannot firmly distinguish between the latter two
-possibilities.  Each would have interesting implications
-(\S~\ref{sec:implications}).  We rule out the possibility that a
-systematic time system offset causes the observed variation
-(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
-further monitoring of the system (\S~\ref{sec:future}).
+In what follows, \S~\ref{sec:observations} presents the new TESS
+observations, and \S~\ref{sec:timing} describes our timing
+analysis.  We tried fitting the data with three models: a constant
+period, a decaying period, and a slightly eccentric precessing orbit.
+A constant period can be ruled out, but we cannot distinguish between
+the latter two possibilities.  Either possibility would have
+interesting implications, described in \S~\ref{sec:implications}.
+Appendix~\ref{sec:verify_tess} considers the possibility that the
+WASP-4 timing anomaly is due to an error in time stamps in the TESS
+data products.  We found this possibility to be unlikely because none
+of the other hot Jupiters we examined show a timing offset with the
+same amplitude as was seen for WASP-4.
 
 \section{New transits and system parameters}
-
-\begin{figure}[t]
+\label{sec:observations}
+
+\begin{figure}[t!]
     \begin{center}
-        \includegraphics[width=0.48\textwidth]{f1.png}
+        \includegraphics[width=0.49\textwidth]{f2.png}
     \end{center}
@@
     \caption{
-        {\bf Phase-folded transit of WASP-4b.} Black points are TESS
-        flux measurements; yellow points are binned measurements.
-        Using the fit to the phase-folded transit (blue line), we
-        measure mid-transit times by fitting the individual transits
-        (Figure~\ref{fig:lightcurves}).
+      {\bf TESS observations of WASP-4b.} On the left, black points
+      are TESS flux measurements, with a vertical offset applied. Blue
+      curves are best-fit models. The numbers printed next to each
+      lightcurve are the approximate transit times expressed in
+      BJD minus 2{,}457{,}000.  The right side shows the
+      residuals.
+       \label{fig:lightcurves}
+    }
+\end{figure}
+
+\begin{figure}[t!]
+    \begin{center}
+        \includegraphics[width=0.49\textwidth]{f1.png}
+    \end{center}
+    \vspace{-0.5cm}
+    \caption{
+        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
+        TESS data. Yellow points are binned measurements.  The bottom
+        panel shows the residuals.  The fit to the phase-folded
+        transit (blue line) is used when measuring mid-transit times for
+        the individual transits (Figure~\ref{fig:lightcurves}).
         \label{fig:phasefold}
@@
 
-\begin{figure}[t]
-    \begin{center}
-        \includegraphics[width=0.48\textwidth]{f2.png}
-    \end{center}
-    \vspace{-0.5cm}
-    \caption{
-        {\bf Individual transits of WASP-4b.} On the left, black
-        points are TESS flux measurements, with a vertical offset
-        applied. Blue curves are best-fit models. Rounded midtimes are
-        printed in BTJD next to the appropriate transits.  The
-        residuals are shown on the right.
-        \label{fig:lightcurves}
-    }
-\end{figure}
-
 \subsection{Observations}
-\label{sec:observations}
-
-WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
-20, during the spacecraft's second sector of science operations.  The
-star is designated by the TESS Input Catalog as TIC 402026209
-\citep{stassun_TIC_2018}.
-
-The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
-array every 2 minutes.   The data were downlinked via the Deep Space
-Network. The timestamps from the spacecraft clock were then converted
-by the Payload Operations Center to TESS Julian Day in the {\it
-Temps Dynamique Barycentrique} (TDB) reference.  The images were then
-reduced to lightcurves by the Science Processing Operations Center (SPOC)
-at NASA Ames~\citep{jenkins_tess_2016}.  During this processing, SPOC used
-the known
-spacecraft trajectory to compute the barycentric time corrections on a
-target by target basis, and reported timestamps in ``BTJD'', where
-${\rm BTJD} = {\rm BJD} - 2457000$, BJD is the barycentric julian
-date, and all times are in the TDB reference (\citealt{urban_explanatory_2012} 
-describe the relevant time systems).
-The lightcurves were then vetted and released by the MIT TESS Science
-Office to the Mikulski Archive for Space
-Telescopes~\citep{ricker_tess_alerts_2018}.
-
-We began our analysis with the Presearch Data Conditioning (PDC) light
-curves from MAST.  The steps used to find an optimal aperture (for
-WASP-4, a $3\times 3$ pixel array) and the PDC-MAP algorithm are
-described by \citet{smith_kepler_apertures_2017} and
-\citet{smith_kepler_PDC_2017} respectively.
-
-We processed the lightcurve as follows.  First, we removed all
-points with non-zero quality flags.  This removed data contaminated by
-coarse spacecraft pointing, cosmic rays, and other annoyances
-\citep{tess_data_product_description_2018}.  We then filtered out
-known bad observing windows.  The data closest to spacecraft perigee
-typically show ramp-like systematics, so we clipped out the first and
-last hour of both orbits.  Next, we focused on the momentum dumps. As
-described in the data release
-notes\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}},
-during the first two TESS sectors, every 2.5~days the spacecraft's
-reaction wheels were reset to maintain pointing stability.  These
-events were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
-were simultaneously set for 54 distinct cadences, and there were 10
-momentum dumps, averaging about 10 minutes of flagged data per dump.
-Simply out of caution, we clipped out an additional 10 minutes before
-and after every momentum dump.  Before applying any filters, we had
-19737 data points. After applying all the filters, we were left with
-18165 measurements, or 92\% of the original data.
-
-After ``cleaning'' the data, we normalized the flux measurements by
-dividing out the median flux.  We converted the timestamps from BTJD
-to BJD by adding the appropriate 2457000 day offset
-\citep{tess_data_product_description_2018}.  Many of these and
+
+WASP-4 was observed by TESS with Camera 2 from August 23 to September
+20, 2018, within the second ``sector'' of science operations.  The
+star is designated as TIC 402026209 in the TESS Input Catalog
+\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
+surrounding WASP-4 were averaged into 2-minute stacks by the onboard
+computer.  The data were downlinked via the Deep Space Network. The
+timestamps from the spacecraft clock were then transformed by the
+Payload Operations Center into the {\it Temps Dynamique Barycentrique}
+(TDB) reference system.  The images were then reduced to lightcurves
+by the Science Processing Operations Center (SPOC) at NASA
+Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
+the known spacecraft trajectory to compute the barycentric time
+corrections on a target by target basis, and expressed the timestamps
+as Barycentric TESS Julian Dates (BTJD), which is simply the
+barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
+vetted and released by the MIT TESS Science Office to the Mikulski
+Archive for Space Telescopes on November 29,
+2018~\citep{ricker_tess_alerts_2018}.
+
+We began our analysis with the Presearch Data Conditioning (PDC)
+lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
+We processed the lightcurve as follows.  First, we removed all points
+with non-zero quality flags.  This removed data contaminated by coarse
+spacecraft pointing, cosmic rays, and other annoyances
+\citep{tess_data_product_description_2018}.  We also removed the data
+within the first and last hours of both orbits, because of ramp-like
+systematic effects that appear during those time ranges. We also
+removed the data that might have been adversely affected by ``momentum
+dumps,'' the firing of thrusters and resetting of reaction wheels that
+took place every 2.5 days during sector
+2.\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}.
+The data during these events were assigned quality flags corresponding
+to ``Reaction Wheel Desaturation Event'' and ``Manual Exclude''.  For
+WASP-4, these flags were simultaneously set for 54 distinct cadences,
+and there were 10 momentum dumps, averaging about 10 minutes of
+flagged data per dump.  Out of caution, we clipped out an additional
+10 minutes before and after every momentum dump.
+
+All told, we removed 8\% of the original data points, and were left
+with 18{,}165 measurements of the relative flux of WASP-4.  We
+normalized the data by dividing out the median flux.  We converted the
+timestamps from BTJD into BJD by adding the appropriate 2457000 day
+offset \citep{tess_data_product_description_2018}.  Many of these and
 subsequent processing steps were performed using
-\texttt{astrobase}~\citep{bhatti_astrobase_2018} We did not
-``flatten'' the lightcurves, as is often done with {\it
-e.g.}, a spline or gaussian process regression.  Such procedures are
-asymmetric about the transit minimum and can skew the midtime
-measurement without being captured by the measurement uncertainties.
+\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
+``flatten'' the lightcurves, as is often done with splines,
+polynomials, or Gaussian processes.  Instead, we modeled the
+out-of-transit flux variations simultaneously with the transit parameters,
+as described below.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
     \begin{center}
         \leavevmode
-        \includegraphics[width=0.48\textwidth]{f3.pdf}
+        \includegraphics[width=0.49\textwidth]{f3.pdf}
     \end{center}
     \vspace{-0.6cm}
-    \caption{
-        {\bf TESS saw WASP-4b transit earlier than expected.} Both
-        plots show observed transit times minus the pre-TESS ephemeris
-        as a function of transit number.  Blue curves are the $\pm
-        1\sigma$ credible interval around the constant-period
-        prediction.  {\it Top:} The full timing dataset spans 10
-        years.  Less precise points are more transparent.  {\it
-        Bottom:} The binned TESS observation (red band) arrived $81.5
-        \pm 11.8\ {\rm seconds}$ before it was expected.
+    \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
+      Both plots show the deviations between the observed and
+      calculated transit times, where the calculation is based only on
+      the pre-TESS data and assumes a constant period.  The blue bands
+      depict the $\pm$$1\sigma$ credible interval of the predicted
+      times.  {\it Top:} The full timing dataset spans 10 years. The
+      darkest points correspond to the most precise data.  {\it
+        Bottom:} Close-up of the TESS observations. The red band shows
+      the average deviation of the TESS data, which is $81.5 \pm
+      11.8\ {\rm seconds}$.
         \label{fig:arrived_early}
@@
 
-Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
-estimate the orbital period, transit duration and a reference epoch
-\citep{kovacs_box-fitting_2002}.  From these parameters, we isolated
-each transit to within $\pm$4 transit durations of its midtime.  We
-then fit the phase-folded transit, using the analytic formulae
-calculated by \citet{mandel_analytic_2002} and implemented by
-\citet{kreidberg_batman_2015}.  Specifically, we fit for a reference
-epoch, the planet to star radius ratio $R_{\rm p}/R_\star$, the
-semi-major axis to stellar radius ratio $a/R_\star$, the inclination
-$i$, two quadratic limb-darkening coefficients $(u_{\rm linear},
-u_{\rm quad})$, and the orbital period $P$.  We set the eccentricity
-to zero.
-
-Having defined the free and fixed parameters, we sampled over the
-posterior using the algorithm proposed by
+Using the cleaned PDC lightcurve, we used the Box Least Squares
+algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
+period, transit duration, and a reference epoch using the TESS data
+alone.  Based on the results, we isolated the data within 4 transit
+durations of each transit midpoint.  We then fitted all of the transit
+data using the analytic formulae given by \citet{mandel_analytic_2002}
+and implemented by \citet{kreidberg_batman_2015}.  We assumed the
+orbit to be circular. The free parameters were the reference epoch,
+the planet to star radius ratio $R_{\rm p}/R_\star$, the orbital
+distance to stellar radius ratio $a/R_\star$, the inclination $i$, two
+quadratic limb-darkening coefficients $(u_{\rm linear}, u_{\rm
+  quad})$, and the orbital period $P$.
+% JNW: Luke, didn't you also have to fit a linear trend for each transit?
+% Otherwise the fit would be poor, for each transit, even if the
+% trends tend to cancel out in the phase folded light curve.
+% And, did you assign the photometric uncertainties as described below?
+
+We sampled the posterior probability distribution for all the
+parameters using the algorithm proposed by
 \citet{goodman_ensemble_2010} and implemented by
@@
   e.g.}, \citet{southworth_high-precision_2009} and
-\citet{huitson_gemini_2017}.
-Figure~\ref{fig:phasefold} shows the phase-folded lightcurve.
-
-To measure the transit midtimes, we then fit  each individual transit
-for a model transit, plus a local linear trend.  We allowed four free 
-parameters per transit: the time of
-mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
-slope and intercept of the line.  We fixed the remaining
-parameters%~---~$a/R_\star$, $i$, $P$, and the limb-darkening
-%coefficients~---~
-to the best-fit values determined from the
-phase-folded TESS lightcurve.
-
-After completing an initial fit for each transit, we set the
-photometric error bars to equal the standard deviation of the
-out-of-transit points of the residual lightcurve.  We then repeated
-the fitting procedure, using the empirically determined photometric
-errors.
+\citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
+phase-folded lightcurve.
+
+To measure the transit times, we fitted each transit separately, with
+four free parameters: the time of mid-transit $t_{\rm tra}$, the
+planet-to-star radius ratio, and the slope and intercept of a linear
+trend to account for any slow variations unrelated to the transit.  We
+fixed the remaining parameters at the values that had been determined
+from the phase-folded TESS lightcurve.  The uncertainty in each
+data point was set equal to the root-mean-square (rms) level of the
+out-of-transit data.
 
@@
 the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
-transit times.  We found that $\chi^2 = 9.2$, with 16 degrees of
-freedom.  Visually inspecting residuals showed that the error variance
-had been overestimated, so we multiplied the measured TESS errors by a
-factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered
-the mean uncertainty of the transit midtimes from $29.8$ to $22.6$
-seconds.
+transit times.  We found that $\chi^2 = 9.2$, with $n=16$ degrees of
+freedom.  The variance of the $\chi^2$ distribution is $2n$, so we
+would expect $\chi^2 16 \pm 5.7$.  Visually inspecting the residuals
+showed that the error variance had been overestimated, so we
+multiplied the measured TESS errors by a factor $f=0.76$, forcing a
+reduced $\chi^2$ of unity.  This lowered the mean uncertainty of the
+transit midtimes from $29.8$ to $22.6$ seconds.
 
@@
 best-fit models, and the residuals.  Table~2 reports the mid-transit
-times and their uncertainties.
-Binning the residuals to 1-hour windows and taking the median absolute
-deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
-hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
-photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  For a
-$T\approx11.8$ star with a 9-pixel aperture, read noise is expected to
-contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
-quadrature.  The predicted zodiacal background contribution is
-$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
-\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
-in planet detection simulations.  This zodiacal noise prediction seems
-to have been an overestimate, given the observed RMS of WASP-4.  A
-broader assessment of TESS's photometric performance is beyond
-the scope of this work.
+times and their uncertainties.  After binning the residuals to 1-hour
+windows, the lightcurves have an rms scatter of 586\,{\rm ppm}.  The
+pre-launch TESS noise model \citep{winn_photonflux_2013,Sullivan_2015}
+\footnote{\url{github.com/lgbouma/tnm}} would have predicted an error
+budget consisting of the following terms added in quadrature:
+410\,{\rm ppm} from photon-counting noise, 202\,{\rm ppm} from
+detector read noise, and 673\,{\rm ppm} from the zodiacal background
+light.  The level of background light appears to have been
+overestimated.
 
@@
   (E) From functions of stellar, transit or RV parameters as discussed in text.
-  (F) {\it Gaia} DR2.
+  (F) \citet{gaia_collaboration_gaia_2018}.
   (G) \citet{stassun_TIC_2018}.
@@
 
-\citet{doyle_accurate_2013} found accurate spectroscopic parameters
-for WASP-4 using high S/N HARPS spectra.  Using their (T$_{\rm eff}$,
-$\log g$, $[{\rm Fe/H}]$) and the available {\it Gaia} parallaxes, we
-calculate the remaining system parameters empirically.
-
-We compute the bolometric flux using broadband magnitudes from
-available all-sky catalogs, as follows: $G$ from {\it Gaia\/} DR2,
-$B_T V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
-2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
-wavelength range 0.4--22~$\mu$m.  Combining the implied angular radius
-with the {\it Gaia} DR2 parallax, and adjusting for the parallax
-offset \citep{stassun_evidence_2018}, we obtain the stellar radius.
-For the stellar mass, we started with the $a/R_\star$ value found from
-the phase-folded TESS lightcurve.  The implied stellar density follows
-\citep{seager_unique_2003}, as do the
-semi-major axis and stellar mass.  We get the planet radius using our
-$R_{\rm p}/R_\star$ measurement.  Finally, for the planet mass we use the
-inclination angle from the transit, and the RV semi-amplitude from 
-\citet{triaud_spin-orbit_2010}.
-
-We list the resulting parameters in Table~1, and adopt them for the
-remaining analysis.  Our parameters broadly agree with those of previous
-authors
+We calculated the stellar and planetary parameters in the following
+way.  We computed the star's spectral energy distribution based on the
+{\it Gaia} DR2 parallax (after making the small correction advocated
+by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
+all the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$
+and $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from
+{\it 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
+wavelength range 0.4--22~$\mu$m.  We adopted the effective temperature
+from the work by \citet{doyle_accurate_2013}, who determined the
+spectroscopic parameters of WASP-4 using high-signal-to-noise
+observations with the High Accuracy Radial-velocity Planet Searcher
+(HARPS).  Then, we determined the stellar radius through the
+combination of the bolometric luminosity and the effective
+temperature, using the Stefan-Boltzmann law.  To determine the stellar
+mass, we first computed the mean stellar density based on the value of
+$a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
+(for the relevant equation, see \citealt{seager_unique_2003} or
+\citealt{winn_exoplanet_2010}).  The mass was calculated from the
+radius and density, and the orbital distance was also calculated from
+the radius and $a/R_\star$.  The planetary radius was calculated as
+the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
+mass was calculated based on the stellar mass, the radial-velocity
+amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
+inclination.
+
+Table~1 gives the resulting parameters, which we adopted for the
+remaining analysis.  Our parameters are in agreement with those of
+previous investigators
 \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
-% The sole exception is that our measured $R_{\rm p}/R_\star$ value is
-% slightly smaller than most ground-based measurements (see Tables 2 and
-% 3 of \citealt{hoyer_tramos_2013}).  However, looking at the I-band
-% measurements of \citet{hoyer_tramos_2013} and taking the mean + std
-% deviation, it's a ~1 sigma effect. So I won't mention it!
-Comparing our empirical result with the Yonsei-Yale evolutionary track
-predicted by the spectroscopic parameters, we see that WASP-4 has an
-age $\approx 7\,{\rm Gyr}$, and is securely in the main-sequence stage
-of evolution.
+By comparing the star's luminosity and spectroscopic parameters with
+the outputs of the Yonsei-Yale stellar-evolutionary models, we found
+that WASP-4 is a main-sequence star with an age of approximately
+$7\,{\rm Gyr}$.
 
@@
 
-\subsection{Pre-TESS timing measurements}
-\label{subsec:times}
-
-\begin{figure*}[t]
+\begin{figure*}[t!]
 	\begin{center}
@@
 	\vspace{-0.5cm}
-    \caption{
-        {\bf Timing residuals and best-fit models for WASP-4b.}
-        Points are observed transit times minus a linear fit to all
-        observed times.  Less precise times are more transparent.  The
-        best constant period model (gray line) is a poor fit to the
-        observations.  The best-fit tidal decay model (blue) and
-        apsidal precession model (orange) can fit the data.  The
-        binned TESS point (red square) is for display purposes only;
-        the individual TESS times are used to fit the models.
+	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
+		The residuals are the observed times minus the calculated times
+		assuming a constant period.  The darkest points correspond to
+		the most precise data. The constant-period model (gray line) is
+		a poor description of the data.  Models with a uniformly
+		decreasing period (blue) or apsidal precession (orange) provide
+		better fits to the data.  The red square represents the
+		combination of all the TESS data and is for display purposes
+		only.  The models were fitted to all of the individual transit
+		times.
 		\label{fig:times}
@@
 
-Table~1 shows the transit times we used in our analysis.  We included
+\subsection{Pre-TESS timing measurements}
+\label{subsec:times}
+
+Table~2 gives the transit times we used in our analysis. We included
 data from peer-reviewed literature for which the analysis was based on
 observations of a single, complete transit, and for which the midpoint
-was fit as a free parameter. We also required that the time system be
-clearly documented.  The occultation times are in Table~3.  There are
-fewer available, so we included all the occultation measurements we
-could find. The tabulated values have been corrected for the
+was allowed to be a free parameter. We also required that the time
+system be clearly documented. Many of the times were previously
+compiled by \citet{hoyer_tramos_2013}. We confirmed that the times in
+that paper were in agreement with the original sources and that
+barycentric corrections had been performed when needed.
+
+The earliest epoch is from EulerCam on the 1.2-m Euler telescope
+\citep{wilson_wasp-4b_2008}.  The second epoch is based on $z$-band
+photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
+with FORS2.  Subsequent observations were performed by
+\citet{winn_transit_2009}, \citet{dragomir_terms_2011},
+\citet{sanchis-ojeda_starspots_2011}, \citet{nikolov_wasp-4b_2012},
+\citet{hoyer_tramos_2013}, and \citet{ranjan_atmospheric_2014}.
+Finally, \citet{huitson_gemini_2017} acquired optical transit spectra
+with the 8.1-m Gemini South telescope between 2011 and 2014, one
+transit per season.  The per-point standard deviation of their
+lightcurves was a few hundred parts per million.  The average
+precision in their reported transit times was 5.6~seconds.  Since
+those data points carry significant weight in the analysis, we
+consulted with those authors to confirm that the time stamps in their
+data represent mid-exposure times, that the barycentric correction was
+performed correctly, and that the time system of the final results was
+BJD$_{\rm TDB}$.  These same authors also used the same instrument
+(and largely the same software) to analyze other hot Jupiters, none of
+which showed a departure from a constant-period model.
+
+We also compiled all the available occultation times, which are given
+in Table~3.  The tabulated values have been corrected for the
 light-travel time across the diameter of the orbit by subtracting
 $2a/c = 22.8$ seconds from the observed time.
-
-We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
-starting point\footnote{We also tried taking times directly from the
-references listed in Table~2; the results did not change, so we opted
-for the homogeneously determined \citet{hoyer_tramos_2013} times.}.
-Though their latest two epochs fell slightly below the prediction of a
-linear ephemeris (Panel C of their Figure~6), they did not find a need
-to fit a quadratic function to the O-C values.  We verified that all
-times were on the BJD$_{\rm TDB}$ time standard using the
-\citet{eastman_achieving_2010} calculator.  We set the zero-point
-epoch to be as close as possible to the average of the midtimes,
-weighted as the inverse midtime variance. This lowered the
-covariance between the transit epoch and the period
-\citep{gibson_gemini_2013}.
-
-Important contributions to the timing dataset were as follows.  Our
-earliest epoch was from EulerCam on the 1.2-m Euler telescope
-\citep{wilson_wasp-4b_2008}.  The second epoch came from $z$-band
-photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
-with FORS2.  We included subsequent transits from
-\citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
-the GPS clock time in UT reference. They then converted this to BJD in
-the TDB reference for their reported midtimes.},
-\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011},
-\citet{nikolov_wasp-4b_2012}, \citet{hoyer_tramos_2013}, and
-\citet{ranjan_atmospheric_2014}.  Finally, \citet{huitson_gemini_2017}
-acquired optical transmission spectra with the 8.1-m Gemini South
-telescope over 2011 to 2014, one transit per season.  The per-point
-standard deviation of their lightcurves was a few hundred parts per
-million, within a factor of 3 from the photon noise limit.  The
-precision on their reported midtimes averaged 5.6~seconds. Though
-extremely precise, this was possible given the quality of their
-data~\citep{carter_analytic_2008}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
-%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS
-%clock time in UT reference. They then converted this to BJD in the
-%TDB reference for their reported midtimes.}, and reported transit
-%midtimes precise to 6~seconds.  \citet{dragomir_terms_2011} updated
-%the ephemeris using two transits measured with the CTIO 0.9-m and
-%1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
-%transits from the Baade telescope at Las Campanas\footnote{
-%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
-%to show that the star's rotation axis is nearly aligned with the
-%planet's orbital axis, in agreement with the Rossiter-McLaughlin
-%measurement by \citet{triaud_spin-orbit_2010}.}.
-%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
-%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
-%the MPG/EOS-2.2~m at La Silla.  \citet{hoyer_tramos_2013} reported
-%nine transit observations taken with the the Y4KCAM on the SMARTS 1-m
-%telescope, and another three with the SOI on the SOAR 4.2-m
-%telescope.  \citet{hoyer_tramos_2013} also homogeneously redetermined
-%the transit times from all previous studies.  After the study by
-%\citet{hoyer_tramos_2013}, \citet{ranjan_atmospheric_2014} acquired
-%near-IR spectra of WASP-4b using HST's WFC3 in both transit and
-%secondary eclipse.  We adopt only their transit time, since the epoch
-%of occultation was not fit as a free parameter.'
-
-% There have been other studies of WASP-4b for which the transit
-% timing data are either hetereogenous, not available, or not useable.
-% We omitted the timing measurements from
-% \citet{southworth_high-precision_2009}, since there were technical
-% problems with the computer clock at the time of
-% observation~\citep{nikolov_wasp-4b_2012}.  The timing study
-% by~\citet{petrucci_no_2013} ruled out transit timing variations
-% larger than 54~seconds over the previous 5 years, but did not report
-% transit times.  \citet{baluev_benchmarking_2015} collected transit
-% times for many hot Jupiters, including WASP-4b, and included times
-% from the Exoplanet Transit Database
-% (ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
-% \citep{poddany_ETD_2010}.  Since the ETD data come from
-% heterogeneous sources, their timestamps are less clearly documented,
-% and the times are thus more prone to systematic errors, we omit them
-% from consideration.  %We consider the effects of including ETD data
-% in %Appendix~\ref{sec:verify_archival_times}.
-% \citet{may_mopss_2018} obtained optical spectra using IMCAS on
-% Magellan, and found a flat transmission spectrum, in agreement with
-% the results reported by \citet{huitson_gemini_2017}.  They did not
-% report transit times.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There were fewer available occultation times.
 \citet{beerer_secondary_2011} observed two occultations of WASP-4b
@@
 ground in the $K_S$ band, and gave a time in HJD, without specifying
-the time standard.  We assumed the standard was UTC, and verified that
-assuming TDB did not alter any of our conclusions.
-%Finally, as previously mentioned, \citet{ranjan_atmospheric_2014}
-%acquired occultation data with HST but kept the epoch as a fixed
-%parameter.
-Finally, \citet{zhou_secondary_2015} acquired occultation data from
-the Anglo-Australian Telescope, focusing on the eclipse depth, but
-also fitting for $e\cos\omega$.  We converted their reported
-$e\cos\omega$ to a midtime, using their ephemeris and the standard
-formula \citep[{\it e.g.},][]{winn_exoplanet_2010}
+the time standard.  We assumed the standard was UTC, and performed the
+appropriate corrections to convert to BJD$_{\rm TDB}$.  We
+also verified that none of our conclusions would be changed if this
+assumption was mistaken. Finally, \citet{zhou_secondary_2015} observed an
+occultation with the the Anglo-Australian Telescope.  They did not
+report the observed midpoint, but they did report a result for
+$e\cos\omega$ based upon the observed midpoint.  We calculated the
+implied midpoint using the formula \citep[{\it
+e.g.},][]{winn_exoplanet_2010}
 \begin{equation}
@@
 \end{equation}
-In total, we had four occultations.  Although the occultation times
-were less precise than the transit times, they nonetheless helped 
-modeling the timing variations.
+In total, there are four occultation times.
 
@@
 
-After collecting the literature times and measuring the TESS times, we
-compared the observed time to the prediction based solely on the
-literature times. The result is shown in
-Figure~\ref{fig:arrived_early}. The TESS observations 
-arrived early.  
-Simply out of caution, we were immediately concerned about systematic
-errors in either the archival or the TESS timestamps. The tests we
-used to mitigate the latter possibility are described in
-Appendix~\ref{sec:verify_tess}. Checks against systematic errors in
-the pre-TESS times are discussed below.
-
-Assuming that the observed timing variation is astrophysical, we
-proceeded by exploring three models for the timing data, identical to
-the study by~\citet{patra_2017}.
+First, we fitted a constant-period model (``linear ephemeris'') to the
+pre-TESS data, and used it to extrapolate to the epochs of the TESS
+observations.  The residuals of the best fitting model are shown in
+Figure~\ref{fig:arrived_early}.  The transits observed by TESS
+occurred earlier than expected.  Because the TESS mission is still in
+an early stage, we were concerned about a possible offset in the TESS
+time stamps due to an error with the TESS clock or the data processing
+pipeline.  Appendix~\ref{sec:verify_tess} describes some tests that
+convinced us that a simple offset is unlikely. Assuming that the
+observed timing variation is astrophysical, we proceeded by exploring
+three models for the timing data in a manner identical to the study
+by~\citet{patra_2017}.
 
@@
 \end{align}
-for $E$ the epoch number.
-The two free parameters are the reference epoch $t_0$ and the period $P$.
-
-The second model assumes a constant period derivative, and a circular
-orbit:
+where $E$ is the epoch number.
+We defined the epoch numbers such that
+$E=0$ is near the weighted average of the observed times.
+This helps to reduce the covariance between $t_0$ and $P$.
+
+The second model assumes the period is changing at a steady rate:
 \begin{align}
@@
 \end{align}
-The three free parameters are the epoch, period at the reference epoch,
-and period derivative, ${\rm d}P/{\rm d}t = (1/P) {\rm d}P/{\rm d}E$.
+The three free parameters are the reference epoch $t_0$, the period at
+the reference epoch, and the period derivative, ${\rm d}P/{\rm d}t =
+(1/P) {\rm d}P/{\rm d}E$.
 
 The third model assumes the planet has a slightly eccentric orbit, and
-that the line of apsides is rotating:
+that the line of apsides is rotating \citep{gimenez_revision_1995}:
 \begin{align}
@@
 \end{align}
-for $P_{\rm s}$ the sidereal period, $e$ the eccentricity, $P_{\rm a}$
-the anomalistic period, and $\omega$ the argument of pericenter.
-Following~\citet{gimenez_revision_1995} and~\citet{patra_2017}, in
-this model the angular velocity of the line of apsides ${\rm
-d}\omega/{\rm d}E$ is constant,
+where $P_{\rm s}$ is the sidereal period, $e$ is the eccentricity,
+$P_{\rm a}$ is the anomalistic period, and $\omega$ is the argument of
+pericenter.  In this model the angular velocity of the line of apsides
+${\rm d}\omega/{\rm d}E$ is constant,
 \begin{equation}
@@
 \end{equation}
-and the sidereal and anomalistic periods are related by
+and the sidereal and anomalistic periods are connected through the
+equation
 \begin{equation}
@@
 \end{equation}
-The five free parameters are the epoch, sidereal period, eccentricity,
-argument of pericenter at the reference epoch, and angular velocity of
-line of apsides:
-$(t_0, P_{\rm s}, e, \omega_0, {\rm d}\omega/{\rm d}E)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Assuming the error bars on the transit midtimes are gaussian and
-% independent, we then proceed by fitting each model to the timing data.
-% For the linear model, we took uniform priors over small windows around
-% the expected parameters.  For the quadratic model we did the same, and
-% took the quadratic term to have a uniform prior corresponding to
-% $
-%   Q_\star' \sim \mathcal{U}[10^3, 10^9],
-% $
-% for $Q_\star'$ the modified quality factor defined in
-% \S~\ref{sec:implications}.  For the precession model, we took the same
-% priors in epoch and sidereal period, and drew the remaining parameters
-% from wide uniform distributions.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%TODO: final numbers
-Figure~\ref{fig:times} shows the residuals with respect to the linear
-model.  The best linear fit has $\chi^2 = 174$ and 61 degrees of
-freedom.  The best quadratic fit has $\chi^2 = 62.6$ and 60
-degrees of freedom, compared to $\chi^2 = 64.4$
-and 58 degrees of freedom for the best-fit precession model.
-
-%TODO: final numbers. note epoch changes...
-The difference in $\chi^2$ between the linear and quadratic fit
-corresponds to $p \approx 10^{-26}$.  A more likely possibility is
-clearly that a systematic error is affecting one of the most important
-sets of times: either the earliest times, the Huitson et al.\ Gemini
-times (epoch numbers 30, 308, 577, and 837), or the TESS times.  We
-have taken every precaution against systematic offsets in the TESS
-times (Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by
-their reported times (J.-M. D\'esert, priv.\
-comm.)\footnote{\citet{huitson_gemini_2017} corrected the times to be
-at mid-exposure. They also analyzed other hot Jupiters, which all
-showed mid-transit times as expected from a constant ephemeris.
-The two Spitzer transits at epochs 362 and 365 are in independent
-agreement with the Gemini times.}.
-Since the linear model provides a poor fit to the data, we discard it
-from further consideration.
-
-The quadratic model provides a slightly better fit to the data than the
-precession model.  It is favored by $\Delta \chi^2 = 1.8$, and has two
-fewer free parameters.  Another useful heuristic for model comparison
-is the Bayesian Information Criterion (BIC),
+The five free parameters of this model are $(t_0, P_{\rm s}, e,
+\omega_0, {\rm d}\omega/{\rm d}E)$, denoting the reference epoch, the
+sidereal period, the eccentricity, the argument of pericenter at the
+reference epoch, and the angular velocity of the line of apsides.
+
+Figure~\ref{fig:times} shows the residuals with respect to the
+constant-period model.  The best-fitting constant-period model has
+$\chi^2 = 174$ and 61 degrees of freedom.  The best-fitting
+decreasing-period model has $\chi^2 = 62.6$ and 60 degrees of freedom.
+The best-fitting precession model has $\chi^2 = 64.4$ and 58 degrees
+of freedom.  Either of the latter two models provides a much better
+fit than the constant-period model.  The difference in $\chi^2$
+between the first and second models corresponds to $p \approx
+10^{-26}$.
+
+The decreasing-period model provides a slightly better fit to the data
+than the precession model.  It is favored by $\Delta \chi^2 = 1.8$,
+and has two fewer free parameters.  A useful heuristic for model
+comparison is the Bayesian Information Criterion (BIC),
 \begin{equation}
@@
 \end{equation}
-for $k$ the number of free parameters, and $n$ the number of data
-points. For us, $n=62$.  The difference in the BIC for the precession
-and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
-BIC}_{\rm quad} = 10$, corrsonding to a Bayes factor $\approx
-1.2\times10^{4}$.  The Akaike Information Criterion similarly favors
-the orbital decay model, by $\Delta {\rm AIC} = 5.8$.  This is strong
-evidence that the quadratic model describes the data better than the
-precession model \citep{kass_bayes_1995}.
-
-If we assume the orbital decay model, we find a period derivative 
+where $k$ is the number of free parameters, and $n$ is the number of
+data points. In this case, $n=62$.  The difference in the BIC between
+the precession and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm
+  prec} - {\rm BIC}_{\rm quad} = 10$, corresponding to a Bayes factor
+of $1.2\times 10^{4}$.  Likewise, the Akaike Information Criterion
+favors the constant-period-derivative model by $\Delta {\rm AIC} =
+5.8$.  Differences of this magnitude are traditionally deemed ``strong
+evidence'' that one model is a better description of the data than the
+other \citep{kass_bayes_1995}, although we prefer to reserve judgment
+until more data can be obtained.
+
+In the decreasing-period model, the period derivative is
 \begin{equation}
@@
   = - (3.98 \pm 0.38)\times 10^{-10}
-  = - (12.6) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
+  = - (12.6 \pm 1.2)~{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
 \end{equation}
-For comparison, the period derivative seen by
-\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b
-is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If both planets
-are decaying, WASP-4b would be inspiraling at about half the rate as
-WASP-12b.
-
-If we instead assume the precession model, the best-fit eccentricity
-is
+For comparison, the best-fitting period derivative of WASP-12b is
+$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$
+\citep{maciejewski_departure_2016,patra_2017}.  If both planets are
+truly falling onto their stars, then WASP-4b is falling at about half
+the rate of WASP-12b.  In the precession model, the best-fit
+eccentricity is
 \begin{equation}
@@
 \end{equation}
-The longitude of periapse correspondingly advances by $\dot{\omega} =
+The longitude of periastron advances by $\dot{\omega} =
 14.0^{+4.9}_{-3.5}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
-period is $26^{+9}_{-7}$ years.
-
-The median parameters and their 68\% confidence intervals are reported
-for the three models in Table~4.  In short, a linear ephemeris does
-not describe the observed times.  Both orbital decay and precession
-can fit the data, but the precession model is statistically
-disfavored.
+period is $26^{+9}_{-7}$ years.  All of the best-fitting parameters
+(the medians of the posterior distributions) and the 68\% credible
+intervals are reported in Table~4.
 
@@
 
-\begin{figure}[t!]
+\begin{figure}[!t]
   \begin{center}
-    \includegraphics[width=0.417\textwidth]{f5.pdf}
+    \includegraphics[width=0.40\textwidth]{f5.pdf}
   \end{center}
@@
   \caption{
-    {\bf WASP-4b should have a short eccentricity damping time and an average
-    decay time compared to other hot Jupiters.  It seems to orbit a
-    main sequence host.
-    }
+    {\bf WASP-4b in the context of other hot Jupiters.}
+    Most of the data in these plots are from \citet{bonomo_gaps_2017},
+    who measured the eccentricities using radial velocities.  The
+    colored symbols highlight WASP-4, WASP-12 (which also shows
+    evidence for a decreasing period), and the collection of hot
+    Jupiters analyzed in Appendix A.
     {\bf \it Top:}
-    Points are transiting giant planets from \citet{bonomo_gaps_2017},
-    who measured the eccentricities using radial velocities.  Solid
-    squares (including WASP-18b) have significant eccentricity
-    detections.  The axes are chosen so that constant eccentricity
-    damping timescales (Equation~\ref{eq:de_dt}) are lines.  The
-    contours shown assume $Q_{\rm p}' = 10^5$.
+    The key parameters relevant to eccentricity damping.  Solid
+    squares represent planets with a securely detected nonzero
+    eccentricity. Lines of constant damping timescale are drawn based
+    on Equation~\ref{eq:de_dt} and assuming $Q_{\rm p}' = 10^5$.
+    WASP-4b has one of the shortest eccentricity damping timescales.
     {\bf \it Middle:} 
-    Axes are chosen so to that constant orbital decay timescales
-    (Equation~\ref{eq:da_dt}) are lines.  The contours shown assume
-    $Q_\star' = 10^7$.
+    Plotted are the key parameters relevant to orbital decay.  Lines
+    of constant decay timescale are drawn based on
+    Equation~\ref{eq:da_dt}), assuming $Q_\star' = 10^7$.  WASP-4b has
+    a relatively short orbital decay timescale, although it is not as
+    extreme a case in this regard as it is for eccentricity damping.
     {\bf \it Bottom:}
-    HR diagram highlighting WASP-12b, which shows timing variations
-    similar to WASP-4b.  We also show hot Jupiters for which we
-    detect no timing variation (see \S~\ref{sec:hj_verification}).
+    A Hertzsprung-Russell diagram for hot Jupiter hosts. WASP-4
+    appears to be on the main sequence.
     \label{fig:context}
@@
 
-The middle panel of Figure~\ref{fig:context} shows the expected
-orbital decay time of WASP-4b compared to other transiting giant
-planets.  Although roughly 20 hot Jupiters have shorter predicted
-orbital decay times than WASP-4b, most have shorter observing
-baselines.  A separate consideration is that some hot Jupiter host
-stars ({\it e.g.}, WASP-18) are fully radiative, and may dissipate
-tidal energy less efficiently than stars with convective zones, like
-WASP-4.
-
-For the sake of argument, let us assume that the timing variation is
-caused entirely by orbital decay.  If the orbital period decreases
-at a fixed rate, then it will reach zero after
+If the timing variation is being caused caused entirely by orbital
+decay, then the characteristic timescale of the decay is
 \begin{equation}
@@
 \end{equation}
-For comparison, the survival time found by \citet{patra_2017} for
-WASP-12b was 3.2\,Myr.
-
-Let us further assume that the ``constant phase lag'' model for tidal
-interaction in a binary applies \citep{zahn_tidal_1977}.  The star's
-response to tidal forcing is parametrized by a modified\footnote{For
-stars, $k_\star \sim \mathcal{O}(10^{-2})$, so it is important to
-explicitly distinguish $Q_\star'$ from $Q_\star$ \citep[{\it
-e.g.},][]{schwarzschild_structure_1958}.} quality factor, $Q_\star' =
-3 Q_\star / (2k_\star)$.  Here $k_\star$ is the stellar Love number,
-which is smaller when the star's density distribution is more
-centrally concentrated. $Q_\star$ is the ratio between the energy
-stored in the equilibrium deformation of the star and the energy lost
-to heat per tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A
-larger $Q_\star$ implies less efficient tidal dissipiation.  Though
-the fluid mechanics responsible for the dissipation should depend on
-the amplitude and frequency of the tidal perturbation
-\citep[][Section~3.3]{ogilvie_tidal_2014}, we will treat $Q_\star'$ as
-a constant.  This should not be taken literally; $Q_\star'$ may in
-fact change sharply as the tidal frequency changes
-\citep{penev_empirical_2018}.
-
-If the planet's spin and orbit are synchronized and the star is slowly
-rotating, the semi-major axis and eccentricity then evolve as
-\citep[Appendix B of][]{metzger_optical_2012}
+For comparison, the corresponding time for WASP-12b is 3.2\,Myr
+\citep{patra_2017}.
+
+If WASP-4 really is undergoing rapid orbital decay, then how many of
+the other known hot Jupiters should have orbits that are decaying at
+detectable (or nearly detectable) rates?  Figure~\ref{fig:context}
+compares some key properties of WASP-4 with those of a larger ensemble
+of hot Jupiters.  The middle panel displays two parameters that
+strongly affect the expected orbital decay timescale, $P
+M_\star/M_{\rm p}$ and $a/R_\star$.  WASP-4 has one of the shortest
+theoretical timescales for orbital decay.  There are about 20 hot
+Jupiters (including WASP-12) for which the theoretical timescale is
+shorter.  In almost all of those cases, though, the planet was
+discovered more recently than WASP-4 and a decade-long baseline of
+observations is not yet available.  A separate consideration not shown
+in Figure~\ref{fig:context} is that the hot Jupiter host stars have a
+variety of different structures, from being fully convective to nearly
+fully radiative, which may lead to widely divergent tidal dissipation
+timescales.
+
+In the simple ``constant phase lag'' model for tidal interaction
+\citep{zahn_tidal_1977}, the rate of dissipation can be parametrized
+by a modified\footnote{For stars, $k_\star \sim \mathcal{O}(10^{-2})$,
+  so it is important to explicitly distinguish $Q_\star'$ from
+  $Q_\star$ \citep[{\it e.g.},][]{schwarzschild_structure_1958}.}
+quality factor, $Q_\star' = 3 Q_\star / (2k_\star)$.
+Here, $Q_\star$ is the ratio between the energy stored in the
+equilibrium deformation of the star and the energy lost to heat per
+tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A larger
+$Q_\star$ implies less efficient tidal dissipiation. The dimensionless
+number $k_\star$ is the stellar Love number, which is smaller when the
+star's density distribution is more centrally concentrated.  In this
+model, once the planet's spin and orbit are synchronized, then the
+semi-major axis and eccentricity evolve as \citep[Appendix B
+of][]{metzger_optical_2012}
 \begin{align}
@@
 The modified quality factor of WASP-4 corresponding to the observed
-value of $\dot{P}$ is then
-\begin{equation}
-	Q_\star' \approx 2.9\times10^4,
-\end{equation}
-which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
-10^5$, based on the orbits of the Galilean moons
-\citep{lainey_strong_2009}.  For stars, studies of the binary
-eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
-\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
-geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
-differ in hot Jupiter systems, since the tidal forcing frequencies and
-amplitudes are different.  Observationally-driven studies of hot
-Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
-different models
+value of $\dot{P}$ is
+\begin{equation}
+	Q_\star' \approx 2.9\times10^4.
+\end{equation}
+This is about an order of magnitude lower than the value that was
+inferred for WASP-12b.  It is also smaller than most theoreticians
+would have expected.  The $Q_{\rm Jup}'$ value of Jupiter is estimated
+to be $\approx$$1.4 \times 10^5$, based on the observed motions of the
+Galilean moons \citep{lainey_strong_2009}.  For stars, studies of the
+binary eccentricity distribution have been interpreted with tidal
+models, giving $Q_\star' \approx 10^5 - 10^7$ \citep[{\it
+    e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
+  geller_direct_2013,milliman_wiyn_2014}.  Population studies of hot
+Jupiter systems have also been undertaken, generally finding $Q_\star'
+\approx 10^5 - 10^8$ using different models
 \citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
 For instance, motivated by the rapid rotation of some hot Jupiter
-hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
+hosts
+\citep{pont_empirical_2009,ciceri_hats-15b_2016,penev_hats-18b_2016},
 \citet{penev_empirical_2018} modeled the evolution of hot Jupiter
@@
 (1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
-quality factor we infer from the apparent period change.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% In fact, for the observed population of hot Jupiters,
-% \citet{penev_constraining_2012} argued that $Q_\star' > 10^7$ would
-% be required at some phase of hot Jupiter evolution at the 99\%
-% level, or else the observed population would not exist (though
-% \citealt{birkby_wts-2_2014} point out more efficient dissipation
-% would be allowed if the hot Jupiters arrived later).  Developing a
-% generative model for the orbital separation distribution of the hot
-% Jupiter population, \citet{cameron_hierarchical_2018} similarly
-% found $10^7 \gtrsim Q_\star' \gtrsim 10^8$.
-
-% Using an empirical approach, \citet{penev_empirical_2018} reported a
-% $Q_\star'$ measurement for WASP-4.  In earlier work,
-% \citet{penev_hats-18b_2016} noted that HATS-18b and WASP-19b orbited
-% host stars whose isochronal ages are consistent with the Sun, but
-% whose stellar rotation periods are about 10 days
-% (\citealt{pont_empirical_2009} pointed out similar evidence for hot
-% Jupiter host spin-up).  The expected rotation periods at solar ages
-% are 20 to 30 days
-% (\citealt{schatzman_theory_1962,skumanich_time_1972}, and
-% measurements by \citealt{barnes_rotation_2016} in the 4\,Gyr-old M67
-% cluster).  Assuming that the difference between the observed and
-% predicted stellar rotation period was due to tides depositing
-% angular momentum in the star, \citet{penev_empirical_2018} modeled
-% the evolution of hot Jupiter systems under the influence of a
-% magnetized wind and a constant phase-lag tide.  They found two-sided
-% limits for $Q_\star'$ in 35 of the 188 systems they examined.  In
-% these 35 systems, they found that $Q_\star'$ varied from $10^5$ at
-% tidal periods of 2 days to $10^7$ at tidal periods of 0.5 days.  For
-% WASP-4 specifically, their method gave $Q_\star' \approx
-% (1.2^{+1.0}_{-0.5})\times10^7$. This is more than 2 orders of
-% magnitude less efficient than the dissipation rate inferred from the
-% decaying orbit would imply.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\citet{essick_orbital_2016} proposed a theoretically-motivated tidal
-dissipation model , in which a coupled network of gravity modes is
-excited at the base of the stellar convective zone, and brakes near
-the stellar core to provide the needed dissipation.  In their
-approach, the stellar quality factors in hot Jupiter systems vary from
-$Q_\star' \approx 10^5 - 10^6$.  From their Equation 26, the
-prediction for WASP-4 is $Q_\star' = 7\times10^5$.  This is an order
-of magnitude larger than implied by the observed period change.
+quality factor we have inferred from the apparent period change.
+
+\citet{essick_orbital_2016} studied the problem of the orbital decay
+of hot Jupiters using a theory in which gravity modes are excited at
+the base of the stellar convective zone, propagate inward through the
+radiative core and break near the stellar core, leading to energy
+dissipation.  They predicted the stellar quality factors in hot
+Jupiter systems to vary from $Q_\star' \approx 10^5 - 10^6$.  From
+their Equation 26, the prediction for WASP-4 is $Q_\star' =
+7\times10^5$, which is an order of of magnitude larger than implied by
+the observed period change.
 
 The applicability of the \citet{essick_orbital_2016} model depends on
-the evolutionary state of the star.  The bottom panel of
-Figure~\ref{fig:context} shows WASP-4 in the context of giant planet
-hosts from the \citet{bonomo_gaps_2017} sample.  On the y-axis is
-$G=g-\mu$, for $g$ the apparent Gaia-band magnitude, and $\mu$ the
-distance modulus reported by \citet{gaia_collaboration_gaia_2018}.
-The x-axis is the effective temperature from \citet{bonomo_gaps_2017},
-which for WASP-4 agrees within $1\sigma$ of that from Table~1.
-Inspecting the HR diagram, WASP-4 shows little evidence of being
-evolved, in agreement with our analysis from
-\S~\ref{sec:system_parameters}.
+the evolutionary state of the star. \citet{weinberg_tidal_2017} showed
+that more rapid dissipation --- enough to account for the period
+change of WASP-12b --- could exist in stars that have begun evolving
+into red giants.  The bottom panel of Figure~\ref{fig:context} shows a
+Hertzsprung-Russell diagram of hot Jupiters hosts, including WASP-4.
+On the y-axis is $G=g-\mu$, for $g$ the apparent Gaia-band magnitude,
+and $\mu$ the distance modulus reported by
+\citet{gaia_collaboration_gaia_2018}.  The x-axis is the effective
+temperature from \citet{bonomo_gaps_2017}, which for WASP-4 agrees
+within $1\sigma$ of that from Table~1.  Inspecting the HR diagram,
+WASP-4 shows little evidence of being evolved, in agreement with our
+analysis from \S~\ref{sec:system_parameters}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To summarize, if the observed period change is caused entirely by
-tidal decay, then the implied stellar tidal dissipation efficiency
-would be stronger than expected.  It might be possible that we are
+tidal orbital decay, then the constant-phase-lag tidal model implies a
+stellar tidal dissipation rate that is higher than expected by at
+least an order of magnitude.  It might be possible that we are
 observing at a special time, shortly after the planet's inward
@@
 oscillation mode.  Tidal dissipation rates might also be increased if
-the star is turning off the main sequence.  A final possibility is
-that an exterior planet could be trapping WASP-4b's spin vector in a
-high-obliquity state, causing planetary obliquity tides as suggested
-for WASP-12b \citep{millholland_obliquity_2018}.  For now, all of
-these explanations seem strained; it is unclear that tidal decay can
-explain the observed timing variations by any mechanism that we have
-considered.
+the star is just turning off the main sequence.  Another hypothesis,
+recently advanced by \citet{millholland_obliquity_2018} for the case
+of WASP-12b, is that an exterior planet could be trapping WASP-4b's
+spin vector in a high-obliquity state, leading to rapid dissipation
+through planetary obliquity tides.
 
@@
 
-Let us instead assume that the timing variation is a portion of an
-apsidal precession cycle.  The required orbital eccentricity is then
-$e\approx 1.8\times10^{-3}$, and one full precession period would take
-about 24 years.
-
+If instead the observed timing variation is just a small portion of an
+apsidal precession cycle, then the orbital eccentricity is a few times
+$10^{-3}$, and the full precession period is about 24 years.
 \citet{ragozzine_probing_2009} calculated apsidal precession periods
-for hot Jupiters ranging between $\approx$10--100 years.  They
-highlighted that for many hot Jupiters, including WASP-4b, over 90\%
-of the apsidal precession comes from the tidal bulge of the planet.
-Precession from general relativity, the planet's rotational bulge, and
-the star's rotational and tidal bulges are minor additions.
-%\citet{ragozzine_probing_2009} also calcuated that WASP-4b was expected
-%to have a relatively short precession period, of 120 years.
-
-\citet{ragozzine_probing_2009} also noted that if the planet has
-non-zero eccentricity, a measurement of $\dot{\omega}$ can be used to
-measure the planet's Love number.  From their Equation 14, the implied
-Love number for WASP-4b is
+for hot Jupiters, finding them to range between about 10 and 100
+years. They highlighted that for many hot Jupiters, including WASP-4b,
+the theoretical precession rate is dominated by the non-Keplerian
+force due to the planet's tidal bulge.  Precession from general
+relativity, the planet's rotational bulge, and the star's rotational
+and tidal bulges contribute at the 10\% level at most.  Thus, a
+measurement of the precession rate can be used to determine the
+planet's Love number.  From their Equation 14, the implied Love number
+for WASP-4b is
 \begin{equation}
@@
 \end{equation}
-For comparison, $k_{2,{\rm Jup}} \approx 0.55$
-\citep{wahl_tidal_2016,ni_empirical_2018}.  A uniform density sphere
-has $k_2 = 1.5$.  Our uncertainties on $k_{2,{\rm p}}$ are large
-because the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$
-are correlated, and the measured occultation times only barely
-constrain them (Figure~\ref{fig:future}).
-
-\subsubsection{Observations of small eccentricities in hot
-Jupiter systems}
-
-Could the eccentricity truly be non-zero?  For WASP-4b, the
-eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
-0.30 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
-$1\,{\rm Gyr}$ old, so unless the planet arrived very recently, any
-initial eccentricity should have been erased.
-
-The top panel of Figure~\ref{fig:context} compares the expected
-eccentricity damping time of WASP-4b with other transiting giant
-planets.  WASP-4b has one of the shortest known eccentricity damping
-times.  The hot Jupiter with the nearest $\tau_e$ and a significant
-measured eccentricity is WASP-18b, with $e = 0.0076 \pm 0.0010$
-\citep{triaud_spin-orbit_2010,bonomo_gaps_2017}.  This is the smallest
-significant eccentricity measured by the RV method, with the
-next-smallest cases being HAT-P-20b ($e =
-0.01580^{+0.0023}_{-0.0018}$), and WASP-38b ($e =
-0.02780^{+0.003}_{-0.0028}$).
-% For the 231 planets studied by \citet{bonomo_gaps_2017}, 2 (13) had
-% $2\sigma$ upper limits of $e < 0.01$ ($e<0.02$).  For WASP-4b, the
-% strongest $2\sigma$ upper-limit on its eccentricity from RVs is
-% $e<0.011$ \citep{husnoo_observational_2012,bonomo_gaps_2017}.
-
-Occultation timing can provide stronger eccentricity limits than RV
-measurements, but they often come with caveats.  For WASP-4b, the
-strongest limit on its eccentricity comes from
-\citet{beerer_secondary_2011}, who observed occultations and derived a
-$2\sigma$ upper limit on $| e\cos\omega | $ of 0.0024.
-% \citet{caceres_ground-based_2011} report $e\cos\omega = 0.0027 \pm
-% 0.0018$.  \citet{zhou_secondary_2015} report a weaker $e\cos\omega =
-% -0.001 \pm 0.003$. 
-Placing eccentricity limits stronger than $e< 10^{-3}$ has only been
-possible for a few hot Jupiters, and with numerous false-starts.  HD
-189733b, for example, was initially reported to have a non-zero
-eccentricty \citep{knutson_map_2007}.  This result was retracted by
-\citet{agol_climate_2010}, who used a larger dataset of 7 Spitzer
-occultations and found a smaller average offset in the occultation
-time than \citet{knutson_map_2007}.  \citet{agol_climate_2010} also
-noted that a larger fraction of the planet's observed occultation time
-offset could be explained by detailed modeling of the planetary
-hot-spot, which can shift occultation times by up to a few minutes
-\citep{williams_resolving_2006}.  Including an updated hot-spot model,
-\citet{agol_climate_2010} reported $e\cos\omega < 0.00023$ at
-2$\sigma$, perhaps the strongest eccentricity limit in place for a hot
-Jupiter.  Similar sagas transpired for HD 209458b and WASP-12b, with
-early reports of small non-zero eccentricities
-\citep{winn_measurement_2005,hebb_wasp-12b_2009,lopez-morales_day-side_2010},
-only to be ruled out by the limits of subsequent followup:
-$|e\cos\omega| < 0.0007$ for HD 209458b, and $| e\cos\omega| < 0.0040$
-($2\sigma$) for WASP-12b
-\citep{crossfield_spitzer_mips_2012,campo_orbit_2011,croll_near-infrared_2011}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% A similar saga occurred for HD 209458b, for which
-% \citet{winn_measurement_2005} combined Spitzer secondary eclipses with
-% RV data and reported a small non-zero eccentricity.  They suspected that
-% the data were consistent with zero eccentricity, and that the lower
-% bound on $e$ was an artifact of an imperfect limb-darkening model.
-% \citet{crossfield_spitzer_mips_2012} eventually ruled out the \citet{winn_measurement_2005} values, by determining
-% $e\cos\omega < 0.0007$ at $2\sigma$.  The final example is WASP-12b, for
-% which the discovery RVs and photometry gave a small finite eccentricity
-% measurement \citep[][at $3\sigma$]{hebb_wasp-12b_2009}.  A prompt
-% ground-based occultation measurement by
-% \citet{lopez-morales_day-side_2010} concurred that a small eccentricity
-% was favored with weak statistical significance. Further follow-up found
-% occultation times consistent with circular orbits; $| e\cos\omega| <
-% 0.0040 $ \citep{campo_orbit_2011,croll_near-infrared_2011}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To summarize, empirical constraints on the ellipticity of hot Jupiter
-orbits at the $e \lesssim 10^{-3}$ level are present for only a
-handful of hot Jupiters, from occultation timing.  Once the timing
-offsets are of order tens of seconds, eccentricities from occultation
-timing may have systematic uncertainties because of the hot-spot issue
-\citep{williams_resolving_2006,agol_climate_2010}.  Except for
-WASP-18b, radial velocity measurements have provided eccentricity
-constraints only at the $e > 10^{-2}$ level.  We therefore cannot
-empirically rule out hot Jupiter eccentricities at the level needed to
-assess WASP-4's putative eccentricity.
-
-\subsubsection{Theoretical expectations for small eccentricities in hot
-Jupiter systems}
-
-Lacking empirical data, we turn to theoretical mechanisms that could
-produce small eccentricities.
+For comparison, the Love number of Jupiter is about 0.55
+\citep{wahl_tidal_2016,ni_empirical_2018}, and a uniform density
+sphere has $k_2 = 1.5$. The uncertainty in $k_{2,{\rm p}}$ for WASP-4b
+is large because the eccentricity, reference time, and ${\rm
+d}\omega/{\rm d}E$ have strongly correlated errors, and the measured
+occultation times only barely constrain these parameters
+(Figure~\ref{fig:future}).
+
+The main problem with the apsidal precession hypothesis is to explain
+why the eccentricity would be as large as $\sim$$10^{-3}$ despite
+rapid tidal circularization.  For WASP-4, Equation~\ref{eq:de_dt}
+gives $\tau_e = 0.30 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is
+several billion years old, so unless the planet arrived very recently,
+any initial eccentricity should have been lowered well below
+$\sim$$10^{-3}$.  The top panel of Figure~\ref{fig:context} compares
+the expected eccentricity damping time of WASP-4b with that of other
+transiting giant planets.  WASP-4b has one of the shortest known
+eccentricity damping times.
 
 \paragraph{Neighboring companion}
+One way to maintain a significant eccentricity is through the
+gravitational perturbations from another planet.
 \citet{mardling_long-term_2007} considered the long-term tidal
@@
 timescales several orders of magnitude longer than the circularization
-time (see their Figures~4 and 5).
+time of an isolated hot Jupiter (see their Figures~4 and 5).
 
@@
 least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
-%\citet{bailey_understanding_2019} point out how the joint requirements
-%of Kozai-Lidov oscillations with a non-detection of an additional RV
-%signal can be combined to produce lower and upper limits on the
-%companion's mass.
 For the Kozai-Lidov mechanism to operate at maximum efficiency, we
@@
 $\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
-(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
-transits along the line of sight, $f$ can be arbitrarily small, and
-both of the preceding limits can be satisfied.
-
-While either of the two preceding mechanisms could produce the needed
-eccentricity, it would be more satifying if long-term radial velocity
-monitoring showed an additional planet in the system.  The chances of
-one existing with a mass between 1-13\,$M_{\rm Jup}$ and orbital
-semi-major axis between 1-20\,AU is about 50\%, and a linear trend is
-currently present at $\sim$2$\sigma$ \citep{knutson_friends_2014}.
+(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b is
+transiting, $f$ can be arbitrarily small, and both of the preceding
+limits can be satisfied.
 
@@
 gravitational fluctuations from stellar convection \citep[][Section
-7]{phinney_pulsars_1992}.  Phinney presents this idea by analogy with
-a pendulum, which is never really at rest. Instead, collisions with
-air molecules randomly perturb its momentum and energy, and the
-equipartition theorem of statistical mechanics tells us the average
-energy is $\langle E_{\rm pend} \rangle = k T_{\rm air}$.  In hot
-Jupiter systems, the ``air molecules'' are convective eddies, and
-their associated kinetic energy perturbs the orbit away from
-circularity.  The mean-squared eccentricity is \citep[][Equation
-7.33]{phinney_pulsars_1992}
+7]{phinney_pulsars_1992}.  From equation 7.33 of that work, the
+mean-squared eccentricity of the orbit is
 \begin{equation}
@@
 \end{equation}
-for $L$ the stellar luminosity, $R_{\rm conv}$ ($M_{\rm conv}$) the
-width (mass) of the convective region,  $\mu$ the reduced mass, $n$
-the orbital frequency, and $a$ the semi-major axis.  For WASP-4, we
-find $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$; this mechanism
-cannot produce the required eccentricity.
+where $L$ is the stellar luminosity, $R_{\rm conv}$ and $M_{\rm conv}$
+are the width and mass of the convective region, $\mu$ is the reduced
+mass, $n$ is the orbital frequency, and $a$ is the semi-major axis.
+For WASP-4, this equation implies $\langle e^2 \rangle^{1/2} \lesssim
+10^{-5}$.  Hence, this mechanism does not seem capable of producing
+the required value of $\sim$$10^{-3}$.
 
@@
 gravitational quadrupole of the star would cause the orbit of the
-planet to precess on the timescale of the stellar activity cycle ("the
-dynamo timescale").  An essential difference between this process and
-apsidal precession is that Applegate timing variations are expected to
-be quasi-periodic \citep[{\it
-e.g.},][Figure~12]{soderhjelm_geometry_1980}.  The transit and
-occultation deviations from the Applegate mechanism would also have
-equal signs, while for apsidal precession they have opposite signs.
-For WASP-4, \citet{watson_orbital_2010} estimated that the Applegate
-effect could produce timing deviations of up to 15 seconds, depending
-on the modulation period of the stellar dynamo, and the corresponding
-level of differential surface shear.  If true, this mechanism cannot
-explain the majority our observed $78$ second variation.
+planet to precess on the timescale of the stellar activity cycle.  An
+essential difference between this process and apsidal precession is
+that Applegate timing variations need not be strictly periodic
+\citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}. This
+mechanism would also produce transit and occultation timing deviations
+of the same sign, while for apsidal precession they would have
+opposite signs.  For WASP-4, \citet{watson_orbital_2010} estimated
+that the Applegate effect could produce timing deviations of up to 15
+seconds, depending on the modulation period of the stellar dynamo, and
+the corresponding level of differential surface shear.  If this
+analysis is accurate, then the Applegate mechanism cannot explain the
+majority our observed $82$ second variation.
 
@@
 
-Another mechanism to produce the observed period change would be if an
-unseen companion in the WASP-4 system accelerated the star towards our
-line of sight.
-This would produce a period derivative
+An apparent change in orbital period can also be caused by an overall
+acceleration of the center of mass of the system.  This would produce
+a period derivative
 \begin{equation}
@@
 \end{equation}
-for $\dot{v}_{\rm r}$ the time derivative of the radial velocity.
-%Recently, \citet{bonomo_gaps_2017} collected all available HARPS data
-%for the system, and reported no detectable slope at $3\sigma$.
-%However,
-\citet{knutson_friends_2014} collected 5 Keck HIRES radial
-velocity measurements for WASP-4, and merged them with measurements
-from CORALIE and HARPS \citep{wilson_wasp-4b_2008,pont_determining_2011,husnoo_observational_2012}.
-Using these datasets, \citet{knutson_friends_2014} found 
+where $\dot{v}_{\rm r}$ is the time derivative of the radial velocity.
+\citet{knutson_friends_2014} combined radial-velocity data from their
+own program as well as those of \citet{wilson_wasp-4b_2008},
+\citet{pont_determining_2011}, and \citet{husnoo_observational_2012}
+to obtain a constraint on any long-term trend:
 \begin{equation}
@@
 \end{equation}
-or that the
-star can be accelerating towards us at most by $\dot{v}_{\rm r} > -
-0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ (2$\sigma$).
-The associated $\dot{P}$ must be less than $9\times 10^{-11}$.  
-Radial acceleration could thus cause at most one-quarter of the
-observed period decrease (Equation~\ref{eq:dP_dt_obs}).
-Improved radial velocity
-monitoring will help clarify the situation,
-particularly given that WASP-4 shows short-term RV trends with
-amplitudes $\sim$$\mathcal{O}(10\,{\rm m\,s}^{-1})$, over tens of days
-\citep{husnoo_observational_2012}.  Given that WASP-4 is a
-magnetically active star \citep{sanchis-ojeda_starspots_2011}, these
-short-timescale variations might conspire to produce spurious
-RV signals over long baselines.
+which can also be expressed as a 2$\sigma$ upper limit of $0.021
+\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ on the acceleration along the
+line of sight.  The associated upper limit on the apparent period
+change resulting from this acceleration is $9\times 10^{-11}$, which
+is only one-quarter of the observed period decrease
+(Equation~\ref{eq:dP_dt_obs}).
+
+Thus, it seems that the observed period decrease cannot be fully
+explained as a constant acceleration of the center of mass of WASP-4.
+Given the limited amount and uneven time coverage of the existing
+radial-velocity data, it remains possible that the center of mass has
+a more complex motion, due to a companion on a highly eccentric orbit
+\citep[similar to {\it e.g.}, WASP-53 or WASP-81][]{triaud_peculiar_2017}.
+It would be useful to gather more radial-velocity data to confirm or
+refute this possibility.
 
 There are two other small effects worth noting.  The first is the
-\citet{shklovskii_possible_1970} effect, in which the star's proper
-motion leads to a changing radial velocity.  This would be apparent in
-the Doppler measurements described above.  The effect contributes a
-change of order $\dot{P} \sim P\mu^2 d/ c \sim 6\times10^{-13}$.  The
-second effect, described by \citet{rafikov_stellar_2009}, comes from
-the star's on-sky motion altering our viewing angle, and leads to an
+\citet{shklovskii_possible_1970} effect due to the star's proper
+motion, which leads to an apparent period change of $P\mu^2 d/ c$,
+which is only $6\times10^{-13}$ for the case of WASP-4.  The second
+effect, described by \citet{rafikov_stellar_2009}, comes from the
+star's on-sky motion altering our viewing angle, and leads to an
 observed apsidal precession. The corresponding period change is
-$\dot{P} \sim (P\mu)^2/2\pi \sim 10^{-21}$. Both effects are too small
-to be of consequence in WASP-4.
-
-Many of the phenomena we have discussed could presumably occur
-simultaneously.  We found though that radial acceleration,
-proper-motion, and the Applegate mechanism are unlikely to explain the
-majority of the timing deviations.  Tidal decay seems implausible,
-given the high level of dissipation required.  This leaves apsidal
-precession as the most plausible explanation, but more data are needed
-for a definitive interpretation.
-
-\section{Future prospects}
+$\dot{P} \sim (P\mu)^2/2\pi$, which is on the order of $10^{-21}$ for
+WASP-4, too small to be of any consequence.
+
+\section{Call for additional observations}
 \label{sec:future}
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f6.pdf}
+		\includegraphics[width=0.49\textwidth]{f6.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Further observations will be needed to
-	    confirm and understand the timing variations of WASP-4b.}
-		Dots are as in Figure~\ref{fig:times}.
-		Lines are 100 random draws from the posteriors of the apsidal
-		precession model (orange), and the orbital decay model (blue).    
+    {\bf Further observations will be needed to confirm and understand
+    the timing variations of WASP-4b.}
+    Dots are as in Figure~\ref{fig:times}.  Lines are 100 random draws
+    from the posteriors of the apsidal precession model (orange), and
+    the orbital decay model (blue).    
 		\label{fig:future}
@@
 
+A primary purpose of this work has been to call attention to the
+timing anomaly of WASP-4 that has been sighted by TESS, and alert
+observers to the need for follow-up transit timing, occultation
+timing, and radial-velocity monitoring.  There is no unique
+interpretation of the current data, and two of the possibilities ---
+orbital decay and apsidal precession --- would be of great interest to
+confirm. Detection of orbital decay would lead to an unusually direct
+determination of a stellar dissipation rate. Detection of apsidal
+precession would give a rare constraint on the interior density
+distribution of an exoplanet.
+
 If the TESS mission is extended after its primary mission, it will
 likely observe additional transits of WASP-4b in the early 2020s
-(Figure~\ref{fig:future}).  High-precision ground-based transit
-measurements could also be useful.  In order to rule between orbital
-decay and apsidal precession, occultation measurements in both the
-near term and also in the mid-2020s will be needed.  Additionally,
-confirmation and extension of the linear RV trend suggested by
-\citet{knutson_friends_2014} would help in understanding whether an
-exterior companion in the system might explain any residual
-eccentricity.  These observations should be carried out because they
-will lead to new knowledge; either a measure of the stellar tidal
-dissipation efficiency, or the eccentricity and Love number of the
-planet.  Either outcome for a single system would be rather
-interesting.  In the coming decade, wide-field photometric surveys
-(TESS, HATPI, NGTS, PLATO) might even enable such measurements at a
-population level, which would be extremely interesting.
-
-\acknowledgements{
+(Figure~\ref{fig:future}).  High-precision transit observations with
+larger telescopes would also be useful.  In order to decide between
+orbital decay and apsidal precession, occultation measurements in both
+the near term and also in the mid-2020s will be needed.  More
+radial-velocity data would help in the search for additional bodies
+that could be causing dynamical perturbations, or an overall acceleration
+of the host star.
+
+TESS will also be monitoring the other known hot Jupiters, which will
+reveal whether the timing anomalies seen in WASP-12 and now WASP-4 are
+commonplace, and may shed some light on the circumstances in which
+they arise. Other wide-field photometric surveys, such as the Next
+Generation Transit Survey \citep{wheatley_next_2018}, HATPI
+(\href{https://hatpi.org}{hatpi.org}) and
+PLATO \citep{rauer_plato_2014} will also extend
+the time baseline of transit timing for a large number of systems.
+
+\acknowledgements
 L.G.B.\ gladly acknowledges helpful discussions with
@@
 people who have turned TESS from an idea into reality.
+%
+J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
+supporting this work.
 %
@@
 	TESS \citep{ricker_transiting_2015},
-	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}}
+	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
 }
@@
 
-%\clearpage
-
-%\newpage
+\clearpage
+\newpage
 \input{WASP-4b_transit_time_table.tex}
@@
 
-Any systematic offset between the TESS times and the true barycentric
-reference would cast doubt on the results of this study.  There is
-historic precedent: Kepler had a systematic error in its timestamps
-that was corrected only in Q14~\citep[][Section
-3.4]{kepler_DR19_2013}.
-
-We devised two checks on the absolute calibration of the TESS time
-system.  First, in \S~\ref{sec:headers}, we recalculate and confirm
-the barycentric correction reported by SPOC in the TESS lightcurve
-file headers.  In \S~\ref{sec:hj_verification}, we measure the transit
-times of other hot Jupiters, and use them to rule out a global TESS
-time offset larger than about a minute.  The latter test confirms that
-WASP-4b is an outlier.
-
-\subsection{Using the headers of the lightcurve files}
+An obvious concern that one might have about the WASP-4 timing anomaly
+is that there might be a systematic offset between the TESS time
+system and the time system in which the previous observations have
+been reported.  There is a precedent for this type of error: data from
+the Kepler mission was afflicted by a systematic timing error that was
+corrected only late in the mission~\citep[][Section
+  3.4]{kepler_DR19_2013}. We performed two checks on the calibration
+of the TESS time system, described below.
+
+\subsection{Checking for consistency in the headers of the lightcurve files}
 \label{sec:headers}
@@
 time, $N=37$, and the offset was as expected.  Then, using the ${\rm
-JD}_{\rm UTC}$ timestamp, we recalculated the barycentric correction
-computed by SPOC, using the \citealt{eastman_achieving_2010}
-calculator.  Due to the first author's ignorance of the spacecraft
-position, we assumed that the observer was located at the Earth's
-geocenter, and used the correct direction for each star.  This gave us
-times in ${\rm BJD}_{\rm TDB}$ that agreed with the archival times to
-within 1.7 seconds.  This offset is comparable to the light-travel
-delay expected for TESS on its orbit from perigee of $\approx
-20R_\oplus$ to apogee of $\approx 60R_\oplus$.
-
-This calculation bounds any error in the SPOC barycentric julian date
-correction to be less than 1.7 seconds.  This is smaller than the
-effect of interest, so we treat the SPOC BJD correction as
-``verified'', and proceed to a subsequent test.
-
-\subsection{Using other hot Jupiters as references}
+JD}_{\rm UTC}$ timestamp, we calculated the barycentric correction
+using the software written by \citealt{eastman_achieving_2010}.  (For
+simplicity, we assumed the TESS spacecraft was located at the Earth's
+geocenter, because the light-travel time between the geocenter and the
+spacecraft is on the order of only one second.)  This gave us times in
+${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
+that are delivered with the TESS data products.
+
+\subsection{Checking for timing anomalies in other hot Jupiter systems}
 \label{sec:hj_verification}
@@
     times and the barycentric reference.} While the WASP-4b transits
-    fell about 75 seconds earlier than expected, other well-observed
+    fell about 82 seconds earlier than expected, other well-observed
     hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
@@
 \end{figure*}
+% JNW: I think the KDE needs to be removed and perhaps replaced with something
+% else.  It gives the impression that the WASP-4 anomaly is only at
+% the 2-sigma level, which is false.  The KDE illustrates the spread
+% in values, but what is relevant is the standard deviation of the
+% mean of the measurements, which is narrower than the KDE implies.
+% I suggest replacing it with a Gaussian with a mean equal to the mean of the
+% deviations, and a width equal to the uncertainty in the mean, i.e., the
+% standard deviation divided by sqrt(N-1).
 
 If the observed timing delay in WASP-4b were caused by a systematic
-offset between the TESS times and the ${\rm
-BJD}_{\rm TDB}$ reference, we might expect that it would apply to
-other hot Jupiters as well.  This test can rule out an important class
-of clock error~--~a systematic global offset.
-
-To explore this possibility, we repeat the timing analysis of the main
-paper, for other hot Jupiters with long preceding observing baselines.
-We first checked which hot Jupiters were observed over the first two
-TESS sectors using a combination of
+global offset between the TESS time system and the ${\rm BJD}_{\rm
+TDB}$ reference, we would expect that it would be apparent in other
+hot Jupiter systems, too. It would also be apparent in eclipsing
+binary observations and any other periodic phenomena that have been
+observed over a long time baseline. Here we examine only hot Jupiters
+because of our greater familiarity with the data.
+
+We repeated all the data reduction and analysis steps described in
+this paper for other hot Jupiters observed by TESS for which timing
+data exists spanning many years.  First, we checked which hot Jupiters
+were observed over the first two TESS sectors using a combination of
 \texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
@@
 
-Using the literature timing data for each hot Jupiter, we then
-performed a least-squares fit to a linear ephemeris.  Using the
-best-fit values and variances, we calculated the precision of the
-predicted transit time during the TESS observing epoch.  This gave 11,
-94, 18, 42, and 59 seconds for WASP-4b, 5b, 6b, 18b, and 46b.
-Comparing the observed and predicted times, Figure~\ref{fig:hjs} shows
-that WASP-4b is the only hot Jupiter that transited significantly
-earlier than expected.
-
-To convert this intuition into a quantitative limit, for each hot
-Jupiter we considered the model
+We determined the best-fitting constant-period ephemeris based on the
+pre-TESS data. Then we used the parameters and uncertainties in the
+best-fitting model to calculate the predicted transit times during
+the TESS observation period, as well as the uncertainty in the
+predicted times.  The uncertainties are 11, 94, 18, 42, and 59 seconds
+for WASP-4b, 5b, 6b, 18b, and 46b, respectively.  By comparing the
+observed and predicted times, Figure~\ref{fig:hjs} shows that WASP-4b
+is the only hot Jupiter that transited significantly earlier than
+expected.
+
+To use these results to place a quantitative limit on any global clock
+offset, for each hot Jupiter we considered the model
 \begin{equation}
@@
 distribution, $(t_0', P')$ are the best-fit reference time and period
-using only the literature transit times, and $(\sigma_{t_0'},
-\sigma_{P'})$ are their uncertainties.
-
-For each planet, we then ask: what fraction of the posterior for
-$t_{\rm offset}$ is consistent with an offset worse than $81.5$
-seconds?  For WASP-4b, the answer is unsurpringly about half.  For
-WASP-6b, the most constraining object, about 1 sample in 2 million is
-consistent with such a timing offset ($4.9\sigma$).  For WASP-18b, 1
-in 80 samples would be consistent with this timing offset
-($2.2\sigma$), and in WASP-46b, the limit is 1 in 44 samples
-($2.0\sigma$).  For WASP-5b, the predicted time is too imprecise to
-rule out timing offsets at the necessary amplitude.  Multiplying the
-three independent probabilities for WASP-5b, 6b, and 18b, we can rule
-out $t_{\rm offset} < -81.5\ {\rm seconds}$ at $6.3\sigma$, or about
-about 1 part in 8 billion.
+using only the pre-TESS transit times, and $(\sigma_{t_0'},
+\sigma_{P'})$ are the corresponding uncertainties.
+
+For each planet, we asked: what fraction of the posterior for $t_{\rm
+offset}$ is consistent with an offset worse than $81.5$ seconds?  For
+WASP-4b, the answer is unsurprisingly 50\%.  For WASP-6b, the most
+constraining object, about 1 sample in 2 million is consistent with
+such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 80 samples
+would be consistent with this timing offset ($2.2\sigma$), and in
+WASP-46b, the limit is 1 in 44 samples ($2.0\sigma$).  For WASP-5b,
+the predicted time is too imprecise to rule out timing offsets at the
+necessary amplitude.  Multiplying the three independent probabilities
+for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.5\
+{\rm seconds}$ at $6.3\sigma$, or about about 1 part in 8 billion.
 

### 2018-12-16 paper/ms.tex
update paper

@@
 \usepackage{comment}
-\usepackage{todonotes}
 
@@
     18 transits of the hot Jupiter WASP-4b.  The sequence of transits
-    occurred $81.5 \pm 11.8$~seconds earlier than had been predicted,
+    occurred $81.5 \pm 11.7$~seconds earlier than had been predicted,
     based on earlier data stretching back to 2007.  This is unlikely
@@
     constant period, ruling out an 81.5-second offset at the
-    6.3$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
+    6.4$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
     to be decreasing at a rate of $\dot{P} = -12.6 \pm 1.2$
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{New transits and system parameters}
@@
 
-\begin{figure}[t!]
+\begin{figure}[t]
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{f2.png}
+        \includegraphics[width=0.49\textwidth]{f1.png}
     \end{center}
@@
 
-\begin{figure}[t!]
+\begin{figure}[t]
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{f1.png}
+        \includegraphics[width=0.49\textwidth]{f2.png}
     \end{center}
@@
 
-\begin{figure}[t!]
-    \begin{center}
-        \leavevmode
-        \includegraphics[width=0.49\textwidth]{f3.pdf}
-    \end{center}
-    \vspace{-0.6cm}
-    \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
-      Both plots show the deviations between the observed and
-      calculated transit times, where the calculation is based only on
-      the pre-TESS data and assumes a constant period.  The blue bands
-      depict the $\pm$$1\sigma$ credible interval of the predicted
-      times.  {\it Top:} The full timing dataset spans 10 years. The
-      darkest points correspond to the most precise data.  {\it
-        Bottom:} Close-up of the TESS observations. The red band shows
-      the average deviation of the TESS data, which is $81.5 \pm
-      11.8\ {\rm seconds}$.
-        \label{fig:arrived_early}
-    }
-\end{figure}
-
 Using the cleaned PDC lightcurve, we used the Box Least Squares
@@
   quad})$, and the orbital period $P$.
-% JNW: Luke, didn't you also have to fit a linear trend for each transit?
-% Otherwise the fit would be poor, for each transit, even if the
-% trends tend to cancel out in the phase folded light curve.
-% And, did you assign the photometric uncertainties as described below?
 
@@
 reduced $\chi^2$ of unity.  This lowered the mean uncertainty of the
-transit midtimes from $29.8$ to $22.6$ seconds.
+transit midtimes from $29.8$ to $22.6$ seconds.  We verified that
+omitting this step did not appreciably alter any of our conclusions.
 
@@
 windows, the lightcurves have an rms scatter of 586\,{\rm ppm}.  The
-pre-launch TESS noise model \citep{winn_photonflux_2013,Sullivan_2015}
-\footnote{\url{github.com/lgbouma/tnm}} would have predicted an error
+pre-launch TESS noise model\footnote{\url{github.com/lgbouma/tnm}} \citep{winn_photonflux_2013,Sullivan_2015}
+would have predicted an error
 budget consisting of the following terms added in quadrature:
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Timing analysis}
@@
 
-\begin{figure*}[t!]
+\begin{figure}[t]
+    \begin{center}
+        \leavevmode
+        \includegraphics[width=0.49\textwidth]{f3.pdf}
+    \end{center}
+    \vspace{-0.6cm}
+    \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
+      Both plots show the deviations between the observed and
+      calculated transit times, where the calculation is based only on
+      the pre-TESS data and assumes a constant period.  The blue bands
+      depict the $\pm$$1\sigma$ credible interval of the predicted
+      times.  {\it Top:} The full timing dataset spans 10 years. The
+      darkest points correspond to the most precise data.  {\it
+        Bottom:} Close-up of the TESS observations. The red band shows
+      the average deviation of the TESS data, which is $81.5 \pm
+      11.7\ {\rm seconds}$.
+        \label{fig:arrived_early}
+    }
+\end{figure}
+
+\begin{figure*}[t]
 	\begin{center}
@@
 \end{align}
-where $E$ is the epoch number.
-We defined the epoch numbers such that
-$E=0$ is near the weighted average of the observed times.
-This helps to reduce the covariance between $t_0$ and $P$.
+where $E$ is the epoch number.  We defined the epoch numbers such that
+$E=0$ is near the weighted average of the observed times.  This helps
+to reduce the covariance between $t_0$ and $P$.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Interpretation}
@@
 
-\begin{figure}[!t]
+\begin{figure}[t]
   \begin{center}
-    \includegraphics[width=0.40\textwidth]{f5.pdf}
+    \includegraphics[width=0.38\textwidth]{f5.pdf}
   \end{center}
@@
 \subsection{Applegate effect}
+
 Some eclipsing binaries exhibit period modulations with amplitudes of
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Call for additional observations}
@@
 
-\begin{figure}[!t]
+\begin{figure}[t]
 	\begin{center}
@@
     {\bf Further observations will be needed to confirm and understand
-    the timing variations of WASP-4b.}
-    Dots are as in Figure~\ref{fig:times}.  Lines are 100 random draws
-    from the posteriors of the apsidal precession model (orange), and
-    the orbital decay model (blue).    
+    the timing variations of WASP-4b.} Dots are as in
+    Figure~\ref{fig:times}.  Lines are 100 random draws from the
+    posteriors of the apsidal precession model (orange), and the
+    orbital decay model (blue).    
 		\label{fig:future}
@@
 
-\clearpage
-\newpage
+%\clearpage
+%\newpage
 \input{WASP-4b_transit_time_table.tex}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Verifying the TESS Time Stamps}
@@
     hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
-    Ticks are observed TESS transit midtimes; the blue distribution is
-    a kernel density estimate; the orange distribution is a gaussian
-    centered on zero with standard deviation ($\sigma_{\rm
-    predicted}$) calculated from the pre-TESS transit times.
+    Ticks are observed TESS transit midtimes; the orange distribution
+    is a gaussian centered on zero with standard deviation
+    ($\sigma_{\rm predicted}$) calculated from the pre-TESS transit
+    times.  The blue distribution is a gaussian centered on the
+    weighted average of the TESS times, with width equal to the
+    uncertainty in the mean, {\it i.e.}, the standard deviation
+    of the TESS residual times divided by $\sqrt{N-1}$, with $N$ the
+    number of transits.
     \label{fig:hjs}
@@
 \end{figure*}
-% JNW: I think the KDE needs to be removed and perhaps replaced with something
-% else.  It gives the impression that the WASP-4 anomaly is only at
-% the 2-sigma level, which is false.  The KDE illustrates the spread
-% in values, but what is relevant is the standard deviation of the
-% mean of the measurements, which is narrower than the KDE implies.
-% I suggest replacing it with a Gaussian with a mean equal to the mean of the
-% deviations, and a width equal to the uncertainty in the mean, i.e., the
-% standard deviation divided by sqrt(N-1).
 
@@
 data exists spanning many years.  First, we checked which hot Jupiters
-were observed over the first two TESS sectors using a combination of
+were observed over the first three TESS sectors using a combination of
 \texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
@@
 constraining object, about 1 sample in 2 million is consistent with
-such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 80 samples
-would be consistent with this timing offset ($2.2\sigma$), and in
-WASP-46b, the limit is 1 in 44 samples ($2.0\sigma$).  For WASP-5b,
+such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 103 samples
+would be consistent with this timing offset ($2.3\sigma$), and in
+WASP-46b, the limit is 1 in 49 samples ($2.0\sigma$).  For WASP-5b,
 the predicted time is too imprecise to rule out timing offsets at the
@@
 for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.5\
-{\rm seconds}$ at $6.3\sigma$, or about about 1 part in 8 billion.
+{\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.
 

### 2018-12-17 paper/ms.tex
update paper; ready for collaborators

@@
     occurred $81.5 \pm 11.7$~seconds earlier than had been predicted,
-    based on earlier data stretching back to 2007.  This is unlikely
+    based on data stretching back to 2007.  This is unlikely
     to be the result of a clock error, because TESS observations of
@@
     shortcomings.  The gravitational influence of a third body is another
-    possibility, although at present there is scant evidence for such a
+    possibility, although at present there is no evidence for such a
     body. Further observations are needed to confirm and
@@
 circular, which may itself be a consequence of star-planet tides.  If
-it could be observed, then \citep{ragozzine_probing_2009} showed that
+it could be observed, then \citet{ragozzine_probing_2009} showed that
 it would be possible to measure the planet's Love number, which would
@@
         transit (blue line) is used when measuring mid-transit times for
-        the individual transits (Figure~\ref{fig:lightcurves}).
+        the individual transits (see text).
         \label{fig:phasefold}
@@
 removed the data that might have been adversely affected by ``momentum
-dumps,'' the firing of thrusters and resetting of reaction wheels that
-took place every 2.5 days during sector
-2.\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}.
+dumps,'' the firing of thrusters and resetting of reaction wheels\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}
+that took place every 2.5 days during sector 2.
 The data during these events were assigned quality flags corresponding
@@
 alone.  Based on the results, we isolated the data within 4 transit
-durations of each transit midpoint.  We then fitted all of the transit
-data using the analytic formulae given by \citet{mandel_analytic_2002}
-and implemented by \citet{kreidberg_batman_2015}.  We assumed the
-orbit to be circular. The free parameters were the reference epoch,
-the planet to star radius ratio $R_{\rm p}/R_\star$, the orbital
-distance to stellar radius ratio $a/R_\star$, the inclination $i$, two
-quadratic limb-darkening coefficients $(u_{\rm linear}, u_{\rm
-  quad})$, and the orbital period $P$.
+durations of each transit midpoint.  To find the transit parameters
+that best fit the data, we first fitted a line to the out-of-transit
+flux measurements surrounding each transit, and divided it out.  We
+then created a phase-folded lightcurve and fitted a standard transit
+model using the analytic formulae given by
+\citet{mandel_analytic_2002} and implemented by
+\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular.
+The free parameters were the reference epoch, the planet to star
+radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
+radius ratio $a/R_\star$, the inclination $i$, two quadratic
+limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
+orbital period $P$.
 
@@
 
-To measure the transit times, we fitted each transit separately, with
+To measure the transit times, we returned to the `cleaned' PDC time
+series and fitted the data surrounding each transit separately, with
 four free parameters: the time of mid-transit $t_{\rm tra}$, the
@@
 fixed the remaining parameters at the values that had been determined
-from the phase-folded TESS lightcurve.  The uncertainty in each
-data point was set equal to the root-mean-square (rms) level of the
+from the phase-folded TESS lightcurve.  The uncertainty in each data
+point was set equal to the root-mean-square (rms) level of the
 out-of-transit data.
@@
 freedom.  The variance of the $\chi^2$ distribution is $2n$, so we
-would expect $\chi^2 16 \pm 5.7$.  Visually inspecting the residuals
+would expect $\chi^2 = 16 \pm 5.7$.  Visually inspecting the residuals
 showed that the error variance had been overestimated, so we
@@
 {\it Transit/RV parameters:} & & & \\
-  $R_{\rm p}/R_\star$                        & $0.15117$              & $+0.00047$, $-0.00037$      & A \\
-  $i$~[deg]                                  & $88.85$                & $+0.77$, $-0.90$            & A \\
-  $a/R_\star$                                & $5.457$                & $+0.030$, $-0.066$          & A \\
-  $u_{\rm linear}$                           & $0.388$                & ---                         & A \\
-  $u_{\rm quad}$                             & $0.212$                & ---                         & A \\
+  $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
+  $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
+  $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
+  $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
+  $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
   $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
@@
   $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
-  $\rho_\star$~[g~cm$^{-3}$]                 & $1.717$                & $+0.029$, $-0.062$          & E \\
-  $M_\star$~[M$_{\odot}$]                    & $0.867$                & $+0.087$, $-0.097$          & E \\
+  $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
+  $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
 {\it Planetary parameters:} & & & \\
-  $a$~[AU]                                   & $0.0227$               & $\pm 0.0008$                & E \\
-  $M_p$~[M$_{\rm Jup}$]                      & $1.189$                & $+0.093$, $-0.104$          & E \\
-  $R_p$~[R$_{\rm Jup}$]                      & $1.314$                & $\pm 0.039$                 & E \\
+  $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
+  $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
+  $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
 \enddata
@@
 by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
-all the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$
+the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$
 and $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from
@@
       depict the $\pm$$1\sigma$ credible interval of the predicted
-      times.  {\it Top:} The full timing dataset spans 10 years. The
+      times.  {\it Top:} The full timing dataset spans 11 years. The
       darkest points correspond to the most precise data.  {\it
         Bottom:} Close-up of the TESS observations. The red band shows
-      the average deviation of the TESS data, which is $81.5 \pm
-      11.7\ {\rm seconds}$.
+      the average deviation of the TESS transits, which arrived $81.5 \pm
+      11.7\ {\rm seconds}$ earlier than predicted.
         \label{fig:arrived_early}
@@
 assumption was mistaken. Finally, \citet{zhou_secondary_2015} observed an
-occultation with the the Anglo-Australian Telescope.  They did not
+occultation with the Anglo-Australian Telescope.  They did not
 report the observed midpoint, but they did report a result for
@@
   = - (3.98 \pm 0.38)\times 10^{-10}
-  = - (12.6 \pm 1.2)~{\rm ms}\,{\rm yr}^{-1}.
+  = - 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
   \begin{center}
-    \includegraphics[width=0.38\textwidth]{f5.pdf}
+    \includegraphics[width=0.385\textwidth]{f5.pdf}
   \end{center}
@@
     of constant decay timescale are drawn based on
-    Equation~\ref{eq:da_dt}), assuming $Q_\star' = 10^7$.  WASP-4b has
+    Equation~\ref{eq:da_dt}, assuming $Q_\star' = 10^7$.  WASP-4b has
     a relatively short orbital decay timescale, although it is not as

### 2018-12-17 src/compute_planet_star_parameters.py
update path for parameter computing

@@ -23,7 +23,7 @@ hoyer_params = [i_hoyer13, a_by_Rs_hoyer13, a_by_Rs_perr_hoyer13,
 
 # get parameters from fit to phase-folded lightcurve
 picklepath = ('/home/luke/Dropbox/proj/tessorbitaldecay/results/'
-              'tess_lightcurve_fit_parameters/402026209/'
+              'tess_lightcurve_fit_parameters/402026209/sector_2/'
               '402026209_phased_mandelagol_fit_empiricalerrs.pickle' )
 
 d = pickle.load(open(picklepath, 'rb'))
@@ -202,5 +202,5 @@ Rs_calcd = (Rstar_by_dist * dist_pc).to(u.Rsun)
 
 print(
     'dist = {:.2f}'.format(dist_pc)+
-    '\nRstar = {:.3f}'.format(Rs_calcd)
+    '\nmy Rstar (but use Keivan\'s) = {:.3f}'.format(Rs_calcd)
 )

### 2018-12-18 src/calculate_wasp4b_eccentricity_damping_timescale.py
new numbers per phase-fold method update

@@ -13,9 +13,9 @@ P = 1.33823251*u.day
 
 #Mstar, Rstar, Mplanet, Rplanet = 0.867, 0.893, 1.189, 1.314 # USED: my table 1
 Mstar, Rstar, Mplanet, Rplanet = (
-    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
+    0.864*u.Msun, 0.893*u.Rsun, 1.186*u.Mjup, 1.321*u.Rjup
 )
-sma = 0.0227*u.au
+sma = 0.0226*u.au
 
 Qplanetprime = 1e5
 

### 2018-12-18 src/kozai_companion_mass_bound.py

@@ -9,11 +9,10 @@ import numpy as np
 from astropy import units as u, constants as const
 
 # parameters from my table
-a_b = 0.0227*u.au
 M_star, R_star, M_b, R_b = (
-    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
+    0.864*u.Msun, 0.893*u.Rsun, 1.186*u.Mjup, 1.321*u.Rjup
 )
-
+a_b = 0.0226*u.au
 
 # This number is made up, but similar to Jupiter's from Juno (Wahl+ 2016).
 k_2b = 0.6

### 2018-12-18 src/rv_companion_mass_bound.py

@@ -9,10 +9,10 @@ import numpy as np
 from astropy import units as u, constants as const
 
 # parameters from my table
-a_b = 0.0227*u.au
 M_star, R_star, M_b, R_b = (
-    0.867*u.Msun, 0.893*u.Rsun, 1.189*u.Mjup, 1.314*u.Rjup
+    0.864*u.Msun, 0.893*u.Rsun, 1.186*u.Mjup, 1.321*u.Rjup
 )
+a_b = 0.0226*u.au
 
 # from equation (52) of the latex'd equation in Avery's email
 

### 2018-12-18 paper/ms.tex
update ms.tex w/ latest numbers

@@
     18 transits of the hot Jupiter WASP-4b.  The sequence of transits
-    occurred $81.5 \pm 11.7$~seconds earlier than had been predicted,
+    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
     based on data stretching back to 2007.  This is unlikely
@@
     other hot Jupiters (WASP-6b, 18b, and 46b) are compatible with a
-    constant period, ruling out an 81.5-second offset at the
+    constant period, ruling out an 81.6-second offset at the
     6.4$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
@@
         Bottom:} Close-up of the TESS observations. The red band shows
-      the average deviation of the TESS transits, which arrived $81.5 \pm
+      the average deviation of the TESS transits, which arrived $81.6 \pm
       11.7\ {\rm seconds}$ earlier than predicted.
@@
 decreasing-period model has $\chi^2 = 62.6$ and 60 degrees of freedom.
-The best-fitting precession model has $\chi^2 = 64.4$ and 58 degrees
+The best-fitting precession model has $\chi^2 = 64.3$ and 58 degrees
 of freedom.  Either of the latter two models provides a much better
@@
 The decreasing-period model provides a slightly better fit to the data
-than the precession model.  It is favored by $\Delta \chi^2 = 1.8$,
+than the precession model.  It is favored by $\Delta \chi^2 = 1.7$,
 and has two fewer free parameters.  A useful heuristic for model
@@
   prec} - {\rm BIC}_{\rm quad} = 10$, corresponding to a Bayes factor
-of $1.2\times 10^{4}$.  Likewise, the Akaike Information Criterion
+of $1.1\times 10^{4}$.  Likewise, the Akaike Information Criterion
 favors the constant-period-derivative model by $\Delta {\rm AIC} =
@@
 \dot{P}
-  = - (3.98 \pm 0.38)\times 10^{-10}
+  = - (4.00 \pm 0.38)\times 10^{-10}
   = - 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}.
@@
 \begin{equation}
-  e = (1.83^{+ 1.88}_{- 0.75})\times10^{-3}
+  e = (1.92^{+ 1.93}_{- 0.76})\times10^{-3}
 \end{equation}
 The longitude of periastron advances by $\dot{\omega} =
-14.0^{+4.9}_{-3.5}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
-period is $26^{+9}_{-7}$ years.  All of the best-fitting parameters
+13.6^{+4.7}_{-3.6}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
+period is $27^{+10}_{-7}$ years.  All of the best-fitting parameters
 (the medians of the posterior distributions) and the 68\% credible
@@
 apsidal precession cycle, then the orbital eccentricity is a few times
-$10^{-3}$, and the full precession period is about 24 years.
+$10^{-3}$, and the full precession period is about 27 years.
 \citet{ragozzine_probing_2009} calculated apsidal precession periods
@@
 rapid tidal circularization.  For WASP-4, Equation~\ref{eq:de_dt}
-gives $\tau_e = 0.30 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is
+gives $\tau_e = 0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is
 several billion years old, so unless the planet arrived very recently,
@@
 \begin{equation}
-  M_{\rm c} > 7.1\,M_\oplus
+  M_{\rm c} > 7.5\,M_\oplus
   \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
@@
   M_{\rm c} &< 
-  23.8\,M_\oplus
+  23.7\,M_\oplus
   \times 
@@
 For each planet, we asked: what fraction of the posterior for $t_{\rm
-offset}$ is consistent with an offset worse than $81.5$ seconds?  For
+offset}$ is consistent with an offset worse than $81.6$ seconds?  For
 WASP-4b, the answer is unsurprisingly 50\%.  For WASP-6b, the most
@@
 necessary amplitude.  Multiplying the three independent probabilities
-for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.5\
+for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.6\
 {\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.

### 2018-12-18 paper/ms.tex
update author list

@@
 %
-\author[0000-0002-5720-4827]{M. Hill}
+\author[0000-0002-5720-4827]{M. L. Hill}
 \affiliation{Department of Earth Sciences, University of California,
@@
 %
-\author[0000-0002-7084-0529]{S. Kane}
+\author[0000-0002-7084-0529]{S. R. Kane}
 \affiliation{Department of Earth Sciences, University of California,
@@
 %
+\author{J. Villasenor}
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+% 
 %-------------------------------------
@@
 %
-\author{G. R. Ricker} % grr@space.mit.edu WAITING
+\author{G. R. Ricker} % grr@space.mit.edu CONFIRMED
 \affiliation{ Department of Physics and Kavli Institute for Astrophysics
@@
 %
-\author[0000-0002-5322-2315]{A. Glidden} % aglidden@mit.edu WAITING
+\author[0000-0002-5322-2315]{Ana Glidden} % aglidden@mit.edu CONFIRMED
 \affiliation{ Department of Physics and Kavli Institute for Astrophysics
@@
 02139, USA }
+\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
+Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
 %
@@
 surrounding WASP-4 were averaged into 2-minute stacks by the onboard
-computer.  The data were downlinked via the Deep Space Network. The
-timestamps from the spacecraft clock were then transformed by the
+computer.  The data were downlinked via the Deep Space Network, and the
+spacecraft time stamps were calibrated against the ground-station clocks.
+The spacecraft clock times were then transformed by the
 Payload Operations Center into the {\it Temps Dynamique Barycentrique}
@@
 \acknowledgements
-L.G.B.\ gladly acknowledges helpful discussions with
-A.~Bailey, F.~Dai and V.~Van Eylen, and is grateful to the
+L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
+A.~Bailey, F.~Dai, K.~Patra and V.~Van Eylen, and are grateful to the
 people who have turned TESS from an idea into reality.
@@
 
-An obvious concern that one might have about the WASP-4 timing anomaly
-is that there might be a systematic offset between the TESS time
-system and the time system in which the previous observations have
-been reported.  There is a precedent for this type of error: data from
-the Kepler mission was afflicted by a systematic timing error that was
-corrected only late in the mission~\citep[][Section
-  3.4]{kepler_DR19_2013}. We performed two checks on the calibration
-of the TESS time system, described below.
-
-\subsection{Checking for consistency in the headers of the lightcurve files}
-\label{sec:headers}
-
-The TESS lightcurve files provide observation start and end times in
-three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
-TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
-lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
-JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
-is the number of leap-seconds since 1961. For the relevant observation
-time, $N=37$, and the offset was as expected.  Then, using the ${\rm
-JD}_{\rm UTC}$ timestamp, we calculated the barycentric correction
-using the software written by \citealt{eastman_achieving_2010}.  (For
-simplicity, we assumed the TESS spacecraft was located at the Earth's
-geocenter, because the light-travel time between the geocenter and the
-spacecraft is on the order of only one second.)  This gave us times in
-${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
-that are delivered with the TESS data products.
-
-\subsection{Checking for timing anomalies in other hot Jupiter systems}
-\label{sec:hj_verification}
-
-\begin{figure*}[ht!]
+\begin{figure*}[t!]
 	\begin{center}
@@
 
+An obvious concern that one might have about the WASP-4 timing anomaly
+is that there might be a systematic offset between the TESS time
+system and the time system in which the previous observations have
+been reported.  There is a precedent for this type of error: data from
+the Kepler mission was afflicted by a systematic timing error that was
+corrected only late in the mission~\citep[][Section
+  3.4]{kepler_DR19_2013}. We performed two checks on the calibration of 
+the TESS time system, described below.
+
+\subsection{Checking for consistency in the headers of the lightcurve files}
+\label{sec:headers}
+
+The TESS lightcurve files provide observation start and end times in
+three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
+TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
+lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
+JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
+is the number of leap-seconds since 1961. For the relevant observation
+time, $N=37$, and the offset was as expected.  Then, using the ${\rm
+JD}_{\rm UTC}$ time stamp, we calculated the barycentric correction
+using the software written by \citealt{eastman_achieving_2010}.  (For
+simplicity, we assumed the TESS spacecraft was located at the Earth's
+geocenter, because the light-travel time between the geocenter and the
+spacecraft is on the order of only one second.)  This gave us times in
+${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
+that are delivered with the TESS data products.
+
+\subsection{Checking for timing anomalies in other hot Jupiter systems}
+\label{sec:hj_verification}
+
 If the observed timing delay in WASP-4b were caused by a systematic
@@
 for $t_{\rm offset}$ a systematic constant offset between the reported
-timestamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
+time stamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
 were
@@
 % 
-% The preceding analysis shows that the TESS timestamps are consistent
+% The preceding analysis shows that the TESS time stamps are consistent
 % at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
@@
 % \citealt{huitson_gemini_2017}, the authors also stand by their
-% timestamps. Since the ETD data come from
-% heterogeneous sources, their timestamps are less clearly documented,
+% time stamps. Since the ETD data come from
+% heterogeneous sources, their time stamps are less clearly documented,
 % and the times are thus more prone to systematic errors, we

### 2018-12-19 paper/ms.tex
remove appendix A.1

@@
 corrected only late in the mission~\citep[][Section
-  3.4]{kepler_DR19_2013}. We performed two checks on the calibration of 
-the TESS time system, described below.
-
-\subsection{Checking for consistency in the headers of the lightcurve files}
-\label{sec:headers}
-
-The TESS lightcurve files provide observation start and end times in
-three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
-TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
-lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
-JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
-is the number of leap-seconds since 1961. For the relevant observation
-time, $N=37$, and the offset was as expected.  Then, using the ${\rm
-JD}_{\rm UTC}$ time stamp, we calculated the barycentric correction
-using the software written by \citealt{eastman_achieving_2010}.  (For
-simplicity, we assumed the TESS spacecraft was located at the Earth's
-geocenter, because the light-travel time between the geocenter and the
-spacecraft is on the order of only one second.)  This gave us times in
-${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
-that are delivered with the TESS data products.
-
-\subsection{Checking for timing anomalies in other hot Jupiter systems}
-\label{sec:hj_verification}
+  3.4]{kepler_DR19_2013}.
 
@@
 \texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
-TEPCat \citep{southworth_homogeneous_2011}.  We then selected hot
-Jupiters for which there were at least five distinct epochs reported
-in the peer-reviewed literature.  We required that each observation be
-of a single transit, that the midpoint be fit as a free parameter, and
-that the time system be clearly documented.  Our final hot Jupiter
-sample included WASP-4b, 5b, 6b, 18b, and 46b.  The collected and
-measured times are given in Tables~5, 6, 7, and 8 for each.
+TEPCat \citep{southworth_homogeneous_2011}.  We recalculated the
+barycentric corrections using the \citet{eastman_achieving_2010} code,
+and found values that agreed with the lightcurve headers to within
+about 1 second.  We then selected hot Jupiters for which there were at
+least five distinct epochs reported in the peer-reviewed literature.
+We required that each observation be of a single transit, that the
+midpoint be fit as a free parameter, and that the time system be
+clearly documented.  Our final hot Jupiter sample included WASP-4b,
+5b, 6b, 18b, and 46b.  The collected and measured times are given in
+Tables~5, 6, 7, and 8 for each.
 
@@
 \input{WASP-46b_transit_time_table.tex}
+
+%\clearpage
+%\newpage
 

### 2018-12-19 paper/ms.tex
recompile

@@
 L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
-A.~Bailey, F.~Dai, K.~Patra and V.~Van Eylen, and are grateful to the
-people who have turned TESS from an idea into reality.
+A.~Bailey, F.~Dai, J.~Goodman, K.~Patra and V.~Van Eylen, and are
+grateful to the people who have turned TESS from an idea into reality.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Verifying the TESS Time Stamps}
+\section{Verifying the TESS Time Stamps Using Other Hot Jupiters}
 \label{sec:verify_tess}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f7.pdf}
+		\includegraphics[width=0.98\textwidth]{f7.pdf}
 	\end{center}

### 2018-12-22 src/plot_future.py
fine-tuning of plots

@@ -140,7 +140,12 @@ def plot_future(
                     - linear_fit(theta_linear, xfit, x_occ=xfit_occ)[1],
                 zorder=-3, alpha=0.17, c='#ff7f0e')
 
-    a0.legend(loc='lower left', fontsize='x-small')
+    a0.text(0.02, 0.04, 'Transits', transform=a0.transAxes, ha='left',
+            va='bottom')
+    a1.text(0.02, 0.04, 'Occultations', transform=a1.transAxes, ha='left',
+            va='bottom')
+
+    a0.legend(loc='lower center', fontsize='x-small')
     for ax in (a0,a1):
         ax.get_yaxis().set_tick_params(which='both', direction='in')
         ax.get_xaxis().set_tick_params(which='both', direction='in')

### 2018-12-22 src/plot_phasefold.py

@@ -58,27 +58,27 @@ def plot_phasefold(picklepath, xlim=None, ylim0=None, ylim1=None):
                                figsize=(0.8*6,0.8*4), gridspec_kw=
                                {'height_ratios':[3, 1]})
 
-    a0.scatter(phase*fit_period*24, phz_flux, c='k', alpha=0.15, label='data',
+    a0.scatter(phase*fit_period*24, phz_flux, c='k', alpha=0.12, label='data',
                zorder=1, s=10, rasterized=True, linewidths=0)
     a0.plot(bin_phase*fit_period*24, bin_fluxs, alpha=1, mew=0.5,
             zorder=8, label='binned', markerfacecolor='yellow',
             markersize=8, marker='.', color='black', lw=0, rasterized=True)
     a0.plot(
-        phase*fit_period*24, fit_phz_flux, c='b',
-        zorder=0, rasterized=True, lw=2, alpha=0.4,
+        phase*fit_period*24, fit_phz_flux, c='#4346ff',
+        zorder=2, rasterized=True, lw=1.5, alpha=0.7,
         label='model'
     )
 
     a1.scatter(
-        phase*fit_period*24, phz_flux-fit_phz_flux, c='k', alpha=0.15,
-        rasterized=True, s=10, linewidths=0
+        phase*fit_period*24, phz_flux-fit_phz_flux, c='k', alpha=0.12,
+        rasterized=True, s=10, linewidths=0, zorder=1
     )
     a1.plot(bin_phase*fit_period*24, bin_fluxs-fit_bin_fluxs, alpha=1, mew=0.5,
             zorder=8, markerfacecolor='yellow', markersize=8,
             marker='.', color='black', lw=0, rasterized=True)
     a1.plot(
-        phase*fit_period*24, fit_phz_flux-fit_phz_flux, c='b',
-        zorder=0, rasterized=True, lw=2, alpha=0.4
+        phase*fit_period*24, fit_phz_flux-fit_phz_flux, c='#4346ff',
+        zorder=2, rasterized=True, lw=1.5, alpha=0.7
     )
 
     a1.set_xlabel('Time from mid-transit [hours]')
@@ -97,7 +97,7 @@ def plot_phasefold(picklepath, xlim=None, ylim0=None, ylim1=None):
 
     f.tight_layout(h_pad=-.3, w_pad=0)
     savpath = '../results/phasefold.png'
-    f.savefig(savpath, dpi=350, bbox_inches='tight')
[…]

### 2018-12-22 src/plot_population_vs_wasp4.py

@@ -357,7 +357,7 @@ def plot_HJs_HR_diagram(catpath, ax):
     ax.scatter(x, y, marker='s', s=5, zorder=2,
                facecolors='none', edgecolors='k', lw=0.4, alpha=0.3)
 
-    marker = itertools.cycle(('o', 'v', '>', 'd', 's', 'P'))
+    marker = itertools.cycle(('o', 'v', '>', 'D', 's', 'P'))
     for tesspl in plnames:
         if tesspl=='WASP-4b':
             s=56

### 2018-12-22 src/plot_stacked_lightcurves.py

@@ -110,9 +110,9 @@ def plot_stacked_lightcurves(
     resultspng = '../results/{:d}_stacked_lightcurves.png'.format(ticid)
     paperpng = '../paper/f1.png'
     paperpdf = '../paper/f1.pdf'
-    f.savefig(resultspng, dpi=400, bbox_inches='tight')
+    f.savefig(resultspng, dpi=600, bbox_inches='tight')
     print('saved {:s}'.format(resultspng))
-    f.savefig(paperpng, dpi=400, bbox_inches='tight')
+    f.savefig(paperpng, dpi=600, bbox_inches='tight')
     print('saved {:s}'.format(paperpng))
     f.savefig(paperpdf, bbox_inches='tight')
     print('saved {:s}'.format(paperpdf))

### 2018-12-22 paper/ms.tex
update author list

@@
 %
-\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu WAITING
+\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu CONFIRMED
 \affiliation{ Department of Physics and Kavli Institute for Astrophysics
@@
 %
-\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu WAITING
-\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
-Street, Cambridge, MA 02138, USA }
-%
-\author{S. Seager} % seager@mit.edu WAITING
+\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu CONFIRMED
+\affiliation{Center for Astrophysics \textbar \ Harvard \&
+Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+%
+\author{S. Seager} % seager@mit.edu CONFIRMED
 \affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
@@
 %
-\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov WAITING
+\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov CONFIRMED
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
@@
 %
-\author[0000-0002-3321-4924]{Z. Berta-Thompson} % Zachory.BertaThompson@Colorado.EDU WAITING
+\author[0000-0002-3321-4924]{Z. Berta-Thompson} % zachory.bertathompson@colorado.EDU CONFIRMED
 \affiliation{Department of Astrophysical and Planetary Sciences,
@@
 %
-\author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.Caldwell@nasa.gov % WAITING
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-%
-\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov WAITING
+% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % WAITING
+% \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+% \affiliation{ SETI Institute, Mountain View, CA 94043, USA}
+%
+\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov CONFIRMED
 \affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
@@
 %
-\author{M. Fausnaugh} % faus@mit.edu WAITING
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
+% \author{M. Fausnaugh} % faus@mit.edu WAITING
+% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
+% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+% 02139, USA }
 %
@@
 %
-\author{N. Guerrero} % nmg@mit.edu WAITING
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0001-8812-0565]{J. Rodriguez} % rodriguez.jr.joey@gmail.com WAITING
-\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden
-Street, Cambridge, MA 02138, USA }
-%
-\author[0000-0002-6778-7552]{J. D. Twicken} % Joseph.Twicken@nasa.gov WAITING
+% \author{N. Guerrero} % nmg@mit.edu WAITING
+% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
+% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+% 02139, USA }
+%
+\author[0000-0001-8812-0565]{J. E. Rodriguez} % rodriguez.jr.joey@gmail.com CONFIRMED
+\affiliation{Center for Astrophysics \textbar \ Harvard \&
+Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+%
+\author[0000-0002-6778-7552]{J. D. Twicken} % joseph.twicken@nasa.gov CONFIRMED
 \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
@@
 %
-\author{B. Wohler} % bill.wohler@nasa.gov WAITING
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
+% \author{B. Wohler} % bill.wohler@nasa.gov WAITING
+% \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+% \affiliation{ SETI Institute, Mountain View, CA 94043, USA}
 %

### 2018-12-22 paper/ms.tex
add comments from all coauthors

@@
     occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
-    based on data stretching back to 2007.  This is unlikely
-    to be the result of a clock error, because TESS observations of
-    other hot Jupiters (WASP-6b, 18b, and 46b) are compatible with a
-    constant period, ruling out an 81.6-second offset at the
-    6.4$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
-    to be decreasing at a rate of $\dot{P} = -12.6 \pm 1.2$
-    milliseconds per year.  This might be caused by tidal orbital
-    decay or apsidal precession, although both interpretations have
-    shortcomings.  The gravitational influence of a third body is another
-    possibility, although at present there is no evidence for such a
-    body. Further observations are needed to confirm and
+    based on data stretching back to 2007.  This is unlikely to be the
+    result of a clock error, because TESS observations of other hot
+    Jupiters (WASP-6b, 18b, and 46b) are compatible with a constant
+    period, ruling out an 81.6-second offset at the 6.4$\sigma$ level.
+    The 1.3-day orbital period of WASP-4b appears to be decreasing at
+    a rate of $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  The
+    apparent period change might be caused by tidal orbital decay or
+    apsidal precession, although both interpretations have
+    shortcomings.  The gravitational influence of a third body is
+    another possibility, although at present there is no evidence for
+    such a body. Further observations are needed to confirm and
     understand the timing variation.
@@
 Although the Transiting Exoplanet Survey Satellite (TESS) is designed
-to detect new planets, it will also perform precise photometric
+to detect new planets, it is also performing precise photometric
 monitoring of most of the planets that have been discovered by
@@
 Second, long-term timing studies may reveal apsidal precession, {\it
-  i.e.}, the rotation of the orbital ellipse within the orbital
-plane. Apsidal precession has long been observed in eclipsing binaries
+i.e.}, the rotation of the orbital ellipse within the orbital plane.
+Apsidal precession has long been observed in eclipsing binaries
 \citep[{\it e.g.},][]{
@@
 \citealt{kostov_kepler-413b_2014}; Kepler-453
-\citealt{welsh_kepler_2015}, planets orbiting rapidly rotating stars
+\citealt{welsh_kepler_2015}), planets orbiting rapidly rotating stars
 (Kepler-13
-\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
+\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015});
 and systems with multiple interacting giant planets (Kepler-108
@@
 circular, which may itself be a consequence of star-planet tides.  If
-it could be observed, then \citet{ragozzine_probing_2009} showed that
-it would be possible to measure the planet's Love number, which would
-be a valuable constraint on the planet's interior.
+it could be observed, then as \citet{ragozzine_probing_2009} showed it
+would be possible to measure the planet's Love number, which would be
+a valuable constraint on the planet's interior.
 
@@
 distance of 5.5 stellar radii
-\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  It is a good target to
-search for departures from a constant period, because transit data
-exists all the way back to 2007.  The orbital eccentricity is less
-than 0.018 (2$\sigma$), based on the work of
-\cite{knutson_friends_2014}, who combined the available transit times,
-occultation times, and Doppler data.  The stellar obliquity (or at
-least its sky projection) is also compatible with zero, within about
-10 degrees
-\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
+\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
+to search for departures from a constant period, because transit data
+exist all the way back to 2007.  The orbital eccentricity is less than
+0.018 (2$\sigma$), based on the work of \cite{knutson_friends_2014},
+who combined the available transit times, occultation times, and
+Doppler data.  The sky projection of the stellar obliquity
+is also compatible with zero, within about 10 degrees
+\citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011}.
 
 In what follows, \S~\ref{sec:observations} presents the new TESS
-observations, and \S~\ref{sec:timing} describes our timing
-analysis.  We tried fitting the data with three models: a constant
-period, a decaying period, and a slightly eccentric precessing orbit.
-A constant period can be ruled out, but we cannot distinguish between
+observations, and \S~\ref{sec:timing} describes our timing analysis.
+We tried fitting the data with three models: a constant period, a
+decaying period, and a slightly eccentric precessing orbit.  A
+constant period can be ruled out, but we cannot distinguish between
 the latter two possibilities.  Either possibility would have
@@
 Appendix~\ref{sec:verify_tess} considers the possibility that the
-WASP-4 timing anomaly is due to an error in time stamps in the TESS
+WASP-4 timing anomaly is due to an error in timestamps in the TESS
 data products.  We found this possibility to be unlikely because none
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{f1.png}
+		\includegraphics[height=0.97\textheight]{f1.png}
     \end{center}
-    \vspace{-0.5cm}
+    \vspace{-0.8cm}
     \caption{
@@
       curves are best-fit models. The numbers printed next to each
-      lightcurve are the approximate transit times expressed in
-      BJD minus 2{,}457{,}000.  The right side shows the
-      residuals.
+      lightcurve are the approximate transit times expressed in BJD
+      minus 2{,}457{,}000.  The right side shows the residuals.
        \label{fig:lightcurves}
     }
-\end{figure}
-
-\begin{figure}[t]
+\end{figure*}
+
+\begin{figure*}[t]
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{f2.png}
+        \includegraphics[width=0.9\textwidth]{f2.png}
     \end{center}
-    \vspace{-0.5cm}
+    \vspace{-0.7cm}
     \caption{
@@
         panel shows the residuals.  The fit to the phase-folded
-        transit (blue line) is used when measuring mid-transit times for
-        the individual transits (see text).
+        transit (blue line) is used when measuring mid-transit times
+        for the individual transits (see text).
         \label{fig:phasefold}
     }
-\end{figure}
+\end{figure*}
 
@@
 surrounding WASP-4 were averaged into 2-minute stacks by the onboard
-computer.  The data were downlinked via the Deep Space Network, and the
-spacecraft time stamps were calibrated against the ground-station clocks.
-The spacecraft clock times were then transformed by the
+computer.  The data were downlinked via the Deep Space Network, and
+the spacecraft timestamps were calibrated against the ground-station
+clocks.  The spacecraft clock times were then transformed by the
 Payload Operations Center into the {\it Temps Dynamique Barycentrique}
@@
 as Barycentric TESS Julian Dates (BTJD), which is simply the
-barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
+barycentric Julian date minus 2{,}457{,}000.  Lightcurves that were
+flagged as Threshold Crossing Events by the SPOC pipeline were then
 vetted and released by the MIT TESS Science Office to the Mikulski
@@
 We processed the lightcurve as follows.  First, we removed all points
-with non-zero quality flags.  This removed data contaminated by coarse
-spacecraft pointing, cosmic rays, and other annoyances
-\citep{tess_data_product_description_2018}.  We also removed the data
-within the first and last hours of both orbits, because of ramp-like
-systematic effects that appear during those time ranges. We also
-removed the data that might have been adversely affected by ``momentum
-dumps,'' the firing of thrusters and resetting of reaction wheels\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}
-that took place every 2.5 days during sector 2.
-The data during these events were assigned quality flags corresponding
-to ``Reaction Wheel Desaturation Event'' and ``Manual Exclude''.  For
-WASP-4, these flags were simultaneously set for 54 distinct cadences,
-and there were 10 momentum dumps, averaging about 10 minutes of
-flagged data per dump.  Out of caution, we clipped out an additional
-10 minutes before and after every momentum dump.
+with non-zero quality flags.  This removed data that might have been
+adversely affected by ``momentum dumps,'' the firing of thrusters and
+resetting of reaction
+wheels\footnote{\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
+that took place every 2.5 days during sector 2
+\citep{tess_data_product_description_2018}.  The data during these
+events were assigned quality flags corresponding to ``Reaction Wheel
+Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
+were simultaneously set for 54 distinct cadences, and there were 10
+momentum dumps, averaging about 10 minutes of flagged data per dump.
+Out of caution, we clipped out an additional 10 minutes before and
+after every momentum dump.  We also removed the data within the first
+and last hours of both orbits, because of ramp-like systematic effects
+that appear during those time ranges.
 
@@
 normalized the data by dividing out the median flux.  We converted the
-timestamps from BTJD into BJD by adding the appropriate 2457000 day
-offset \citep{tess_data_product_description_2018}.  Many of these and
+timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
+appropriate 2457000 day offset
+\citep{tess_data_product_description_2018}.  Many of these and
 subsequent processing steps were performed using
@@
 polynomials, or Gaussian processes.  Instead, we modeled the
-out-of-transit flux variations simultaneously with the transit parameters,
-as described below.
+out-of-transit flux variations simultaneously with the transit
+parameters, as described below.
 
@@
 \citet{mandel_analytic_2002} and implemented by
-\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular.
+\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular,
+consistent with the limits from radial velocities and occultation
+timing
+\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
 The free parameters were the reference epoch, the planet to star
@@
 To measure the transit times, we returned to the `cleaned' PDC time
-series and fitted the data surrounding each transit separately, with
-four free parameters: the time of mid-transit $t_{\rm tra}$, the
-planet-to-star radius ratio, and the slope and intercept of a linear
-trend to account for any slow variations unrelated to the transit.  We
-fixed the remaining parameters at the values that had been determined
-from the phase-folded TESS lightcurve.  The uncertainty in each data
-point was set equal to the root-mean-square (rms) level of the
-out-of-transit data.
-
-To verify that the measured uncertainties are plausible, we computed
-the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
-transit times.  We found that $\chi^2 = 9.2$, with $n=16$ degrees of
-freedom.  The variance of the $\chi^2$ distribution is $2n$, so we
-would expect $\chi^2 = 16 \pm 5.7$.  Visually inspecting the residuals
-showed that the error variance had been overestimated, so we
-multiplied the measured TESS errors by a factor $f=0.76$, forcing a
-reduced $\chi^2$ of unity.  This lowered the mean uncertainty of the
-transit midtimes from $29.8$ to $22.6$ seconds.  We verified that
-omitting this step did not appreciably alter any of our conclusions.
-
-Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
-best-fit models, and the residuals.  Table~2 reports the mid-transit
-times and their uncertainties.  After binning the residuals to 1-hour
-windows, the lightcurves have an rms scatter of 586\,{\rm ppm}.  The
-pre-launch TESS noise model\footnote{\url{github.com/lgbouma/tnm}} \citep{winn_photonflux_2013,Sullivan_2015}
-would have predicted an error
-budget consisting of the following terms added in quadrature:
-410\,{\rm ppm} from photon-counting noise, 202\,{\rm ppm} from
-detector read noise, and 673\,{\rm ppm} from the zodiacal background
-light.  The level of background light appears to have been
-overestimated.
+series and fitted the data within four transit durations of each
+transit separately. We used four free parameters: the time of
+mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
+slope and intercept of a linear trend to account for any slow
+variations unrelated to the transit.  We fixed the remaining
+parameters at the values that had been determined from the
+phase-folded TESS lightcurve.  The uncertainty in each data point was
+set equal to the root-mean-square (rms) level of the out-of-transit
+data.
+
+To verify that the measured uncertainties are estimated accurately, we
+computed the reduced $\chi^2$ for a linear ephemeris fit to the
+measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
+$n=16$ degrees of freedom.  The variance of the $\chi^2$ distribution
+is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.  Visually
+inspecting the residuals showed that the error variance had been
+overestimated, so we multiplied the measured TESS errors by a factor
+$f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
+uncertainty of the transit midtimes from $29.8$ to $22.6$ seconds.  We
+verified that omitting this step did not appreciably alter any of our
+conclusions.
+
+Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
+transit, the best-fit models, and the residuals.  Table~2 reports the
+mid-transit times and their uncertainties.  After binning the
+residuals to 1-hour windows, the lightcurves have an rms scatter of
+586\,{\rm ppm}.  The pre-launch TESS noise
+model\footnote{\url{github.com/lgbouma/tnm}}
+(\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
+would have predicted an error budget consisting of the following terms
+added in quadrature: 410\,{\rm ppm} from photon-counting noise,
+202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
+zodiacal background light.  The level of background light appears to
+have been overestimated.
 
@@
   $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
+  $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
   $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
@@
   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
+  %Rotation period~[days]                     & $\approx$34            & ---                         & H \\
 {\it Planetary parameters:} & & & \\
@@
   (D) \citet{stassun_accurate_2017}.
-  (E) From functions of stellar, transit or RV parameters as discussed in text.
+  (E) This work, see \S~\ref{sec:system_parameters}.
   (F) \citet{gaia_collaboration_gaia_2018}.
   (G) \citet{stassun_TIC_2018}.
+  %(H) \citet{hoyer_tramos_2013}, from star-spot occultation events.
 }
 
+\vspace{-1cm}
 \end{deluxetable}
@@
 remaining analysis.  Our parameters are in agreement with those of
-previous investigators
-\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
+previous investigators, but have the benefit of incorporating the {\it
+Gaia} parallax
+\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
 By comparing the star's luminosity and spectroscopic parameters with
@@
 	\end{center}
-	\vspace{-0.5cm}
+	\vspace{-0.7cm}
 	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
@@
 precision in their reported transit times was 5.6~seconds.  Since
-those data points carry significant weight in the analysis, we
-consulted with those authors to confirm that the time stamps in their
-data represent mid-exposure times, that the barycentric correction was
-performed correctly, and that the time system of the final results was
-BJD$_{\rm TDB}$.  These same authors also used the same instrument
-(and largely the same software) to analyze other hot Jupiters, none of
-which showed a departure from a constant-period model.
-
-We also compiled all the available occultation times, which are given
-in Table~3.  The tabulated values have been corrected for the
+those data points carry significant weight in the analysis, we checked
+that the timestamps in their data represent mid-exposure times, that
+the barycentric correction was performed correctly, and that the time
+system of the final results was BJD$_{\rm TDB}$.  These same authors
+also used the same instrument and method to analyze other hot
+Jupiters, none of which showed a departure from a constant-period
+model.  Finally, these authors also measured two transit midpoints
+using Spitzer (Baxter et al., in prep); these
+times agree with the Gemini South results.
+
+We also compiled the available occultation times, which are given in
+Table~3.  The tabulated values have been corrected for the
 light-travel time across the diameter of the orbit by subtracting
@@
 the time standard.  We assumed the standard was UTC, and performed the
-appropriate corrections to convert to BJD$_{\rm TDB}$.  We
-also verified that none of our conclusions would be changed if this
-assumption was mistaken. Finally, \citet{zhou_secondary_2015} observed an
-occultation with the Anglo-Australian Telescope.  They did not
-report the observed midpoint, but they did report a result for
-$e\cos\omega$ based upon the observed midpoint.  We calculated the
-implied midpoint using the formula \citep[{\it
-e.g.},][]{winn_exoplanet_2010}
+appropriate corrections to convert to BJD$_{\rm TDB}$.  We verified
+that none of our conclusions would be changed if this assumption was
+mistaken. Finally, \citet{zhou_secondary_2015} observed an occultation
+with the Anglo-Australian Telescope.  They did not report the observed
+midpoint, but they did report a result for $e\cos\omega$ based upon
+the observed midpoint.  We calculated the implied midpoint using the
+formula \citep[{\it e.g.},][]{winn_exoplanet_2010}
 \begin{equation}
@@
 an early stage, we were concerned about a possible offset in the TESS
-time stamps due to an error with the TESS clock or the data processing
+timestamps due to an error with the TESS clock or the data processing
 pipeline.  Appendix~\ref{sec:verify_tess} describes some tests that
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \begin{center}
-    \includegraphics[width=0.385\textwidth]{f5.pdf}
+    \includegraphics[height=0.9\textheight]{f5.pdf}
   \end{center}
-  \vspace{-0.5cm}
+  \vspace{-0.6cm}
   \caption{
@@
   }
-\end{figure}
+\end{figure*}
 
@@
 \begin{equation}
-	Q_\star' \approx 2.9\times10^4.
+  Q_\star' = (2.9 \pm 0.3)\times10^4.
 \end{equation}
@@
 phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
-(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
-quality factor we have inferred from the apparent period change.
+(1.2^{+1.0}_{-0.5})\times10^7$, which would correspond to $\dot{P}
+\approx -30$ microseconds per year.  This strongly disagrees with the
+the period change that we have observed.
 
@@
 Hertzsprung-Russell diagram of hot Jupiters hosts, including WASP-4.
-On the y-axis is $G=g-\mu$, for $g$ the apparent Gaia-band magnitude,
+On the y-axis is $G=g-\mu$, for $g$ the apparent {\it Gaia}-band magnitude,
 and $\mu$ the distance modulus reported by
@@
 where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
-of WASP-4b and the hypothetical WASP-4c, and we have assumed WASP-4b's
-Love number is $k_{2,{\rm b}}\approx 0.6$.  For the RV signal of the
-companion to remain undetected, it would need to be in the residual
-$({\rm O-C})_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
-\citet{triaud_spin-orbit_2010}.  Again following
-\citet{bailey_understanding_2019}, this implies
+of WASP-4b and the hypothetical WASP-4c, and owing to our imprecise
+measurement we have assumed WASP-4b's Love number is $k_{2,{\rm
+b}}\approx 0.6$.  For the RV signal of the companion to remain
+undetected, it would need to be in the residual $({\rm O-C})_{\rm RV}
+= 15.2\,{\rm m s}^{-1}$ reported by \citet{triaud_spin-orbit_2010}.
+Again following \citet{bailey_understanding_2019}, this implies
 \begin{align}
@@
 10^{-5}$.  Hence, this mechanism does not seem capable of producing
-the required value of $\sim$$10^{-3}$.
+the required eccentricity of $\sim$$10^{-3}$.
 
@@
 that the Applegate effect could produce timing deviations of up to 15
-seconds, depending on the modulation period of the stellar dynamo, and
-the corresponding level of differential surface shear.  If this
+seconds, depending on the modulation period of the stellar dynamo.
+If this
 analysis is accurate, then the Applegate mechanism cannot explain the
@@
 \subsection{Other possible explanations}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Further observations will be needed to confirm and understand
+			the timing variations of WASP-4b.} Dots are as in
+		Figure~\ref{fig:times}.  Lines are 100 random draws from the
+		posteriors of the apsidal precession model (orange), and the
+		orbital decay model (blue).    
+		\label{fig:future}
+	}
+\end{figure*}
 
@@
 
-There are two other small effects worth noting.  The first is the
+There are two final small effects worth noting.  The first is the
 \citet{shklovskii_possible_1970} effect due to the star's proper
@@
 \label{sec:future}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-    {\bf Further observations will be needed to confirm and understand
-    the timing variations of WASP-4b.} Dots are as in
-    Figure~\ref{fig:times}.  Lines are 100 random draws from the
-    posteriors of the apsidal precession model (orange), and the
-    orbital decay model (blue).    
-		\label{fig:future}
-	}
-\end{figure}
 
@@
 
-If the TESS mission is extended after its primary mission, it will
-likely observe additional transits of WASP-4b in the early 2020s
+If TESS is extended beyond its primary mission, it will likely observe
+additional transits of WASP-4b in the early 2020s
 (Figure~\ref{fig:future}).  High-precision transit observations with
-larger telescopes would also be useful.  In order to decide between
-orbital decay and apsidal precession, occultation measurements in both
-the near term and also in the mid-2020s will be needed.  More
-radial-velocity data would help in the search for additional bodies
-that could be causing dynamical perturbations, or an overall acceleration
-of the host star.
+larger telescopes would also be useful. 
+In order to decide between orbital decay and
+apsidal precession, occultation measurements in both the near term and
+also in the mid-2020s will be needed.  More radial-velocity data would
+help in the search for additional bodies that could be causing
+dynamical perturbations, or an overall acceleration of the host star.  
+The transit duration variations are expected to be of order 10 seconds
+\citep{pal_periastron_2008}, and so may remain out of reach.
 
@@
 %
-J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
-supporting this work.
-%
 WASP-4 was included on the ``short-cadence'' target list thanks to the
@@
 and G011183 respectively). 
+%
+J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
+supporting this work.
+%
+J.M.D. acknowledges funding from the European Research Council (ERC)
+under the European Union's Horizon 2020 research and innovation
+programme (grant agreement no. 679633; Exo-Atmos), and support from
+the NWO TOP Grant Module 2 (Project Number 614.001.601).
+%
+J.E.R. was supported by the Harvard Future Faculty Leaders
+Postdoctoral fellowship.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Verifying the TESS Time Stamps Using Other Hot Jupiters}
+\section{Verifying the TESS Timestamps Using Other Hot Jupiters}
 \label{sec:verify_tess}
@@
 for $t_{\rm offset}$ a systematic constant offset between the reported
-time stamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
+timestamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
 were
@@
 % 
-% The preceding analysis shows that the TESS time stamps are consistent
+% The preceding analysis shows that the TESS timestamps are consistent
 % at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
@@
 % \citealt{huitson_gemini_2017}, the authors also stand by their
-% time stamps. Since the ETD data come from
-% heterogeneous sources, their time stamps are less clearly documented,
+% timestamps. Since the ETD data come from
+% heterogeneous sources, their timestamps are less clearly documented,
 % and the times are thus more prone to systematic errors, we

### 2019-02-20 review/referee_response_0.txt
add referee report

Dear Dr. Endl,

We thank you for organizing the review of our paper, and the referee
for constructive criticism.  We have considered the referee's remarks
carefully and revised our manuscript accordingly.

Below, we have reproduced the referee's report, along with our
point-by-point responses. A list of changes is also appended to the
manuscript.

Sincerely,

Luke Bouma

********************************************************************************

Referee comment:

The authors present a number of new and high-quality transit observations of
the hot Jupiter WASP-4b observed by TESS. After comparing the predicted
mid-transit times (obtained only considering the pre-existing transits in the
literature) with the measured values from the TESS light curves, the authors
conclude that the TESS transits occur ~ 82 seconds earlier than predicted. They
examine different possibilities and find that the 2 most favored scenarios to
explain this deviation are orbital decay and apsidal precession. However, the
authors caution that more observations are needed to confirm these results.
Also, they redetermine various parameters of the system, finding them in
general agreement with literature values.

TODO TODO TODO
* TODO: add more weight to the distant companion idea.
* ADD: our own analysis of the RV data, and the resulting implied period
  derivative.
* ADD: RV analysis emphasis (!!).

In general, the manuscript is scientifically accurate and the results presented
in it certainly will be of interest to other investigators in
the field. Then, I consider this paper worthy of being published after the
concerns described below be addressed.

Below are my comments, beginning with the major points. I recommend the authors
take them into consideration when revising their manuscript.

-- Major points:

1) In section 3.1, the authors claim "..We included data from peer-reviewed
literature for which the analysis was based on observations of a single,
complete transit,..". However, the transits listed in Table 2 corresponding to
the epochs E=-827 (ttra= 2454697.79817; Winn et al. 2009), E=-804 (ttra=
2454728.57767; Hoyer et al. 2013), E=-537 (ttra= 2455085.88418; Dragomir et al.
2011), E=-208 (ttra= 2455526.16356; Ranjan et al. 2014) are incomplete light
curves, and those from the epochs E=-795 (2454740.62125; Hoyer et al. 2013),
E=-251 (2455468.61943; Hoyer et al. 2013) and E=21 (2455832.61815; Hoyer et al.
2013) present scarcity of after ingress/before egress data-points. On one hand,
given that several partial light curves have been included in their study, the
use of the word "complete" to indicate the characteristics of the transits in
their sample, must be avoided. 
On the other hand, at least two other transits included in their sample show
visible anomalies, probably produced by the pass of the planet in front of one
or several cold spots during transit (E=-561, ttra=2455053.76774 and E=-526,
ttra=2455100.60595, both from Sanchis-Ojeda et al. 2011). Different studies
have proven that mid transit-times measured from partial light curves (Sada et
al. 2012, Barros et al. 2013, Nikolov et al. 2015, Mancini et al. 2018) or
light curves with anomalies due to the presence of spots (Oshagh et al. 2013,
Mazeh et al. 2015, Ioannidis et al. 2016), usually present larger uncertainties
and untrustable values. Taking in consideration the conclusions of these
previous works, it is possible that the exclusion of incomplete/with anomalies
transits might affect the outcome of their analysis and conclusions,
particularly with respect to orbital decay and apsidal precession. Then, in
order to warrant confident results, I would like to see the same analysis but
without including partial transits or with visible spot-crossing events on
them.

TODO TODO TODO
* USE: best-judgement.
* ENSURE: do not include partial transits.
* OMITTING THE ONES WITH VISIBLE SPOT-CROSSINGS: meh. Maybe focus on the ones
  with excess reisduals during partial phases.

2) As mentioned in 1), another possible interpretation for explaining the
deviation from a linear ephemeris of the TESS mid-transit times could be the
presence of anomalies in the light curves due to stellar activity. It has been
well established that WASP-4 is an active star with a rotation period (P) in
between 20-40 days (P ~ 22 days and P ~ 34 days were determined by
Sanchis-Ojeda et al. 2011 and Hoyer et al. 2013, respectively), which is
expected for a main-sequence G7 star. Given that the TESS mid-transits span
around 30 days, wouldn't it be possible that these measurements be affected by
the existence of unseen stellar spots? Regarding this point, I would like to
see an extra paragraph in the Section "Interpretation" with a discussion about
this scenario.

TODO TODO TODO
* ONE IDEA: consider ADDING the spots (e.g., from the hubble data, at the
  observed amplitude). Run it through the lightcurves, over a full transit
  phase. QUANITFY: How much does the transit mid-time change?
* ALSO QUANTIFY: what is the peak-to-peak stellar variability? (Can we confirm
  the previously measured rotation periods?)
* The period is the same order of time as the observations...

-- Minor points:

0. TITLE:

i) As it is, the title implies a very strong asseveration. It would be more
appropriate "Possible Early Arrival of WASP-4b for the TESS Mission" or
something similar.

TODO TODO TODO
* Pending spot studies: no.

1. INTRODUCTION:

i) Please, include here a reference for TESS.

  > Response: Done.

2. NEW TRANSITS AND SYSTEM PARAMETERS:

Section 2.1:

i) It could be informative to provide a reference for the "Deep Space Network".

  > Response: Barring any specific suggestion by the referee, we will leave
  > this one for the reader to pursue at their leisure.

ii) barycentric Julian date --> Barycentric Julian Date

  > Response: Done.

iii) For the casual reader, it would be important to briefly specify what
"Threshold Crossing Events" means, and what the flags "Reaction Wheel
Desaturation Event" and "Manual Exclude" represent. Footnotes including this
information would be appropriate. Also, could the authors describe what a
"Presearch Data Conditioning" lightcurve is?

  > Response: We have clarified that a "Presearch Data
  > Conditioning" lightcurve is at a stage of process in which
  > systematic trends have been removed through the decorrelation
  > process described by Smith et al (2017).
  >
  > We now omit explicit mention of "Treshold Crossing Events" in
  > favor of a more qualitative description.
  >
  > As described in the text, the flags "Reaction Wheel
  > Desaturation Event" and "Manual Exclude" together mean that the
  > data are potentially of poor quality due to "momentum dumps".
  > We referred the reader to the data release notes for an
  > explanation of what a "momentum dump" is, and have added text
  > to this footnote to improve its clarity.

iv) It might be worthwhile to mention what "ramp-like systematic effects" are
and also include a reference.

  > Response: We have changed the wording from "ramp-like
  > systematic effects" to "correlated red noise". No reference is
  > needed.

v) It would be important to include in this section an explanation of how
photometric data-point errors are estimated.

  > Response: Section 2.2, paragraph 3, final sentence explains
  > this: "The uncertainty in each data point was set equal to the
  > root-mean-square (rms) level of the out-of-transit data." We
  > have added the word "photometric" to the above sentence, to
  > clarify that we mean the photometric data points.

Section 2.2:

i) Could the authors explicitly indicate, in the first paragraph, the number of
transits observed by TESS finally used to create the phase-folded light curve?
Furthermore, the manuscript would be more readable if the main text refers
first to Figure 1 instead of Figure 2.

  > Response: We have clarified that the phase-folded lightcurve
  > comes from 18 transits.  We have left the figure ordering
  > as-is.

ii) The authors say that the phase-folded lightcurve was fitted by using the
implementation of Kreidberg (2015) of the analytic model of Mandel & Agol
(2002). Is this referring to the BATMAN code? If it is so, It would be good to
mention the code's name in the text. Also, I would appreciate if, in the main
text, the authors clarify from which source are the values used as initial
photometric parameters (planet to star radius ratio, orbital distance to
stellar radius ratio, inclination, etc). On the other hand, for highlighting
the results obtained with TESS data, it would be important to visualize all the
transit parameters derived from section 2.2 in an independent table.
Furthermore, to see the agreement between the transit parameters reported in
previous studies and those resulting from this work, it would be useful to
present together the values obtained in both cases.

  > Response: We now mention the name "BATMAN". We do the same with
  > "emcee".
  >
  > We did not explicitly describe our choice for the initial
  > parameters, because the converged best-fit values from the
  > chain do not depend on them.  For reproducibility purposes,
  > similar best-fit values can be obtained independent of initial
  > parameter choices. Barring an overruling request from the
  > referee, we will not mention our explicit choice in the text
  > (it was to use values from Hoyer et al., 2013).
  >
  > The transit parameters from the phase-folded fit: Rp/Rstar,
  > inclination, a/Rstar, and limb darkening parameters are given
  > in the first five lines of Table 1.
  >
  > The transit mid-times from the independent fits to each transit
  > are given in Table 2.
  >
  > The parameters which are not reported are the depths of the
  > individual transits, and the linear fit parameters. Since the
  > focus of this study is the transit times, we would prefer to
  > omit these additional parameters for the sake of concision.
  >
  > For the same reason, we leave the interested reader to explore
  > differences in the stellar and planetary parameters by citing
  > relevant work in the last paragraph of Section 2.3. From e.g.,
  > Table 8 of Southworth+2009, we see that our values agree with
  > those of previous studies, and so do not merit further
  > attention, given the main focus of the present work.

iii) In the caption of Figure 2, it would be important to specify which is the
bin size used to determine the yellow points. Furthermore, in Figure 2, the
photometric error of the unbinned data-points should be included.

  > Response: We have added a representative error bar to Figure 2
  > which shows the photometric error of the unbinned data points.
  > 
  > We have opted to not explicitly specify the bin size. It can be
  > inferred from inspection of the plot, and is only relevant for
  > visualization purposes.

iv) To estimate transit times for each individual lightcurve, did the authors
use the same code as for the phase-folded light curve? In this case, how were
the mid-transit times' uncertainties determined?

  > Response: This is discussed in paragraphs 3 and 4 of Section 2.2:

    > """To measure the transit times, we returned to the `cleaned' PDC
    > time series and fitted the data within four transit durations of
    > each transit separately. We used four free parameters: the time of
    > mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
    > slope and intercept of a linear trend to account for any slow
    > variations unrelated to the transit.  We fixed the remaining
    > parameters at the values that had been determined from the
    > phase-folded TESS lightcurve.  The uncertainty in each
    > \added{photometric} data point was set equal to the
    > root-mean-square (rms) level of the out-of-transit data.
    > 
    > To verify that the measured uncertainties are estimated accurately,
    > we computed the reduced $\chi^2$ for a linear ephemeris fit to the
    > measured TESS mid-transit times.  We found that $\chi^2 = 9.2$,
    > with $n=16$ degrees of freedom.  The variance of the $\chi^2$
    > distribution is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.
    > Visually inspecting the residuals showed that the error variance
    > had been overestimated, so we multiplied the measured TESS errors
    > by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This
    > lowered the mean uncertainty of the transit midtimes from $29.8$ to
    > $22.6$ seconds.  We verified that omitting this step did not
    > appreciably alter any of our conclusions."""

  > So yes, we used the same transit model as before (the BATMAN
  > code), but with the addition of the slope and intercept of a
  > linear trend as free parameters.  Our proceduring for
  > determining the mid-transit time uncertainties is also
  > described.

v) Is χ^2=9.2 the reduced chi-square? If it is so, It should be clarified by
using a different symbol or by adding a sub-index, for example. Please, fix
this in the rest of the manuscript when necessary.

  > Response: $\chi^2 = 9.2$ is the $\chi^2$ value. We have
  > corrected the typo in the text, in which we wrote "reduced
  > $\chi^2$'.

vi) χ^2 = 16 {plus minus} 5.7 --> χ^2 = 9.2 {plus minus} 5.7

  > Response: For 18 transits, and 2 degrees of freedom, we
  > would expect $\chi^2 = 16 \pm 5.7$.

vii) The note at the end of Table 2 explains what the parameter sigmat0
represents, but it's not clear which is its relation with the mid-transit
times' measurements presented in the same table.

  > Response: This was a typo -- $\sigma_{t_0}$ should have read
  > $\sigma_{t_{\rm tra}}$. We have added a sentence to summarize
  > the complete procedure in the note at the end of Table 2.

Section 2.3:

i) The authors should include the references for each of the mentioned catalogs
and also for the Yonsei Yale isochrones.

  > Response: Done, thank you.

ii) How did the authors estimate the errors of the stellar and planet
parameters?

  > Response: We have added the following sentence in the last
  > paragraph of Section 2.3: "The uncertainties in our derived
  > stellar and planetary parameters are propagated according to
  > standard analytic formulae, under the assumption of
  > uncorrelated and normally distributed initial measurements."

3. TIMING ANALYSIS

Section 3.1:

i) The sentence "..Since those data points carry significant weight in the
analysis, we checked that the timestamps in their data represent mid-exposure
times, that the barycentric correction was performed correctly, and that the
time system of the final results was BJD(TDB).." is confusing, because seems to
imply that these checks were only performed on the Huitson et al. mid-transit
times and not in all the literature mid-transit times. It would be
recommendable to clarify this point.

  > Response: we have modified the wording slightly to clarify. The point is
  > that for these Gemini South observations, we confirmed with the authors
  > through back and forth correspondence that the details of their
  > time-keeping were sound.  We made this extra effort because of the weight
  > of the Huitson+(2017) times in the analysis.
  >
  > For the remaining data points, we used the references stated in Section 3.1
  > and Table 2 to confirm that we were using with mid-exposure times, and
  > corrected from HJD_UTC to BJD_TDB as appropriate. However the point is
  > indeed that because of the weight of the Huitson et al (2017) times, we
  > made an particular extra effort to confirm their accuracy with the original
  > authors.

ii) Could the authors specify what is the meaning of each of the variables in
Equation (1)?

  > Response: Done.

Section 3.2:

i) How did you fit the mid times of the preTESS measurements? Did you use a
weighted least-square fit?

  > Response: Yes, this was through weighted least-squares. We have changed the
  > wording accordingly.

ii) Figure 3:

I would recommend changing the "title" indicated in bold font in the caption,
by a more explicit sentence about what is shown in the plot. To give an
example, a possibility could be "Observed minus predicted mid-transit times".

  > Response: Though we appreciate the suggestion, we have opted to keep the
  > "title" as is. The nuance of what "predicted" means for this plot (a linear
  > ephemeris, based only on the pre-TESS times) is sufficient that we find it
  > to be better left in the body of the caption.

Also: ii-1) In the legends of both panels, could the authors specify which is
the bin size of the TESS light curves? (Please, fix this in all the
manuscript).

  > Response: We have added the following sentence in the caption: "The binned
  > TESS point is the weighted average of 18 TESS transits".  The caption of
  > Figure 4 already stated the same.  We have changed the words "Dots" to
  > "Symbols" in the caption of Figure 6 in order to clarify this for the
  > reader.

ii-2) "The red band shows the average deviation of the TESS transits,..." -->
"The red band shows the average deviation of the TESS transits {plus
minus}1σ,..."

  > Response: Done.

iii) It would be important to introduce references for equations (4) and (5).
Also, for the casual reader, It might be informative to give a brief
explanation of what sidereal and anomalistic periods are.

  > Response: Equations (4) and (5) are generic equations for the Taylor series
  > expansion of the transit and occultation times as a function of epoch, and
  > do not require references.  We have added the following sentence to describe the
  > sidereal and anomalistic periods: "The sidereal period is the duration
  > required to return to the same orientation with respect to the stars; the
  > slightly longer anomalistic period is the duration required to reach a
  > fixed longitude with respect to the rotating line of apsides."

iv) Figure 4: In the caption, could the authors explicitly indicate what the
plot shows in each of its panels (top and bottom)?

  > Response: Done.

v) Similar to i) How did you fit the mid-transit times and occultations shown
in Figure 4?

  > Response: We have added the following sentence in Section 3.2:
  > "We fit each model by assuming a Gaussian likelihood and
  > sampling over the posterior probability distributions."

vi) In the caption of Table 4, please be more explicit, "Best-fit model
parameters" of what?

  > Response: The caption now reads "Best-fit transit timing model
  > parameters".

vii) Given that the mid-transit times values used in this work are taken from
different authors with different methodologies to perform the fitting and
detrending of the light curves, could the authors mention how this could affect
their results?

TODO TODO TODO
* We are trusting that people fit transits.
* (Maybe cite some of the Espinoza studies about detailed transit fitting
  systematics)

********************************************************************************
4. INTERPRETATION

Referee comment:

Section 4.1:

i) First line: The word "caused" is repeated twice.

  > Response: Corrected.

ii) In the caption of Figure 5 should be indicated what open circles represent.

  > Response: Done.

iii) The characteristic timescale value calculated in equation (13), is it
computed by considering a theoretical dP/dt value or from the one estimated
through the observations?

  > Response: Changed wording to remove any ambiguity.

iv) The authors mention that there are about 20 Hot Jupiters for which the
theoretical timescale is shorter, where are the timescale values obtained from?
Are they estimated from equation (13)? If this is the case, how do the
theoretical dP/dts are computed?

  > Response: As noted in the caption of Figure 5, the decay timescale values
  > are obtained from Equation 15, assuming Qstar'=10^7.  We have slightly
  > modified the text for clarity on this point.

v) In the top panels of Figures 4 and 6, the orbital decay model evolution
shows that at the beginning the O-C values increase until, at some point, they
reach a maximum and after that, they start decreasing. What is the physical
explanation for that behavior? If the orbital period is diminishing, wouldn't
it be expected to see only a decreasing in the O-C values? Why the O-C values
would increase at these first epochs?

  > Response: To make Figure 4 and 6, we start with the observed transit and
  > occultation mid-times, as a function of epoch number.  We then fit a line
  > to these data.  If the data were generated by a constant period orbit (with
  > gaussian uncertainties) the residuals to such a fit would show no
  > correlation with epoch number.  However if the data were generated by a
  > planet with a constant period derivative (with gaussian uncertainties), the
  > residuals to such a fit would always be a parabola.  This is because the
  > linear fit to the data will have contributions from the beginning, middle,
  > and end of the transit timeseries. Thus the best-fit linear ephemeris (to
  > data that actually have a decreasing period) has a best-fit period
  > somewhere between the minimum and maximum period. At the earliest epochs,
  > the best-fit period is an underestimate, so you see the increase in O-C. At
  > the latest epochs, the best-fit period is an overestimate, so you see the
  > decrease.

Section 4.2:

i) It might be informative to explicitly include the equation (14) of Ragozzine
& Wolf (2009) to see of which parameters is depending on.

  > Response: we have opted against this suggestion, chiefly to encourage the
  > interested reader to refer to the original study!

ii) Similar to Figure 3, I would recommend changing the title indicated in bold
font in Figure 6, by a more explicit sentence about what is shown in the plot.
For eample "Different apsidal precession and orbital decay models for the O-C
values" or something like that.

  > Response: As in Figure 3, we appreciate the suggestion, but have opted to
  > keep the "title" as is. We have appended the sentence "The two models may
  > begin to diverge in the mid-2020s" as further clarification for the reader.

iii) Mb is indicated as the mass of the hypothetical planet WASP-4c, however,
the Mb symbol is not present in equation (19).

  > Response: We have correcly the wording accordingly.

iv) Why do the authors assume k2b = 0.6 for the Love number? Is this value
randomly chosen?

  > Response: We have added the phrase "similar to Jupiter". Section 4.2
  > already notes that our measured value for k_2p under the precession
  > hypothesis (Equation 18) is imprecise.

v) Would it be possible to introduce a definition of the reduced mass presented
in equation 21? Also, could the authors specify which values they used for L,
Mc, and Rc and provide references?

  > Response: We have clarified the origin of the values used when evaluating
  > Equation 21.  The tricky ones -- the mass and radius of the convective
  > envelope -- came from the MESA/MIST projects, for which we have added
  > appropriate citations.  Our thanks to the referee for pointing out this
  > omission.  We have opted to not explicitly define the reduced mass, as it
  > is standard knowledge in the gravitational two-body problem.

********************************************************************************

Referee comment:

i) "WASP-4 timing anomaly" --> "WASP-4b timing anomaly"

  > Response: Corrected.

ii) In the caption of Figure 7: "with standard deviation (σpredicted)" -->
"with standard deviation (σpre-TESS)"

  > Response: Corrected.

### 2019-02-20 paper/ms.tex
remove unused ETD appendix
WIP referee report update to paper
almost all of the minor changes. major suggestions still to be
incorporated, with the additional analysis of the RV data.

@@
-\documentclass[12pt,twocolumn,tighten]{aastex62}
+\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 %\pdfoutput=1 %for arXiv submission
@@
 %\NewPageAfterKeywords
+\listofchanges
 
@@
 %
-% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % WAITING
+% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % 
+%%%WAITING (LIKELY YES)
 % \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
@@
 %
-% \author{M. Fausnaugh} % faus@mit.edu WAITING
-% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
-% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-% 02139, USA }
+\added{\author{M. Fausnaugh} % faus@mit.edu CONFIRMED
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }}
 %
@@
 %
-% \author{N. Guerrero} % nmg@mit.edu WAITING
-% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
-% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-% 02139, USA }
+\added{
+\author{N. Guerrero} % nmg@mit.edu CONFIRMED
+\affiliation{ Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA }
+}
 %
@@
 %
-% \author{B. Wohler} % bill.wohler@nasa.gov WAITING
-% \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-% \affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-%
+\added{
+\author{B. Wohler} % bill.wohler@nasa.gov YES
+\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
+\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
+}
 
@@
     shortcomings.  The gravitational influence of a third body is
-    another possibility, although at present there is no evidence for
+    another possibility, although at present there is \replaced{no}{minimal} 
+    evidence for
     such a body. Further observations are needed to confirm and
@@
 
-Although the Transiting Exoplanet Survey Satellite (TESS) is designed
+Although the Transiting Exoplanet Survey Satellite (TESS, \added{
+\citealt{ricker_transiting_2015}}) is designed
 to detect new planets, it is also performing precise photometric
@@
     \caption{
-        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
-        TESS data. Yellow points are binned measurements.  The bottom
+        {\bf Phase-folded lightcurve of WASP-4b.} Gray points are
+        TESS \replaced{data}{flux measurements, with median 
+        $1\sigma$ error shown in the lower 
+        right.} 
+        Yellow points are binned measurements.  The bottom
         panel shows the residuals.  The fit to the phase-folded
         transit (blue line) is used when measuring mid-transit times
-        for the individual transits (see text).
+        \replaced{for the}{of} individual transits (see 
+        \S~\ref{sec:measurement}).
         \label{fig:phasefold}
@@
 as Barycentric TESS Julian Dates (BTJD), which is simply the
-barycentric Julian date minus 2{,}457{,}000.  Lightcurves that were
-flagged as Threshold Crossing Events by the SPOC pipeline were then
+\replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus 2{,}457{,}000.  
+Lightcurves that 
+were
+flagged \deleted{as Threshold Crossing Events} by the SPOC pipeline \added{as 
+crossing a transit detection threshold} were then
 vetted and released by the MIT TESS Science Office to the Mikulski
@@
 We began our analysis with the Presearch Data Conditioning (PDC)
-lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
-We processed the lightcurve as follows.  First, we removed all points
+lightcurve\replaced{ 
+\citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}{, which has had 
+systematic trends removed 
+through the decorrelation process described by 
+\citet{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}.
+We \added{then} processed the lightcurve as follows.  First, we removed all 
+points
 with non-zero quality flags.  This removed data that might have been
@@
 resetting of reaction
-wheels\footnote{\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
+wheels\footnote{\added{Details concerning the spacecraft pointing and momentum 
+dumps are described in the data release 
+notes:}\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
 that took place every 2.5 days during sector 2
-\citep{tess_data_product_description_2018}.  The data during these
+\deleted{\citep{tess_data_product_description_2018}}.  The data during these
 events were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
+Desaturation Event'' and ``Manual Exclude'' 
+\added{\citep[][Table 28]{tess_data_product_description_2018}}.  For WASP-4, 
+these flags
 were simultaneously set for 54 distinct cadences, and there were 10
@@
 after every momentum dump.  We also removed the data within the first
-and last hours of both orbits, because of ramp-like systematic effects
-that appear during those time ranges.
+and last hours of both orbits, because of \replaced{ramp-like systematic 
+effects that appear}{correlated red noise that appears} during those time 
+ranges.
 
@@
 
-Using the cleaned PDC lightcurve, we used the Box Least Squares
+Using the cleaned PDC lightcurve, we applied the Box Least Squares
 algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
@@
 flux measurements surrounding each transit, and divided it out.  We
-then created a phase-folded lightcurve and fitted a standard transit
+then created a phase-folded lightcurve \added{from all 18 transits}
+and fitted a standard transit
 model using the analytic formulae given by
 \citet{mandel_analytic_2002} and implemented by
-\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular,
+\replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
+We assumed 
+the orbit to be circular,
 consistent with the limits from radial velocities and occultation
@@
 \citet{goodman_ensemble_2010} and implemented by
-\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
+\replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
+  Table~1 gives the 
+results, which
 are in reasonable agreement with the parameters reported by {\it
@@
 parameters at the values that had been determined from the
-phase-folded TESS lightcurve.  The uncertainty in each data point was
-set equal to the root-mean-square (rms) level of the out-of-transit
-data.
+phase-folded TESS lightcurve.  The uncertainty in each \added{photometric} 
+data point was set equal to the root-mean-square (rms) level of the 
+out-of-transit data.
 
 To verify that the measured uncertainties are estimated accurately, we
-computed the reduced $\chi^2$ for a linear ephemeris fit to the
+computed the \deleted{reduced} $\chi^2$ \added{value} for a linear ephemeris 
+fit to the
 measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
@@
 Table~1 gives the resulting parameters, which we adopted for the
-remaining analysis.  Our parameters are in agreement with those of
+remaining analysis.
+\added{The uncertainties in our derived stellar and planetary parameters
+    are propagated according to standard analytic formulae, under the 
+    assumption of uncorrelated and normally distributed initial measurements.}
+Our \added{system} parameters are in agreement with those of
 previous investigators, but have the benefit of incorporating the {\it
@@
 that WASP-4 is a main-sequence star with an age of approximately
-$7\,{\rm Gyr}$.
+$7\,{\rm Gyr}$ \added{\citep{demarque_y2_2004}}.
 
@@
       calculated transit times, where the calculation is based only on
-      the pre-TESS data and assumes a constant period.  The blue bands
+      \deleted{the} pre-TESS data and assumes a constant period.  The blue 
+      bands
       depict the $\pm$$1\sigma$ credible interval of the predicted
       times.  {\it Top:} The full timing dataset spans 11 years. The
-      darkest points correspond to the most precise data.  {\it
+      darkest points correspond to the most precise data. \added{The binned 
+      TESS point is the weighted average of 18 TESS transits.}  {\it
         Bottom:} Close-up of the TESS observations. The red band shows
-      the average deviation of the TESS transits, which arrived $81.6 \pm
-      11.7\ {\rm seconds}$ earlier than predicted.
+      the average deviation of the TESS transits \added{($\pm 1\sigma$)}, 
+      which arrived $81.6 \pm 11.7\ {\rm seconds}$ earlier than predicted.
         \label{fig:arrived_early}
@@
 	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-		The residuals are the observed times minus the calculated times
-		assuming a constant period.  The darkest points correspond to
-		the most precise data. The constant-period model (gray line) is
+		\replaced{The residuals are the observed times minus the calculated 
+		times assuming a constant period. The darkest points correspond to
+        the most precise data.}{The vertical axis shows the 
+		observed times minus the calculated times assuming a constant period 
+		for transits ({\it top}) and occultations ({\it bottom}). Darker 
+		points correspond to more precise data.}
+        The constant-period model (gray line) is
 		a poor description of the data.  Models with a uniformly
 		decreasing period (blue) or apsidal precession (orange) provide
-		better fits to the data.  The red square represents the
+		better fits\deleted{ to the data}.  The red square represents the
 		combination of all the TESS data and is for display purposes
@@
 precision in their reported transit times was 5.6~seconds.  Since
-those data points carry significant weight in the analysis, we checked
+these data points carry significant weight in the analysis, we 
+\replaced{checked}{corresponded with the authors to confirm}
 that the timestamps in their data represent mid-exposure times, that
@@
   t_0 +  P E  +
-  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
+  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right),
   \label{eq:occultation_time}
 \end{equation}
-In total, there are four occultation times.
+\added{for $E$ the transit number, $t_0$ the reference epoch, 
+$e$ the eccentricity, and $\omega$ the argument of pericenter.}
+In total, there are four \added{available} occultation times.
 
@@
 
-First, we fitted a constant-period model (``linear ephemeris'') to the
+First, we \replaced{fitted}{performed a weighted least-squares fit of} a 
+constant-period model (\added{a} ``linear ephemeris'') to the
 pre-TESS data, and used it to extrapolate to the epochs of the TESS
@@
 $P_{\rm a}$ is the anomalistic period, and $\omega$ is the argument of
-pericenter.  In this model the angular velocity of the line of apsides
+pericenter.
+In this model the angular velocity of the line of apsides
 ${\rm d}\omega/{\rm d}E$ is constant,
@@
 \end{equation}
+\added{The sidereal period is the duration required to return to the 
+    same orientation with respect to the stars; the slightly longer 
+    anomalistic period is the duration required to reach a fixed 
+    longitude with respect to the rotating line of apsides.}
 The five free parameters of this model are $(t_0, P_{\rm s}, e,
@@
 
+\added{We fit each model by assuming a Gaussian likelihood and sampling over 
+the posterior probability distributions.}
 Figure~\ref{fig:times} shows the residuals with respect to the
@@
   \begin{center}
-    \includegraphics[height=0.9\textheight]{f5.pdf}
+    \includegraphics[height=0.88\textheight]{f5.pdf}
   \end{center}
@@
     colored symbols highlight WASP-4, WASP-12 (which also shows
-    evidence for a decreasing period), and the collection of hot
-    Jupiters analyzed in Appendix A.
+    evidence for a decreasing period), and the \deleted{collection of}
+    hot Jupiters analyzed in Appendix A.
     {\bf \it Top:}
-    The key parameters relevant to eccentricity damping.  Solid
-    squares represent planets with a securely detected nonzero
-    eccentricity. Lines of constant damping timescale are drawn based
-    on Equation~\ref{eq:de_dt} and assuming $Q_{\rm p}' = 10^5$.
-    WASP-4b has one of the shortest eccentricity damping timescales.
+    Key parameters relevant to eccentricity damping.  Solid squares
+    represent planets with a securely detected nonzero eccentricity.
+    \added{Circles represent planets whose orbits are consistent with 
+    circular.}
+    Lines of constant damping timescale are drawn based on
+    Equation~\ref{eq:de_dt} and assuming $Q_{\rm p}' = 10^5$.  WASP-4b
+    has one of the shortest eccentricity damping times.
     {\bf \it Middle:} 
-    Plotted are the key parameters relevant to orbital decay.  Lines
-    of constant decay timescale are drawn based on
-    Equation~\ref{eq:da_dt}, assuming $Q_\star' = 10^7$.  WASP-4b has
-    a relatively short orbital decay timescale, although it is not as
-    extreme a case in this regard as it is for eccentricity damping.
+    Key parameters relevant to orbital decay.  Lines of constant decay
+    timescale are drawn based on Equation~\ref{eq:da_dt}, assuming
+    $Q_\star' = 10^7$.  WASP-4b has a relatively short orbital decay
+    timescale, although it is not as extreme a case in this regard as
+    it is for eccentricity damping.
     {\bf \it Bottom:}
@@
 
-If the timing variation is being caused caused entirely by orbital
-decay, then the characteristic timescale of the decay is
+If the timing variation is \deleted{being caused} caused entirely by
+orbital decay, then \replaced{the}{the best-fit model parameters yield a}
+characteristic \replaced{timescale of the decay is}{decay timescale of}
 \begin{equation}
@@
 M_\star/M_{\rm p}$ and $a/R_\star$.  WASP-4 has one of the shortest
-theoretical timescales for orbital decay.  There are about 20 hot
+theoretical timescales for orbital decay.  \replaced{There 
+are}{Figure~\ref{fig:context} shows} about 20 hot
 Jupiters (including WASP-12) for which the theoretical timescale is
-shorter.  In almost all of those cases, though, the planet was
+shorter.  In almost all of those cases, though, the 
+planet was
 discovered more recently than WASP-4 and a decade-long baseline of
@@
   M_{\rm c} > 7.5\,M_\oplus
-  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
+  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2}.
   \label{eq:kozai_bound}
 \end{equation}
-where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
-of WASP-4b and the hypothetical WASP-4c, and owing to our imprecise
-measurement we have assumed WASP-4b's Love number is $k_{2,{\rm
-b}}\approx 0.6$.  For the RV signal of the companion to remain
+\replaced{where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major 
+axis of WASP-4b and the hypothetical WASP-4c, and owing}{
+where $a_{\rm b}$ is the semi-major axis of WASP-4b, $M_{\rm c}$
+is the mass of the hypothetical WASP-4c, and $a_{\rm c}$ is WASP-4c's 
+semi-major axis. Owing}
+to our imprecise
+measurement\added{, in Equation~\ref{eq:kozai_bound}} we have assumed 
+WASP-4b's Love number is $k_{2,{\rm
+b}}\approx 0.6$\added{, similar to Jupiter}.  For the RV signal of the 
+companion to remain
 undetected, it would need to be in the residual $({\rm O-C})_{\rm RV}
@@
 mass, $n$ is the orbital frequency, and $a$ is the semi-major axis.
-For WASP-4, this equation implies $\langle e^2 \rangle^{1/2} \lesssim
+\added{For the luminosity, reduced mass, and semi-major axis, we used values 
+from Table~1, combined with the Stefan-Boltzmann law and standard 
+definitions. To estimate the 
+width and mass of the convective region we ran the MESA code
+for a star with mass and metallicity matched to WASP-4, and the 
+input physics detailed in the MIST isochrones project
+\citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
+We identified the tachocline boundary using the mixing 
+types specified in the resulting radial profiles, and found $R_{\rm conv} 
+\approx 0.33 R_\odot$, and $M_{\rm conv} \approx 9\times10^{-4} M_\odot$. }
+For WASP-4, this implies $\langle e^2 \rangle^{1/2} \lesssim
 10^{-5}$.  Hence, this mechanism does not seem capable of producing
@@
 		{\bf Further observations will be needed to confirm and understand
-			the timing variations of WASP-4b.} Dots are as in
+			the timing variations of WASP-4b.} \replaced{Dots}{Symbols} are as 
+			in
 		Figure~\ref{fig:times}.  Lines are 100 random draws from the
@@
 		orbital decay model (blue).
+        \added{The two models may begin to diverge in the mid-2020s}.    
 		\label{fig:future}
@@
   \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
+  \texttt{MESA} 
+  \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
   \texttt{numpy} \citep{walt_numpy_2011}, 
@@
 		is a gaussian centered on zero with standard deviation
-		($\sigma_{\rm predicted}$) calculated from the pre-TESS transit
+		(\replaced{$\sigma_{\rm predicted}$}{$\sigma_{\rm pre-TESS}$}) calculated from the pre-TESS transit
 		times.  The blue distribution is a gaussian centered on the
@@
 
-An obvious concern that one might have about the WASP-4 timing anomaly
-is that there might be a systematic offset between the TESS time
-system and the time system in which the previous observations have
-been reported.  There is a precedent for this type of error: data from
-the Kepler mission was afflicted by a systematic timing error that was
-corrected only late in the mission~\citep[][Section
-  3.4]{kepler_DR19_2013}.
+An obvious concern that one might have about the WASP-4\added{b}
+timing anomaly is that there might be a systematic offset between the
+TESS time system and the time system in which the previous
+observations have been reported.  There is a precedent for this type
+of error: data from the Kepler mission was afflicted by a systematic
+timing error that was corrected only late in the
+mission~\citep[][Section 3.4]{kepler_DR19_2013}.
 
@@
 
-% \section{Verifying the WASP-4 Archival Times}
-% \label{sec:verify_archival_times}
-% 
-% \begin{figure*}[ht!]
-%   \begin{center}
-%     \leavevmode
-%     \includegraphics[width=0.9\textwidth]{f7.pdf}
-%   \end{center}
-%   \caption{
-%     {\bf Extra times for WASP-4b neither confirm nor refute the
-%     	period-decay.}
-%     Our timing analysis (Section~\ref{sec:timing}) imposes stringent
-%     selection criteria for lightcurves.
-%     Here we overplot additional points from the Exoplanet Transit
-%     Database (ETD), and the discovery epoch from \citet{wilson_wasp-4b_2008}.
-%     \label{fig:verify}
-%   }
-% \end{figure*}
-% 
-% The preceding analysis shows that the TESS timestamps are consistent
-% at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
-% Another explanation for the WASP-4b timing variation could be
-% incorrect archival times.
-% As a sanity check on the times we included for our analysis in
-% Table~1, we collected times from the Exoplanet Transit Database
-% \citep[ETD; ][]{poddany_ETD_2010}.  We considered only the 12
-% lightcurves of the highest self-reported quality (${\rm DQ} = 1$).  We
-% then inspected each lightcurve by eye.  If the lightcurve showed
-% substantial red noise, we discarded it.  This left 9 midtimes,
-% reported in ${\rm HJD}_{\rm UTC}$. We converted the times to ${\rm
-% BJD}_{\rm TDB}$ \citep{eastman_achieving_2010}.
-% 
-% The resulting times are shown in Figure \ref{fig:verify}.  Focusing on
-% the epochs that overlap with \citet{huitson_gemini_2017}'s study, four
-% of six possible times agree with the \citet{huitson_gemini_2017}
-% times; the other two fall about 90 seconds early.  With their stated
-% uncertainties, these 6 ETD data points are inconsistent with a
-% constant period.  The 6 points came from 5 different observers.
-% Systematic errors in absolute clock offsets by different observers
-% should be uncorrelated; if there were such systematic errors in the
-% ETD times, this could explain the many-sigma outliers.  Though the
-% majority of the reported times do support the accuracy of the
-% \citet{huitson_gemini_2017} times, a systematic procedure would call
-% for averaging the ETD times.  If we did this, and discarded the four
-% \citet{huitson_gemini_2017} epochs, it is likely that the evidence for
-% period decay would become much weaker.  Since
-% \citet{huitson_gemini_2017} documented their time system clearly, and
-% produced lightcurves of extremely high quality, we included their times
-% in our analysis.  From private correspondence with
-% \citealt{huitson_gemini_2017}, the authors also stand by their
-% timestamps. Since the ETD data come from
-% heterogeneous sources, their timestamps are less clearly documented,
-% and the times are thus more prone to systematic errors, we
-% categorically omit them from consideration.
-% 
-% We also overplotted the discovery epoch reported by
-% \citet{wilson_wasp-4b_2008}. 
-% This epoch folded together WASP-S survey data, a partial FTS
-% lightcurve, and a complete EulerCam lightcurve. 
-% Though not explicitly stated by \citet{wilson_wasp-4b_2008}, the
-% archival WASP-S lightcurves are in ${\rm HJD}_{\rm UTC}$, so they can
-% be converted to ${\rm BJD}_{\rm TDB}$ without uncertainty in the
-% absolute time reference (Collier-Cameron, priv.\ comm.).
-% The discovery epoch falls about 2 minutes earlier than the epochs we
-% used for the fit, and is visible in the lower left of Figure
-% \ref{fig:verify}.
-
-% Another independent line of evidence supporting the
-% \citet{huitson_gemini_2017} times is that \citet{patra_2017} showed
-% for WASP-12b that precise transit times reported by
-% \citet{stevenson_transmission_2014} from the Gemini telescopes with
-% GMOS are congruent with those from other observers.  This indicates
-% that the telescopes do not have a long-standing clock offset.
-
+\newpage
+
+\listofchanges
 \end{document}

### 2019-02-24 paper/ms.tex
minor paper updates

@@
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%
+\added{
+\author[0000-0002-8958-0683]{F. Dai}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+}
 %
@@
 \added{We fit each model by assuming a Gaussian likelihood and sampling over 
-the posterior probability distributions.}
+the posterior probability distributions. The prior for the quadratic
+model allowed the period derivative to have any sign.}
 Figure~\ref{fig:times} shows the residuals with respect to the
@@
 $\chi^2 = 174$ and 61 degrees of freedom.  The best-fitting
-decreasing-period model has $\chi^2 = 62.6$ and 60 degrees of freedom.
+\replaced{decreasing-period}{quadratic} model has $\chi^2 = 62.6$ and 60 degrees of freedom.
 The best-fitting precession model has $\chi^2 = 64.3$ and 58 degrees
@@
 
-The decreasing-period model provides a slightly better fit to the data
+The \replaced{decreasing-period}{quadratic} model provides a slightly better fit to the data
 than the precession model.  It is favored by $\Delta \chi^2 = 1.7$,
@@
 
-In the decreasing-period model, the period derivative is
+In the \replaced{decreasing-period}{quadratic} model, the period derivative is
 \begin{equation}
@@
 grateful to the people who have turned TESS from an idea into reality.
+\added{L.G.B. thanks A.~Bixel and E.~May for clarifying details
+concerning the available IMACS lightcurves.}
 %

### 2019-02-24 review/referee_response_0.txt
minor response update

@@
 
-  > Response: We have added the following sentence in Section 3.2:
-  > "We fit each model by assuming a Gaussian likelihood and
-  > sampling over the posterior probability distributions."
+  > Response: We have added the following sentences in Section 3.2:
+  > "We fit each model by assuming a Gaussian likelihood and sampling over the
+  > posterior probability distributions. The prior for the quadratic model
+  > allowed the period derivative to have any sign." We have clarified the
+  > wording throughout from "decreasing-period model" to "quadratic model" to
+  > reflect the latter point.
 
@@
   > Response: Corrected.
+
+********************************************************************************
+Other changes:
+
+* Added the following coauthors: Fei Dai, Natalia Guerrero, Bill Wohler,
+  Douglas Caldwell.

### 2019-02-24 paper/ms.tex
address "major comment" #1 from referee

@@
 data from peer-reviewed literature for which the analysis was based on
-observations of a single, complete transit, and for which the midpoint
-was allowed to be a free parameter. We also required that the time
-system be clearly documented. Many of the times were previously
-compiled by \citet{hoyer_tramos_2013}. We confirmed that the times in
-that paper were in agreement with the original sources and that
-barycentric corrections had been performed when needed.
+observations of a single\deleted{, complete} transit, and for which
+the midpoint was allowed to be a free parameter. We also required that
+the time system be clearly documented. Many of the times were
+previously compiled by \citet{hoyer_tramos_2013}. We confirmed that
+the times in that paper were in agreement with the original sources
+and that barycentric corrections had been performed when needed.
 
@@
   e = (1.92^{+ 1.93}_{- 0.76})\times10^{-3}
+  \label{eq:e_obs}
 \end{equation}
@@
 intervals are reported in Table~4.
+
+\added{
+To assess the robustness of these results, we considered a few
+possible systematic effects in the timing dataset.  One concern in the
+pre-TESS timing measurements was incomplete phase coverage: a handful
+of the transit times in Table~2 are from lightcurves with partial
+gaps.  A separate issue is the effect of spot-crossing anomalies on
+mid-transit time measurements.  Both effects lead to lightcurve
+asymmetries which skew transit-time measurements.
+
+We repeated the analysis desribed above, but omitted epochs -827,
+-804, -537, and -208 because of gaps in their coverage.  We also
+omitted epochs -526 and -561 because of visible spot anomalies during
+the transits. (All epoch numbers are as in Table~2.)  The resulting
+best-fit transit timing model parameters were all within $1\sigma$ of
+the values quoted in Table~4.  The confidence intervals,
+goodness-of-fit statistics $\chi^2$, and model comparison statistics
+$(AIC, BIC)$ did not appreciably change.
+}
 

### 2019-02-24 review/referee_response_0.txt
ditto

@@
 
-TODO TODO TODO
-* USE: best-judgement.
-* ENSURE: do not include partial transits.
-* OMITTING THE ONES WITH VISIBLE SPOT-CROSSINGS: meh. Maybe focus on the ones
-  with excess reisduals during partial phases.
+  > Response: We thank the referee for this suggestion.  We have removed
+  > the word "complete" when describing the transit timing dataset.  We
+  > have performed the check requested: We repeated the analysis, but
+  > omitting epochs -827, -804, -537, and -208 because of gaps in their
+  > coverage.  We also omitted epochs -526 and -561 because of the
+  > visible spot anomalies during the transits.
+  > 
+  > The resulting best-fit transit timing model parameters were all
+  > within $1\sigma$ of the values quoted in Table 4.  The confidence
+  > intervals, goodness-of-fit statistics $\chi^2$, and model comparison
+  > statistics $(AIC, BIC)$ did not appreciably change. 
+  > 
+  > The most important of these statistics are summarized below for the
+  > referee's inspection, and this exercise is described in the text
+  > at the end of Section 3.
+  > 
+  > ORIGINAL NUMBERS:
+  > ##########################################
+  > 
+  > quadratic model:
+  > dP/dt = -4.001e-10 +(3.748e-11) -(3.812e-11)
+  > dP/dt = -12.62 +(1.18) -(1.20) millisec/yr
+  > 
+  > -----chi2 values for best fit parameters-----
+  > 
+  > linear model:	    chi2: 174.446	dof: 61	AIC: 1.78e+02	BIC: 1.83e+02
+  > quadratic model:	chi2: 62.5811	dof: 60	AIC: 6.86e+01	BIC: 7.50e+01
+  > precession model: chi2: 64.329	dof: 58	AIC: 7.43e+01	BIC: 8.50e+01
+  > 
+  > -----quad vs precession-----
+  > delta_AIC = AIC_prec-AIC_quad = 5.75
+  > delta_BIC = BIC_prec-BIC_quad = 10.03
+  > approx Bayes Factor = exp(deltaBIC/2) = 1.14e+04
+  > 
+  > 
+  > NUMBERS WHEN OMITTING PARTIAL TRANSITS AND OBVIOUS SPOT CROSSINGS:
+  > ##########################################
+  > 
+  > quadratic model:
+  > dP/dt = -4.098e-10 +(3.909e-11) -(3.862e-11)
+  > dP/dt = -12.93 +(1.23) -(1.22) millisec/yr
+  > 
+  > -----chi2 values for best fit parameters-----
+  > 
+  > linear model:	     chi2: 163.285	dof: 55	AIC: 1.67e+02	BIC: 1.71e+02
+  > quadratic model:	 chi2: 53.5164	dof: 54	AIC: 5.95e+01	BIC: 6.56e+01
+  > precession model:  chi2: 56.4648	dof: 52	AIC: 6.65e+01	BIC: 7.67e+01
+  > 
+  > -----quad vs precession-----
+  > delta_AIC = AIC_prec-AIC_quad = 6.95
+  > delta_BIC = BIC_prec-BIC_quad = 11.03
+  > approx Bayes Factor = exp(deltaBIC/2) = 3.10e+04
 

### 2019-02-27 src/check_spot_crossing_time_variation_amplitude.py
add script to look at spot deviation distribution

@@ -1,0 +1,147 @@
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from glob import glob
+import os, pickle
+from numpy import array as nparr
+from parse import search
+
+def _make_transit_time_csv(fnames, ticid, outdir='../data/', hasspots=False):
+
+    t0_list, t0_merrs, t0_perrs, t0_bigerrs, picklepaths = (
+        [],[],[],[],[]
+    )
+    seed_list = []
+
+    transit_ix = 0
+    for fname in fnames:
+        transit_ix += 1
+
+        try:
+            d = pickle.load(open(fname, 'rb'))
+
+            if hasspots:
+                seed_list.append(int(search('{}_seed{}/{}',fname)[1]))
+
+            fitparams = d['fitinfo']['finalparams']
+            fiterrs = d['fitinfo']['finalparamerrs']
+
+            t0_list.append(fitparams['t0'])
+            t0_merrs.append(fiterrs['std_merrs']['t0'])
+            t0_perrs.append(fiterrs['std_perrs']['t0'])
+            t0_bigerrs.append(max(
+                fiterrs['std_merrs']['t0'],fiterrs['std_perrs']['t0']))
+            picklepaths.append(fname)
+
+        except Exception as e:
+            print(e)
+            print('transit {:d} failed, continue'.format(transit_ix))
+            continue
+
+    t0, t0_merr, t0_perr, t0_bigerr = (
+        nparr(t0_list),nparr(t0_merrs),nparr(t0_perrs),nparr(t0_bigerrs)
+    )
+    seedarr = nparr(seed_list)
+
+    df = pd.DataFrame({
+        't0_BTJD':t0, 't0_merr':t0_merr, 't0_perr':t0_perr,
+        't0_bigerr':t0_bigerr, 'BJD_TDB':t0+2457000,
+        'picklepath':picklepaths
+    })
+    if hasspots:
+        df['seed'] = seedarr
+        outname = (
+            str(ticid)+'_has_spots_{}_transits.csv'.format(transit_ix)
+        )
+    else:
+        outname = (
+            str(ticid)+'_true_times_{:d}_transits.csv'.format(transit_ix)
+        )
+
+    outpath = os.path.join(outdir,outname)
+    df.to_csv(outpath, index=False)
+    print('saved to {:s}'.format(outpath))
+
+    return outpath
+
+
+def main(ticid=402026209,
[…]

### 2019-02-27 review/referee_response_0.txt
referee response: done with major point #2

@@
 
-We thank you for organizing the review of our paper, and the referee
-for constructive criticism.  We have considered the referee's remarks
-carefully and revised our manuscript accordingly.
+We thank you for organizing the review of our paper, and thank the
+referee for constructive criticism.  We have considered the referee's
+remarks carefully and revised our manuscript accordingly.
 
@@
 
+COMMENT:
+
 -- Major points:
@@
 
-  > Response: We thank the referee for this suggestion.  We have removed
+  > RESPONSE: We thank the referee for this suggestion.  We have removed
   > the word "complete" when describing the transit timing dataset.  We
@@
 
-TODO TODO TODO
-* ONE IDEA: consider ADDING the spots (e.g., from the hubble data, at the
-  observed amplitude). Run it through the lightcurves, over a full transit
-  phase. QUANITFY: How much does the transit mid-time change?
-* ALSO QUANTIFY: what is the peak-to-peak stellar variability? (Can we confirm
-  the previously measured rotation periods?)
-* The period is the same order of time as the observations...
+%FIXME: put in however long the time is!!
+
+  > RESPONSE: We have considered this possibility, and have now added
+  > the following text in Section 3.2:
+
+  > To explore the effect of possible spot-crossing events on the TESS
+  > transit time measurements, we performed a separate test.  We
+  > injected triangular spot-anomalies with amplitude $0.03\%$ and
+  > duration 30 minutes at random phases into each transit.  The
+  > amplitudes were larger than the spot-crossing anomalies observed
+  > by \citet{southworth_high-precision_2009} and
+  > \citet{sanchis-ojeda_starspots_2011}, and the durations were
+  > comparable.  Spots of these amplitudes would be visible
+  > in the lightcurve residuals, and resemble the anomalies present in
+  > transits ``1360.54'' and ``1372.58'' (Figure
+  > \ref{fig:lightcurves}).
+  > 
+  > With spots injected, we repeated our measurement of the transit
+  > times.  On average, the meaured transit times did not change with
+  > spots injected, because the flux deviations are symmetric about
+  > the transit minimum.  For individual transits, there were no cases
+  > for which the timing deviation was larger than one minute, but
+  > about one-eighth of the measured transit times were shifted by
+  > more than 30 seconds early.  For the TESS observations to all be
+  > skewed early, it would be necessary for spot-crossing events to
+  > occur in the latter half of every transit \citep[{\it
+  > e.g.},][]{ioannidis_how_2016}.  The lightcurve residuals are on
+  > average symmetric, and do not support this possibility
+  > (Figure~\ref{fig:lightcurves}).  Further, since the stellar
+  > rotation period is between 20 and 40 days
+  > \citep{sanchis-ojeda_starspots_2011,hoyer_tramos_2013}, and the
+  > TESS data span one lunar month, the spot distribution required by
+  > this hypothesis is also rather convoluted.  We conclude that
+  > neither spot-crossing events nor incomplete phase coverage seem to
+  > explain the observed timing variations, and proceed to possible
+  > astrophysical implications.
 
@@
 
-  > Response: Done.
+  > RESPONSE: Done.
 
@@
 
-  > Response: Barring any specific suggestion by the referee, we will leave
+  > RESPONSE: Barring any specific suggestion by the referee, we will leave
   > this one for the reader to pursue at their leisure.
@@
 
-  > Response: We have clarified that a "Presearch Data
+  > RESPONSE: We have clarified that a "Presearch Data
   > Conditioning" lightcurve is at a stage of process in which
@@
 
-  > Response: We have changed the wording from "ramp-like
+  > RESPONSE: We have changed the wording from "ramp-like
   > systematic effects" to "correlated red noise". No reference is
@@
 
-  > Response: Section 2.2, paragraph 3, final sentence explains
+  > RESPONSE: Section 2.2, paragraph 3, final sentence explains
   > this: "The uncertainty in each data point was set equal to the
@@
 
-  > Response: We have clarified that the phase-folded lightcurve
+  > RESPONSE: We have clarified that the phase-folded lightcurve
   > comes from 18 transits.  We have left the figure ordering
@@
 
-  > Response: We now mention the name "BATMAN". We do the same with
+  > RESPONSE: We now mention the name "BATMAN". We do the same with
   > "emcee".
@@
 
-  > Response: We have added a representative error bar to Figure 2
+  > RESPONSE: We have added a representative error bar to Figure 2
   > which shows the photometric error of the unbinned data points.
@@
 
-  > Response: This is discussed in paragraphs 3 and 4 of Section 2.2:
+  > RESPONSE: This is discussed in paragraphs 3 and 4 of Section 2.2:
 
@@
 
-  > Response: $\chi^2 = 9.2$ is the $\chi^2$ value. We have
+  > RESPONSE: $\chi^2 = 9.2$ is the $\chi^2$ value. We have
   > corrected the typo in the text, in which we wrote "reduced
@@
 
-  > Response: For 18 transits, and 2 degrees of freedom, we
+  > RESPONSE: For 18 transits, and 2 degrees of freedom, we
   > would expect $\chi^2 = 16 \pm 5.7$.
@@
 
-  > Response: This was a typo -- $\sigma_{t_0}$ should have read
+  > RESPONSE: This was a typo -- $\sigma_{t_0}$ should have read
   > $\sigma_{t_{\rm tra}}$. We have added a sentence to summarize
@@
 
-  > Response: Done, thank you.
+  > RESPONSE: Done, thank you.
 
@@
 
-  > Response: We have added the following sentence in the last
+  > RESPONSE: We have added the following sentence in the last
   > paragraph of Section 2.3: "The uncertainties in our derived
@@
 
-  > Response: we have modified the wording slightly to clarify. The point is
+  > RESPONSE: we have modified the wording slightly to clarify. The point is
   > that for these Gemini South observations, we confirmed with the authors
@@
 
-  > Response: Yes, this was through weighted least-squares. We have changed the
+  > RESPONSE: Yes, this was through weighted least-squares. We have changed the
   > wording accordingly.
@@
 
-  > Response: Though we appreciate the suggestion, we have opted to keep the
+  > RESPONSE: Though we appreciate the suggestion, we have opted to keep the
   > "title" as is. The nuance of what "predicted" means for this plot (a linear
@@
 
-  > Response: We have added the following sentence in the caption: "The binned
+  > RESPONSE: We have added the following sentence in the caption: "The binned
   > TESS point is the weighted average of 18 TESS transits".  The caption of
@@
 
-  > Response: Equations (4) and (5) are generic equations for the Taylor series
+  > RESPONSE: Equations (4) and (5) are generic equations for the Taylor series
   > expansion of the transit and occultation times as a function of epoch, and
@@
 
-  > Response: We have added the following sentences in Section 3.2:
+  > RESPONSE: We have added the following sentences in Section 3.2:
   > "We fit each model by assuming a Gaussian likelihood and sampling over the
@@
 
-  > Response: The caption now reads "Best-fit transit timing model
+  > RESPONSE: The caption now reads "Best-fit transit timing model
   > parameters".
@@
 
-  > Response: Corrected.
+  > RESPONSE: Corrected.
 
@@
 
-  > Response: Changed wording to remove any ambiguity.
+  > RESPONSE: Changed wording to remove any ambiguity.
 
@@
 
-  > Response: As noted in the caption of Figure 5, the decay timescale values
+  > RESPONSE: As noted in the caption of Figure 5, the decay timescale values
   > are obtained from Equation 15, assuming Qstar'=10^7.  We have slightly
@@
 
-  > Response: To make Figure 4 and 6, we start with the observed transit and
+  > RESPONSE: To make Figure 4 and 6, we start with the observed transit and
   > occultation mid-times, as a function of epoch number.  We then fit a line
@@
 
-  > Response: we have opted against this suggestion, chiefly to encourage the
+  > RESPONSE: we have opted against this suggestion, chiefly to encourage the
   > interested reader to refer to the original study!
@@
 
-  > Response: As in Figure 3, we appreciate the suggestion, but have opted to
+  > RESPONSE: As in Figure 3, we appreciate the suggestion, but have opted to
   > keep the "title" as is. We have appended the sentence "The two models may
@@
 
-  > Response: We have correcly the wording accordingly.
+  > RESPONSE: We have correcly the wording accordingly.
 
@@
 
-  > Response: We have added the phrase "similar to Jupiter". Section 4.2
+  > RESPONSE: We have added the phrase "similar to Jupiter". Section 4.2
   > already notes that our measured value for k_2p under the precession
@@
 
-  > Response: We have clarified the origin of the values used when evaluating
+  > RESPONSE: We have clarified the origin of the values used when evaluating
   > Equation 21.  The tricky ones -- the mass and radius of the convective

### 2019-02-27 paper/ms.tex
referee response: done with major point #2

@@
 gaps.  A separate issue is the effect of spot-crossing anomalies on
-mid-transit time measurements.  Both effects lead to lightcurve
-asymmetries which skew transit-time measurements.
-
-We repeated the analysis desribed above, but omitted epochs -827,
+mid-transit time measurements.  Both effects can lead to lightcurve
+asymmetries, and skewed transit-time measurements.
+
+We repeated the model-fitting desribed above, but omitted epochs -827,
 -804, -537, and -208 because of gaps in their coverage.  We also
@@
 best-fit transit timing model parameters were all within $1\sigma$ of
-the values quoted in Table~4.  The confidence intervals,
-goodness-of-fit statistics $\chi^2$, and model comparison statistics
-$(AIC, BIC)$ did not appreciably change.
+the values quoted in Table~4.  The uncertainties, goodness-of-fit
+statistics $\chi^2$, and model comparison statistics $(AIC, BIC)$ did
+not appreciably change.
+
+To explore the effect of possible spot-crossing events on the TESS
+transit time measurements, we performed a separate test.  We injected
+triangular spot-anomalies with amplitude $0.03\%$ and duration 30
+minutes at random phases into each transit.  The amplitudes were
+larger than the spot-crossing anomalies observed by
+\citet{southworth_high-precision_2009} and
+\citet{sanchis-ojeda_starspots_2011}, and the durations were
+comparable.  Spots of these amplitudes would be visible in
+the lightcurve residuals, and resemble the anomalies present in
+transits ``1360.54'' and ``1372.58'' (Figure \ref{fig:lightcurves}).
+
+With spots injected, we repeated our measurement of the transit times.
+On average, the meaured transit times did not change with spots
+injected, because the flux deviations are symmetric about the transit
+minimum.  For individual transits, there were no cases for which the
+timing deviation was larger than one minute, but about one-eighth of
+the measured transit times were shifted by more than 30 seconds early.
+For the TESS observations to all be skewed early, it would be
+necessary for spot-crossing events to occur in the latter half of
+every transit \citep[{\it e.g.},][]{ioannidis_how_2016}.  The
+lightcurve residuals are on average symmetric, and do not support this
+possibility (Figure~\ref{fig:lightcurves}).  Further, since the
+stellar rotation period is between 20 and 40 days
+\citep{sanchis-ojeda_starspots_2011,hoyer_tramos_2013}, and the TESS
+data span one lunar month, the spot distribution required by this
+hypothesis is also rather convoluted.  We conclude that neither
+spot-crossing events nor incomplete phase coverage seem to explain the
+observed timing variations, and proceed by considering the
+astrophysical implications.
 }

### 2019-03-02 src/plot_O_minus_C.py
minor plot changes

@@ -14,6 +14,9 @@ import os, pickle
 
 from numpy import array as nparr
 
+from astropy.time import Time
+from astropy import units as units, constants as const
+
 def get_data(
     datacsv='../data/WASP-18b_transits_and_TESS_times_O-C_vs_epoch_selected.csv',
     is_occultation=False
@@ -252,6 +255,16 @@ def plot_O_minus_C(
         a1.text(0.98,0.05, 'Occultations', transform=a1.transAxes, color='k',
                 fontsize='medium', va='bottom', ha='right')
 
+        # add "time" axis on top
+        # make twin axis to show year on top
+        period = 1.338231466*units.day
+        t0 = 2455804.515752*units.day
+        transittimes = x*period + t0
+        times = Time(transittimes, format='jd', scale='tdb')
+        a_top = a0.twiny()
+        a_top.scatter(times.decimalyear, np.zeros_like(times), s=0)
+        a_top.set_xlabel('Year', fontsize='large')
+
         # hidden point for a1 legend
         #a1.plot(1500, 3, alpha=1, mew=0.5,
         #        zorder=-3, label='binned TESS time', markerfacecolor='yellow',
@@ -266,11 +279,13 @@ def plot_O_minus_C(
             ax.get_yaxis().set_tick_params(which='both', direction='in')
             ax.get_xaxis().set_tick_params(which='both', direction='in')
             ax.set_xlim(xlim)
+        a_top.get_yaxis().set_tick_params(which='both', direction='in')
+        a_top.get_xaxis().set_tick_params(which='both', direction='in')
         a0.set_ylim(ylim)
         a1.set_ylim(ylim1)
 
-        fig.text(0.5,0, xlabel, ha='center')
-        fig.text(-0.025,0.5, ylabel, va='center', rotation=90)
+        fig.text(0.5,0, xlabel, ha='center', fontsize='large')
+        fig.text(-0.02,0.5, ylabel, va='center', rotation=90, fontsize='large')
 
         fig.tight_layout(h_pad=0, w_pad=0)
         fig.savefig(savpath, bbox_inches='tight', dpi=400)
@@ -487,10 +502,10 @@ def main(plname, xlim=None, ylim=None, include_all_points=False, ylim1=None):
         copyfile(savpath.replace('.png','.pdf'), '../paper/f4.pdf')
         print('saved ../paper/f4.pdf')
     else:
[…]

### 2019-03-02 src/plot_future.py

@@ -267,15 +267,15 @@ def main(plname, xlim=None, ylim=None, savname=None, ylim1=None):
         refs,
         tcol, tcol_occ,
         savpath=savpath,
-        xlabel='Gregorian year',
+        xlabel='Year',
         ylabel='Deviation from constant period [minutes]',
         xlim=xlim, ylim=ylim, ylim1=ylim1)
 
     if savpath == os.path.join(pkldir, 'future.png'):
-        copyfile(savpath, '../paper/f6.png')
-        print('saved ../paper/f6.png')
-        copyfile(savpath.replace('.png','.pdf'), '../paper/f6.pdf')
-        print('saved ../paper/f6.pdf')
+        copyfile(savpath, '../paper/f7.png')
+        print('saved ../paper/f7.png')
+        copyfile(savpath.replace('.png','.pdf'), '../paper/f7.pdf')
+        print('saved ../paper/f7.pdf')
 
 
 if __name__=="__main__":

### 2019-03-02 src/plot_hjs.py

@@ -224,8 +224,8 @@ def plot_hjs():
 
     fig.tight_layout(h_pad=0, w_pad=0)
 
-    for savname in ['../results/hjs.png', '../paper/f7.png',
-                    '../results/hjs.pdf', '../paper/f7.pdf']:
+    for savname in ['../results/hjs.png', '../paper/f8.png',
+                    '../results/hjs.pdf', '../paper/f8.pdf']:
        fig.savefig(savname, dpi=500, bbox_inches='tight')
        print('saved {:s}'.format(savname))
 

### 2019-03-02 paper/ms.tex
all referee changes added, ready for JNW review

--- 2019-02-24 paper/ms.tex
@@
-\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
+\documentclass[12pt,twocolumn,tighten]{aastex62}
+%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 %\pdfoutput=1 %for arXiv submission
@@
 \usepackage{hyperref}
-\usepackage{comment}
 
@@
 %% Add "Submitted to " argument.
-\received{\today}
-\revised{---}
+\received{December 22, 2018}
+\revised{\today}
 \accepted{---}
@@
 
-%\NewPageAfterKeywords
+\NewPageAfterKeywords
 \listofchanges
@@
 %
-\added{
 \author[0000-0002-8958-0683]{F. Dai}
@@
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
-}
 %
 \author[0000-0002-6939-9211]{T. Daylan}
+\altaffiliation{ Kavli Fellow }
 \affiliation{ Department of Physics and Kavli Institute for Astrophysics
@@
 %
-\added{\author{M. Fausnaugh} % faus@mit.edu CONFIRMED
+\author{M. Fausnaugh} % faus@mit.edu CONFIRMED
 \affiliation{ Department of Physics and Kavli Institute for Astrophysics
 and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }}
+02139, USA }
 %
@@
 %
-\added{
 \author{N. Guerrero} % nmg@mit.edu CONFIRMED
@@
 02139, USA }
-}
 %
@@
 %
-\added{
 \author{B. Wohler} % bill.wohler@nasa.gov YES
@@
 \affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-}
 
@@
     shortcomings.  The gravitational influence of a third body is
-    another possibility, although at present there is \replaced{no}{minimal} 
-    evidence for
-    such a body. Further observations are needed to confirm and
-    understand the timing variation.
+    another possibility, though at present there is
+    \replaced{no}{minimal} evidence for such a body. Further
+    observations are needed to confirm and understand the timing
+    variation.
 \end{abstract}
@@
 
-Although the Transiting Exoplanet Survey Satellite (TESS, \added{
-\citealt{ricker_transiting_2015}}) is designed
-to detect new planets, it is also performing precise photometric
-monitoring of most of the planets that have been discovered by
-ground-based transit surveys over the last two decades.  One of the
-applications of the new TESS data is to search for timing anomalies in
-hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
-occultation times should eventually reveal two distinct processes.
-
-First, whenever the orbital angular momentum is less than 3 times the
-rotational angular momentum, the system is vulnerable to tidal orbital
-decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
-Jupiters satisfy this condition and their orbits should be shrinking
-\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay,
-which depends on how friction dissipates the energy of tidal
-disturbances, is difficult to observe directly or calculate from first
-principles. This has been a stubborn source of uncertainty in the
-interpretation of many phenomena in stellar and planetary astrophysics
-(as reviewed by \citealt{Mazeh2008} and \citealt{ogilvie_tidal_2014}).
-
-Second, long-term timing studies may reveal apsidal precession, {\it
-i.e.}, the rotation of the orbital ellipse within the orbital plane.
-Apsidal precession has long been observed in eclipsing binaries
-\citep[{\it e.g.},][]{
-  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
-For exoplanets, precessing orbits have been detected in a handful of
-cases, including circumbinary systems (Kepler-413
-\citealt{kostov_kepler-413b_2014}; Kepler-453
-\citealt{welsh_kepler_2015}), planets orbiting rapidly rotating stars
-(Kepler-13
-\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015});
-and systems with multiple interacting giant planets (Kepler-108
-\citealt{mills_kepler-108_2017}; Kepler-693
-\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
-observed in hot Jupiter systems because their orbits are nearly
-circular, which may itself be a consequence of star-planet tides.  If
-it could be observed, then as \citet{ragozzine_probing_2009} showed it
-would be possible to measure the planet's Love number, which would be
-a valuable constraint on the planet's interior.
-
-So far, the most convincing direct evidence for either orbital decay
-or apsidal precession has come from WASP-12b.
-\citet{maciejewski_departure_2016} and \citet{patra_2017} showed that
-the transit period of WASP-12b has been steadily decreasing by about
-30~milliseconds per year over the last decade.  Here, we present
-evidence for a timing anomaly in the WASP-4 system.  The hot Jupiter
-WASP-4b orbits a G7V star every 1.34 days, corresponding to an orbital
-distance of 5.5 stellar radii
+Although the Transiting Exoplanet Survey Satellite (TESS\added{,
+\citealt{ricker_transiting_2015}}) is designed to detect new planets,
+it is also \replaced{performing precise photometric monitoring
+of}{precisely monitoring} most of the planets that have been
+discovered by ground-based transit surveys over the last two decades.
+One \replaced{of the applications}{application} of the new TESS data
+is to search for timing anomalies in \added{previously known }hot
+Jupiter systems.  Long-term monitoring of hot Jupiter transit and
+occultation times should eventually reveal \replaced{two
+distinct}{three select} processes.
+
+\explain{The following four paragraphs have been modified to add
+emphasis to the possibility of an outer planet explaining the timing
+deviations. Some background information given in the original
+submission has been removed.}
+
+First, the orbits of most hot Jupiters should shrink because of tidal
+orbital decay
+\citep{counselman_outcomes_1973,hut_stability_1980,levrard_falling_2009,matsumura_tidal_2010}.
+Directly measuring the rate of decay might lead to an improved
+understanding of how friction dissipates the energy of tidal
+disturbances (a problem reviewed by \citealt{Mazeh2008} and
+\citealt{ogilvie_tidal_2014}).
+
+Second, if hot Jupiter orbits are at all eccentric, then long-term
+timing studies should reveal rotation of the orbital ellipse within
+the orbital plane (``apsidal precession'').  If this effect were
+observed, it could yield a measure of the planet's Love number, which
+would constrain the planet's interior structure
+\citep{ragozzine_probing_2009}.
+
+The most convincing direct evidence yet found for either orbital decay
+or apsidal precession of a hot Jupiter is the case of WASP-12b, which
+has a transit period that has decreased by about 30~milliseconds per
+year over the last decade
+\citep{maciejewski_departure_2016,patra_2017}.
+
+The final effect of interest that can produce period changes in hot
+Jupiter systems is gravitational acceleration caused by massive outer
+companions \citep[{\it e.g.},][Section~4]{agol_detecting_2005}.
+Prototypes include WASP-53 and WASP-81, systems in which the inner hot
+Jupiters are periodically perturbed by eccentric brown dwarf
+companions with semimajor axes of a few astronomical units
+\citep{triaud_peculiar_2017}.
+
+% DELETED THE FOLLOWING TEXT FROM SUBMITTED VERSION:
+% First, whenever the orbital angular momentum is less than 3 times
+% the rotational angular momentum, the system is vulnerable to tidal
+% orbital decay \citep{counselman_outcomes_1973,hut_stability_1980}.
+% Most hot Jupiters satisfy this condition and their orbits should be
+% shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The
+% rate of decay, which depends on how friction dissipates the energy
+% of tidal disturbances, is difficult to observe directly or calculate
+% from first principles. This has been a stubborn source of
+% uncertainty in the interpretation of many phenomena in stellar and
+% planetary astrophysics (as reviewed by \citealt{Mazeh2008} and
+% \citealt{ogilvie_tidal_2014}).
+%
+% Second, long-term timing studies may reveal apsidal precession, {\it
+% i.e.}, the rotation of the orbital ellipse within the orbital plane.
+% Apsidal precession has long been observed in eclipsing binaries
+% \citep[{\it
+% e.g.},][]{schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
+% For exoplanets, precessing orbits have been detected in a handful of
+% cases, including circumbinary systems (Kepler-413
+% \citealt{kostov_kepler-413b_2014}; Kepler-453
+% \citealt{welsh_kepler_2015}), planets orbiting rapidly rotating
+% stars (Kepler-13
+% \citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015});
+% and systems with multiple interacting giant planets (Kepler-108
+% \citealt{mills_kepler-108_2017}; Kepler-693
+% \citealt{masuda_eccentric_2017}).  Apsidal precession is not
+% regularly observed in hot Jupiter systems because their orbits are
+% nearly circular, which may itself be a consequence of star-planet
+% tides.  If it could be observed, then as
+% \citet{ragozzine_probing_2009} showed it would be possible to
+% measure the planet's Love number, which would be a valuable
+% constraint on the planet's interior.
+%
+% So far, the most convincing direct evidence for either orbital decay
+% or apsidal precession has come from WASP-12b.
+% \citet{maciejewski_departure_2016} and \citet{patra_2017} showed
+% that the transit period of WASP-12b has been steadily decreasing by
+% about 30~milliseconds per year over the last decade.
+
+Here, we present evidence for a timing anomaly in the WASP-4 system.
+The hot Jupiter WASP-4b orbits a G7V star every 1.34 days,
+corresponding to an orbital distance of 5.5 stellar radii
 \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
@@
 who combined the available transit times, occultation times, and
-Doppler data.  The sky projection of the stellar obliquity
-is also compatible with zero, within about 10 degrees
+Doppler data.  The sky projection of the stellar obliquity is also
+compatible with zero, within about 10 degrees
 \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011}.
@@
 decaying period, and a slightly eccentric precessing orbit.  A
-constant period can be ruled out, but we cannot distinguish between
-the latter two possibilities.  Either possibility would have
-interesting implications, described in \S~\ref{sec:implications}.
+constant period can be ruled out\replaced{, but we}{. We} cannot
+distinguish between the \replaced{latter two possibilities}{
+  possibilities of a decaying orbit, a precessing orbit, and the
+  unmodeled possibility of an orbit being gravitationally perturbed by
+  an outer companion}.  \replaced{Either possibility}{Any of the three
+scenarios} would have interesting implications\replaced{, described in
+\S~\ref{sec:implications}}{ (\S~\ref{sec:implications}), and more data
+are required for a definitive ruling (\S~\ref{sec:future}).}
 Appendix~\ref{sec:verify_tess} considers the possibility that the
@@
         TESS \replaced{data}{flux measurements, with median 
-        $1\sigma$ error shown in the lower 
+        $1\sigma$ uncertainty shown in the lower 
         right.} 
@@
 as Barycentric TESS Julian Dates (BTJD), which is simply the
-\replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus 2{,}457{,}000.  
-Lightcurves that 
-were
-flagged \deleted{as Threshold Crossing Events} by the SPOC pipeline \added{as 
-crossing a transit detection threshold} were then
-vetted and released by the MIT TESS Science Office to the Mikulski
-Archive for Space Telescopes on November 29,
-2018~\citep{ricker_tess_alerts_2018}.
+\replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus
+2{,}457{,}000.  Lightcurves that were flagged \deleted{as Threshold
+Crossing Events} by the SPOC pipeline \added{as crossing a transit
+detection threshold} were then vetted and released by the MIT TESS
+Science Office to the Mikulski Archive for Space Telescopes on
+November 29, 2018~\citep{ricker_tess_alerts_2018}.
 
@@
 lightcurve\replaced{
-\citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}{, which has had 
-systematic trends removed 
-through the decorrelation process described by 
-\citet{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}.
-We \added{then} processed the lightcurve as follows.  First, we removed all 
-points
-with non-zero quality flags.  This removed data that might have been
-adversely affected by ``momentum dumps,'' the firing of thrusters and
-resetting of reaction
-wheels\footnote{\added{Details concerning the spacecraft pointing and momentum 
-dumps are described in the data release 
-notes:}\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
-that took place every 2.5 days during sector 2
-\deleted{\citep{tess_data_product_description_2018}}.  The data during these
+  \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}{, which
+has had non-astrophysical variability removed through the methods
+discussed by \citet{smith_kepler_apertures_2017} and
+\citet{smith_kepler_PDC_2017}}.  We \added{then} processed the
+lightcurve as follows.  First, we removed all points with non-zero
+quality flags.  This removed data that might have been adversely
+affected by ``momentum dumps,'' the firing of thrusters and resetting
+of reaction wheels\footnote{\added{The spacecraft pointing and
+momentum dumps are described in the data release notes:
+}\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
+that took place every 2.5 days during sector 2\deleted{
+  \citep{tess_data_product_description_2018}}.  The data during these
 events were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude'' 
-\added{\citep[][Table 28]{tess_data_product_description_2018}}.  For WASP-4, 
-these flags
+Desaturation Event'' and ``Manual Exclude'' \added{\citep[][Table
+28]{tess_data_product_description_2018}}.  For WASP-4, these flags
 were simultaneously set for 54 distinct cadences, and there were 10
@@
 after every momentum dump.  We also removed the data within the first
-and last hours of both orbits, because of \replaced{ramp-like systematic 
-effects that appear}{correlated red noise that appears} during those time 
-ranges.
+and last hours of both orbits, because of \replaced{ramp-like
+systematic effects that appear}{correlated red noise that appears}
+during those time ranges.
 
@@
 then created a phase-folded lightcurve \added{from all 18 transits}
-and fitted a standard transit
-model using the analytic formulae given by
-\citet{mandel_analytic_2002} and implemented by
+and fitted a standard transit model using the analytic formulae given
+by \citet{mandel_analytic_2002} and implemented by
 \replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
-We assumed 
-the orbit to be circular,
-consistent with the limits from radial velocities and occultation
-timing
+We assumed the orbit to be circular, consistent with the limits from
+radial velocities and occultation timing
 \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
@@
 \replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
-  Table~1 gives the 
-results, which
-are in reasonable agreement with the parameters reported by {\it
-e.g.}, \citet{southworth_high-precision_2009} and
+Table~1 gives the results, which are in reasonable agreement with the
+parameters reported by {\it e.g.},
+\citet{southworth_high-precision_2009} and
 \citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
@@
 parameters at the values that had been determined from the
-phase-folded TESS lightcurve.  The uncertainty in each \added{photometric} 
-data point was set equal to the root-mean-square (rms) level of the 
-out-of-transit data.
+phase-folded TESS lightcurve.  The uncertainty in each
+\added{photometric} data point was set equal to the root-mean-square
+(rms) level of the out-of-transit data.
 
 To verify that the measured uncertainties are estimated accurately, we
-computed the \deleted{reduced} $\chi^2$ \added{value} for a linear ephemeris 
-fit to the
-measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
-$n=16$ degrees of freedom.  The variance of the $\chi^2$ distribution
-is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.  Visually
-inspecting the residuals showed that the error variance had been
-overestimated, so we multiplied the measured TESS errors by a factor
-$f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
-uncertainty of the transit midtimes from $29.8$ to $22.6$ seconds.  We
-verified that omitting this step did not appreciably alter any of our
-conclusions.
+computed the \deleted{reduced }$\chi^2$\added{ value} for a linear
+ephemeris fit to the measured TESS mid-transit times.  We found that
+$\chi^2 = 9.2$, with $n=16$ degrees of freedom.  The variance of the
+$\chi^2$ distribution is $2n$, so we would expect $\chi^2 = 16 \pm
+5.7$.  Visually inspecting the residuals showed that the error
+variance had been overestimated, so we multiplied the measured TESS
+errors by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.
+This lowered the mean uncertainty of the transit midtimes from $29.8$
+to $22.6$ seconds.  We verified that omitting this step did not
+appreciably alter any of our conclusions.
 
@@
   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
-  %Rotation period~[days]                     & $\approx$34            & ---                         & H \\
 {\it Planetary parameters:} & & & \\
@@
   (G) \citet{stassun_TIC_2018}.
-  %(H) \citet{hoyer_tramos_2013}, from star-spot occultation events.
 }
@@
 by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
-the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$
-and $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from
-{\it 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
+the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$ and
+$V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
+2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
 wavelength range 0.4--22~$\mu$m.  We adopted the effective temperature
@@
 Table~1 gives the resulting parameters, which we adopted for the
-remaining analysis.
-\added{The uncertainties in our derived stellar and planetary parameters
-    are propagated according to standard analytic formulae, under the 
-    assumption of uncorrelated and normally distributed initial measurements.}
-Our \added{system} parameters are in agreement with those of
-previous investigators, but have the benefit of incorporating the {\it
-Gaia} parallax
+remaining analysis.  \added{The uncertainties in our derived stellar
+and planetary parameters are propagated according to standard analytic
+formulae, under the assumption that the variables are uncorrelated and
+normally distributed.} Our \added{system} parameters are in agreement
+with those of previous investigators, but have the benefit of
+incorporating the {\it Gaia} parallax
 \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
@@
     \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
-      Both plots show the deviations between the observed and
-      calculated transit times, where the calculation is based only on
-      \deleted{the} pre-TESS data and assumes a constant period.  The blue 
-      bands
-      depict the $\pm$$1\sigma$ credible interval of the predicted
-      times.  {\it Top:} The full timing dataset spans 11 years. The
-      darkest points correspond to the most precise data. \added{The binned 
-      TESS point is the weighted average of 18 TESS transits.}  {\it
-        Bottom:} Close-up of the TESS observations. The red band shows
-      the average deviation of the TESS transits \added{($\pm 1\sigma$)}, 
-      which arrived $81.6 \pm 11.7\ {\rm seconds}$ earlier than predicted.
+      Both plots show the deviations between observed and calculated
+      transit times, where the calculation is based only on
+      \deleted{the} pre-TESS data and assumes a constant period.  The
+      blue bands depict the $\pm$$1\sigma$ credible interval of the
+      predicted times.  {\it Top:} The full timing dataset spans 11
+      years. The darkest points correspond to the most precise data.
+      \added{The binned TESS point is the weighted average of 18 TESS
+      transits.}  {\it Bottom:} Close-up of the TESS observations. The
+      red band shows the average deviation of the TESS transits
+      \added{($\pm 1\sigma$)}, which arrived $81.6 \pm 11.7\ {\rm
+      seconds}$ earlier than predicted.
       \label{fig:arrived_early}
@@
   times assuming a constant period. The darkest points correspond to
-        the most precise data.}{The vertical axis shows the 
-		observed times minus the calculated times assuming a constant period 
-		for transits ({\it top}) and occultations ({\it bottom}). Darker 
-		points correspond to more precise data.}
-        The constant-period model (gray line) is
-		a poor description of the data.  Models with a uniformly
-		decreasing period (blue) or apsidal precession (orange) provide
-		better fits\deleted{ to the data}.  The red square represents the
-		combination of all the TESS data and is for display purposes
-		only.  The models were fitted to all of the individual transit
-		times.
+  the most precise data.}{The vertical axis shows the observed times
+  minus the calculated times assuming a constant period for transits
+  ({\it top}) and occultations ({\it bottom}). Darker points
+  correspond to more precise data.} The constant-period model (gray
+  line) is a poor description of the data.  Models with a uniformly
+  decreasing period (blue) or an eccentric, precessing orbit (orange)
+  provide	better fits\deleted{ to the data}.  The red square
+  represents the combination of all the TESS data and is for display
+  purposes only.  The models were fitted to all of the individual
+  transit times.
   \label{fig:times}
@@
 these data points carry significant weight in the analysis, we
-\replaced{checked}{corresponded with the authors to confirm}
-that the timestamps in their data represent mid-exposure times, that
-the barycentric correction was performed correctly, and that the time
+\replaced{checked}{corresponded with the authors to confirm} that the
+timestamps in their data represent mid-exposure times, that the
+barycentric correction was performed correctly, and that the time
 system of the final results was BJD$_{\rm TDB}$.  These same authors
@@
 model.  Finally, these authors also measured two transit midpoints
-using Spitzer (Baxter et al., in prep); these
-times agree with the Gemini South results.
+using Spitzer (Baxter et al., in prep); these times agree with the
+Gemini South results.
 
@@
 \end{equation}
-\added{for $E$ the transit number, $t_0$ the reference epoch, 
-$e$ the eccentricity, and $\omega$ the argument of pericenter.}
-In total, there are four \added{available} occultation times.
+\added{for $E$ the transit number, $t_0$ the reference epoch, $e$ the
+eccentricity, and $\omega$ the argument of pericenter.} In total,
+there are four \added{available} occultation times.
 
@@
 
-First, we \replaced{fitted}{performed a weighted least-squares fit of} a 
-constant-period model (\added{a} ``linear ephemeris'') to the
+First, we \replaced{fitted}{performed a weighted least-squares fit of}
+a constant-period model (\added{a }``linear ephemeris'') to the
 pre-TESS data, and used it to extrapolate to the epochs of the TESS
@@
 $P_{\rm a}$ is the anomalistic period, and $\omega$ is the argument of
-pericenter.
-In this model the angular velocity of the line of apsides
+pericenter.  In this model the angular velocity of the line of apsides
 ${\rm d}\omega/{\rm d}E$ is constant,
@@
 same orientation with respect to the stars; the slightly longer
-    anomalistic period is the duration required to reach a fixed 
-    longitude with respect to the rotating line of apsides.}
-The five free parameters of this model are $(t_0, P_{\rm s}, e,
-\omega_0, {\rm d}\omega/{\rm d}E)$, denoting the reference epoch, the
-sidereal period, the eccentricity, the argument of pericenter at the
-reference epoch, and the angular velocity of the line of apsides.
-
-\added{We fit each model by assuming a Gaussian likelihood and sampling over 
-the posterior probability distributions. The prior for the quadratic
-model allowed the period derivative to have any sign.}
+anomalistic period is the duration required to reach a fixed longitude
+with respect to the rotating line of apsides.} The five free
+parameters of this model are $(t_0, P_{\rm s}, e, \omega_0, {\rm
+d}\omega/{\rm d}E)$, denoting the reference epoch, the sidereal
+period, the eccentricity, the argument of pericenter at the reference
+epoch, and the angular velocity of the line of apsides.
+
+\added{We fitted each model by assuming a Gaussian likelihood and
+sampling over the posterior probability distributions. The prior for
+the quadratic model allowed the period derivative to have any sign.}
 Figure~\ref{fig:times} shows the residuals with respect to the
@@
 $\chi^2 = 174$ and 61 degrees of freedom.  The best-fitting
-\replaced{decreasing-period}{quadratic} model has $\chi^2 = 62.6$ and 60 degrees of freedom.
-The best-fitting precession model has $\chi^2 = 64.3$ and 58 degrees
-of freedom.  Either of the latter two models provides a much better
-fit than the constant-period model.  The difference in $\chi^2$
-between the first and second models corresponds to $p \approx
-10^{-26}$.
-
-The \replaced{decreasing-period}{quadratic} model provides a slightly better fit to the data
-than the precession model.  It is favored by $\Delta \chi^2 = 1.7$,
-and has two fewer free parameters.  A useful heuristic for model
-comparison is the Bayesian Information Criterion (BIC),
+\replaced{decreasing-period}{quadratic} model has $\chi^2 = 62.6$ and
+60 degrees of freedom.  The best-fitting precession model has $\chi^2
+= 64.3$ and 58 degrees of freedom.  Either of the latter two models
+provides a much better fit than the constant-period model.  The
+difference in $\chi^2$ between the first and second models corresponds
+to $p \approx 10^{-26}$.
+
+The \replaced{decreasing-period}{quadratic} model provides a slightly
+better fit to the data than the precession model.  It is favored by
+$\Delta \chi^2 = 1.7$, and has two fewer free parameters.  A useful
+heuristic for model comparison is the Bayesian Information Criterion
+(BIC),
 \begin{equation}
@@
 the precession and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm
-  prec} - {\rm BIC}_{\rm quad} = 10$, corresponding to a Bayes factor
-of $1.1\times 10^{4}$.  Likewise, the Akaike Information Criterion
-favors the constant-period-derivative model by $\Delta {\rm AIC} =
-5.8$.  Differences of this magnitude are traditionally deemed ``strong
+prec} - {\rm BIC}_{\rm quad} = 10$, corresponding to a Bayes factor of
+$1.1\times 10^{4}$.  Likewise, the Akaike Information Criterion favors
+the constant-period-derivative model by $\Delta {\rm AIC} = 5.8$.
+Differences of this magnitude are traditionally deemed ``strong
 evidence'' that one model is a better description of the data than the
@@
 
-In the \replaced{decreasing-period}{quadratic} model, the period derivative is
+In the \replaced{decreasing-period}{quadratic} model, the period
+derivative is
 \begin{equation}
@@
 truly falling onto their stars, then WASP-4b is falling at about half
-the rate of WASP-12b.  In the precession model, the best-fit
-eccentricity is
+the rate of WASP-12b.
+
+In the precession model, the best-fit eccentricity is
 \begin{equation}
@@
 
-\added{
+\explain{The following subsection has been added since the original
+submission, per the referee's suggestions.}
+\subsection{Systematic concerns}
+
+\paragraph{Omitting suspect pre-TESS transits}
 To assess the robustness of these results, we considered a few
 possible systematic effects in the timing dataset.  One concern in the
-pre-TESS timing measurements was incomplete phase coverage: a handful
+pre-TESS measurements was incomplete phase coverage: a handful
 of the transit times in Table~2 are from lightcurves with partial
 gaps.  A separate issue is the effect of spot-crossing anomalies on
-mid-transit time measurements.  Both effects lead to lightcurve
-asymmetries which skew transit-time measurements.
-
-We repeated the analysis desribed above, but omitted epochs -827,
--804, -537, and -208 because of gaps in their coverage.  We also
-omitted epochs -526 and -561 because of visible spot anomalies during
-the transits. (All epoch numbers are as in Table~2.)  The resulting
+mid-transit time measurements.  To address these concerns, we repeated
+the model-fitting described above, but omitted epochs -827, -804,
+-537, and -208 because of gaps in their coverage.  We also omitted
+epochs -526 and -561 because of visible spot anomalies during the
+transits. (All epoch numbers are as in Table~2.)  The resulting
 best-fit transit timing model parameters were all within $1\sigma$ of
-the values quoted in Table~4.  The confidence intervals,
-goodness-of-fit statistics $\chi^2$, and model comparison statistics
-$(AIC, BIC)$ did not appreciably change.
-}
+the values quoted in Table~4.  The uncertainties, goodness-of-fit
+statistics, and model comparison statistics did not appreciably
+change.
+
+\paragraph{Spot-crossing events in TESS data}
+To explore the effect of possible spot-crossing events on the TESS
+transit time measurements, we performed a separate test.  We injected
+triangular spot-anomalies with amplitude $0.03\%$ and duration 30
+minutes at random phases into each transit.  The amplitudes were
+larger than the spot-crossing anomalies observed by
+\citet{southworth_high-precision_2009} and
+\citet{sanchis-ojeda_starspots_2011}, and the durations were
+comparable.  Spots of these amplitudes resemble the anomalies present
+in transits ``1360.54'' and ``1372.58'' of Figure
+\ref{fig:lightcurves}: they are visible, but at low confidence.
+
+With spots injected, we repeated our measurement of the transit times.
+On average, the measured transit times did not change with spots
+injected, because the flux deviations are symmetric about the transit
+minimum.  For individual transits, there were no cases for which the
+timing deviation was larger than one minute.  The most concerning
+shift occurs when the spot anomaly occurs during transit egress, in
+which case the measured mid-time is shifted early by between 30 and 50
+seconds \citep[qualitatively similar to results found
+by][]{ioannidis_how_2016}.
+
+Therefore the TESS observations could hypothetically all be skewed
+early if there were spot-crossing events during every egress.  Two
+arguments rule out this possibility.  (1) The lightcurve residuals do
+not show evidence for these events.  (2) The stellar rotation period
+is between 20 and 40 days, and the sky-projected stellar obliquity is
+less than 10 degrees
+\citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
+Since the planet orbits every 1.3 days, requiring that spot anomalies
+always occur during egress would be equivalent to requiring a stellar
+spot distribution that is exquisitely (and thus implausibly)
+distributed to match the planet egress times.
+
+\paragraph{Detrending choices in pre-TESS data}
+There is a final systematic concern in our global methodology.  We
+have opted to use mid-transit time values derived by different
+authors, who used heterogeneous methods to fit and detrend their
+lightcurves.  This of course assumes that these authors
+have correctly documented their time systems.  Further, though many
+choices in transit-fitting ({\it e.g.}, parametrization of
+limb-darkening and eccentricity) do not affect transit mid-time
+measurements, different detrending approaches can asymmetrically warp
+transits and shift mid-transit times.  The magnitude of this
+systematic effect is hard to quantify, but the situation is fairly
+clear from Figure~\ref{fig:times}.  Many independent authors provided
+transit measurements shortly after WASP-4b's discovery, so the data
+should on average be acceptable.  \citet{huitson_gemini_2017} provided
+the most important data from epochs 0-1000, and so if their data were
+systematically affected by detrending choices at the level of many
+times their reported uncertainties, our timing anomaly would
+be suspect.  For this reason, we paid careful attention to the
+\citet{huitson_gemini_2017} data set, and corresponded with the authors
+to confirm that their results are not affected by systematic 
+effects of the required amplitude.
+
+None of the concerns mentioned in this subsection are likely
+explanations for the observed timing variations. We proceed by
+considering possible astrophysical explanations.
 
@@
     colored symbols highlight WASP-4, WASP-12 (which also shows
-    evidence for a decreasing period), and the \deleted{collection of}
+    evidence for a decreasing period), and the\deleted{ collection of}
     hot Jupiters analyzed in Appendix A.
@@
     \added{Circles represent planets whose orbits are consistent with
-    circular.}
-    Lines of constant damping timescale are drawn based on
+    circular.} Lines of constant damping timescale are drawn based on
     Equation~\ref{eq:de_dt} and assuming $Q_{\rm p}' = 10^5$.  WASP-4b
@@
 
-If the timing variation is \deleted{being caused} caused entirely by
+If the timing variation is \deleted{being caused }caused entirely by
 orbital decay, then \replaced{the}{the best-fit model parameters yield a}
-characteristic \replaced{timescale of the decay is}{decay timescale of}
+characteristic\replaced{ timescale of the decay is}{ decay timescale of}
 \begin{equation}
@@
 theoretical timescales for orbital decay.  \replaced{There
-are}{Figure~\ref{fig:context} shows} about 20 hot
-Jupiters (including WASP-12) for which the theoretical timescale is
-shorter.  In almost all of those cases, though, the 
-planet was
-discovered more recently than WASP-4 and a decade-long baseline of
-observations is not yet available.  A separate consideration not shown
-in Figure~\ref{fig:context} is that the hot Jupiter host stars have a
+are}{Figure~\ref{fig:context} shows} about 20 hot Jupiters (including
+WASP-12) for which the theoretical timescale is shorter.  In almost
+all of those cases, though, the planet was discovered more recently
+than WASP-4 and a decade-long baseline of observations is not yet
+available.  A separate consideration not shown in
+Figure~\ref{fig:context} is that the hot Jupiter host stars have a
 variety of different structures, from being fully convective to nearly
@@
 by a modified\footnote{For stars, $k_\star \sim \mathcal{O}(10^{-2})$,
-  so it is important to explicitly distinguish $Q_\star'$ from
-  $Q_\star$ \citep[{\it e.g.},][]{schwarzschild_structure_1958}.}
-quality factor, $Q_\star' = 3 Q_\star / (2k_\star)$.
-Here, $Q_\star$ is the ratio between the energy stored in the
-equilibrium deformation of the star and the energy lost to heat per
-tidal period \citep[{\it e.g.},][]{goldreich_q_1966}.  A larger
-$Q_\star$ implies less efficient tidal dissipiation. The dimensionless
-number $k_\star$ is the stellar Love number, which is smaller when the
-star's density distribution is more centrally concentrated.  In this
-model, once the planet's spin and orbit are synchronized, then the
-semi-major axis and eccentricity evolve as \citep[Appendix B
-of][]{metzger_optical_2012}
+so it is important to explicitly distinguish $Q_\star'$ from $Q_\star$
+\citep[{\it e.g.},][]{schwarzschild_structure_1958}.} quality factor,
+$Q_\star' = 3 Q_\star / (2k_\star)$.  Here, $Q_\star$ is the ratio
+between the energy stored in the equilibrium deformation of the star
+and the energy lost to heat per tidal period \citep[{\it
+e.g.},][]{goldreich_q_1966}.  A larger $Q_\star$ implies less
+efficient tidal dissipiation. The dimensionless number $k_\star$ is
+the stellar Love number, which is smaller when the star's density
+distribution is more centrally concentrated.  In this model, once the
+planet's spin and orbit are synchronized, then the semi-major axis and
+eccentricity evolve as \citep[Appendix B of][]{metzger_optical_2012}
 \begin{align}
@@
 
-A different way to excite the hot Jupiter's eccentricity is the
+A separate way a neighboring companion could excite the hot Jupiter's
+eccentricity is through the
 Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
@@
   M_{\rm c} > 7.5\,M_\oplus
-  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2}.
+  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
   \label{eq:kozai_bound}
@@
 \replaced{where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major 
-axis of WASP-4b and the hypothetical WASP-4c, and owing}{
-where $a_{\rm b}$ is the semi-major axis of WASP-4b, $M_{\rm c}$
-is the mass of the hypothetical WASP-4c, and $a_{\rm c}$ is WASP-4c's 
-semi-major axis. Owing}
+axis of WASP-4b and the hypothetical WASP-4c, and owing}{where $a_{\rm b}$ is 
+the semi-major axis of WASP-4b, $M_{\rm c}$ is the mass of the hypothetical 
+WASP-4c, and $a_{\rm c}$ is WASP-4c's semi-major axis. Owing}
 to our imprecise
@@
 \paragraph{Fluctuations in the gravitational potential from convection}
-A separate mechanism to pump the eccentricity invokes the
+An independent mechanism to pump the eccentricity invokes the
 gravitational fluctuations from stellar convection \citep[][Section
@@
 mass, $n$ is the orbital frequency, and $a$ is the semi-major axis.
-\added{For the luminosity, reduced mass, and semi-major axis, we used values 
-from Table~1, combined with the Stefan-Boltzmann law and standard 
-definitions. To estimate the 
-width and mass of the convective region we ran the MESA code
-for a star with mass and metallicity matched to WASP-4, and the 
-input physics detailed in the MIST isochrones project
+\added{For the luminosity, reduced mass, and semi-major axis, we used
+values from Table~1, combined with the Stefan-Boltzmann law and
+standard definitions. To estimate the width and mass of the convective
+region we ran the MESA code for a star with mass and metallicity
+matched to WASP-4, and the input physics detailed in the MIST
+isochrones project
 \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
-We identified the tachocline boundary using the mixing 
-types specified in the resulting radial profiles, and found $R_{\rm conv} 
-\approx 0.33 R_\odot$, and $M_{\rm conv} \approx 9\times10^{-4} M_\odot$. }
-For WASP-4, this implies $\langle e^2 \rangle^{1/2} \lesssim
-10^{-5}$.  Hence, this mechanism does not seem capable of producing
-the required eccentricity of $\sim$$10^{-3}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% LB: I calculated the above using MESA grids from Penev+ 2018 to
-% estimate the core/envelope masses, but mentioning this is too
-% detailed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We identified the tachocline boundary using the mixing types specified
+in the resulting radial profiles, and found $R_{\rm conv} \approx 0.33
+R_\odot$, and $M_{\rm conv} \approx 9\times10^{-4} M_\odot$. } For
+WASP-4, this implies $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$.
+Hence, this mechanism does not seem capable of producing the required
+eccentricity of $\sim$$10^{-3}$.
+
+\explain{The subsequent subsection adds a description of our RV
+analysis, which we did not include in the original submission. Some of
+the material was taken from what was previously the "Other possible
+explanations" section.}
+\subsection{Timing variation due to line of sight acceleration}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Radial velocities of WASP-4 (top), and residuals from the
+    best-fit Keplerian model (bottom).} The lower panel shows the
+    best-fit linear trend inferred from the RV data (black line,
+    $1\sigma$ errors in gray), and the trend that would be needed to
+    produce the period decrease seen in transits (purple dotted line).
+    Since both the RV and transit timing datasets are sparse after
+    2013, a distant massive companion on an eccentric orbit might
+    still explain the observations.
+	\label{fig:rvs}
+  \vspace{-0.3cm}
+	}
+\end{figure}
+
+An acceleration of the center of mass of the system towards our line
+of sight could cause a decrease in the apparent orbital period.  The
+period derivative would be
+\begin{equation}
+	\dot{P} = \frac{\dot{v}_{\rm r} P}{c},
+\end{equation}
+where $\dot{v}_{\rm r}$ is the time derivative of the radial velocity.
+
+\citet{knutson_friends_2014}, combining radial-velocity data from
+their own program with those of \citet{wilson_wasp-4b_2008},
+\citet{pont_determining_2011}, and \citet{husnoo_observational_2012},
+found a weak long-term trend in WASP-4:
+\begin{equation}
+  \dot{v}_{\rm r} =
+     -0.0099^{+0.0052}_{-0.0054}
+     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}.
+\end{equation}
+This acceleration translates to an expected $\dot{P} = -(4.4 \pm 2.4)
+\times10^{-11}$. The period decrease from the observed RV trend is an
+order of magnitude smaller than our observed $\dot{P}$ from transit
+times, $(4.0\pm 0.4)\times10^{-10}$ (Table~4).
+
+Nonetheless, it is intriguing that the host star shows a weakly
+significant acceleration, and with the correct sign needed to explain
+the transit timing variations.  Since the radial velocities could
+affect the interpretation of the system, we redetermined the line of
+sight acceleration as follows.
+
+First, we collected the usable RV measurements from CORALIE, HARPS,
+and HIRES.  We included the CORALIE measurements from
+\citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010}, using
+the homogeneous radial velocities calculated by the latter authors. We
+included the HARPS values reported by \citet{pont_determining_2011},
+which are identical to those from \citet{husnoo_observational_2012}.
+We omitted the HARPS data points taken over three nights by
+\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
+because they were calculated using a different pipeline than the
+longer-baseline \citealt{pont_determining_2011} HARPS measurements,
+and necessary inclusion of an extra offset term would nullify their
+statistical value.  Finally, we included the five HIRES measurements
+taken over many years by \citet{knutson_friends_2014}.
+
+We then fitted a single Keplerian orbit, plus instrument offsets,
+jitters, and a long-term trend
+\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
+on the period and time of inferior conjunction using the values from
+Table~4, and fixed the eccentricity to zero, consistent with results
+from \citet{beerer_secondary_2011}, \citet{knutson_friends_2014} and
+\citet{bonomo_gaps_2017}.  The remaining free parameters were the
+velocity semi-amplitude, the instrument zero-points, the instrument
+jitters (an additive white noise term for each instrument), linear
+($\dot{v_{\rm r}}$), and optionally second-order ($\ddot{v_{\rm r}}$)
+acceleration terms.
+
+We found that the best-fitting model with both linear and quadratic
+radial velocity terms was marginally preferred (by $\Delta
+\mathrm{BIC} = 5.8$) over the best-fitting model with only a linear
+term.  Regardless, for consistency with \citet{knutson_friends_2014},
+who fixed the quadratic component of the long-term trend to zero, in
+Figure~\ref{fig:rvs} we show best-fitting models for the linear-trend
+case.  The best-fit value for the line of sight acceleration,
+\begin{equation}
+  \dot{v}_{\rm r} =
+     -0.0077^{+0.0052}_{-0.0047}
+     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
+\end{equation}
+is within $1\sigma$ of the value reported by
+\citet{knutson_friends_2014}. (Note that the CORALIE data used in our
+and their analyses differ, as we included additional measurements
+reported by \citealt{triaud_spin-orbit_2010}).  The implied period
+derivative is still therefore about an order of magnitude smaller than
+our observed $\dot{P}$ from transit timing.
+
+To summarize, only about one tenth of the observed period decrease can
+be explained through a constant acceleration of the WASP-4 system's
+center of mass.  However, given the limited amount and uneven time
+coverage of the existing radial-velocity data (Figure~\ref{fig:rvs}),
+it remains possible that the center of mass has a more complex motion,
+perhaps due to a companion on an eccentric orbit \citep[{\it e.g.},
+WASP-53 or WASP-81][]{triaud_peculiar_2017}.  It would be useful to
+gather more radial-velocity data to confirm or refute this
+possibility.
 
@@
 
-Some eclipsing binaries exhibit period modulations with amplitudes of
-$\lesssim 0.05\,{\rm days}$ over timescales of decades \citep[{\it
-    e.g.},][]{soderhjelm_geometry_1980,hall_relation_1989}.
-\citet{applegate_mechanism_1992} proposed a mechanism to explain these
-modulations in which the internal structure of a magnetically active
-star changes shape via cyclic exchange of angular momentum between the
-inner and outer zones of the star.  This model could also apply to a
-hot Jupiter orbiting a star with a convective zone.  The changing
-gravitational quadrupole of the star would cause the orbit of the
-planet to precess on the timescale of the stellar activity cycle.  An
-essential difference between this process and apsidal precession is
-that Applegate timing variations need not be strictly periodic
-\citep[{\it e.g.},][Figure~12]{soderhjelm_geometry_1980}. This
-mechanism would also produce transit and occultation timing deviations
-of the same sign, while for apsidal precession they would have
-opposite signs.  For WASP-4, \citet{watson_orbital_2010} estimated
-that the Applegate effect could produce timing deviations of up to 15
-seconds, depending on the modulation period of the stellar dynamo.
-If this
-analysis is accurate, then the Applegate mechanism cannot explain the
-majority our observed $82$ second variation.
-
+A separate candidate explanation for the timing deviations is the
+\citet{applegate_mechanism_1992} effect.  Some eclipsing binaries
+exhibit period modulations with amplitudes of $\lesssim 0.05\,{\rm
+days}$ over timescales of decades \citep[{\it
+e.g.},][]{soderhjelm_geometry_1980,hall_relation_1989}.  The Applegate
+mechanism explains these modulations by positing that the internal
+structure of a magnetically active star changes shape via cyclic
+exchange of angular momentum between the inner and outer zones of the
+star.  This model could also apply to a hot Jupiter orbiting a star
+with a convective zone.  The changing gravitational quadrupole of the
+star would cause the orbit of the planet to precess on the timescale
+of the stellar activity cycle.  An essential difference between this
+process and apsidal precession is that Applegate timing variations
+need not be strictly periodic \citep[{\it
+e.g.},][Figure~12]{soderhjelm_geometry_1980}. This mechanism would
+also produce transit and occultation timing deviations of the same
+sign, while for apsidal precession they would have opposite signs.
+For WASP-4, \citet{watson_orbital_2010} estimated that the Applegate
+effect could produce timing deviations of up to 15 seconds, depending
+on the modulation period of the stellar dynamo.  If this analysis is
+accurate, then the Applegate mechanism cannot explain the majority our
+observed $82$ second variation.
+
+\explain{The following subsection had material moved into the ``Timing
+variation due to line of sight acceleration'' subsection.}
 \subsection{Other possible explanations}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f6.pdf}
+		\includegraphics[width=0.9\textwidth]{f7.pdf}
 	\end{center}
@@
     the timing variations of WASP-4b.} \replaced{Dots}{Symbols} are as
-			in
-		Figure~\ref{fig:times}.  Lines are 100 random draws from the
+    in Figure~\ref{fig:times}.  Lines are 100 random draws from the
     posteriors of the apsidal precession model (orange), and the
-		orbital decay model (blue).
-        \added{The two models may begin to diverge in the mid-2020s}.    
+    orbital decay model (blue).  \added{The two models may begin to
+    diverge in the mid-2020s}.    
 		\label{fig:future}
@@
 
-An apparent change in orbital period can also be caused by an overall
-acceleration of the center of mass of the system.  This would produce
-a period derivative
-\begin{equation}
-	\dot{P} \approx \frac{\dot{v}_{\rm r} P}{c},
-\end{equation}
-where $\dot{v}_{\rm r}$ is the time derivative of the radial velocity.
-\citet{knutson_friends_2014} combined radial-velocity data from their
-own program as well as those of \citet{wilson_wasp-4b_2008},
-\citet{pont_determining_2011}, and \citet{husnoo_observational_2012}
-to obtain a constraint on any long-term trend:
-\begin{equation}
-\dot{v}_{\rm r} =
-   -0.0099^{+0.0052}_{-0.0054}
-   \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
-\end{equation}
-which can also be expressed as a 2$\sigma$ upper limit of $0.021
-\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ on the acceleration along the
-line of sight.  The associated upper limit on the apparent period
-change resulting from this acceleration is $9\times 10^{-11}$, which
-is only one-quarter of the observed period decrease
-(Equation~\ref{eq:dP_dt_obs}).
-
-Thus, it seems that the observed period decrease cannot be fully
-explained as a constant acceleration of the center of mass of WASP-4.
-Given the limited amount and uneven time coverage of the existing
-radial-velocity data, it remains possible that the center of mass has
-a more complex motion, due to a companion on a highly eccentric orbit
-\citep[similar to {\it e.g.}, WASP-53 or WASP-81][]{triaud_peculiar_2017}.
-It would be useful to gather more radial-velocity data to confirm or
-refute this possibility.
-
-There are two final small effects worth noting.  The first is the
-\citet{shklovskii_possible_1970} effect due to the star's proper
-motion, which leads to an apparent period change of $P\mu^2 d/ c$,
-which is only $6\times10^{-13}$ for the case of WASP-4.  The second
-effect, described by \citet{rafikov_stellar_2009}, comes from the
-star's on-sky motion altering our viewing angle, and leads to an
+There are two \replaced{final}{other} small effects worth noting.  The
+first is the \citet{shklovskii_possible_1970} effect due to the star's
+proper motion, which leads to an apparent period change of $P\mu^2 d/
+c$, which is only $6\times10^{-13}$ for the case of WASP-4.  The
+second effect, described by \citet{rafikov_stellar_2009}, comes from
+the star's on-sky motion altering our viewing angle, and leads to an
 observed apsidal precession. The corresponding period change is
@@
 precession would give a rare constraint on the interior density
-distribution of an exoplanet.
+distribution of an exoplanet.\added{ The third possibility --- a
+massive outer companion --- would be the least exotic option, but
+nonetheless a valuable discovery.}
 
@@
 (Figure~\ref{fig:future}).  High-precision transit observations with
-larger telescopes would also be useful. 
-In order to decide between orbital decay and
-apsidal precession, occultation measurements in both the near term and
-also in the mid-2020s will be needed.  More radial-velocity data would
-help in the search for additional bodies that could be causing
-dynamical perturbations, or an overall acceleration of the host star.  
-The transit duration variations are expected to be of order 10 seconds
-\citep{pal_periastron_2008}, and so may remain out of reach.
+larger telescopes would also be useful.  In order to decide between
+orbital decay and apsidal precession, occultation measurements in both
+the near term and also in the mid-2020s will be needed.  More
+radial-velocity data would help in the search for additional bodies
+that could be causing dynamical perturbations, or an overall
+acceleration of the host star.  The transit duration variations are
+expected to be of order 10 seconds \citep{pal_periastron_2008}, and so
+may remain out of reach.
 
@@
 Generation Transit Survey \citep{wheatley_next_2018}, HATPI
-(\href{https://hatpi.org}{hatpi.org}) and
-PLATO \citep{rauer_plato_2014} will also extend
-the time baseline of transit timing for a large number of systems.
+(\href{https://hatpi.org}{hatpi.org}) and PLATO
+\citep{rauer_plato_2014} will also extend the time baseline of transit
+timing for a large number of systems.
 
@@
 supporting this work.
+%
+T.D. acknowledges support from MIT's Kavli Institute as a Kavli
+postdoctoral fellow.
 %
@@
   \texttt{pandas} \citep{mckinney-proc-scipy-2010},
+  \texttt{radvel} \citep{fulton_radvel_2018},
   \texttt{scikit-learn} \citep{scikit-learn},
@@
 		\leavevmode
-		\includegraphics[width=0.98\textwidth]{f7.pdf}
+		\includegraphics[width=0.98\textwidth]{f8.pdf}
 	\end{center}
@@
 
-\newpage
+%\newpage
 

### 2019-03-02 review/referee_response_0.txt
update referee response (ready for JNW)

@@
 
-Referee comment:
+REFEREE COMMENT:
 
@@
 
-TODO TODO TODO
-* TODO: add more weight to the distant companion idea.
-* ADD: our own analysis of the RV data, and the resulting implied period
-  derivative.
-* ADD: RV analysis emphasis (!!).
-
 In general, the manuscript is scientifically accurate and the results presented
-in it certainly will be of interest to other investigators in
-the field. Then, I consider this paper worthy of being published after the
-concerns described below be addressed.
+in it certainly will be of interest to other investigators in the field. Then,
+I consider this paper worthy of being published after the concerns described
+below be addressed.
 
@@
 
-COMMENT:
+  > RESPONSE: We thank the referee for this summary. In our revision, we
+  > have added weight to a third possible explanation that we previously
+  > mentioned but did not sufficiently emphasize: a massive outer
+  > companion, perhaps on an eccentric orbit, might be able to explain
+  > both the transit and radial velocity observations. We have
+  > emphasized this possibility throughout the text where appropriate.
+  > We have opted against developing a joint model (transits+RVs) for
+  > the putative 3-body system, because the RV data show only a weak
+  > linear trend. Detection of a periastron passage or any curvature in
+  > the radial velocities would be needed for this scenario to be truly
+  > convincing.  Instead, we added our own re-analysis of the RV data
+  > (now section 4.3, and figure 6) which confirms the Knutson et al
+  > (2014) result that the star's acceleration towards our line of sight
+  > is only weakly detected, and is an order of magnitude too small to
+  > explain the observed period derivative from the transit times.
 
--- Major points:
+********************************************************************************
+
+REFEREE COMMENT:
+
+Major points:
 
@@
 their sample, must be avoided.
+
 On the other hand, at least two other transits included in their sample show
@@
   > The most important of these statistics are summarized below for the
-  > referee's inspection, and this exercise is described in the text
-  > at the end of Section 3.
+  > referee's inspection, and this exercise is described in the new
+  > subsection "Systematic Concerns" at the end of Section 3.
   > 
@@
 
+REFEREE COMMENT
+
 2) As mentioned in 1), another possible interpretation for explaining the
@@
 
-%FIXME: put in however long the time is!!
-
   > RESPONSE: We have considered this possibility, and have now added
-  > the following text in Section 3.2:
+  > the following text in Section 3.3:
 
   > To explore the effect of possible spot-crossing events on the TESS
-  > transit time measurements, we performed a separate test.  We
-  > injected triangular spot-anomalies with amplitude $0.03\%$ and
-  > duration 30 minutes at random phases into each transit.  The
-  > amplitudes were larger than the spot-crossing anomalies observed
-  > by \citet{southworth_high-precision_2009} and
+  > transit time measurements, we performed a separate test.  We injected
+  > triangular spot-anomalies with amplitude $0.03\%$ and duration 30
+  > minutes at random phases into each transit.  The amplitudes were
+  > larger than the spot-crossing anomalies observed by
+  > \citet{southworth_high-precision_2009} and
   > \citet{sanchis-ojeda_starspots_2011}, and the durations were
-  > comparable.  Spots of these amplitudes would be visible
-  > in the lightcurve residuals, and resemble the anomalies present in
-  > transits ``1360.54'' and ``1372.58'' (Figure
-  > \ref{fig:lightcurves}).
+  > comparable.  Spots of these amplitudes resemble the anomalies present
+  > in transits ``1360.54'' and ``1372.58'' of Figure
+  > \ref{fig:lightcurves}: they are visible, but at low confidence.
   > 
-  > With spots injected, we repeated our measurement of the transit
-  > times.  On average, the meaured transit times did not change with
-  > spots injected, because the flux deviations are symmetric about
-  > the transit minimum.  For individual transits, there were no cases
-  > for which the timing deviation was larger than one minute, but
-  > about one-eighth of the measured transit times were shifted by
-  > more than 30 seconds early.  For the TESS observations to all be
-  > skewed early, it would be necessary for spot-crossing events to
-  > occur in the latter half of every transit \citep[{\it
-  > e.g.},][]{ioannidis_how_2016}.  The lightcurve residuals are on
-  > average symmetric, and do not support this possibility
-  > (Figure~\ref{fig:lightcurves}).  Further, since the stellar
-  > rotation period is between 20 and 40 days
-  > \citep{sanchis-ojeda_starspots_2011,hoyer_tramos_2013}, and the
-  > TESS data span one lunar month, the spot distribution required by
-  > this hypothesis is also rather convoluted.  We conclude that
-  > neither spot-crossing events nor incomplete phase coverage seem to
-  > explain the observed timing variations, and proceed to possible
-  > astrophysical implications.
+  > With spots injected, we repeated our measurement of the transit times.
+  > On average, the measured transit times did not change with spots
+  > injected, because the flux deviations are symmetric about the transit
+  > minimum.  For individual transits, there were no cases for which the
+  > timing deviation was larger than one minute.  The most concerning
+  > shift occurs when the spot anomaly occurs during transit ingress, in
+  > which case the measured mid-time is shifted late by between 30 and 50
+  > seconds \citep[qualitatively similar to results found
+  > by][]{ioannidis_how_2016}.
+  > 
+  > Therefore the TESS observations could hypothetically all be skewed
+  > late if there were spot-crossing events during every ingress.  Two
+  > arguments rule out this possibility.  (1) The lightcurve residuals do
+  > not show evidence for these events.  (2) The stellar rotation period
+  > is between 20 and 40 days, and the sky-projected stellar obliquity is
+  > less than 10 degrees
+  > \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
+  > Since the planet orbits every 1.3 days, requiring that spot anomalies
+  > always occur during ingress would be equivalent to requiring a stellar
+  > spot distribution that is exquisitely selected to match the planet
+  > ingress times.  Such a spot distribution is rather implausible.
+
+REFEREE COMMENT
 
@@
 
-TODO TODO TODO
-* Pending spot studies: no.
+  > RESPONSE: The main result of the paper is that TESS saw WASP-4b
+  > transit 81.6 +/- 11.7 seconds early (~= 7 sigma). Given the
+  > statistical significance, we would prefer to keep the title as-is.
+
+REFEREE COMMENT
 
@@
   > RESPONSE: Done.
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 ii) barycentric Julian date --> Barycentric Julian Date
@@
 
-  > RESPONSE: We have clarified that a "Presearch Data
-  > Conditioning" lightcurve is at a stage of process in which
-  > systematic trends have been removed through the decorrelation
-  > process described by Smith et al (2017).
+  > RESPONSE: We have clarified that a "Presearch Data Conditioning"
+  > lightcurve is at a stage of process in which systematic trends
+  > have been removed through the decorrelation process described by
+  > Smith et al (2017).
   >
@@
   >
-  > As described in the text, the flags "Reaction Wheel
-  > Desaturation Event" and "Manual Exclude" together mean that the
-  > data are potentially of poor quality due to "momentum dumps".
-  > We referred the reader to the data release notes for an
-  > explanation of what a "momentum dump" is, and have added text
-  > to this footnote to improve its clarity.
+  > As described in the text, the flags "Reaction Wheel Desaturation
+  > Event" and "Manual Exclude" together mean that the data are
+  > potentially of poor quality due to "momentum dumps".  We referred
+  > the reader to the data release notes for an explanation of what a
+  > "momentum dump" is, and have added text to this footnote to
+  > improve its clarity.
+
+REFEREE COMMENT
 
@@
   > needed.
+
+REFEREE COMMENT
 
@@
   > clarify that we mean the photometric data points.
+
+REFEREE COMMENT
 
@@
   > as-is.
+
+REFEREE COMMENT
 
@@
   > We did not explicitly describe our choice for the initial
-  > parameters, because the converged best-fit values from the
-  > chain do not depend on them.  For reproducibility purposes,
-  > similar best-fit values can be obtained independent of initial
-  > parameter choices. Barring an overruling request from the
-  > referee, we will not mention our explicit choice in the text
-  > (it was to use values from Hoyer et al., 2013).
+  > parameters, because the converged best-fit values from the chain
+  > do not depend on them.  For reproducibility purposes, similar
+  > best-fit values can be obtained independent of initial parameter
+  > choices. Barring an overruling request from the referee, we will
+  > not mention our explicit choice in the text (it was to use values
+  > from Hoyer et al., 2013).
   >
   > The transit parameters from the phase-folded fit: Rp/Rstar,
-  > inclination, a/Rstar, and limb darkening parameters are given
-  > in the first five lines of Table 1.
+  > inclination, a/Rstar, and limb darkening parameters are given in
+  > the first five lines of Table 1.
   >
@@
   > individual transits, and the linear fit parameters. Since the
-  > focus of this study is the transit times, we would prefer to
-  > omit these additional parameters for the sake of concision.
+  > focus of this study is the transit times, we omitted these
+  > additional parameters for the sake of concision.
   >
-  > For the same reason, we leave the interested reader to explore
-  > differences in the stellar and planetary parameters by citing
-  > relevant work in the last paragraph of Section 2.3. From e.g.,
-  > Table 8 of Southworth+2009, we see that our values agree with
-  > those of previous studies, and so do not merit further
-  > attention, given the main focus of the present work.
+  > Regarding the suggestion to add a table comparing our
+  > stellar/planet parameters with those found by other investigators:
+  > again, because the focus of the paper is the transit times, we
+  > have left the interested reader to explore these differences by
+  > pointing them to the relevant studies in the last paragraph of
+  > Section 2.3. From e.g., Table 8 of Southworth+2009, we see that
+  > our values do agree with those of previous studies, and so do not
+  > merit further attention, given the main focus of the present work.
+
+REFEREE COMMENT
 
@@
   > visualization purposes.
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 v) Is χ^2=9.2 the reduced chi-square? If it is so, It should be clarified by
@@
   > corrected the typo in the text, in which we wrote "reduced
-  > $\chi^2$'.
+  > $\chi^2$", which likely led to the confusion!
+
+REFEREE COMMENT
 
@@
   > would expect $\chi^2 = 16 \pm 5.7$.
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 Section 2.3:
@@
   > RESPONSE: Done, thank you.
+
+REFEREE COMMENT
 
@@
   > uncorrelated and normally distributed initial measurements."
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 ii) Could the authors specify what is the meaning of each of the variables in
@@
   > wording accordingly.
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 Also: ii-1) In the legends of both panels, could the authors specify which is
@@
 
+REFEREE COMMENT
+
 ii-2) "The red band shows the average deviation of the TESS transits,..." -->
@@
 
+REFEREE COMMENT
+
 iii) It would be important to introduce references for equations (4) and (5).
@@
 
-  > RESPONSE: Equations (4) and (5) are generic equations for the Taylor series
-  > expansion of the transit and occultation times as a function of epoch, and
-  > do not require references.  We have added the following sentence to describe the
-  > sidereal and anomalistic periods: "The sidereal period is the duration
-  > required to return to the same orientation with respect to the stars; the
-  > slightly longer anomalistic period is the duration required to reach a
-  > fixed longitude with respect to the rotating line of apsides."
+  > RESPONSE: Equations (4) and (5) are generic equations for the
+  > Taylor series expansion of the transit and occultation times as a
+  > function of epoch, and so do not require references.  We have
+  > added the following sentence to describe the sidereal and
+  > anomalistic periods: "The sidereal period is the duration required
+  > to return to the same orientation with respect to the stars; the
+  > slightly longer anomalistic period is the duration required to
+  > reach a fixed longitude with respect to the rotating line of
+  > apsides."
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 v) Similar to i) How did you fit the mid-transit times and occultations shown
@@
   > RESPONSE: We have added the following sentences in Section 3.2:
-  > "We fit each model by assuming a Gaussian likelihood and sampling over the
-  > posterior probability distributions. The prior for the quadratic model
-  > allowed the period derivative to have any sign." We have clarified the
-  > wording throughout from "decreasing-period model" to "quadratic model" to
-  > reflect the latter point.
+  > "We fit each model by assuming a Gaussian likelihood and sampling
+  > over the posterior probability distributions. The prior for the
+  > quadratic model allowed the period derivative to have any sign."
+  > We have clarified the wording throughout from "decreasing-period
+  > model" to "quadratic model" to reflect the latter point.
+
+REFEREE COMMENT
 
@@
 
+REFEREE COMMENT
+
 vii) Given that the mid-transit times values used in this work are taken from
@@
 
-TODO TODO TODO
-* We are trusting that people fit transits.
-* (Maybe cite some of the Espinoza studies about detailed transit fitting
-  systematics)
+  > RESPONSE: We have added the following paragraph at the end of
+  > Section 3.3:
 
-********************************************************************************
+  > There is a final systematic concern in our global methodology.  We
+  > have opted to use mid-transit time values derived by different
+  > authors, who used heterogeneous methods to fit and detrend their
+  > lightcurves.  One important assumption is of course that these authors
+  > have correctly documented their time systems.  Further, though many
+  > assumptions in transit-fitting ({\it e.g.}, concering limb-darkening
+  > and eccentricity) do not affect transit-time measurements, different
+  > detrending approaches can asymmetrically warp transits and shift
+  > mid-transit times.  The magnitude of this systematic effect is hard to
+  > quantify, but the situation is fairly clear from
+  > Figure~\ref{fig:times}.  Many independent authors provided transit
+  > measurements shortly after WASP-4b's discovery, so the data should on
+  > average be acceptable.  \citet{huitson_gemini_2017} provided the most
+  > important data from epochs 0-1000, and so if their data were
+  > systematically affected by detrending choices at the level of many
+  > times their reported uncertainties, our reported timing anomaly would
+  > be suspect.  For this reason, we have paid extra attention to the
+  > \citet{huitson_gemini_2017} data set, corresponding with the authors
+  > to confirm that such systematic concerns would not affect their
+  > results at the many-$\sigma$ level.
+
+REFEREE COMMENT
+
 4. INTERPRETATION
@@
 
-  > RESPONSE: Corrected.
+  > RESPONSE: Corrected, thank you.
+
+REFEREE COMMENT:
 
@@
   > RESPONSE: Done.
+
+REFEREE COMMENT:
 
@@
   > RESPONSE: Changed wording to remove any ambiguity.
+
+REFEREE COMMENT:
 
@@
   > modified the text for clarity on this point.
+
+REFEREE COMMENT:
 
@@
 
+REFEREE COMMENT:
+
 Section 4.2:
@@
   > interested reader to refer to the original study!
+
+REFEREE COMMENT:
 
@@
 
+REFEREE COMMENT:
+
 iii) Mb is indicated as the mass of the hypothetical planet WASP-4c, however,
@@
   > RESPONSE: We have correcly the wording accordingly.
+
+REFEREE COMMENT:
 
@@
   > hypothesis (Equation 18) is imprecise.
+
+REFEREE COMMENT:
 
@@
 
-********************************************************************************
-
-Referee comment:
+REFEREE COMMENT:
 
@@
   > RESPONSE: Corrected.
+
+REFEREE COMMENT:
 
@@
 ********************************************************************************
-Other changes:
+OTHER CHANGES:
 
@@
   Douglas Caldwell.
+
+* Minor edits in text throughout, as indicated with the \added{},
+  \deleted{}, and \replaced{}{} commands.
+
+* Minor tweaks to figures.

### 2019-03-02 review/referee_response_0.txt
referee response -- LGB done

@@
   > timing deviation was larger than one minute.  The most concerning
-  > shift occurs when the spot anomaly occurs during transit ingress, in
-  > which case the measured mid-time is shifted late by between 30 and 50
+  > shift occurs when the spot anomaly occurs during transit egress, in
+  > which case the measured mid-time is shifted early by between 30 and 50
   > seconds \citep[qualitatively similar to results found
@@
   > Therefore the TESS observations could hypothetically all be skewed
-  > late if there were spot-crossing events during every ingress.  Two
+  > early if there were spot-crossing events during every egress.  Two
   > arguments rule out this possibility.  (1) The lightcurve residuals do
@@
   > Since the planet orbits every 1.3 days, requiring that spot anomalies
-  > always occur during ingress would be equivalent to requiring a stellar
-  > spot distribution that is exquisitely selected to match the planet
-  > ingress times.  Such a spot distribution is rather implausible.
+  > always occur during egress would be equivalent to requiring a stellar
+  > spot distribution that is exquisitely (and thus implausibly)
+  > distributed to match the planet egress times.
 
@@
   > stellar and planetary parameters are propagated according to
-  > standard analytic formulae, under the assumption of
-  > uncorrelated and normally distributed initial measurements."
+  > standard analytic formulae, under the assumption of uncorrelated
+  > and normally distributed initial measurements."
 
@@
 
-  > RESPONSE: we have modified the wording slightly to clarify. The point is
-  > that for these Gemini South observations, we confirmed with the authors
-  > through back and forth correspondence that the details of their
-  > time-keeping were sound.  We made this extra effort because of the weight
-  > of the Huitson+(2017) times in the analysis.
+  > RESPONSE: we have modified the wording slightly to clarify. The
+  > point is that for these Gemini South observations, we confirmed
+  > with the authors through back and forth correspondence that the
+  > details of their time-keeping were sound.  We made this extra
+  > effort because of the weight of the Huitson+(2017) times in the
+  > analysis.
   >
-  > For the remaining data points, we used the references stated in Section 3.1
-  > and Table 2 to confirm that we were using with mid-exposure times, and
-  > corrected from HJD_UTC to BJD_TDB as appropriate. However the point is
-  > indeed that because of the weight of the Huitson et al (2017) times, we
-  > made an particular extra effort to confirm their accuracy with the original
-  > authors.
+  > For the remaining data points, we used the references stated in
+  > Section 3.1 and Table 2 to confirm that we were using with
+  > mid-exposure times, and corrected from HJD_UTC to BJD_TDB as
+  > appropriate. However the point is indeed that because of the
+  > weight of the Huitson et al (2017) times, we made an particular
+  > extra effort to confirm their accuracy with the original authors.
 
@@
 
-  > RESPONSE: Though we appreciate the suggestion, we have opted to keep the
-  > "title" as is. The nuance of what "predicted" means for this plot (a linear
-  > ephemeris, based only on the pre-TESS times) is sufficient that we find it
-  > to be better left in the body of the caption.
+  > RESPONSE: Though we appreciate the suggestion, we have opted to
+  > keep the "title" as is. The nuance of what "predicted" means for
+  > this plot (a linear ephemeris, based only on the pre-TESS times)
+  > is sufficient that we find it to be better left in the body of the
+  > caption.
 
@@
 
-  > RESPONSE: We have added the following sentence in the caption: "The binned
-  > TESS point is the weighted average of 18 TESS transits".  The caption of
-  > Figure 4 already stated the same.  We have changed the words "Dots" to
-  > "Symbols" in the caption of Figure 6 in order to clarify this for the
-  > reader.
+  > RESPONSE: We have added the following sentence in the caption:
+  > "The binned TESS point is the weighted average of 18 TESS
+  > transits".  The caption of Figure 4 already stated the same.  We
+  > have changed the words "Dots" to "Symbols" in the caption of
+  > Figure 6 in order to clarify this for the reader.
 
@@
 
-  > RESPONSE: We have added the following paragraph at the end of
+  > RESPONSE: We have added the following paragraphs at the end of
   > Section 3.3:
@@
   > authors, who used heterogeneous methods to fit and detrend their
-  > lightcurves.  One important assumption is of course that these authors
-  > have correctly documented their time systems.  Further, though many
-  > assumptions in transit-fitting ({\it e.g.}, concering limb-darkening
-  > and eccentricity) do not affect transit-time measurements, different
-  > detrending approaches can asymmetrically warp transits and shift
-  > mid-transit times.  The magnitude of this systematic effect is hard to
-  > quantify, but the situation is fairly clear from
-  > Figure~\ref{fig:times}.  Many independent authors provided transit
-  > measurements shortly after WASP-4b's discovery, so the data should on
-  > average be acceptable.  \citet{huitson_gemini_2017} provided the most
-  > important data from epochs 0-1000, and so if their data were
-  > systematically affected by detrending choices at the level of many
-  > times their reported uncertainties, our reported timing anomaly would
-  > be suspect.  For this reason, we have paid extra attention to the
-  > \citet{huitson_gemini_2017} data set, corresponding with the authors
-  > to confirm that such systematic concerns would not affect their
-  > results at the many-$\sigma$ level.
+  > lightcurves.  This of course assumes that these authors have
+  > correctly documented their time systems.  Further, though many
+  > choices in transit-fitting ({\it e.g.}, parametrization of
+  > limb-darkening and eccentricity) do not affect transit mid-time
+  > measurements, different detrending approaches can asymmetrically
+  > warp transits and shift mid-transit times.  The magnitude of this
+  > systematic effect is hard to quantify, but the situation is fairly
+  > clear from Figure~\ref{fig:times}.  Many independent authors
+  > provided transit measurements shortly after WASP-4b's discovery,
+  > so the data should on average be acceptable.
+  > \citet{huitson_gemini_2017} provided the most important data from
+  > epochs 0-1000, and so if their data were systematically affected
+  > by detrending choices at the level of many times their reported
+  > uncertainties, our timing anomaly would be suspect.  For this
+  > reason, we paid careful attention to the
+  > \citet{huitson_gemini_2017} data set, and corresponded with the
+  > authors to confirm that their results are not affected by
+  > systematic effects of the required amplitude.
+  > 
+  > None of the concerns mentioned in this subsection are likely
+  > explanations for the observed timing variations. We proceed by
+  > considering possible astrophysical explanations.
 
@@
 
-  > RESPONSE: To make Figure 4 and 6, we start with the observed transit and
-  > occultation mid-times, as a function of epoch number.  We then fit a line
-  > to these data.  If the data were generated by a constant period orbit (with
-  > gaussian uncertainties) the residuals to such a fit would show no
-  > correlation with epoch number.  However if the data were generated by a
-  > planet with a constant period derivative (with gaussian uncertainties), the
-  > residuals to such a fit would always be a parabola.  This is because the
-  > linear fit to the data will have contributions from the beginning, middle,
-  > and end of the transit timeseries. Thus the best-fit linear ephemeris (to
-  > data that actually have a decreasing period) has a best-fit period
-  > somewhere between the minimum and maximum period. At the earliest epochs,
-  > the best-fit period is an underestimate, so you see the increase in O-C. At
-  > the latest epochs, the best-fit period is an overestimate, so you see the
+  > RESPONSE: To make Figure 4 and 6, we start with the observed
+  > transit and occultation mid-times, as a function of epoch number.
+  > We then fit a line to these data.  If the data were generated by a
+  > constant period orbit (with gaussian uncertainties) the residuals
+  > to such a fit would show no correlation with epoch number.
+  > However if the data were generated by a planet with a constant
+  > period derivative (with gaussian uncertainties), the residuals to
+  > such a fit would always be a parabola.  This is because the linear
+  > fit to the data will have contributions from the beginning,
+  > middle, and end of the transit timeseries. Thus the best-fit
+  > linear ephemeris (to data that actually have a decreasing period)
+  > has a best-fit period somewhere between the minimum and maximum
+  > period. At the earliest epochs, the best-fit period is an
+  > underestimate, so you see the increase in O-C. At the latest
+  > epochs, the best-fit period is an overestimate, so you see the
   > decrease.
@@
 
-  > RESPONSE: we have opted against this suggestion, chiefly to encourage the
-  > interested reader to refer to the original study!
+  > RESPONSE: we have opted against this suggestion, chiefly to
+  > encourage the interested reader to refer to the original study!
 
@@
 
-  > RESPONSE: As in Figure 3, we appreciate the suggestion, but have opted to
-  > keep the "title" as is. We have appended the sentence "The two models may
-  > begin to diverge in the mid-2020s" as further clarification for the reader.
+  > RESPONSE: As in Figure 3, we appreciate the suggestion, but have
+  > opted to keep the "title" as is. We have appended the sentence
+  > "The two models may begin to diverge in the mid-2020s" as further
+  > clarification for the reader.
 
@@
 
-  > RESPONSE: We have clarified the origin of the values used when evaluating
-  > Equation 21.  The tricky ones -- the mass and radius of the convective
-  > envelope -- came from the MESA/MIST projects, for which we have added
-  > appropriate citations.  Our thanks to the referee for pointing out this
-  > omission.  We have opted to not explicitly define the reduced mass, as it
-  > is standard knowledge in the gravitational two-body problem.
+  > RESPONSE: We have clarified the origin of the values used when
+  > evaluating Equation 21.  The tricky ones -- the mass and radius of
+  > the convective envelope -- came from the MESA/MIST projects, for
+  > which we have added appropriate citations.  Our thanks to the
+  > referee for pointing out this omission.  We have opted to not
+  > explicitly define the reduced mass, as it is standard knowledge in
+  > the gravitational two-body problem.
 

### 2019-03-06 paper/ms.tex
Update ms.tex

@@
 \affiliation{ Department of Astrophysical Sciences, Princeton
-University, 4 Ivy Lane, Princeton, NJ 08540, USA}
 %
@@
 occultation times should eventually reveal \replaced{two
-distinct}{three select} processes.
+distinct}{variations caused by three different} \replaced{processes}{phenomena}.
 
@@
 disturbances (a problem reviewed by \citealt{Mazeh2008} and
-\citealt{ogilvie_tidal_2014}).
-
-Second, if hot Jupiter orbits are at all eccentric, then long-term
+\citealt{ogilvie_tidal_2014}). Second, if hot Jupiter orbits are appreciably eccentric, then long-term
 timing studies should reveal rotation of the orbital ellipse within
@@
 \citep{ragozzine_probing_2009}.
-
 The most convincing direct evidence yet found for either orbital decay
@@
 \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
-to search for departures from a constant period, because transit data
-exist all the way back to 2007.  The orbital eccentricity is less than
+to search for departures from a constant period, because transits
+have been observed since 2007.  The orbital eccentricity is less than
 0.018 (2$\sigma$), based on the work of \cite{knutson_friends_2014},
@@
 observations, and \S~\ref{sec:timing} describes our timing analysis.
-We tried fitting the data with three models: a constant period, a
-decaying period, and a slightly eccentric precessing orbit.  A
+We tried fitting the data with three models: a constant period; a
+steadily shrinking period; and a slightly eccentric, precessing orbit.  A
 constant period can be ruled out\replaced{, but we}{. We} cannot
@@
 surrounding WASP-4 were averaged into 2-minute stacks by the onboard
-computer.  The data were downlinked via the Deep Space Network, and
+computer.  The data were downlinked via the Deep Space Network\footnote{\url{https://deepspace.jpl.nasa.gov}}, and
 the spacecraft timestamps were calibrated against the ground-station
@@
 timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
-appropriate 2457000 day offset
+appropriate 2{,}457{,}000 day offset
 \citep{tess_data_product_description_2018}.  Many of these and
@@
 zodiacal background light.  The level of background light appears to
-have been overestimated.
+have been overestimated in the model.
 
@@
 submission, per the referee's suggestions.}
-\subsection{Systematic concerns}
-
-\paragraph{Omitting suspect pre-TESS transits}
+\subsection{Possible systematic errors}
+
 To assess the robustness of these results, we considered a few
-possible systematic effects in the timing dataset.  One concern in the
-pre-TESS measurements was incomplete phase coverage: a handful
-of the transit times in Table~2 are from lightcurves with partial
+possible systematic effects in the timing dataset.
+
+\paragraph{Suspect pre-TESS light curves}
+Some of the
+pre-TESS light curves have incomplete phase coverage: a handful
+of the transit times in Table~2 are from lightcurves with 
 gaps.  A separate issue is the effect of spot-crossing anomalies on
-mid-transit time measurements.  To address these concerns, we repeated
+transit timings.  To address these concerns, we repeated
 the model-fitting described above, but omitted epochs -827, -804,
@@
 triangular spot-anomalies with amplitude $0.03\%$ and duration 30
-minutes at random phases into each transit.  The amplitudes were
-larger than the spot-crossing anomalies observed by
+minutes at random phases into each transit.  The amplitude was
+chosen to be larger than the spot-crossing anomalies observed by
 \citet{southworth_high-precision_2009} and
-\citet{sanchis-ojeda_starspots_2011}, and the durations were
-comparable.  Spots of these amplitudes resemble the anomalies present
+\citet{sanchis-ojeda_starspots_2011}, and the duration was chosen
+to be comparable to those of previously observed events.
+Spots of these amplitudes resemble the possible anomalies present
 in transits ``1360.54'' and ``1372.58'' of Figure
-\ref{fig:lightcurves}: they are visible, but at low confidence.
+\ref{fig:lightcurves}.
 
 With spots injected, we repeated our measurement of the transit times.
-On average, the measured transit times did not change with spots
-injected, because the flux deviations are symmetric about the transit
-minimum.  For individual transits, there were no cases for which the
-timing deviation was larger than one minute.  The most concerning
-shift occurs when the spot anomaly occurs during transit egress, in
+On average, the measured transit times did not change after injecting spots,
+because the flux deviations are equally likely to occur
+in the first and second halves of the transit. 
+For individual transits, there were no cases for which the
+timing deviation was larger than one minute.  The largest 
+shifts occur when the spot anomaly occurs during transit ingress or egress, in
 which case the measured mid-time is shifted early by between 30 and 50
@@
 \paragraph{Detrending choices in pre-TESS data}
-There is a final systematic concern in our global methodology.  We
-have opted to use mid-transit time values derived by different
+There is a final concern that is difficult to address.  We
+collected the mid-transit time values derived by different
 authors, who used heterogeneous methods to fit and detrend their
-lightcurves.  This of course assumes that these authors
-have correctly documented their time systems.  Further, though many
+lightcurves.  We have also assumed that these authors
+have correctly documented the time systems in which the data are
+reported.  Further, though many
 choices in transit-fitting ({\it e.g.}, parametrization of
@@
 clear from Figure~\ref{fig:times}.  Many independent authors provided
-transit measurements shortly after WASP-4b's discovery, so the data
-should on average be acceptable.  \citet{huitson_gemini_2017} provided
-the most important data from epochs 0-1000, and so if their data were
-systematically affected by detrending choices at the level of many
-times their reported uncertainties, our timing anomaly would
-be suspect.  For this reason, we paid careful attention to the
+transit measurements shortly after WASP-4b's discovery, and the data are
+consistent with each other. \citet{huitson_gemini_2017} provided
+the most important data from epochs 0-1000. If their data were
+systematically affected by detrending choices or time-system confusion
+at the level of several
+times their reported uncertainties, then it possible that the orbital
+period is constant despite the evidence in the TESS data.
+For this reason, we paid careful attention to the
 \citet{huitson_gemini_2017} data set, and corresponded with the authors
@@
 
-None of the concerns mentioned in this subsection are likely
-explanations for the observed timing variations. We proceed by
+None of the concerns mentioned in this subsection seem likely
+to explain the observed timing variations. We proceed by
 considering possible astrophysical explanations.
@@
 \begin{equation}
-  \frac{P}{ \dot{P} } = 9.2 \, {\rm Myr}.
+  \frac{P}{ {\rm d}P/{\rm d}t } = 9.2 \, {\rm Myr}.
 \end{equation}
@@
 \citet{pont_determining_2011}, and \citet{husnoo_observational_2012},
-found a weak long-term trend in WASP-4:
+found evidence for long-term trend in WASP-4, with low statistical significance:
 \begin{equation}
@@
 \times10^{-11}$. The period decrease from the observed RV trend is an
-order of magnitude smaller than our observed $\dot{P}$ from transit
+order of magnitude smaller than the observed $\dot{P}$ from transit
 times, $(4.0\pm 0.4)\times10^{-10}$ (Table~4).
@@
 significant acceleration, and with the correct sign needed to explain
-the transit timing variations.  Since the radial velocities could
-affect the interpretation of the system, we redetermined the line of
-sight acceleration as follows.
+the transit timing variations.  Given the potential importance of the radial velocities for
+interpreting this system, we performed an independent analysis,
+as follows.
 

### 2019-03-06 review/referee_response_0.txt
Update referee_response_0.txt

@@
-Dear Dr. Endl,
+bDear Dr. Endl,
 
@@
   > both the transit and radial velocity observations. We have
-  > emphasized this possibility throughout the text where appropriate.
-  > We have opted against developing a joint model (transits+RVs) for
-  > the putative 3-body system, because the RV data show only a weak
-  > linear trend. Detection of a periastron passage or any curvature in
-  > the radial velocities would be needed for this scenario to be truly
-  > convincing.  Instead, we added our own re-analysis of the RV data
-  > (now section 4.3, and figure 6) which confirms the Knutson et al
+  > described this possibility throughout the text where appropriate.
+  > We also added our own re-analysis of the extant RV data
+  > (now section 4.3, and figure 6) to check on the Knutson et al
   > (2014) result that the star's acceleration towards our line of sight
@@
 
-  > RESPONSE: Barring any specific suggestion by the referee, we will leave
-  > this one for the reader to pursue at their leisure.
+  > RESPONSE: We are not aware of any paper that suits the purpose here,
+  > so we added a citation to the URL of the Deep Space Network.
 
@@
 
-  > So yes, we used the same transit model as before (the BATMAN
+  > In short, we used the same transit model as before (the BATMAN
   > code), but with the addition of the slope and intercept of a

### 2019-03-06 paper/ms.tex
fix typos in d1984d4d5df426e0f423d67c48df9de8f4627960

@@
 \affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
 %
@@
 is to search for timing anomalies in \added{previously known }hot
-Jupiter systems.  Long-term monitoring of hot Jupiter transit and
-occultation times should eventually reveal \replaced{two
-distinct}{variations caused by three different} \replaced{processes}{phenomena}.
+Jupiter sytems.  Long-term monitoring of hot Jupiter transit and
+occultation times should eventually reveal \replaced{two distinct
+processes}{variations caused by three different phenomena}.
 
@@
 disturbances (a problem reviewed by \citealt{Mazeh2008} and
-\citealt{ogilvie_tidal_2014}). Second, if hot Jupiter orbits are appreciably eccentric, then long-term
-timing studies should reveal rotation of the orbital ellipse within
-the orbital plane (``apsidal precession'').  If this effect were
-observed, it could yield a measure of the planet's Love number, which
-would constrain the planet's interior structure
-\citep{ragozzine_probing_2009}.
-The most convincing direct evidence yet found for either orbital decay
-or apsidal precession of a hot Jupiter is the case of WASP-12b, which
-has a transit period that has decreased by about 30~milliseconds per
-year over the last decade
-\citep{maciejewski_departure_2016,patra_2017}.
+\citealt{ogilvie_tidal_2014}). Second, if hot Jupiter orbits are
+appreciably eccentric, then long-term timing studies should reveal
+rotation of the orbital ellipse within the orbital plane (``apsidal
+precession'').  If this effect were observed, it could yield a measure
+of the planet's Love number, which would constrain the planet's
+interior structure \citep{ragozzine_probing_2009}.  The most
+convincing direct evidence yet found for either orbital decay or
+apsidal precession of a hot Jupiter is the case of WASP-12b, which has
+a transit period that has decreased by about 30~milliseconds per year
+over the last decade \citep{maciejewski_departure_2016,patra_2017}.
 
@@
 \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
-to search for departures from a constant period, because transits
-have been observed since 2007.  The orbital eccentricity is less than
-0.018 (2$\sigma$), based on the work of \cite{knutson_friends_2014},
-who combined the available transit times, occultation times, and
-Doppler data.  The sky projection of the stellar obliquity is also
-compatible with zero, within about 10 degrees
+to search for departures from a constant period, because transits have
+been observed since 2007.  The orbital eccentricity is less than 0.018
+(2$\sigma$), based on the work of \cite{knutson_friends_2014}, who
+combined the available transit times, occultation times, and Doppler
+data.  The sky projection of the stellar obliquity is also compatible
+with zero, within about 10 degrees
 \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011}.
@@
 We tried fitting the data with three models: a constant period; a
-steadily shrinking period; and a slightly eccentric, precessing orbit.  A
-constant period can be ruled out\replaced{, but we}{. We} cannot
+steadily shrinking period; and a slightly eccentric, precessing orbit.
+A constant period can be ruled out\replaced{, but we}{. We} cannot
 distinguish between the \replaced{latter two possibilities}{
@@
 surrounding WASP-4 were averaged into 2-minute stacks by the onboard
-computer.  The data were downlinked via the Deep Space Network\footnote{\url{https://deepspace.jpl.nasa.gov}}, and
-the spacecraft timestamps were calibrated against the ground-station
+computer.  The data were downlinked via the Deep Space
+Network\footnote{\url{https://deepspace.jpl.nasa.gov}}, and the
+spacecraft timestamps were calibrated against the ground-station
 clocks.  The spacecraft clock times were then transformed by the
@@
 \paragraph{Suspect pre-TESS light curves}
-Some of the
-pre-TESS light curves have incomplete phase coverage: a handful
-of the transit times in Table~2 are from lightcurves with 
+Some of the pre-TESS light curves have incomplete phase coverage: a
+handful of the transit times in Table~2 are from lightcurves with
 gaps.  A separate issue is the effect of spot-crossing anomalies on
-transit timings.  To address these concerns, we repeated
-the model-fitting described above, but omitted epochs -827, -804,
--537, and -208 because of gaps in their coverage.  We also omitted
-epochs -526 and -561 because of visible spot anomalies during the
-transits. (All epoch numbers are as in Table~2.)  The resulting
-best-fit transit timing model parameters were all within $1\sigma$ of
-the values quoted in Table~4.  The uncertainties, goodness-of-fit
-statistics, and model comparison statistics did not appreciably
-change.
+transit timings.  To address these concerns, we repeated the
+model-fitting described above, but omitted epochs -827, -804, -537,
+and -208 because of gaps in their coverage.  We also omitted epochs
+-526 and -561 because of visible spot anomalies during the transits.
+(All epoch numbers are as in Table~2.)  The resulting best-fit transit
+timing model parameters were all within $1\sigma$ of the values quoted
+in Table~4.  The uncertainties, goodness-of-fit statistics, and model
+comparison statistics did not appreciably change.
 
@@
 triangular spot-anomalies with amplitude $0.03\%$ and duration 30
-minutes at random phases into each transit.  The amplitude was
-chosen to be larger than the spot-crossing anomalies observed by
+minutes at random phases into each transit.  The amplitude was chosen
+to be larger than the spot-crossing anomalies observed by
 \citet{southworth_high-precision_2009} and
-\citet{sanchis-ojeda_starspots_2011}, and the duration was chosen
-to be comparable to those of previously observed events.
-Spots of these amplitudes resemble the possible anomalies present
-in transits ``1360.54'' and ``1372.58'' of Figure
-\ref{fig:lightcurves}.
+\citet{sanchis-ojeda_starspots_2011}, and the duration was chosen to
+be comparable to those of previously observed events.  Spots of these
+amplitudes resemble the possible anomalies present in transits
+``1360.54'' and ``1372.58'' of Figure \ref{fig:lightcurves}.
 
 With spots injected, we repeated our measurement of the transit times.
-On average, the measured transit times did not change after injecting spots,
-because the flux deviations are equally likely to occur
-in the first and second halves of the transit. 
-For individual transits, there were no cases for which the
-timing deviation was larger than one minute.  The largest 
-shifts occur when the spot anomaly occurs during transit ingress or egress, in
-which case the measured mid-time is shifted early by between 30 and 50
-seconds \citep[qualitatively similar to results found
+On average, the measured transit times did not change after injecting
+spots, because the flux deviations are equally likely to occur in the
+first and second halves of the transit.  For individual transits,
+there were no cases for which the timing deviation was larger than one
+minute.  The largest shifts occur when the spot anomaly occurs during
+transit ingress or egress, in which case the measured mid-time is
+shifted either late or early by between 30 and 50 seconds
+\citep[qualitatively similar to results found
 by][]{ioannidis_how_2016}.
@@
 \paragraph{Detrending choices in pre-TESS data}
-There is a final concern that is difficult to address.  We
-collected the mid-transit time values derived by different
-authors, who used heterogeneous methods to fit and detrend their
-lightcurves.  We have also assumed that these authors
-have correctly documented the time systems in which the data are
-reported.  Further, though many
-choices in transit-fitting ({\it e.g.}, parametrization of
-limb-darkening and eccentricity) do not affect transit mid-time
-measurements, different detrending approaches can asymmetrically warp
-transits and shift mid-transit times.  The magnitude of this
-systematic effect is hard to quantify, but the situation is fairly
-clear from Figure~\ref{fig:times}.  Many independent authors provided
-transit measurements shortly after WASP-4b's discovery, and the data are
-consistent with each other. \citet{huitson_gemini_2017} provided
-the most important data from epochs 0-1000. If their data were
+There is a final concern that is difficult to address.  We collected
+the mid-transit time values derived by different authors, who used
+heterogeneous methods to fit and detrend their lightcurves.  We have
+also assumed that these authors have correctly documented the time
+systems in which the data are reported.  Further, though many choices
+in transit-fitting ({\it e.g.}, parametrization of limb-darkening and
+eccentricity) do not affect transit mid-time measurements, different
+detrending approaches can asymmetrically warp transits and shift
+mid-transit times.  The magnitude of this systematic effect is hard to
+quantify, but the situation is fairly clear from
+Figure~\ref{fig:times}.  Many independent authors provided transit
+measurements shortly after WASP-4b's discovery, and the data are
+consistent with each other. \citet{huitson_gemini_2017} provided the
+most important data from epochs 0-1000. If their data were
 systematically affected by detrending choices or time-system confusion
-at the level of several
-times their reported uncertainties, then it possible that the orbital
-period is constant despite the evidence in the TESS data.
-For this reason, we paid careful attention to the
-\citet{huitson_gemini_2017} data set, and corresponded with the authors
-to confirm that their results are not affected by systematic 
+at the level of several times their reported uncertainties, then it
+possible that the orbital period is constant despite the evidence in
+the TESS data.  For this reason, we paid careful attention to the
+\citet{huitson_gemini_2017} data set, and corresponded with the
+authors to confirm that their results are not affected by systematic
 effects of the required amplitude.
 
-None of the concerns mentioned in this subsection seem likely
-to explain the observed timing variations. We proceed by
-considering possible astrophysical explanations.
+None of the concerns mentioned in this subsection seem likely to
+explain the observed timing variations. We proceed by considering
+possible astrophysical explanations.
 
@@
 \citet{pont_determining_2011}, and \citet{husnoo_observational_2012},
-found evidence for long-term trend in WASP-4, with low statistical significance:
+found evidence for long-term trend in WASP-4, with low statistical
+significance:
 \begin{equation}
@@
 significant acceleration, and with the correct sign needed to explain
-the transit timing variations.  Given the potential importance of the radial velocities for
-interpreting this system, we performed an independent analysis,
-as follows.
+the transit timing variations.  Given the potential importance of the
+radial velocities for interpreting this system, we performed an
+independent analysis, as follows.
 

### 2019-03-06 review/referee_response_0.txt
fix typos in bd1c6634d4620f19dce3c4596df1ece60ed3885b

@@
-bDear Dr. Endl,
+Dear Dr. Endl,
 
@@
 
-  > RESPONSE: We thank the referee for this summary. In our revision, we
-  > have added weight to a third possible explanation that we previously
-  > mentioned but did not sufficiently emphasize: a massive outer
-  > companion, perhaps on an eccentric orbit, might be able to explain
-  > both the transit and radial velocity observations. We have
+  > RESPONSE: We thank the referee for this summary. In our revision,
+  > we have added weight to a third possible explanation that we
+  > previously mentioned but did not sufficiently emphasize: a massive
+  > outer companion, perhaps on an eccentric orbit, might be able to
+  > explain both the transit and radial velocity observations. We have
   > described this possibility throughout the text where appropriate.
-  > We also added our own re-analysis of the extant RV data
-  > (now section 4.3, and figure 6) to check on the Knutson et al
-  > (2014) result that the star's acceleration towards our line of sight
-  > is only weakly detected, and is an order of magnitude too small to
+  > We also added our own re-analysis of the extant RV data (now
+  > section 4.3, and figure 6) to check on the Knutson et al (2014)
+  > result that the star's acceleration towards our line of sight is
+  > only weakly detected, and is an order of magnitude too small to
   > explain the observed period derivative from the transit times.
@@
 
-  > RESPONSE: We are not aware of any paper that suits the purpose here,
-  > so we added a citation to the URL of the Deep Space Network.
+  > RESPONSE: We are not aware of any paper that suits the purpose
+  > here, so we added a citation to the URL of the Deep Space Network.
 

### 2019-03-06 paper/ms.tex
update gitignore

@@
 computer.  The data were downlinked via the Deep Space
-Network\footnote{\url{https://deepspace.jpl.nasa.gov}}, and the
+Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
 spacecraft timestamps were calibrated against the ground-station
@@
 586\,{\rm ppm}.  The pre-launch TESS noise
-model\footnote{\url{github.com/lgbouma/tnm}}
+model\footnote{\url{github.com/lgbouma/tnm}, commit \texttt{be06f09}}
 (\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
@@
 L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
-A.~Bailey, F.~Dai, J.~Goodman, K.~Patra and V.~Van Eylen, and are
+A.~Bailey, J.~Goodman, K.~Patra and V.~Van Eylen, and are
 grateful to the people who have turned TESS from an idea into reality.
@@
 were observed over the first three TESS sectors using a combination of
-\texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}} and
+\texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}, 
+	commit \texttt{569bbc2}} and
 TEPCat \citep{southworth_homogeneous_2011}.  We recalculated the

### 2019-03-06 review/referee_response_0.txt
update and recompile referee response

@@
   > To explore the effect of possible spot-crossing events on the TESS
-  > transit time measurements, we performed a separate test.  We injected
-  > triangular spot-anomalies with amplitude $0.03\%$ and duration 30
-  > minutes at random phases into each transit.  The amplitudes were
-  > larger than the spot-crossing anomalies observed by
-  > \citet{southworth_high-precision_2009} and
+  > transit time measurements, we performed a separate test.  We
+  > injected triangular spot-anomalies with amplitude $0.03\%$ and
+  > duration 30 minutes at random phases into each transit.  The
+  > amplitudes were larger than the spot-crossing anomalies observed
+  > by \citet{southworth_high-precision_2009} and
   > \citet{sanchis-ojeda_starspots_2011}, and the durations were
-  > comparable.  Spots of these amplitudes resemble the anomalies present
-  > in transits ``1360.54'' and ``1372.58'' of Figure
+  > comparable.  Spots of these amplitudes resemble the anomalies
+  > present in transits ``1360.54'' and ``1372.58'' of Figure
   > \ref{fig:lightcurves}: they are visible, but at low confidence.
   > 
-  > With spots injected, we repeated our measurement of the transit times.
-  > On average, the measured transit times did not change with spots
-  > injected, because the flux deviations are symmetric about the transit
-  > minimum.  For individual transits, there were no cases for which the
-  > timing deviation was larger than one minute.  The most concerning
-  > shift occurs when the spot anomaly occurs during transit egress, in
-  > which case the measured mid-time is shifted early by between 30 and 50
-  > seconds \citep[qualitatively similar to results found
-  > by][]{ioannidis_how_2016}.
+  > With spots injected, we repeated our measurement of the transit
+  > times.  On average, the measured transit times did not change
+  > after injecting spots, because the flux deviations are equally
+  > likely to occur in the first and second halves of the transit.
+  > For individual transits, there were no cases for which the timing
+  > deviation was larger than one minute.  The largest shifts occur
+  > when the spot anomaly occurs during transit ingress or egress, in
+  > which case the measured mid-time is shifted either late or early
+  > by between 30 and 50 seconds \citep[qualitatively similar to
+  > results found by][]{ioannidis_how_2016}.
   > 
@@
   > early if there were spot-crossing events during every egress.  Two
-  > arguments rule out this possibility.  (1) The lightcurve residuals do
-  > not show evidence for these events.  (2) The stellar rotation period
-  > is between 20 and 40 days, and the sky-projected stellar obliquity is
-  > less than 10 degrees
+  > arguments rule out this possibility.  (1) The lightcurve residuals
+  > do not show evidence for these events.  (2) The stellar rotation
+  > period is between 20 and 40 days, and the sky-projected stellar
+  > obliquity is less than 10 degrees
   > \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
-  > Since the planet orbits every 1.3 days, requiring that spot anomalies
-  > always occur during egress would be equivalent to requiring a stellar
-  > spot distribution that is exquisitely (and thus implausibly)
-  > distributed to match the planet egress times.
+  > Since the planet orbits every 1.3 days, requiring that spot
+  > anomalies always occur during egress would be equivalent to
+  > requiring a stellar spot distribution that is exquisitely (and
+  > thus implausibly) distributed to match the planet egress times.
 
@@
 
-  > There is a final systematic concern in our global methodology.  We
-  > have opted to use mid-transit time values derived by different
-  > authors, who used heterogeneous methods to fit and detrend their
-  > lightcurves.  This of course assumes that these authors have
-  > correctly documented their time systems.  Further, though many
-  > choices in transit-fitting ({\it e.g.}, parametrization of
-  > limb-darkening and eccentricity) do not affect transit mid-time
-  > measurements, different detrending approaches can asymmetrically
-  > warp transits and shift mid-transit times.  The magnitude of this
-  > systematic effect is hard to quantify, but the situation is fairly
-  > clear from Figure~\ref{fig:times}.  Many independent authors
-  > provided transit measurements shortly after WASP-4b's discovery,
-  > so the data should on average be acceptable.
-  > \citet{huitson_gemini_2017} provided the most important data from
-  > epochs 0-1000, and so if their data were systematically affected
-  > by detrending choices at the level of many times their reported
-  > uncertainties, our timing anomaly would be suspect.  For this
-  > reason, we paid careful attention to the
+  > There is a final concern that is difficult to address.  We collected
+  > the mid-transit time values derived by different authors, who used
+  > heterogeneous methods to fit and detrend their lightcurves.  We have
+  > also assumed that these authors have correctly documented the time
+  > systems in which the data are reported.  Further, though many choices
+  > in transit-fitting ({\it e.g.}, parametrization of limb-darkening and
+  > eccentricity) do not affect transit mid-time measurements, different
+  > detrending approaches can asymmetrically warp transits and shift
+  > mid-transit times.  The magnitude of this systematic effect is hard to
+  > quantify, but the situation is fairly clear from
+  > Figure~\ref{fig:times}.  Many independent authors provided transit
+  > measurements shortly after WASP-4b's discovery, and the data are
+  > consistent with each other. \citet{huitson_gemini_2017} provided the
+  > most important data from epochs 0-1000. If their data were
+  > systematically affected by detrending choices or time-system confusion
+  > at the level of several times their reported uncertainties, then it
+  > possible that the orbital period is constant despite the evidence in
+  > the TESS data.  For this reason, we paid careful attention to the
   > \citet{huitson_gemini_2017} data set, and corresponded with the
-  > authors to confirm that their results are not affected by
-  > systematic effects of the required amplitude.
+  > authors to confirm that their results are not affected by systematic
+  > effects of the required amplitude.
   > 
-  > None of the concerns mentioned in this subsection are likely
-  > explanations for the observed timing variations. We proceed by
-  > considering possible astrophysical explanations.
+  > None of the concerns mentioned in this subsection seem likely to
+  > explain the observed timing variations. We proceed by considering
+  > possible astrophysical explanations.
 

### 2019-03-06 paper/ms.tex
arxiv submission ready

@@
   minus the calculated times assuming a constant period for transits
-  ({\it top}) and occultations ({\it bottom}). Darker points
-  correspond to more precise data.} The constant-period model (gray
-  line) is a poor description of the data.  Models with a uniformly
+  ({\it top}) and occultations ({\it bottom}). In the upper panel,
+  darker points correspond to more precise data.} The constant-period
+  model (gray line) is a poor description of the data.  Models with a
   decreasing period (blue) or an eccentric, precessing orbit (orange)
@@
 because they were calculated using a different pipeline than the
-longer-baseline \citealt{pont_determining_2011} HARPS measurements,
+longer-baseline \citeauthor{pont_determining_2011} HARPS measurements,
 and necessary inclusion of an extra offset term would nullify their
@@
 perhaps due to a companion on an eccentric orbit \citep[{\it e.g.},
-WASP-53 or WASP-81][]{triaud_peculiar_2017}.  It would be useful to
+WASP-53 or WASP-81,][]{triaud_peculiar_2017}.  It would be useful to
 gather more radial-velocity data to confirm or refute this
@@
 necessary amplitude.  Multiplying the three independent probabilities
-for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.6\
+for WASP-6b, 18b, and 46b, we can rule out $t_{\rm offset} < -81.6\
 {\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.

### 2019-03-06 paper/ms.tex
arxiv submitted

@@
 %\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
-%\pdfoutput=1 %for arXiv submission
+\pdfoutput=1 %for arXiv submission
 \usepackage{amsmath,amstext,amssymb}
@@
 A constant period can be ruled out\replaced{, but we}{. We} cannot
-distinguish between the \replaced{latter two possibilities}{
+distinguish between the\replaced{ latter two possibilities}{
   possibilities of a decaying orbit, a precessing orbit, and the

### 2019-03-27 paper/ms.tex
caught bug in k2p reported errors
reported just the 84th and 16th percentiles in the arxiv paper, instead
of 84-50, and 50-16.

(not good!)
update k2p discussion

@@
     \begin{center}
-        \includegraphics[width=0.9\textwidth]{f2.png}
+        \includegraphics[width=0.7\textwidth]{f2.png}
     \end{center}
@@
 \subsection{Analysis}
+\label{subsec:analysis}
 
@@
 sampling over the posterior probability distributions. The prior for
-the quadratic model allowed the period derivative to have any sign.}
+the quadratic model allowed the period derivative to have any sign.
+We considered two possible priors for the precession model: the first
+is a wide prior that allows non-physical values of the planetary
+Love number (Equation~\ref{eq:love_number}). The second requires
+that the planetary Love number is less than that of a sphere with a
+homogeneous density profile.  }
+
 Figure~\ref{fig:times} shows the residuals with respect to the
@@
 \replaced{decreasing-period}{quadratic} model has $\chi^2 = 62.6$ and
-60 degrees of freedom.  The best-fitting precession model has $\chi^2
-= 64.3$ and 58 degrees of freedom.  Either of the latter two models
-provides a much better fit than the constant-period model.  The
-difference in $\chi^2$ between the first and second models corresponds
-to $p \approx 10^{-26}$.
+60 degrees of freedom.  The best-fitting precession model \added{under
+the wide prior} has \replaced{$\chi^2 = 64.3$}{$\chi^2 = 64.0$} and 58
+degrees of freedom.\added{ Under the physical prior, the fit is slightly
+worse, with $\chi^2=64.8$}. The precession and quadratic models both
+provide much better fits than the constant-period model.  The
+difference in $\chi^2$ between the linear and quadratic models
+corresponds to $p \approx 10^{-26}$.
 
@@
 better fit to the data than the precession model.  It is favored by
-$\Delta \chi^2 = 1.7$, and has two fewer free parameters.  A useful
-heuristic for model comparison is the Bayesian Information Criterion
-(BIC),
+\replaced{$\Delta \chi^2 = 1.7$}{$\Delta \chi^2 = 1.4$}, and has two
+fewer free parameters.  A useful heuristic for model comparison is the
+Bayesian Information Criterion,
 \begin{equation}
@@
 the precession and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm
-prec} - {\rm BIC}_{\rm quad} = 10$, corresponding to a Bayes factor of
+prec} - {\rm BIC}_{\rm quad} = 10$\added{, assuming a wide prior for
+the precession model}. This corresponds to a Bayes factor of
 $1.1\times 10^{4}$.  Likewise, the Akaike Information Criterion favors
@@
 
-To assess the robustness of these results, we considered a few
+To assess the overall robustness of our results, we considered a few
 possible systematic effects in the timing dataset.
@@
 transit ingress or egress, in which case the measured mid-time is
-shifted either late or early by between 30 and 50 seconds
-\citep[qualitatively similar to results found
-by][]{ioannidis_how_2016}.
-
-Therefore the TESS observations could hypothetically all be skewed
-early if there were spot-crossing events during every egress.  Two
+shifted either late or early by between 30 and 50 seconds.
+
+The TESS observations could therefore be skewed
+early if there were spot-crossing events during {\it every} egress.  Two
 arguments rule out this possibility.  (1) The lightcurve residuals do
@@
 
+\explain{The result given in Equation~\ref{eq:love_number} has been
+corrected from the submitted version. The previously stated result, k2
+= $1.5^{+2.1}_{-1.1}$ had a whopping three errors. The first was a
+typo that led to printing of 84th and 16th percentiles, instead of the
+difference of these percentiles from the median. The second was an
+insufficiently precise specification of the assumed priors, which are
+now described.  The third was ignoring the uncertainty in the planet
+radius and semimajor axis, which enter at the fifth power and are now
+included.  }
+
 If instead the observed timing variation is just a small portion of an
@@
 for WASP-4b is
-\begin{equation}
-k_{2,{\rm p}} = 1.5^{+2.1}_{-1.1}.
-\end{equation}
-For comparison, the Love number of Jupiter is about 0.55
-\citep{wahl_tidal_2016,ni_empirical_2018}, and a uniform density
-sphere has $k_2 = 1.5$. The uncertainty in $k_{2,{\rm p}}$ for WASP-4b
-is large because the eccentricity, reference time, and ${\rm
-d}\omega/{\rm d}E$ have strongly correlated errors, and the measured
-occultation times only barely constrain these parameters
+\begin{align}
+  k_{2,{\rm p}} &= 1.59^{+0.70}_{-0.47}
+    \quad{\rm (wide\ prior),} \\
+  k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
+    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5])
+  \label{eq:love_number}
+\end{align}
+where the uncertainties in the semimajor axis and planet radius have been
+propagated through our Markov chains, and $\mathcal{U}$ denotes the uniform 
+distribution.  For comparison, the Love number
+of Jupiter is about 0.55 \citep{wahl_tidal_2016,ni_empirical_2018},
+and a uniform density sphere has $k_2 = 1.5$. The uncertainty in
+$k_{2,{\rm p}}$ for WASP-4b is large because the eccentricity,
+reference time, and ${\rm d}\omega/{\rm d}E$ have strongly correlated
+errors, and the \added{four available}\deleted{measured} occultation
+times only weakly constrain these parameters
 (Figure~\ref{fig:future}).
 
-The main problem with the apsidal precession hypothesis is to explain
-why the eccentricity would be as large as $\sim$$10^{-3}$ despite
-rapid tidal circularization.  For WASP-4, Equation~\ref{eq:de_dt}
-gives $\tau_e = 0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is
-several billion years old, so unless the planet arrived very recently,
-any initial eccentricity should have been lowered well below
-$\sim$$10^{-3}$.  The top panel of Figure~\ref{fig:context} compares
-the expected eccentricity damping time of WASP-4b with that of other
-transiting giant planets.  WASP-4b has one of the shortest known
-eccentricity damping times.
+\added{ The results under both priors are larger than the Love number
+of Jupiter, at weak statistical signifiance.  Imposing the
+physically-motivated prior truncates the allowed values of $k_{2,{\rm
+p}}$, and drives the precession period towards larger values.
+Different priors also change the predicted evolution of the
+orbit.  While the precession model matches the orbital decay model for
+the next two decades in transits (Figure~\ref{fig:future}), in
+occultations the precession model can deviate from the orbital decay
+model, particularly if one imposes stricter priors on the planetary
+Love number. This suggests that promptly obtaining occultation time
+measurements for the system could help rule out the apsidal precession
+model.  For the time being, the data are insufficient to make a
+stronger ruling.  }
+
+The main problem with the apsidal precession hypothesis \added{under
+either of the priors} is to explain why the eccentricity would be as
+large as $\sim$$10^{-3}$ despite rapid tidal circularization.  For
+WASP-4, Equation~\ref{eq:de_dt} gives $\tau_e = 0.29 (Q_{\rm p}'/10^5)
+\, {\rm Myr}$.  The star is several billion years old, so unless the
+planet arrived very recently, any initial eccentricity should have
+been lowered well below $\sim$$10^{-3}$.  The top panel of
+Figure~\ref{fig:context} compares the expected eccentricity damping
+time of WASP-4b with that of other transiting giant planets.  WASP-4b
+has one of the shortest known eccentricity damping times.
 
@@
 \explain{The following subsection had material moved into the ``Timing
-variation due to line of sight acceleration'' subsection.}
+variation due to line of sight acceleration'' subsection. Figure 7 was
+updated to clarify the Love number prior for the precession model.}
 \subsection{Other possible explanations}
 
-\begin{figure*}[t]
+\begin{figure*}[!ht]
 	\begin{center}
@@
 	\caption{
-    {\bf Further observations will be needed to confirm and understand
+    {\bf Further observations are needed to confirm and understand
     the timing variations of WASP-4b.} \replaced{Dots}{Symbols} are as
@@
     posteriors of the apsidal precession model (orange), and the
-    orbital decay model (blue).  \added{The two models may begin to
-    diverge in the mid-2020s}.    
+    orbital decay model (blue).  \added{The prior for the precession
+    model assumes $k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5]$; dashed orange
+    lines emphasize samples with planetary Love numbers below 0.75}.    
 		\label{fig:future}
@@
 If TESS is extended beyond its primary mission, it will likely observe
-additional transits of WASP-4b in the early 2020s
-(Figure~\ref{fig:future}).  High-precision transit observations with
+additional transits of WASP-4b in the early 2020s.
+High-precision transit observations with
 larger telescopes would also be useful.  In order to decide between
 orbital decay and apsidal precession, occultation measurements in both
-the near term and also in the mid-2020s will be needed.  More
+the near term and also in the mid-2020s will be needed 
+(Figure~\ref{fig:future}).  More
 radial-velocity data would help in the search for additional bodies
@@
 L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
-A.~Bailey, J.~Goodman, K.~Patra and V.~Van Eylen, and are
-grateful to the people who have turned TESS from an idea into reality.
-\added{L.G.B. thanks A.~Bixel and E.~May for clarifying details
-concerning the available IMACS lightcurves.}
+A.~Bailey, J.~Goodman, K.~Patra\added{, D. Ragozzine,} and V.~Van
+Eylen, and are grateful to the people who have turned TESS from an
+idea into reality.  \added{L.G.B. thanks A.~Bixel and E.~May for
+clarifying details concerning the available IMACS lightcurves.}
 %

### 2019-03-27 paper/ms.tex
minor text edits

@@
 
-\NewPageAfterKeywords
+%\NewPageAfterKeywords
 \listofchanges
@@
 \replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus
-2{,}457{,}000.  Lightcurves that were flagged \deleted{as Threshold
+2{,}457{,}000.  Lightcurves that were flagged\deleted{ as Threshold
 Crossing Events} by the SPOC pipeline \added{as crossing a transit
@@
 is a wide prior that allows non-physical values of the planetary
-Love number (Equation~\ref{eq:love_number}). The second requires
-that the planetary Love number is less than that of a sphere with a
-homogeneous density profile.  }
+Love number (Equation~\ref{eq:love_number}). The second prior requires
+that the planetary Love number is less than that of a sphere of
+constant density.  }
 
@@
 60 degrees of freedom.  The best-fitting precession model \added{under
-the wide prior} has \replaced{$\chi^2 = 64.3$}{$\chi^2 = 64.0$} and 58
-degrees of freedom.\added{ Under the physical prior, the fit is slightly
-worse, with $\chi^2=64.8$}. The precession and quadratic models both
-provide much better fits than the constant-period model.  The
-difference in $\chi^2$ between the linear and quadratic models
+the wide prior} has $\chi^2 = 64.3$ and 58
+degrees of freedom.\added{ Under the physical prior, the best-fit is
+slightly worse, with $\chi^2=64.8$}. The precession and quadratic
+models both provide much better fits than the constant-period model.
+The difference in $\chi^2$ between the linear and quadratic models
 corresponds to $p \approx 10^{-26}$.
 
-The \replaced{decreasing-period}{quadratic} model provides a slightly
-better fit to the data than the precession model.  It is favored by
-\replaced{$\Delta \chi^2 = 1.7$}{$\Delta \chi^2 = 1.4$}, and has two
-fewer free parameters.  A useful heuristic for model comparison is the
-Bayesian Information Criterion,
+The \replaced{decreasing-period}{quadratic} model provides a
+marginally better fit to the data than the precession model.  It is
+favored by $\Delta \chi^2 = 1.7$, and has two fewer free parameters.
+A useful heuristic for model comparison is the Bayesian Information
+Criterion,
 \begin{equation}
@@
 corrected from the submitted version. The previously stated result, k2
-= $1.5^{+2.1}_{-1.1}$ had a whopping three errors. The first was a
+= $1.5^{+2.1}_{-1.1}$ had a
 typo that led to printing of 84th and 16th percentiles, instead of the
-difference of these percentiles from the median. The second was an
-insufficiently precise specification of the assumed priors, which are
-now described.  The third was ignoring the uncertainty in the planet
-radius and semimajor axis, which enter at the fifth power and are now
-included.  }
+difference of these percentiles from the median. 
+The uncertainties also previously did not propagate uncertainty in the
+planet radius and semimajor axis, which enter at the fifth power and
+are now included. 
+Finally, we now also specify the assumed priors, as they add a
+nuance in the interpretation.   }
 
@@
   k_{2,{\rm p}} &= 1.59^{+0.70}_{-0.47}
-    \quad{\rm (wide\ prior),} \\
+    \quad{\rm (wide\ prior),} \label{eq:love_number} \\
   k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
     \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5])
-  \label{eq:love_number}
 \end{align}
@@
 p}}$, and drives the precession period towards larger values.
-Different priors also change the predicted evolution of the
+Different priors therefore change the predicted evolution of the
 orbit.  While the precession model matches the orbital decay model for
@@
 occultations the precession model can deviate from the orbital decay
-model, particularly if one imposes stricter priors on the planetary
-Love number. This suggests that promptly obtaining occultation time
+model, particularly if one requires the planetary Love number to be
+smaller. This suggests that promptly obtaining occultation time
 measurements for the system could help rule out the apsidal precession
 model.  For the time being, the data are insufficient to make a
-stronger ruling.  }
-
-The main problem with the apsidal precession hypothesis \added{under
-either of the priors} is to explain why the eccentricity would be as
-large as $\sim$$10^{-3}$ despite rapid tidal circularization.  For
-WASP-4, Equation~\ref{eq:de_dt} gives $\tau_e = 0.29 (Q_{\rm p}'/10^5)
-\, {\rm Myr}$.  The star is several billion years old, so unless the
-planet arrived very recently, any initial eccentricity should have
-been lowered well below $\sim$$10^{-3}$.  The top panel of
-Figure~\ref{fig:context} compares the expected eccentricity damping
-time of WASP-4b with that of other transiting giant planets.  WASP-4b
-has one of the shortest known eccentricity damping times.
+stronger judgement.  }
+
+\added{Regardless of the detailed model assumptions, the}
+\deleted{The} main \added{physical }problem with the apsidal precession
+hypothesis is to explain why the
+eccentricity would be as large as $\sim$$10^{-3}$ despite rapid tidal
+circularization.  For WASP-4, Equation~\ref{eq:de_dt} gives $\tau_e =
+0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is several billion
+years old, so unless the planet arrived very recently, any initial
+eccentricity should have been lowered well below $\sim$$10^{-3}$.  The
+top panel of Figure~\ref{fig:context} compares the expected
+eccentricity damping time of WASP-4b with that of other transiting
+giant planets.  WASP-4b has one of the shortest known eccentricity
+damping times.
 

### 2019-03-27 review/referee_response_0.txt
minor updates to review
ugh

@@
 
-  > RESPONSE: We thank the referee for this summary. In our revision,
-  > we have added weight to a third possible explanation that we
-  > previously mentioned but did not sufficiently emphasize: a massive
-  > outer companion, perhaps on an eccentric orbit, might be able to
-  > explain both the transit and radial velocity observations. We have
-  > described this possibility throughout the text where appropriate.
-  > We also added our own re-analysis of the extant RV data (now
+  > RESPONSE: We thank the referee for this summary. The major and minor
+  > points are addressed individually below. There are also three larger
+  > changes that we highlight here.
+  >
+  > 1) In our revision, we have added weight to a third possible
+  > explanation that we previously mentioned but did not sufficiently
+  > emphasize: a massive outer companion, perhaps on an eccentric orbit,
+  > might be able to explain both the transit and radial velocity
+  > observations. We have described this possibility throughout the text
+  > where appropriate.
+  >
+  > 2) We also added our own re-analysis of the extant RV data (now
   > section 4.3, and figure 6) to check on the Knutson et al (2014)
@@
   > explain the observed period derivative from the transit times.
+  >
+  > 3) Finally, we have also added a more detailed exploration of how
+  > the assumed prior in the precession model affects the inferred Love
+  > number, and the predicted transit and occultation times.
 
@@
   > duration 30 minutes at random phases into each transit.  The
-  > amplitudes were larger than the spot-crossing anomalies observed
-  > by \citet{southworth_high-precision_2009} and
+  > amplitudes were larger than the spot-crossing anomalies observed by
+  > \citet{southworth_high-precision_2009} and
   > \citet{sanchis-ojeda_starspots_2011}, and the durations were
@@
   > With spots injected, we repeated our measurement of the transit
-  > times.  On average, the measured transit times did not change
-  > after injecting spots, because the flux deviations are equally
-  > likely to occur in the first and second halves of the transit.
-  > For individual transits, there were no cases for which the timing
-  > deviation was larger than one minute.  The largest shifts occur
-  > when the spot anomaly occurs during transit ingress or egress, in
-  > which case the measured mid-time is shifted either late or early
-  > by between 30 and 50 seconds \citep[qualitatively similar to
-  > results found by][]{ioannidis_how_2016}.
+  > times.  On average, the measured transit times did not change after
+  > injecting spots, because the flux deviations are equally likely to
+  > occur in the first and second halves of the transit.  For individual
+  > transits, there were no cases for which the timing deviation was
+  > larger than one minute.  The largest shifts occur when the spot
+  > anomaly occurs during transit ingress or egress, in which case the
+  > measured mid-time is shifted either late or early by between 30 and
+  > 50 seconds \citep[qualitatively similar to results found
+  > by][]{ioannidis_how_2016}.
   > 
@@
   > anomalies always occur during egress would be equivalent to
-  > requiring a stellar spot distribution that is exquisitely (and
-  > thus implausibly) distributed to match the planet egress times.
+  > requiring a stellar spot distribution that is exquisitely (and thus
+  > implausibly) distributed to match the planet egress times.
 
@@
   > RESPONSE: We have clarified that a "Presearch Data Conditioning"
-  > lightcurve is at a stage of process in which systematic trends
-  > have been removed through the decorrelation process described by
-  > Smith et al (2017).
+  > lightcurve is at a stage of process in which systematic trends have
+  > been removed through the decorrelation process described by Smith et
+  > al (2017).
   >
-  > We now omit explicit mention of "Treshold Crossing Events" in
-  > favor of a more qualitative description.
+  > We now omit explicit mention of "Treshold Crossing Events" in favor
+  > of a more qualitative description.
   >
@@
   > the reader to the data release notes for an explanation of what a
-  > "momentum dump" is, and have added text to this footnote to
-  > improve its clarity.
+  > "momentum dump" is, and have added text to this footnote to improve
+  > its clarity.
 
@@
 
-  > RESPONSE: We have changed the wording from "ramp-like
-  > systematic effects" to "correlated red noise". No reference is
-  > needed.
+  > RESPONSE: We have changed the wording from "ramp-like systematic
+  > effects" to "correlated red noise". No reference is needed.
 
@@
 
-  > RESPONSE: Yes, this was through weighted least-squares. We have changed the
-  > wording accordingly.
+  > RESPONSE: Yes, this was through weighted least-squares. We have
+  > changed the wording accordingly.
 
@@
   > RESPONSE: We have added the following sentences in Section 3.2:
-  > "We fit each model by assuming a Gaussian likelihood and sampling
+  > """
+  > We fitted each model by assuming a Gaussian likelihood and sampling
   > over the posterior probability distributions. The prior for the
-  > quadratic model allowed the period derivative to have any sign."
-  > We have clarified the wording throughout from "decreasing-period
-  > model" to "quadratic model" to reflect the latter point.
+  > quadratic model allowed the period derivative to have any sign.  We
+  > considered two possible priors for the precession model: the first
+  > is a wide prior that allows non-physical values of the planetary
+  > Love number (Equation~\ref{eq:love_number}). The second requires
+  > that the planetary Love number is less than that of a sphere with a
+  > homogeneous density profile.  
+  > """
 
@@
 
-  > RESPONSE: As noted in the caption of Figure 5, the decay timescale values
-  > are obtained from Equation 15, assuming Qstar'=10^7.  We have slightly
-  > modified the text for clarity on this point.
+  > RESPONSE: As noted in the caption of Figure 5, the decay timescale
+  > values are obtained from Equation 15, assuming Qstar'=10^7.  We have
+  > slightly modified the text for clarity on this point.
 
@@
 
-  > RESPONSE: To make Figure 4 and 6, we start with the observed
+  > RESPONSE: To make Figure 4 and 6 (now 7), we start with the observed
   > transit and occultation mid-times, as a function of epoch number.
   > We then fit a line to these data.  If the data were generated by a
-  > constant period orbit (with gaussian uncertainties) the residuals
-  > to such a fit would show no correlation with epoch number.
-  > However if the data were generated by a planet with a constant
-  > period derivative (with gaussian uncertainties), the residuals to
-  > such a fit would always be a parabola.  This is because the linear
-  > fit to the data will have contributions from the beginning,
-  > middle, and end of the transit timeseries. Thus the best-fit
-  > linear ephemeris (to data that actually have a decreasing period)
-  > has a best-fit period somewhere between the minimum and maximum
-  > period. At the earliest epochs, the best-fit period is an
-  > underestimate, so you see the increase in O-C. At the latest
-  > epochs, the best-fit period is an overestimate, so you see the
-  > decrease.
+  > constant period orbit (with gaussian uncertainties) the residuals to
+  > such a fit would show no correlation with epoch number.  However if
+  > the data were generated by a planet with a constant period
+  > derivative (with gaussian uncertainties), the residuals to such a
+  > fit would always be a parabola.  This is because the linear fit to
+  > the data will have contributions from the beginning, middle, and end
+  > of the transit timeseries. Thus the best-fit linear ephemeris (to
+  > data that actually have a decreasing period) has a best-fit period
+  > somewhere between the minimum and maximum period. At the earliest
+  > epochs, the best-fit period is an underestimate, so you see the
+  > increase in O-C. At the latest epochs, the best-fit period is an
+  > overestimate, so you see the decrease.
 
@@
 ********************************************************************************
-OTHER CHANGES:
+OTHER MINOR CHANGES:
 
@@
 * Minor tweaks to figures.
+
+* Added acknowledgement to D. Ragozzine.
+
+* Updated Figure 7 to use physical prior, and to show subset of samples
+  with k2 < 0.75.

### 2019-04-11 paper/ms.tex
aj resubmit
aj accepted

@@
 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
-wavelength range 0.4--22~$\mu$m.  We adopted the effective temperature
+wavelength range 0.4--22~$\mu$m.
+We adopted the effective temperature
 from the work by \citet{doyle_accurate_2013}, who determined the
@@
   decreasing period (blue) or an eccentric, precessing orbit (orange)
-  provide	better fits\deleted{ to the data}.  The red square
-  represents the combination of all the TESS data and is for display
-  purposes only.  The models were fitted to all of the individual
-  transit times.
+  provide	better fits\deleted{ to the data}.  The binned TESS
+  point (red square) is the weighted average of 18 TESS transits
+  and is for display purposes only.
+  The models were fitted to all of the individual transit times.
   \label{fig:times}
@@
   k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
-    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5])
+    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5]),
 \end{align}
@@
 	TESS \citep{ricker_transiting_2015},
-	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
+	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018},
+    Tycho-2 \citep{hog_tycho-2_2000},
+    2MASS \citep{skrutskie_two_2006},
+    APASS \citep{henden_data_2012},
+    WISE \citep{wright_wide-field_2010}.
 }

### 2019-09-02 rnaas_note/ms.tex
[WIP] add RNAAS note sketch

% TODO:
%
% * email to JNW
% * email to full collaborator list
% * add any missing CPS coauthors
% * check if it was really only 5 CPS points in knutson's paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[RNAAS]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}

\begin{document}

\title{WASP-4 is accelerating towards the Earth}

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0001-8638-0320]{A. W. Howard}
\affiliation{Cahill Center for Astrophysics, California Institute of
Technology, Pasadena, CA 91125, USA}
%
\author{H. Knutson}
\affiliation{Division of Geological and Planetary Sciences, California
Institute of Technology, Pasadena, CA 91125, USA}
%
\author[0000-0002-0531-1073]{H. Isaacson}
\affiliation{Astronomy Department, University of California, Berkeley,
CA 94720, USA}
%

%% Note that RNAAS manuscripts DO NOT have abstracts.
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
(1332), Transit timing variation method (1710)}

%% Start the main body of the article. If no sections in the 
%% research note leave the \section call blank to make the title.

\section{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In recent work we showed that the TESS mission observed the hot
Jupiter WASP-4b transit $\approx 82$ seconds earlier than expected,
under the assumption of a constant orbital period
\citep{bouma_wasp-4b_2019}.  This result was based on transit
times\footnote{ Our timing analysis included data from peer-reviewed
literature for which the times were measured from a single transit,
and for which the midpoint was allowed to be a free parameter. We also
required that the time system be clearly documented. In other words,
we needed to know
whether any heliocentric/barycentric corrections had been performed,
and whether the absolute time system was UTC or TDB.} acquired since
2007 (CITE ALL).  We pointed out that the transit times were best fit
by a timing model with a constant period derivative, and reported a
best-fit decay rate of $\dot{P} = -12.6 \pm 1.2\,{\rm ms\ per\ year}$.
Our interpretation was that the apparent period change could be caused
by any of three scenarios: a decaying orbit, a precessing orbit, and
an orbit being gravitationally perturbed by an outer companion.

Thereafter, \citet{southworth_transit_2019} reported 22 new transit
times for the system, and also found that the entire series of transit
times was consistent with a quadratic ephemeris.  Their interpretation
of the timing variations did not differ in any major aspects from our own.  A
separate follow-up study by \citet{baluev_homogeneously_2019} also reported
new light curves, and analyzed these plus light curves that were
purposefully omitted from our analysis due to systematic uncertainties
in the absolute time system.  \citeauthor{baluev_homogeneously_2019}
found that when they used all the available TTV data, the need for a
quadratic ephemeris was present ``at the high $\sim 5-7$ sigma
level''.  However, they pointed out that if they used lower-precision
subsets of the available timing data, the necessity for the quadratic
term decreased.  \citeauthor{baluev_homogeneously_2019} also pointed
out that the precise transit times reported by
\citet{huitson_gemini_2017} were quite important in the time-series.
Overall, \citeauthor{baluev_homogeneously_2019} were not convinced.

We, too, wanted to confirm and understand the timing variation.  One
of the lines of follow-up we originally mentioned
was the need for additional radial velocity
observations.  We acquired them this past season with Keck HIRES.
After fitting out the hot Jupiter, the residuals now
show a strong linear trend  (Figure~\ref{fig:rv_o_minus_c}).
Specifically, the best-fit radial velocity derivative, $\dot{v}_{\rm
r}$, is
\begin{equation}
  \dot{v}_{\rm r} = - X.XX \pm Y.YY\,{\rm m\,s^{-1}\,day^{-1}}.
\end{equation}
This acceleration towards our line of sight is XXX times faster than 
reported from the 5 points available during the Friends of Hot
Jupiters project \citep{knutson_friends_2014}.
The updated radial velocity values are available in the
data-behind-the-figure online version of this {\it Note}.

Under the assumption of constant acceleration, the implied period
derivative is $\dot{P} = \dot{v}_{\rm r} P / c$, yielding X.XX
milliseconds per year for WASP-4.

With respect to WASP-4b's quadratic ephemeris, it seems rather
likely that this line of sight acceleration is at fault.
Further radial velocity observations of the system should help in
resolving the nature of the companion.

%FIXME FIXME

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RNAAS: you get one figure.

\begin{figure}
    \begin{center}
		\includegraphics{20190716_rv_fit.png}
    \end{center}
    \vspace{-0.8cm}
    \caption{
      {\bf Radial velocity observations of WASP-4.}
      The orbit of WASP-4b has been subtracted.
      %FIXME: finish writing.
       \label{fig:rv_o_minus_c}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgements

%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\software{
  \texttt{radvel} \citep{fulton_radvel_2018},
}

\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\end{document}

### 2019-09-10 rnaas_note/ms.tex
rnaas note update

@@
 %
+% * update the RV numbers
+% * redo the RV fit with the latest CKS points
 % * email to JNW
@@
 
-In recent work we showed that the TESS mission observed the hot
+In recent work we showed that the TESS spacecraft observed the hot
 Jupiter WASP-4b transit $\approx 82$ seconds earlier than expected,
@@
 times was consistent with a quadratic ephemeris.  Their interpretation
-of the timing variations did not differ in any major aspects from our own.  A
-separate follow-up study by \citet{baluev_homogeneously_2019} also reported
-new light curves, and analyzed these plus light curves that were
-purposefully omitted from our analysis due to systematic uncertainties
-in the absolute time system.  \citeauthor{baluev_homogeneously_2019}
+of the timing variations did not differ in any major respects from our
+own, though with their additional data a lower best-fit decay rate was
+reported ($\dot{P} = -XX.X \pm X.X\,{\rm ms\ per\ year}$).
+
+A separate study by \citet{baluev_homogeneously_2019} also reported
+new light curves. \citeauthor{baluev_homogeneously_2019} analyzed
+their newly obtained light curves, along with archival light curves
+that were omitted from our analysis due to systematic uncertainties in
+the absolute time system.  \citeauthor{baluev_homogeneously_2019}
 found that when they used all the available TTV data, the need for a
@@
 We, too, wanted to confirm and understand the timing variation.  One
-of the lines of follow-up we originally mentioned
-was the need for additional radial velocity
-observations.  We acquired them this past season with Keck HIRES.
-After fitting out the hot Jupiter, the residuals now
-show a strong linear trend  (Figure~\ref{fig:rv_o_minus_c}).
+line of follow-up we mentioned was the need for additional radial
+velocity observations.  We acquired them this past season with Keck
+HIRES.
+After fitting out the hot Jupiter, the residuals now show a strong
+linear trend  (Figure~\ref{fig:rv_o_minus_c}).
 Specifically, the best-fit radial velocity derivative, $\dot{v}_{\rm
@@
 Jupiters project \citep{knutson_friends_2014}.
+%FoHJ: gammadot = -0.0099^{+0.0052}_{-0.0054}
 The updated radial velocity values are available in the
@@
 
-Under the assumption of constant acceleration, the implied period
-derivative is $\dot{P} = \dot{v}_{\rm r} P / c$, yielding X.XX
-milliseconds per year for WASP-4.
+Under the assumption of constant acceleration, $\dot{P} = \dot{v}_{\rm
+r} P / c$, so the implied period decrease that should be seen in
+transits is $\approx$X.X milliseconds per year.
+All the available TTV data show a period decrease of $\approx
+8-12\,{\rm ms\ per\ year}$
+\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.
 
-With respect to WASP-4b's quadratic ephemeris, it seems rather
-likely that this line of sight acceleration is at fault.
-Further radial velocity observations of the system should help in
-resolving the nature of the companion.
-
-%FIXME FIXME
+It seems quite likely that the line of sight acceleration is to blame
+for the apparent decrease of WASP-4b's orbital period.  While further
+radial velocity observations of the system should help clarify the
+mass and semi-major axis of the companion, it seems unlikely that
+further transit observations will yield near-term constraints on
+orbital decay.
 

### 2019-09-10 src/compare_pdots.py

@@ -1,0 +1,17 @@
+import numpy as np
+from astropy import units as u, constants as c
+
+period = 1.3382324*u.day
+
+# Baluev Figure 4.  His T_d = -P/Pdot.
+T_d_inv_min = 0.075*(1/u.Myr)
+T_d_inv_max = 0.100*(1/u.Myr)
+
+T_d_max = 1/T_d_inv_min
+T_d_min = 1/T_d_inv_max
+
+Pdot_max = (-period/T_d_min).to(u.millisecond/u.yr)
+Pdot_min = (-period/T_d_max).to(u.millisecond/u.yr)
+
+print(Pdot_max)
+print(Pdot_min)

### 2019-09-11 src/WASP4.py
updates and new code for RNAAS RV analysis

@@ -57,7 +57,7 @@ params['dvdt'].vary = True
 # the velocities are expected to be in m/s
 datadir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/data"
 #datapath = os.path.join(datadir,'RVs_all_WASP4b_for_fitting_USEDINPAPER.csv')
-datapath = os.path.join(datadir,'RVs_all_minustwo_WASP4b_for_fitting_20190716.csv')
+datapath = os.path.join(datadir,'RVs_all_WASP4b_for_fitting_20190911.csv')
 data = pd.read_csv(datapath, sep=',')
 
 # Define prior shapes and widths here.

### 2019-09-11 src/is_it_doppler_shift.py

@@ -42,7 +42,7 @@ dP_dt_minus1sigma = (
 )
 dP_dt_limit = gamma_dot_limit * P / c.c
 dP_dt_measured = -4.00e-10
-dP_dt_measured_err = 3.8e-10
+dP_dt_measured_err = 3.8e-11
 
 print('\nfor WASP-4b')
 print('Knutson+ 2014 reported {:d} CPS data points, and used {:d} data sets'.
@@ -136,6 +136,43 @@ print('\nfor WASP-4b')
 print('Fei Dai\'s analysis gives...')
 print('gamma_dot_value: {:.6g} +/-({:.6g}) --> gamma_dot_limit: {:.6g}'.
       format(gamma_dot_value, gamma_dot_sig, gamma_dot_limit))
+print('2σ upper limit on dP/dt change from Doppler shift: {:.6g}'.format(dP_dt_limit.cgs))
+print('dP/dt measured: {:.6g}, +/-({:.6g})'.
+      format(dP_dt_measured, dP_dt_measured_err))
+print('(dP/dt_Dopplerlimit)/(dP/dt_measured): {:.6g}'.format(dP_dt_limit.cgs.value/dP_dt_measured))
+print('(dP/dt_Dopplerquoted)/(dP/dt_measured): {:.6g}'.format(dP_dt_preferred.cgs.value/dP_dt_measured))
+
+
+# WASP-4b. 2019/09/11 LGB analysis, for RNAAS
+P = 1.3382324*u.day
+gamma_dot_value, gamma_dot_sig = -0.0422, 0.0028
+
+gamma_dot_preferred = gamma_dot_value * (u.m/u.s)/u.day
+gamma_dot_limit = (gamma_dot_value - 2*gamma_dot_sig) * (u.m/u.s)/u.day
+
+gamma_dot_upper = (gamma_dot_value + gamma_dot_sig) * (u.m/u.s)/u.day
+gamma_dot_lower = (gamma_dot_value - gamma_dot_sig) * (u.m/u.s)/u.day
+
+dP_dt_preferred = gamma_dot_preferred * P / c.c
+dP_dt_upper = gamma_dot_upper * P / c.c
+dP_dt_lower = gamma_dot_lower * P / c.c
+dP_dt_limit = gamma_dot_limit * P / c.c
+
+print('\nfor WASP-4b')
+print('LGB\'s RNAAS analysis gives...')
[…]

### 2019-09-11 src/plot_rv_vs_CaHK_check.py

@@ -1,0 +1,37 @@
+"""
+do we need to worry about stellar activity?
+"""
+import pandas as pd, numpy as np, matplotlib.pyplot as plt
+
+df = pd.read_csv('../data/20190911_jump_wasp4_rv.csv', comment='#')
+
+f,axs = plt.subplots(nrows=2,ncols=1,figsize=(4,6), sharex=True)
+
+axs[0].scatter(df['bjd'], df['mnvel'])
+axs[0].set_ylabel('mnvel')
+
+axs[1].scatter(df['bjd'], df['svalue'])
+axs[1].set_ylabel('svalue')
+axs[1].set_xlabel('bjd tdb')
+
+f.tight_layout()
+outpath = '../results/20190911_rv_vs_CaHK_check_timeseries.png'
+f.savefig(outpath, dpi=300, bbox_inches='tight')
+print('made {}'.format(outpath))
+
+##########################################
+
+plt.close('all')
+
+f,ax = plt.subplots(nrows=1,ncols=1,figsize=(4,3))
+
+ax.scatter(df['mnvel'], df['svalue'])
+ax.set_ylabel('svalue')
+ax.set_xlabel('mnvel')
+
+f.tight_layout()
+outpath = '../results/20190911_rv_vs_CaHK_check_mnvel_vs_svalue.png'
+f.savefig(outpath, dpi=300, bbox_inches='tight')
+print('made {}'.format(outpath))
+
+

### 2019-09-11 src/plot_rvs_RNAAS.py

@@ -24,7 +24,7 @@ from radvel.driver import save_status, load_status
 import configparser
 
 ##########################################
-# a bunch of stuff needed to hack at radvel objects
+# stuff needed to hack at radvel objects
 
 class args_object(object):
     """
@@ -40,95 +40,16 @@ class args_object(object):
         self.plotkw = {}
         self.gp = False
 
-def _hack_radvel_plots(args):
-    """
-    Hacked version of radvel.driver.plots
-
-    Args:
-        args (ArgumentParser): command line arguments
-    """
-
-    config_file = args.setupfn
-    conf_base = os.path.basename(config_file).split('.')[0]
-    statfile = os.path.join(
-        args.inputdir, "{}_radvel.stat".format(conf_base)
-    )
-
-    status = load_status(statfile)
-
-    assert status.getboolean('fit', 'run'), \
-        "Must perform max-liklihood fit before plotting"
-    post = radvel.posterior.load(status.get('fit', 'postfile'))
-
-    # update the plotted posterior to match the median best-fit parameters.
-    summarycsv = "../results/rv_fitting/t10onlytwo_LGB_20190228_fix_gammaddot/WASP4_post_summary.csv"
-    sdf = pd.read_csv(summarycsv)
-    for param in [c for c in sdf.columns if 'Unnamed' not in c]:
-        post.params[param] = radvel.Parameter(value=sdf.ix[1][param])
-
-    for ptype in args.type:
-        print("Creating {} plot for {}".format(ptype, conf_base))
-
-        if ptype == 'rv':
-            args.plotkw['uparams'] = post.uparams
-            saveto = os.path.join(
-                args.outputdir,conf_base+'_rv_multipanel.pdf'
-            )
-            P, _ = radvel.utils.initialize_posterior(config_file)
-            if hasattr(P, 'bjd0'):
-                args.plotkw['epoch'] = P.bjd0
-
-            #FIXME FIXME: see line 103 of this for how to access residuals
-            model = post.likelihood.model
-            rvtimes = post.likelihood.x
-            rverr = post.likelihood.errorbars()
-            num_planets = model.num_planets
-
-            #FIXME FIXME
-            rawresid = post.likelihood.residuals()
-
-            resid = (
-                rawresid + post.params['dvdt'].value*(rvtimes-model.time_base)
-                + post.params['curv'].value*(rvtimes-model.time_base)**2
[…]

### 2019-09-11 src/run_rv_fitting.py

@@ -23,7 +23,8 @@ setupfn = "/home/luke/Dropbox/proj/WASP-4b_anomaly/src/WASP4.py"
 #outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/t10onlytwo_LGB_20190228_fix_gammaddot"
 #outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/LGB_20190716_fix_gammaddot"
 #outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/LGB_20190716_fix_gammaddot_PAPERDATA"
-outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/LGB_20190716_fix_gammaddot_minustwo"
+#outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/LGB_20190716_fix_gammaddot_minustwo"
+outputdir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/results/rv_fitting/LGB_20190911_fix_gammaddot"
 if not os.path.exists(outputdir):
     os.mkdir(outputdir)
 

### 2019-09-11 rnaas_note/ms.tex
update RNAAS note

--- 2019-09-02 rnaas_note/ms.tex
@@
 %
+% * pdf figures...
 % * email to JNW
@@
 
-In recent work we showed that the TESS mission observed the hot
-Jupiter WASP-4b transit $\approx 82$ seconds earlier than expected,
-under the assumption of a constant orbital period
-\citep{bouma_wasp-4b_2019}.  This result was based on transit
-times\footnote{ Our timing analysis included data from peer-reviewed
-literature for which the times were measured from a single transit,
-and for which the midpoint was allowed to be a free parameter. We also
-required that the time system be clearly documented. In other words,
-we needed to know
-whether any heliocentric/barycentric corrections had been performed,
-and whether the absolute time system was UTC or TDB.} acquired since
-2007 (CITE ALL).  We pointed out that the transit times were best fit
-by a timing model with a constant period derivative, and reported a
-best-fit decay rate of $\dot{P} = -12.6 \pm 1.2\,{\rm ms\ per\ year}$.
-Our interpretation was that the apparent period change could be caused
-by any of three scenarios: a decaying orbit, a precessing orbit, and
-an orbit being gravitationally perturbed by an outer companion.
+We recently showed that under the assumption of a constant orbital
+period, the hot Jupiter WASP-4b transited
+$\approx$82 seconds early for the TESS spacecraft 
+\citep{bouma_wasp-4b_2019}.  Our timing
+analysis included data from peer-reviewed literature for which the
+times were measured from a single transit, and for which the midpoint
+was allowed to be a free parameter. We also required the time system
+to be clearly documented. In other words, we needed to know whether
+any heliocentric or barycentric corrections had been performed, and
+whether the absolute time system was UTC or TDB. The transit times
+we analyzed spanned 2007 to 2019 
+\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,dragomir_terms_2011,sanchis-ojeda_starspots_2011,nikolov_wasp-4b_2012,hoyer_tramos_2013,ranjan_atmospheric_2014,huitson_gemini_2017}.
+The combined timing data were best fit by
+an ephemeris with a constant negative period derivative; our best-fit
+decay rate was $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  Our
+interpretation was that the apparent period change could be caused by
+any of three scenarios: a decaying orbit, a precessing orbit, and an
+orbit being gravitationally perturbed by an outer companion.
 
 Thereafter, \citet{southworth_transit_2019} reported 22 new transit
-times for the system, and also found that the entire series of transit
+times for the system, and confirmed that the updated series of transit
 times was consistent with a quadratic ephemeris.  Their interpretation
-of the timing variations did not differ in any major aspects from our own.  A
-separate follow-up study by \citet{baluev_homogeneously_2019} also reported
-new light curves, and analyzed these plus light curves that were
-purposefully omitted from our analysis due to systematic uncertainties
-in the absolute time system.  \citeauthor{baluev_homogeneously_2019}
-found that when they used all the available TTV data, the need for a
-quadratic ephemeris was present ``at the high $\sim 5-7$ sigma
-level''.  However, they pointed out that if they used lower-precision
-subsets of the available timing data, the necessity for the quadratic
-term decreased.  \citeauthor{baluev_homogeneously_2019} also pointed
-out that the precise transit times reported by
-\citet{huitson_gemini_2017} were quite important in the time-series.
-Overall, \citeauthor{baluev_homogeneously_2019} were not convinced.
+of the timing variations did not differ in any major respects from our
+own, though with additional data they found a lower best-fit decay
+rate of $\dot{P} = -XX.X \pm X.X$ milliseconds per year.
 
-We, too, wanted to confirm and understand the timing variation.  One
-of the lines of follow-up we originally mentioned
-was the need for additional radial velocity
-observations.  We acquired them this past season with Keck HIRES.
-After fitting out the hot Jupiter, the residuals now
-show a strong linear trend  (Figure~\ref{fig:rv_o_minus_c}).
-Specifically, the best-fit radial velocity derivative, $\dot{v}_{\rm
-r}$, is
+A separate study by \citet{baluev_homogeneously_2019} reported
+additional light curves. \citeauthor{baluev_homogeneously_2019}
+analyzed their newly obtained photometry, along with archival light
+curves that we omitted from our analysis due to systematic
+uncertainties in the absolute time system.
+\citeauthor{baluev_homogeneously_2019} found that when they used all
+the available TTV data, the need for a quadratic ephemeris was present
+``at the high $\sim 5-7$ sigma level''.  However, they pointed out
+that if they used lower-precision subsets of the available timing
+data, the necessity for the quadratic term decreased.
+\citeauthor{baluev_homogeneously_2019} also pointed out that the
+precise transit times reported by \citet{huitson_gemini_2017} were
+quite important in the time-series.  Overall,
+\citeauthor{baluev_homogeneously_2019} did not find the claim of a
+period decrease convincing.
+
+One line of follow-up needed to confirm and understand the timing
+variation was additional radial velocity observations.  The longest RV
+baseline previously available was six observations acquired by
+\citet{knutson_friends_2014} and the CKS team on Keck HIRES from
+2010 to 2013.  This past season we acquired four additional
+observations with Keck HIRES. Even before fitting out the hot Jupiter,
+the residuals showed a strong linear trend.
+
+For completeness, in our analysis we included CORALIE measurements
+from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
+using the homogeneous velocities reported by the latter authors.  We
+included the HARPS values reported by \citet{pont_determining_2011},
+which are identical to those from \citet{husnoo_observational_2012}.
+We omitted the HARPS data points taken over three nights by
+\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
+because they were calculated using a different pipeline than the
+longer-baseline \citeauthor{pont_determining_2011} measurements, and
+necessarily inclusion of an extra offset term would nullify their
+statistical value.
+
+We then fitted a single Keplerian orbit, plus instrument offsets,
+jitters, and a linear trend
+\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
+on the period and time of inferior conjunction using the values from
+\citet{bouma_wasp-4b_2019}, and fixed the eccentricity to zero,
+consistent with results from \citet{beerer_secondary_2011},
+\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The
+remaining free parameters were the velocity semi-amplitude, the
+instrument zero-points, the instrument jitters (an additive white
+noise term for each instrument), and optionally a linear acceleration
+term ($\dot{v_{\rm r}}$).
+
+%FoHJ: gammadot = -0.0099^{+0.0052}_{-0.0054}
+The AIC and BIC strongly favored a model with a linear trend ($\Delta
+{\rm BIC}=73$ compared to a model without the linear trend).
+Figure~\ref{fig:rv_o_minus_c} shows the best-fitting model, with the
+orbit of WASP-4b subtracted.  The best-fit radial velocity derivative,
+$\dot{v}_{\rm r}$, is
 \begin{equation}
-  \dot{v}_{\rm r} = - X.XX \pm Y.YY\,{\rm m\,s^{-1}\,day^{-1}}.
+  \dot{v}_{\rm r} = -0.0422^{+0.0028}_{-0.0027}\,{\rm m\,s^{-1}\,day^{-1}}.
 \end{equation}
-This acceleration towards our line of sight is XXX times faster than 
-reported from the 5 points available during the Friends of Hot
-Jupiters project \citep{knutson_friends_2014}.
-The updated radial velocity values are available in the
-data-behind-the-figure online version of this {\it Note}.
+This acceleration towards our line of sight is 4.3 times faster than
+the suggested $\approx 2\sigma$ trend \citet{knutson_friends_2014}
+found from the shorter baseline.  The radial velocity values are
+available in the data-behind-the-figure online version of this {\it
+Note}.
 
-Under the assumption of constant acceleration, the implied period
-derivative is $\dot{P} = \dot{v}_{\rm r} P / c$, yielding X.XX
-milliseconds per year for WASP-4.
+Under the assumption of constant acceleration, $\dot{P} = \dot{v}_{\rm
+r} P / c$, the implied period decrease that should be seen in
+transits is roughly $5.9 \pm 0.4$ milliseconds per year.  All the
+available TTV data show a period decrease of $\approx 8-12$
+milliseconds per year~\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.
 
-With respect to WASP-4b's quadratic ephemeris, it seems rather
-likely that this line of sight acceleration is at fault.
-Further radial velocity observations of the system should help in
-resolving the nature of the companion.
-
-%FIXME FIXME
+While the quantitative agreement is not perfect, Occam's razor would
+suggest that the line of sight acceleration is probably a sufficient
+explanation for the apparent decrease of WASP-4b's orbital period.
+While further radial velocity observations of the system should help
+clarify the mass and semi-major axis of the companion, it seems
+unlikely that additional transit observations will yield near-term
+constraints on orbital decay.
 
@@
     \begin{center}
-		\includegraphics{20190716_rv_fit.png}
+		\includegraphics{rv_fit.pdf}
     \end{center}
@@
       {\bf Radial velocity observations of WASP-4.}
-      The orbit of WASP-4b has been subtracted.
-      %FIXME: finish writing.
+      The orbit of WASP-4b has been subtracted.  The best-fit linear
+      trend from the RVs is shown with the black line, with $1\sigma$
+      errors in gray. The trend needed to produce the period decrease
+      of 12.6 milliseconds per year seen by \citet{bouma_wasp-4b_2019}
+      in transits is the purple dotted line in the bottom left.
      \label{fig:rv_o_minus_c}
@@
 \acknowledgements
-
 %
@@
 directorate.
-%
-This research has made use of the NASA Exoplanet Archive, which is
-operated by the California Institute of Technology, under contract
-with the National Aeronautics and Space Administration under the
-Exoplanet Exploration Program.
 %
@@
 \software{
-  \texttt{radvel} \citep{fulton_radvel_2018},
+  \texttt{radvel} \citep{fulton_radvel_2018}
+  %FIXME numerical python packages
 }

### 2019-09-11 rnaas_note/ms.tex
rnaas minor

@@
 This acceleration towards our line of sight is 4.3 times faster than
-the suggested $\approx 2\sigma$ trend \citet{knutson_friends_2014}
+the $\approx$$2\sigma$ trend \citet{knutson_friends_2014}
 found from the shorter baseline.  The radial velocity values are
@@
 
-While the quantitative agreement is not perfect, Occam's razor would
+Though the quantitative agreement is not perfect, Occam's razor would
 suggest that the line of sight acceleration is probably a sufficient
@@
 While further radial velocity observations of the system should help
-clarify the mass and semi-major axis of the companion, it seems
+in determining the mass and semi-major axis of the companion, it seems
 unlikely that additional transit observations will yield near-term

### 2020-01-16 first_paper_review/referee_response_0.txt
rename review->first_paper_review

--- 2019-03-27 review/referee_response_0.txt
@@
 
-* Added the following coauthors: Fei Dai, Natalia Guerrero, Bill Wohler,
-  Douglas Caldwell.
+* Added the following coauthors: Fei Dai, Natalia Guerrero, Bill Wohler.
+
+  - Fei Dai contributed an independent analysis of the RV dataset.
+
+  - Natalia Guerrero is a member of the TESS Science Office at MIT, and has
+    been included in accordance with the TESS collaboration's coauthor listing
+    policy, and for contributions to the manuscript.
+
+  - Bill Wohler is a member of the Science Processing Operations Center at NASA
+    Ames, and has been included in accordance with the TESS collaboration's
+    coauthor listing policy, and for contributions to the manuscript.
 

### 2020-01-16 paper/ms.tex
[wip] stage second paper
on WASP-4's linear trend

--- 2019-03-27 paper/ms.tex
@@
-\documentclass[12pt,twocolumn,tighten]{aastex62}
-%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
-\pdfoutput=1 %for arXiv submission
+% TODO:
+%
+% * email to full collaborator list
+% * add any missing CPS coauthors
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[RNAAS]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 
-\renewcommand*{\sectionautorefname}{Section} %for \autoref
-\renewcommand*{\subsectionautorefname}{Section} %for \autoref
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
-%% Add "Submitted to " argument.
-\received{December 22, 2018}
-\revised{\today}
-\accepted{---}
-\submitjournal{AAS journals.}
-
-\shortauthors{Bouma et al.}
-\shorttitle{Early Arrival of WASP-4\lowercase{b}}
-
-%\NewPageAfterKeywords
-\listofchanges
-
 \begin{document}
 
-\title{WASP-4\lowercase{b} Arrived Early for the TESS Mission}
+% NOTE: may want to improve title
+\title{WASP-4 is accelerating towards the Earth}
 
@@
 
+%
+% key authors: presumably Rachel did the Zorro reductions?
+%
 \author[0000-0002-0514-5538]{L. G. Bouma}
@@
 %
-\author[0000-0003-3438-843X]{C. Baxter}
-\affiliation{ API, University of Amsterdam, P.O. Box 94249, 1090 GE
-    Amsterdam, The Netherlands}
+\author{R. A. Matson}
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
 %
-\author[0000-0002-0628-0088]{W. Bhatti}
-\affiliation{ Department of Astrophysical Sciences, Princeton
-University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+
 %
-\author[0000-0002-8958-0683]{F. Dai}
-\affiliation{ Department of Astrophysical Sciences, Princeton
-University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+% contributing authors: alphabetical
 %
-\author[0000-0002-6939-9211]{T. Daylan}
-\altaffiliation{ Kavli Fellow }
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
+\author[0000-0002-0531-1073]{H. Isaacson}
+\affiliation{Astronomy Department, University of California, Berkeley,
+CA 94720, USA}
 %
-\author[0000-0002-0875-8401]{J.-M. D\'esert}
-\affiliation{ API, University of Amsterdam, P.O. Box 94249, 1090 GE
-Amsterdam, The Netherlands}
+\author[0000-0001-8638-0320]{A. W. Howard}
+\affiliation{Cahill Center for Astrophysics, California Institute of
+Technology, Pasadena, CA 91125, USA}
 %
-\author[0000-0002-5720-4827]{M. L. Hill}
-\affiliation{Department of Earth Sciences, University of California,
-Riverside, CA 92521, USA}
+\author{S. B. Howell}
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
 %
-\author[0000-0002-7084-0529]{S. R. Kane}
-\affiliation{Department of Earth Sciences, University of California,
-Riverside, CA 92521, USA}
+\author{H. Knutson}
+\affiliation{Division of Geological and Planetary Sciences, California
+Institute of Technology, Pasadena, CA 91125, USA}
 %
-\author[0000-0002-3481-9052]{K. G. Stassun}
-\affiliation{Vanderbilt University, Department of Physics \& Astronomy,
-6301 Stevenson Center Lane, Nashville, TN 37235, USA}
-\affiliation{Fisk University, Department of Physics, 1000 17th Avenue
-N., Nashville, TN 37208, USA}
-%
-\author{J. Villasenor}
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-% 
-%-------------------------------------
-% TESS Mission Architects:
-% These authors should be listed in this order
-% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
-%-------------------------------------
-%
-\author{G. R. Ricker} % grr@space.mit.edu CONFIRMED
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu CONFIRMED
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu CONFIRMED
-\affiliation{Center for Astrophysics \textbar \ Harvard \&
-Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
-%
-\author{S. Seager} % seager@mit.edu CONFIRMED
-\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
-Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
-%
-\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov CONFIRMED
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-%
-%-------------------------------------
-% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
-%These 9 authors should be listed in alphabetical order
-%-------------------------------------
-% 3 TSO COAUTHORS
-% 3 SPOC COAUTHORS
-% 3 POC COAUTHORS
-%
-\author[0000-0002-3321-4924]{Z. Berta-Thompson} % zachory.bertathompson@colorado.EDU CONFIRMED
-\affiliation{Department of Astrophysical and Planetary Sciences,
-University of Colorado, Boulder, CO 80309, USA}
-%
-% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % 
-%%%WAITING (LIKELY YES)
-% \affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-% \affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-%
-\author[0000-0001-8020-7121]{K. Col\'on} % knicole.colon@nasa.gov CONFIRMED
-\affiliation{NASA Goddard Space Flight Center, Exoplanets and Stellar
-Astrophysics Laboratory (Code 667), Greenbelt, MD 20771, USA}
-%
-\author{M. Fausnaugh} % faus@mit.edu CONFIRMED
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0002-5322-2315]{Ana Glidden} % aglidden@mit.edu CONFIRMED
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-\affiliation{ Department of Earth, Atmospheric, and Planetary Sciences,
-Massachusetts Institute of Technology, Cambridge, MA 02139, USA }
-%
-\author{N. Guerrero} % nmg@mit.edu CONFIRMED
-\affiliation{ Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA }
-%
-\author[0000-0001-8812-0565]{J. E. Rodriguez} % rodriguez.jr.joey@gmail.com CONFIRMED
-\affiliation{Center for Astrophysics \textbar \ Harvard \&
-Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
-%
-\author[0000-0002-6778-7552]{J. D. Twicken} % joseph.twicken@nasa.gov CONFIRMED
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
-%
-\author{B. Wohler} % bill.wohler@nasa.gov YES
-\affiliation{ NASA Ames Research Center, Moffett Field, CA 94035, USA }
-\affiliation{ SETI Institute, Mountain View, CA 94043, USA}
 
-\begin{abstract}
-    The Transiting Exoplanet Survey Satellite (TESS) recently observed
-    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
-    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
-    based on data stretching back to 2007.  This is unlikely to be the
-    result of a clock error, because TESS observations of other hot
-    Jupiters (WASP-6b, 18b, and 46b) are compatible with a constant
-    period, ruling out an 81.6-second offset at the 6.4$\sigma$ level.
-    The 1.3-day orbital period of WASP-4b appears to be decreasing at
-    a rate of $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  The
-    apparent period change might be caused by tidal orbital decay or
-    apsidal precession, although both interpretations have
-    shortcomings.  The gravitational influence of a third body is
-    another possibility, though at present there is
-    \replaced{no}{minimal} evidence for such a body. Further
-    observations are needed to confirm and understand the timing
-    variation.
-\end{abstract}
+%% Note that RNAAS manuscripts DO NOT have abstracts.
+%% See the online documentation for the full list of available subject
+%% keywords and the rules for their use.
+\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
+(1332), Transit timing variation method (1710)}
 
-\keywords{
-  planet-star interactions ---
-  planets and satellites: individual (WASP-4b, WASP-5b, WASP-6b,
-    WASP-12b, WASP-18b, WASP-46b) ---
-  binaries: close
-}
+%% Start the main body of the article. If no sections in the 
+%% research note leave the \section call blank to make the title.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
-\label{sec:intro}
 
-Although the Transiting Exoplanet Survey Satellite (TESS\added{,
-\citealt{ricker_transiting_2015}}) is designed to detect new planets,
-it is also \replaced{performing precise photometric monitoring
-of}{precisely monitoring} most of the planets that have been
-discovered by ground-based transit surveys over the last two decades.
-One \replaced{of the applications}{application} of the new TESS data
-is to search for timing anomalies in \added{previously known }hot
-Jupiter sytems.  Long-term monitoring of hot Jupiter transit and
-occultation times should eventually reveal \replaced{two distinct
-processes}{variations caused by three different phenomena}.
+Since the first discovery of transiting hot Jupiters, the hunt for
+tidal orbital decay has received much attention (CITE LIKE 20 PAPERS).
+These efforts are gradually beginning to yield fruit: by combining
+transit and occultation timing and radial velocity measurements, CITET
+Yee have shown that the data for the hot Jupiter WASP-12b are most
+compatible with orbital decay.
 
-\explain{The following four paragraphs have been modified to add
-emphasis to the possibility of an outer planet explaining the timing
-deviations. Some background information given in the original
-submission has been removed.}
+This study highlights a point that, while obvious, has perhaps not
+yet received due attention.
+The point is that the hunt for orbital decay of hot Jupiters will be
+crippled without long-term radial velocity monitoring programs.
+The reason is simple: when the timescale of orbital period change far
+exceeds the observing baseline, the first
+deviation from constant periodicity is always quadratic in the transit
+number:
+\begin{align}
+  t_{\rm tra} = t_0 + P E + \frac{1}{2}\frac{dP}{dE}E^2,
+\end{align}
+where
+\begin{align}
+  \frac{dP}{dE} = P \frac{dP}{dt},
+\end{align}
+$E$ is the transit number, $P$ is the orbital period, $t_0$ is the
+reference epoch, and $t_{\rm tra}$ is the transt mid-time.
 
-First, the orbits of most hot Jupiters should shrink because of tidal
-orbital decay
-\citep{counselman_outcomes_1973,hut_stability_1980,levrard_falling_2009,matsumura_tidal_2010}.
-Directly measuring the rate of decay might lead to an improved
-understanding of how friction dissipates the energy of tidal
-disturbances (a problem reviewed by \citealt{Mazeh2008} and
-\citealt{ogilvie_tidal_2014}). Second, if hot Jupiter orbits are
-appreciably eccentric, then long-term timing studies should reveal
-rotation of the orbital ellipse within the orbital plane (``apsidal
-precession'').  If this effect were observed, it could yield a measure
-of the planet's Love number, which would constrain the planet's
-interior structure \citep{ragozzine_probing_2009}.  The most
-convincing direct evidence yet found for either orbital decay or
-apsidal precession of a hot Jupiter is the case of WASP-12b, which has
-a transit period that has decreased by about 30~milliseconds per year
-over the last decade \citep{maciejewski_departure_2016,patra_2017}.
+There are ample reasons to expect orbital decay to be common
+(CITE).
+However there are reasons to expect the R\'omer delay to perhaps be
+even {\it more} common (CITE Knutson 14, Bryan 16).
+Both manifest to first order identically in transit times (up to the
+sign of the line-of-sight acceleration) -- as a non-zero period
+derivative.
 
-The final effect of interest that can produce period changes in hot
-Jupiter systems is gravitational acceleration caused by massive outer
-companions \citep[{\it e.g.},][Section~4]{agol_detecting_2005}.
-Prototypes include WASP-53 and WASP-81, systems in which the inner hot
-Jupiters are periodically perturbed by eccentric brown dwarf
-companions with semimajor axes of a few astronomical units
-\citep{triaud_peculiar_2017}.
+Specifically, \citet{knutson_friends_2014} showed that X\% of hot
+Jupiters show some form of radial acceleration in radial
+velocity time-series with baselines of YY years.
+Assuming a separation distribution of (WHATEVER), this implies that
+ZZ\% of hot Jupiters will {\it always} show a changing orbital period,
+over timescale of say 10 years.
 
-% DELETED THE FOLLOWING TEXT FROM SUBMITTED VERSION:
-% First, whenever the orbital angular momentum is less than 3 times
-% the rotational angular momentum, the system is vulnerable to tidal
-% orbital decay \citep{counselman_outcomes_1973,hut_stability_1980}.
-% Most hot Jupiters satisfy this condition and their orbits should be
-% shrinking \citep{levrard_falling_2009,matsumura_tidal_2010}.  The
-% rate of decay, which depends on how friction dissipates the energy
-% of tidal disturbances, is difficult to observe directly or calculate
-% from first principles. This has been a stubborn source of
-% uncertainty in the interpretation of many phenomena in stellar and
-% planetary astrophysics (as reviewed by \citealt{Mazeh2008} and
-% \citealt{ogilvie_tidal_2014}).
-%
-% Second, long-term timing studies may reveal apsidal precession, {\it
-% i.e.}, the rotation of the orbital ellipse within the orbital plane.
-% Apsidal precession has long been observed in eclipsing binaries
-% \citep[{\it
-% e.g.},][]{schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
-% For exoplanets, precessing orbits have been detected in a handful of
-% cases, including circumbinary systems (Kepler-413
-% \citealt{kostov_kepler-413b_2014}; Kepler-453
-% \citealt{welsh_kepler_2015}), planets orbiting rapidly rotating
-% stars (Kepler-13
-% \citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015});
-% and systems with multiple interacting giant planets (Kepler-108
-% \citealt{mills_kepler-108_2017}; Kepler-693
-% \citealt{masuda_eccentric_2017}).  Apsidal precession is not
-% regularly observed in hot Jupiter systems because their orbits are
-% nearly circular, which may itself be a consequence of star-planet
-% tides.  If it could be observed, then as
-% \citet{ragozzine_probing_2009} showed it would be possible to
-% measure the planet's Love number, which would be a valuable
-% constraint on the planet's interior.
-%
-% So far, the most convincing direct evidence for either orbital decay
-% or apsidal precession has come from WASP-12b.
-% \citet{maciejewski_departure_2016} and \citet{patra_2017} showed
-% that the transit period of WASP-12b has been steadily decreasing by
-% about 30~milliseconds per year over the last decade.
+This is specific to hot Jupiters. The abundance of giant planets
+exterior to Super-Earth sized systems is also high (CITE Bryan 2019).
 
-Here, we present evidence for a timing anomaly in the WASP-4 system.
-The hot Jupiter WASP-4b orbits a G7V star every 1.34 days,
-corresponding to an orbital distance of 5.5 stellar radii
-\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
-to search for departures from a constant period, because transits have
-been observed since 2007.  The orbital eccentricity is less than 0.018
-(2$\sigma$), based on the work of \cite{knutson_friends_2014}, who
-combined the available transit times, occultation times, and Doppler
-data.  The sky projection of the stellar obliquity is also compatible
-with zero, within about 10 degrees
-\citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011}.
+The main example of this study is the hot Jupiter WASP-4b which has an
+orbital period that from transits appears to be decreasing by about 10
+milliseconds per year.
+After discovering the timing variations in the TESS data, we obtained
+N additional radial velocity measurements using Keck-HIRES,
+extending the RV baseline by Y years.
+Previously, the radial velocities were sparse and were consistent with
+not showing any linear trend.
+Our new measurements reveal a line of sight acceleration of
+$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
+This translates to an expected period decrease simply from the light
+travel time effect of ZZ milliseconds per year --- about what is
+observed.
 
-In what follows, \S~\ref{sec:observations} presents the new TESS
-observations, and \S~\ref{sec:timing} describes our timing analysis.
-We tried fitting the data with three models: a constant period; a
-steadily shrinking period; and a slightly eccentric, precessing orbit.
-A constant period can be ruled out\replaced{, but we}{. We} cannot
-distinguish between the\replaced{ latter two possibilities}{
-  possibilities of a decaying orbit, a precessing orbit, and the
-  unmodeled possibility of an orbit being gravitationally perturbed by
-  an outer companion}.  \replaced{Either possibility}{Any of the three
-scenarios} would have interesting implications\replaced{, described in
-\S~\ref{sec:implications}}{ (\S~\ref{sec:implications}), and more data
-are required for a definitive ruling (\S~\ref{sec:future}).}
-Appendix~\ref{sec:verify_tess} considers the possibility that the
-WASP-4 timing anomaly is due to an error in timestamps in the TESS
-data products.  We found this possibility to be unlikely because none
-of the other hot Jupiters we examined show a timing offset with the
-same amplitude as was seen for WASP-4.
+\section{WASP-4's acceleration towards the Earth}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{New transits and system parameters}
-\label{sec:observations}
+Caveats: we of course checked the RV vs bisector scatter diagram, as
+well as the RV vs other activity indicators, to ensure we weren't
+seeing a spurious multi-hundred meter per second change.
 
-\begin{figure*}[t]
-    \begin{center}
-		\includegraphics[height=0.97\textheight]{f1.png}
-    \end{center}
-    \vspace{-0.8cm}
-    \caption{
-      {\bf TESS observations of WASP-4b.} On the left, black points
-      are TESS flux measurements, with a vertical offset applied. Blue
-      curves are best-fit models. The numbers printed next to each
-      lightcurve are the approximate transit times expressed in BJD
-      minus 2{,}457{,}000.  The right side shows the residuals.
-       \label{fig:lightcurves}
-    }
-\end{figure*}
+\section{Do other hot Jupiters show the Romer delay?}
+yes, HAT-P-7b. but omg utc to tdb.
 
-\begin{figure*}[t]
-    \begin{center}
-        \includegraphics[width=0.7\textwidth]{f2.png}
-    \end{center}
-    \vspace{-0.7cm}
-    \caption{
-        {\bf Phase-folded lightcurve of WASP-4b.} Gray points are
-        TESS \replaced{data}{flux measurements, with median 
-        $1\sigma$ uncertainty shown in the lower 
-        right.} 
-        Yellow points are binned measurements.  The bottom
-        panel shows the residuals.  The fit to the phase-folded
-        transit (blue line) is used when measuring mid-transit times
-        \replaced{for the}{of} individual transits (see 
-        \S~\ref{sec:measurement}).
-        \label{fig:phasefold}
-    }
-\end{figure*}
+\section{What is the expected Pdot distribution?}
 
-\subsection{Observations}
+SUMMARY OF LITERATURE
 
-WASP-4 was observed by TESS with Camera 2 from August 23 to September
-20, 2018, within the second ``sector'' of science operations.  The
-star is designated as TIC 402026209 in the TESS Input Catalog
-\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
-surrounding WASP-4 were averaged into 2-minute stacks by the onboard
-computer.  The data were downlinked via the Deep Space
-Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
-spacecraft timestamps were calibrated against the ground-station
-clocks.  The spacecraft clock times were then transformed by the
-Payload Operations Center into the {\it Temps Dynamique Barycentrique}
-(TDB) reference system.  The images were then reduced to lightcurves
-by the Science Processing Operations Center (SPOC) at NASA
-Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
-the known spacecraft trajectory to compute the barycentric time
-corrections on a target by target basis, and expressed the timestamps
-as Barycentric TESS Julian Dates (BTJD), which is simply the
-\replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus
-2{,}457{,}000.  Lightcurves that were flagged\deleted{ as Threshold
-Crossing Events} by the SPOC pipeline \added{as crossing a transit
-detection threshold} were then vetted and released by the MIT TESS
-Science Office to the Mikulski Archive for Space Telescopes on
-November 29, 2018~\citep{ricker_tess_alerts_2018}.
+This is a major challenge in the hunt for orbital decay.  The main
+example of this study is the hot Jupiter WASP-4b, which we recently
+showed transited $\approx$82 seconds early for the TESS spacecraft
+assuming a constant orbital period \citep{bouma_wasp-4b_2019}.  Our
+timing analysis included data from peer-reviewed literature for which
+the times were measured from a single transit, and for which the
+midpoint was allowed to be a free parameter. We also required the time
+system to be clearly documented\footnote{In other words, we needed to
+know whether any heliocentric or barycentric corrections had been
+performed, and whether the absolute time system was UTC or TDB.}. The
+transit times we analyzed spanned 2007 to 2019
+\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,dragomir_terms_2011,sanchis-ojeda_starspots_2011,nikolov_wasp-4b_2012,hoyer_tramos_2013,ranjan_atmospheric_2014,huitson_gemini_2017}.
+The combined timing data were best fit by an ephemeris with a constant
+negative period derivative; our best-fit decay rate was $\dot{P} =
+-12.6 \pm 1.2$ milliseconds per year.  Our interpretation was that the
+apparent period change could be caused by any of three scenarios: a
+decaying orbit, a precessing orbit, and an orbit being gravitationally
+perturbed by an outer companion.
 
-We began our analysis with the Presearch Data Conditioning (PDC)
-lightcurve\replaced{
-  \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}{, which
-has had non-astrophysical variability removed through the methods
-discussed by \citet{smith_kepler_apertures_2017} and
-\citet{smith_kepler_PDC_2017}}.  We \added{then} processed the
-lightcurve as follows.  First, we removed all points with non-zero
-quality flags.  This removed data that might have been adversely
-affected by ``momentum dumps,'' the firing of thrusters and resetting
-of reaction wheels\footnote{\added{The spacecraft pointing and
-momentum dumps are described in the data release notes:
-}\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
-that took place every 2.5 days during sector 2\deleted{
-  \citep{tess_data_product_description_2018}}.  The data during these
-events were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude'' \added{\citep[][Table
-28]{tess_data_product_description_2018}}.  For WASP-4, these flags
-were simultaneously set for 54 distinct cadences, and there were 10
-momentum dumps, averaging about 10 minutes of flagged data per dump.
-Out of caution, we clipped out an additional 10 minutes before and
-after every momentum dump.  We also removed the data within the first
-and last hours of both orbits, because of \replaced{ramp-like
-systematic effects that appear}{correlated red noise that appears}
-during those time ranges.
+Thereafter, \citet{southworth_transit_2019} reported 22 new transit
+times for the system, and confirmed that the updated series of transit
+times was consistent with a quadratic ephemeris.  Their interpretation
+of the timing variations did not differ in any major respects from our
+own, though with additional data they found a lower best-fit decay
+rate of $\dot{P} = -XX.X \pm X.X$ milliseconds per year.
 
-All told, we removed 8\% of the original data points, and were left
-with 18{,}165 measurements of the relative flux of WASP-4.  We
-normalized the data by dividing out the median flux.  We converted the
-timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
-appropriate 2{,}457{,}000 day offset
-\citep{tess_data_product_description_2018}.  Many of these and
-subsequent processing steps were performed using
-\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
-``flatten'' the lightcurves, as is often done with splines,
-polynomials, or Gaussian processes.  Instead, we modeled the
-out-of-transit flux variations simultaneously with the transit
-parameters, as described below.
+A separate study by \citet{baluev_homogeneously_2019} reported
+additional light curves. \citeauthor{baluev_homogeneously_2019}
+analyzed their newly obtained photometry, along with archival light
+curves that we omitted from our analysis due to systematic
+uncertainties in the absolute time system.
+\citeauthor{baluev_homogeneously_2019} found that when they used all
+the available TTV data, the need for a quadratic ephemeris was present
+``at the high $\sim 5-7$ sigma level''.  However, they pointed out
+that if they used lower-precision subsets of the available timing
+data, the necessity for the quadratic term decreased.
+\citeauthor{baluev_homogeneously_2019} also pointed out that the
+precise transit times reported by \citet{huitson_gemini_2017} were
+quite important in the time-series.  Overall,
+\citeauthor{baluev_homogeneously_2019} did not find the claim of a
+period decrease convincing.
 
-\subsection{Measuring the transit times}
-\label{sec:measurement}
+One line of follow-up needed to confirm and understand the timing
+variation was additional radial velocity observations.  The longest RV
+baseline previously available was six observations acquired by
+\citet{knutson_friends_2014} and the CKS team on Keck HIRES from
+2010 to 2013.  This past season we acquired four additional
+observations with Keck HIRES. Even before fitting out the hot Jupiter,
+the residuals showed a strong linear trend.
 
-Using the cleaned PDC lightcurve, we applied the Box Least Squares
-algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
-period, transit duration, and a reference epoch using the TESS data
-alone.  Based on the results, we isolated the data within 4 transit
-durations of each transit midpoint.  To find the transit parameters
-that best fit the data, we first fitted a line to the out-of-transit
-flux measurements surrounding each transit, and divided it out.  We
-then created a phase-folded lightcurve \added{from all 18 transits}
-and fitted a standard transit model using the analytic formulae given
-by \citet{mandel_analytic_2002} and implemented by
-\replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
-We assumed the orbit to be circular, consistent with the limits from
-radial velocities and occultation timing
-\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
-The free parameters were the reference epoch, the planet to star
-radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
-radius ratio $a/R_\star$, the inclination $i$, two quadratic
-limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
-orbital period $P$.
-
-We sampled the posterior probability distribution for all the
-parameters using the algorithm proposed by
-\citet{goodman_ensemble_2010} and implemented by
-\replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
-Table~1 gives the results, which are in reasonable agreement with the
-parameters reported by {\it e.g.},
-\citet{southworth_high-precision_2009} and
-\citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
-phase-folded lightcurve.
-
-To measure the transit times, we returned to the `cleaned' PDC time
-series and fitted the data within four transit durations of each
-transit separately. We used four free parameters: the time of
-mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
-slope and intercept of a linear trend to account for any slow
-variations unrelated to the transit.  We fixed the remaining
-parameters at the values that had been determined from the
-phase-folded TESS lightcurve.  The uncertainty in each
-\added{photometric} data point was set equal to the root-mean-square
-(rms) level of the out-of-transit data.
-
-To verify that the measured uncertainties are estimated accurately, we
-computed the \deleted{reduced }$\chi^2$\added{ value} for a linear
-ephemeris fit to the measured TESS mid-transit times.  We found that
-$\chi^2 = 9.2$, with $n=16$ degrees of freedom.  The variance of the
-$\chi^2$ distribution is $2n$, so we would expect $\chi^2 = 16 \pm
-5.7$.  Visually inspecting the residuals showed that the error
-variance had been overestimated, so we multiplied the measured TESS
-errors by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.
-This lowered the mean uncertainty of the transit midtimes from $29.8$
-to $22.6$ seconds.  We verified that omitting this step did not
-appreciably alter any of our conclusions.
-
-Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
-transit, the best-fit models, and the residuals.  Table~2 reports the
-mid-transit times and their uncertainties.  After binning the
-residuals to 1-hour windows, the lightcurves have an rms scatter of
-586\,{\rm ppm}.  The pre-launch TESS noise
-model\footnote{\url{github.com/lgbouma/tnm}, commit \texttt{be06f09}}
-(\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
-would have predicted an error budget consisting of the following terms
-added in quadrature: 410\,{\rm ppm} from photon-counting noise,
-202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
-zodiacal background light.  The level of background light appears to
-have been overestimated in the model.
-
-\subsection{Star and planet parameters}
-\label{sec:system_parameters}
-
-\begin{deluxetable}{lccc}
-\tabletypesize{\scriptsize}
-\tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
-\tablenum{1}
-
-\tablehead{
-\colhead{Parameter} & \colhead{Value} & \colhead{68\% Confidence Interval} & \colhead{Comment}
-}
-
-\startdata
-{\it Transit/RV parameters:} & & & \\
-  $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
-  $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
-  $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
-  $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
-  $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
-  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
-{\it Stellar parameters:} & & & \\
-  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
-  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
-  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
-  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
-  $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
-  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
-  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
-  $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
-  $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
-  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
-{\it Planetary parameters:} & & & \\
-  $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
-  $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
-  $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
-\enddata
-
-\tablecomments{
-  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
-  Orbital periods are in Table~4. The limb darkening parameters were
-  allowed to float around the \citet{claret_limb_2017} prediction, but
-  were unconstrained.
-  (B) \citet{triaud_spin-orbit_2010}.
-  (C) From HARPS spectra \citep{doyle_accurate_2013}.
-  (D) \citet{stassun_accurate_2017}.
-  (E) This work, see \S~\ref{sec:system_parameters}.
-  (F) \citet{gaia_collaboration_gaia_2018}.
-  (G) \citet{stassun_TIC_2018}.
-}
-
-\vspace{-1cm}
-\end{deluxetable}
-
-We calculated the stellar and planetary parameters in the following
-way.  We computed the star's spectral energy distribution based on the
-{\it Gaia} DR2 parallax (after making the small correction advocated
-by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
-the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$ and
-$V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
-2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
-wavelength range 0.4--22~$\mu$m.  We adopted the effective temperature
-from the work by \citet{doyle_accurate_2013}, who determined the
-spectroscopic parameters of WASP-4 using high-signal-to-noise
-observations with the High Accuracy Radial-velocity Planet Searcher
-(HARPS).  Then, we determined the stellar radius through the
-combination of the bolometric luminosity and the effective
-temperature, using the Stefan-Boltzmann law.  To determine the stellar
-mass, we first computed the mean stellar density based on the value of
-$a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
-(for the relevant equation, see \citealt{seager_unique_2003} or
-\citealt{winn_exoplanet_2010}).  The mass was calculated from the
-radius and density, and the orbital distance was also calculated from
-the radius and $a/R_\star$.  The planetary radius was calculated as
-the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
-mass was calculated based on the stellar mass, the radial-velocity
-amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
-inclination.
-
-Table~1 gives the resulting parameters, which we adopted for the
-remaining analysis.  \added{The uncertainties in our derived stellar
-and planetary parameters are propagated according to standard analytic
-formulae, under the assumption that the variables are uncorrelated and
-normally distributed.} Our \added{system} parameters are in agreement
-with those of previous investigators, but have the benefit of
-incorporating the {\it Gaia} parallax
-\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
-By comparing the star's luminosity and spectroscopic parameters with
-the outputs of the Yonsei-Yale stellar-evolutionary models, we found
-that WASP-4 is a main-sequence star with an age of approximately
-$7\,{\rm Gyr}$ \added{\citep{demarque_y2_2004}}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timing analysis}
-\label{sec:timing}
-
-\begin{figure}[t]
-    \begin{center}
-        \leavevmode
-        \includegraphics[width=0.49\textwidth]{f3.pdf}
-    \end{center}
-    \vspace{-0.6cm}
-    \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
-      Both plots show the deviations between observed and calculated
-      transit times, where the calculation is based only on
-      \deleted{the} pre-TESS data and assumes a constant period.  The
-      blue bands depict the $\pm$$1\sigma$ credible interval of the
-      predicted times.  {\it Top:} The full timing dataset spans 11
-      years. The darkest points correspond to the most precise data.
-      \added{The binned TESS point is the weighted average of 18 TESS
-      transits.}  {\it Bottom:} Close-up of the TESS observations. The
-      red band shows the average deviation of the TESS transits
-      \added{($\pm 1\sigma$)}, which arrived $81.6 \pm 11.7\ {\rm
-      seconds}$ earlier than predicted.
-      \label{fig:arrived_early}
-    }
-\end{figure}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-  \replaced{The residuals are the observed times minus the calculated
-  times assuming a constant period. The darkest points correspond to
-  the most precise data.}{The vertical axis shows the observed times
-  minus the calculated times assuming a constant period for transits
-  ({\it top}) and occultations ({\it bottom}). In the upper panel,
-  darker points correspond to more precise data.} The constant-period
-  model (gray line) is a poor description of the data.  Models with a
-  decreasing period (blue) or an eccentric, precessing orbit (orange)
-  provide	better fits\deleted{ to the data}.  The red square
-  represents the combination of all the TESS data and is for display
-  purposes only.  The models were fitted to all of the individual
-  transit times.
-  \label{fig:times}
-	}
-\end{figure*}
-
-\subsection{Pre-TESS timing measurements}
-\label{subsec:times}
-
-Table~2 gives the transit times we used in our analysis. We included
-data from peer-reviewed literature for which the analysis was based on
-observations of a single\deleted{, complete} transit, and for which
-the midpoint was allowed to be a free parameter. We also required that
-the time system be clearly documented. Many of the times were
-previously compiled by \citet{hoyer_tramos_2013}. We confirmed that
-the times in that paper were in agreement with the original sources
-and that barycentric corrections had been performed when needed.
-
-The earliest epoch is from EulerCam on the 1.2-m Euler telescope
-\citep{wilson_wasp-4b_2008}.  The second epoch is based on $z$-band
-photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
-with FORS2.  Subsequent observations were performed by
-\citet{winn_transit_2009}, \citet{dragomir_terms_2011},
-\citet{sanchis-ojeda_starspots_2011}, \citet{nikolov_wasp-4b_2012},
-\citet{hoyer_tramos_2013}, and \citet{ranjan_atmospheric_2014}.
-Finally, \citet{huitson_gemini_2017} acquired optical transit spectra
-with the 8.1-m Gemini South telescope between 2011 and 2014, one
-transit per season.  The per-point standard deviation of their
-lightcurves was a few hundred parts per million.  The average
-precision in their reported transit times was 5.6~seconds.  Since
-these data points carry significant weight in the analysis, we
-\replaced{checked}{corresponded with the authors to confirm} that the
-timestamps in their data represent mid-exposure times, that the
-barycentric correction was performed correctly, and that the time
-system of the final results was BJD$_{\rm TDB}$.  These same authors
-also used the same instrument and method to analyze other hot
-Jupiters, none of which showed a departure from a constant-period
-model.  Finally, these authors also measured two transit midpoints
-using Spitzer (Baxter et al., in prep); these times agree with the
-Gemini South results.
-
-We also compiled the available occultation times, which are given in
-Table~3.  The tabulated values have been corrected for the
-light-travel time across the diameter of the orbit by subtracting
-$2a/c = 22.8$ seconds from the observed time.
-\citet{beerer_secondary_2011} observed two occultations of WASP-4b
-using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
-\citet{caceres_ground-based_2011} detected an occultation from the
-ground in the $K_S$ band, and gave a time in HJD, without specifying
-the time standard.  We assumed the standard was UTC, and performed the
-appropriate corrections to convert to BJD$_{\rm TDB}$.  We verified
-that none of our conclusions would be changed if this assumption was
-mistaken. Finally, \citet{zhou_secondary_2015} observed an occultation
-with the Anglo-Australian Telescope.  They did not report the observed
-midpoint, but they did report a result for $e\cos\omega$ based upon
-the observed midpoint.  We calculated the implied midpoint using the
-formula \citep[{\it e.g.},][]{winn_exoplanet_2010}
-\begin{equation}
-  t_{\rm occ}(E) =
-  t_0 +  P E  +
-  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right),
-  \label{eq:occultation_time}
-\end{equation}
-\added{for $E$ the transit number, $t_0$ the reference epoch, $e$ the
-eccentricity, and $\omega$ the argument of pericenter.} In total,
-there are four \added{available} occultation times.
-
-\subsection{Analysis}
-\label{subsec:analysis}
-
-First, we \replaced{fitted}{performed a weighted least-squares fit of}
-a constant-period model (\added{a }``linear ephemeris'') to the
-pre-TESS data, and used it to extrapolate to the epochs of the TESS
-observations.  The residuals of the best fitting model are shown in
-Figure~\ref{fig:arrived_early}.  The transits observed by TESS
-occurred earlier than expected.  Because the TESS mission is still in
-an early stage, we were concerned about a possible offset in the TESS
-timestamps due to an error with the TESS clock or the data processing
-pipeline.  Appendix~\ref{sec:verify_tess} describes some tests that
-convinced us that a simple offset is unlikely. Assuming that the
-observed timing variation is astrophysical, we proceeded by exploring
-three models for the timing data in a manner identical to the study
-by~\citet{patra_2017}.
-
-The first model assumes a constant orbital period on a circular orbit:
-\begin{align}
-  t_{\rm tra}(E) &= t_0 + PE,\\
-  t_{\rm occ}(E) &= t_0 + \frac{P}{2} + PE,
-\end{align}
-where $E$ is the epoch number.  We defined the epoch numbers such that
-$E=0$ is near the weighted average of the observed times.  This helps
-to reduce the covariance between $t_0$ and $P$.
-
-The second model assumes the period is changing at a steady rate:
-\begin{align}
-  t_{\rm tra}(E) &=
-    t_0 + PE +
-    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2, \\
-  t_{\rm occ}(E) &=
-    t_0 + \frac{P}{2} + PE +
-    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
-\end{align}
-The three free parameters are the reference epoch $t_0$, the period at
-the reference epoch, and the period derivative, ${\rm d}P/{\rm d}t =
-(1/P) {\rm d}P/{\rm d}E$.
-
-The third model assumes the planet has a slightly eccentric orbit, and
-that the line of apsides is rotating \citep{gimenez_revision_1995}:
-\begin{align}
-  t_{\rm tra}(E) &= 
-		t_0 + P_{\rm s}E
-    - \frac{e P_{\rm a}}{\pi} \cos\omega,\\
-  t_{\rm occ}(E) &= 
-    t_0 + \frac{P_{\rm a}}{2} + P_{\rm s}E
-    + \frac{e P_{\rm a}}{\pi} \cos\omega,
-\end{align}
-where $P_{\rm s}$ is the sidereal period, $e$ is the eccentricity,
-$P_{\rm a}$ is the anomalistic period, and $\omega$ is the argument of
-pericenter.  In this model the angular velocity of the line of apsides
-${\rm d}\omega/{\rm d}E$ is constant,
-\begin{equation}
-  \omega(E) = \omega_0 + \frac{{\rm d}\omega}{{\rm d}E} E,
-\end{equation}
-and the sidereal and anomalistic periods are connected through the
-equation
-\begin{equation}
-  P_{\rm s} = P_{\rm a} \left(
-    1 - \frac{1}{2\pi}\frac{{\rm d}\omega}{{\rm d}E}
-    \right).
-\end{equation}
-\added{The sidereal period is the duration required to return to the
-same orientation with respect to the stars; the slightly longer
-anomalistic period is the duration required to reach a fixed longitude
-with respect to the rotating line of apsides.} The five free
-parameters of this model are $(t_0, P_{\rm s}, e, \omega_0, {\rm
-d}\omega/{\rm d}E)$, denoting the reference epoch, the sidereal
-period, the eccentricity, the argument of pericenter at the reference
-epoch, and the angular velocity of the line of apsides.
-
-\added{We fitted each model by assuming a Gaussian likelihood and
-sampling over the posterior probability distributions. The prior for
-the quadratic model allowed the period derivative to have any sign.
-We considered two possible priors for the precession model: the first
-is a wide prior that allows non-physical values of the planetary
-Love number (Equation~\ref{eq:love_number}). The second prior requires
-that the planetary Love number is less than that of a sphere of
-constant density.  }
-
-Figure~\ref{fig:times} shows the residuals with respect to the
-constant-period model.  The best-fitting constant-period model has
-$\chi^2 = 174$ and 61 degrees of freedom.  The best-fitting
-\replaced{decreasing-period}{quadratic} model has $\chi^2 = 62.6$ and
-60 degrees of freedom.  The best-fitting precession model \added{under
-the wide prior} has $\chi^2 = 64.3$ and 58
-degrees of freedom.\added{ Under the physical prior, the best-fit is
-slightly worse, with $\chi^2=64.8$}. The precession and quadratic
-models both provide much better fits than the constant-period model.
-The difference in $\chi^2$ between the linear and quadratic models
-corresponds to $p \approx 10^{-26}$.
-
-The \replaced{decreasing-period}{quadratic} model provides a
-marginally better fit to the data than the precession model.  It is
-favored by $\Delta \chi^2 = 1.7$, and has two fewer free parameters.
-A useful heuristic for model comparison is the Bayesian Information
-Criterion,
-\begin{equation}
-  {\rm BIC} = \chi^2 + k\log n,
-\end{equation}
-where $k$ is the number of free parameters, and $n$ is the number of
-data points. In this case, $n=62$.  The difference in the BIC between
-the precession and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm
-prec} - {\rm BIC}_{\rm quad} = 10$\added{, assuming a wide prior for
-the precession model}. This corresponds to a Bayes factor of
-$1.1\times 10^{4}$.  Likewise, the Akaike Information Criterion favors
-the constant-period-derivative model by $\Delta {\rm AIC} = 5.8$.
-Differences of this magnitude are traditionally deemed ``strong
-evidence'' that one model is a better description of the data than the
-other \citep{kass_bayes_1995}, although we prefer to reserve judgment
-until more data can be obtained.
-
-In the \replaced{decreasing-period}{quadratic} model, the period
-derivative is
-\begin{equation}
-\dot{P}
-  = - (4.00 \pm 0.38)\times 10^{-10}
-  = - 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}.
-  \label{eq:dP_dt_obs}
-\end{equation}
-For comparison, the best-fitting period derivative of WASP-12b is
-$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$
-\citep{maciejewski_departure_2016,patra_2017}.  If both planets are
-truly falling onto their stars, then WASP-4b is falling at about half
-the rate of WASP-12b.
-
-In the precession model, the best-fit eccentricity is
-\begin{equation}
-  e = (1.92^{+ 1.93}_{- 0.76})\times10^{-3}
-  \label{eq:e_obs}
-\end{equation}
-The longitude of periastron advances by $\dot{\omega} =
-13.6^{+4.7}_{-3.6}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
-period is $27^{+10}_{-7}$ years.  All of the best-fitting parameters
-(the medians of the posterior distributions) and the 68\% credible
-intervals are reported in Table~4.
-
-\explain{The following subsection has been added since the original
-submission, per the referee's suggestions.}
-\subsection{Possible systematic errors}
-
-To assess the overall robustness of our results, we considered a few
-possible systematic effects in the timing dataset.
-
-\paragraph{Suspect pre-TESS light curves}
-Some of the pre-TESS light curves have incomplete phase coverage: a
-handful of the transit times in Table~2 are from lightcurves with
-gaps.  A separate issue is the effect of spot-crossing anomalies on
-transit timings.  To address these concerns, we repeated the
-model-fitting described above, but omitted epochs -827, -804, -537,
-and -208 because of gaps in their coverage.  We also omitted epochs
--526 and -561 because of visible spot anomalies during the transits.
-(All epoch numbers are as in Table~2.)  The resulting best-fit transit
-timing model parameters were all within $1\sigma$ of the values quoted
-in Table~4.  The uncertainties, goodness-of-fit statistics, and model
-comparison statistics did not appreciably change.
-
-\paragraph{Spot-crossing events in TESS data}
-To explore the effect of possible spot-crossing events on the TESS
-transit time measurements, we performed a separate test.  We injected
-triangular spot-anomalies with amplitude $0.03\%$ and duration 30
-minutes at random phases into each transit.  The amplitude was chosen
-to be larger than the spot-crossing anomalies observed by
-\citet{southworth_high-precision_2009} and
-\citet{sanchis-ojeda_starspots_2011}, and the duration was chosen to
-be comparable to those of previously observed events.  Spots of these
-amplitudes resemble the possible anomalies present in transits
-``1360.54'' and ``1372.58'' of Figure \ref{fig:lightcurves}.
-
-With spots injected, we repeated our measurement of the transit times.
-On average, the measured transit times did not change after injecting
-spots, because the flux deviations are equally likely to occur in the
-first and second halves of the transit.  For individual transits,
-there were no cases for which the timing deviation was larger than one
-minute.  The largest shifts occur when the spot anomaly occurs during
-transit ingress or egress, in which case the measured mid-time is
-shifted either late or early by between 30 and 50 seconds.
-
-The TESS observations could therefore be skewed
-early if there were spot-crossing events during {\it every} egress.  Two
-arguments rule out this possibility.  (1) The lightcurve residuals do
-not show evidence for these events.  (2) The stellar rotation period
-is between 20 and 40 days, and the sky-projected stellar obliquity is
-less than 10 degrees
-\citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
-Since the planet orbits every 1.3 days, requiring that spot anomalies
-always occur during egress would be equivalent to requiring a stellar
-spot distribution that is exquisitely (and thus implausibly)
-distributed to match the planet egress times.
-
-\paragraph{Detrending choices in pre-TESS data}
-There is a final concern that is difficult to address.  We collected
-the mid-transit time values derived by different authors, who used
-heterogeneous methods to fit and detrend their lightcurves.  We have
-also assumed that these authors have correctly documented the time
-systems in which the data are reported.  Further, though many choices
-in transit-fitting ({\it e.g.}, parametrization of limb-darkening and
-eccentricity) do not affect transit mid-time measurements, different
-detrending approaches can asymmetrically warp transits and shift
-mid-transit times.  The magnitude of this systematic effect is hard to
-quantify, but the situation is fairly clear from
-Figure~\ref{fig:times}.  Many independent authors provided transit
-measurements shortly after WASP-4b's discovery, and the data are
-consistent with each other. \citet{huitson_gemini_2017} provided the
-most important data from epochs 0-1000. If their data were
-systematically affected by detrending choices or time-system confusion
-at the level of several times their reported uncertainties, then it
-possible that the orbital period is constant despite the evidence in
-the TESS data.  For this reason, we paid careful attention to the
-\citet{huitson_gemini_2017} data set, and corresponded with the
-authors to confirm that their results are not affected by systematic
-effects of the required amplitude.
-
-None of the concerns mentioned in this subsection seem likely to
-explain the observed timing variations. We proceed by considering
-possible astrophysical explanations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Interpretation}
-\label{sec:implications}
-
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[height=0.88\textheight]{f5.pdf}
-  \end{center}
-  \vspace{-0.6cm}
-  \caption{
-    {\bf WASP-4b in the context of other hot Jupiters.}
-    Most of the data in these plots are from \citet{bonomo_gaps_2017},
-    who measured the eccentricities using radial velocities.  The
-    colored symbols highlight WASP-4, WASP-12 (which also shows
-    evidence for a decreasing period), and the\deleted{ collection of}
-    hot Jupiters analyzed in Appendix A.
-    {\bf \it Top:}
-    Key parameters relevant to eccentricity damping.  Solid squares
-    represent planets with a securely detected nonzero eccentricity.
-    \added{Circles represent planets whose orbits are consistent with
-    circular.} Lines of constant damping timescale are drawn based on
-    Equation~\ref{eq:de_dt} and assuming $Q_{\rm p}' = 10^5$.  WASP-4b
-    has one of the shortest eccentricity damping times.
-    {\bf \it Middle:} 
-    Key parameters relevant to orbital decay.  Lines of constant decay
-    timescale are drawn based on Equation~\ref{eq:da_dt}, assuming
-    $Q_\star' = 10^7$.  WASP-4b has a relatively short orbital decay
-    timescale, although it is not as extreme a case in this regard as
-    it is for eccentricity damping.
-    {\bf \it Bottom:}
-    A Hertzsprung-Russell diagram for hot Jupiter hosts. WASP-4
-    appears to be on the main sequence.
-    \label{fig:context}
-  }
-\end{figure*}
-
-\subsection{Orbital decay}
-
-If the timing variation is \deleted{being caused }caused entirely by
-orbital decay, then \replaced{the}{the best-fit model parameters yield a}
-characteristic\replaced{ timescale of the decay is}{ decay timescale of}
-\begin{equation}
-  \frac{P}{ {\rm d}P/{\rm d}t } = 9.2 \, {\rm Myr}.
-\end{equation}
-For comparison, the corresponding time for WASP-12b is 3.2\,Myr
-\citep{patra_2017}.
-
-If WASP-4 really is undergoing rapid orbital decay, then how many of
-the other known hot Jupiters should have orbits that are decaying at
-detectable (or nearly detectable) rates?  Figure~\ref{fig:context}
-compares some key properties of WASP-4 with those of a larger ensemble
-of hot Jupiters.  The middle panel displays two parameters that
-strongly affect the expected orbital decay timescale, $P
-M_\star/M_{\rm p}$ and $a/R_\star$.  WASP-4 has one of the shortest
-theoretical timescales for orbital decay.  \replaced{There
-are}{Figure~\ref{fig:context} shows} about 20 hot Jupiters (including
-WASP-12) for which the theoretical timescale is shorter.  In almost
-all of those cases, though, the planet was discovered more recently
-than WASP-4 and a decade-long baseline of observations is not yet
-available.  A separate consideration not shown in
-Figure~\ref{fig:context} is that the hot Jupiter host stars have a
-variety of different structures, from being fully convective to nearly
-fully radiative, which may lead to widely divergent tidal dissipation
-timescales.
-
-In the simple ``constant phase lag'' model for tidal interaction
-\citep{zahn_tidal_1977}, the rate of dissipation can be parametrized
-by a modified\footnote{For stars, $k_\star \sim \mathcal{O}(10^{-2})$,
-so it is important to explicitly distinguish $Q_\star'$ from $Q_\star$
-\citep[{\it e.g.},][]{schwarzschild_structure_1958}.} quality factor,
-$Q_\star' = 3 Q_\star / (2k_\star)$.  Here, $Q_\star$ is the ratio
-between the energy stored in the equilibrium deformation of the star
-and the energy lost to heat per tidal period \citep[{\it
-e.g.},][]{goldreich_q_1966}.  A larger $Q_\star$ implies less
-efficient tidal dissipiation. The dimensionless number $k_\star$ is
-the stellar Love number, which is smaller when the star's density
-distribution is more centrally concentrated.  In this model, once the
-planet's spin and orbit are synchronized, then the semi-major axis and
-eccentricity evolve as \citep[Appendix B of][]{metzger_optical_2012}
-\begin{align}
-  \frac{1}{\tau_{\rm e}} &=
-  \frac{|\dot{e}|}{e} =
-    \frac{63 \pi } {2 Q_{\rm p}' }
-    \left( \frac{R_{\rm p}}{a} \right)^5
-    \left( \frac{M_\star}{M_{\rm p}} \right)
-    \left( \frac{1}{P} \right)
-  \label{eq:de_dt}
-  \\
-  \frac{1}{\tau_{\rm a}} &=
-  \frac{|\dot{a}|}{a} =
-    \frac{9 \pi } {Q_\star' }
-    \left( \frac{R_\star}{a} \right)^5
-    \left( \frac{M_{\rm p}}{M_\star} \right)
-    \left( \frac{1}{P} \right).
-  \label{eq:da_dt}
-\end{align}
-The orbital period evolves as
-\begin{equation}
-\label{eq:dP_dt}
-  \dot{P} = -\frac{27\pi}{2 Q_\star'}
-            \left(\frac{M_{\rm p}}{M_\star}\right)
-            \left(\frac{R_\star}{a}\right)^5.
-\end{equation}
-The modified quality factor of WASP-4 corresponding to the observed
-value of $\dot{P}$ is
-\begin{equation}
-  Q_\star' = (2.9 \pm 0.3)\times10^4.
-\end{equation}
-This is about an order of magnitude lower than the value that was
-inferred for WASP-12b.  It is also smaller than most theoreticians
-would have expected.  The $Q_{\rm Jup}'$ value of Jupiter is estimated
-to be $\approx$$1.4 \times 10^5$, based on the observed motions of the
-Galilean moons \citep{lainey_strong_2009}.  For stars, studies of the
-binary eccentricity distribution have been interpreted with tidal
-models, giving $Q_\star' \approx 10^5 - 10^7$ \citep[{\it
-e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
-geller_direct_2013,milliman_wiyn_2014}.  Population studies of hot
-Jupiter systems have also been undertaken, generally finding $Q_\star'
-\approx 10^5 - 10^8$ using different models
-\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
-For instance, motivated by the rapid rotation of some hot Jupiter
-hosts
-\citep{pont_empirical_2009,ciceri_hats-15b_2016,penev_hats-18b_2016},
-\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
-systems under the influence of a magnetized wind and a constant
-phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
-(1.2^{+1.0}_{-0.5})\times10^7$, which would correspond to $\dot{P}
-\approx -30$ microseconds per year.  This strongly disagrees with the
-the period change that we have observed.
-
-\citet{essick_orbital_2016} studied the problem of the orbital decay
-of hot Jupiters using a theory in which gravity modes are excited at
-the base of the stellar convective zone, propagate inward through the
-radiative core and break near the stellar core, leading to energy
-dissipation.  They predicted the stellar quality factors in hot
-Jupiter systems to vary from $Q_\star' \approx 10^5 - 10^6$.  From
-their Equation 26, the prediction for WASP-4 is $Q_\star' =
-7\times10^5$, which is an order of of magnitude larger than implied by
-the observed period change.
-
-The applicability of the \citet{essick_orbital_2016} model depends on
-the evolutionary state of the star. \citet{weinberg_tidal_2017} showed
-that more rapid dissipation --- enough to account for the period
-change of WASP-12b --- could exist in stars that have begun evolving
-into red giants.  The bottom panel of Figure~\ref{fig:context} shows a
-Hertzsprung-Russell diagram of hot Jupiters hosts, including WASP-4.
-On the y-axis is $G=g-\mu$, for $g$ the apparent {\it Gaia}-band magnitude,
-and $\mu$ the distance modulus reported by
-\citet{gaia_collaboration_gaia_2018}.  The x-axis is the effective
-temperature from \citet{bonomo_gaps_2017}, which for WASP-4 agrees
-within $1\sigma$ of that from Table~1.  Inspecting the HR diagram,
-WASP-4 shows little evidence of being evolved, in agreement with our
-analysis from \S~\ref{sec:system_parameters}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% By a similar argument though, WASP-12 would not clearly stand out as a
-% candidate subgiant, and \citet{weinberg_tidal_2017} suggest it may in
-% fact be a subgiant (though \citealt{bailey_understanding_2019} do not
-% find evidence that supports this suggestion).  A more thorough
-% isochronal analysis would be of interest, particularly if the transit
-% and occultation times continue to vary.  Such an analysis is beyond
-% the scope of this study.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To summarize, if the observed period change is caused entirely by
-tidal orbital decay, then the constant-phase-lag tidal model implies a
-stellar tidal dissipation rate that is higher than expected by at
-least an order of magnitude.  It might be possible that we are
-observing at a special time, shortly after the planet's inward
-migration, or when the planet is near resonance with a stellar
-oscillation mode.  Tidal dissipation rates might also be increased if
-the star is just turning off the main sequence.  Another hypothesis,
-recently advanced by \citet{millholland_obliquity_2018} for the case
-of WASP-12b, is that an exterior planet could be trapping WASP-4b's
-spin vector in a high-obliquity state, leading to rapid dissipation
-through planetary obliquity tides.
-
-\subsection{Apsidal precession}
-\label{sec:apsidal_precession}
-
-\explain{The result given in Equation~\ref{eq:love_number} has been
-corrected from the submitted version. The previously stated result, k2
-= $1.5^{+2.1}_{-1.1}$ had a
-typo that led to printing of 84th and 16th percentiles, instead of the
-difference of these percentiles from the median. 
-The uncertainties also previously did not propagate uncertainty in the
-planet radius and semimajor axis, which enter at the fifth power and
-are now included. 
-Finally, we now also specify the assumed priors, as they add a
-nuance in the interpretation.   }
-
-If instead the observed timing variation is just a small portion of an
-apsidal precession cycle, then the orbital eccentricity is a few times
-$10^{-3}$, and the full precession period is about 27 years.
-\citet{ragozzine_probing_2009} calculated apsidal precession periods
-for hot Jupiters, finding them to range between about 10 and 100
-years. They highlighted that for many hot Jupiters, including WASP-4b,
-the theoretical precession rate is dominated by the non-Keplerian
-force due to the planet's tidal bulge.  Precession from general
-relativity, the planet's rotational bulge, and the star's rotational
-and tidal bulges contribute at the 10\% level at most.  Thus, a
-measurement of the precession rate can be used to determine the
-planet's Love number.  From their Equation 14, the implied Love number
-for WASP-4b is
-\begin{align}
-  k_{2,{\rm p}} &= 1.59^{+0.70}_{-0.47}
-    \quad{\rm (wide\ prior),} \label{eq:love_number} \\
-  k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
-    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5])
-\end{align}
-where the uncertainties in the semimajor axis and planet radius have been
-propagated through our Markov chains, and $\mathcal{U}$ denotes the uniform 
-distribution.  For comparison, the Love number
-of Jupiter is about 0.55 \citep{wahl_tidal_2016,ni_empirical_2018},
-and a uniform density sphere has $k_2 = 1.5$. The uncertainty in
-$k_{2,{\rm p}}$ for WASP-4b is large because the eccentricity,
-reference time, and ${\rm d}\omega/{\rm d}E$ have strongly correlated
-errors, and the \added{four available}\deleted{measured} occultation
-times only weakly constrain these parameters
-(Figure~\ref{fig:future}).
-
-\added{ The results under both priors are larger than the Love number
-of Jupiter, at weak statistical signifiance.  Imposing the
-physically-motivated prior truncates the allowed values of $k_{2,{\rm
-p}}$, and drives the precession period towards larger values.
-Different priors therefore change the predicted evolution of the
-orbit.  While the precession model matches the orbital decay model for
-the next two decades in transits (Figure~\ref{fig:future}), in
-occultations the precession model can deviate from the orbital decay
-model, particularly if one requires the planetary Love number to be
-smaller. This suggests that promptly obtaining occultation time
-measurements for the system could help rule out the apsidal precession
-model.  For the time being, the data are insufficient to make a
-stronger judgement.  }
-
-\added{Regardless of the detailed model assumptions, the}
-\deleted{The} main \added{physical }problem with the apsidal precession
-hypothesis is to explain why the
-eccentricity would be as large as $\sim$$10^{-3}$ despite rapid tidal
-circularization.  For WASP-4, Equation~\ref{eq:de_dt} gives $\tau_e =
-0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The star is several billion
-years old, so unless the planet arrived very recently, any initial
-eccentricity should have been lowered well below $\sim$$10^{-3}$.  The
-top panel of Figure~\ref{fig:context} compares the expected
-eccentricity damping time of WASP-4b with that of other transiting
-giant planets.  WASP-4b has one of the shortest known eccentricity
-damping times.
-
-\paragraph{Neighboring companion}
-One way to maintain a significant eccentricity is through the
-gravitational perturbations from another planet.
-\citet{mardling_long-term_2007} considered the long-term tidal
-evolution of hot Jupiters with companions.  The companion in their
-model is coplanar, and can have a mass down to an Earth-mass; the main
-requirement is that both the hot Jupiter and the outer companion start
-on eccentric orbits.  They found that although the early phases of the
-two-planet eccentricity evolution occur quickly, the final phase of
-the joint eccentricity evolution towards circularity would occur on
-timescales several orders of magnitude longer than the circularization
-time of an isolated hot Jupiter (see their Figures~4 and 5).
-
-A separate way a neighboring companion could excite the hot Jupiter's
-eccentricity is through the
-Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
-In this case, the orbital plane of the outer companion, ``c'', would
-need to be inclined relative to that of the hot Jupiter, ``b'', by at
-least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
-For the Kozai-Lidov mechanism to operate at maximum efficiency, we
-need \citep[][Equation 20]{bailey_understanding_2019}
-\begin{equation}
-  M_{\rm c} > 7.5\,M_\oplus
-  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
-  \label{eq:kozai_bound}
-\end{equation}
-\replaced{where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major 
-axis of WASP-4b and the hypothetical WASP-4c, and owing}{where $a_{\rm b}$ is 
-the semi-major axis of WASP-4b, $M_{\rm c}$ is the mass of the hypothetical 
-WASP-4c, and $a_{\rm c}$ is WASP-4c's semi-major axis. Owing}
-to our imprecise
-measurement\added{, in Equation~\ref{eq:kozai_bound}} we have assumed 
-WASP-4b's Love number is $k_{2,{\rm
-b}}\approx 0.6$\added{, similar to Jupiter}.  For the RV signal of the 
-companion to remain
-undetected, it would need to be in the residual $({\rm O-C})_{\rm RV}
-= 15.2\,{\rm m s}^{-1}$ reported by \citet{triaud_spin-orbit_2010}.
-Again following \citet{bailey_understanding_2019}, this implies
-\begin{align}
-  M_{\rm c} &<
-  ({\rm O-C})_{\rm RV}
-  \left( \frac{M_\star a_b}{G} \right)^{1/2}
-  \left( \frac{a_c}{a_b} \right)^{1/2}
-  f^{-1/2}
-  \nonumber
-  \\
-  M_{\rm c} &< 
-  23.7\,M_\oplus
-  \times 
-  \left( \frac{a_c}{a_b} \right)^{1/2}
-  f^{-1/2},
-  \label{eq:rv_bound}
-\end{align}
-for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto \sin^2 i_{\rm c}$
-a geometric prefactor that depends on the argument of periastron
-$\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
-(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b is
-transiting, $f$ can be arbitrarily small, and both of the preceding
-limits can be satisfied.
-
-\paragraph{Fluctuations in the gravitational potential from convection}
-An independent mechanism to pump the eccentricity invokes the
-gravitational fluctuations from stellar convection \citep[][Section
-7]{phinney_pulsars_1992}.  From equation 7.33 of that work, the
-mean-squared eccentricity of the orbit is
-\begin{equation}
-  \langle e^2 \rangle =
-  \frac{ 2 \langle E_{\rm e} \rangle }{\mu n^2 a^2}
-  = 6.8\times10^{-5}
-  \frac{(L^2 R_{\rm conv}^2 M_{\rm conv}^2)^{1/3}}{\mu n^2 a^2},
-\end{equation}
-where $L$ is the stellar luminosity, $R_{\rm conv}$ and $M_{\rm conv}$
-are the width and mass of the convective region, $\mu$ is the reduced
-mass, $n$ is the orbital frequency, and $a$ is the semi-major axis.
-\added{For the luminosity, reduced mass, and semi-major axis, we used
-values from Table~1, combined with the Stefan-Boltzmann law and
-standard definitions. To estimate the width and mass of the convective
-region we ran the MESA code for a star with mass and metallicity
-matched to WASP-4, and the input physics detailed in the MIST
-isochrones project
-\citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
-We identified the tachocline boundary using the mixing types specified
-in the resulting radial profiles, and found $R_{\rm conv} \approx 0.33
-R_\odot$, and $M_{\rm conv} \approx 9\times10^{-4} M_\odot$. } For
-WASP-4, this implies $\langle e^2 \rangle^{1/2} \lesssim 10^{-5}$.
-Hence, this mechanism does not seem capable of producing the required
-eccentricity of $\sim$$10^{-3}$.
-
-\explain{The subsequent subsection adds a description of our RV
-analysis, which we did not include in the original submission. Some of
-the material was taken from what was previously the "Other possible
-explanations" section.}
-\subsection{Timing variation due to line of sight acceleration}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Radial velocities of WASP-4 (top), and residuals from the
-    best-fit Keplerian model (bottom).} The lower panel shows the
-    best-fit linear trend inferred from the RV data (black line,
-    $1\sigma$ errors in gray), and the trend that would be needed to
-    produce the period decrease seen in transits (purple dotted line).
-    Since both the RV and transit timing datasets are sparse after
-    2013, a distant massive companion on an eccentric orbit might
-    still explain the observations.
-	\label{fig:rvs}
-  \vspace{-0.3cm}
-	}
-\end{figure}
-
-An acceleration of the center of mass of the system towards our line
-of sight could cause a decrease in the apparent orbital period.  The
-period derivative would be
-\begin{equation}
-	\dot{P} = \frac{\dot{v}_{\rm r} P}{c},
-\end{equation}
-where $\dot{v}_{\rm r}$ is the time derivative of the radial velocity.
-
-\citet{knutson_friends_2014}, combining radial-velocity data from
-their own program with those of \citet{wilson_wasp-4b_2008},
-\citet{pont_determining_2011}, and \citet{husnoo_observational_2012},
-found evidence for long-term trend in WASP-4, with low statistical
-significance:
-\begin{equation}
-  \dot{v}_{\rm r} =
-     -0.0099^{+0.0052}_{-0.0054}
-     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}.
-\end{equation}
-This acceleration translates to an expected $\dot{P} = -(4.4 \pm 2.4)
-\times10^{-11}$. The period decrease from the observed RV trend is an
-order of magnitude smaller than the observed $\dot{P}$ from transit
-times, $(4.0\pm 0.4)\times10^{-10}$ (Table~4).
-
-Nonetheless, it is intriguing that the host star shows a weakly
-significant acceleration, and with the correct sign needed to explain
-the transit timing variations.  Given the potential importance of the
-radial velocities for interpreting this system, we performed an
-independent analysis, as follows.
-
-First, we collected the usable RV measurements from CORALIE, HARPS,
-and HIRES.  We included the CORALIE measurements from
-\citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010}, using
-the homogeneous radial velocities calculated by the latter authors. We
+For completeness, in our analysis we included CORALIE measurements
+from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
+using the homogeneous velocities reported by the latter authors.  We
 included the HARPS values reported by \citet{pont_determining_2011},
@@
 because they were calculated using a different pipeline than the
-longer-baseline \citeauthor{pont_determining_2011} HARPS measurements,
-and necessary inclusion of an extra offset term would nullify their
-statistical value.  Finally, we included the five HIRES measurements
-taken over many years by \citet{knutson_friends_2014}.
+longer-baseline \citeauthor{pont_determining_2011} measurements, and
+necessarily inclusion of an extra offset term would nullify their
+statistical value.
 
 We then fitted a single Keplerian orbit, plus instrument offsets,
-jitters, and a long-term trend
+jitters, and a linear trend
 \citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
 on the period and time of inferior conjunction using the values from
-Table~4, and fixed the eccentricity to zero, consistent with results
-from \citet{beerer_secondary_2011}, \citet{knutson_friends_2014} and
-\citet{bonomo_gaps_2017}.  The remaining free parameters were the
-velocity semi-amplitude, the instrument zero-points, the instrument
-jitters (an additive white noise term for each instrument), linear
-($\dot{v_{\rm r}}$), and optionally second-order ($\ddot{v_{\rm r}}$)
-acceleration terms.
+\citet{bouma_wasp-4b_2019}, and fixed the eccentricity to zero,
+consistent with results from \citet{beerer_secondary_2011},
+\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The
+remaining free parameters were the velocity semi-amplitude, the
+instrument zero-points, the instrument jitters (an additive white
+noise term for each instrument), and optionally a linear acceleration
+term ($\dot{v_{\rm r}}$).
 
-We found that the best-fitting model with both linear and quadratic
-radial velocity terms was marginally preferred (by $\Delta
-\mathrm{BIC} = 5.8$) over the best-fitting model with only a linear
-term.  Regardless, for consistency with \citet{knutson_friends_2014},
-who fixed the quadratic component of the long-term trend to zero, in
-Figure~\ref{fig:rvs} we show best-fitting models for the linear-trend
-case.  The best-fit value for the line of sight acceleration,
+%FoHJ: gammadot = -0.0099^{+0.0052}_{-0.0054}
+The AIC and BIC strongly favored a model with a linear trend ($\Delta
+{\rm BIC}=73$ compared to a model without the linear trend).
+Figure~\ref{fig:rv_o_minus_c} shows the best-fitting model, with the
+orbit of WASP-4b subtracted.  The best-fit radial velocity derivative,
+$\dot{v}_{\rm r}$, is
 \begin{equation}
-  \dot{v}_{\rm r} =
-     -0.0077^{+0.0052}_{-0.0047}
-     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
+  \dot{v}_{\rm r} = -0.0422^{+0.0028}_{-0.0027}\,{\rm m\,s^{-1}\,day^{-1}}.
 \end{equation}
-is within $1\sigma$ of the value reported by
-\citet{knutson_friends_2014}. (Note that the CORALIE data used in our
-and their analyses differ, as we included additional measurements
-reported by \citealt{triaud_spin-orbit_2010}).  The implied period
-derivative is still therefore about an order of magnitude smaller than
-our observed $\dot{P}$ from transit timing.
+This acceleration towards our line of sight is 4.3 times faster than
+the $\approx$$2\sigma$ trend \citet{knutson_friends_2014}
+found from the shorter baseline.  The radial velocity values are
+available in the data-behind-the-figure online version of this {\it
+Note}.
 
-To summarize, only about one tenth of the observed period decrease can
-be explained through a constant acceleration of the WASP-4 system's
-center of mass.  However, given the limited amount and uneven time
-coverage of the existing radial-velocity data (Figure~\ref{fig:rvs}),
-it remains possible that the center of mass has a more complex motion,
-perhaps due to a companion on an eccentric orbit \citep[{\it e.g.},
-WASP-53 or WASP-81,][]{triaud_peculiar_2017}.  It would be useful to
-gather more radial-velocity data to confirm or refute this
-possibility.
+Under the assumption of constant acceleration, $\dot{P} = \dot{v}_{\rm
+r} P / c$, the implied period decrease that should be seen in
+transits is roughly $5.9 \pm 0.4$ milliseconds per year.  All the
+available TTV data show a period decrease of $\approx 8-12$
+milliseconds per year~\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.
 
-\subsection{Applegate effect}
+Though the quantitative agreement is not perfect, Occam's razor would
+suggest that the line of sight acceleration is probably a sufficient
+explanation for the apparent decrease of WASP-4b's orbital period.
+While further radial velocity observations of the system should help
+in determining the mass and semi-major axis of the companion, it seems
+unlikely that additional transit observations will yield near-term
+constraints on orbital decay.
 
-A separate candidate explanation for the timing deviations is the
-\citet{applegate_mechanism_1992} effect.  Some eclipsing binaries
-exhibit period modulations with amplitudes of $\lesssim 0.05\,{\rm
-days}$ over timescales of decades \citep[{\it
-e.g.},][]{soderhjelm_geometry_1980,hall_relation_1989}.  The Applegate
-mechanism explains these modulations by positing that the internal
-structure of a magnetically active star changes shape via cyclic
-exchange of angular momentum between the inner and outer zones of the
-star.  This model could also apply to a hot Jupiter orbiting a star
-with a convective zone.  The changing gravitational quadrupole of the
-star would cause the orbit of the planet to precess on the timescale
-of the stellar activity cycle.  An essential difference between this
-process and apsidal precession is that Applegate timing variations
-need not be strictly periodic \citep[{\it
-e.g.},][Figure~12]{soderhjelm_geometry_1980}. This mechanism would
-also produce transit and occultation timing deviations of the same
-sign, while for apsidal precession they would have opposite signs.
-For WASP-4, \citet{watson_orbital_2010} estimated that the Applegate
-effect could produce timing deviations of up to 15 seconds, depending
-on the modulation period of the stellar dynamo.  If this analysis is
-accurate, then the Applegate mechanism cannot explain the majority our
-observed $82$ second variation.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% RNAAS: you get one figure.
 
-\explain{The following subsection had material moved into the ``Timing
-variation due to line of sight acceleration'' subsection. Figure 7 was
-updated to clarify the Love number prior for the precession model.}
-\subsection{Other possible explanations}
+\begin{figure}
+    \begin{center}
+		\includegraphics{rv_fit.pdf}
+    \end{center}
+    \vspace{-0.8cm}
+    \caption{
+      {\bf Radial velocity observations of WASP-4.}
+      The orbit of WASP-4b has been subtracted.  The best-fit linear
+      trend from the RVs is shown with the black line, with $1\sigma$
+      errors in gray. The trend needed to produce the period decrease
+      of 12.6 milliseconds per year seen by \citet{bouma_wasp-4b_2019}
+      in transits is the purple dotted line in the bottom left.
+     \label{fig:rv_o_minus_c}
+    }
+\end{figure}
 
-\begin{figure*}[!ht]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f7.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Further observations are needed to confirm and understand
-    the timing variations of WASP-4b.} \replaced{Dots}{Symbols} are as
-    in Figure~\ref{fig:times}.  Lines are 100 random draws from the
-    posteriors of the apsidal precession model (orange), and the
-    orbital decay model (blue).  \added{The prior for the precession
-    model assumes $k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5]$; dashed orange
-    lines emphasize samples with planetary Love numbers below 0.75}.    
-		\label{fig:future}
-	}
-\end{figure*}
-
-There are two \replaced{final}{other} small effects worth noting.  The
-first is the \citet{shklovskii_possible_1970} effect due to the star's
-proper motion, which leads to an apparent period change of $P\mu^2 d/
-c$, which is only $6\times10^{-13}$ for the case of WASP-4.  The
-second effect, described by \citet{rafikov_stellar_2009}, comes from
-the star's on-sky motion altering our viewing angle, and leads to an
-observed apsidal precession. The corresponding period change is
-$\dot{P} \sim (P\mu)^2/2\pi$, which is on the order of $10^{-21}$ for
-WASP-4, too small to be of any consequence.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Call for additional observations}
-\label{sec:future}
-
-A primary purpose of this work has been to call attention to the
-timing anomaly of WASP-4 that has been sighted by TESS, and alert
-observers to the need for follow-up transit timing, occultation
-timing, and radial-velocity monitoring.  There is no unique
-interpretation of the current data, and two of the possibilities ---
-orbital decay and apsidal precession --- would be of great interest to
-confirm. Detection of orbital decay would lead to an unusually direct
-determination of a stellar dissipation rate. Detection of apsidal
-precession would give a rare constraint on the interior density
-distribution of an exoplanet.\added{ The third possibility --- a
-massive outer companion --- would be the least exotic option, but
-nonetheless a valuable discovery.}
-
-If TESS is extended beyond its primary mission, it will likely observe
-additional transits of WASP-4b in the early 2020s.
-High-precision transit observations with
-larger telescopes would also be useful.  In order to decide between
-orbital decay and apsidal precession, occultation measurements in both
-the near term and also in the mid-2020s will be needed 
-(Figure~\ref{fig:future}).  More
-radial-velocity data would help in the search for additional bodies
-that could be causing dynamical perturbations, or an overall
-acceleration of the host star.  The transit duration variations are
-expected to be of order 10 seconds \citep{pal_periastron_2008}, and so
-may remain out of reach.
-
-TESS will also be monitoring the other known hot Jupiters, which will
-reveal whether the timing anomalies seen in WASP-12 and now WASP-4 are
-commonplace, and may shed some light on the circumstances in which
-they arise. Other wide-field photometric surveys, such as the Next
-Generation Transit Survey \citep{wheatley_next_2018}, HATPI
-(\href{https://hatpi.org}{hatpi.org}) and PLATO
-\citep{rauer_plato_2014} will also extend the time baseline of transit
-timing for a large number of systems.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \acknowledgements
-L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
-A.~Bailey, J.~Goodman, K.~Patra\added{, D. Ragozzine,} and V.~Van
-Eylen, and are grateful to the people who have turned TESS from an
-idea into reality.  \added{L.G.B. thanks A.~Bixel and E.~May for
-clarifying details concerning the available IMACS lightcurves.}
-%
-WASP-4 was included on the ``short-cadence'' target list thanks to the
-Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
-and G011183 respectively). 
-%
-J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
-supporting this work.
-%
-T.D. acknowledges support from MIT's Kavli Institute as a Kavli
-postdoctoral fellow.
-%
-J.M.D. acknowledges funding from the European Research Council (ERC)
-under the European Union's Horizon 2020 research and innovation
-programme (grant agreement no. 679633; Exo-Atmos), and support from
-the NWO TOP Grant Module 2 (Project Number 614.001.601).
-%
-J.E.R. was supported by the Harvard Future Faculty Leaders
-Postdoctoral fellowship.
 %
@@
 %
-\facility{
-	TESS \citep{ricker_transiting_2015},
-	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
+\software{
+	\texttt{astroplan} \citep{astroplan2018},
+  \texttt{radvel} \citep{fulton_radvel_2018}
+  %FIXME numerical python packages
 }
-%
-\software{
-  \texttt{astrobase} \citep{bhatti_astrobase_2018},
-  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
-  \texttt{astroquery} \citep{astroquery_2018},
-  \texttt{BATMAN} \citep{kreidberg_batman_2015},
-  \texttt{corner} \citep{corner_2016},
-  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
-  \texttt{IPython} \citep{perez_2007},
-  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
-  \texttt{MESA} 
-  \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
-  \texttt{numpy} \citep{walt_numpy_2011}, 
-  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
-  \texttt{radvel} \citep{fulton_radvel_2018},
-  \texttt{scikit-learn} \citep{scikit-learn},
-  \texttt{scipy} \citep{jones_scipy_2001}.
-}
-
-%\clearpage
-%\newpage
-\input{WASP-4b_transit_time_table.tex}
-
-%\clearpage
-%\newpage
-\input{occultation_time_table.tex}
-
-%\clearpage
-%\newpage
-\input{model_fit_table.tex}
-
-\clearpage
 
@@
 
-%\clearpage
-%\newpage
-
-\appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Verifying the TESS Timestamps Using Other Hot Jupiters}
-\label{sec:verify_tess}
-
-\begin{figure*}[t!]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.98\textwidth]{f8.pdf}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-		{\bf There is no evidence for a systematic offset between TESS
-			times and the barycentric reference.} While the WASP-4b transits
-		fell about 82 seconds earlier than expected, other well-observed
-		hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
-		Ticks are observed TESS transit midtimes; the orange distribution
-		is a gaussian centered on zero with standard deviation
-		(\replaced{$\sigma_{\rm predicted}$}{$\sigma_{\rm pre-TESS}$}) calculated from the pre-TESS transit
-		times.  The blue distribution is a gaussian centered on the
-		weighted average of the TESS times, with width equal to the
-		uncertainty in the mean, {\it i.e.}, the standard deviation
-		of the TESS residual times divided by $\sqrt{N-1}$, with $N$ the
-		number of transits.
-		\label{fig:hjs}
-	}
-\end{figure*}
-
-An obvious concern that one might have about the WASP-4\added{b}
-timing anomaly is that there might be a systematic offset between the
-TESS time system and the time system in which the previous
-observations have been reported.  There is a precedent for this type
-of error: data from the Kepler mission was afflicted by a systematic
-timing error that was corrected only late in the
-mission~\citep[][Section 3.4]{kepler_DR19_2013}.
-
-If the observed timing delay in WASP-4b were caused by a systematic
-global offset between the TESS time system and the ${\rm BJD}_{\rm
-TDB}$ reference, we would expect that it would be apparent in other
-hot Jupiter systems, too. It would also be apparent in eclipsing
-binary observations and any other periodic phenomena that have been
-observed over a long time baseline. Here we examine only hot Jupiters
-because of our greater familiarity with the data.
-
-We repeated all the data reduction and analysis steps described in
-this paper for other hot Jupiters observed by TESS for which timing
-data exists spanning many years.  First, we checked which hot Jupiters
-were observed over the first three TESS sectors using a combination of
-\texttt{tessmaps}\footnote{\url{github.com/lgbouma/tessmaps}, 
-	commit \texttt{569bbc2}} and
-TEPCat \citep{southworth_homogeneous_2011}.  We recalculated the
-barycentric corrections using the \citet{eastman_achieving_2010} code,
-and found values that agreed with the lightcurve headers to within
-about 1 second.  We then selected hot Jupiters for which there were at
-least five distinct epochs reported in the peer-reviewed literature.
-We required that each observation be of a single transit, that the
-midpoint be fit as a free parameter, and that the time system be
-clearly documented.  Our final hot Jupiter sample included WASP-4b,
-5b, 6b, 18b, and 46b.  The collected and measured times are given in
-Tables~5, 6, 7, and 8 for each.
-
-We determined the best-fitting constant-period ephemeris based on the
-pre-TESS data. Then we used the parameters and uncertainties in the
-best-fitting model to calculate the predicted transit times during
-the TESS observation period, as well as the uncertainty in the
-predicted times.  The uncertainties are 11, 94, 18, 42, and 59 seconds
-for WASP-4b, 5b, 6b, 18b, and 46b, respectively.  By comparing the
-observed and predicted times, Figure~\ref{fig:hjs} shows that WASP-4b
-is the only hot Jupiter that transited significantly earlier than
-expected.
-
-To use these results to place a quantitative limit on any global clock
-offset, for each hot Jupiter we considered the model
-\begin{equation}
-  t_{\rm tra}(E) = t_0 + PE + t_{\rm offset},
-\end{equation}
-for $t_{\rm offset}$ a systematic constant offset between the reported
-timestamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
-were
-\begin{align}
-  t_0 &\sim \mathcal{N}[t_0', \sigma_{t_0'}], \\
-  P &\sim \mathcal{N}[P', \sigma_{P'}], \\
-  t_{\rm offset} &\sim \mathcal{U}[-20\sigma_{t_0'},20\sigma_{t_0'}],
-\end{align}
-where $\mathcal{N}$ and $\mathcal{U}$ denote a normal and uniform
-distribution, $(t_0', P')$ are the best-fit reference time and period
-using only the pre-TESS transit times, and $(\sigma_{t_0'},
-\sigma_{P'})$ are the corresponding uncertainties.
-
-For each planet, we asked: what fraction of the posterior for $t_{\rm
-offset}$ is consistent with an offset worse than $81.6$ seconds?  For
-WASP-4b, the answer is unsurprisingly 50\%.  For WASP-6b, the most
-constraining object, about 1 sample in 2 million is consistent with
-such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 103 samples
-would be consistent with this timing offset ($2.3\sigma$), and in
-WASP-46b, the limit is 1 in 49 samples ($2.0\sigma$).  For WASP-5b,
-the predicted time is too imprecise to rule out timing offsets at the
-necessary amplitude.  Multiplying the three independent probabilities
-for WASP-6b, 18b, and 46b, we can rule out $t_{\rm offset} < -81.6\
-{\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.
-
-%\clearpage
-%\newpage
-\input{WASP-5b_transit_time_table.tex}
-\input{WASP-6b_transit_time_table.tex}
-\input{WASP-18b_transit_time_table.tex}
-\input{WASP-46b_transit_time_table.tex}
-
-%\clearpage
-%\newpage
-
-%\newpage
-
-\listofchanges
 \end{document}

### 2020-01-16 paper/ms.tex
add rv_vs_activity_checks

@@
 
-Caveats: we of course checked the RV vs bisector scatter diagram, as
-well as the RV vs other activity indicators, to ensure we weren't
-seeing a spurious multi-hundred meter per second change.
+\subsection{Trend due to stellar activity?}
+
+First, we considered whether the measured RV trend is correlated with
+WASP-4's emission in the Ca II H \& K lines, as quantified with the
+chromospheric $S$-index (CITE: Wright 2004).  We only analyzed the
+HIRES velocities for this step, since they are the main source of
+signal in our analysis.  First, we subtracted the orbital solution
+from the Keck-HIRES velocities.  Then, following CITE Bryan 2016 and
+2019, we calculated the Spearman rank correlation coefficient between
+the $S$-index and the orbit-subtracted velocities.  We found a
+correlation coefficient of 0.16, which weakly hints at a correlation,
+but is not statistically significant (the corresponding $p$-value is
+0.65).  We therefore do not think that the linear trend can be
+attributed to stellar activity.
 

### 2020-01-16 src/plot_rvs_RNAAS.py

@@ -111,7 +111,10 @@ def main(make_my_plot=1):
     """
 
     # initialization script used to make the fix_gammadot fits
-    basedir = "/home/luke/Dropbox/proj/WASP-4b_anomaly/"
+    basedir = os.path.join(
+        os.path.expanduser('~'),
+        "Dropbox/proj/WASP-4b_anomaly/"
+    )
 
     setupfn = os.path.join(basedir,"src/WASP4.py")
 

### 2020-01-16 src/rv_vs_activity_checks.py

@@ -1,0 +1,70 @@
+"""
+(environment): py37_emcee2
+"""
+import os
+import pandas as pd, numpy as np, matplotlib.pyplot as plt
+
+from plot_rvs_RNAAS import _get_fit_results
+
+from scipy.stats import spearmanr
+
+def check_rv_resid_vs_CaHK():
+
+    df = pd.read_csv('../data/20190911_jump_wasp4_rv.csv', comment='#')
+
+    # initialization script used to make the fix_gammadot fits
+    basedir = os.path.join(
+        os.path.expanduser('~'),
+        "Dropbox/proj/WASP-4b_anomaly/"
+    )
+
+    setupfn = os.path.join(basedir,"src/WASP4.py")
+
+    outputdir = os.path.join(
+        basedir,
+        "results/rv_fitting/LGB_20190911_fix_gammaddot"
+    )
+
+    (rvtimes, rvs, rverrs, resid, telvec, dvdt,
+     curv, dvdt_merr, dvdt_perr, time_base) = _get_fit_results(
+         setupfn, outputdir
+    )
+
+    sel = (telvec == 'HIRES')
+
+    offset=2450000
+    times  = rvtimes[sel] - offset
+    hires_rv_resid = resid[sel]
+    hires_rv_err = rverrs[sel]
+    # ensure order between df and the rv/time vectors is matched
+    np.testing.assert_array_equal(df.bjd-offset, times)
+    hires_CaHK = np.array(df.svalue)
+    hires_CaHK_err = np.array(df.svalue_err)
+
+    r_value, p_value = spearmanr(hires_rv_resid, hires_CaHK)
+
+    ########################################## 
+
+    savpath='../results/20190911_rv_resid_vs_CaHK_svalue.png'
+
+    plt.close('all')
+    f,ax = plt.subplots(nrows=1,ncols=1,figsize=(4,3))
+
+    ax.errorbar(hires_rv_resid, hires_CaHK, xerr=hires_rv_err,
+                yerr=hires_CaHK_err, fmt='ko')
+
+    ax.text(
+        0.02, 0.98, 'Spearman R: {:.2f} (p={:.3f})'.format(r_value, p_value),
+        va='top', ha='left', transform=ax.transAxes, color='black'
+    )
+
+    ax.set_ylabel('CaHK Svalue')
+    ax.set_xlabel('HIRES RV - orbit model [m/s]')
+
+    f.tight_layout()
[…]

### 2020-01-16 src/get_knutson_population_pdots.py
add get_knutson_population_pdots

@@ -1,0 +1,146 @@
+import os, json, pickle
+
+import numpy as np, pandas as pd
+from astropy import units as u, constants as c
+from numpy import array as nparr
+from astroquery.nasa_exoplanet_archive import NasaExoplanetArchive
+
+from astropy.coordinates import SkyCoord
+
+from astroquery.mast import Tesscut
+
+from astrobase.services.identifiers import simbad_to_tic
+from astrobase.services.mast import tic_objectsearch
+from astrobase.services.tesslightcurves import (
+    get_tess_visibility_given_ticids,
+    is_two_minute_spoc_lightcurve_available
+)
+
+def main():
+
+    fdf = calculate_and_format_knutson_pdot_table()
+    tdf = check_tess_visibility_and_availability(fdf)
+
+    import IPython; IPython.embed()
+
+def check_tess_visibility_and_availability(fdf):
+
+    outpath = '../results/knutson_all_pdots_tess_visibility.csv'
+
+    if not os.path.exists(outpath):
+
+        ticids = [simbad_to_tic(n.rstrip(' b')) for n in nparr(fdf['planet'])]
+        fdf["TICID"] = ticids
+
+        badnames = ['TrES-2 b','TrES-3 b','TrES-4 b']
+        badnameticids = [399860444, 116264089, 159742538]
+        for n, ticid in zip(badnames, badnameticids):
+            fdf.loc[fdf['planet']==n,'TICID'] = ticid
+
+        sector_strs, full_sector_strs = (
+            get_tess_visibility_given_ticids(nparr(fdf.TICID))
+        )
+
+        fdf['sector_str'] = sector_strs
+        fdf['full_sector_str'] = full_sector_strs
+
+        have_2min = list(
+            map(is_two_minute_spoc_lightcurve_available,
+                nparr(fdf.TICID).astype(str)
+               )
+        )
+        fdf['2min_spoc_lc_available'] = have_2min
+
+        fdf.to_csv(outpath, sep=';', index=False)
+        print('made {}'.format(outpath))
+
+    else:
+        fdf = pd.read_csv(outpath, sep=';')
+
+    return fdf
+
+
+
+
+def calculate_and_format_knutson_pdot_table():
+
+    sig_df = calculate_pdots(signifiant_trends=1)
[…]

### 2020-01-16 paper/ms.tex
wrote procedure to do the M/a^2 constraints

@@
 attributed to stellar activity.
+
+\subsection{Constraints on companion masses and semi-major axes}
+
+Given a linear radial velocity trend, we can place limits on the mass
+and semi-major axis of any planetary or stellar companions in the
+system.
+The constraints on the two parameters are degenerate: $\dot{\gamma}
+\propto M_{\rm p} a^{-2}$.
+Nonetheless, the additional information from the Zorro images can help
+limit the available parameter space by setting an outer limit on the
+semimajor axis, and a maximum brightness (and thereby mass) of any
+putative companions.
+
+The general procedure we use to derive these constraints has been developed by
+\citet{wright_linear_trends_2007}, \citet{montet_trends_2014},
+\citet{knutson_friends_2014}, \citet{bryan_statistics_2016,bryan_excess_2019},
+and others.
+For this study, we closely follow the procedure of \citet{bryan_excess_2019}.
+
+We begin by defining a $50\times50$ grid in true planetary mass and
+semimajor axis, evenly spaced from 3 to 300$\,$$M_{\rm Jup}$ and 0.5 to
+500$\,$${\rm AU}$.
+
+In each cell, we inject 500 simulated companions.
+
+For each companion, we draw a set of orbital parameters for WASP-4b
+from {\bf Table X} of \citet{bouma_wasp-4b_2019}.
+Then, this orbital solution is subtracted, leaving the linear trend.
+
+We then draw a mass and semimajor axis from log-uniform
+distributions. We draw the inclination from a uniform distribution in
+$\cos i$.
+The eccentricity is drawn from \citet{kipping_beta_2013}'s long-period
+exoplanet Beta distribution ($a=1.12$, $b=3.09$).
+
+Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulation companion, we then
+fit for the time and argument of periastron, and velocity zero-points.
+
+We then calculate the log-likelihood value of the best-fit solution.
+
+We convert the resulting $50\times50\times500$ cube of log-likelihood
+values to probability, and marginalize over the samples in each grid
+cell to derive a probability distribution in mass and semi-major axis.
+
+The distribution is shown in {\bf Figure~X}.
 

### 2020-01-20 src/derive_mass_semimaj_constraints.py
add mass vs semimaj constraint plot

@@ -37,10 +37,11 @@ The distribution is shown in {\bf Figure~X}.
 ###########
 
 import os, pickle
-import numpy as np, pandas as pd
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
 from astropy import units as u, constants as c
 from itertools import product
 from numpy import array as nparr
+from datetime import datetime
 
 import radvel
 from radvel.mcmc import statevars
@@ -56,6 +57,9 @@ from scipy.stats import beta
 
 from WASP4 import time_base
 
+from cdips.plotting import savefig
+import matplotlib as mpl
+
 ##########
 # driver #
 ##########
@@ -64,7 +68,7 @@ def loguniform(low=0, high=1, size=None, base=10):
     return np.power(base, np.random.uniform(low, high, size))
 
 def derive_mass_semimaj_constraints(
-    setupfn, rvfitdir, chainpath, verbose=True
+    setupfn, rvfitdir, chainpath, verbose=False
     ):
 
     np.random.seed(42)
@@ -77,7 +81,7 @@ def derive_mass_semimaj_constraints(
     )
 
     n_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-    n_injections_per_cell = 500 # want: 500
+    n_injections_per_cell = 10 # want: 500
 
     mass_grid = (
         np.logspace(np.log10(1), np.log10(300), num=n_grid_edges)*u.Mjup
@@ -108,8 +112,12 @@ def derive_mass_semimaj_constraints(
             mass_lower = mass_grid[mass_left_ind].squeeze()
             sma_lower = sma_grid[sma_left_ind].squeeze()
 
-            if verbose:
-                print(mass_lower, mass_upper, sma_lower, sma_upper)
+            pstr = (
+                '{:s} {:.2f}, {:.2f}, {:.2f}, {:.2f}'.
+                format(datetime.now().isoformat(),
+                       mass_lower, mass_upper, sma_lower, sma_upper)
+            )
+            print(pstr)
 
             # sample the models from the chain
             rv_model_single_planet_and_linear_trend, chain_sample_params = (
@@ -184,13 +192,44 @@ def derive_mass_semimaj_constraints(
 
         log_like_arr = pickle.load(open(outpath, 'rb'))
 
[…]

### 2020-01-20 src/radvel_utils.py
minor

@@ -151,10 +151,6 @@ def initialize_sim_posterior(data, mass_c, sma_c, incl_c, ecc_c,
 
     k_c = semi_amplitude(Msini, period_c, Mtotal, ecc_c, Msini_units='jupiter')
 
-    if verbose:
-        print('period: {:.1f} days'.format(period_c))
-        print('k_c: {:.1f} m/s'.format(k_c))
-
     P = templateWASP4(data, period_c, ecc_c, k_c, tc_c=2455470,
                       w_c=0.42*np.pi/2, gammajit_dict=gammajit_dict)
 

### 2020-01-20 paper/ms.tex
minor
better words

--- 2020-01-16 paper/ms.tex
@@
 attributed to stellar activity.
+
+\subsection{Constraints on companion masses and semi-major axes}
+
+Given a linear radial velocity trend, we can place limits on the mass
+and semi-major axis of any planetary or stellar companions in the
+system.
+The constraints on the two parameters are degenerate: $\dot{\gamma}
+\propto M_{\rm p} a^{-2}$.
+Nonetheless, the additional information from the Zorro images can help
+limit the available parameter space by setting an outer limit on the
+semimajor axis, and a maximum brightness (and thereby mass) of any
+putative companions.
+
+The general procedure we use to derive these constraints has been developed by
+\citet{wright_linear_trends_2007}, \citet{montet_trends_2014},
+\citet{knutson_friends_2014}, \citet{bryan_statistics_2016,bryan_excess_2019},
+and others.
+For this study, we closely follow the procedure of \citet{bryan_excess_2019}.
+
+We begin by defining a $50\times50$ grid in true planetary mass and semimajor
+axis, evenly spaced from 1 to 300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.
+
+We then consider the possibility that an additional companion in any particular
+cell can explain the observed linear trend.
+
+In each cell, we consider 500 hypothetical companions.
+
+We assign each companion a mass and semimajor axis from log-uniform
+distributions within the grid cell. We draw the inclination from a uniform
+distribution in $\cos i$.  The eccentricity is drawn from
+\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution ($a=1.12$,
+$b=3.09$).
+
+For each simulated companion, we then draw a sample from the converged chains
+of our initial model of WASP-4b, plus its linear trend. We subtract
+the planet's orbital solution, leaving the linear trend.
+
+Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulated outer companion,
+and the fixed instrument offsets and jitters from the MCMC chains, we then
+perform a maximum likelihood fit for the time and argument of periastron of the
+outer simulated companion.
+
+We then convert the resulting $50\times50\times500$ cube of best-fit
+log-likelihood values to probabilities, and average over the samples in
+each grid cell to derive a probability distribution in mass and semi-major
+axis.
+
+The distribution is shown in {\bf Figure~X}.
 

### 2020-01-20 paper/ms.tex
update paper

@@
 attributed to stellar activity.
+
+\subsection{Zorro imaging}
+
+Convinced that the linear trend warranted closer examination, we
+performed speckle imaging using Zorro at Gemini to attempt to detect
+any nearby companions.  To our knowledge, this is the first high
+resolution imaging that has been performed for WASP-4 (FIXME: IS THIS
+TRUE?).
+
+We observed WASP-4 twice with Gemini-South, 11/12 Sept. (8 sets of
+1000 frames each) and 28/29 Sept. (7 sets of 1000 frames each). The
+second set had better seeing and produced a slightly better result.
+The spatial scale is from the diffraction limit of Gemini at the
+wavelength observed to about 1.25'' plus a bit more.
+
+% TODO: might Howell suggest a paragraph for how the contrast ratios
+% were derived?
+To derive contrast ratios from these images, we did XXXXX.
+
+We would like to convert the contrast ratios obtained through the
+Zorro imaging to constraints on the masses of putative companions, and
+their separations from the host star.
+
+To do this, we followed the methodology of CITE Montet+2014, and opted
+to employ the CITE Baraffe+2003 models for substellar mass objects and
+the MESA isochrones for stellar mass objects (CITE Choi et al XXXX).
+In the intermediate regime, near 100 Mjup, we interpolated between the
+two sets of models.
+
+We assumed that companions would have contracted, and asserted the
+system age to be 5 Gyr.
+
+Due to the narrow-band filters of the Zorro imager, we further assumed
+that all sources had blackbody spectra. While this is clearly false,
+we do not readily have access to the planetary and stellar atmosphere
+models needed for the correct calculation. However, the CITE
+Baraffe+2003 and MESA models do report effective temperatures and
+bolometric luminosities.  We opt to use these theoretical quantities,
+combined with the measured Zorro filter wheels, to the absolute
+magnitudes in the 562 and 832 nm Zorro bandpasses.  The apparent
+magnitudes follow from information about the distance of the star.
+
+Applying the same assumption to WASP-4 itself, we compute its apparent
+magnitude in each of the Zorro filters, and finally derive the
+transformation from contrast to companion mass.
 

### 2020-01-20 src/contrast_to_masslimit.py
[WIP] add contrast_to_masslimit
to convert zorro contrasts to limits on the mass.

WIP because it's 5:30pm

@@ -1,0 +1,166 @@
+"""
+We would like to convert the contrast ratios obtained through the Zorro imaging
+to constraints on the masses of putative companions, and their separations from
+the host star.
+
+To do this, we followed the methodology of Montet+2014, and opted to employ the
+Baraffe+2003 models for substellar mass objects and the MESA isochrones for
+stellar mass objects (Choi et al XXXX).
+
+We assumed that the system age was 5 Gyr, so that companions would have fully
+contracted.
+
+Due to the narrow-band filters of the Zorro imager, we further assumed that all
+sources had blackbody spectra. While this is clearly false, we do not readily
+have access to the planetary and stellar atmosphere models needed for the
+correct calculation. However, the Baraffe+2003 and MESA models do report
+effective temperatures and bolometric luminosities.  We opt to use these
+theoretical quantities, combined with the measured Zorro filter wheels, to the
+absolute magnitudes in the 562 and 832 nm Zorro bandpasses.  The apparent
+magnitudes follow from information about the distance of the star.
+
+Applying the same assumption to WASP-4 itself, we compute its apparent
+magnitude in each of the Zorro filters, and finally derive the transformation
+from contrast to companion mass.
+"""
+
+##########
+# config #
+##########
+
+import os
+import pandas as pd, numpy as np
+from numpy import array as nparr
+
+from read_mist_model import ISO
+
+from astropy.modeling.models import BlackBody
+from astropy import units as u
+
+
+datadir = '../data/companion_isochrones/'
+
+def get_merged_companion_isochrone():
+
+    outpath = '../data/companion_isochrones/MIST_plus_Baraffe_merged.csv'
+
+    if not os.path.exists(outpath):
+
+        #
+        # Get Teff, L for the stellar models of MIST at 5 Gyr.
+        #
+
+        # MIST isochrones, v1.2 from the website
+        mistpath = os.path.join(
+            datadir, 'MIST_v1.2_feh_p0.00_afe_p0.0_vvcrit0.4_basic.iso'
+        )
+        iso = ISO(mistpath)
+
+        # 10**9.7 = 5.01 Gyr. I'm OK with not interpolating further here.
+        mist_age_ind = iso.age_index(9.7)
+        mist_logTeff = iso.isos[mist_age_ind]['log_Teff']
[…]

### 2020-01-22 src/contrast_to_masslimit.py
contrast_to_masslimit: now working
so we can go from the Zorro observations to useful limits on the mass!
turns out: <~ 200 Mjup at ~>27AU is the number.

@@ -29,14 +29,18 @@ from contrast to companion mass.
 ##########
 
 import os
-import pandas as pd, numpy as np
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
 from numpy import array as nparr
+from numpy import trapz
+from scipy.interpolate import interp1d
 
 from read_mist_model import ISO
 
 from astropy.modeling.models import BlackBody
 from astropy import units as u
+from astropy import constants as const
 
+from cdips.plotting import savefig
 
 datadir = '../data/companion_isochrones/'
 
@@ -73,7 +77,7 @@ def get_merged_companion_isochrone():
         # There is one point that overlaps: Mstar = 0.1Msun. For that point,
         # use the Baraffe model. 
         #
-        sel = (mist_df.mass < 0.3) & (mist_df.mass > 0.1)
+        sel = (mist_df.mass < 0.9) & (mist_df.mass > 0.1)
         mist_df = mist_df[sel]
 
         #
@@ -128,39 +132,168 @@ def abs_mag_in_zorro_bandpass(lum, teff, bandpass='562'):
     # by Howell in priv. comm.) 
 
     width = 100 # nanometers, around the bandpass middle
-    sel = np.abs(float(bandpass) - df.nm) < width
+    sel = np.abs(float(bandpass) - bpdf.nm) < width
 
     bpdf = bpdf[sel]
 
-    #FIXME: now create the blackbody flux objects, multiply by the transmission
-    #filter, and figure out how to convert the counts to ABSOLUTE magntudes.
+    #
+    # see /doc/20200121_blackbody_mag_derivn.pdf for relevant discussion of
+    # units and where the equations come from.
+    #
+
+    M_Xs = []
+    for temperature, luminosity in zip(teff*u.K, lum*u.Lsun):
+
+        bb = BlackBody(temperature=temperature)
+
+        wvlen = nparr(bpdf.nm)*u.nm
+        B_nu_vals = bb(wvlen)
+        B_lambda_vals = B_nu_vals * (const.c / wvlen**2)
+
+        T_lambda = nparr(bpdf.Transmission)
+
+        F_X = 4*np.pi*u.sr * trapz(B_lambda_vals * T_lambda, wvlen)
+
+        F = const.sigma_sb * temperature**4
+
+        # https://nssdc.gsfc.nasa.gov/planetary/factsheet/sunfact.html
+        M_bol_sun = 4.83
+        M_bol_star = (
[…]

### 2020-01-22 paper/ms.tex
paper minor

@@
 
-\documentclass[RNAAS]{aastex62}
+\documentclass[12pt,twocolumn,tighten]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 
+\renewcommand*{\sectionautorefname}{Section} %for \autoref
+\renewcommand*{\subsectionautorefname}{Section} %for \autoref
+
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
+%% Add "Submitted to " argument.
+\received{\today}
+\revised{---}
+\accepted{---}
+\submitjournal{AAS journals.}
+\shorttitle{WASP-4 is accelerating towards the Earth}
+
 \begin{document}
 
-% NOTE: may want to improve title
+% TODO: may want to improve title
 \title{WASP-4 is accelerating towards the Earth}
@@
 
-%% Note that RNAAS manuscripts DO NOT have abstracts.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
+\begin{abstract}
+  We recently noted that combining measurements from the Transiting
+  Exoplanet Survey Satellite (TESS) with ground-based transit
+  observations, the orbital period of WASP-4b seemed to be decreasing.
+  Possible explanations included tidal orbital decay, orbital
+  precession, and light-travel time effects.
+  In this work we present new observations of the radial velocity of
+  WASP-4, acquired with Keck-HIRES.
+  They show that the system is accelerating towards our line of sight,
+  at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm s}^{-1}\,{\rm
+  yr}^{-1}$.
+  This implies a period decrease of X.XX milliseconds per year,
+  compared to the observed value of Y.YY.
+  Combining the radial velocities with new speckle imaging from
+  Gemini-Zorro, we show that the acceleration is best explained by a
+  10-200$\,M_{\rm Jup}$ companion between 10 and 100$\,$AU.
+  We encourage further long-term monitoring of hot Jupiter radial
+  velocities; these observations are critical in the ongoing hunt
+  for tidal orbital decay.
+\end{abstract}
+
 \keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
 (1332), Transit timing variation method (1710)}
-
-%% Start the main body of the article. If no sections in the 
-%% research note leave the \section call blank to make the title.
 
@@
 These efforts are gradually beginning to yield fruit: by combining
-transit and occultation timing and radial velocity measurements, CITET
-Yee have shown that the data for the hot Jupiter WASP-12b are most
-compatible with orbital decay.
+transit and occultation timing and radial velocity measurements, CITE
+Yee et al{.} have shown that the data for the hot Jupiter WASP-12b are
+most compatible with orbital decay.
 
-This study highlights a point that, while obvious, has perhaps not
-yet received due attention.
-The point is that the hunt for orbital decay of hot Jupiters will be
-crippled without long-term radial velocity monitoring programs.
+The present study highlights a point that, while obvious, has perhaps
+not yet received due attention.  The point is that the hunt for
+orbital decay of hot Jupiters will be crippled without long-term
+radial velocity monitoring programs.
 The reason is simple: when the timescale of orbital period change far
-exceeds the observing baseline, the first
-deviation from constant periodicity is always quadratic in the transit
-number:
+exceeds the observing baseline, the first deviation from constant
+periodicity is always quadratic in the transit number:
 \begin{align}
@@
 
-The main example of this study is the hot Jupiter WASP-4b which has an
+The main focus of this study is the hot Jupiter WASP-4b, which has an
 orbital period that from transits appears to be decreasing by about 10
 milliseconds per year.
-After discovering the timing variations in the TESS data, we obtained
-N additional radial velocity measurements using Keck-HIRES,
-extending the RV baseline by Y years.
+After discovering the timing variations in the TESS data (CITE Bouma
+2019), we obtained N additional radial velocity measurements using
+Keck-HIRES, extending the RV baseline by Y years.
 Previously, the radial velocities were sparse and were consistent with
@@
 
+% FIXME: figure out correct sections
+\S~\ref{sec:observations} presents the new observations...
+\S~\ref{sec:timing} describes our timing analysis...
+\S~\ref{sec:implications} ...
+\S~\ref{sec:future}...
+
 \section{WASP-4's acceleration towards the Earth}
+\label{sec:observations}
 

### 2020-01-23 src/is_it_doppler_shift.py
minor

@@ -143,7 +143,7 @@ print('(dP/dt_Dopplerlimit)/(dP/dt_measured): {:.6g}'.format(dP_dt_limit.cgs.val
 print('(dP/dt_Dopplerquoted)/(dP/dt_measured): {:.6g}'.format(dP_dt_preferred.cgs.value/dP_dt_measured))
 
 
-# WASP-4b. 2019/09/11 LGB analysis, for RNAAS
+# WASP-4b. 2019/09/11 LGB analysis, for paper #2
 P = 1.3382324*u.day
 gamma_dot_value, gamma_dot_sig = -0.0422, 0.0028
 

### 2020-01-23 paper/ms.tex
wrote some words

@@
 \usepackage{hyperref}
+\usepackage{natbib}
 
@@
 
+\defcitealias{bouma_wasp4b_2019}{B19}
+
 % TODO: may want to improve title
-\title{WASP-4 is accelerating towards the Earth}
+\title{WASP-4 is accelerating towards the Earth: A Cautionary Tale in
+the Search for Tidal Orbital Decay}
+% "An Explanation for WASP-4's Decreasing Orbital Period".
+% "WASP-4's Decreasing Orbital Period: Solved."
 
@@
 milliseconds per year.
-After discovering the timing variations in the TESS data (CITE Bouma
-2019), we obtained N additional radial velocity measurements using
+After discovering the timing variations in the TESS data 
+\citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, we obtained N additional
+radial velocity measurements using
 Keck-HIRES, extending the RV baseline by Y years.
@@
 observed.
-
-% FIXME: figure out correct sections
-\S~\ref{sec:observations} presents the new observations...
-\S~\ref{sec:timing} describes our timing analysis...
-\S~\ref{sec:implications} ...
-\S~\ref{sec:future}...
-
-\section{WASP-4's acceleration towards the Earth}
-\label{sec:observations}
-
-\subsection{Trend due to stellar activity?}
-
-First, we considered whether the measured RV trend is correlated with
-WASP-4's emission in the Ca II H \& K lines, as quantified with the
-chromospheric $S$-index (CITE: Wright 2004).  We only analyzed the
-HIRES velocities for this step, since they are the main source of
-signal in our analysis.  First, we subtracted the orbital solution
-from the Keck-HIRES velocities.  Then, following CITE Bryan 2016 and
-2019, we calculated the Spearman rank correlation coefficient between
-the $S$-index and the orbit-subtracted velocities.  We found a
-correlation coefficient of 0.16, which weakly hints at a correlation,
-but is not statistically significant (the corresponding $p$-value is
-0.65).  We therefore do not think that the linear trend can be
-attributed to stellar activity.
-
-\subsection{Zorro imaging}
-
-Convinced that the linear trend warranted closer examination, we
-performed speckle imaging using Zorro at Gemini to attempt to detect
-any nearby companions.  To our knowledge, this is the first high
-resolution imaging that has been performed for WASP-4 (FIXME: IS THIS
-TRUE?).
-
-We observed WASP-4 twice with Gemini-South, 11/12 Sept. (8 sets of
-1000 frames each) and 28/29 Sept. (7 sets of 1000 frames each). The
-second set had better seeing and produced a slightly better result.
-The spatial scale is from the diffraction limit of Gemini at the
-wavelength observed to about 1.25'' plus a bit more.
-
-% TODO: might Howell suggest a paragraph for how the contrast ratios
-% were derived?
-To derive contrast ratios from these images, we did XXXXX.
-
-We would like to convert the contrast ratios obtained through the
-Zorro imaging to constraints on the masses of putative companions, and
-their separations from the host star.
-
-To do this, we followed the methodology of CITE Montet+2014, and opted
-to employ the CITE Baraffe+2003 models for substellar mass objects and
-the MESA isochrones for stellar mass objects (CITE Choi et al XXXX).
-In the intermediate regime, near 100 Mjup, we interpolated between the
-two sets of models.
-
-We assumed that companions would have contracted, and asserted the
-system age to be 5 Gyr.
-
-Due to the narrow-band filters of the Zorro imager, we further assumed
-that all sources had blackbody spectra. While this is clearly false,
-we do not readily have access to the planetary and stellar atmosphere
-models needed for the correct calculation. However, the CITE
-Baraffe+2003 and MESA models do report effective temperatures and
-bolometric luminosities.  We opt to use these theoretical quantities,
-combined with the measured Zorro filter wheels, to the absolute
-magnitudes in the 562 and 832 nm Zorro bandpasses.  The apparent
-magnitudes follow from information about the distance of the star.
-
-Applying the same assumption to WASP-4 itself, we compute its apparent
-magnitude in each of the Zorro filters, and finally derive the
-transformation from contrast to companion mass.
-
-\subsection{Constraints on companion masses and semi-major axes}
-
-Given a linear radial velocity trend, we can place limits on the mass
-and semi-major axis of any planetary or stellar companions in the
-system.
-The constraints on the two parameters are degenerate: $\dot{\gamma}
-\propto M_{\rm p} a^{-2}$.
-Nonetheless, the additional information from the Zorro images can help
-limit the available parameter space by setting an outer limit on the
-semimajor axis, and a maximum brightness (and thereby mass) of any
-putative companions.
-
-The general procedure we use to derive these constraints has been developed by
-\citet{wright_linear_trends_2007}, \citet{montet_trends_2014},
-\citet{knutson_friends_2014}, \citet{bryan_statistics_2016,bryan_excess_2019},
-and others.
-For this study, we closely follow the procedure of \citet{bryan_excess_2019}.
-
-We begin by defining a $50\times50$ grid in true planetary mass and semimajor
-axis, evenly spaced from 1 to 300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.
-
-We then consider the possibility that an additional companion in any particular
-cell can explain the observed linear trend.
-
-In each cell, we consider 500 hypothetical companions.
-
-We assign each companion a mass and semimajor axis from log-uniform
-distributions within the grid cell. We draw the inclination from a uniform
-distribution in $\cos i$.  The eccentricity is drawn from
-\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution ($a=1.12$,
-$b=3.09$).
-
-For each simulated companion, we then draw a sample from the converged chains
-of our initial model of WASP-4b, plus its linear trend. We subtract
-the planet's orbital solution, leaving the linear trend.
-
-Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulated outer companion,
-and the fixed instrument offsets and jitters from the MCMC chains, we then
-perform a maximum likelihood fit for the time and argument of periastron of the
-outer simulated companion.
-
-We then convert the resulting $50\times50\times500$ cube of best-fit
-log-likelihood values to probabilities, and average over the samples in
-each grid cell to derive a probability distribution in mass and semi-major
-axis.
-
-The distribution is shown in {\bf Figure~X}.
-
-\section{Do other hot Jupiters show the Romer delay?}
-yes, HAT-P-7b. but omg utc to tdb.
-
-\section{What is the expected Pdot distribution?}
-
-SUMMARY OF LITERATURE
-
-This is a major challenge in the hunt for orbital decay.  The main
-example of this study is the hot Jupiter WASP-4b, which we recently
-showed transited $\approx$82 seconds early for the TESS spacecraft
-assuming a constant orbital period \citep{bouma_wasp-4b_2019}.  Our
-timing analysis included data from peer-reviewed literature for which
-the times were measured from a single transit, and for which the
-midpoint was allowed to be a free parameter. We also required the time
-system to be clearly documented\footnote{In other words, we needed to
-know whether any heliocentric or barycentric corrections had been
-performed, and whether the absolute time system was UTC or TDB.}. The
-transit times we analyzed spanned 2007 to 2019
-\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,dragomir_terms_2011,sanchis-ojeda_starspots_2011,nikolov_wasp-4b_2012,hoyer_tramos_2013,ranjan_atmospheric_2014,huitson_gemini_2017}.
-The combined timing data were best fit by an ephemeris with a constant
-negative period derivative; our best-fit decay rate was $\dot{P} =
--12.6 \pm 1.2$ milliseconds per year.  Our interpretation was that the
-apparent period change could be caused by any of three scenarios: a
-decaying orbit, a precessing orbit, and an orbit being gravitationally
-perturbed by an outer companion.
 
@@
 
-One line of follow-up needed to confirm and understand the timing
-variation was additional radial velocity observations.  The longest RV
-baseline previously available was six observations acquired by
-\citet{knutson_friends_2014} and the CKS team on Keck HIRES from
-2010 to 2013.  This past season we acquired four additional
-observations with Keck HIRES. Even before fitting out the hot Jupiter,
-the residuals showed a strong linear trend.
+% FIXME: figure out correct sections
+\S~\ref{sec:observations} presents the new observations...
+\S~\ref{sec:analysis} describes our analysis...
+\S~\ref{sec:discussion} discusses...
+\S~\ref{sec:conclusion} gives conclusions...
 
-For completeness, in our analysis we included CORALIE measurements
+\section{Observations}
+\label{sec:observations}
+
+\subsection{Transits}
+
+{\bf Table~N} lists the transit times we used in our analysis.  We
+include data from the peer-reviewed literature for which (a) the
+analysis was based on observations of a single transit, (b) the
+midpoint was fitted as a free parameter, and (c) the time system
+specified both the absolute reference (TDB or UTC) and also whether
+any barycentric or heliocentric corrections had been performed.
+
+The majority of times are identical to those we collected in
+\citetalias{bouma_wasp4b_2019}.  Twenty-two new times reported by
+\citet{southworth_transit_2019} are included.  These transits were
+observed from the 3.58m NTT and Danish 1.54m telescopes at La Silla,
+and the SAAO 1.0m telescope.
+
+% FIXME : verify with Michael Gillon and Phil Evans that this is true.
+A number of additional times were also recently reported by
+\citet{baluev_2019}, based on a homogeneous analysis of archival
+ground-based observations.  We included their ``high quality'' transit
+times with provenance from the TRAPPIST and El Sauce telescopes, for
+which we verified with the original observers that appropriate
+time-system corrections had been performed (M.~Gillon, P.~Evans,
+priv{.} comm{.}).
+
+The four available occultation data points tabulated by
+\citetalias{bouma_wasp4b_2019} are sparse and have neglegible
+statistical value due to their large uncertainties, and we forgo
+their use in this analysis.
+
+\subsection{Radial velocities}
+
+% TODO: Details of reduction from Howard or Knutson
+After identifying the period decrease in
+\citetalias{bouma_wasp4b_2019}, we acquired four additional radial
+velocity measurements with the Keck High Resolution Echelle
+Spectrometer (HIRES, \citealt{vogt_hires_1994}).  Our observations
+were acquired under the purview of the California Planet Survey
+\citep{howard_cps_2010}.  The spectra were reduced using {\bf software
+X}.  Previously, the HIRES data-points spanned 2010 to 2013
+\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
+observing baseline to nine years.
+
+%FIXME: I should ask Didier if he has any new CORALIE observations...
+The complete set of radial velocity observations is given in {\bf
+Table~M}.  Along with the 2010-2019 HIRES observations, there are also
+earlier measurements from CORALIE and HARPS.  Following
+\citetalias{bouma_wasp4b_2019}, we included the CORALIE measurements
 from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
-using the homogeneous velocities reported by the latter authors.  We
-included the HARPS values reported by \citet{pont_determining_2011},
-which are identical to those from \citet{husnoo_observational_2012}.
+using the homogeneous radial velocities calculated by the latter
+authors. We included the HARPS values reported by
+\citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
 We omitted the HARPS data points taken over three nights by
 \citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
-because they were calculated using a different pipeline than the
-longer-baseline \citeauthor{pont_determining_2011} measurements, and
-necessarily inclusion of an extra offset term would nullify their
-statistical value.
+because they have a systematic offset from the remaining datasets.
 
-We then fitted a single Keplerian orbit, plus instrument offsets,
-jitters, and a linear trend
+\subsection{Speckle imaging}
+
+A cursory analysis of the new HIRES observations led to our detection
+of a linear trend in the residuals after fitting out the orbit of
+WASP-4b.  This prompted us to acquire speckle imaging using Zorro at
+Gemini-South \citep[see][, and the instrument
+web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
+Zorro (and its counterpart, 'Alopeke, on Gemini-North) are
+dual-channel speckle interferometers that can be used to observe in
+narrow-band filters centered at 562$\,$nm and 832$\,$nm.
+
+We observed WASP-4 twice, on the night of September 11-12 (8 sets of
+1000 frames each) and also on the night of September 28-29 (7 sets of
+1000 frames each). The second night had better seeing and produced a
+slightly better result, which we opt to use for our analysis.
+
+% TODO: might Howell suggest a paragraph for how the contrast ratios
+% were derived?
+{\bf We reduced the speckle images to contrast curves by ...}
+
+\section{Analysis}
+\label{sec:analysis}
+
+\subsection{Transits}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f1.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
+  The vertical axis shows the observed times minus the calculated
+  times assuming a constant orbital period .  Darker points correspond
+  to more precise data.  The quadratic ephemeris and its $1\sigma$
+  uncertainties are shown in blue.  The binned TESS point (red square)
+  is the weighted average of 18 TESS transits and is for display
+  purposes only.  The models were fitted to all of the individual
+  transit times.
+  \label{fig:times}
+	}
+\end{figure*}
+
+We considered two possible models for the observed transit times.
+The first model assumes a constant orbital period on a circular orbit:
+\begin{align}
+  t_{\rm tra}(E) &= t_0 + PE,\\
+\end{align}
+where $E$ is the epoch number and $t_0$ is a reference epoch.
+The second model assumes the period changes at a steady rate:
+\begin{align}
+  t_{\rm tra}(E) &=
+    t_0 + PE +
+    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
+\end{align}
+The free parameters are the reference epoch $t_0$, the period at
+the reference epoch, and the period derivative, ${\rm d}P/{\rm d}t =
+(1/P) {\rm d}P/{\rm d}E$.
+We defined the epoch numbers such that $E=0$ is near the weighted
+average of the observed times.  This helps to reduce the covariance
+between $t_0$ and $P$.
+A third possible model that we did not consider for reasons that will
+become apparent is a precessing, eccentric orbit \citep[{\it
+e.g.},][]{gimenez_revision_1995,patra_2017}.
+
+We fitted each of the two models by assuming a Gaussian likelihood and
+sampling over the posterior probability distributions.  We sampled the
+posterior using the algorithm proposed by
+\citet{goodman_ensemble_2010} and implemented by
+\citet{foreman-mackey_emcee_2013} in \texttt{emcee}.  The prior for
+the quadratic model allowed the period derivative to have any sign.
+
+Figure~\ref{fig:times} shows the observed transit times, minus the
+best-fit constant period model. 
+The best-fitting constant-period model has $\chi^2 = XXX$ and YY
+degrees of freedom.  The best-fitting
+quadratic model has $\chi^2 = YYY$ and ZZ degrees of freedom. 
+The difference in the BIC \citep{kass_bayes_1995} between the linear
+and quadratic and models is $\Delta {\rm BIC} = ZZ.Z$, strongly
+favoring the quadratic case.
+
+The best-fit period derivative in the quadratic model is
+\begin{equation}
+\dot{P}
+  = - (X.XX \pm 0.XX)\times 10^{-10}
+  = - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}.
+  \label{eq:dP_dt_obs}
+\end{equation}
+This agrees with the value reported by \citet{southworth_transit_2019}
+after their including an additional 22 archival transit measurements
+($\dot{P} = - 9.2 \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$).
+It is smaller than the rate of period decrease found by
+\citetalias{bouma_wasp4b_2019} ($- 12.6 \pm 1.2~{\rm ms}\,{\rm
+yr}^{-1}$)., presumably because of missing coverage filled in by
+\citeauthor{southworth_transit_2019}'s observations.
+The other best-fit transit timing model parameters are reported in
+{\bf Table~X}.
+
+\subsection{Radial Velocities: WASP-4's acceleration towards the Earth}
+
+%FIXME: is the purple here from my Pdot, or Southworth's?
+%FIXME: perhaps include the WASP-4b phase-fold here too??
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f2.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Radial velocity observations of WASP-4.} The best-fit Keplerian orbit
+    of WASP-4b has been subtracted.  The linear trend inferred from the
+    RV data is shown with a black line, with $1\sigma$ errors in gray.
+    The trend that would be needed to produce the period decrease seen
+    in transits ($\dot{P}= - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}$) is
+    indicated with the purple dotted line.  The four new RV
+    measurements presented in this work increased the significance of
+    the slope from $\approx$2$\sigma$ to 15$\sigma$.
+	\label{fig:rvs}
+  \vspace{-0.3cm}
+	}
+\end{figure}
+
+%FIXME: this entire paragraph copies Bouma+19 too heavily.
+We began by fitting a single Keplerian orbit, plus instrument offsets,
+jitters, and a long-term trend
 \citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
 on the period and time of inferior conjunction using the values from
-\citet{bouma_wasp-4b_2019}, and fixed the eccentricity to zero,
-consistent with results from \citet{beerer_secondary_2011},
-\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The
-remaining free parameters were the velocity semi-amplitude, the
-instrument zero-points, the instrument jitters (an additive white
-noise term for each instrument), and optionally a linear acceleration
-term ($\dot{v_{\rm r}}$).
+Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed the eccentricity
+to zero, per the results of \citet{beerer_secondary_2011},
+\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The free
+parameters were the velocity semi-amplitude, the instrument
+zero-points, the instrument jitters (an additive white noise term for
+each instrument), linear ($\dot{v_{\rm r}}$), and optionally
+second-order ($\ddot{v_{\rm r}}$) acceleration terms.
 
-%FoHJ: gammadot = -0.0099^{+0.0052}_{-0.0054}
-The AIC and BIC strongly favored a model with a linear trend ($\Delta
-{\rm BIC}=73$ compared to a model without the linear trend).
-Figure~\ref{fig:rv_o_minus_c} shows the best-fitting model, with the
-orbit of WASP-4b subtracted.  The best-fit radial velocity derivative,
-$\dot{v}_{\rm r}$, is
+%FIXME: this is a janky approach if so??
+We found that the best-fitting model with both linear and quadratic
+radial velocity terms was marginally preferred (by $\Delta
+\mathrm{BIC} = 5.8$) over the best-fitting model with only a linear
+term.  Regardless, for consistency with \citet{knutson_friends_2014},
+who fixed the quadratic component of the long-term trend to zero, in
+Figure~\ref{fig:rvs} we show best-fitting models for the linear-trend
+case.
+
+WASP-4 was found to be accelerating towards our line of sight at high
+confidence,
 \begin{equation}
-  \dot{v}_{\rm r} = -0.0422^{+0.0028}_{-0.0027}\,{\rm m\,s^{-1}\,day^{-1}}.
+  \dot{v}_{\rm r} =
+     -0.0422^{+0.0028}_{-0.0027}
+     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}.
 \end{equation}
-This acceleration towards our line of sight is 4.3 times faster than
-the $\approx$$2\sigma$ trend \citet{knutson_friends_2014}
-found from the shorter baseline.  The radial velocity values are
-available in the data-behind-the-figure online version of this {\it
-Note}.
+For comparison, before our recent observing run, 
+$\dot{v}_{\rm r}$ was thought to be about five times smaller, and was
+only marginally significant
+\citep{knutson_friends_2014,bouma_wasp4b_2019}.
 
-Under the assumption of constant acceleration, $\dot{P} = \dot{v}_{\rm
-r} P / c$, the implied period decrease that should be seen in
-transits is roughly $5.9 \pm 0.4$ milliseconds per year.  All the
+The system's acceleration towards our line of sight causes a decrease
+in the apparent orbital period (the R{\o}mer delay).  The period
+derivative expected from radial velocities is
+\begin{align}
+  \dot{P}_{\rm RV} &= \frac{\dot{v}_{\rm r} P}{c} \\
+  \dot{P}_{\rm RV} &= -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
+\end{align}
+In other words, the majority of the period decrease observed in
+transits ($\dot{P} = - X.X \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$) seems to
+be explained by the acceleration of the host star.
+% FIXME
+All the
 available TTV data show a period decrease of $\approx 8-12$
@@
 
+An important consideration is whether the measured RV trend is at all
+correlated with stellar activity.  We investigate this by analyzing
+WASP-4's emission in the Ca II H \& K lines, as quantified with the
+chromospheric $S$-index \citep{wright_chromospheric_2004}.  We only
+examined the HIRES velocities for this step, since they are the main
+source of signal for our analysis.  First, we subtracted the orbital
+solution from the Keck-HIRES velocities.  Then, following
+\citet{bryan_statistics_2016,bryan_excess_2019}, we calculated the
+Spearman rank correlation coefficient between the $S$-index and the
+orbit-subtracted velocities.  We found a correlation coefficient of
+0.16. Though suggestive, it is not statistically significant (the
+corresponding $p$-value is 0.65).  Furthermore, inspection of the
+$S$-index timeseries showed no secular or sinusoidal trends, as would
+be expected if we were observing a long-term magnetic activity cycle.
+The $S$-index values are included in {\bf Table X}.  We conclude that
+it would be highly unlikely for the linear trend to be caused by
+stellar activity.
+
+\subsection{Constraints on companion masses and semi-major axes}
+
+%TODO: describe 3b better
+\begin{figure*}[!t]
+	\gridline{
+		\fig{f3a.png}{0.33\textwidth}{}
+		\fig{f3b.pdf} {0.5\textwidth}{}
+	}
+	\vspace{-0.7cm}
+    \caption{
+      {\it Left:\,} 
+      Speckle image reconstructed from 1000 40 millisecond frames in an 832 nm bandpass, and acquired on September 28, 2019.
+      The image scale is $2.46''\times2.46''$.
+      {\it Right:\,} 
+      Contrast curve derived from point-source injection-recovery
+      experiments.
+    }
+    \label{fig:zorro}
+\end{figure*}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Masses and semi-major axes of companions that meet
+    requirements of both the radial velocity and the speckle imaging.}
+    The likelihood inferred from radial velocities is shown in
+    grayscale, and the region excluded from the imaging is shown with
+    a cross-hatch pattern.
+	\label{fig:mass_sma}
+  \vspace{-0.3cm}
+	}
+\end{figure}
+
+Given a linear radial velocity trend, we can place lower-limits on the
+mass and semi-major axis of additional bodies in the system.  The
+constraints on the two parameters are degenerate: $\dot{\gamma}
+\propto M_{\rm p} a^{-2}$.  Nonetheless, the Zorro images and contrast
+curves can limit the available parameter space by setting an upper
+limit on the semimajor axis, and a maximum brightness (and thereby
+mass) of any putative companions.
+
+The general procedure we use to derive constraints on possible
+companion masses and semi-major axes has been developed by
+\citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
+\citet{montet_trends_2014}, \citet{knutson_friends_2014},
+\citet{bryan_statistics_2016,bryan_excess_2019}, and others.
+
+\paragraph{Speckle imaging constraints}
+
+First, we would like to convert the contrast ratios obtained through
+the Zorro imaging (Figure~\ref{fig:zorro}) to limits on the masses of
+putative companions and their separations from the host star.
+
+To do this, we followed \citet{montet_trends_2014}, and opted to
+employ the \citet{baraffe_evolutionary_2003} models for substellar
+mass objects and the MIST isochrones for stellar mass objects
+\citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
+We assumed that the system age was 5 Gyr, so that companions would
+have fully contracted.
+
+Due to the custom filters of the Zorro imager, and corresponding lack
+of synthetic photometry, we further assumed that all sources had
+blackbody spectra. While this is clearly false, we do not readily have
+access to the planetary and stellar atmosphere models needed for the
+consistent calculation with the \texttt{COND03} and \texttt{MESA}
+models.  We therefore adopted the effective temperatures and
+bolometric luminosities from the \citet{baraffe_evolutionary_2003} and
+MIST isochrones.  Using these theoretical quantities and the empircal
+Zorro bandpass functions, we calculated absolute magnitudes in the 562
+and 832 nm Zorro bandpasses for stellar and planetary mass companions.
+Applying the same calculation to WASP-4 itself using the effective
+temperature and bolometric luminosity from
+\citetalias{bouma_wasp4b_2019}, we derived the transformation from
+contrast ratio to companion mass.
+The resulting limits are shown as the cross-hatched region in
+Figure~\ref{fig:mass_sma}.
+
+\paragraph{Radial velocity constraints}
+
+To derive constraints on possible companion masses and separations
+from the radial velocities, we followed the procedure of
+\citet{bryan_excess_2019}.  We began by defining a $50\times50$ grid
+in true planetary mass and semimajor axis, evenly spaced from 1 to
+300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the
+possibility that an additional companion in any particular cell could
+explain the observed linear trend.  In each cell, we simulated 500
+hypothetical companions.
+
+We assigned each companion a mass and semimajor axis from log-uniform
+distributions within the grid cell. We drew the inclination from a
+uniform distribution in $\cos i$.  The eccentricity was drawn from
+\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
+($a=1.12$, $b=3.09$).  For each simulated companion, we then drew a
+sample from the converged chains of our initial model of WASP-4b, plus
+its linear trend. We subtracted the planet's orbital solution, leaving
+the linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each
+simulated outer companion, and the fixed instrument offsets and
+jitters from the MCMC chains, we then performed a maximum likelihood
+fit for the time and argument of periastron of the outer simulated
+companion.  We converted the resulting $50\times50\times500$ cube of
+best-fit log-likelihood values to probabilities, and averaged over the
+samples in each grid cell to derive a probability distribution in mass
+and semi-major axis.
+
+The result is shown as grayscale background in
+Figure~\ref{fig:mass_sma}.
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Implications for WASP-4}
+Previous possibilities suggested to explain the period decrease of
+WASP-4b included tidal orbital decay, orbital precession, and
+light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
+velocity measurements have now shown the least exotic
+option---light-travel time effects---is also the most likely
+explanation.  Transits show the obital period to decrease by $YY.Y \pm
+Z.Z~{\rm ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed
+in radial velocities implies an expected period decrease of $YY.Y \pm
+Z.Z~{\rm ms}\,{\rm yr}^{-1}$.  
 Though the quantitative agreement is not perfect, Occam's razor would
@@
 explanation for the apparent decrease of WASP-4b's orbital period.
-While further radial velocity observations of the system should help
-in determining the mass and semi-major axis of the companion, it seems
-unlikely that additional transit observations will yield near-term
-constraints on orbital decay.
+% While further radial velocity observations of the system should help
+% in determining the mass and semi-major axis of the companion, it seems
+% unlikely that additional transit observations will yield near-term
+% constraints on orbital decay.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% RNAAS: you get one figure.
+The corresponding requirements for the companion causing the
+acceleration are that it is likely either a brown-dwarf or low mass
+star, and that it is between $10$-$100\,{\rm AU}$ from the host star
+(Figure~\ref{fig:mass_sma}).
+Given such a mass, this companion could at one time have influenced
+the orbital evolution of the inner giant.
+Further radial velocity monitoring should eventually reveal the
+orbital parameters and minimum mass of the companion.
 
-\begin{figure}
+\subsection{How many other hot Jupiters will show the R\'omer delay?}
+
+We identified WASP-4b's decreasing orbital period as part of a search
+for tidal orbital decay.  However, over half of hot Jupiters have
+companions outside of $1\,{\rm AU}$ with super-Jovian masses
+\citep{knutson_friends_2014}.  Line-of-sight accelerations are
+therefore relatively common in hot Jupiter systems.
+
+To evaluate the importance of these effects in future transit timing
+searches, we collected the linear trends reported by
+\citet{knutson_friends_2014}, and computed the expected orbital period
+changes $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
+systems.
+The results are given in {\bf Table~X}, and visualized for hot
+Jupiters with significant linear trends in Figure~\ref{fig:pdot_pop}.
+
+\begin{figure}[t]
 	\begin{center}
-		\includegraphics{rv_fit.pdf}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f4.pdf}
 	\end{center}
-    \vspace{-0.8cm}
+	\vspace{-0.7cm}
 	\caption{
-      {\bf Radial velocity observations of WASP-4.}
-      The orbit of WASP-4b has been subtracted.  The best-fit linear
-      trend from the RVs is shown with the black line, with $1\sigma$
-      errors in gray. The trend needed to produce the period decrease
-      of 12.6 milliseconds per year seen by \citet{bouma_wasp-4b_2019}
-      in transits is the purple dotted line in the bottom left.
-     \label{fig:rv_o_minus_c}
+    {\bf ...}
+	\label{fig:pdot_pop}
+  \vspace{-0.3cm}
 	}
 \end{figure}
+
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 Services.
+%
+Based on observations obtained at the Gemini Observatory, which is
+operated by the Association of Universities for Research in Astronomy,
+Inc., under a cooperative agreement with the NSF on behalf of the
+Gemini partnership: the National Science Foundation (United States),
+National Research Council (Canada), CONICYT (Chile), Ministerio de
+Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
+Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
+and Korea Astronomy and Space Science Institute (Republic of Korea).
+%
+Observations in the paper made use of the High-Resolution Imaging
+instrument Zorro at Gemini-South. Zorro was funded by the NASA
+Exoplanet Exploration Program and built at the NASA Ames Research
+Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
+Quigley.
 %
@@
 \software{
+  \texttt{astrobase} \citep{bhatti_astrobase_2018},
 	\texttt{astroplan} \citep{astroplan2018},
-  \texttt{radvel} \citep{fulton_radvel_2018}
-  %FIXME numerical python packages
+  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
+  \texttt{astroquery} \citep{astroquery_2018},
+  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
+  \texttt{corner} \citep{corner_2016},
+  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
+  \texttt{IPython} \citep{perez_2007},
+  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
+	\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
+  \texttt{numpy} \citep{walt_numpy_2011}, 
+  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
+  \texttt{radvel} \citep{fulton_radvel_2018},
+  % \texttt{scikit-learn} \citep{scikit-learn},
+  \texttt{scipy} \citep{jones_scipy_2001}.
 }
+
+%\clearpage
+%\newpage
+\input{model_fit_table.tex}
 

### 2020-01-23 paper/collect.sh

@@ -1,0 +1,5 @@
+cp ../results/O_minus_C_includeallpoints_onlytransits.pdf f1.pdf
+cp ../results/20190911_rv_fit.pdf f2.pdf
+cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png f3a.png
+cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832.pdf f3b.pdf
+cp ../results/mass_semimaj_constraints.pdf f4.pdf

### 2020-01-24 src/get_knutson_population_pdots.py
update get_knutson_population_pdots

@@ -18,10 +18,14 @@ from astrobase.services.tesslightcurves import (
 
 def main():
 
-    fdf = calculate_and_format_knutson_pdot_table()
-    tdf = check_tess_visibility_and_availability(fdf)
+    k14_table = 1
+    tess_vis = 0
 
-    import IPython; IPython.embed()
+    if k14_table:
+        fdf = calculate_and_format_knutson_pdot_table()
+    if tess_vis:
+        tdf = check_tess_visibility_and_availability(fdf)
+
 
 def check_tess_visibility_and_availability(fdf):
 
@@ -60,12 +64,32 @@ def check_tess_visibility_and_availability(fdf):
     return fdf
 
 
-
-
 def calculate_and_format_knutson_pdot_table():
 
     sig_df = calculate_pdots(signifiant_trends=1)
     limit_df = calculate_pdots(signifiant_trends=0)
+
+    #
+    # update WASP-4 -- add it to significant table!
+    #
+    limit_df = limit_df.drop(22, axis=0)
+
+    w4_dict = {
+        'planet': 'WASP-4 b',
+        'gammadot': -0.0422,
+        'gammadot_pluserr': 0.0028,
+        'gammadot_minuserr': 0.0027,
+        'comment': 'NaN',
+        'pl_name': 'WASP-4 b',
+        'pl_orbper': 1.338231466,
+        'Pdot': -5.94,
+        'Pdot_upper_limit': -5.94+0.39,
+        'Pdot_lower_limit': -5.94-0.39,
+        'Pdot_perr': 0.39,
+        'Pdot_merr': 0.39
+    }
+    w4_df = pd.DataFrame(w4_dict, index=[0])
+    sig_df = pd.concat((sig_df, w4_df))
 
     sig_df['abs_Pdot'] = np.abs(sig_df.Pdot)
     sig_df['K14_significant'] = True
@@ -80,6 +104,7 @@ def calculate_and_format_knutson_pdot_table():
 
     return fdf
 
+
 def calculate_pdots(signifiant_trends=1):
 
     if signifiant_trends:
@@ -93,8 +118,8 @@ def calculate_pdots(signifiant_trends=1):
     gamma_dot_value = nparr(df.gammadot) * (u.m/u.s)/u.day
     gamma_dot_perr = nparr(df.gammadot_pluserr) * (u.m/u.s)/u.day
[…]

### 2020-01-24 src/plot_k14_pdot.py
add plot_k14_pdot

@@ -1,0 +1,82 @@
+"""
+Plot significant period derivatives expected from Knutson+14.
+"""
+
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from numpy import array as nparr
+import matplotlib.patheffects as pe
+
+#
+# manually added WASP-4 to the output
+# \dot{P}_{\rm RV} &= -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
+#
+
+df = pd.read_csv('../results/knutson_all_pdots.csv', sep=';')
+sel = (df.K14_significant) & (df.Pdot != 0)
+df = df[sel]
+
+savpath = '../results/k14_pdot.png'
+fig, ax = plt.subplots(figsize=(4,3))
+
+yval = np.arange(0, len(df), 1) + 0.1
+xval = nparr(df.Pdot)
+x_perr = nparr(df.Pdot_perr)
+x_merr = nparr(df.Pdot_merr)
+dotlabels = nparr(df.planet)
+
+ax.errorbar(xval, yval, xerr=np.vstack([x_merr, x_perr]),
+            fmt='.k', ecolor='black', zorder=2, alpha=1, mew=1,
+            markersize=2.5, capsize=3)
+
+ix = 0
+for _x, _y, _l in zip(xval+x_perr+2, yval, dotlabels):
+
+    if 'WASP-4' in _l:
+        c = 'C0'
+        ax.errorbar(xval[ix], yval[ix],
+                    xerr=np.vstack([x_merr[ix], x_perr[ix]]),
+                    fmt='.', color=c, ecolor=c, zorder=3, alpha=1,
+                    mew=1, markersize=2.5, capsize=3)
+
+    else:
+        c = 'k'
+
+    ax.text(
+        _x, _y, _l.replace(' ','$\,$'),
+        fontsize=7.5, ha='left', va='center',
+        path_effects=[pe.withStroke(linewidth=2, foreground="white")],
+        color=c
+    )
+
+    ix += 1
+
+# axes etc
+xlabel = 'Expected d$P$/d$t$ from RV [millisecond/year]'
+ax.set_xlabel(xlabel)
+
+xlim = ax.get_xlim()
+xmax = max(np.abs(xlim))
+ax.set_xlim([-xmax, xmax])
+
+ax.set_ylim([-1, len(df)])
+ylim = ax.get_ylim()
+ax.set_ylim((min(ylim), max(ylim)))
[…]

### 2020-01-24 paper/ms.tex
update paper

@@
 % TODO: may want to improve title
-\title{WASP-4 is accelerating towards the Earth: A Cautionary Tale in
-the Search for Tidal Orbital Decay}
+\title{WASP-4 is Accelerating Towards the Earth}
+% ": A Cautionary Tale in the Search for Tidal Orbital Decay"
 % "An Explanation for WASP-4's Decreasing Orbital Period".
@@
 \begin{abstract}
-  We recently noted that combining measurements from the Transiting
-  Exoplanet Survey Satellite (TESS) with ground-based transit
-  observations, the orbital period of WASP-4b seemed to be decreasing.
+  We recently reported that space and ground-based transit observations
+  showed the orbital period of WASP-4b to be decreasing.
   Possible explanations included tidal orbital decay, orbital
@@
   WASP-4, acquired with Keck-HIRES.
-  They show that the system is accelerating towards our line of sight,
-  at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm s}^{-1}\,{\rm
-  yr}^{-1}$.
+  The new data show that the system is accelerating towards our line
+  of sight, at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm
+  s}^{-1}\,{\rm yr}^{-1}$.
   This implies a period decrease of X.XX milliseconds per year,
@@
   Combining the radial velocities with new speckle imaging from
-  Gemini-Zorro, we show that the acceleration is best explained by a
-  10-200$\,M_{\rm Jup}$ companion between 10 and 100$\,$AU.
-  We encourage further long-term monitoring of hot Jupiter radial
-  velocities; these observations are critical in the ongoing hunt
-  for tidal orbital decay.
+  Gemini-Zorro, we find that the WASP-4 system likely has a
+  10-200$\,M_{\rm Jup}$ companion with semi-major axis between 10 and
+  100$\,$AU.
+  As observing baselines continue to extend, long-term tracking of hot
+  Jupiter radial velocities will be necessary to distinguish tidal
+  orbital decay from the R{\o}mer delay.
 \end{abstract}
@@
 orbital decay of hot Jupiters will be crippled without long-term
-radial velocity monitoring programs.
-The reason is simple: when the timescale of orbital period change far
-exceeds the observing baseline, the first deviation from constant
-periodicity is always quadratic in the transit number:
-\begin{align}
-  t_{\rm tra} = t_0 + P E + \frac{1}{2}\frac{dP}{dE}E^2,
-\end{align}
-where
-\begin{align}
-  \frac{dP}{dE} = P \frac{dP}{dt},
-\end{align}
-$E$ is the transit number, $P$ is the orbital period, $t_0$ is the
-reference epoch, and $t_{\rm tra}$ is the transt mid-time.
+radial velocity monitoring programs.  The reason is simple: when the
+timescale of orbital period change far exceeds the observing baseline,
+the first deviation from constant periodicity is always quadratic in
+the transit number.
 
-There are ample reasons to expect orbital decay to be common
-(CITE).
-However there are reasons to expect the R\'omer delay to perhaps be
-even {\it more} common (CITE Knutson 14, Bryan 16).
-Both manifest to first order identically in transit times (up to the
-sign of the line-of-sight acceleration) -- as a non-zero period
-derivative.
+There are ample reasons to expect orbital decay to be common (CITE).
+However there are reasons to expect the R{\o}mer delay to perhaps be
+even {\it more} common (CITE Knutson 14, Bryan 16).  Both manifest to
+first order identically in transit times (up to the sign of the
+line-of-sight acceleration) -- as a non-zero period derivative.
 
 Specifically, \citet{knutson_friends_2014} showed that X\% of hot
-Jupiters show some form of radial acceleration in radial
-velocity time-series with baselines of YY years.
-Assuming a separation distribution of (WHATEVER), this implies that
-ZZ\% of hot Jupiters will {\it always} show a changing orbital period,
-over timescale of say 10 years.
-
-This is specific to hot Jupiters. The abundance of giant planets
-exterior to Super-Earth sized systems is also high (CITE Bryan 2019).
+Jupiters show some form of radial acceleration in radial velocity
+time-series with baselines of YY years.  Assuming a separation
+distribution of (WHATEVER), this implies that ZZ\% of hot Jupiters
+will {\it always} show a changing orbital period, over timescale of
+say 10 years.
+The abundance of giant planets
+exterior to Super-Earth sized systems is also high (CITE Bryan 2019),
+but this is not so much of a concern to us, as their planetary masses
+and therefore tidal timescales are typically expected to be at least
+an order of magnitude shorter than hot Jupiters.
 
@@
 orbital period that from transits appears to be decreasing by about 10
-milliseconds per year.
-After discovering the timing variations in the TESS data 
-\citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, we obtained N additional
-radial velocity measurements using
-Keck-HIRES, extending the RV baseline by Y years.
-Previously, the radial velocities were sparse and were consistent with
-not showing any linear trend.
-Our new measurements reveal a line of sight acceleration of
-$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
-This translates to an expected period decrease simply from the light
-travel time effect of ZZ milliseconds per year --- about what is
-observed.
+milliseconds per year.  We discovered the timing variations using
+data from TESS and a decade of ground-based monitoring
+\citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.
 
@@
 times for the system, and confirmed that the updated series of transit
-times was consistent with a quadratic ephemeris.  Their interpretation
-of the timing variations did not differ in any major respects from our
-own, though with additional data they found a lower best-fit decay
-rate of $\dot{P} = -XX.X \pm X.X$ milliseconds per year.
+times was consistent with a quadratic ephemeris.  With additional
+data, they found a lower best-fit decay rate of $\dot{P} = -XX.X \pm
+X.X$ milliseconds per year.
 
-A separate study by \citet{baluev_homogeneously_2019} reported
-additional light curves. \citeauthor{baluev_homogeneously_2019}
-analyzed their newly obtained photometry, along with archival light
-curves that we omitted from our analysis due to systematic
-uncertainties in the absolute time system.
-\citeauthor{baluev_homogeneously_2019} found that when they used all
+A separate study by \citet{baluev_2019} reported
+additional archival transit light curves of WASP-4b, most notably from
+the TRAPPIST archive, and a select few other observers.
+% \citeauthor{baluev_homogeneously_2019} analyzed their newly obtained
+% photometry, along with archival light curves that we omitted from our
+% analysis due to systematic uncertainties in the absolute time system.
+\citeauthor{baluev_2019} found that when they used all
 the available TTV data, the need for a quadratic ephemeris was present
@@
 that if they used lower-precision subsets of the available timing
-data, the necessity for the quadratic term decreased.
-\citeauthor{baluev_homogeneously_2019} also pointed out that the
-precise transit times reported by \citet{huitson_gemini_2017} were
-quite important in the time-series.  Overall,
-\citeauthor{baluev_homogeneously_2019} did not find the claim of a
-period decrease convincing.
+data, the necessity for the quadratic term decreased. This is not
+particularly surprising.
+% \citeauthor{baluev_homogeneously_2019} also pointed out that the
+% precise transit times reported by \citet{huitson_gemini_2017} were
+% quite important in the time-series.  Overall,
+% \citeauthor{baluev_homogeneously_2019} did not find the claim of a
+% period decrease convincing.
+
+To determine the origin of the period change, we acquired four
+additional radial velocity measurements using Keck-HIRES, extending
+the RV baseline from 3 to 9 years.  Previously, the radial velocities
+were sparse and showed a weak ($<2\sigma$) linear trend.  Our new
+measurements reveal a line of sight acceleration of $\dot{\gamma} =
+-XX {\rm m\,s^{-1}\,yr^{-1}}$.  This translates to an expected period
+decrease simply from the light travel time effect of ZZ milliseconds
+per year --- about commensurate with what is observed.
 
@@
 \S~\ref{sec:discussion} discusses...
-\S~\ref{sec:conclusion} gives conclusions...
+\S~\ref{sec:conclusions} gives conclusions...
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f1.pdf}
+		\includegraphics[width=0.5\textwidth]{f1.pdf}
 	\end{center}
@@
 	}
-\end{figure*}
+\end{figure}
 
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f2.pdf}
+		\includegraphics[width=0.5\textwidth]{f2.pdf}
 	\end{center}
@@
 %TODO: describe 3b better
-\begin{figure*}[!t]
+%TODO: also, put 3b in as the bottom right inset to the contrast curve
+\begin{figure}[!t]
 	\gridline{
 		\fig{f3a.png}{0.33\textwidth}{}
+	}
+	\vspace{-1cm}
+	\gridline{
 		\fig{f3b.pdf} {0.5\textwidth}{}
@@
     \caption{
-      {\it Left:\,} 
+      {\it Top:\,} 
       Speckle image reconstructed from 1000 40 millisecond frames in an 832 nm bandpass, and acquired on September 28, 2019.
       The image scale is $2.46''\times2.46''$.
-      {\it Right:\,} 
+      {\it Bottom:\,} 
       Contrast curve derived from point-source injection-recovery
@@
     \label{fig:zorro}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{How many other hot Jupiters will show the R\'omer delay?}
+\subsection{How many other hot Jupiters will show the R{\o}mer delay?}
 
@@
 \citet{knutson_friends_2014}, and computed the expected orbital period
-changes $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
+derivatives $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
 systems.
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f4.pdf}
+		\includegraphics[width=0.48\textwidth]{f5.pdf}
 	\end{center}
@@
 	\caption{
-    {\bf ...}
+  {\bf Predicted hot Jupiter period changes from linear radial
+  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
+  \citet{knutson_friends_2014} have shown long-term radial velocity
+  trends.  The HAT-P-11 signal is plotted, though its radial velocity
+  signal may be caused by stellar activity.  Three hot Jupiters are
+  not plotted because their radial velocity curves better described as
+  quadratic trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects
+  are ordered in the $y$ dimension by the absolute value of d$P$/d$t$.
 	\label{fig:pdot_pop}
@@
 
+We found that WASP-4 is accelerating towards the Earth, at
+X.XX m/s/yr.
+This seems like the most plausible explanation for the orbital 
+period decrease observed in transits~---~not tidal
+orbital decay, or apsidal precession.
+The most likely parameters for the companion are XXX.
+Over half of hot Jupiters have super-Jovian outer companions,
+so the R{\o}mer delay will become an increasingly large nuisance in
+the hunt for orbital decay as
+the observing baselines for hot Jupiters increase.
+The requirement to avoid missing out is to have long-term
+radial velocity monitoring with the same spectrograph (or else to
+calibrate instrumental offsets by observing the same stars simultaneously).
+
+%TODO: what's the scaling? sigma_P ~ 1/baseline? baseline^2?
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\acknowledgements
+% \acknowledgements
+% %
+% This paper includes data collected by the TESS mission, which are
+% publicly available from the Mikulski Archive for Space Telescopes
+% (MAST).
+% %
+% Funding for the TESS mission is provided by NASA's Science Mission
+% directorate.
+% %
+% This work made use of NASA's Astrophysics Data System Bibliographic
+% Services.
+% %
+% Based on observations obtained at the Gemini Observatory, which is
+% operated by the Association of Universities for Research in Astronomy,
+% Inc., under a cooperative agreement with the NSF on behalf of the
+% Gemini partnership: the National Science Foundation (United States),
+% National Research Council (Canada), CONICYT (Chile), Ministerio de
+% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
+% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
+% and Korea Astronomy and Space Science Institute (Republic of Korea).
+% %
+% Observations in the paper made use of the High-Resolution Imaging
+% instrument Zorro at Gemini-South. Zorro was funded by the NASA
+% Exoplanet Exploration Program and built at the NASA Ames Research
+% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
+% Quigley.
+% %
+% This research has made use of the VizieR catalogue access tool, CDS,
+% Strasbourg, France. The original description of the VizieR service was
+% published in A\&AS 143, 23.
+% %
+% This work has made use of data from the European Space Agency (ESA)
+% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
+% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
+% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
+% for the DPAC has been provided by national institutions, in particular
+% the institutions participating in the {\it Gaia} Multilateral
+% Agreement.
 %
-This paper includes data collected by the TESS mission, which are
-publicly available from the Mikulski Archive for Space Telescopes
-(MAST).
+% \newline
 %
-Funding for the TESS mission is provided by NASA's Science Mission
-directorate.
-%
-This work made use of NASA's Astrophysics Data System Bibliographic
-Services.
-%
-Based on observations obtained at the Gemini Observatory, which is
-operated by the Association of Universities for Research in Astronomy,
-Inc., under a cooperative agreement with the NSF on behalf of the
-Gemini partnership: the National Science Foundation (United States),
-National Research Council (Canada), CONICYT (Chile), Ministerio de
-Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
-Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
-and Korea Astronomy and Space Science Institute (Republic of Korea).
-%
-Observations in the paper made use of the High-Resolution Imaging
-instrument Zorro at Gemini-South. Zorro was funded by the NASA
-Exoplanet Exploration Program and built at the NASA Ames Research
-Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
-Quigley.
-%
-This research has made use of the VizieR catalogue access tool, CDS,
-Strasbourg, France. The original description of the VizieR service was
-published in A\&AS 143, 23.
-%
-This work has made use of data from the European Space Agency (ESA)
-mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
-by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
-\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
-for the DPAC has been provided by national institutions, in particular
-the institutions participating in the {\it Gaia} Multilateral
-Agreement.
-%
-\newline
-%
+
 \software{

### 2020-01-24 paper/collect.sh

@@ -3,3 +3,4 @@ cp ../results/20190911_rv_fit.pdf f2.pdf
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png f3a.png
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832.pdf f3b.pdf
 cp ../results/mass_semimaj_constraints.pdf f4.pdf
+cp ../results/k14_pdot.pdf f5.pdf

### 2020-01-24 src/derive_mass_semimaj_constraints.py
add plot_mass_semimaj_constraints
now as dedicated "module"

@@ -60,6 +60,8 @@ from WASP4 import time_base
 from cdips.plotting import savefig
 import matplotlib as mpl
 
+from plot_mass_semimaj_constraints import plot_mass_semimaj_constraints
+
 ##########
 # driver #
 ##########
@@ -204,32 +206,6 @@ def derive_mass_semimaj_constraints(
     prob_arr = np.exp(log_like)/np.exp(log_like).sum().sum()
 
     return prob_arr, mass_grid, sma_grid
-
-
-def plot_mass_semimaj_constraints(prob_arr, mass_grid, sma_grid):
-
-    fig, ax = plt.subplots(figsize=(4,3))
-
-    X,Y = np.meshgrid(sma_grid[:-1].value, mass_grid[:-1].value)
-
-    norm = mpl.colors.Normalize(vmin=np.log(prob_arr).max(), vmax=-30)
-
-    im = ax.pcolor(X, Y, np.log(prob_arr), cmap='gray_r', norm=norm)
-
-    ax.set_xscale('log')
-    ax.set_yscale('log')
-
-    ax.set_xlabel('Semi-major axis [AU]')
-    ax.set_ylabel('Companion mass [M$_\mathrm{{jup}}$]')
-
-    fig.colorbar(im, orientation='vertical', extend='min')
-
-    ax.get_yaxis().set_tick_params(which='both', direction='in')
-    ax.get_xaxis().set_tick_params(which='both', direction='in')
-    fig.tight_layout(h_pad=0, w_pad=0)
-
-    figpath = '../results/mass_semimaj_constraints.png'
-    savefig(fig, figpath)
 
 
 if __name__=="__main__":

### 2020-01-24 src/plot_mass_semimaj_constraints.py

@@ -1,0 +1,117 @@
+import os, pickle
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from numpy import array as nparr
+from cdips.plotting import savefig
+import matplotlib as mpl
+from astropy import units as u, constants as c
+
+def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
+                                  sma_grid=None, with_contrast=False,
+                                  discrete_color=False, linear_z=False):
+
+    if prob_arr is None:
+
+        n_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+        n_injections_per_cell = 500 # want: 500
+
+        mass_grid = (
+            np.logspace(np.log10(1), np.log10(300), num=n_grid_edges)*u.Mjup
+        )
+        sma_grid = (
+            np.logspace(np.log10(3), np.log10(500), num=n_grid_edges)*u.AU
+        )
+
+        sizestr = '{}x{}x{}'.format(n_grid_edges-1,
+                                    n_grid_edges-1,
+                                    n_injections_per_cell)
+        pklpath = (
+            '../data/rv_simulations/mass_semimaj_loglikearr_{}.pickle'.
+            format(sizestr)
+        )
+        log_like_arr = pickle.load(open(pklpath, 'rb'))
+
+        #
+        # Convert log-likelihood values to relative probability by taking the exp.
+        # Then average out the "sample" dimension (mass, sma, eccentricity, etc).
+        #
+        log_like = np.log(np.exp(log_like_arr).mean(axis=2))
+
+        # -2*logprob == chi^2
+        # Convert likelihood values to a normalized probability via
+        #   P ~ -exp(-chi^2/2)
+        prob_arr = np.exp(log_like)/np.exp(log_like).sum().sum()
+
+    #################
+    # make the plot #
+    #################
+
+    fig, ax = plt.subplots(figsize=(4,3))
+
+    X,Y = np.meshgrid(sma_grid[:-1].value, mass_grid[:-1].value)
+
+    cmap = plt.cm.gray_r
+
+    if not linear_z:
+
+        if discrete_color:
+            bounds = np.round(np.linspace(-10, np.log(prob_arr).max(), 5),1)
+            norm = mpl.colors.BoundaryNorm(bounds, cmap.N)
+        else:
[…]

### 2020-01-24 paper/collect.sh
update paper

@@ -2,5 +2,5 @@ cp ../results/O_minus_C_includeallpoints_onlytransits.pdf f1.pdf
 cp ../results/20190911_rv_fit.pdf f2.pdf
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png f3a.png
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832.pdf f3b.pdf
-cp ../results/mass_semimaj_constraints.pdf f4.pdf
+cp ../results/mass_semimaj_constraints_with_contrast_discretecolor.pdf f4.pdf
 cp ../results/k14_pdot.pdf f5.pdf

### 2020-01-25 paper/ms.tex
update paper
minor

@@
 %
-% key authors: presumably Rachel did the Zorro reductions?
+% key authors: did Rachel do the Zorro reductions?
 %
@@
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
-%
-\author{R. A. Matson}
-\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
-%
 
@@
 Institute of Technology, Pasadena, CA 91125, USA}
+%
+%
+%\author{R. A. Matson}
+% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
 %
@@
 
-%FIXME: I should ask Didier if he has any new CORALIE observations...
 The complete set of radial velocity observations is given in {\bf
@@
       {\it Top:\,} 
-      Speckle image reconstructed from 1000 40 millisecond frames in an 832 nm bandpass, and acquired on September 28, 2019.
-      The image scale is $2.46''\times2.46''$.
+      Speckle image reconstructed from 1000 40 millisecond frames in
+      an 832 nm bandpass, and acquired on September 28, 2019.  The
+      image scale is $2.46''\times2.46''$.
       {\it Bottom:\,} 
@@
 \citetalias{bouma_wasp4b_2019}, we derived the transformation from
-contrast ratio to companion mass.
-The resulting limits are shown as the cross-hatched region in
-Figure~\ref{fig:mass_sma}.
+contrast ratio to companion mass.  The resulting limits are shown as
+the cross-hatched region in Figure~\ref{fig:mass_sma}.
 
@@
 from the radial velocities, we followed the procedure of
-\citet{bryan_excess_2019}.  We began by defining a $50\times50$ grid
-in true planetary mass and semimajor axis, evenly spaced from 1 to
-300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the
-possibility that an additional companion in any particular cell could
-explain the observed linear trend.  In each cell, we simulated 500
-hypothetical companions.
+\citet{bryan_excess_2019}.  We began by defining a $75\times50$ grid
+in true planetary mass and semimajor axis, with even logarithmic
+spacing from 1 to 900$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We
+then considered the possibility that an additional companion in any
+particular cell could explain the observed linear trend.  In each
+cell, we simulated 500 hypothetical companions.
 

### 2020-01-25 src/derive_mass_semimaj_constraints.py

@@ -82,22 +82,23 @@ def derive_mass_semimaj_constraints(
          setupfn, rvfitdir
     )
 
-    n_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-    n_injections_per_cell = 500 # want: 500
+    n_mass_grid_edges = 76 # a 4x4 grid has 5 edges. want: 51
+    n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+    n_injections_per_cell = 1 # 500 # want: 500
 
     mass_grid = (
-        np.logspace(np.log10(1), np.log10(300), num=n_grid_edges)*u.Mjup
+        np.logspace(np.log10(1), np.log10(900), num=n_mass_grid_edges)*u.Mjup
     )
     sma_grid = (
-        np.logspace(np.log10(3), np.log10(500), num=n_grid_edges)*u.AU
+        np.logspace(np.log10(3), np.log10(500), num=n_sma_grid_edges)*u.AU
     )
 
     log_like_arr = np.zeros(
-        (n_grid_edges-1, n_grid_edges-1, n_injections_per_cell)
+        (n_mass_grid_edges-1, n_sma_grid_edges-1, n_injections_per_cell)
     )
 
-    sizestr = '{}x{}x{}'.format(n_grid_edges-1,
-                                n_grid_edges-1,
+    sizestr = '{}x{}x{}'.format(n_mass_grid_edges-1,
+                                n_sma_grid_edges-1,
                                 n_injections_per_cell)
     outpath = (
         '../data/rv_simulations/mass_semimaj_loglikearr_{}.pickle'.

### 2020-01-25 paper/ms.tex
eccentricities: planets and stars be different

@@
 To derive constraints on possible companion masses and separations
-from the radial velocities, we followed the procedure of
-\citet{bryan_excess_2019}.  We began by defining a $75\times50$ grid
-in true planetary mass and semimajor axis, with even logarithmic
-spacing from 1 to 900$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We
-then considered the possibility that an additional companion in any
-particular cell could explain the observed linear trend.  In each
-cell, we simulated 500 hypothetical companions.
+from the radial velocities, we mostly followed the procedure of
+\citet{bryan_excess_2019}.
 
+We began by defining a $75\times50$ grid in true planetary mass and
+semimajor axis, with even logarithmic spacing from 1 to 900$\,$$M_{\rm
+Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the possibility
+that an additional companion in any particular cell could explain the
+observed linear trend.  In each cell, we simulated 500 hypothetical
+companions.
+
+%FIXME: may want to verify that changing the cutoff location, to say
+%80Mjup or whatever the minimum BD mass is doesn't drastically alter
+%the posterior
 We assigned each companion a mass and semimajor axis from log-uniform
 distributions within the grid cell. We drew the inclination from a
-uniform distribution in $\cos i$.  The eccentricity was drawn from
+uniform distribution in $\cos i$.  For companion masses less than
+$10\,M_{\rm Jup}$, we drew the eccentricity from
 \citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
-($a=1.12$, $b=3.09$).  For each simulated companion, we then drew a
-sample from the converged chains of our initial model of WASP-4b, plus
-its linear trend. We subtracted the planet's orbital solution, leaving
-the linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each
-simulated outer companion, and the fixed instrument offsets and
-jitters from the MCMC chains, we then performed a maximum likelihood
-fit for the time and argument of periastron of the outer simulated
-companion.  We converted the resulting $50\times50\times500$ cube of
-best-fit log-likelihood values to probabilities, and averaged over the
-samples in each grid cell to derive a probability distribution in mass
-and semi-major axis.
+($a=1.12$, $b=3.09$).  If the companion mass exceeded $10\,M_{\rm
+Jup}$, we drew the eccentricity distribution from the power-law $p_e
+\propto e^\eta$ reported by \citet{moe_mind_2017} in their
+Equation~17.  We emphasize that the long-period exoplanet and
+long-period binary distributions are quite different: the exoplanet
+distribution is ``bottom-heavy'', with eccentricities preferentially
+close to zero. The binary star distribution is ``top-heavy'', with
+more eccentricities close to one.  We chose a mass cutoff of
+$10\,M_{\rm Jup}$ based on the bound observed by
+\citet{schlaufman_evidence_2018} between giant planets, presumably
+formed through core accretion, and brown dwarfs, presumably formed via
+gravitational instability.
+
+For each simulated companion, we then drew a sample from the converged
+chains of our initial model of WASP-4b, plus its linear trend. We
+subtracted the planet's orbital solution, leaving the linear trend.
+Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulated outer
+companion, and the fixed instrument offsets and jitters from the MCMC
+chains, we then performed a maximum likelihood fit for the time and
+argument of periastron of the outer simulated companion.  We converted
+the resulting $75\times50\times500$ cube of best-fit log-likelihood
+values to probabilities, and averaged over the samples in each grid
+cell to derive a probability distribution in mass and semi-major axis.
 

### 2020-01-25 src/beta_vs_powerlaw.py
very different

@@ -1,0 +1,21 @@
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from scipy.stats import beta, powerlaw
+
+n_draws = int(1e4)
+a, b = 1.12, 3.09
+ecc_pl = beta.rvs(a, b, size=n_draws)
+
+eta = 0.5 # roughly
+
+# f(x,a) = ax^{a-1} for scipy's built-in powerlaw distribution.
+ecc_stellar = powerlaw.rvs(eta+1, size=n_draws)
+
+plt.hist(ecc_pl, color='C0', label='beta (1.12,3.09)', alpha=0.5)
+
+plt.hist(ecc_stellar, color='C1', label='powerlaw (e^{0.5})',
+         alpha=0.5)
+plt.xlabel('ecc')
+plt.legend()
+outpath = '../results/beta_vs_powerlaw.png'
+plt.savefig(outpath, dpi=350)
+print('made {}'.format(outpath))

### 2020-01-25 src/derive_mass_semimaj_constraints.py
[WIP] update ecc draw procedure
WIP because we're gonna multiprocess the max-like fitting

@@ -38,7 +38,7 @@ The distribution is shown in {\bf Figure~X}.
 
 import os, pickle
 import numpy as np, pandas as pd, matplotlib.pyplot as plt
-from astropy import units as u, constants as c
+from astropy import units as u, constants as const
 from itertools import product
 from numpy import array as nparr
 from datetime import datetime
@@ -53,7 +53,7 @@ import configparser
 from radvel_utils import (
     draw_models, _get_fit_results, args_object, initialize_sim_posterior
 )
-from scipy.stats import beta
+from scipy.stats import beta, powerlaw
 
 from WASP4 import time_base
 
@@ -84,7 +84,7 @@ def derive_mass_semimaj_constraints(
 
     n_mass_grid_edges = 76 # a 4x4 grid has 5 edges. want: 51
     n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-    n_injections_per_cell = 1 # 500 # want: 500
+    n_injections_per_cell = 500 # 500 # want: 500
 
     mass_grid = (
         np.logspace(np.log10(1), np.log10(900), num=n_mass_grid_edges)*u.Mjup
@@ -150,12 +150,36 @@ def derive_mass_semimaj_constraints(
             cos_incl_c = np.random.uniform(0, 1, size=n_injections_per_cell)
             incl_c = np.rad2deg(np.arccos(cos_incl_c))
 
+            ecc_c = np.zeros(n_injections_per_cell)
+
+            # case: companion is a planet. use Kipping (2013).
             a, b = 1.12, 3.09
-            ecc_c = beta.rvs(a, b, size=n_injections_per_cell)
+            _ecc_c_planetary = beta.rvs(a, b, size=n_injections_per_cell)
+
+            # case: companion is a star. go for Moe & Di Stefano (2017), Eq 17.
+
+            mstar = 0.864*u.Msun # WASP-4
+            period_c = ((
+                (sma_c*u.AU)**(3) *
+                4*np.pi**2 / (const.G * (mstar+mass_c*u.Mjup) )
+            )**(1/2)).to(u.day)
+
+            # Moe & Di Stefano (2017), Eq 17.
+            eta = 0.6 - 0.7 / ( np.log10(period_c.to(u.day).value) - 0.5 )
+
+            # f(x,a) = ax^{a-1} for scipy's built-in powerlaw distribution.
[…]

### 2020-01-25 src/plot_mass_semimaj_constraints.py

@@ -11,18 +11,19 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     if prob_arr is None:
 
-        n_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-        n_injections_per_cell = 500 # want: 500
+        n_mass_grid_edges = 76 # a 4x4 grid has 5 edges. want: 51
+        n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+        n_injections_per_cell = 1 # 500 # want: 500
 
         mass_grid = (
-            np.logspace(np.log10(1), np.log10(300), num=n_grid_edges)*u.Mjup
+            np.logspace(np.log10(1), np.log10(900), num=n_mass_grid_edges)*u.Mjup
         )
         sma_grid = (
-            np.logspace(np.log10(3), np.log10(500), num=n_grid_edges)*u.AU
+            np.logspace(np.log10(3), np.log10(500), num=n_sma_grid_edges)*u.AU
         )
 
-        sizestr = '{}x{}x{}'.format(n_grid_edges-1,
-                                    n_grid_edges-1,
+        sizestr = '{}x{}x{}'.format(n_mass_grid_edges-1,
+                                    n_sma_grid_edges-1,
                                     n_injections_per_cell)
         pklpath = (
             '../data/rv_simulations/mass_semimaj_loglikearr_{}.pickle'.

### 2020-01-25 src/minimum_mass_given_linear_trend.py
fix bug

@@ -5,13 +5,16 @@ Wright's blog says:
 "the lowest mass object that could be contributing a linear trend turns out to
 be one in an eccentric orbit with e~0.5"
 https://sites.psu.edu/astrowright/2015/01/06/loooooooong-period-exoplanets/
+
+This calculation was published in https://arxiv.org/pdf/1501.00633.pdf,
+Katherina Feng et al (2015).
 """
 
 import numpy as np
 from astropy import units as u, constants as c
 
 baseline = 10*u.year
-gammadot = -0.04 *u.m/u.s/u.year
+gammadot = -0.04 *u.m/u.s/u.day
 mstar = 0.864*u.Msun
 
 M_min = (

### 2020-01-25 paper/ms.tex
update words

@@
 \begin{abstract}
-  We recently reported that space and ground-based transit observations
-  showed the orbital period of WASP-4b to be decreasing.
-  Possible explanations included tidal orbital decay, orbital
-  precession, and light-travel time effects.
-  In this work we present new observations of the radial velocity of
-  WASP-4, acquired with Keck-HIRES.
-  The new data show that the system is accelerating towards our line
-  of sight, at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm
-  s}^{-1}\,{\rm yr}^{-1}$.
-  This implies a period decrease of X.XX milliseconds per year,
-  compared to the observed value of Y.YY.
-  Combining the radial velocities with new speckle imaging from
-  Gemini-Zorro, we find that the WASP-4 system likely has a
-  10-200$\,M_{\rm Jup}$ companion with semi-major axis between 10 and
-  100$\,$AU.
-  As observing baselines continue to extend, long-term tracking of hot
-  Jupiter radial velocities will be necessary to distinguish tidal
-  orbital decay from the R{\o}mer delay.
+  Space and ground-based transit observations of the hot Jupiter
+  WASP-4b have recently shown that its orbital period is decreasing.
+  Potential causes of the period change include tidal orbital decay,
+  orbital precession, and light-travel time effects.
+  In this work we present new radial velocity measurements of WASP-4,
+  acquired with Keck-HIRES.
+  The new data show that the system is accelerating towards the Earth
+  at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm s}^{-1}\,{\rm
+  yr}^{-1}$.
+  The corresponding period decrease explains WASP-4's
+  changing orbital period as a light-travel time effect.
+  Combining the radial velocities with new speckle imaging limits, we
+  find that system's acceleration is likely caused by a
+  10-200$\,M_{\rm Jup}$ companion with semi-major axis between 10-100$\,$AU.
+  Based on statistics from \citet{knutson_friends_2014}, about 1 in 6
+  hot Jupiters are expected to show period decreases commensurate with
+  WASP-4 due to acceleration by outer companions.
+  Continued long-term radial velocity monitoring of hot Jupiters is
+  therefore essential to distinguish tidal orbital decay from
+  line-of-sight accelerations.
 \end{abstract}
@@
 
-Since the first discovery of transiting hot Jupiters, the hunt for
-tidal orbital decay has received much attention (CITE LIKE 20 PAPERS).
-These efforts are gradually beginning to yield fruit: by combining
-transit and occultation timing and radial velocity measurements, CITE
-Yee et al{.} have shown that the data for the hot Jupiter WASP-12b are
-most compatible with orbital decay.
+The issue of whether and how the orbits of hot Jupiters decay due to
+tidal interaction with their host stars has drawn significant
+attention over the past 25 years
+\citep{rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
+Numerous observations have also been performed (CITE).  These efforts
+are now yielding fruit: by combining transit and occultation timing
+with radial velocity measurements, \citet{yee_orbit_2020} have shown
+that the data for the hot Jupiter WASP-12b are most compatible with
+orbital decay.
 
-The present study highlights a point that, while obvious, has perhaps
-not yet received due attention.  The point is that the hunt for
-orbital decay of hot Jupiters will be crippled without long-term
-radial velocity monitoring programs.  The reason is simple: when the
-timescale of orbital period change far exceeds the observing baseline,
-the first deviation from constant periodicity is always quadratic in
-the transit number.
+The present study highlights a point that, though obvious, has perhaps
+not yet received due attention.  The point is that observational
+programs aimed at identifying orbital decay in hot Jupiters will be
+crippled without long-term radial velocity monitoring programs.  The
+reason is simple: when the timescale of orbital period change exceeds
+the observing baseline, the first deviation from constant periodicity
+is always quadratic in the transit number.
 
-There are ample reasons to expect orbital decay to be common (CITE).
-However there are reasons to expect the R{\o}mer delay to perhaps be
-even {\it more} common (CITE Knutson 14, Bryan 16).  Both manifest to
-first order identically in transit times (up to the sign of the
-line-of-sight acceleration) -- as a non-zero period derivative.
-
+While orbital decay may be common (CITE).  line-of-sight accelerations
+could be even {\it more} common (CITE Knutson 14, Bryan 16).  Both
+manifest to first order identically in transit times (up to the sign
+of the line-of-sight acceleration) -- as a non-zero period derivative.
 Specifically, \citet{knutson_friends_2014} showed that X\% of hot
@@
 say 10 years.
-The abundance of giant planets
-exterior to Super-Earth sized systems is also high (CITE Bryan 2019),
-but this is not so much of a concern to us, as their planetary masses
-and therefore tidal timescales are typically expected to be at least
-an order of magnitude shorter than hot Jupiters.
+
+% The abundance of giant planets
+% exterior to Super-Earth sized systems is also high (CITE Bryan 2019),
+% but this is not so much of a concern to us, as their planetary masses
+% and therefore tidal timescales are typically expected to be at least
+% an order of magnitude shorter than hot Jupiters.
 
@@
 \citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.
-
 Thereafter, \citet{southworth_transit_2019} reported 22 new transit
@@
 X.X$ milliseconds per year.
-
 A separate study by \citet{baluev_2019} reported
@@
 % analysis due to systematic uncertainties in the absolute time system.
-\citeauthor{baluev_2019} found that when they used all
-the available TTV data, the need for a quadratic ephemeris was present
-``at the high $\sim 5-7$ sigma level''.  However, they pointed out
-that if they used lower-precision subsets of the available timing
-data, the necessity for the quadratic term decreased. This is not
-particularly surprising.
+% \citeauthor{baluev_2019} found that when they used all the available
+% TTV data, the need for a quadratic ephemeris was present.  However,
+% they pointed out that if they used lower-precision subsets of the
+% available timing data, the necessity for the quadratic term decreased.
+% This is not particularly surprising.
 % \citeauthor{baluev_homogeneously_2019} also pointed out that the
@@
 were sparse and showed a weak ($<2\sigma$) linear trend.  Our new
-measurements reveal a line of sight acceleration of $\dot{\gamma} =
+measurements reveal a line-of-sight acceleration of $\dot{\gamma} =
 -XX {\rm m\,s^{-1}\,yr^{-1}}$.  This translates to an expected period
@@
 per year --- about commensurate with what is observed.
+In the following, we avoid the term ``R{\o}mer delay'' as this phrase
+typically signifies arrival time delay due to observatory motion. We
+are discussing an analog of the Doppler effect for a source with
+constant line-of-sight acceleration.
 
@@
 
-WASP-4 was found to be accelerating towards our line of sight at high
+WASP-4 was found to be accelerating towards our line-of-sight at high
 confidence,
@@
 
-The system's acceleration towards our line of sight causes a decrease
-in the apparent orbital period (the R{\o}mer delay).  The period
+The system's acceleration towards our line-of-sight causes a decrease
+in the apparent orbital period.  The period
 derivative expected from radial velocities is
@@
 Given a linear radial velocity trend, we can place lower-limits on the
-mass and semi-major axis of additional bodies in the system.  The
-constraints on the two parameters are degenerate: $\dot{\gamma}
-\propto M_{\rm p} a^{-2}$.  Nonetheless, the Zorro images and contrast
-curves can limit the available parameter space by setting an upper
-limit on the semimajor axis, and a maximum brightness (and thereby
-mass) of any putative companions.
+mass and semi-major axis of additional bodies in the system.  As a
+quick estimate of the minimum mass required to explain the linear
+trend in WASP-4, we turn to \citet{feng_california_2015}.  As they
+discuss, the scenario that yields the minimum companion mass for a
+system with a linear trend is a companion with $e\approx0.5$ and
+$\omega=90^\circ$.  From their Equation~1,
+\begin{equation}
+ M_{\rm min} \approx 0.0164\,M_{\rm Jup}
+  \left( \frac{\tau}{{\rm yr}} \right)^{4/3}
+  \left| \frac{\dot{\gamma}}{{\rm m\,s^{-1}\,yr^{-1}}} \right|
+  \left( \frac{M_\star}{{M_\odot}} \right)^{2/3},
+\end{equation}
+where $\tau$ is the observing baseline.
+For WASP-4, this yields $M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher mass
+companions are allowed, presuming that they orbit further from the
+star; at fixed $\dot{\gamma}$, $M_{\rm comp} \propto a^2$
+\citep{torres_substellar_1999,liu_crossing_2002}.
 
-The general procedure we use to derive constraints on possible
-companion masses and semi-major axes has been developed by
+High-resolution images can further limit the available parameter space
+by setting an upper limit on the semimajor axis, and a maximum
+brightness (and thereby mass) of any putative companions.
+The procedure we use to combine constraints from both radial
+velocities and high resolution imaging has been developed by
 \citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
@@
 
-\subsection{How many other hot Jupiters will show the R{\o}mer delay?}
+\subsection{How many other hot Jupiters are accelerating towards the
+Earth?}
 
@@
 Jupiters with significant linear trends in Figure~\ref{fig:pdot_pop}.
+
+Including WASP-4b, 16 of 51 hot Jupiters from
+\citet{knutson_friends_2014} show a non-zero radial velocity trend.
+About half are positive, and half are negative, and therefore
+$\approx$1 in 6 hot Jupiters are expected to show period decreases
+commensurate with WASP-4 due to acceleration by outer companions.
 
@@
 Over half of hot Jupiters have super-Jovian outer companions,
-so the R{\o}mer delay will become an increasingly large nuisance in
+so the light-travel time effect will become an increasingly large nuisance in
 the hunt for orbital decay as

### 2020-01-27 src/plot_O_minus_C_paperII.py
plot_O_minus_C_postpub -> plot_O_minus_C_paperII
and i made the plot better

@@ -1,0 +1,309 @@
+# -*- coding: utf-8 -*-
+'''
+make O-C model figure
+'''
+from __future__ import division, print_function
+
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt, pandas as pd, numpy as np
+
+from glob import glob
+from shutil import copyfile
+import os, pickle
+
+from numpy import array as nparr
+
+from astropy.time import Time
+from astropy import units as units, constants as const
+
+def get_data(
+    datacsv='../data/WASP-18b_transits_and_TESS_times_O-C_vs_epoch_selected.csv',
+    is_occultation=False
+    ):
+    # need to run make_parameter_vs_epoch_plots.py first; this generates the
+    # SELECTED epochs (x values), mid-times (y values), and mid-time errors
+    # (sigma_y).
+
+    df = pd.read_csv(datacsv, sep=';')
+
+    tcol = [c for c in nparr(df.columns) if
+            ('times_BJD_TDB_minus_' in c) and ('minutes' in c)]
+    if len(tcol) != 1:
+        raise AssertionError('unexpected input file')
+    else:
+        tcol = tcol[0]
+
+    if is_occultation:
+        err_column = 'err_sel_occ_times_BJD_TDB_minutes'
+    else:
+        err_column = 'err_sel_transit_times_BJD_TDB_minutes'
+
+    data = nparr( [
+        nparr(df['sel_epoch']),
+        nparr(df[tcol]),
+        nparr(df[err_column])
+    ])
+
+    x, y, sigma_y = data
+    refs = nparr(df['original_reference'])
+
+    return x, y, sigma_y, data, tcol, refs
+
+
+def linear_fit(theta, x, x_occ=None):
+    """
+    Linear model. Parameters (t0, P).
+    Must pass transit times.
+
+    If x_occ is none, returns model t_tra array.
+    If x_occ is a numpy array, returns tuple of model t_tra and t_occ arrays.
+    """
+    t0, period = theta
+    if not isinstance(x_occ,np.ndarray):
+        return t0 + period*x
+    else:
+        return t0 + period*x, t0 + period/2 + period*x_occ
+
+def quadratic_fit(theta, x, x_occ=None):
+    """
+    Quadratic model. Parameters (t0, P, 0.5dP/dE).
+    Must pass transit times.
+
[…]

### 2020-01-27 src/plot_O_minus_C_postpub.py

@@ -1,704 +1,0 @@
-# -*- coding: utf-8 -*-
-'''
-make O-C model figure
-'''
-from __future__ import division, print_function
-
-import matplotlib as mpl
-mpl.use('Agg')
-import matplotlib.pyplot as plt, pandas as pd, numpy as np
-
-from glob import glob
-from shutil import copyfile
-import os, pickle
-
-from numpy import array as nparr
-
-from astropy.time import Time
-from astropy import units as units, constants as const
-
-def get_data(
-    datacsv='../data/WASP-18b_transits_and_TESS_times_O-C_vs_epoch_selected.csv',
-    is_occultation=False
-    ):
-    # need to run make_parameter_vs_epoch_plots.py first; this generates the
-    # SELECTED epochs (x values), mid-times (y values), and mid-time errors
-    # (sigma_y).
-
-    df = pd.read_csv(datacsv, sep=';')
-
-    tcol = [c for c in nparr(df.columns) if
-            ('times_BJD_TDB_minus_' in c) and ('minutes' in c)]
-    if len(tcol) != 1:
-        raise AssertionError('unexpected input file')
-    else:
-        tcol = tcol[0]
-
-    if is_occultation:
-        err_column = 'err_sel_occ_times_BJD_TDB_minutes'
-    else:
-        err_column = 'err_sel_transit_times_BJD_TDB_minutes'
-
-    data = nparr( [
-        nparr(df['sel_epoch']),
-        nparr(df[tcol]),
-        nparr(df[err_column])
-    ])
-
-    x, y, sigma_y = data
-    refs = nparr(df['original_reference'])
-
-    return x, y, sigma_y, data, tcol, refs
-
-
-def linear_fit(theta, x, x_occ=None):
-    """
-    Linear model. Parameters (t0, P).
-    Must pass transit times.
-
-    If x_occ is none, returns model t_tra array.
-    If x_occ is a numpy array, returns tuple of model t_tra and t_occ arrays.
-    """
-    t0, period = theta
-    if not isinstance(x_occ,np.ndarray):
-        return t0 + period*x
-    else:
-        return t0 + period*x, t0 + period/2 + period*x_occ
-
-def quadratic_fit(theta, x, x_occ=None):
-    """
-    Quadratic model. Parameters (t0, P, 0.5dP/dE).
-    Must pass transit times.
-
-    If x_occ is none, returns model t_tra array.
-    If x_occ is a numpy array, returns tuple of model t_tra and t_occ arrays.
[…]

### 2020-01-27 src/derive_mass_semimaj_constraints.py
figure improvements
* full O-C revamp with the Southworth and Baluev transit points
* propagated through the RV O-C diagram
* running bigger grid for mass vs sma constraints

@@ -97,8 +97,8 @@ def derive_mass_semimaj_constraints(
          setupfn, rvfitdir
     )
 
-    n_mass_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-    n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+    n_mass_grid_edges = 129 # a 4x4 grid has 5 edges. want: 64+1, 128+1...
+    n_sma_grid_edges = 129 # a 4x4 grid has 5 edges.
     n_injections_per_cell = 512 # 500 # want: 500
 
     mass_grid = (

### 2020-01-27 src/plot_mass_semimaj_constraints.py

@@ -11,9 +11,9 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     if prob_arr is None:
 
-        n_mass_grid_edges = 26 # a 4x4 grid has 5 edges. want: 51
-        n_sma_grid_edges = 26 # a 4x4 grid has 5 edges. want: 51
-        n_injections_per_cell = 16 # 500 # want: 500
+        n_mass_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+        n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+        n_injections_per_cell = 512 # 500 # want: 500
 
         mass_grid = (
             np.logspace(np.log10(1), np.log10(900), num=n_mass_grid_edges)*u.Mjup
@@ -98,6 +98,8 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     ax.get_yaxis().set_tick_params(which='both', direction='in')
     ax.get_xaxis().set_tick_params(which='both', direction='in')
+    ax.tick_params(right=True, which='both', direction='in')
+    ax.tick_params(top=True, which='both', direction='in')
     fig.tight_layout(h_pad=0, w_pad=0)
 
     dstr = '_discretecolor' if discrete_color else ''

### 2020-01-27 src/plot_rvs_2020.py

@@ -75,9 +75,9 @@ def main(make_my_plot=1):
 
     # what would explain the Pdot from transits?
     period = 1.338231466*units.day
-    Pdot_tra = -4e-10
-    Pdot_tra_perr = -4e-10 + 0.4e-10
-    Pdot_tra_merr = -4e-10 - 0.4e-10
+    Pdot_tra = -2.736e-10
+    Pdot_tra_perr = Pdot_tra + 2.83e-11
+    Pdot_tra_merr = Pdot_tra - 2.83e-11
     dvdt_tra = (Pdot_tra * const.c / period).to(
         (units.m/units.s)/units.day).value
     dvdt_tra_perr = (Pdot_tra_perr * const.c / period).to(
@@ -121,6 +121,7 @@ def main(make_my_plot=1):
 
     ax.get_yaxis().set_tick_params(which='both', direction='in')
     ax.get_xaxis().set_tick_params(which='both', direction='in')
+    ax.tick_params(right=True, which='both', direction='in')
     a_top.get_yaxis().set_tick_params(which='both', direction='in')
     a_top.get_xaxis().set_tick_params(which='both', direction='in')
 

### 2020-01-27 paper/ms.tex
update paper

@@
 the TRAPPIST archive, and a select few other observers.
-% \citeauthor{baluev_homogeneously_2019} analyzed their newly obtained
+% \citeauthor{baluev_2019} analyzed their newly obtained
 % photometry, along with archival light curves that we omitted from our
@@
 % This is not particularly surprising.
-% \citeauthor{baluev_homogeneously_2019} also pointed out that the
+% \citeauthor{baluev_2019} also pointed out that the
 % precise transit times reported by \citet{huitson_gemini_2017} were
 % quite important in the time-series.  Overall,
-% \citeauthor{baluev_homogeneously_2019} did not find the claim of a
+% \citeauthor{baluev_2019} did not find the claim of a
 % period decrease convincing.
@@
 midpoint was fitted as a free parameter, and (c) the time system
-specified both the absolute reference (TDB or UTC) and also whether
-any barycentric or heliocentric corrections had been performed.
+specified both the leap second correction (TDB or UTC) and also
+whether any barycentric or heliocentric corrections had been
+performed.
 
+% FIXME : verify with Michael Gillon that this is true. (P. Evans
+% done)
 The majority of times are identical to those we collected in
@@
 observed from the 3.58m NTT and Danish 1.54m telescopes at La Silla,
-and the SAAO 1.0m telescope.
+and the SAAO 1.0m telescope.  Additional timimg
+measurements have also recently been reported by \citet{baluev_2019},
+based on a homogeneous analysis of archival ground-based observations
+taken professional and amateur observers.  We included
+twelve of their ``high quality'' transit times acquired by TRAPPIST
+(six transits), El~Sauce (four transits), and
+\citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
+the original observers that correct barycentric and leap-second
+corrections had been performed (M.~Gillon, P.~Evans, priv{.} comm{.}).
+We omitted the fourteen remaining \citeauthor{baluev_2019}
+ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
+whether leap-second corrections had or had not been performed.
 
-% FIXME : verify with Michael Gillon and Phil Evans that this is true.
-A number of additional times were also recently reported by
-\citet{baluev_2019}, based on a homogeneous analysis of archival
-ground-based observations.  We included their ``high quality'' transit
-times with provenance from the TRAPPIST and El Sauce telescopes, for
-which we verified with the original observers that appropriate
-time-system corrections had been performed (M.~Gillon, P.~Evans,
-priv{.} comm{.}).
-
-The four available occultation data points tabulated by
-\citetalias{bouma_wasp4b_2019} are sparse and have neglegible
-statistical value due to their large uncertainties, and we forgo
-their use in this analysis.
+The four available occultations tabulated by
+\citetalias{bouma_wasp4b_2019} have neglegible statistical value due
+to their large uncertainties, and we forgo their use in this analysis.
 
@@
 \subsection{Transits}
+\label{sec:transit_analysis}
 
@@
   The vertical axis shows the observed times minus the calculated
-  times assuming a constant orbital period .  Darker points correspond
+  times assuming a constant orbital period.  More opaque points correspond
   to more precise data.  The quadratic ephemeris and its $1\sigma$
-  uncertainties are shown in blue.  The binned TESS point (red square)
-  is the weighted average of 18 TESS transits and is for display
-  purposes only.  The models were fitted to all of the individual
-  transit times.
+  statistical uncertainties are shown in blue.  Neither model fully
+  explains the transit residuals (see~\ref{sec:transit_analysis}).
+  The binned TESS point (yellow star) is the weighted average of 18
+  TESS transits and is for display purposes only.  Its uncertainty is
+  slightly larger than the marker size.  The models were fitted to all
+  of the individual transit times.
   \label{fig:times}
@@
 
-We considered two possible models for the observed transit times.
-The first model assumes a constant orbital period on a circular orbit:
+We considered two possible models for the observed transit times.  The
+first model assumes a constant orbital period on a circular orbit:
 \begin{align}
@@
 \end{align}
-where $E$ is the epoch number and $t_0$ is a reference epoch.
-The second model assumes the period changes at a steady rate:
+where $E$ is the epoch number and $t_0$ is a reference epoch.  The
+second model assumes the period changes at a steady rate:
 \begin{align}
@@
 \end{align}
-The free parameters are the reference epoch $t_0$, the period at
-the reference epoch, and the period derivative, ${\rm d}P/{\rm d}t =
-(1/P) {\rm d}P/{\rm d}E$.
-We defined the epoch numbers such that $E=0$ is near the weighted
-average of the observed times.  This helps to reduce the covariance
-between $t_0$ and $P$.
-A third possible model that we did not consider for reasons that will
-become apparent is a precessing, eccentric orbit \citep[{\it
+The free parameters are the reference epoch $t_0$, the period at the
+reference epoch, and the period derivative, ${\rm d}P/{\rm d}t = (1/P)
+{\rm d}P/{\rm d}E$.  We defined the epoch numbers such that $E=0$ is
+near the weighted average of the observed times.  This helps to reduce
+the covariance between $t_0$ and $P$.  A third possible model that we
+did not consider for reasons that will become apparent is a
+precessing, eccentric orbit \citep[{\it
 e.g.},][]{gimenez_revision_1995,patra_2017}.
@@
 Figure~\ref{fig:times} shows the observed transit times, minus the
-best-fit constant period model. 
-The best-fitting constant-period model has $\chi^2 = XXX$ and YY
-degrees of freedom.  The best-fitting
-quadratic model has $\chi^2 = YYY$ and ZZ degrees of freedom. 
+best-fit constant period model.  The best-fitting constant-period
+model has $\chi^2 = 276$, 91 degrees of freedom, and $\chi^2_{\rm
+red}=3.0$.  The best-fitting quadratic model has $\chi^2 = 183$, 90
+degrees of freedom, and $\chi^2_{\rm red}=2.0$.  From the reduced
+$\chi^2$ values, we can surmise that neither model entirely describes
+the transit data that we have collected---there must be some
+additional source of signal or potentially noise. In
+\citet{bouma_wasp4b_2019}, we found that the quadratic model for the
+(less complete) transit data gave $\chi^2_{\rm red}=1.0$.  The only
+difference between our old and new transit datasets are the addition
+of transit times from \citet{southworth_transit_2019} and
+\citet{baluev_2019}.  The worsened $\chi^2_{\rm red}$ could reflect
+underestimated statistical uncertainties in these or in fact any of
+the transit measurements. It could also reflect systematic errors in
+the time-systems in which the transit measurements were recorded,
+though we have taken every caution against the latter possibility.
+% For purposes of simplicity, we simply accept that the additional data
+% have effectively increased the scatter in the transit O-C diagram.
 The difference in the BIC \citep{kass_bayes_1995} between the linear
-and quadratic and models is $\Delta {\rm BIC} = ZZ.Z$, strongly
+and quadratic and models is $\Delta {\rm BIC} = 89.1$, strongly
 favoring the quadratic case.
@@
 \dot{P}
-  = - (X.XX \pm 0.XX)\times 10^{-10}
-  = - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}.
+  = - (2.74 \pm 0.28)\times 10^{-10}
+  = - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
 \end{equation}
-This agrees with the value reported by \citet{southworth_transit_2019}
-after their including an additional 22 archival transit measurements
-($\dot{P} = - 9.2 \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$).
-It is smaller than the rate of period decrease found by
-\citetalias{bouma_wasp4b_2019} ($- 12.6 \pm 1.2~{\rm ms}\,{\rm
-yr}^{-1}$)., presumably because of missing coverage filled in by
-\citeauthor{southworth_transit_2019}'s observations.
-The other best-fit transit timing model parameters are reported in
-{\bf Table~X}.
+This agrees to within $1\sigma$ of the value reported by
+\citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
+ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
+of period decrease we found earlier \citep[$- 12.6 \pm 1.2~{\rm
+ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of
+missing coverage filled in by \citeauthor{southworth_transit_2019}'s
+and \citeauthor{baluev_2019}'s observations.  The other best-fit
+transit timing model parameters are reported in {\bf Table~X}.
 
@@
 
-%FIXME: is the purple here from my Pdot, or Southworth's?
 %FIXME: perhaps include the WASP-4b phase-fold here too??
@@
     The trend that would be needed to produce the period decrease seen
-    in transits ($\dot{P}= - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}$) is
+    in transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is
     indicated with the purple dotted line.  The four new RV
@@
 In other words, the majority of the period decrease observed in
-transits ($\dot{P} = - X.X \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$) seems to
+transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) seems to
 be explained by the acceleration of the host star.
-% FIXME
-All the
-available TTV data show a period decrease of $\approx 8-12$
-milliseconds per year~\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.
 

### 2020-01-27 paper/collect.sh

@@ -1,4 +1,4 @@
-cp ../results/O_minus_C_includeallpoints_onlytransits.pdf f1.pdf
+cp ../results/O_minus_C_20200127_onlytransits.pdf f1.pdf
 cp ../results/20190911_rv_fit.pdf f2.pdf
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png f3a.png
 cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832.pdf f3b.pdf

### 2020-01-28 src/plot_O_minus_C_paperII.py
plot improvements

@@ -90,7 +90,7 @@ def plot_O_minus_C(
     include_all_points=False,
     x_extra=None, y_extra=None, sigma_y_extra=None,
     x_occ=None, y_occ=None, sigma_y_occ=None,
-    onlytransits=False, theta_quad_merr=None, theta_quad_perr=None):
+    onlytransits=False, perrs_3d=None, merrs_3d=None):
 
     xfit = np.linspace(10*np.min(x), 10*np.max(x), 10000)
 
@@ -147,30 +147,87 @@ def plot_O_minus_C(
     if include_all_points:
         raise NotImplementedError
 
-    a0.plot(xfit,
-            quadratic_fit(theta_quadratic, xfit)
-                - linear_fit(theta_linear, xfit),
-            zorder=-1, color='C0')
+    #a0.plot(xfit,
+    #        quadratic_fit(theta_quadratic, xfit)
+    #            - linear_fit(theta_linear, xfit),
+    #        zorder=-1, color='C0')
     a0.plot(xfit,
             linear_fit(theta_linear, xfit)
                 - linear_fit(theta_linear, xfit),
             zorder=-3, color='gray')
 
+    #
+    # Do the error bands
+    #
+    sigma_t0 = max(
+        [perrs_3d['t0 [min]'], merrs_3d['t0 [min]']]
+    )
+    sigma_P = max(
+        [perrs_3d['P [min]'], merrs_3d['P [min]']]
+    )
+    sigma_quad = max(
+        [perrs_3d['0.5 dP/dE [min]'], merrs_3d['0.5 dP/dE [min]']]
+    )
+
+    model_tmid_upper = np.maximum.reduce([
+        (
+            (theta_quadratic[0]+sigma_t0) +
+            (theta_quadratic[1]+sigma_P)*xfit +
+            (theta_quadratic[2]+sigma_quad)*xfit**2
+        ),
+        (
+            (theta_quadratic[0]+sigma_t0) +
+            (theta_quadratic[1]-sigma_P)*xfit +
+            (theta_quadratic[2]+sigma_quad)*xfit**2
+        ),
+        (
+            (theta_quadratic[0]+sigma_t0) +
+            (theta_quadratic[1]+sigma_P)*xfit +
+            (theta_quadratic[2]-sigma_quad)*xfit**2
+        ),
+        (
+            (theta_quadratic[0]+sigma_t0) +
+            (theta_quadratic[1]-sigma_P)*xfit +
[…]

### 2020-01-28 src/plot_mass_semimaj_constraints.py

@@ -4,6 +4,7 @@ from numpy import array as nparr
 from cdips.plotting import savefig
 import matplotlib as mpl
 from astropy import units as u, constants as c
+import matplotlib.patheffects as pe
 
 def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
                                   sma_grid=None, with_contrast=False,
@@ -11,8 +12,8 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     if prob_arr is None:
 
-        n_mass_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
-        n_sma_grid_edges = 51 # a 4x4 grid has 5 edges. want: 51
+        n_mass_grid_edges = 129 # a 4x4 grid has 5 edges. want: 51
+        n_sma_grid_edges = 129 # a 4x4 grid has 5 edges. want: 51
         n_injections_per_cell = 512 # 500 # want: 500
 
         mass_grid = (
@@ -54,11 +55,14 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     if not linear_z:
 
+        cutoff = -12
+
         if discrete_color:
-            bounds = np.round(np.linspace(-10, np.log(prob_arr).max(), 5),1)
+            bounds = np.round(np.linspace(cutoff, np.log(prob_arr).max(), 5),1)
             norm = mpl.colors.BoundaryNorm(bounds, cmap.N)
         else:
-            norm = mpl.colors.Normalize(vmin=np.log(prob_arr).max(), vmax=-10)
+            norm = mpl.colors.Normalize(vmin=np.log(prob_arr).max(),
+                                        vmax=cutoff)
 
         im = ax.pcolormesh(X, Y, np.log(prob_arr), cmap=cmap, norm=norm,
                            shading='flat') # vs 'gouraud'
@@ -66,10 +70,11 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
     else:
 
         if discrete_color:
-            bounds = np.linspace(1e-10, prob_arr.max(), 5)
+            bounds = np.linspace(10**(cutoff), prob_arr.max(), 5)
             norm = mpl.colors.BoundaryNorm(bounds, cmap.N)
         else:
-            norm = mpl.colors.Normalize(vmin=prob_arr.max(), vmax=1e-10)
+            norm = mpl.colors.Normalize(vmin=prob_arr.max(),
[…]

### 2020-01-28 src/plot_zorro_speckle.py
add plot_zorro_speckle

@@ -1,0 +1,51 @@
+import os, pickle
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from numpy import array as nparr
+from cdips.plotting import savefig
+import matplotlib as mpl
+from astropy import units as u, constants as c
+import matplotlib.patheffects as pe
+import matplotlib.image as mpimg
+
+def plot_zorro_speckle():
+
+    datapath = (
+        '../data/WASP4_zorro_speckle/WASP-4_20190928_832_companionbounds.csv'
+    )
+    df = pd.read_csv(datapath)
+
+    fig, ax = plt.subplots(figsize=(4,3))
+
+    ax.plot(
+        df.ang_sep, df.delta_mag, color='k'
+    )
+
+
+    img = mpimg.imread(
+        '../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png'
+    )
+
+    # [left, bottom, width, height]
+    inset = fig.add_axes([0.55, 0.23, .4, .4])
+    inset.imshow(img)
+    inset.axis('off')
+    plt.setp(inset, xticks=[], yticks=[])
+
+    ax.set_ylabel('$\Delta$m (832 nm)')
+    ax.set_xlabel('Angular separation [arcsec]')
+
+    ax.get_yaxis().set_tick_params(which='both', direction='in')
+    ax.get_xaxis().set_tick_params(which='both', direction='in')
+    ax.tick_params(right=True, which='both', direction='in')
+    ax.tick_params(top=True, which='both', direction='in')
+    fig.tight_layout(h_pad=0, w_pad=0)
+
+    figpath = (
+        '../results/zorro_speckle.png'
+    )
+    savefig(fig, figpath)
+
+
+if __name__ == "__main__":
+
+    plot_zorro_speckle()

### 2020-01-28 paper/ms.tex
update paper

@@
   The new data show that the system is accelerating towards the Earth
-  at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm s}^{-1}\,{\rm
-  yr}^{-1}$.
+  at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
+  s}^{-1}\,{\rm yr}^{-1}$.
   The corresponding period decrease explains WASP-4's
@@
 the TRAPPIST archive, and a select few other observers.
-% \citeauthor{baluev_2019} analyzed their newly obtained
-% photometry, along with archival light curves that we omitted from our
-% analysis due to systematic uncertainties in the absolute time system.
-% \citeauthor{baluev_2019} found that when they used all the available
-% TTV data, the need for a quadratic ephemeris was present.  However,
-% they pointed out that if they used lower-precision subsets of the
-% available timing data, the necessity for the quadratic term decreased.
-% This is not particularly surprising.
-% \citeauthor{baluev_2019} also pointed out that the
-% precise transit times reported by \citet{huitson_gemini_2017} were
-% quite important in the time-series.  Overall,
-% \citeauthor{baluev_2019} did not find the claim of a
-% period decrease convincing.
 
@@
 measurements reveal a line-of-sight acceleration of $\dot{\gamma} =
--XX {\rm m\,s^{-1}\,yr^{-1}}$.  This translates to an expected period
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,yr^{-1}}$.  This
+translates to an expected period
 decrease simply from the light travel time effect of ZZ milliseconds
@@
   \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-  The vertical axis shows the observed times minus the calculated
+  The vertical axis shows the observed transit times minus the calculated
   times assuming a constant orbital period.  More opaque points correspond
-  to more precise data.  The quadratic ephemeris and its $1\sigma$
+  to more precise data.  The quadratic ephemeris and its $\pm1\sigma$
   statistical uncertainties are shown in blue.  Neither model fully
-  explains the transit residuals (see~\ref{sec:transit_analysis}).
+  explains the scatter in transit residuals, perhaps due to
+  underestimated uncertainties (see~\ref{sec:transit_analysis}).
   The binned TESS point (yellow star) is the weighted average of 18
-  TESS transits and is for display purposes only.  Its uncertainty is
+  TESS transits and is binned for display purposes only.  Its uncertainty is
   slightly larger than the marker size.  The models were fitted to all
@@
 \begin{align}
-  t_{\rm tra}(E) &= t_0 + PE,\\
+  t_{\rm tra}(E) &= t_0 + PE,
 \end{align}
@@
     measurements presented in this work increased the significance of
-    the slope from $\approx$2$\sigma$ to 15$\sigma$.
+    the radial velocity trend from $\approx$2$\sigma$ to 15$\sigma$.
 	\label{fig:rvs}
@@
 \begin{equation}
-  \dot{v}_{\rm r} =
+  \dot{v}_{\rm r} = \dot{\gamma} = 
      -0.0422^{+0.0028}_{-0.0027}
@@
 
-%TODO: describe 3b better
-%TODO: also, put 3b in as the bottom right inset to the contrast curve
 \begin{figure}[!t]
-	\gridline{
-		\fig{f3a.png}{0.33\textwidth}{}
-	}
-	\vspace{-1cm}
-	\gridline{
-		\fig{f3b.pdf} {0.5\textwidth}{}
-	}
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.5\textwidth]{f3.pdf}
+	\end{center}
 	\vspace{-0.7cm}
     \caption{
-      {\it Top:\,} 
-      Speckle image reconstructed from 1000 40 millisecond frames in
-      an 832 nm bandpass, and acquired on September 28, 2019.  The
-      image scale is $2.46''\times2.46''$.
-      {\it Bottom:\,} 
-      Contrast curve derived from point-source injection-recovery
-      experiments.
+      {\bf Zorro contrast curve derived from point-source
+      injection-recovery experiments.}
+      Sources below the curve would have been detected.
+      The inset shows the speckle image reconstructed from 1000 40
+      millisecond frames in an 832 nm bandpass, and acquired on
+      September 28, 2019.  The image scale is $2.46''\times2.46''$.
     }
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f4.pdf}
+		\includegraphics[width=0.78\textwidth]{f4.pdf}
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.5cm}
 	\caption{
@@
 	}
-\end{figure}
+\end{figure*}
 
@@
 
-We began by defining a $75\times50$ grid in true planetary mass and
+We began by defining a $128\times128$ grid in true planetary mass and
 semimajor axis, with even logarithmic spacing from 1 to 900$\,$$M_{\rm
@@
 that an additional companion in any particular cell could explain the
-observed linear trend.  In each cell, we simulated 500 hypothetical
+observed linear trend.  In each cell, we simulated 512 hypothetical
 companions.
@@
 argument of periastron of the outer simulated companion.  We converted
-the resulting $75\times50\times500$ cube of best-fit log-likelihood
+the resulting $128\times128\times512$ cube of best-fit log-likelihood
 values to probabilities, and averaged over the samples in each grid
@@
 orbital parameters and minimum mass of the companion.
+
+% TODO: add a paragraph noting that while the skepticism
+% is fine "I was right".
+
+% \citeauthor{baluev_2019} analyzed their newly obtained
+% photometry, along with archival light curves that we omitted from our
+% analysis due to systematic uncertainties in the absolute time system.
+% \citeauthor{baluev_2019} found that when they used all the available
+% TTV data, the need for a quadratic ephemeris was present.  However,
+% they pointed out that if they used lower-precision subsets of the
+% available timing data, the necessity for the quadratic term decreased.
+% This is not particularly surprising.
+% \citeauthor{baluev_2019} also pointed out that the
+% precise transit times reported by \citet{huitson_gemini_2017} were
+% quite important in the time-series.  Overall,
+% \citeauthor{baluev_2019} did not find the claim of a
+% period decrease convincing.
 

### 2020-01-28 paper/collect.sh

@@ -1,6 +1,5 @@
 cp ../results/O_minus_C_20200127_onlytransits.pdf f1.pdf
 cp ../results/20190911_rv_fit.pdf f2.pdf
-cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832_img.png f3a.png
-cp ../data/WASP4_zorro_speckle/WASP-4_20190928_832.pdf f3b.pdf
+cp ../results/zorro_speckle.pdf f3.pdf
 cp ../results/mass_semimaj_constraints_with_contrast_discretecolor.pdf f4.pdf
 cp ../results/k14_pdot.pdf f5.pdf

### 2020-01-28 src/plot_mass_semimaj_constraints.py
minor

@@ -120,7 +120,7 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
     ax.set_ylabel('Companion mass [M$_\mathrm{{jup}}$]')
 
     cbar = fig.colorbar(im, orientation='vertical', extend='min',
-                        label='log(likelihood)')
+                        label='$\ln$(likelihood)')
     cbar.ax.tick_params(labelsize=6)
 
     ax.set_ylim([1,1e3])

### 2020-01-28 paper/ms.tex
update the paper

@@
-% TODO:
-%
-% * email to full collaborator list
-% * add any missing CPS coauthors
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% TODO: may want to improve title
 \title{WASP-4 is Accelerating Towards the Earth}
+% ", Explaining its Decreasing Orbital Period"
 % ": A Cautionary Tale in the Search for Tidal Orbital Decay"
@@
 %
-% key authors: did Rachel do the Zorro reductions?
+% key authors:
 %
@@
 %
-\author[0000-0002-0531-1073]{H. Isaacson}
-\affiliation{Astronomy Department, University of California, Berkeley,
-CA 94720, USA}
-%
 \author[0000-0001-8638-0320]{A. W. Howard}
@@
 %
+\author[0000-0002-0531-1073]{H. Isaacson}
+\affiliation{Astronomy Department, University of California, Berkeley,
+CA 94720, USA}
+%
 \author{H. Knutson}
@@
 %
-%
-%\author{R. A. Matson}
-% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\author{R. A. Matson}
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
 %
@@
   WASP-4b have recently shown that its orbital period is decreasing.
-  Potential causes of the period change include tidal orbital decay,
-  orbital precession, and light-travel time effects.
-  In this work we present new radial velocity measurements of WASP-4,
-  acquired with Keck-HIRES.
-  The new data show that the system is accelerating towards the Earth
-  at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
-  s}^{-1}\,{\rm yr}^{-1}$.
-  The corresponding period decrease explains WASP-4's
-  changing orbital period as a light-travel time effect.
-  Combining the radial velocities with new speckle imaging limits, we
-  find that system's acceleration is likely caused by a
-  10-200$\,M_{\rm Jup}$ companion with semi-major axis between 10-100$\,$AU.
-  Based on statistics from \citet{knutson_friends_2014}, about 1 in 6
-  hot Jupiters are expected to show period decreases commensurate with
-  WASP-4 due to acceleration by outer companions.
-  Continued long-term radial velocity monitoring of hot Jupiters is
-  therefore essential to distinguish tidal orbital decay from
-  line-of-sight accelerations.
+  The period change could be caused by tidal orbital decay, orbital
+  precession, or light-travel time effects.  In this work we present
+  new radial velocity measurements of WASP-4, acquired with
+  Keck-HIRES.  The data show that the system is accelerating towards
+  the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
+  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  This implied period decrease
+  probably explains WASP-4's changing orbital period as a light-travel
+  time effect.  By combining the radial velocities with new speckle
+  imaging limits, we find that system's acceleration is likely caused
+  by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis between
+  10-100$\,$AU.  Based on statistics from
+  \citet{knutson_friends_2014}, about 1 in 6 hot Jupiters are expected
+  to show period decreases commensurate with WASP-4 due to
+  acceleration by outer companions.  Continued long-term radial
+  velocity monitoring of hot Jupiters is therefore essential to
+  distinguish tidal orbital decay from line-of-sight accelerations.
 \end{abstract}
@@
 The main focus of this study is the hot Jupiter WASP-4b, which has an
-orbital period that from transits appears to be decreasing by about 10
-milliseconds per year.  We discovered the timing variations using
-data from TESS and a decade of ground-based monitoring
-\citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.
-Thereafter, \citet{southworth_transit_2019} reported 22 new transit
-times for the system, and confirmed that the updated series of transit
-times was consistent with a quadratic ephemeris.  With additional
-data, they found a lower best-fit decay rate of $\dot{P} = -XX.X \pm
-X.X$ milliseconds per year.
-A separate study by \citet{baluev_2019} reported
-additional archival transit light curves of WASP-4b, most notably from
-the TRAPPIST archive, and a select few other observers.
+orbital period that appears to be decreasing by about 10 milliseconds
+per year.  We discovered the timing variations using data from TESS
+and a decade of ground-based observations \citep[][hereafter
+\citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.  Thereafter,
+\citet{southworth_transit_2019} reported 22 new transit times for the
+system, and found an updated decay rate of $\dot{P} = -9.2 \pm 1.1$
+milliseconds per year. The \citeauthor{southworth_transit_2019} decay rate
+was $\approx$3$\sigma$ less rapid than that found by
+\citetalias{bouma_wasp4b_2019}, but the conclusions of the
+studies were otherwise the same.  A separate study by
+\citet{baluev_2019} reported additional archival transit light curves
+of WASP-4b, most notably from the TRAPPIST archive, as well as from
+select other observers.  \citet{baluev_2019} pointed out that
+when using lower-precision subsets of the available transit data,
+the case for a decreasing period worsened.  We agree with this
+statement.
 
@@
 additional radial velocity measurements using Keck-HIRES, extending
-the RV baseline from 3 to 9 years.  Previously, the radial velocities
-were sparse and showed a weak ($<2\sigma$) linear trend.  Our new
-measurements reveal a line-of-sight acceleration of $\dot{\gamma} =
+the RV baseline from 3 to 9 years.  Previously, the five available
+HIRES radial velocities suggested a weak ($<2\sigma$) linear trend
+\citep{knutson_friends_2014}.  Our new measurements reveal a
+line-of-sight acceleration of $\dot{\gamma} =
 -0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,yr^{-1}}$.  This
-translates to an expected period
-decrease simply from the light travel time effect of ZZ milliseconds
-per year --- about commensurate with what is observed.
-In the following, we avoid the term ``R{\o}mer delay'' as this phrase
-typically signifies arrival time delay due to observatory motion. We
-are discussing an analog of the Doppler effect for a source with
-constant line-of-sight acceleration.
+translates to an expected period decrease from the light-travel time
+effect of -5.9 milliseconds per year---about commensurate with what
+is observed from transits.  In the following, we avoid the term
+``R{\o}mer delay'', which typically signifies arrival time
+delay due to observatory motion. We are discussing the Doppler effect
+seen by a stationary observer for a source with constant line-of-sight
+acceleration.
 
-% FIXME: figure out correct sections
-\S~\ref{sec:observations} presents the new observations...
-\S~\ref{sec:analysis} describes our analysis...
-\S~\ref{sec:discussion} discusses...
-\S~\ref{sec:conclusions} gives conclusions...
+In the following, \S~\ref{sec:observations} collects the available
+transit data and presents the new radial velocity and speckle imaging
+observations.  \S~\ref{sec:analysis} analyzes the data, and finds that
+they yield a picture in which the WASP-4 system is accelerating
+towards our line-of-sight, likely due to the pull of a brown or
+M-dwarf companion.
+\S~\ref{sec:discussion} places this result in the broader context of
+orbital decay searches, and points out that line-of-sight
+accelerations, {\it i.e.}, ``false positive orbital decay signals'',
+are relatively common for the hot Jupiter population.
+\S~\ref{sec:conclusions} offers brief concluding remarks.
 
@@
 
-{\bf Table~N} lists the transit times we used in our analysis.  We
-include data from the peer-reviewed literature for which (a) the
+{\bf Table~N} lists the transit times we collected for our analysis.
+We include data from the peer-reviewed literature for which (a) the
 analysis was based on observations of a single transit, (b) the
@@
 observed from the 3.58m NTT and Danish 1.54m telescopes at La Silla,
-and the SAAO 1.0m telescope.  Additional timimg
-measurements have also recently been reported by \citet{baluev_2019},
-based on a homogeneous analysis of archival ground-based observations
-taken professional and amateur observers.  We included
-twelve of their ``high quality'' transit times acquired by TRAPPIST
-(six transits), El~Sauce (four transits), and
+and the SAAO 1.0m telescope.
+
+Additional timimg measurements have also recently been reported by
+\citet{baluev_2019}, based on a homogeneous analysis of archival
+ground-based observations taken professional and amateur observers.
+We included twelve of their ``high quality'' transit times acquired by
+TRAPPIST (six transits), El~Sauce (four transits), and
 \citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
@@
 of a linear trend in the residuals after fitting out the orbit of
-WASP-4b.  This prompted us to acquire speckle imaging using Zorro at
-Gemini-South \citep[see][, and the instrument
+WASP-4b.  This prompted us to acquire speckle images using Zorro at
+Gemini-South \citep[see][and the instrument
 web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
-Zorro (and its counterpart, 'Alopeke, on Gemini-North) are
-dual-channel speckle interferometers that can be used to observe in
-narrow-band filters centered at 562$\,$nm and 832$\,$nm.
-
-We observed WASP-4 twice, on the night of September 11-12 (8 sets of
-1000 frames each) and also on the night of September 28-29 (7 sets of
-1000 frames each). The second night had better seeing and produced a
-slightly better result, which we opt to use for our analysis.
+Zorro is a dual-channel speckle interferometer with narrow-band
+filters centered at 562$\,$nm and 832$\,$nm.
 
@@
 % were derived?
-{\bf We reduced the speckle images to contrast curves by ...}
+We observed WASP-4 twice, on the night of September 11-12 (8 sets of
+1000 frames each) and also on the night of September 28-29 (7 sets of
+1000 frames each).
+{\bf We reconstructed the images following...}
+{\bf We reduced the reconstructed speckle images to contrast curves by
+injecting and recovering point-sources...}
+The contrast curves showed that the second night, which had better
+seeing, also produces the more constraining result.
+The 832$\,$nm results were also typically the most constraining, since
+faint companions are typically redder.
+We therefore opt for the 832$\,$nm September 28-29 results for the
+remaining analysis.
 
@@
   \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-  The vertical axis shows the observed transit times minus the calculated
-  times assuming a constant orbital period.  More opaque points correspond
-  to more precise data.  The quadratic ephemeris and its $\pm1\sigma$
-  statistical uncertainties are shown in blue.  Neither model fully
-  explains the scatter in transit residuals, perhaps due to
-  underestimated uncertainties (see~\ref{sec:transit_analysis}).
-  The binned TESS point (yellow star) is the weighted average of 18
-  TESS transits and is binned for display purposes only.  Its uncertainty is
-  slightly larger than the marker size.  The models were fitted to all
-  of the individual transit times.
+  The vertical axis shows the observed transit times minus the
+  calculated times assuming a constant orbital period.  More opaque
+  points correspond to more precise data.  The quadratic ephemeris and
+  its $\pm1\sigma$ statistical uncertainties are shown in blue.
+  Neither model fully explains the scatter in transit residuals,
+  perhaps due to underestimated uncertainties
+  (see~\ref{sec:transit_analysis}).  The binned TESS point (yellow
+  star) is the weighted average of 18 TESS transits and is binned for
+  display purposes only.  Its uncertainty is slightly larger than the
+  marker size.  The models were fitted to all of the individual
+  transit times.
   \label{fig:times}
@@
 
-We considered two possible models for the observed transit times.  The
-first model assumes a constant orbital period on a circular orbit:
+We considered two models for the observed transit times.  The first
+model assumes a constant orbital period on a circular orbit:
 \begin{align}
@@
 where $E$ is the epoch number and $t_0$ is a reference epoch.  The
-second model assumes the period changes at a steady rate:
+second model assumes that the period changes at a steady rate:
 \begin{align}
@@
 The free parameters are the reference epoch $t_0$, the period at the
-reference epoch, and the period derivative, ${\rm d}P/{\rm d}t = (1/P)
-{\rm d}P/{\rm d}E$.  We defined the epoch numbers such that $E=0$ is
-near the weighted average of the observed times.  This helps to reduce
-the covariance between $t_0$ and $P$.  A third possible model that we
-did not consider for reasons that will become apparent is a
-precessing, eccentric orbit \citep[{\it
+reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
+(1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
+$E=0$ is near the weighted average of the observed times.  This helps
+to reduce the covariance between $t_0$ and $P$.  A third possible
+model that we did not consider for reasons that will become apparent
+is a precessing, eccentric orbit \citep[{\it
 e.g.},][]{gimenez_revision_1995,patra_2017}.
@@
 red}=3.0$.  The best-fitting quadratic model has $\chi^2 = 183$, 90
-degrees of freedom, and $\chi^2_{\rm red}=2.0$.  From the reduced
-$\chi^2$ values, we can surmise that neither model entirely describes
-the transit data that we have collected---there must be some
-additional source of signal or potentially noise. In
+degrees of freedom, and $\chi^2_{\rm red}=2.0$.  The difference in the
+Bayesian information criteria (BIC) between the linear and quadratic
+and models is $\Delta {\rm BIC} = 89.1$, strongly favoring the
+quadratic case \citep{kass_bayes_1995}.
+
+From the reduced $\chi^2$ values, we can surmise that neither model
+entirely describes the transit data that we have collected---there
+must be some additional source of signal or noise. In
 \citet{bouma_wasp4b_2019}, we found that the quadratic model for the
-(less complete) transit data gave $\chi^2_{\rm red}=1.0$.  The only
-difference between our old and new transit datasets are the addition
-of transit times from \citet{southworth_transit_2019} and
-\citet{baluev_2019}.  The worsened $\chi^2_{\rm red}$ could reflect
-underestimated statistical uncertainties in these or in fact any of
-the transit measurements. It could also reflect systematic errors in
-the time-systems in which the transit measurements were recorded,
-though we have taken every caution against the latter possibility.
-% For purposes of simplicity, we simply accept that the additional data
-% have effectively increased the scatter in the transit O-C diagram.
-The difference in the BIC \citep{kass_bayes_1995} between the linear
-and quadratic and models is $\Delta {\rm BIC} = 89.1$, strongly
-favoring the quadratic case.
+(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsened
+$\chi^2_{\rm red}$ could reflect underestimated statistical
+uncertainties in any of our transit measurements.  It could also
+reflect systematic errors in the time-systems in which the transit
+measurements were recorded, though we have taken every caution against
+this latter possibility.
+
+%FIXME: might just do the example calculation?
+One approach to analyzing such data would be to introduce a fudge
+factor that inflates the transit measurement uncertainties, and lowers
+the reduced $\chi^2$.  We do not think that such an approach is
+warranted, since it would not change the result that the quadratic
+model is strongly preferred.
 
@@
 
-%FIXME: perhaps include the WASP-4b phase-fold here too??
 \begin{figure}[t]
@@
 	\caption{
-    {\bf Radial velocity observations of WASP-4.} The best-fit Keplerian orbit
-    of WASP-4b has been subtracted.  The linear trend inferred from the
-    RV data is shown with a black line, with $1\sigma$ errors in gray.
-    The trend that would be needed to produce the period decrease seen
-    in transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is
-    indicated with the purple dotted line.  The four new RV
-    measurements presented in this work increased the significance of
-    the radial velocity trend from $\approx$2$\sigma$ to 15$\sigma$.
+    {\bf Radial velocity observations of WASP-4.} The best-fit
+    Keplerian orbit of WASP-4b has been subtracted.  The linear trend
+    inferred from the RV data is shown with a black line, with
+    $1\sigma$ errors in gray.  The trend that would be needed to
+    produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
+    0.89~{\rm ms}\,{\rm yr}^{-1}$) is indicated with the purple dotted
+    line.  The four new RV measurements presented in this work
+    increased the significance of the radial velocity trend from
+    $\approx$2$\sigma$ to 15$\sigma$.
 	\label{fig:rvs}
@@
 
-%FIXME: this entire paragraph copies Bouma+19 too heavily.
-We began by fitting a single Keplerian orbit, plus instrument offsets,
-jitters, and a long-term trend
+Our initial model for the radial velocity data was a single Keplerian
+orbit, plus instrument offsets, jitters, and a long-term trend
 \citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
-on the period and time of inferior conjunction using the values from
-Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed the eccentricity
-to zero, per the results of \citet{beerer_secondary_2011},
-\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The free
-parameters were the velocity semi-amplitude, the instrument
-zero-points, the instrument jitters (an additive white noise term for
-each instrument), linear ($\dot{v_{\rm r}}$), and optionally
-second-order ($\ddot{v_{\rm r}}$) acceleration terms.
+on the orbital period and time of inferior conjunction using the
+values from Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed
+WASP-4b's eccentricity to zero
+\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
+The free parameters were the velocity semi-amplitude, the instrument
+zero-points, an additive ``white noise'' instrument jitter for each
+instrument, linear ($\dot{v_{\rm r}}$), and optionally second-order
+($\ddot{v_{\rm r}}$) acceleration terms.
 
@@
 \end{equation}
-For comparison, before our recent observing run, 
-$\dot{v}_{\rm r}$ was thought to be about five times smaller, and was
-only marginally significant
-\citep{knutson_friends_2014,bouma_wasp4b_2019}.
+For comparison, before our new measurements, $\dot{v}_{\rm r}$ was
+thought to be about five times smaller, and was only marginally
+significant \citep{knutson_friends_2014,bouma_wasp4b_2019}.
 
 The system's acceleration towards our line-of-sight causes a decrease
-in the apparent orbital period.  The period
-derivative expected from radial velocities is
+in the apparent orbital period.  The period derivative expected from
+radial velocities is
 \begin{align}
-  \dot{P}_{\rm RV} &= \frac{\dot{v}_{\rm r} P}{c} \\
-  \dot{P}_{\rm RV} &= -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
+  \dot{P}_{\rm RV}
+  = \frac{\dot{v}_{\rm r} P}{c}
+  = -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
 \end{align}
@@
 
-An important consideration is whether the measured RV trend is at all
+An important consideration is whether the measured RV trend is
 correlated with stellar activity.  We investigate this by analyzing
@@
 orbit-subtracted velocities.  We found a correlation coefficient of
-0.16. Though suggestive, it is not statistically significant (the
-corresponding $p$-value is 0.65).  Furthermore, inspection of the
+0.16. This correlation is not statistically significant; the
+corresponding $p$-value is 0.65.  Furthermore, inspection of the
 $S$-index timeseries showed no secular or sinusoidal trends, as would
@@
 The $S$-index values are included in {\bf Table X}.  We conclude that
-it would be highly unlikely for the linear trend to be caused by
+it is highly unlikely that the linear trend is caused by
 stellar activity.
@@
       {\bf Zorro contrast curve derived from point-source
-      injection-recovery experiments.}
-      Sources below the curve would have been detected.
-      The inset shows the speckle image reconstructed from 1000 40
-      millisecond frames in an 832 nm bandpass, and acquired on
-      September 28, 2019.  The image scale is $2.46''\times2.46''$.
+      injection-recovery experiments.} Sources below the curve would
+      have been detected.  The inset shows the speckle image
+      reconstructed from 1000 40 millisecond frames in an 832 nm
+      bandpass, and acquired on September 28, 2019.  The image scale
+      is $2.46''\times2.46''$.
     }
@@
 	\end{center}
-	\vspace{-0.5cm}
+	\vspace{-0.6cm}
 	\caption{
@@
     The likelihood inferred from radial velocities is shown in
-    grayscale, and the region excluded from the imaging is shown with
-    a cross-hatch pattern.
+    grayscale, and the region excluded from the speckle imaging is
+    indicated (blue line).  The expected mass to semimajor-axis
+    degeneracy is shown with a black line.
 	\label{fig:mass_sma}
-  \vspace{-0.3cm}
+  \vspace{-0cm}
 	}
@@
 bolometric luminosities from the \citet{baraffe_evolutionary_2003} and
-MIST isochrones.  Using these theoretical quantities and the empircal
-Zorro bandpass functions, we calculated absolute magnitudes in the 562
-and 832 nm Zorro bandpasses for stellar and planetary mass companions.
-Applying the same calculation to WASP-4 itself using the effective
-temperature and bolometric luminosity from
+MIST isochrones.  Using these theoretical quantities and the
+empirically-measured Zorro bandpasses, we calculated absolute
+magnitudes in the 562 and 832 nm Zorro bands for stellar and planetary
+mass companions.  Applying the same calculation to WASP-4 itself using
+the effective temperature and bolometric luminosity from
 \citetalias{bouma_wasp4b_2019}, we derived the transformation from
-contrast ratio to companion mass.  The resulting limits are shown as
-the cross-hatched region in Figure~\ref{fig:mass_sma}.
+contrast ratio to companion mass.  The resulting limits are shown in
+Figure~\ref{fig:mass_sma}.
 
@@
 
-%FIXME: may want to verify that changing the cutoff location, to say
-%80Mjup or whatever the minimum BD mass is doesn't drastically alter
-%the posterior
 We assigned each companion a mass and semimajor axis from log-uniform
@@
 Equation~17.  We emphasize that the long-period exoplanet and
-long-period binary distributions are quite different: the exoplanet
-distribution is ``bottom-heavy'', with eccentricities preferentially
-close to zero. The binary star distribution is ``top-heavy'', with
-more eccentricities close to one.  We chose a mass cutoff of
-$10\,M_{\rm Jup}$ based on the bound observed by
-\citet{schlaufman_evidence_2018} between giant planets, presumably
-formed through core accretion, and brown dwarfs, presumably formed via
-gravitational instability.
+long-period binary eccentricity distributions are quite different: the
+exoplanet distribution is ``bottom-heavy'', with eccentricities
+preferentially close to zero. The binary star distribution is
+``top-heavy'', with eccentricities closer to one.  We chose a mass
+cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
+bound observed by \citet{schlaufman_evidence_2018} between giant
+planets, presumably formed through core accretion, and brown dwarfs,
+presumably formed via gravitational instability.
 
@@
 chains of our initial model of WASP-4b, plus its linear trend. We
-subtracted the planet's orbital solution, leaving the linear trend.
-Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulated outer
-companion, and the fixed instrument offsets and jitters from the MCMC
-chains, we then performed a maximum likelihood fit for the time and
-argument of periastron of the outer simulated companion.  We converted
-the resulting $128\times128\times512$ cube of best-fit log-likelihood
-values to probabilities, and averaged over the samples in each grid
-cell to derive a probability distribution in mass and semi-major axis.
-
-The result is shown as grayscale background in
-Figure~\ref{fig:mass_sma}.
+subtracted the planet's orbital solution, leaving RV data points with
+a linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each
+simulated outer companion, and the fixed instrument offsets and
+jitters from the MCMC chains, we then performed a maximum likelihood
+fit for the time and argument of periastron of the outer simulated
+companion.  We converted the resulting $128\times128\times512$ cube of
+best-fit log-likelihood values to probabilities, and averaged over the
+samples in each grid cell to derive a probability distribution in mass
+and semi-major axis.  Figure~\ref{fig:mass_sma} shows the result in
+grayscale.
 
@@
 \subsection{Implications for WASP-4}
-Previous possibilities suggested to explain the period decrease of
-WASP-4b included tidal orbital decay, orbital precession, and
+Previous potential explanations for
+WASP-4b's decreasing orbital period included tidal orbital decay, orbital precession, and
 light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
-velocity measurements have now shown the least exotic
-option---light-travel time effects---is also the most likely
-explanation.  Transits show the obital period to decrease by $YY.Y \pm
-Z.Z~{\rm ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed
-in radial velocities implies an expected period decrease of $YY.Y \pm
-Z.Z~{\rm ms}\,{\rm yr}^{-1}$.  
+velocity measurements strongly indicate that the least exotic
+option---light-travel time effects---is also the most likely.  Transits show the obital period 
+decreasing by 
+$- 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$; the line-of-sight
+acceleration observed
+in radial velocities would predict a period decrease of
+$-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$.
 Though the quantitative agreement is not perfect, Occam's razor would
@@
 star, and that it is between $10$-$100\,{\rm AU}$ from the host star
-(Figure~\ref{fig:mass_sma}).
-Given such a mass, this companion could at one time have influenced
-the orbital evolution of the inner giant.
+(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
+at one time have influenced the orbital evolution of the inner giant.
+The fact that most hot Jupiters have similar massive outer
+companions \citep{knutson_friends_2014,bryan_statistics_2016} is
+circumstantial evidence for certain high-eccentricity formation
+channels (see \citealt{dawson_johnson_2018}).
 Further radial velocity monitoring should eventually reveal the
-orbital parameters and minimum mass of the companion.
+orbital parameters and minimum mass of WASP-4's companion.
 
-% TODO: add a paragraph noting that while the skepticism
-% is fine "I was right".
-
-% \citeauthor{baluev_2019} analyzed their newly obtained
-% photometry, along with archival light curves that we omitted from our
-% analysis due to systematic uncertainties in the absolute time system.
+% \citeauthor{baluev_2019} analyzed their newly obtained photometry,
+% along with archival light curves that we omitted from our analysis
+% due to systematic uncertainties in the absolute time system.
 % \citeauthor{baluev_2019} found that when they used all the available
@@
 % they pointed out that if they used lower-precision subsets of the
-% available timing data, the necessity for the quadratic term decreased.
-% This is not particularly surprising.
-% \citeauthor{baluev_2019} also pointed out that the
-% precise transit times reported by \citet{huitson_gemini_2017} were
-% quite important in the time-series.  Overall,
-% \citeauthor{baluev_2019} did not find the claim of a
-% period decrease convincing.
+% available timing data, the necessity for the quadratic term
+% decreased.  This is not particularly surprising.
+% \citeauthor{baluev_2019} also pointed out that the precise transit
+% times reported by \citet{huitson_gemini_2017} were quite important
+% in the time-series.  Overall, \citeauthor{baluev_2019} did not find
+% the claim of a period decrease convincing.
 
@@
 We identified WASP-4b's decreasing orbital period as part of a search
-for tidal orbital decay.  However, over half of hot Jupiters have
-companions outside of $1\,{\rm AU}$ with super-Jovian masses
+for tidal orbital decay.  However, most hot Jupiters have companions
+outside of $1\,{\rm AU}$ with super-Jovian masses
 \citep{knutson_friends_2014}.  Line-of-sight accelerations are
-therefore relatively common in hot Jupiter systems.
+correspondingly relatively common in hot Jupiter systems.
 
 To evaluate the importance of these effects in future transit timing
-searches, we collected the linear trends reported by
+searches, we collected the linear radial velocity trends reported by
 \citet{knutson_friends_2014}, and computed the expected orbital period
 derivatives $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
-systems.
-The results are given in {\bf Table~X}, and visualized for hot
-Jupiters with significant linear trends in Figure~\ref{fig:pdot_pop}.
+systems.  The results are given in {\bf Table~X}, and visualized for
+hot Jupiters with significant linear trends in
+Figure~\ref{fig:pdot_pop}.
 
-Including WASP-4b, 16 of 51 hot Jupiters from
+Including WASP-4b, 16 of 51 hot Jupiters surveyed by
 \citet{knutson_friends_2014} show a non-zero radial velocity trend.
-About half are positive, and half are negative, and therefore
-$\approx$1 in 6 hot Jupiters are expected to show period decreases
-commensurate with WASP-4 due to acceleration by outer companions.
+About half are positive, and half are negative.  Therefore around 1 in
+3 hot Jupiters are expected to show period changes commensurate with
+WASP-4 due to acceleration by outer companions.  Half of these will be
+period decreases, and will be an astrophysical false positives in
+searches for tidal orbital decay.
+
+For hot Jupiters that have been monitored for over 10 years, secular
+changes in the orbital period be measured to a precision of $\lesssim
+10~{\rm ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020,
+submitted).
+
+%FIXME
+As observations from TESS continue...
+As the Transiting Exoplanet
+
+Continued long-term radial
+  velocity monitoring of hot Jupiters is therefore essential to
+  distinguish tidal orbital decay from line-of-sight accelerations.
 
@@
 
-We found that WASP-4 is accelerating towards the Earth, at
-X.XX m/s/yr.
-This seems like the most plausible explanation for the orbital 
-period decrease observed in transits~---~not tidal
-orbital decay, or apsidal precession.
-The most likely parameters for the companion are XXX.
-Over half of hot Jupiters have super-Jovian outer companions,
-so the light-travel time effect will become an increasingly large nuisance in
-the hunt for orbital decay as
-the observing baselines for hot Jupiters increase.
-The requirement to avoid missing out is to have long-term
-radial velocity monitoring with the same spectrograph (or else to
-calibrate instrumental offsets by observing the same stars simultaneously).
+From newly acquired radial velocity
+measurements, we found that WASP-4 is accelerating towards the Earth at
+$\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
+s}^{-1}\,{\rm yr}^{-1}$.  The corresponding light-travel time effect
+predicts a period decrease $\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm
+ms}\,{\rm yr}^{-1}$.  The majority of the period decrease observed in
+transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is
+probably explained by the acceleration of the host star~---~not tidal
+orbital decay, or apsidal precession.  The companion causing the
+acceleration is most likely a brown dwarf or low-mass star, and its
+semi-major axis is likely 10-100$\,$AU.
+
+Over half of hot Jupiters have outer companions with masses larger
+than Jupiter outside of 1$\,$AU. The accelerations and period changes
+induced by these outer companions will become an increasingly large
+nuisance in the hunt for orbital decay as the observational baselines
+get longer.  This issue can be alleviated through long-term radial
+velocity monitoring of hot Jupiters.
 

### 2020-01-29 paper/ms.tex
update paper
derived a scaling relation

@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently shown that its orbital period is decreasing.
+  WASP-4b have recently indicated that its orbital period is decreasing.
   The period change could be caused by tidal orbital decay, orbital
@@
   probably explains WASP-4's changing orbital period as a light-travel
-  time effect.  By combining the radial velocities with new speckle
-  imaging limits, we find that system's acceleration is likely caused
-  by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis between
-  10-100$\,$AU.  Based on statistics from
+  time effect.  By combining the radial velocities with new limits
+  from speckle imaging, we find that system's acceleration is likely
+  caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
+  between 10-100$\,$AU.  Based on statistics from
   \citet{knutson_friends_2014}, about 1 in 6 hot Jupiters are expected
   to show period decreases commensurate with WASP-4 due to
-  acceleration by outer companions.  Continued long-term radial
-  velocity monitoring of hot Jupiters is therefore essential to
-  distinguish tidal orbital decay from line-of-sight accelerations.
+  acceleration by outer companions.  These period decreases will be
+  increasingly measureable in coming years, because the precision of
+  period derivative measurements scales as the fourth power of the
+  observing baseline.  Continued long-term radial velocity monitoring
+  of hot Jupiters is therefore essential to distinguish tidal orbital
+  decay from line-of-sight accelerations.
 \end{abstract}
@@
     \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
+  \label{eq:quadratic_model}
 \end{align}
@@
 
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+  {\bf Predicted hot Jupiter period changes from linear radial
+  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
+  \citet{knutson_friends_2014} have shown long-term radial velocity
+  trends.  The HAT-P-11 signal is plotted, though its radial velocity
+  signal may be caused by stellar activity.  Three hot Jupiters are
+  not plotted because their radial velocity curves better described as
+  quadratic trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects
+  are ordered in the $y$ dimension by the absolute value of d$P$/d$t$.
+	\label{fig:pdot_pop}
+  \vspace{-0.3cm}
+	}
+\end{figure}
+
 We identified WASP-4b's decreasing orbital period as part of a search
@@
 
-For hot Jupiters that have been monitored for over 10 years, secular
-changes in the orbital period be measured to a precision of $\lesssim
-10~{\rm ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020,
-submitted).
+\subsection{At what rate is the measurement precision of d$P$/d$t$
+increasing?}
+\label{sec:fisher}
 
-%FIXME
-As observations from TESS continue...
-As the Transiting Exoplanet
+For hot Jupiters that have been monitored for over baselines $\tau$ of
+$\approx$10 years, secular changes in their orbital periods are
+currently being constrained to a precision of $\lesssim 10~{\rm
+ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
+This is roughly commensurate with level of signal many outer
+companions are expected to induce from radial velocity accelerations
+(Figure~\ref{fig:pdot_pop}).
 
-Continued long-term radial
-  velocity monitoring of hot Jupiters is therefore essential to
-  distinguish tidal orbital decay from line-of-sight accelerations.
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-  {\bf Predicted hot Jupiter period changes from linear radial
-  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
-  \citet{knutson_friends_2014} have shown long-term radial velocity
-  trends.  The HAT-P-11 signal is plotted, though its radial velocity
-  signal may be caused by stellar activity.  Three hot Jupiters are
-  not plotted because their radial velocity curves better described as
-  quadratic trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects
-  are ordered in the $y$ dimension by the absolute value of d$P$/d$t$.
-	\label{fig:pdot_pop}
-  \vspace{-0.3cm}
-	}
-\end{figure}
+It is therefore naturally of interest to ask at what point in time
+additional detections of the light-travel time effect can become
+common-place among hot Jupiters.  This question is the same as
+asking: at what rate does the uncertainty in the quadratic term of
+Equation~\label{eq:quadratic_model} scale with the observing baseline?
+We take a Fisher analysis approach to the problem, following the notes
+by \citet{gould_chi2_2003}.
+Rewriting Equation~\ref{eq:quadratic_model} in simpler terms,
+\begin{align}
+  t_{\rm tra} = a_0 + a_1\cdot E + a_2\cdot E^2,
+\end{align}
+where $a_0=t_0$, $a_1=P$, $a_2=0.5\cdot {\rm d}P/{\rm d}E$.
+One can then show that if $N$ transit time measurements are taken at
+uniform times across a baseline of duration $\tau$ with constant
+precision $\sigma_{\rm tra}$, then the uncertainty of the quadratic
+term is given by
+\begin{align}
+  \sigma_{a_2} =
+  \left(\frac{25{,}920}{N}\right)^{1/2} \frac{\sigma_{\rm tra}}{\tau^4}
+\end{align}
+This result implies that if a decade passes, doubling the observing
+baseline, we gain a factor of 16 in precision on d$P$/d$t$.  Assuming
+regular observations continue from ground-based telescopes (CITE) and
+space-based observatories (CITE TESS, CITE CHEOPS), this will enable
+measurements of period derivatives precise to less than $1~{\rm
+ms}\,{\rm yr}^{-1}$ before 2030.
+%FIXME the units above make no sense????
 
@@
 
-From newly acquired radial velocity
-measurements, we found that WASP-4 is accelerating towards the Earth at
-$\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
-s}^{-1}\,{\rm yr}^{-1}$.  The corresponding light-travel time effect
-predicts a period decrease $\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm
-ms}\,{\rm yr}^{-1}$.  The majority of the period decrease observed in
-transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is
-probably explained by the acceleration of the host star~---~not tidal
-orbital decay, or apsidal precession.  The companion causing the
-acceleration is most likely a brown dwarf or low-mass star, and its
-semi-major axis is likely 10-100$\,$AU.
+From newly acquired radial velocity measurements, we found that WASP-4
+is accelerating towards the Earth at $\dot{\gamma} =
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.
+The corresponding light-travel time effect predicts a period decrease
+$\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$.  The
+majority of the period decrease observed in transits ($\dot{P}= - 8.64
+\pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is probably explained by the
+acceleration of the host star~---~not tidal orbital decay, or apsidal
+precession.  The companion causing the acceleration is most likely a
+brown dwarf or low-mass star, and its semi-major axis is likely
+10-100$\,$AU.
 
 Over half of hot Jupiters have outer companions with masses larger
-than Jupiter outside of 1$\,$AU. The accelerations and period changes
-induced by these outer companions will become an increasingly large
-nuisance in the hunt for orbital decay as the observational baselines
-get longer.  This issue can be alleviated through long-term radial
+than Jupiter that orbit beyond 1$\,$AU
+\citep{knutson_friends_2014,bryan_statistics_2016}. The accelerations
+and period changes induced by these outer companions will become an
+increasingly large nuisance in the hunt for orbital decay as the
+observational baselines get longer.  In particular, the precision with
+which the period derivative can be measured scales as the fourth power
+of the baseline duration (\S~\ref{sec:fisher}), and so within a decade
+many more hot Jupiters should show orbital period changes due to
+accelerations induced by their outer companions.  This effect can be
+distinguished from tidal orbital decay through long-term radial
 velocity monitoring of hot Jupiters.
-
-%TODO: what's the scaling? sigma_P ~ 1/baseline? baseline^2?
 

### 2020-01-29 paper/ms.tex
update paper

--- 2020-01-28 paper/ms.tex
@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently shown that its orbital period is decreasing.
-  The period change could be caused by tidal orbital decay, orbital
-  precession, or light-travel time effects.  In this work we present
+  WASP-4b have recently indicated that its orbital period is
+  decreasing.  The period change could be caused by tidal orbital
+  decay, orbital precession, or light-travel time effects.  We present
   new radial velocity measurements of WASP-4, acquired with
@@
   the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
-  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  This implied period decrease
+  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease
   probably explains WASP-4's changing orbital period as a light-travel
-  time effect.  By combining the radial velocities with new speckle
-  imaging limits, we find that system's acceleration is likely caused
-  by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis between
-  10-100$\,$AU.  Based on statistics from
+  time effect.  Combining the radial velocities with new limits from
+  speckle imaging, we find that the system's acceleration is likely
+  caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
+  between 10-100$\,$AU.  Using statistics from
   \citet{knutson_friends_2014}, about 1 in 6 hot Jupiters are expected
-  to show period decreases commensurate with WASP-4 due to
-  acceleration by outer companions.  Continued long-term radial
-  velocity monitoring of hot Jupiters is therefore essential to
-  distinguish tidal orbital decay from line-of-sight accelerations.
+  to show similar period decreases as WASP-4, due to acceleration by
+  outer companions.  We find that these period decreases will be
+  increasingly measureable in coming years, because the precision of
+  period derivative measurements scales as the fourth power of the
+  observing baseline.  Continued radial velocity monitoring
+  of hot Jupiters is therefore essential to distinguish tidal orbital
+  decay from line-of-sight accelerations.
 \end{abstract}
@@
 
-The issue of whether and how the orbits of hot Jupiters decay due to
-tidal interaction with their host stars has drawn significant
-attention over the past 25 years
-\citep{rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
-Numerous observations have also been performed (CITE).  These efforts
-are now yielding fruit: by combining transit and occultation timing
-with radial velocity measurements, \citet{yee_orbit_2020} have shown
-that the data for the hot Jupiter WASP-12b are most compatible with
-orbital decay.
+The orbits of most hot Jupiters are unstable to tidal decay
+\citep{counselman_outcomes_1973,hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
+The relevant issue is whether the timescale for tidal orbital decay is
+shorter or longer than the timescale for main-sequence stellar
+evolution.  This question depends on the uncertain rate at which
+friction inside the star can damp the energy of tidal oscillations and
+thereby shrink the orbit (as reviewed by \citealt{Mazeh2008} and
+\citealt{ogilvie_tidal_2014}).
 
-The present study highlights a point that, though obvious, has perhaps
-not yet received due attention.  The point is that observational
-programs aimed at identifying orbital decay in hot Jupiters will be
-crippled without long-term radial velocity monitoring programs.  The
-reason is simple: when the timescale of orbital period change exceeds
-the observing baseline, the first deviation from constant periodicity
-is always quadratic in the transit number.
+Indirect studies of age and angular momentum indicators including the
+hot Jupiter semi-major axis distribution, host star spin rates, and
+galactic velocity dispersions have led to estimates for the
+inspiral timescale that vary from much less to much greater than
+the main-sequence evolution time  ({\it e.g.},
+\citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
+\citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
+\citealt{hamer_schlaufman_2019}).  Direct
+measurements of tidal orbital decay through transit and occultation
+timing could provide an empirical solution.  For instance, combined
+transit timing and radial velocity measurements for WASP-12b have
+shown that its secular period decrease of $\approx$30 milliseconds per
+year is almost certainly due to tidal orbital decay
+\citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
 
-While orbital decay may be common (CITE).  line-of-sight accelerations
-could be even {\it more} common (CITE Knutson 14, Bryan 16).  Both
-manifest to first order identically in transit times (up to the sign
-of the line-of-sight acceleration) -- as a non-zero period derivative.
-Specifically, \citet{knutson_friends_2014} showed that X\% of hot
-Jupiters show some form of radial acceleration in radial velocity
-time-series with baselines of YY years.  Assuming a separation
-distribution of (WHATEVER), this implies that ZZ\% of hot Jupiters
-will {\it always} show a changing orbital period, over timescale of
-say 10 years.
-
-% The abundance of giant planets
-% exterior to Super-Earth sized systems is also high (CITE Bryan 2019),
-% but this is not so much of a concern to us, as their planetary masses
-% and therefore tidal timescales are typically expected to be at least
-% an order of magnitude shorter than hot Jupiters.
+This study highlights a point that, though obvious, has perhaps not
+yet received due attention.  The point is that observational programs
+aimed at identifying orbital decay in hot Jupiters through transit
+timing will be crippled without concurrent long-term radial velocity
+monitoring.  The reason is that line-of-sight accelerations due to
+outer companions \citep[{\it e.g.},][]{agol_detecting_2005} and tidal
+orbital decay both initially manifest identically in transit times, as
+a non-zero period derivative.  Massive outer companions to hot
+Jupiters are the norm; \citet{bryan_statistics_2016} calculated an
+occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
+Jupiters with masses from 1-20$M_{\rm Jup}$ and semi-major axes from
+5-100$\,$AU.  Therefore it could very well be that more hot Jupiters
+will show shrinking orbital periods due to long term accelerations
+than due to tidal orbital decay.
 
@@
 per year.  We discovered the timing variations using data from TESS
+\citep{ricker_transiting_2015}
 and a decade of ground-based observations \citep[][hereafter
@@
 system, and found an updated decay rate of $\dot{P} = -9.2 \pm 1.1$
-milliseconds per year. The \citeauthor{southworth_transit_2019} decay rate
-was $\approx$3$\sigma$ less rapid than that found by
-\citetalias{bouma_wasp4b_2019}, but the conclusions of the
-studies were otherwise the same.  A separate study by
-\citet{baluev_2019} reported additional archival transit light curves
-of WASP-4b, most notably from the TRAPPIST archive, as well as from
-select other observers.  \citet{baluev_2019} pointed out that
-when using lower-precision subsets of the available transit data,
-the case for a decreasing period worsened.  We agree with this
-statement.
+milliseconds per year. The \citeauthor{southworth_transit_2019} decay
+rate was $\approx$3$\sigma$ less rapid than that found by
+\citetalias{bouma_wasp4b_2019}, but the conclusions of the studies
+were otherwise similar.  A separate study by \citet{baluev_2019}
+reported additional archival transit light curves of WASP-4b, most
+notably from TRAPPIST, and also from select other
+observers.  \citet{baluev_2019} pointed out that when using
+lower-precision subsets of the available transit data, the case for a
+decreasing period worsened.  We agree with this statement.
 
@@
 the RV baseline from 3 to 9 years.  Previously, the five available
-HIRES radial velocities suggested a weak ($<2\sigma$) linear trend
+HIRES radial velocities suggested a weak ($\approx$$2\sigma$) linear trend
 \citep{knutson_friends_2014}.  Our new measurements reveal a
@@
 translates to an expected period decrease from the light-travel time
-effect of -5.9 milliseconds per year---about commensurate with what
-is observed from transits.  In the following, we avoid the term
-``R{\o}mer delay'', which typically signifies arrival time
-delay due to observatory motion. We are discussing the Doppler effect
-seen by a stationary observer for a source with constant line-of-sight
+effect of -5.9 milliseconds per year---about commensurate with what is
+observed from transits.  In the following, we avoid the term
+``R{\o}mer delay'', which typically signifies arrival time delay due
+to observatory motion. We are discussing the Doppler effect seen by a
+stationary observer for a source with constant line-of-sight
 acceleration.
@@
 towards our line-of-sight, likely due to the pull of a brown or
-M-dwarf companion.
-\S~\ref{sec:discussion} places this result in the broader context of
-orbital decay searches, and points out that line-of-sight
-accelerations, {\it i.e.}, ``false positive orbital decay signals'',
-are relatively common for the hot Jupiter population.
-\S~\ref{sec:conclusions} offers brief concluding remarks.
+M-dwarf companion.  \S~\ref{sec:discussion} places this result in the
+broader context of orbital decay searches, and points out that
+line-of-sight accelerations, {\it i.e.}, ``false positive orbital
+decay signals'', are relatively common in the hot Jupiter population.
+\S~\ref{sec:conclusions} offers concluding remarks.
 
@@
 
-{\bf Table~N} lists the transit times we collected for our analysis.
-We include data from the peer-reviewed literature for which (a) the
-analysis was based on observations of a single transit, (b) the
-midpoint was fitted as a free parameter, and (c) the time system
-specified both the leap second correction (TDB or UTC) and also
+Table~1 lists the transit times we collected for our analysis.  We
+include data from the peer-reviewed literature for which {\it (i)} the
+analysis was based on observations of a single transit, {\it (ii)} the
+midpoint was fitted as a free parameter, and {\it (iii)} the time
+system specified both the leap second correction (TDB or UTC) and also
 whether any barycentric or heliocentric corrections had been
@@
 
-Additional timimg measurements have also recently been reported by
+Additional timing measurements were also recently made available by
 \citet{baluev_2019}, based on a homogeneous analysis of archival
 ground-based observations taken professional and amateur observers.
-We included twelve of their ``high quality'' transit times acquired by
+We included twelve of their ``high quality'' transit times from
 TRAPPIST (six transits), El~Sauce (four transits), and
@@
 
-The complete set of radial velocity observations is given in {\bf
-Table~M}.  Along with the 2010-2019 HIRES observations, there are also
+The complete set of radial velocity observations is given in 
+Table~2.  Along with the 2010-2019 HIRES observations, there are also
 earlier measurements from CORALIE and HARPS.  Following
@@
 We considered two models for the observed transit times.  The first
-model assumes a constant orbital period on a circular orbit:
+model assumes a constant orbital period $P$ on a circular orbit:
 \begin{align}
@@
 best-fit constant period model.  The best-fitting constant-period
-model has $\chi^2 = 276$, 91 degrees of freedom, and $\chi^2_{\rm
-red}=3.0$.  The best-fitting quadratic model has $\chi^2 = 183$, 90
-degrees of freedom, and $\chi^2_{\rm red}=2.0$.  The difference in the
+model has 91 degrees of freedom, $\chi^2 = 276$,  and $\chi^2_{\rm
+red}=3.0$.  The best-fitting quadratic model has 90 degrees of freedom,
+$\chi^2 = 183$, and $\chi^2_{\rm red}=2.0$.  The difference in the
 Bayesian information criteria (BIC) between the linear and quadratic
@@
 
-%FIXME: might just do the example calculation?
 One approach to analyzing such data would be to introduce a fudge
-factor that inflates the transit measurement uncertainties, and lowers
-the reduced $\chi^2$.  We do not think that such an approach is
-warranted, since it would not change the result that the quadratic
-model is strongly preferred.
+factor to inflate the transit measurement uncertainties, and lower the
+reduced $\chi^2$.  We do not think that such an approach is warranted,
+since it would not change the result that the quadratic model is
+strongly preferred.
 
@@
 ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
-of period decrease we found earlier \citep[$- 12.6 \pm 1.2~{\rm
+of period decrease we reported earlier \citep[$- 12.6 \pm 1.2~{\rm
 ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of
-missing coverage filled in by \citeauthor{southworth_transit_2019}'s
-and \citeauthor{baluev_2019}'s observations.  The other best-fit
-transit timing model parameters are reported in {\bf Table~X}.
+the new data from \citeauthor{southworth_transit_2019}
+and \citeauthor{baluev_2019}  The other best-fit
+transit timing model parameters are reported in Table~3.
 
@@
 	\caption{
-    {\bf Radial velocity observations of WASP-4.} The best-fit
+    {\bf Radial velocity residuals of WASP-4.} The best-fit
     Keplerian orbit of WASP-4b has been subtracted.  The linear trend
@@
 \begin{align}
-  \dot{P}_{\rm RV}
+  \dot{P}_{\rm\,RV}
   = \frac{\dot{v}_{\rm r} P}{c}
@@
 be expected if we were observing a long-term magnetic activity cycle.
-The $S$-index values are included in {\bf Table X}.  We conclude that
+The $S$-index values are included in Table~2.  We conclude that
 it is highly unlikely that the linear trend is caused by
@@
 where $\tau$ is the observing baseline.
-For WASP-4, this yields $M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher mass
-companions are allowed, presuming that they orbit further from the
-star; at fixed $\dot{\gamma}$, $M_{\rm comp} \propto a^2$
+For WASP-4, this yields $M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher masses
+are allowed for companions that orbit further from the
+star: at fixed $\dot{\gamma}$, $M_{\rm comp} \propto a^2$
 \citep{torres_substellar_1999,liu_crossing_2002}.
@@
 of synthetic photometry, we further assumed that all sources had
-blackbody spectra. While this is clearly false, we do not readily have
+blackbody spectra. While this is a bold simplification, we do not readily have
 access to the planetary and stellar atmosphere models needed for the
@@
 cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
-bound observed by \citet{schlaufman_evidence_2018} between giant
-planets, presumably formed through core accretion, and brown dwarfs,
-presumably formed via gravitational instability.
+bound observed by \citet{schlaufman_evidence_2018} between
+giant planets and brown dwarfs, though this value is also close to the
+$13\,M{\rm Jup}$ required for a solar-metallicity body to
+significantly burn deuterium \citep{burrows_nongray_1997}.
 
 For each simulated companion, we then drew a sample from the converged
-chains of our initial model of WASP-4b, plus its linear trend. We
-subtracted the planet's orbital solution, leaving RV data points with
+chains of our initial model of WASP-4b. We
+subtracted the planet's orbital solution, leaving RV points with
 a linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each
@@
 companion.  We converted the resulting $128\times128\times512$ cube of
-best-fit log-likelihood values to probabilities, and averaged over the
+log-likelihood values to probabilities, and averaged over the
 samples in each grid cell to derive a probability distribution in mass
-and semi-major axis.  Figure~\ref{fig:mass_sma} shows the result in
-grayscale.
+and semi-major axis.  Figure~\ref{fig:mass_sma} shows the result.
 
@@
 \subsection{Implications for WASP-4}
-Previous potential explanations for
-WASP-4b's decreasing orbital period included tidal orbital decay, orbital precession, and
+Previous potential explanations for WASP-4b's decreasing orbital
+period included tidal orbital decay, orbital precession, and
 light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
 velocity measurements strongly indicate that the least exotic
-option---light-travel time effects---is also the most likely.  Transits show the obital period 
-decreasing by 
-$- 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$; the line-of-sight
-acceleration observed
-in radial velocities would predict a period decrease of
-$-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$.
-Though the quantitative agreement is not perfect, Occam's razor would
-suggest that the line of sight acceleration is probably a sufficient
-explanation for the apparent decrease of WASP-4b's orbital period.
+option---light-travel time effects---is also the most likely.
+Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
+ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
+velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
+ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
+perfect, Occam's razor would suggest that the line of sight
+acceleration is probably a sufficient explanation for the apparent
+decrease of WASP-4b's orbital period.
 % While further radial velocity observations of the system should help
@@
 acceleration are that it is likely either a brown-dwarf or low mass
-star, and that it is between $10$-$100\,{\rm AU}$ from the host star
+star, orbiting between $10$-$100\,{\rm AU}$ from the host star
 (Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
@@
 
-We identified WASP-4b's decreasing orbital period as part of a search
-for tidal orbital decay.  However, most hot Jupiters have companions
-outside of $1\,{\rm AU}$ with super-Jovian masses
-\citep{knutson_friends_2014}.  Line-of-sight accelerations are
-correspondingly relatively common in hot Jupiter systems.
-
-To evaluate the importance of these effects in future transit timing
-searches, we collected the linear radial velocity trends reported by
-\citet{knutson_friends_2014}, and computed the expected orbital period
-derivatives $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
-systems.  The results are given in {\bf Table~X}, and visualized for
-hot Jupiters with significant linear trends in
-Figure~\ref{fig:pdot_pop}.
-
-Including WASP-4b, 16 of 51 hot Jupiters surveyed by
-\citet{knutson_friends_2014} show a non-zero radial velocity trend.
-About half are positive, and half are negative.  Therefore around 1 in
-3 hot Jupiters are expected to show period changes commensurate with
-WASP-4 due to acceleration by outer companions.  Half of these will be
-period decreases, and will be an astrophysical false positives in
-searches for tidal orbital decay.
-
-For hot Jupiters that have been monitored for over 10 years, secular
-changes in the orbital period be measured to a precision of $\lesssim
-10~{\rm ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020,
-submitted).
-
-%FIXME
-As observations from TESS continue...
-As the Transiting Exoplanet
-
-Continued long-term radial
-  velocity monitoring of hot Jupiters is therefore essential to
-  distinguish tidal orbital decay from line-of-sight accelerations.
-
 \begin{figure}[t]
@@
 
+We identified WASP-4b's decreasing orbital period as part of a search
+for tidal orbital decay.  However, most hot Jupiters have companions
+outside of $5\,{\rm AU}$ with super-Jovian masses
+\citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight accelerations are
+correspondingly common in hot Jupiter systems.
+
+To evaluate the importance of these effects for future transit timing
+analyses, we collected the linear radial velocity trends reported by
+\citet{knutson_friends_2014}, and computed the expected orbital period
+derivatives $\dot{P}_{\rm\,RV} = \dot{v}_{\rm r} P/c$ for each
+system.  The results are given in Table~4, and visualized for
+hot Jupiters with significant ($>$$3\sigma$) linear trends in
+Figure~\ref{fig:pdot_pop}.
+
+Including WASP-4b, 16 of 51 hot Jupiters surveyed by
+\citet{knutson_friends_2014} show a non-zero radial velocity trend.
+Therefore around 1 in
+3 hot Jupiters is expected to show period changes commensurate with
+WASP-4 due to acceleration by outer companions.  Half of these will be
+period decreases, and will be an astrophysical false positive in
+searches for tidal orbital decay.
+
+\subsection{At what rate is the measurement precision of d$P$/d$t$
+increasing?}
+\label{sec:fisher}
+
+For hot Jupiters that have been monitored for over baselines of
+$\approx$10 years, secular changes in their orbital periods are
+currently being constrained to a precision of $\lesssim 10~{\rm
+ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
+This is roughly commensurate with the level of signal many outer
+companions are expected to induce (Figure~\ref{fig:pdot_pop}).
+
+It is therefore pertinent to ask at what point in time further
+detections of the light-travel time effect will become routine for hot
+Jupiters.  This question is the same as asking: at what rate does the
+uncertainty in the quadratic term of Equation~\ref{eq:quadratic_model}
+scale with the observing baseline?  We take a Fisher analysis approach
+to the problem.
+First, we rewrite Equation~\ref{eq:quadratic_model} as
+\begin{align}
+  t_{\rm tra} = a_0 + a_1 E + a_2 E^2,
+\end{align}
+where $a_0\equiv t_0$, $a_1\equiv P$, and $a_2\equiv 0.5\cdot {\rm d}P/{\rm d}E$.  
+By following \citet{gould_chi2_2003},
+we derived
+that if $N$ transit timing measurements are taken uniformly
+across a baseline of $\Delta E$ epochs with constant precision
+$\sigma_{\rm tra}$, then the uncertainty of the quadratic term is
+given by
+\begin{align}
+  \sigma_{a_2} =
+  \left(\frac{25{,}920}{N}\right)^{1/2}
+  \frac{\sigma_{\rm tra}}{(\Delta E)^4}
+\end{align}
+This result implies that a doubled observing baseline yields
+a sixteen-fold improvement in precision on d$P$/d$t$.  If regular
+observations continue from ground and space-based observatories,
+period derivatives with precision below $1~{\rm ms}\,{\rm yr}^{-1}$
+will be measureable before 2030.
+
 \section{Conclusions}
@@
 
-From newly acquired radial velocity
-measurements, we found that WASP-4 is accelerating towards the Earth at
-$\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
-s}^{-1}\,{\rm yr}^{-1}$.  The corresponding light-travel time effect
-predicts a period decrease $\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm
-ms}\,{\rm yr}^{-1}$.  The majority of the period decrease observed in
-transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is
-probably explained by the acceleration of the host star~---~not tidal
-orbital decay, or apsidal precession.  The companion causing the
-acceleration is most likely a brown dwarf or low-mass star, and its
-semi-major axis is likely 10-100$\,$AU.
+From newly acquired radial velocity measurements, we found that WASP-4
+is accelerating towards the Earth at $\dot{\gamma} =
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.
+The corresponding light-travel time effect predicts a period decrease
+$\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$.  The
+majority of the period decrease observed in transits ($\dot{P}= - 8.64
+\pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is therefore explained by the
+acceleration of the host star~---~not tidal orbital decay, or apsidal
+precession.  The companion causing the acceleration is most likely a
+brown dwarf or low-mass star with semi-major axis between
+10-100$\,$AU.
 
-Over half of hot Jupiters have outer companions with masses larger
-than Jupiter outside of 1$\,$AU. The accelerations and period changes
-induced by these outer companions will become an increasingly large
-nuisance in the hunt for orbital decay as the observational baselines
-get longer.  This issue can be alleviated through long-term radial
-velocity monitoring of hot Jupiters.
-
-%TODO: what's the scaling? sigma_P ~ 1/baseline? baseline^2?
+Most hot Jupiters have outer companions with masses larger than
+Jupiter beyond 5$\,$AU
+\citep{knutson_friends_2014,bryan_statistics_2016}. The accelerations
+and period changes induced by these outer companions will become an
+increasingly large nuisance in the hunt for orbital decay as the
+observational baselines get longer.  In particular, the precision with
+which the period derivative can be measured from transits scales as
+the fourth power of the baseline duration (\S~\ref{sec:fisher}), and
+so within a decade many more hot Jupiters should show orbital period
+changes due to accelerations induced by their outer companions.  This
+effect can be distinguished from tidal orbital decay through long-term
+radial velocity monitoring of hot Jupiters.
 
@@
 
-%\clearpage
-%\newpage
-\input{model_fit_table.tex}
-
 \bibliographystyle{yahapj}                            
@@
 
+\input{WASP-4b_transit_time_table.tex}
+\input{WASP-4b_rv_table.tex}
+\input{model_fit_table.tex}
+\input{pdot_table.tex}
+
 \end{document}

### 2020-01-29 paper/pdot_table.tex
update paper

\startlongtable
\begin{deluxetable*}{lllllllll}
%
\tablecaption{
	Predicted hot Jupiter period changes from linear radial
	velocity trends reported by \citet{knutson_friends_2014}.
\label{tab:pdot_table}
}
%
\tablenum{4}
%
% planet;gammadot;gammadot_pluserr;gammadot_minuserr;comment;pl_name;pl_orbper;Pdot;Pdot_upper_limit;Pdot_lower_limit;Pdot_perr;Pdot_merr;abs_Pdot;K14_significant
\tablehead{
  \colhead{Planet} &
  \colhead{$\dot{\gamma}$ [m$\,$s$^{-1}$$\,$yr$^{-1}$]} &
  \colhead{$+\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
  \colhead{$-\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
  \colhead{$P$ [days]} &
  \colhead{$\dot{P}_{\,{\rm RV}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{$+\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{$-\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{Significant?}
}
%
%FIXME precision
\startdata
HAT-P-2 b & -0.0938 & 0.0067 & 0.0069 & 5.6335158 & -55.62 & 3.97 & 4.09 & 1
\enddata
%
\tablecomments{
  Table~4 is published in its entirety in a machine-readable format.
  A single row is shown here for guidance regarding form and content.
}
%
\end{deluxetable*}

### 2020-01-29 src/abbreviate_the_bibliography.py
add abbreviate_the_bibliography

@@ -1,0 +1,34 @@
+"""
+Zotero generates long journal names. Abbreviate them in bulk to conform to the
+AAS style guide.
+"""
+
+from collections import OrderedDict
+
+mdict = OrderedDict({})
+
+mdict['The Astrophysical Journal Letters'] = r'\apjl'
+mdict['The Astrophysical Journal Supplement Series'] = r'\apjs'
+mdict['The Astronomical Journal'] = r'\aj'
+mdict['Annual Review of Astronomy and Astrophysics'] = r'\araa'
+mdict['The Astrophysical Journal'] = r'\apj'
+mdict['Publications of the Astronomical Society of the Pacific'] = r'\pasp'
+mdict['Astronomy and Astrophysics'] = r'\aap'
+mdict['Monthly Notices of the Royal Astronomical Society: Letters'] = r'\mnras:l'
+mdict['Monthly Notices of the Royal Astronomical Society'] = r'\mnras'
+mdict['Nature'] = r'\nat'
+
+with open('bibliography.bib', 'r') as f:
+    lines = f.readlines()
+
+for k,v in mdict.items():
+    for ix, l in enumerate(lines):
+        if k in l:
+            _newline = l.replace(k, v)
+            lines[ix] = _newline
+
+outpath = 'abbrevbib.bib'
+with open(outpath, 'w') as f:
+    f.writelines(lines)
+
+print('made {}'.format(outpath))

### 2020-01-29 src/plot_mass_semimaj_constraints.py
tweaking

@@ -95,7 +95,7 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
         ax.text(
             100, 450, 'Ruled out by\nspeckle imaging',
             fontsize=7.5, ha='center', va='center',
-            path_effects=[pe.withStroke(linewidth=0.8, foreground="white")],
+            path_effects=[pe.withStroke(linewidth=0.5, foreground="white")],
             color='C0'
         )
 
@@ -109,7 +109,7 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
         ax.text(
             np.percentile(_sma, 60), np.percentile(_mass, 40), '$M\propto a^2$',
             fontsize=7.5, ha='left', va='center',
-            path_effects=[pe.withStroke(linewidth=0.8, foreground="white")],
+            path_effects=[pe.withStroke(linewidth=0.5, foreground="white")],
             color='black'
         )
 

### 2020-01-29 src/plot_zorro_speckle.py

@@ -31,7 +31,7 @@ def plot_zorro_speckle():
     inset.axis('off')
     plt.setp(inset, xticks=[], yticks=[])
 
-    ax.set_ylabel('$\Delta$m (832 nm)')
+    ax.set_ylabel('$\Delta$mag (832 nm)')
     ax.set_xlabel('Angular separation [arcsec]')
 
     ax.get_yaxis().set_tick_params(which='both', direction='in')

### 2020-01-29 paper/ms.tex
contd fixing of the bibliography
tweaking

@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently indicated that its orbital period is
+  WASP-4b have recently shown that its orbital period is
   decreasing.  The period change could be caused by tidal orbital
@@
   caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
-  between 10-100$\,$AU.  Using statistics from
-  \citet{knutson_friends_2014}, about 1 in 6 hot Jupiters are expected
+  between 10-100$\,$AU.  The statistics from
+  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters are expected
   to show similar period decreases as WASP-4, due to acceleration by
@@
 corresponding $p$-value is 0.65.  Furthermore, inspection of the
-$S$-index timeseries showed no secular or sinusoidal trends, as would
+$S$-index timeseries did not show secular or sinusoidal trends, as would
 be expected if we were observing a long-term magnetic activity cycle.
@@
       have been detected.  The inset shows the speckle image
-      reconstructed from 1000 40 millisecond frames in an 832 nm
+      reconstructed from 1000 40 millisecond frames in an 832$\,$nm
       bandpass, and acquired on September 28, 2019.  The image scale
@@
   \texttt{astrobase} \citep{bhatti_astrobase_2018},
-	\texttt{astroplan} \citep{astroplan2018},
-  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
+  % \texttt{astroplan} \citep{astroplan2018},
+  \texttt{astropy} \citep{astropy_2018},
   \texttt{astroquery} \citep{astroquery_2018},
@@
 
+%
+% The following are entries from Table 1 that are not otherwise cited
+% in the text
+%
+\nocite{wilson_wasp-4b_2008}
+\nocite{gillon_improved_2009}
+\nocite{winn_transit_2009}
+\nocite{hoyer_tramos_2013}
+\nocite{dragomir_terms_2011}
+\nocite{sanchis-ojeda_starspots_2011}
+\nocite{nikolov_wasp-4b_2012}
+\nocite{ranjan_atmospheric_2014}
+\nocite{huitson_gemini_2017}
+
 \bibliographystyle{yahapj}                            

### 2020-01-29 paper/ms.tex
fix tex justification

@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently shown that its orbital period is
-  decreasing.  The period change could be caused by tidal orbital
-  decay, orbital precession, or light-travel time effects.  We present
-  new radial velocity measurements of WASP-4, acquired with
-  Keck-HIRES.  The data show that the system is accelerating towards
-  the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
-  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease
-  probably explains WASP-4's changing orbital period as a light-travel
-  time effect.  Combining the radial velocities with new limits from
+  WASP-4b have recently shown that its orbital period is decreasing.
+  The period change could be caused by tidal orbital decay, orbital
+  precession, or light-travel time effects.  We present new radial
+  velocity measurements of WASP-4, acquired with Keck-HIRES.  The data
+  show that the system is accelerating towards the Earth at
+  $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
+  s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease probably
+  explains WASP-4's changing orbital period as a light-travel time
+  effect.  Combining the radial velocities with new limits from
   speckle imaging, we find that the system's acceleration is likely
@@
   between 10-100$\,$AU.  The statistics from
-  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters are expected
-  to show similar period decreases as WASP-4, due to acceleration by
-  outer companions.  We find that these period decreases will be
-  increasingly measureable in coming years, because the precision of
-  period derivative measurements scales as the fourth power of the
-  observing baseline.  Continued radial velocity monitoring
-  of hot Jupiters is therefore essential to distinguish tidal orbital
-  decay from line-of-sight accelerations.
+  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters
+  are expected to show similar period decreases as WASP-4, due to
+  acceleration by outer companions.  We find that these period
+  decreases will be increasingly measureable in coming years, because
+  the precision of period derivative measurements scales as the fourth
+  power of the observing baseline.  Continued radial velocity
+  monitoring of hot Jupiters is therefore essential to distinguish
+  tidal orbital decay from line-of-sight accelerations.
 \end{abstract}
@@
 hot Jupiter semi-major axis distribution, host star spin rates, and
-galactic velocity dispersions have led to estimates for the
-inspiral timescale that vary from much less to much greater than
-the main-sequence evolution time  ({\it e.g.},
+galactic velocity dispersions have led to estimates for the inspiral
+timescale that vary from much less to much greater than the
+main-sequence evolution time  ({\it e.g.},
 \citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
 \citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
-\citealt{hamer_schlaufman_2019}).  Direct
-measurements of tidal orbital decay through transit and occultation
-timing could provide an empirical solution.  For instance, combined
-transit timing and radial velocity measurements for WASP-12b have
-shown that its secular period decrease of $\approx$30 milliseconds per
-year is almost certainly due to tidal orbital decay
+\citealt{hamer_schlaufman_2019}).  Direct measurements of tidal
+orbital decay through transit and occultation timing could provide an
+empirical solution.  For instance, combined transit timing and radial
+velocity measurements for WASP-12b have shown that its secular period
+decrease of $\approx$30 milliseconds per year is almost certainly due
+to tidal orbital decay
 \citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
@@
 per year.  We discovered the timing variations using data from TESS
-\citep{ricker_transiting_2015}
-and a decade of ground-based observations \citep[][hereafter
+\citep{ricker_transiting_2015} and a decade of ground-based
+observations \citep[][hereafter
 \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.  Thereafter,
@@
 reported additional archival transit light curves of WASP-4b, most
-notably from TRAPPIST, and also from select other
-observers.  \citet{baluev_2019} pointed out that when using
-lower-precision subsets of the available transit data, the case for a
-decreasing period worsened.  We agree with this statement.
+notably from TRAPPIST, and also from select other observers.
+\citet{baluev_2019} pointed out that when using lower-precision
+subsets of the available transit data, the case for a decreasing
+period worsened.  We agree with this statement.
 
@@
 the RV baseline from 3 to 9 years.  Previously, the five available
-HIRES radial velocities suggested a weak ($\approx$$2\sigma$) linear trend
-\citep{knutson_friends_2014}.  Our new measurements reveal a
+HIRES radial velocities suggested a weak ($\approx$$2\sigma$) linear
+trend \citep{knutson_friends_2014}.  Our new measurements reveal a
 line-of-sight acceleration of $\dot{\gamma} =
@@
 
-The complete set of radial velocity observations is given in 
-Table~2.  Along with the 2010-2019 HIRES observations, there are also
-earlier measurements from CORALIE and HARPS.  Following
+The complete set of radial velocity observations is given in Table~2.
+Along with the 2010-2019 HIRES observations, there are also earlier
+measurements from CORALIE and HARPS.  Following
 \citetalias{bouma_wasp4b_2019}, we included the CORALIE measurements
@@
 1000 frames each) and also on the night of September 28-29 (7 sets of
-1000 frames each).
-{\bf We reconstructed the images following...}
+1000 frames each).  {\bf We reconstructed the images following...}
 {\bf We reduced the reconstructed speckle images to contrast curves by
-injecting and recovering point-sources...}
-The contrast curves showed that the second night, which had better
-seeing, also produces the more constraining result.
-The 832$\,$nm results were also typically the most constraining, since
-faint companions are typically redder.
-We therefore opt for the 832$\,$nm September 28-29 results for the
+injecting and recovering point-sources...} The contrast curves showed
+that the second night, which had better seeing, also produces the more
+constraining result.  The 832$\,$nm results were also typically the
+most constraining, since faint companions are typically redder.  We
+therefore opt for the 832$\,$nm September 28-29 results for the
 remaining analysis.
@@
 of period decrease we reported earlier \citep[$- 12.6 \pm 1.2~{\rm
-ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of
-the new data from \citeauthor{southworth_transit_2019}
-and \citeauthor{baluev_2019}  The other best-fit
-transit timing model parameters are reported in Table~3.
+ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of the
+new data from \citeauthor{southworth_transit_2019} and
+\citeauthor{baluev_2019}  The other best-fit transit timing model
+parameters are reported in Table~3.
 
@@
 corresponding $p$-value is 0.65.  Furthermore, inspection of the
-$S$-index timeseries did not show secular or sinusoidal trends, as would
-be expected if we were observing a long-term magnetic activity cycle.
-The $S$-index values are included in Table~2.  We conclude that
-it is highly unlikely that the linear trend is caused by
-stellar activity.
+$S$-index timeseries did not show secular or sinusoidal trends, as
+would be expected if we were observing a long-term magnetic activity
+cycle.  The $S$-index values are included in Table~2.  We conclude
+that it is highly unlikely that the linear trend is caused by stellar
+activity.
 
@@
 \end{equation}
-where $\tau$ is the observing baseline.
-For WASP-4, this yields $M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher masses
-are allowed for companions that orbit further from the
-star: at fixed $\dot{\gamma}$, $M_{\rm comp} \propto a^2$
+where $\tau$ is the observing baseline.  For WASP-4, this yields
+$M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher masses are allowed for
+companions that orbit further from the star: at fixed $\dot{\gamma}$,
+$M_{\rm comp} \propto a^2$
 \citep{torres_substellar_1999,liu_crossing_2002}.
@@
 by setting an upper limit on the semimajor axis, and a maximum
-brightness (and thereby mass) of any putative companions.
-The procedure we use to combine constraints from both radial
-velocities and high resolution imaging has been developed by
+brightness (and thereby mass) of any putative companions.  The
+procedure we use to combine constraints from both radial velocities
+and high resolution imaging has been developed by
 \citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
@@
 of synthetic photometry, we further assumed that all sources had
-blackbody spectra. While this is a bold simplification, we do not readily have
-access to the planetary and stellar atmosphere models needed for the
-consistent calculation with the \texttt{COND03} and \texttt{MESA}
-models.  We therefore adopted the effective temperatures and
-bolometric luminosities from the \citet{baraffe_evolutionary_2003} and
-MIST isochrones.  Using these theoretical quantities and the
+blackbody spectra. While this is a bold simplification, we do not
+readily have access to the planetary and stellar atmosphere models
+needed for the consistent calculation with the \texttt{COND03} and
+\texttt{MESA} models.  We therefore adopted the effective temperatures
+and bolometric luminosities from the \citet{baraffe_evolutionary_2003}
+and MIST isochrones.  Using these theoretical quantities and the
 empirically-measured Zorro bandpasses, we calculated absolute
@@
 cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
-bound observed by \citet{schlaufman_evidence_2018} between
-giant planets and brown dwarfs, though this value is also close to the
+bound observed by \citet{schlaufman_evidence_2018} between giant
+planets and brown dwarfs, though this value is also close to the
 $13\,M{\rm Jup}$ required for a solar-metallicity body to
@@
 For each simulated companion, we then drew a sample from the converged
-chains of our initial model of WASP-4b. We
-subtracted the planet's orbital solution, leaving RV points with
-a linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each
-simulated outer companion, and the fixed instrument offsets and
-jitters from the MCMC chains, we then performed a maximum likelihood
-fit for the time and argument of periastron of the outer simulated
-companion.  We converted the resulting $128\times128\times512$ cube of
-log-likelihood values to probabilities, and averaged over the
-samples in each grid cell to derive a probability distribution in mass
-and semi-major axis.  Figure~\ref{fig:mass_sma} shows the result.
+chains of our initial model of WASP-4b. We subtracted the planet's
+orbital solution, leaving RV points with a linear trend.  Given
+$(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each simulated outer
+companion, and the fixed instrument offsets and jitters from the MCMC
+chains, we then performed a maximum likelihood fit for the time and
+argument of periastron of the outer simulated companion.  We converted
+the resulting $128\times128\times512$ cube of log-likelihood values to
+probabilities, and averaged over the samples in each grid cell to
+derive a probability distribution in mass and semi-major axis.
+Figure~\ref{fig:mass_sma} shows the result.
 
@@
 decrease of WASP-4b's orbital period.
-% While further radial velocity observations of the system should help
-% in determining the mass and semi-major axis of the companion, it seems
-% unlikely that additional transit observations will yield near-term
-% constraints on orbital decay.
 
@@
 at one time have influenced the orbital evolution of the inner giant.
-The fact that most hot Jupiters have similar massive outer
-companions \citep{knutson_friends_2014,bryan_statistics_2016} is
-circumstantial evidence for certain high-eccentricity formation
-channels (see \citealt{dawson_johnson_2018}).
-Further radial velocity monitoring should eventually reveal the
-orbital parameters and minimum mass of WASP-4's companion.
+The fact that most hot Jupiters have similar massive outer companions
+\citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
+evidence for certain high-eccentricity formation channels (see
+\citealt{dawson_johnson_2018}).  Further radial velocity monitoring
+should eventually reveal the orbital parameters and minimum mass of
+WASP-4's companion.
 
@@
 outside of $5\,{\rm AU}$ with super-Jovian masses
-\citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight accelerations are
-correspondingly common in hot Jupiter systems.
+\citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight
+accelerations are correspondingly common in hot Jupiter systems.
 
@@
 \citet{knutson_friends_2014}, and computed the expected orbital period
-derivatives $\dot{P}_{\rm\,RV} = \dot{v}_{\rm r} P/c$ for each
-system.  The results are given in Table~4, and visualized for
-hot Jupiters with significant ($>$$3\sigma$) linear trends in
-Figure~\ref{fig:pdot_pop}.
+derivatives $\dot{P}_{\rm\,RV} = \dot{v}_{\rm r} P/c$ for each system.
+The results are given in Table~4, and visualized for hot Jupiters with
+significant ($>$$3\sigma$) linear trends in Figure~\ref{fig:pdot_pop}.
 
@@
 \citet{knutson_friends_2014} show a non-zero radial velocity trend.
-Therefore around 1 in
-3 hot Jupiters is expected to show period changes commensurate with
-WASP-4 due to acceleration by outer companions.  Half of these will be
-period decreases, and will be an astrophysical false positive in
-searches for tidal orbital decay.
+Therefore around 1 in 3 hot Jupiters is expected to show period
+changes commensurate with WASP-4 due to acceleration by outer
+companions.  Half of these will be period decreases, and will be an
+astrophysical false positive in searches for tidal orbital decay.
 
@@
 scale with the observing baseline?  We take a Fisher analysis approach
-to the problem.
-First, we rewrite Equation~\ref{eq:quadratic_model} as
+to the problem.  First, we rewrite Equation~\ref{eq:quadratic_model}
+as
 \begin{align}
@@
 \end{align}
-where $a_0\equiv t_0$, $a_1\equiv P$, and $a_2\equiv 0.5\cdot {\rm d}P/{\rm d}E$.  
-By following \citet{gould_chi2_2003},
-we derived
-that if $N$ transit timing measurements are taken uniformly
-across a baseline of $\Delta E$ epochs with constant precision
-$\sigma_{\rm tra}$, then the uncertainty of the quadratic term is
-given by
+where $a_0\equiv t_0$, $a_1\equiv P$, and $a_2\equiv 0.5\cdot {\rm
+d}P/{\rm d}E$.  By following \citet{gould_chi2_2003}, we derived that
+if $N$ transit timing measurements are taken uniformly across a
+baseline of $\Delta E$ epochs with constant precision $\sigma_{\rm
+tra}$, then the uncertainty of the quadratic term is given by
 \begin{align}
@@
 \end{align}
-This result implies that a doubled observing baseline yields
-a sixteen-fold improvement in precision on d$P$/d$t$.  If regular
+This result implies that a doubled observing baseline yields a
+sixteen-fold improvement in precision on d$P$/d$t$.  If regular
 observations continue from ground and space-based observatories,

### 2020-01-29 paper/ms.tex
minor

@@
 Bayesian information criteria (BIC) between the linear and quadratic
-and models is $\Delta {\rm BIC} = 89.1$, strongly favoring the
+and models is $\Delta {\rm BIC} = 89$, strongly favoring the
 quadratic case \citep{kass_bayes_1995}.
@@
 Our initial model for the radial velocity data was a single Keplerian
-orbit, plus instrument offsets, jitters, and a long-term trend
+orbit, plus instrument offsets, jitters, and a long-term linear trend
 \citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
@@
 zero-points, an additive ``white noise'' instrument jitter for each
-instrument, linear ($\dot{v_{\rm r}}$), and optionally second-order
-($\ddot{v_{\rm r}}$) acceleration terms.
+instrument, and a linear ($\dot{v_{\rm r}}$) acceleration term.  We
+also considered a model without the linear trend, and found that it
+was disfavored by $\Delta {\rm BIC}=73$.
 
-%FIXME: this is a janky approach if so??
-We found that the best-fitting model with both linear and quadratic
-radial velocity terms was marginally preferred (by $\Delta
-\mathrm{BIC} = 5.8$) over the best-fitting model with only a linear
-term.  Regardless, for consistency with \citet{knutson_friends_2014},
-who fixed the quadratic component of the long-term trend to zero, in
-Figure~\ref{fig:rvs} we show best-fitting models for the linear-trend
-case.
-
-WASP-4 was found to be accelerating towards our line-of-sight at high
-confidence,
+In the preferred ``planet+trend'' model, WASP-4 is accelerating
+towards our line-of-sight at high confidence,
 \begin{equation}
@@
 \end{equation}
-For comparison, before our new measurements, $\dot{v}_{\rm r}$ was
-thought to be about five times smaller, and was only marginally
-significant \citep{knutson_friends_2014,bouma_wasp4b_2019}.
+For comparison, before our new measurements, $\dot{\gamma}$ was
+thought to be about five times smaller, and had marginal statistical significance
+\citep{knutson_friends_2014,bouma_wasp4b_2019}.
 

### 2020-01-29 paper/ms.tex
update paper
ready for JNW!

@@
   WASP-4b have recently shown that its orbital period is decreasing.
-  The period change could be caused by tidal orbital decay, orbital
-  precession, or light-travel time effects.  We present new radial
-  velocity measurements of WASP-4, acquired with Keck-HIRES.  The data
-  show that the system is accelerating towards the Earth at
-  $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
-  s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease probably
-  explains WASP-4's changing orbital period as a light-travel time
-  effect.  Combining the radial velocities with new limits from
+  Proposed explanations for the period change include tidal orbital
+  decay, orbital precession, and light-travel time effects.  We
+  present new radial velocity measurements of WASP-4, acquired with
+  Keck-HIRES.  The data show that the system is accelerating towards
+  the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
+  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease
+  probably explains WASP-4's changing orbital period as a light-travel
+  time effect.  Combining the radial velocities with new limits from
   speckle imaging, we find that the system's acceleration is likely
@@
 orbital decay through transit and occultation timing could provide an
-empirical solution.  For instance, combined transit timing and radial
-velocity measurements for WASP-12b have shown that its secular period
-decrease of $\approx$30 milliseconds per year is almost certainly due
-to tidal orbital decay
+empirical resolution.  For instance, combined transit timing and
+radial velocity measurements for WASP-12b have shown that its secular
+period decrease of $\approx$30 milliseconds per year is almost
+certainly due to tidal orbital decay
 \citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
@@
 occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
-Jupiters with masses from 1-20$M_{\rm Jup}$ and semi-major axes from
-5-100$\,$AU.  Therefore it could very well be that more hot Jupiters
-will show shrinking orbital periods due to long term accelerations
-than due to tidal orbital decay.
+Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
+from 5-100$\,$AU.  Therefore it could very well be that more hot
+Jupiters will show shrinking orbital periods due to long term
+accelerations than due to tidal orbital decay.
 
@@
 orbital period that appears to be decreasing by about 10 milliseconds
-per year.  We discovered the timing variations using data from TESS
-\citep{ricker_transiting_2015} and a decade of ground-based
+per year.  We discovered the timing variations by combining data from
+TESS \citep{ricker_transiting_2015} with a decade of ground-based
 observations \citep[][hereafter
@@
 were otherwise similar.  A separate study by \citet{baluev_2019}
-reported additional archival transit light curves of WASP-4b, most
-notably from TRAPPIST, and also from select other observers.
-\citet{baluev_2019} pointed out that when using lower-precision
-subsets of the available transit data, the case for a decreasing
-period worsened.  We agree with this statement.
+reported additional archival transit light curves of WASP-4b from
+TRAPPIST and select other observers.  \citet{baluev_2019} pointed out
+that when using lower-precision subsets of the available transit data,
+the case for a decreasing period worsened.  We agree with this
+statement.
 
@@
 
-In the following, \S~\ref{sec:observations} collects the available
-transit data and presents the new radial velocity and speckle imaging
-observations.  \S~\ref{sec:analysis} analyzes the data, and finds that
-they yield a picture in which the WASP-4 system is accelerating
-towards our line-of-sight, likely due to the pull of a brown or
-M-dwarf companion.  \S~\ref{sec:discussion} places this result in the
-broader context of orbital decay searches, and points out that
-line-of-sight accelerations, {\it i.e.}, ``false positive orbital
-decay signals'', are relatively common in the hot Jupiter population.
-\S~\ref{sec:conclusions} offers concluding remarks.
+In the following, Section~\ref{sec:observations} collects the
+available transit data and presents the new radial velocity and
+speckle imaging observations.  Section~\ref{sec:analysis} analyzes the
+data, and finds that they yield a picture in which the WASP-4 system
+is accelerating towards our line-of-sight, likely due to the pull of a
+brown or M-dwarf companion.  Section~\ref{sec:discussion} places this
+result in the broader context of orbital decay searches, and points
+out that line-of-sight accelerations, {\it i.e.}, ``false positive
+orbital decay signals'', are relatively common in the hot Jupiter
+population.  Section~\ref{sec:conclusions} offers concluding remarks.
 
@@
 
-% TODO: Details of reduction from Howard or Knutson
 After identifying the period decrease in
@@
 velocity measurements with the Keck High Resolution Echelle
-Spectrometer (HIRES, \citealt{vogt_hires_1994}).  Our observations
+Spectrometer (HIRES; \citealt{vogt_hires_1994}).  Our observations
 were acquired under the purview of the California Planet Survey
@@
 % were derived?
-We observed WASP-4 twice, on the night of September 11-12 (8 sets of
-1000 frames each) and also on the night of September 28-29 (7 sets of
-1000 frames each).  {\bf We reconstructed the images following...}
+We observed WASP-4 twice, on the night of September 11-12 and also
+on the night of September 28-29. 
+{\bf We reconstructed the images following...}
 {\bf We reduced the reconstructed speckle images to contrast curves by
-injecting and recovering point-sources...} The contrast curves showed
-that the second night, which had better seeing, also produces the more
-constraining result.  The 832$\,$nm results were also typically the
-most constraining, since faint companions are typically redder.  We
-therefore opt for the 832$\,$nm September 28-29 results for the
-remaining analysis.
+injecting and recovering point-sources...}
+The contrast curves showed that the second night, which had better
+seeing, also produced the more constraining result.  The 832$\,$nm
+results were also the most useful, since faint companions are redder than
+the host star.  We therefore opted for the 832$\,$nm September 28-29
+results for the remaining analysis.
 
@@
   its $\pm1\sigma$ statistical uncertainties are shown in blue.
-  Neither model fully explains the scatter in transit residuals,
-  perhaps due to underestimated uncertainties
-  (see~\ref{sec:transit_analysis}).  The binned TESS point (yellow
-  star) is the weighted average of 18 TESS transits and is binned for
-  display purposes only.  Its uncertainty is slightly larger than the
-  marker size.  The models were fitted to all of the individual
-  transit times.
+  The scatter in the residuals could be due to underestimated
+  measurement uncertainties (see Section~\ref{sec:transit_analysis}).
+  The binned TESS point (yellow star) is the weighted average of 18
+  TESS transits and is binned for display purposes only.  Its
+  uncertainty is slightly larger than the marker size.  The models
+  were fitted to all of the individual transit times.
   \label{fig:times}
@@
 \end{align}
-where $E$ is the epoch number and $t_0$ is a reference epoch.  The
-second model assumes that the period changes at a steady rate:
+where $E$ is the integer transit number and $t_0$ is a reference
+epoch.  The second model assumes that the period changes at a steady
+rate:
 \begin{align}
@@
 must be some additional source of signal or noise. In
-\citet{bouma_wasp4b_2019}, we found that the quadratic model for the
+\citetalias{bouma_wasp4b_2019}, we found that the quadratic model for the
 (sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsened
@@
 
-One approach to analyzing such data would be to introduce a fudge
-factor to inflate the transit measurement uncertainties, and lower the
-reduced $\chi^2$.  We do not think that such an approach is warranted,
-since it would not change the result that the quadratic model is
-strongly preferred.
+One approach to analyzing such data would be to inflate the transit
+measurement uncertainties, and lower the reduced $\chi^2$.  We do not
+think that such an approach is warranted, since it would not change
+the result that the quadratic model is strongly preferred.
 
@@
 ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
-of period decrease we reported earlier \citep[$- 12.6 \pm 1.2~{\rm
+of period decrease we previously reported \citepalias[$- 12.6 \pm 1.2~{\rm
 ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of the
@@
 For comparison, before our new measurements, $\dot{\gamma}$ was
-thought to be about five times smaller, and had marginal statistical significance
-\citep{knutson_friends_2014,bouma_wasp4b_2019}.
+thought to be about five times smaller, and had marginal statistical
+significance \citep{knutson_friends_2014,bouma_wasp4b_2019}.
 
@@
     \caption{
-      {\bf Zorro contrast curve derived from point-source
+      {\bf Zorro contrast limits derived from point-source
       injection-recovery experiments.} Sources below the curve would
@@
     {\bf Masses and semi-major axes of companions that meet
-    requirements of both the radial velocity and the speckle imaging.}
-    The likelihood inferred from radial velocities is shown in
-    grayscale, and the region excluded from the speckle imaging is
-    indicated (blue line).  The expected mass to semimajor-axis
-    degeneracy is shown with a black line.
+    requirements of both the radial velocities and the speckle
+    imaging.} The likelihood inferred from radial velocities is shown
+    in grayscale, and the region excluded from the speckle imaging is
+    indicated.  The expected mass to semimajor-axis degeneracy is
+    shown with a black line.
 	\label{fig:mass_sma}
@@
 planets and brown dwarfs, though this value is also close to the
-$13\,M{\rm Jup}$ required for a solar-metallicity body to
-significantly burn deuterium \citep{burrows_nongray_1997}.
+$13\,M_{\rm Jup}$ deuterium-burning limit \citep[{\it
+e.g.},][]{burrows_nongray_1997}.
 
@@
 ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
-perfect, Occam's razor would suggest that the line of sight
+perfect, Occam's razor would suggest that the line-of-sight
 acceleration is probably a sufficient explanation for the apparent
@@
 \citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
-evidence for certain high-eccentricity formation channels (see
+evidence for certain high-eccentricity formation pathways (see
 \citealt{dawson_johnson_2018}).  Further radial velocity monitoring
@@
   \citet{knutson_friends_2014} have shown long-term radial velocity
-  trends.  The HAT-P-11 signal is plotted, though its radial velocity
-  signal may be caused by stellar activity.  Three hot Jupiters are
-  not plotted because their radial velocity curves better described as
-  quadratic trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects
-  are ordered in the $y$ dimension by the absolute value of d$P$/d$t$.
+  trends.  HAT-P-11 signal is plotted, though its signal may be caused
+  by stellar activity.  Three hot Jupiters are not plotted because
+  their radial velocity curves better described as quadratic trends in
+  time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered in the $y$
+  dimension by the absolute value of d$P$/d$t$.
 	\label{fig:pdot_pop}
@@
 which the period derivative can be measured from transits scales as
-the fourth power of the baseline duration (\S~\ref{sec:fisher}), and
+the fourth power of the baseline duration (Section~\ref{sec:fisher}), and
 so within a decade many more hot Jupiters should show orbital period
-changes due to accelerations induced by their outer companions.  This
-effect can be distinguished from tidal orbital decay through long-term
-radial velocity monitoring of hot Jupiters.
+changes due to accelerations induced by their outer companions.  
+To distinguish this effect from tidal decay, further long-term radial
+velocity measurements of hot Jupiters are required.
 

### 2020-01-30 paper/ms.tex
fix math error in dP/dt vs baseline scaling

@@
   decreases will be increasingly measureable in coming years, because
-  the precision of period derivative measurements scales as the fourth
-  power of the observing baseline.  Continued radial velocity
+  the precision of period derivative measurements scales as
+  (observing baseline)$^{5/2}$.  Continued radial velocity
   monitoring of hot Jupiters is therefore essential to distinguish
@@
 radial velocity measurements for WASP-12b have shown that its secular
-period decrease of $\approx$30 milliseconds per year is almost
-certainly due to tidal orbital decay
+period decrease of $\approx$30 milliseconds per year is probably due
+to tidal orbital decay
 \citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
@@
 \begin{align}
-  \sigma_{a_2} =
-  \left(\frac{25{,}920}{N}\right)^{1/2}
-  \frac{\sigma_{\rm tra}}{(\Delta E)^4}
+  \sigma_{a_2} = 6\sqrt{5}
+   \frac{\sigma}{N^{1/2} (\Delta E)^2} \propto (\Delta E)^{-5/2}.
 \end{align}
-This result implies that a doubled observing baseline yields a
-sixteen-fold improvement in precision on d$P$/d$t$.  If regular
+This result implies that a doubled observing baseline yields an
+$\approx$5.7-fold improvement in precision on d$P$/d$t$.  If regular
 observations continue from ground and space-based observatories,
 period derivatives with precision below $1~{\rm ms}\,{\rm yr}^{-1}$
-will be measureable before 2030.
+will become much more common over the coming decade.
 
@@
 observational baselines get longer.  In particular, the precision with
-which the period derivative can be measured from transits scales as
-the fourth power of the baseline duration (Section~\ref{sec:fisher}), and
+which the period derivative can be measured from transits scales with
+the baseline duration to the power of 5/2 (Section~\ref{sec:fisher}), and
 so within a decade many more hot Jupiters should show orbital period
-changes due to accelerations induced by their outer companions.  
+changes due to accelerations from their outer companions.  
 To distinguish this effect from tidal decay, further long-term radial

### 2020-01-30 src/make_WASP-4b_rv_table.py
add publication-ready Pdot and RV tables

@@ -1,0 +1,74 @@
+'''
+make table of RVs
+'''
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt, pandas as pd, numpy as np
+
+from numpy import array as nparr
+
+from glob import glob
+import os, pickle
+
+def main():
+
+    df = pd.read_csv(
+        '../data/RVs_all_WASP4b_for_fitting_20190911.csv', sep=','
+    )
+    jump_df = pd.read_csv(
+        '../data/20190911_jump_wasp4_rv.csv', comment='#', sep=','
+    )
+
+    svalue = np.hstack((
+        np.ones(len(df[df.tel != 'HIRES']))*np.nan,
+        nparr(jump_df.svalue)
+    ))
+
+    original_references = np.array(df['Source'])
+    references = []
+    for ref in original_references:
+        if ref == 'Triaud+2010_vizier':
+            references.append('\citet{triaud_spin-orbit_2010}')
+        elif ref == 'Pont+2011_vizierMNRAS':
+            references.append('\citet{pont_determining_2011}')
+        elif ref == 'jump_20190911':
+            references.append('This work')
+    references = np.array(references)
+
+    outdf = pd.DataFrame(
+        {'time': np.round(nparr(df.time),8),
+         'mnvel': np.round(nparr(df.mnvel),5),
+         'errvel': np.round(nparr(df.errvel),5),
+         'svalue': svalue,
+         'tel': nparr(df.tel),
+         'provenance': references
+        }
+    )
+
+    outpath = '../paper/WASP-4b_rv_data.tex'
+    with open(outpath,'w') as tf:
+        tf.write(outdf.to_latex(index=False, escape=False))
+    print('wrote {:s}'.format(outpath))
+
+    outdf.to_csv('../paper/WASP-4b_rv_data.csv', index=False)
+    print('also wrote ../paper/WASP-4b_rv_data.csv')
+
+
+    #
+    # trim this tex output to remove the annoying first rows, so you
+    # can \input{WASP-4b_rv_data.tex} from the table file.
+    #
+    with open(outpath,'r') as f:
+        lines = f.readlines()
+    midrule_ix = [ix for ix,l in enumerate(lines)
+                  if r'\midrule' in l][0]
+    bottomrule_ix = [ix for ix,l in enumerate(lines)
+                     if r'\bottomrule' in l][0]
[…]

### 2020-01-30 src/make_pdot_table.py

@@ -1,0 +1,59 @@
+'''
+make table of Pdots
+'''
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt, pandas as pd, numpy as np
+
+from numpy import array as nparr
+
+from glob import glob
+import os, pickle
+
+def main():
+
+    df = pd.read_csv(
+        '../results/knutson_all_pdots.csv', sep=';'
+    )
+
+    # \colhead{Planet} &
+    # \colhead{$\dot{\gamma}$ [m$\,$s$^{-1}$$\,$yr$^{-1}$]} &
+    # \colhead{$+\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
+    # \colhead{$-\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
+    # \colhead{$P$ [days]} &
+    # \colhead{$\dot{P}_{\,{\rm RV}}$ [ms$\,$yr$^{-1}$]} &
+    # \colhead{$+\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
+    # \colhead{$-\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
+    # \colhead{Significant?}
+
+    # 'planet', 'gammadot', 'gammadot_pluserr', 'gammadot_minuserr', 'comment',
+    # 'pl_name', 'pl_orbper', 'Pdot', 'Pdot_upper_limit', 'Pdot_lower_limit',
+    # 'Pdot_perr', 'Pdot_merr', 'abs_Pdot', 'K14_significant'
+
+    scols = [
+        'planet', 'gammadot', 'gammadot_pluserr', 'gammadot_minuserr',
+        'pl_orbper', 'Pdot', 'Pdot_perr', 'Pdot_merr', 'K14_significant',
+        'comment'
+    ]
+
+    outdf = df[scols]
+
+    threeptcols = ['Pdot', 'Pdot_perr', 'Pdot_merr']
+    for c in threeptcols:
+        outdf[c] = np.round(nparr(outdf[c]),3)
+
+    sixptcols = ['gammadot', 'gammadot_pluserr', 'gammadot_minuserr']
+    for c in sixptcols:
+        outdf[c] = np.round(nparr(outdf[c]),6)
+
[…]

### 2020-01-30 paper/ms.tex
add Howard Isaacson's suggestion

@@
 Spectrometer (HIRES; \citealt{vogt_hires_1994}).  Our observations
-were acquired under the purview of the California Planet Survey
-\citep{howard_cps_2010}.  The spectra were reduced using {\bf software
-X}.  Previously, the HIRES data-points spanned 2010 to 2013
-\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
-observing baseline to nine years.
+were acquired using the standard setup and reduction techniques of the
+California Planet Survey \citep{howard_cps_2010}.  Previously, the
+HIRES data-points spanned 2010 to 2013 \citep{knutson_friends_2014}.
+Our new measurements triple the HIRES observing baseline to nine
+years.
 

### 2020-01-31 paper/ms.tex
fix gammadot unit typos; check Feng+15 result

@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently shown that its orbital period is decreasing.
+  WASP-4b have recently shown that its orbital period appears to be decreasing.
   Proposed explanations for the period change include tidal orbital
@@
   the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
-  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease
+  m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.  The implied period decrease
   probably explains WASP-4's changing orbital period as a light-travel
@@
 line-of-sight acceleration of $\dot{\gamma} =
--0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,yr^{-1}}$.  This
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,day^{-1}}$.  This
 translates to an expected period decrease from the light-travel time
@@
 Given a linear radial velocity trend, we can place lower-limits on the
-mass and semi-major axis of additional bodies in the system.  As a
+mass and semi-major axis of additional bodies in the system.  For a
 quick estimate of the minimum mass required to explain the linear
@@
 system with a linear trend is a companion with $e\approx0.5$ and
-$\omega=90^\circ$.  From their Equation~1,
+$\omega=90^\circ$.  Substituting $P\approx 1.25\tau$, $K \approx
+0.5\tau \gammadot$ into the mass function \citep[{\it
+e.g.},][]{wright_efficient_2009} yields
 \begin{equation}
- M_{\rm min} \approx 0.0164\,M_{\rm Jup}
+ M_{\rm min} \approx 5.99\,M_{\rm Jup}
   \left( \frac{\tau}{{\rm yr}} \right)^{4/3}
-  \left| \frac{\dot{\gamma}}{{\rm m\,s^{-1}\,yr^{-1}}} \right|
+  \left| \frac{\dot{\gamma}}{{\rm m\,s^{-1}\,day^{-1}}} \right|
   \left( \frac{M_\star}{{M_\odot}} \right)^{2/3},
 \end{equation}
-where $\tau$ is the observing baseline.  For WASP-4, this yields
-$M_{\rm min} = 4.7 M_{\rm Jup}$.  Higher masses are allowed for
+where $\tau$ is the observing baseline.  For WASP-4, this gives
+$M_{\rm min} = 4.9 M_{\rm Jup}$.  Higher masses are allowed for
 companions that orbit further from the star: at fixed $\dot{\gamma}$,
@@
 is accelerating towards the Earth at $\dot{\gamma} =
--0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
 The corresponding light-travel time effect predicts a period decrease

### 2020-01-31 src/minimum_mass_given_linear_trend.py

@@ -11,10 +11,10 @@ Katherina Feng et al (2015).
 """
 
 import numpy as np
-from astropy import units as u, constants as c
+from astropy import units as u, constants as const
 
 baseline = 10*u.year
-gammadot = -0.04 *u.m/u.s/u.day
+gammadot = -0.0422 *u.m/u.s/u.day
 mstar = 0.864*u.Msun
 
 M_min = (
@@ -25,3 +25,37 @@ M_min = (
 
 print(M_min.to(u.Mjup))
 print(M_min.to(u.Mearth))
+
+####################
+
+#
+# now double check the prefactor
+#
+ecc = 0.5
+prefactor = (
+    (1.25/(2*np.pi*const.G))**(1/3)
+    * (1-ecc**2)**(1/2)
+    * ( (1*u.Msun)**(2/3)*(1*u.m/u.s/u.year)*(1*u.year)**(4/3) )
+)
+print(prefactor.to(u.Mjup))
+print((prefactor/2).to(u.Mjup))
+
+
+#
+# what if we used better units?
+#
+ecc = 0.5
+prefactor = (
+    (1.25/(2*np.pi*const.G))**(1/3)
+    * (1-ecc**2)**(1/2)
+    * ( (1*u.Msun)**(2/3)*(1*u.m/u.s/u.day)*(1*u.year)**(4/3) )
+)
+print(prefactor.to(u.Mjup))
+print((prefactor/2).to(u.Mjup))
+
+M_min = (
+    (5.992106 * u.Mjup) * (baseline/(1*u.yr))**(4/3)
+    * np.abs(gammadot/(1*u.m/u.s/u.day))
+    * (mstar/u.Msun)**(2/3)
+)
+print(M_min)

### 2020-01-31 paper/ms.tex
text updates
per JNW comments

@@
   Space and ground-based transit observations of the hot Jupiter
-  WASP-4b have recently shown that its orbital period appears to be decreasing.
-  Proposed explanations for the period change include tidal orbital
-  decay, orbital precession, and light-travel time effects.  We
-  present new radial velocity measurements of WASP-4, acquired with
-  Keck-HIRES.  The data show that the system is accelerating towards
-  the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
-  m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.  The implied period decrease
-  probably explains WASP-4's changing orbital period as a light-travel
-  time effect.  Combining the radial velocities with new limits from
-  speckle imaging, we find that the system's acceleration is likely
-  caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
-  between 10-100$\,$AU.  The statistics from
-  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters
-  are expected to show similar period decreases as WASP-4, due to
-  acceleration by outer companions.  We find that these period
-  decreases will be increasingly measureable in coming years, because
-  the precision of period derivative measurements scales as
-  (observing baseline)$^{5/2}$.  Continued radial velocity
-  monitoring of hot Jupiters is therefore essential to distinguish
-  tidal orbital decay from line-of-sight accelerations.
+  WASP-4b have recently shown that its orbital period appears to be
+  decreasing.  Proposed explanations for the period change include
+  tidal orbital decay, orbital precession, and light-travel time
+  effects.  We present new radial velocity measurements of WASP-4,
+  acquired with Keck-HIRES.  The data show that the system is
+  accelerating towards the Earth at $\dot{\gamma} =
+  -0.0422^{+0.0028}_{-0.0027}\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
+  The implied period decrease explains most or all of WASP-4's
+  changing orbital period as a light-travel time effect ($-5.94 \pm
+  0.39~{\rm msec}\,{\rm yr}^{-1}$ implied, compared to $-8.64 \pm
+  0.89~{\rm msec}\,{\rm yr}^{-1}$ observed).  Combining the radial
+  velocities with new upper limits from speckle imaging, we find that
+  the system's acceleration is likely caused by a 10-200$\,M_{\rm
+  Jup}$ companion with semi-major axis between 10-100$\,$AU.  The
+  statistics from \citet{knutson_friends_2014} imply that about 1 in 6
+  hot Jupiters are expected to show comparable period decreases to
+  WASP-4, due to acceleration by outer companions.  We find that these
+  period decreases will be increasingly measureable in coming years,
+  because the precision of period derivative measurement scales as
+  (observing baseline)$^{5/2}$.  Continued radial velocity monitoring
+  of hot Jupiters is therefore essential to distinguish tidal orbital
+  decay from line-of-sight accelerations.
 \end{abstract}
@@
 that when using lower-precision subsets of the available transit data,
-the case for a decreasing period worsened.  We agree with this
-statement.
+the case for a decreasing period worsened.
 
@@
 effect of -5.9 milliseconds per year---about commensurate with what is
-observed from transits.  In the following, we avoid the term
-``R{\o}mer delay'', which typically signifies arrival time delay due
-to observatory motion. We are discussing the Doppler effect seen by a
-stationary observer for a source with constant line-of-sight
-acceleration.
+observed from transits.  While the apparent period change caused by a
+line-of-sight acceleration has been referred to as the ``R{\o}mer
+effect'' \citep{yee_orbit_2020}, we prefer to avoid this term, which
+typically signifies arrival time delay due to observatory motion. We
+are discussing the Doppler effect seen by a stationary observer for a
+source with constant line-of-sight acceleration.
 
@@
   = - (2.74 \pm 0.28)\times 10^{-10}
-  = - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}.
+  = - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
     produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
-    0.89~{\rm ms}\,{\rm yr}^{-1}$) is indicated with the purple dotted
+    0.89~{\rm msec}\,{\rm yr}^{-1}$) is indicated with the purple dotted
     line.  The four new RV measurements presented in this work
@@
   = \frac{\dot{v}_{\rm r} P}{c}
-  = -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
+  = -5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}.
 \end{align}
 In other words, the majority of the period decrease observed in
-transits ($\dot{P}= - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) seems to
+transits ($\dot{P}= - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) seems to
 be explained by the acceleration of the host star.
@@
 observations continue from ground and space-based observatories,
-period derivatives with precision below $1~{\rm ms}\,{\rm yr}^{-1}$
+period derivatives with precision below $1~{\rm msec}\,{\rm yr}^{-1}$
 will become much more common over the coming decade.
@@
 The corresponding light-travel time effect predicts a period decrease
-$\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$.  The
+$\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$.  The
 majority of the period decrease observed in transits ($\dot{P}= - 8.64
-\pm 0.89~{\rm ms}\,{\rm yr}^{-1}$) is therefore explained by the
+\pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
 acceleration of the host star~---~not tidal orbital decay, or apsidal

### 2020-01-31 paper/ms.tex
minor updates

@@
 web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
-Zorro is a dual-channel speckle interferometer with narrow-band
+Zorro is a dual-channel speckle interferometer employing narrow-band
 filters centered at 562$\,$nm and 832$\,$nm.
 
-% TODO: might Howell suggest a paragraph for how the contrast ratios
-% were derived?
-We observed WASP-4 twice, on the night of September 11-12 and also
-on the night of September 28-29. 
-{\bf We reconstructed the images following...}
-{\bf We reduced the reconstructed speckle images to contrast curves by
-injecting and recovering point-sources...}
-The contrast curves showed that the second night, which had better
-seeing, also produced the more constraining result.  The 832$\,$nm
-results were also the most useful, since faint companions are redder than
-the host star.  We therefore opted for the 832$\,$nm September 28-29
-results for the remaining analysis.
+We observed WASP-4 twice, on the night of September 11-12 with
+relatively poor seeing (1.2$''$) and also on the night of September
+28-29.  Over each observation, we acquired three sets of 1000
+60$\,$msec exposures.  If a companion is present, the autocorrelation
+functions of these speckle images reveal a characteristic
+interference pattern. This pattern in then used to determine the
+properties of the detected companion and to produce a reconstructed
+image.  Using the reduced images and their
+reconstructed speckle images,  contrast curves are produced yielding
+5-sigma detection limits  (see \citealt{howell_speckle_2011}). 
+No companions were detected,
+and the second night, which had better seeing
+(0.6$''$), also produced the more constraining result.  The 832$\,$nm
+limits were the most useful to us given that faint companions
+are redder than the host star.  We therefore opted to use the
+832$\,$nm September 28-29 contrast limits for the remaining analysis.
 
@@
 The system's acceleration towards our line-of-sight causes a decrease
-in the apparent orbital period.  The period derivative expected from
-radial velocities is
+in the apparent orbital period.  The expected period derivative is
 \begin{align}
@@
 system with a linear trend is a companion with $e\approx0.5$ and
-$\omega=90^\circ$.  Substituting $P\approx 1.25\tau$, $K \approx
-0.5\tau \gammadot$ into the mass function \citep[{\it
+$\omega=90^\circ$.  Substituting $P\approx 1.25\tau$ and $K \approx
+0.5\tau \dot{\gamma}$ into the mass function \citep[{\it
 e.g.},][]{wright_efficient_2009} yields
@@
 of synthetic photometry, we further assumed that all sources had
-blackbody spectra. While this is a bold simplification, we do not
+blackbody spectra. While this is a simplification, we do not
 readily have access to the planetary and stellar atmosphere models
@@
 ($a=1.12$, $b=3.09$).  If the companion mass exceeded $10\,M_{\rm
-Jup}$, we drew the eccentricity distribution from the power-law $p_e
-\propto e^\eta$ reported by \citet{moe_mind_2017} in their
-Equation~17.  We emphasize that the long-period exoplanet and
+Jup}$, we drew the eccentricity from the power-law $p_e \propto
+e^\eta$ reported by \citet{moe_mind_2017} in their Equation~17 ($\eta
+\approx 0.5$ for most orbital periods).  The long-period exoplanet and
 long-period binary eccentricity distributions are quite different: the

### 2020-01-31 src/plot_mass_semimaj_constraints.py

@@ -55,16 +55,16 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
 
     if not linear_z:
 
-        cutoff = -12
+        cutoff = -6.3
 
         if discrete_color:
-            bounds = np.round(np.linspace(cutoff, np.log(prob_arr).max(), 5),1)
+            bounds = np.round(np.linspace(cutoff, np.log10(prob_arr).max(), 5),1)
             norm = mpl.colors.BoundaryNorm(bounds, cmap.N)
         else:
-            norm = mpl.colors.Normalize(vmin=np.log(prob_arr).max(),
+            norm = mpl.colors.Normalize(vmin=np.log10(prob_arr).max(),
                                         vmax=cutoff)
 
-        im = ax.pcolormesh(X, Y, np.log(prob_arr), cmap=cmap, norm=norm,
+        im = ax.pcolormesh(X, Y, np.log10(prob_arr), cmap=cmap, norm=norm,
                            shading='flat') # vs 'gouraud'
 
     else:
@@ -89,14 +89,24 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
         ax.plot(
             nparr(zdf['sma_AU'])*u.AU,
             (nparr(zdf['m_comp/m_sun'])*u.Msun).to(u.Mjup),
-            color='C0', lw=1
+            color='C0', lw=1,
+            zorder=4
         )
 
-        ax.text(
+        t = ax.text(
             100, 450, 'Ruled out by\nspeckle imaging',
             fontsize=7.5, ha='center', va='center',
-            path_effects=[pe.withStroke(linewidth=0.5, foreground="white")],
-            color='C0'
+            # path_effects=[pe.withStroke(linewidth=0.5, foreground="white")],
+            color='C0', zorder=3
+        )
+
+        ax.fill_between(
+            nparr(zdf['sma_AU'])*u.AU,
+            (nparr(zdf['m_comp/m_sun'])*u.Msun).to(u.Mjup),
+            1000,
+            color='white',
+            alpha=0.8,
+            zorder=2
         )
 
         _sma = np.logspace(np.log10(90), np.log10(180))
@@ -117,13 +127,14 @@ def plot_mass_semimaj_constraints(prob_arr=None, mass_grid=None,
     ax.set_yscale('log')
 
     ax.set_xlabel('Semi-major axis [AU]')
[…]

### 2020-01-31 paper/ms.tex
more word-smithing

@@
 The system's acceleration towards our line-of-sight causes a decrease
-in the apparent orbital period.  The expected period derivative is
+in the apparent orbital period:
 \begin{align}
-  \dot{P}_{\rm\,RV}
-  = \frac{\dot{v}_{\rm r} P}{c}
-  = -5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}.
+  \dot{P}_{\rm\,RV} &= \frac{\dot{v}_{\rm r} P}{c}.
+\end{align}
+Expressed in convenient units,
+\begin{align}
+  \dot{P}_{\rm\,RV} &= 105{.}3\,{\rm msec}\,{\rm yr}^{-1} \
+  \left( \frac{P}{{\rm day}} \right)
+  \left( \frac{\dot{\gamma}}{ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}} \right).
+\end{align}
+For WASP-4, this yields
+\begin{align}
+  \dot{P}_{\rm\,RV} &= -5.94 \pm 0.39\,{\rm msec}\,{\rm yr}^{-1}.
 \end{align}
@@
 ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
-perfect, Occam's razor would suggest that the line-of-sight
-acceleration is probably a sufficient explanation for the apparent
-decrease of WASP-4b's orbital period.
+perfect, Occam's razor would suggest that most or all of the apparent
+decrease of WASP-4b's orbital period is caused by the line-of-sight
+acceleration.
 
@@
 WASP-4's companion.
-
-% \citeauthor{baluev_2019} analyzed their newly obtained photometry,
-% along with archival light curves that we omitted from our analysis
-% due to systematic uncertainties in the absolute time system.
-% \citeauthor{baluev_2019} found that when they used all the available
-% TTV data, the need for a quadratic ephemeris was present.  However,
-% they pointed out that if they used lower-precision subsets of the
-% available timing data, the necessity for the quadratic term
-% decreased.  This is not particularly surprising.
-% \citeauthor{baluev_2019} also pointed out that the precise transit
-% times reported by \citet{huitson_gemini_2017} were quite important
-% in the time-series.  Overall, \citeauthor{baluev_2019} did not find
-% the claim of a period decrease convincing.
 
@@
 changes commensurate with WASP-4 due to acceleration by outer
-companions.  Half of these will be period decreases, and will be an
+companions.  Half of these will be period decreases\footnote{ At a
+population level, one would expect half of the line-of-sight
+accelerations to be positive, and half negative.  Tidal orbital decay
+however is usually negative, and might manifest as an asymmetry in the
+$\dot{P}$ distribution of hot Jupiters.}, and will be an
 astrophysical false positive in searches for tidal orbital decay.

### 2020-01-31 src/calculate_pdot_scaling_relations.py

@@ -1,0 +1,6 @@
+from astropy import units as u, constants as const
+import numpy as np
+
+const = (1*u.day)/(const.c) * (1*u.m/u.s/u.day)
+
+print(const.to(u.millisecond/u.year))

### 2020-02-03 paper/ms.tex
text: update observation descriptions

@@
 and the SAAO 1.0m telescope.
+% Five of the twenty-two occurred at simultaneous epochs as transits
+% previously reported by \citet{hoyer_tramos_2013}.
 
@@
 The complete set of radial velocity observations is given in Table~2.
-Along with the 2010-2019 HIRES observations, there are also earlier
+Along with the 2010-2019 HIRES observations, there are also
 measurements from CORALIE and HARPS.  Following
@@
 using the homogeneous radial velocities calculated by the latter
-authors. We included the HARPS values reported by
+authors.
+
+For HARPS, we included the values reported by
 \citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
-We omitted the HARPS data points taken over three nights by
-\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
-because they have a systematic offset from the remaining datasets.
+While \citet{triaud_spin-orbit_2010} also acquired HARPS data over
+three nights for Rossiter-McLaughlin observations, these data were
+reduced with a non-standard pipeline, and so are systematically offset
+from the remaining HARPS data.  In our final fit, we therefore omitted
+the three \citet{triaud_spin-orbit_2010} nights.  We did nonetheless
+experiment with using the recent \citet{trifonov_public_2020}
+re-reductions, which allowed us to include two calibration RVs taken
+as part of the \citet{triaud_spin-orbit_2010} program. The decision
+regarding whether to include or omit these points did not
+substantially affect any of our conclusions.
 
@@
 60$\,$msec exposures.  If a companion is present, the autocorrelation
-functions of these speckle images reveal a characteristic
-interference pattern. This pattern in then used to determine the
-properties of the detected companion and to produce a reconstructed
-image.  Using the reduced images and their
-reconstructed speckle images,  contrast curves are produced yielding
-5-sigma detection limits  (see \citealt{howell_speckle_2011}). 
-No companions were detected,
-and the second night, which had better seeing
-(0.6$''$), also produced the more constraining result.  The 832$\,$nm
-limits were the most useful to us given that faint companions
-are redder than the host star.  We therefore opted to use the
-832$\,$nm September 28-29 contrast limits for the remaining analysis.
+functions of these speckle images reveal a characteristic interference
+pattern. This pattern in then used to determine the properties of the
+detected companion and to produce a reconstructed image.  Using the
+reduced images and their reconstructed speckle images,  contrast
+curves are produced yielding 5-sigma detection limits  (see
+\citealt{howell_speckle_2011}).  No companions were detected, and the
+second night, which had better seeing (0.6$''$), also produced the
+more constraining result.  The 832$\,$nm limits were the most useful
+to us given that faint companions are redder than the host star.  We
+therefore opted to use the 832$\,$nm September 28-29 contrast limits
+for the remaining analysis.
 

### 2020-02-03 src/inflate_uncertainties.py
inflate pdot uncertainties s.t. chisq_red = 1

@@ -1,0 +1,36 @@
+"""
+One approach to analyzing such data would be to inflate the transit
+measurement uncertainties, and lower the reduced $\chi^2$.  We do not
+think that such an approach is warranted, because it would not change
+the result that the quadratic model is strongly preferred.  Instead,
+we opt to simply inflate the uncertainties reported for each model by
+a factor of $\sqrt{\chi^2_{\rm red}}$ ($\approx$1.73$\times$ for the
+linear model, and $\approx$1.41$\times$ for the quadtratic).
+"""
+
+from numpy import sqrt
+from astropy import units as u, constants as c
+
+# error provenance:
+# /Users/luke/Dropbox/proj/tessorbitaldecay/results/model_comparison/WASP-4b_20200127/model_comparison_output.txt
+
+linear_chisq_red = 3.0
+
+t0_err = 14
+P_err = 15
+print('t0_err: {} -> {}'.format(t0_err, sqrt(linear_chisq_red)*t0_err ))
+print('P_err: {} -> {}'.format(P_err, sqrt(linear_chisq_red)*P_err ))
+
+
+quad_chisq_red = 2.0
+t0_err = 22
+P_err = 17
+Pdot_err = 2.815
+print('t0_err: {} -> {}'.format(t0_err, sqrt(quad_chisq_red)*t0_err ))
+print('P_err: {} -> {}'.format(P_err, sqrt(quad_chisq_red)*P_err ))
+print('Pdot_err: {} -> {}'.format(Pdot_err, sqrt(quad_chisq_red)*Pdot_err ))
+
+Pdot_err = sqrt(quad_chisq_red)*Pdot_err * 1e-11* (u.m/u.m)
+print('Pdot_err: {} -> {}'.format(Pdot_err,
+                                  Pdot_err.to(u.millisecond/u.year) ))
+

### 2020-02-03 src/plot_O_minus_C_paperII.py

@@ -159,15 +159,16 @@ def plot_O_minus_C(
     #
     # Do the error bands
     #
+    chi2red = (182.708/90.) # /Users/luke/Dropbox/proj/tessorbitaldecay/results/model_comparison/WASP-4b_20200127/model_comparison_output.txt
     sigma_t0 = max(
         [perrs_3d['t0 [min]'], merrs_3d['t0 [min]']]
-    )
+    )*np.sqrt(chi2red)
     sigma_P = max(
         [perrs_3d['P [min]'], merrs_3d['P [min]']]
-    )
+    )*np.sqrt(chi2red)
     sigma_quad = max(
         [perrs_3d['0.5 dP/dE [min]'], merrs_3d['0.5 dP/dE [min]']]
-    )
+    )*np.sqrt(chi2red)
 
     model_tmid_upper = np.maximum.reduce([
         (

### 2020-02-03 paper/ms.tex
propagate uncertainty inflation thru text

@@
   0.39~{\rm msec}\,{\rm yr}^{-1}$ implied, compared to $-8.64 \pm
-  0.89~{\rm msec}\,{\rm yr}^{-1}$ observed).  Combining the radial
+  1.26~{\rm msec}\,{\rm yr}^{-1}$ observed).  Combining the radial
   velocities with new upper limits from speckle imaging, we find that
@@
 measurement uncertainties, and lower the reduced $\chi^2$.  We do not
-think that such an approach is warranted, since it would not change
-the result that the quadratic model is strongly preferred.
+think that such an approach is warranted, because it would not change
+the result that the quadratic model is strongly preferred.  Instead,
+we opt to simply inflate the uncertainties reported for each model by
+a factor of $\sqrt{\chi^2_{\rm red}}$ ($\approx$1.73$\times$ for the
+linear model, and $\approx$1.41$\times$ for the quadtratic).
 
-The best-fit period derivative in the quadratic model is
+The best-fit period derivative in the quadratic model is then
 \begin{equation}
@@
   = - (2.74 \pm 0.28)\times 10^{-10}
-  = - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}.
+  = - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
     produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
-    0.89~{\rm msec}\,{\rm yr}^{-1}$) is indicated with the purple dotted
+    1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated with the purple dotted
     line.  The four new RV measurements presented in this work
@@
 In other words, the majority of the period decrease observed in
-transits ($\dot{P}= - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) seems to
+transits ($\dot{P}= - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) seems to
 be explained by the acceleration of the host star.
@@
 option---light-travel time effects---is also the most likely.
-Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
+Transits show the obital period decreasing by $- 8.64 \pm 1.26~{\rm
 ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
@@
 where $a_0\equiv t_0$, $a_1\equiv P$, and $a_2\equiv 0.5\cdot {\rm
-d}P/{\rm d}E$.  By following \citet{gould_chi2_2003}, we derived that
+d}P/{\rm d}E$.  Following \citet{gould_chi2_2003}, one can show that
 if $N$ transit timing measurements are taken uniformly across a
@@
 observations continue from ground and space-based observatories,
-period derivatives with precision below $1~{\rm msec}\,{\rm yr}^{-1}$
-will become much more common over the coming decade.
+period derivatives will begin to be measured with precision exceeding
+$1~{\rm msec}\,{\rm yr}^{-1}$ over the coming decade.
 
@@
 majority of the period decrease observed in transits ($\dot{P}= - 8.64
-\pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
+\pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
 acceleration of the host star~---~not tidal orbital decay, or apsidal

### 2020-02-03 src/plot_rvs_2020.py
rv plot: minor changes
* update the pdot error bar to be inflated
* add a WIP plot that I'm not convinced is necessary, showing the RV
  orbit

@@ -14,15 +14,11 @@ from numpy import array as nparr
 from astropy.time import Time
 from astropy import units as units, constants as const
 
+from astrobase.lcmath import phase_magseries
+
 from radvel_utils import args_object, _get_fit_results
 
-def main(make_my_plot=1):
-    """
-    args:
-
-        make_my_plot: assumes you have run the radvel fit. reads in the output
-        values to plot.
-    """
+def plot_rvs_2020_single_residual():
 
     # initialization script used to make the fix_gammadot fits
     basedir = os.path.join(
@@ -76,8 +72,9 @@ def main(make_my_plot=1):
     # what would explain the Pdot from transits?
     period = 1.338231466*units.day
     Pdot_tra = -2.736e-10
-    Pdot_tra_perr = Pdot_tra + 2.83e-11
-    Pdot_tra_merr = Pdot_tra - 2.83e-11
+    Pdot_err = 2**(1/2.)*2.83e-11 # inflating appropriately
+    Pdot_tra_perr = Pdot_tra + Pdot_err
+    Pdot_tra_merr = Pdot_tra - Pdot_err
     dvdt_tra = (Pdot_tra * const.c / period).to(
         (units.m/units.s)/units.day).value
     dvdt_tra_perr = (Pdot_tra_perr * const.c / period).to(
@@ -133,7 +130,126 @@ def main(make_my_plot=1):
     print('saved {:s}'.format(savpath))
 
 
+
+def plot_rvs_2020_orbit_and_residual():
+    # [WIP] is this really necessary?
+
+    # initialization script used to make the fix_gammadot fits
+    basedir = os.path.join(
+        os.path.expanduser('~'),
+        "Dropbox/proj/WASP-4b_anomaly/"
+    )
+
+    setupfn = os.path.join(basedir,"src/WASP4.py")
+
+    outputdir = os.path.join(
+        basedir,
+        "results/rv_fitting/LGB_20190911_fix_gammaddot"
+    )
+    savpath='../results/20190911_rv_fit_orbit_and_residual.png'
+
+    (rvtimes, rvs, rverrs, resid, telvec, dvdt,
+     curv, dvdt_merr, dvdt_perr, time_base) = _get_fit_results(
+         setupfn, outputdir
+    )
+
+    #
+    # make the plot
+    #
+    offset=2450000
+
[…]

### 2020-02-03 paper/ms.tex
paper smithing

@@
 \citet{baluev_2019}, based on a homogeneous analysis of archival
-ground-based observations taken professional and amateur observers.
+ground-based observations.
 We included twelve of their ``high quality'' transit times from
@@
 from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
-using the homogeneous radial velocities calculated by the latter
-authors.
+using the homogeneous values calculated by the latter authors.
 
@@
 regarding whether to include or omit these points did not
-substantially affect any of our conclusions.
+noticeably affect our results.
 
@@
 
-A cursory analysis of the new HIRES observations led to our detection
+An initial analysis of the new HIRES observations led to our detection
 of a linear trend in the residuals after fitting out the orbit of
@@
 detected companion and to produce a reconstructed image.  Using the
-reduced images and their reconstructed speckle images,  contrast
-curves are produced yielding 5-sigma detection limits  (see
-\citealt{howell_speckle_2011}).  No companions were detected, and the
-second night, which had better seeing (0.6$''$), also produced the
-more constraining result.  The 832$\,$nm limits were the most useful
-to us given that faint companions are redder than the host star.  We
-therefore opted to use the 832$\,$nm September 28-29 contrast limits
-for the remaining analysis.
+reconstructed speckle images,  contrast curves are produced yielding
+5-sigma detection limits  (see \citealt{howell_speckle_2011}).  No
+companions were detected, and the second night, which had better
+seeing (0.6$''$), also produced the more constraining result.  The
+832$\,$nm limits were the most useful to us given that faint
+companions are redder than the host star.  We therefore opted to use
+the 832$\,$nm September 28-29 contrast limits for the remaining
+analysis.
 
@@
   calculated times assuming a constant orbital period.  More opaque
-  points correspond to more precise data.  The quadratic ephemeris and
-  its $\pm1\sigma$ statistical uncertainties are shown in blue.
-  The scatter in the residuals could be due to underestimated
-  measurement uncertainties (see Section~\ref{sec:transit_analysis}).
+  points correspond to more precise data.  The $\pm1\sigma$
+  uncertainties of the quadratic ephemeris 
+  are shown in blue.
   The binned TESS point (yellow star) is the weighted average of 18
@@
 and models is $\Delta {\rm BIC} = 89$, strongly favoring the
-quadratic case \citep{kass_bayes_1995}.
+quadratic model \citep{kass_bayes_1995}.
 
 From the reduced $\chi^2$ values, we can surmise that neither model
-entirely describes the transit data that we have collected---there
+entirely describes the transit data---there
 must be some additional source of signal or noise. In
@@
 we opt to simply inflate the uncertainties reported for each model by
-a factor of $\sqrt{\chi^2_{\rm red}}$ ($\approx$1.73$\times$ for the
-linear model, and $\approx$1.41$\times$ for the quadtratic).
+a factor of $(\chi^2_{\rm red})^{1/2}$, or $\approx$1.73$\times$ for the
+linear model, and $\approx$1.41$\times$ for the quadratic.
 
-The best-fit period derivative in the quadratic model is then
+The resulting best-fit period derivative for the quadratic model is 
 \begin{equation}
@@
 \citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
-ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
+ms}\,{\rm yr}^{-1}$).  It is $\approx$2.3$\sigma$ larger than the rate
 of period decrease we previously reported \citepalias[$- 12.6 \pm 1.2~{\rm
@@
 \begin{align}
-  \dot{P}_{\rm\,RV} &= \frac{\dot{v}_{\rm r} P}{c}.
+  \dot{P}_{\rm\,RV} &= \frac{\dot{v}_{\rm r} P}{c},
 \end{align}
-Expressed in convenient units,
+or in more convenient units,
 \begin{align}
@@
 \end{align}
-In other words, the majority of the period decrease observed in
-transits ($\dot{P}= - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) seems to
-be explained by the acceleration of the host star.
+The majority of the period decrease observed in transits ($\dot{P}= -
+8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) seems to be explained by the
+acceleration of the host star.
 
@@
 \input{WASP-4b_transit_time_table.tex}
+\vspace{-2cm}
 \input{WASP-4b_rv_table.tex}
+\vspace{-2cm}
 \input{model_fit_table.tex}
+\vspace{-2cm}
 \input{pdot_table.tex}

### 2020-02-03 paper/ms.tex
minor paper edits

@@
 \begin{abstract}
-  Space and ground-based transit observations of the hot Jupiter
+  Space and ground-based transit measurements of the hot Jupiter
   WASP-4b have recently shown that its orbital period appears to be
@@
   hot Jupiters are expected to show comparable period decreases to
-  WASP-4, due to acceleration by outer companions.  We find that these
+  WASP-4, due to acceleration by outer companions.  These
   period decreases will be increasingly measureable in coming years,
-  because the precision of period derivative measurement scales as
-  (observing baseline)$^{5/2}$.  Continued radial velocity monitoring
+  because the precision of the period derivative measurement scales with
+  the observing baseline to the power of 5/2.  Continued radial velocity monitoring
   of hot Jupiters is therefore essential to distinguish tidal orbital
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f1.pdf}
+		\includegraphics[width=0.47\textwidth]{f1.pdf}
 	\end{center}
@@
   The binned TESS point (yellow star) is the weighted average of 18
-  TESS transits and is binned for display purposes only.  Its
-  uncertainty is slightly larger than the marker size.  The models
+  TESS transits and is binned for display purposes only.  The models
   were fitted to all of the individual transit times.
@@
   = - (2.74 \pm 0.28)\times 10^{-10}
-  = - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
+  = - 8w.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f2.pdf}
+		\includegraphics[width=0.47\textwidth]{f2.pdf}
 	\end{center}
@@
     produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
-    1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated with the purple dotted
-    line.  The four new RV measurements presented in this work
-    increased the significance of the radial velocity trend from
-    $\approx$2$\sigma$ to 15$\sigma$.
+    1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated in purple.  The four
+    new RV measurements presented in this work increase the
+    significance of the radial velocity trend from $\approx$2$\sigma$
+    to 15$\sigma$.
 	\label{fig:rvs}
@@
 
-In the preferred ``planet+trend'' model, WASP-4 is accelerating
+In the preferred ``planet+linear trend'' model, WASP-4 is accelerating
 towards our line-of-sight at high confidence,
@@
 \end{align}
-The majority of the period decrease observed in transits ($\dot{P}= -
-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) seems to be explained by the
+The majority of the period decrease seen in transits ($\dot{P}= -
+8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) therefore seems to be caused by the
 acceleration of the host star.
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f3.pdf}
+		\includegraphics[width=0.47\textwidth]{f3.pdf}
 	\end{center}
@@
       have been detected.  The inset shows the speckle image
-      reconstructed from 1000 40 millisecond frames in an 832$\,$nm
+      reconstructed from 1000 60 millisecond frames in an 832$\,$nm
       bandpass, and acquired on September 28, 2019.  The image scale
@@
 		\leavevmode
-		\includegraphics[width=0.78\textwidth]{f4.pdf}
+		\includegraphics[width=0.73\textwidth]{f4.pdf}
 	\end{center}
@@
 velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
-ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
-perfect, Occam's razor would suggest that most or all of the apparent
+ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is $\approx$2$\sigma$
+discrepant, Occam's razor would suggest that most or all of the apparent
 decrease of WASP-4b's orbital period is caused by the line-of-sight
@@
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.4cm}
 	\caption{
@@
   \citet{knutson_friends_2014} have shown long-term radial velocity
-  trends.  HAT-P-11 signal is plotted, though its signal may be caused
-  by stellar activity.  Three hot Jupiters are not plotted because
-  their radial velocity curves better described as quadratic trends in
-  time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered in the $y$
-  dimension by the absolute value of d$P$/d$t$.
+  trends.  HAT-P-11 is shown, though its signal may be caused
+  by stellar activity.  Three hot Jupiters are not shown because
+  their radial velocity curves are better described as quadratic
+  trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered
+  in the $y$ dimension by the absolute value of d$P$/d$t$.
 	\label{fig:pdot_pop}
@@
 accelerations to be positive, and half negative.  Tidal orbital decay
-however is usually negative, and might manifest as an asymmetry in the
+however is usually negative, and could evenutally manifest as an asymmetry in the
 $\dot{P}$ distribution of hot Jupiters.}, and will be an
@@
 
-For hot Jupiters that have been monitored for over baselines of
-$\approx$10 years, secular changes in their orbital periods are
-currently being constrained to a precision of $\lesssim 10~{\rm
-ms}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
+For hot Jupiters that have been monitored over baselines
+exceeding 10 years, secular changes in their orbital periods are
+currently being constrained to a precision of $\lesssim$$10~{\rm
+msec}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
 This is roughly commensurate with the level of signal many outer
@@
 if $N$ transit timing measurements are taken uniformly across a
-baseline of $\Delta E$ epochs with constant precision $\sigma_{\rm
-tra}$, then the uncertainty of the quadratic term is given by
+baseline of $\Delta E$ epochs with constant precision $\sigma$, then
+the uncertainty of the quadratic term is given by
 \begin{align}
@@
 observations continue from ground and space-based observatories,
-period derivatives will begin to be measured with precision exceeding
-$1~{\rm msec}\,{\rm yr}^{-1}$ over the coming decade.
+period derivatives will be measured with precision exceeding
+$1~{\rm msec}\,{\rm yr}^{-1}$ within the coming decade.
 
@@
 and period changes induced by these outer companions will become an
-increasingly large nuisance in the hunt for orbital decay as the
+increasingly large nuisance in the hunt for tidal orbital decay as the
 observational baselines get longer.  In particular, the precision with
@@
 To distinguish this effect from tidal decay, further long-term radial
-velocity measurements of hot Jupiters are required.
+velocity measurements of hot Jupiters are strongly encouraged.
 

### 2020-02-03 src/plot_rvs_paperII.py
add plot_rvs_paperII
because the old versions were doing the residuals wrong, so it made more
sense to go the radvel path

@@ -1,0 +1,743 @@
+# env: py37_emcee2
+
+import radvel
+from radvel import driver
+from radvel.plot import orbit_plots, mcmc_plots
+from radvel.mcmc import statevars
+from radvel import plot
+from radvel.utils import t_to_phase, fastbin, sigfig
+from radvel.driver import load_status, save_status
+import configparser, os, emcee
+
+import numpy as np
+import matplotlib
+from matplotlib import rcParams, gridspec
+from matplotlib import pyplot as plt
+from matplotlib.cm import nipy_spectral
+from matplotlib.ticker import MaxNLocator
+from matplotlib.offsetbox import AnchoredText
+from astropy.time import Time
+
+from astropy import units as units, constants as const
+
+##########
+# config #
+##########
+latex = {
+    'ms': r'm s$^{\mathregular{-1}}$',
+    'BJDTDB': r'BJD$_{\mathregular{TDB}}$'
+}
+
+telfmts_default = {
+    'j': dict(color='C2', marker=u'o', label='HIRES', mew=1),
+    'k': dict(color='k', fmt='s', mfc='none', label='HIRES pre 2004', mew=1),
+    'a': dict(color='g', fmt='d', label='APF'),
+    'pfs': dict(color='magenta', fmt='p', label='PFS'),
+    'CORALIE': dict(color='C0', fmt='d', label='CORALIE'),
+    'h': dict(color='C1', fmt="s", label='HARPS'),
+    'harps-n': dict(color='firebrick', fmt='^', label='HARPS-N'),
+    'l': dict(color='g', fmt='*', label='LICK'),
+}
+telfmts_default['lick'] = telfmts_default['l']
+telfmts_default['hires_rj'] = telfmts_default['j']
+telfmts_default['hires'] = telfmts_default['j']
+telfmts_default['hires_rk'] = telfmts_default['k']
+telfmts_default['apf'] = telfmts_default['a']
+telfmts_default['harps'] = telfmts_default['h']
+telfmts_default['LICK'] = telfmts_default['l']
+telfmts_default['HIRES_RJ'] = telfmts_default['j']
+telfmts_default['HIRES'] = telfmts_default['j']
+telfmts_default['HIRES_RK'] = telfmts_default['k']
+telfmts_default['APF'] = telfmts_default['a']
+telfmts_default['HARPS'] = telfmts_default['h']
[…]

### 2020-02-03 src/plot_rvs_2020.py
cleanup old rv plots

@@ -1,281 +1,0 @@
-"""
-plot RV vs time, and residual (with best fit line model)
-"""
-
-import matplotlib as mpl
-mpl.use('Agg')
-import matplotlib.pyplot as plt, pandas as pd, numpy as np
-
-from glob import glob
-from shutil import copyfile
-import os, pickle
-
-from numpy import array as nparr
-from astropy.time import Time
-from astropy import units as units, constants as const
-
-from astrobase.lcmath import phase_magseries
-
-from radvel_utils import args_object, _get_fit_results
-
-def plot_rvs_2020_data_and_residual():
-
-    # initialization script used to make the fix_gammadot fits
-    basedir = os.path.join(
-        os.path.expanduser('~'),
-        "Dropbox/proj/WASP-4b_anomaly/"
-    )
-
-    setupfn = os.path.join(basedir,"src/WASP4.py")
-
-    outputdir = os.path.join(
-        basedir,
-        "results/rv_fitting/LGB_20190911_fix_gammaddot"
-    )
-    savpath='../results/20190911_rv_data_and_residual.png'
-
-    (rvtimes, rvs, rverrs, resid, telvec, dvdt,
-     curv, dvdt_merr, dvdt_perr, time_base) = _get_fit_results(
-         setupfn, outputdir
-    )
-
-    #
-    # make the plot
-    #
-    offset=2450000
-
-    fig, (a0, a1) = plt.subplots(nrows=2, ncols=1, sharex=True,
-                               figsize=(0.8*6,0.8*5.5), gridspec_kw=
-                               {'height_ratios':[3, 2]})
-
-    utel = np.unique(telvec)
-    markers = ['o','s','^']
-    for ix, tel in enumerate(utel):
-        sel = (telvec == tel)
-        a0.errorbar(rvtimes[sel]-offset, rvs[sel], rverrs[sel], marker=markers[ix],
-                    ecolor='gray', zorder=10, mew=0, ms=4, elinewidth=1,
-                    color='C{}'.format(ix), label=tel, lw=0)
-
-        a1.errorbar(rvtimes[sel]-offset, resid[sel], rverrs[sel], marker=markers[ix],
-                    ecolor='gray', zorder=10, mew=0, ms=4, elinewidth=1,
-                    color='C{}'.format(ix), lw=0)
-
[…]

### 2020-02-03 src/plot_rvs_NEWDATA.py

@@ -1,311 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-plot RV vs time, and residual (with best fit line model)
-"""
-from __future__ import division, print_function
-
-import matplotlib as mpl
-mpl.use('Agg')
-import matplotlib.pyplot as plt, pandas as pd, numpy as np
-
-from glob import glob
-from shutil import copyfile
-import os, pickle
-
-from astropy.time import Time
-from astropy import units as units, constants as const
-
-from numpy import array as nparr
-
-import radvel
-from radvel.plot import orbit_plots, mcmc_plots
-from radvel.mcmc import statevars
-from radvel.driver import save_status, load_status
-import configparser
-
-##########################################
-# a bunch of stuff needed to hack at radvel objects
-
-class args_object(object):
-    """
-    a minimal version of the "parser" object that lets you work with the
-    high-level radvel API from python. (without directly using the command line
-    interface)
-    """
-    def __init__(self, setupfn, outputdir):
-        # return args object with the following parameters set
-        self.setupfn = setupfn
-        self.outputdir = outputdir
-        self.decorr = False
-        self.plotkw = {}
-        self.gp = False
-
-def _hack_radvel_plots(args):
-    """
-    Hacked version of radvel.driver.plots
-
-    Args:
-        args (ArgumentParser): command line arguments
-    """
-
-    config_file = args.setupfn
-    conf_base = os.path.basename(config_file).split('.')[0]
-    statfile = os.path.join(
-        args.inputdir, "{}_radvel.stat".format(conf_base)
-    )
-
-    status = load_status(statfile)
-
-    assert status.getboolean('fit', 'run'), \
-        "Must perform max-liklihood fit before plotting"
-    post = radvel.posterior.load(status.get('fit', 'postfile'))
-
-    # update the plotted posterior to match the median best-fit parameters.
-    summarycsv = "../results/rv_fitting/t10onlytwo_LGB_20190228_fix_gammaddot/WASP4_post_summary.csv"
-    sdf = pd.read_csv(summarycsv)
-    for param in [c for c in sdf.columns if 'Unnamed' not in c]:
[…]

### 2020-02-03 paper/ms.tex
compile paper

@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f1.pdf}
+		\includegraphics[width=0.7\textwidth]{f1.pdf}
 	\end{center}
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f2.pdf}
+		\includegraphics[width=0.7\textwidth]{f2.pdf}
 	\end{center}
@@
 	\caption{
-    {\bf Radial velocity residuals of WASP-4.} The best-fit
-    Keplerian orbit of WASP-4b has been subtracted.  The linear trend
-    inferred from the RV data is shown with a black line, with
-    $1\sigma$ errors in gray.  The trend that would be needed to
+    {\bf Radial velocities of WASP-4.}
+    {\it Top:} RV measurements, with best-fit instrument offsets
+    added. 
+    {\it Middle:}
+    Residuals, after subtracting the
+    maximum-likelihood orbit of WASP-4b.  The linear trend
+    inferred from the RV data is shown with a black line.
+    The trend that would be needed to
     produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
     1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated in purple.  The four
-    new RV measurements presented in this work increase the
-    significance of the radial velocity trend from $\approx$2$\sigma$
+    new RV measurements from this work increase the
+    significance of the linear trend from $\approx$2$\sigma$
     to 15$\sigma$.
+    {\it Bottom:}
+    Phased orbit of WASP-4b.
 	\label{fig:rvs}
@@
 		\leavevmode
-		\includegraphics[width=0.73\textwidth]{f4.pdf}
+		\includegraphics[width=0.8\textwidth]{f4.pdf}
 	\end{center}
@@
 
-\bibliographystyle{yahapj}                            
-\bibliography{bibliography}
-
 \input{WASP-4b_transit_time_table.tex}
@@
 
+%\clearpage
+\bibliographystyle{yahapj}                            
+\bibliography{bibliography}
+
 \end{document}

### 2020-02-03 paper/collect.sh

@@ -1,5 +1,5 @@
 cp ../results/O_minus_C_20200127_onlytransits.pdf f1.pdf
-cp ../results/20190911_rv_fit.pdf f2.pdf
+cp ../results/rv_fitting/LGB_20190911_fix_gammaddot_paperII/WASP4_rv_multipanel.pdf f2.pdf
 cp ../results/zorro_speckle.pdf f3.pdf
 cp ../results/mass_semimaj_constraints_with_contrast_discretecolor.pdf f4.pdf
 cp ../results/k14_pdot.pdf f5.pdf

### 2020-02-03 paper/ms.tex
tweaked tables

@@
   statistics from \citet{knutson_friends_2014} imply that about 1 in 6
-  hot Jupiters are expected to show comparable period decreases to
-  WASP-4, due to acceleration by outer companions.  These
-  period decreases will be increasingly measureable in coming years,
-  because the precision of the period derivative measurement scales with
-  the observing baseline to the power of 5/2.  Continued radial velocity monitoring
-  of hot Jupiters is therefore essential to distinguish tidal orbital
-  decay from line-of-sight accelerations.
+  hot Jupiters are expected to show period decreases comparable to
+  WASP-4, due to acceleration by outer companions.  These period
+  changes will be increasingly measureable in coming years, because
+  the precision of the period derivative measurement scales with the
+  observing baseline to the power of 5/2.  Continued radial velocity
+  monitoring of hot Jupiters is therefore essential to distinguish
+  tidal orbital decay from line-of-sight accelerations.
 \end{abstract}
@@
 
-We fitted each of the two models by assuming a Gaussian likelihood and
+We fitted each model by assuming a Gaussian likelihood and
 sampling over the posterior probability distributions.  We sampled the
@@
   = - (2.74 \pm 0.28)\times 10^{-10}
-  = - 8w.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
+  = - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
   \label{eq:dP_dt_obs}
@@
 
-\subsection{Radial Velocities: WASP-4's acceleration towards the Earth}
+\subsection{Radial velocities}
 
@@
 	\label{fig:rvs}
-  \vspace{-0.3cm}
+  \vspace{-0.0cm}
 	}
@@
 \end{equation}
+The remaining radial velocity model parameters are listed in Table~4.
 For comparison, before our new measurements, $\dot{\gamma}$ was
@@
 	\end{center}
-	\vspace{-0.6cm}
+	\vspace{-0.8cm}
 	\caption{
@@
 derivatives $\dot{P}_{\rm\,RV} = \dot{v}_{\rm r} P/c$ for each system.
-The results are given in Table~4, and visualized for hot Jupiters with
+The results are given in Table~5, and visualized for hot Jupiters with
 significant ($>$$3\sigma$) linear trends in Figure~\ref{fig:pdot_pop}.
@@
 \input{WASP-4b_transit_time_table.tex}
-\vspace{-2cm}
 \input{WASP-4b_rv_table.tex}
-\vspace{-2cm}
 \input{model_fit_table.tex}
-\vspace{-2cm}
+\input{rv_model_posterior_table.tex}
 \input{pdot_table.tex}

### 2020-02-03 paper/pdot_table.tex
add publication-ready Pdot and RV tables
tweaked tables

@@
 %
-\tablenum{4}
+\tablenum{5}
 %
@@
 \tablecomments{
-  Table~4 is published in its entirety in a machine-readable format.
-  A single row is shown here for guidance regarding form and content.
+  Table~5 is published in its entirety in a machine-readable format.
+  The first entry is shown for guidance regarding form and content.
+  Orbital periods were retrieved from NASA's Exoplanet Archive.
+  Additional comments regarding non-linear trends and stellar activity
+  are included in the MRT.
 }

### 2020-02-04 paper/ms.tex
minor text edits

@@
   -0.0422^{+0.0028}_{-0.0027}\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
-  The implied period decrease explains most or all of WASP-4's
+  The implied period decrease explains most or all of WASP-4b's
   changing orbital period as a light-travel time effect ($-5.94 \pm

### 2020-02-04 paper/ms.tex
Update ms.tex

--- 2020-02-03 paper/ms.tex
@@
 
-\title{WASP-4 is Accelerating Towards the Earth}
+\title{WASP-4 is Accelerating Towards the Sun}
 % ", Explaining its Decreasing Orbital Period"
@@
 \begin{abstract}
-  Space and ground-based transit measurements of the hot Jupiter
-  WASP-4b have recently shown that its orbital period appears to be
-  decreasing.  Proposed explanations for the period change include
-  tidal orbital decay, orbital precession, and light-travel time
-  effects.  We present new radial velocity measurements of WASP-4,
-  acquired with Keck-HIRES.  The data show that the system is
-  accelerating towards the Earth at $\dot{\gamma} =
-  -0.0422^{+0.0028}_{-0.0027}\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
-  The implied period decrease explains most or all of WASP-4's
-  changing orbital period as a light-travel time effect ($-5.94 \pm
-  0.39~{\rm msec}\,{\rm yr}^{-1}$ implied, compared to $-8.64 \pm
-  1.26~{\rm msec}\,{\rm yr}^{-1}$ observed).  Combining the radial
-  velocities with new upper limits from speckle imaging, we find that
-  the system's acceleration is likely caused by a 10-200$\,M_{\rm
-  Jup}$ companion with semi-major axis between 10-100$\,$AU.  The
-  statistics from \citet{knutson_friends_2014} imply that about 1 in 6
-  hot Jupiters are expected to show period decreases comparable to
-  WASP-4, due to acceleration by outer companions.  These period
-  changes will be increasingly measureable in coming years, because
-  the precision of the period derivative measurement scales with the
-  observing baseline to the power of 5/2.  Continued radial velocity
+  The orbital period of the hot Jupiter WASP-4b appears to be decreasing at
+  a rate of $-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$,
+  based on transit-timing measurements spanning 12 years.
+  Proposed explanations for the period change include
+  tidal orbital decay, apsidal precession, and acceleration of the system
+  along the line of sight.
+  To investigate further, we performed new radial velocity measurements and
+  speckle imaging of WASP-4.
+  The radial-velocity data show that the system is
+  accelerating towards the Sun at a rate of
+  $-0.0422\pm 0.0028\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
+  The associated Doppler effect would cause
+  the apparent period to shrink at a rate of
+  $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$, comparable
+  to the observed rate.
+  Thus, the observed change in the transit period is mostly
+  or entirely produced by the line-of-sight acceleration of the system.
+  This acceleration is probably caused by a wide-orbiting companion
+  of mass 10-200$\,M_{\rm Jup}$ and orbital distance 10-100$\,$AU,
+  based on the magnitude of the radial-velocity trend and the non-detection of any
+  companion in the speckle image.  We expect that 1 out of 6 hot Jupiters
+  will show apparent period changes of the same order of magnitude as WASP-4,
+  based on the hot-Jupiter companion statistics of  
+  \citet{knutson_friends_2014}. Continued radial velocity
   monitoring of hot Jupiters is therefore essential to distinguish
-  tidal orbital decay from line-of-sight accelerations.
+  the effects of tidal orbital decay or apsidal precession from line-of-sight acceleration.
 \end{abstract}
@@
 
-The orbits of most hot Jupiters are unstable to tidal decay
-\citep{counselman_outcomes_1973,hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
-The relevant issue is whether the timescale for tidal orbital decay is
+The orbits of most hot Jupiters are formally unstable to tidal decay
+\citep{counselman_outcomes_1973,
+hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
+It is not clear, though, whether the timescale for tidal orbital decay is
 shorter or longer than the timescale for main-sequence stellar
-evolution.  This question depends on the uncertain rate at which
-friction inside the star can damp the energy of tidal oscillations and
-thereby shrink the orbit (as reviewed by \citealt{Mazeh2008} and
+evolution.  This answer to this question depends on the uncertain rate at which
+friction inside the star damps the tidal oscillations, which is ultimately
+what causes the orbit to shrink
+(as reviewed by \citealt{Mazeh2008} and
 \citealt{ogilvie_tidal_2014}).
-
-Indirect studies of age and angular momentum indicators including the
-hot Jupiter semi-major axis distribution, host star spin rates, and
-galactic velocity dispersions have led to estimates for the inspiral
-timescale that vary from much less to much greater than the
-main-sequence evolution time  ({\it e.g.},
+Population studies of hot Jupiters --- based on ages, rotation rates,
+orbital distances, and Galactic kinematics --- have led to differing
+conclusions, with estimated lifetimes ranging from less than a gigayear
+to much longer than main-sequence lifetimes
+(see, {\it e.g.},
 \citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
 \citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
-\citealt{hamer_schlaufman_2019}).  Direct measurements of tidal
-orbital decay through transit and occultation timing could provide an
-empirical resolution.  For instance, combined transit timing and
-radial velocity measurements for WASP-12b have shown that its secular
-period decrease of $\approx$30 milliseconds per year is probably due
-to tidal orbital decay
+\citealt{hamer_schlaufman_2019}).
+
+An empirical resolution might be possible through long-term timing
+of transits and occultations, seeking evidence for changes in the orbital
+period.  For instance, long-term transit timing and
+radial velocity measurements for WASP-12b have revealed a secular
+decrease in the period at a rate of $\approx$30 milliseconds per year, which has
+been interpreted as the effect of tidal orbital decay
 \citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
 
-This study highlights a point that, though obvious, has perhaps not
-yet received due attention.  The point is that observational programs
+This study draws attention to a confounding factor that, while elementary,
+does not seem to have
+received the attention it deserves.
+The point is that observational programs
 aimed at identifying orbital decay in hot Jupiters through transit
-timing will be crippled without concurrent long-term radial velocity
-monitoring.  The reason is that line-of-sight accelerations due to
-outer companions \citep[{\it e.g.},][]{agol_detecting_2005} and tidal
-orbital decay both initially manifest identically in transit times, as
-a non-zero period derivative.  Massive outer companions to hot
-Jupiters are the norm; \citet{bryan_statistics_2016} calculated an
+timing must be accompanied by concurrent long-term radial velocity
+monitoring.  The reason is that an apparent change in period
+can be produced by the Doppler effect associated with acceleration
+of the hot-Jupiter host star along the line of sight, such as the acceleration that
+might be produced by a massive, wide-orbiting companion.
+Massive outer companions to hot Jupiters are common.
+\citet{bryan_statistics_2016} calculated an
 occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
 Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
-from 5-100$\,$AU.  Therefore it could very well be that more hot
-Jupiters will show shrinking orbital periods due to long term
-accelerations than due to tidal orbital decay.
+from 5-100$\,$AU.  Therefore, we expect that many hot Jupiters will display
+secular trends in orbital period that are unrelated to tidal orbital decay.
+This possibility can be checked by performing long-term radial-velocity monitoring
+at a level sensitive enough to detect or rule out the relevant amplitude of
+acceleration.
 
-The main focus of this study is the hot Jupiter WASP-4b, which has an
+The focus of this study is the hot Jupiter WASP-4b, which has an
 orbital period that appears to be decreasing by about 10 milliseconds
-per year.  We discovered the timing variations by combining data from
-TESS \citep{ricker_transiting_2015} with a decade of ground-based
-observations \citep[][hereafter
-\citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.  Thereafter,
-\citet{southworth_transit_2019} reported 22 new transit times for the
-system, and found an updated decay rate of $\dot{P} = -9.2 \pm 1.1$
-milliseconds per year. The \citeauthor{southworth_transit_2019} decay
-rate was $\approx$3$\sigma$ less rapid than that found by
-\citetalias{bouma_wasp4b_2019}, but the conclusions of the studies
-were otherwise similar.  A separate study by \citet{baluev_2019}
-reported additional archival transit light curves of WASP-4b from
-TRAPPIST and select other observers.  \citet{baluev_2019} pointed out
-that when using lower-precision subsets of the available transit data,
-the case for a decreasing period worsened.
+per year.  The period decrease was identified by \citet{bouma_wasp4b_2019}]{bouma_wasp4b_2019},
+who combined data from
+the NASA TESS mission \citep{ricker_transiting_2015} and a decade of ground-based
+transit observations. Soon thereafter, \citet{southworth_transit_2019} reported an additional
+22 transit times and recalculated the period derivative to be $\dot{P} = -9.2 \pm 1.1$
+milliseconds per year.  A separate study by \citet{baluev_2019}
+reported on additional transit times, and pointed out
+that the period decrease was statistically significant only when
+analyzing the data with the highest precision.
 
 To determine the origin of the period change, we acquired four
-additional radial velocity measurements using Keck-HIRES, extending
-the RV baseline from 3 to 9 years.  Previously, the five available
-HIRES radial velocities suggested a weak ($\approx$$2\sigma$) linear
-trend \citep{knutson_friends_2014}.  Our new measurements reveal a
+additional radial velocity measurements with the Keck I 10m telescope
+and the High Resolution Echelle
+Spectrometer (HIRES; \citealt{vogt_hires_1994}).
+In doing so, we extended
+the time baseline of HIRES measurements from 3 to 9 years.
+The previously available HIRES data led to the marginal ($\approx$$2\sigma$)
+detection of a radial-velocity trend \citep{knutson_friends_2014}.
+Our new measurements reveal a
 line-of-sight acceleration of $\dot{\gamma} =
--0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,day^{-1}}$.  This
-translates to an expected period decrease from the light-travel time
-effect of -5.9 milliseconds per year---about commensurate with what is
-observed from transits.  While the apparent period change caused by a
+-0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,day^{-1}}$.
+Through the Doppler effect\footnote{While the apparent period change caused by a
 line-of-sight acceleration has been referred to as the ``R{\o}mer
-effect'' \citep{yee_orbit_2020}, we prefer to avoid this term, which
-typically signifies arrival time delay due to observatory motion. We
-are discussing the Doppler effect seen by a stationary observer for a
-source with constant line-of-sight acceleration.
+effect'' \citep{yee_orbit_2020}, a simpler and probably better term is
+the Doppler effect. The R{\o}mer effect refers to the {\it delay} in
+the reception of a signal due to a change in the time required for
+light to traverse the distance between the source and observer. The Doppler
+effect refers to the change in the apparent {\it rate} of a process due
+to changes in the relative motion of the source and observer, such
+as the rate of transits.}, this translates into an expected period decrease
+$-5.9$ milliseconds per year, which is comparable to the period decrease
+that was measured from transit timing.  We undertook high-resolution
+(speckle) imaging to search for evidence of a companion that could
+be responsible for the acceleration of WASP-4.
 
-In the following, Section~\ref{sec:observations} collects the
-available transit data and presents the new radial velocity and
-speckle imaging observations.  Section~\ref{sec:analysis} analyzes the
-data, and finds that they yield a picture in which the WASP-4 system
-is accelerating towards our line-of-sight, likely due to the pull of a
-brown or M-dwarf companion.  Section~\ref{sec:discussion} places this
-result in the broader context of orbital decay searches, and points
-out that line-of-sight accelerations, {\it i.e.}, ``false positive
-orbital decay signals'', are relatively common in the hot Jupiter
-population.  Section~\ref{sec:conclusions} offers concluding remarks.
+Section~\ref{sec:observations} of this paper presents all of the
+available transit data as well as the new radial velocity and
+speckle imaging observations.  Section~\ref{sec:analysis} describes
+our analysis of the data, and our interpretation
+that WASP-4 is being pulled around by a brown dwarf or low-mass star.
+Section~\ref{sec:discussion} places this
+result within the context of orbital decay searches, and points
+out that line-of-sight accelerations will be a relatively
+common type of ``false positive.''
+Section~\ref{sec:conclusions} offers concluding remarks.
 
@@
 Table~1 lists the transit times we collected for our analysis.  We
-include data from the peer-reviewed literature for which {\it (i)} the
-analysis was based on observations of a single transit, {\it (ii)} the
-midpoint was fitted as a free parameter, and {\it (iii)} the time
-system specified both the leap second correction (TDB or UTC) and also
-whether any barycentric or heliocentric corrections had been
-performed.
+included data from the peer-reviewed literature for which {\it (i)} the
+reported time was based on the data from a single transit (as opposed
+to fitting the data spanning multiple transits and assuming the
+period to be constant), {\it (ii)} the
+central transit time was allowed to be a free parameter, and {\it (iii)} the time
+system was documented clearly, in particular specifying
+whether barycentric or heliocentric corrections had been
+performed and whether leap seconds had been taken into account (TDB or UTC).
 
@@
 % done)
-The majority of times are identical to those we collected in
+Most of these data are identical to the data presented by
 \citetalias{bouma_wasp4b_2019}.  Twenty-two new times reported by
 \citet{southworth_transit_2019} are included.  These transits were
-observed from the 3.58m NTT and Danish 1.54m telescopes at La Silla,
-and the SAAO 1.0m telescope.
+observed from the 3.58m New Technology Telescope and the Danish 1.54m telescope at La Silla, Chile,
+and the South African Astronomical Observatory 1.0m telescope.
 % Five of the twenty-two occurred at simultaneous epochs as transits
 % previously reported by \citet{hoyer_tramos_2013}.
-
-Additional timing measurements were also recently made available by
+Additional timing measurements were also reported recently by
 \citet{baluev_2019}, based on a homogeneous analysis of archival
@@
 We included twelve of their ``high quality'' transit times from
-TRAPPIST (six transits), El~Sauce (four transits), and
+the TRAPPIST telescope (six transits), El~Sauce (four transits), and
 \citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
@@
 ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
-whether leap-second corrections had or had not been performed.
-
-The four available occultations tabulated by
-\citetalias{bouma_wasp4b_2019} have neglegible statistical value due
-to their large uncertainties, and we forgo their use in this analysis.
+whether leap-second corrections had been performed.
+We did not include in our analysis the four occultation times tabulated by
+\citetalias{bouma_wasp4b_2019}, because of the large timing
+uncertainties and negligible statistical power.
 
@@
 
-After identifying the period decrease in
-\citetalias{bouma_wasp4b_2019}, we acquired four additional radial
-velocity measurements with the Keck High Resolution Echelle
-Spectrometer (HIRES; \citealt{vogt_hires_1994}).  Our observations
-were acquired using the standard setup and reduction techniques of the
+We acquired four new radial
+velocity measurements with Keck/HIRES.  Our observations
+were performed using the standard setup and reduction techniques of the
 California Planet Survey \citep{howard_cps_2010}.  Previously, the
@@
 The complete set of radial velocity observations is given in Table~2.
-Along with the 2010-2019 HIRES observations, there are also
-measurements from CORALIE and HARPS.  Following
-\citetalias{bouma_wasp4b_2019}, we included the CORALIE measurements
+Along with the 2010--2019 HIRES observations are early
+measurements with two different spectrographs.
+We included data from CORALIE ({\bf Spell it out on first use)}
 from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
-using the homogeneous values calculated by the latter authors.
-
-For HARPS, we included the values reported by
-\citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
+adopting the values from the homogeneous analysis of the latter authors.
+We included data from HARPS ({\bf spell it out})
+from \citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
 While \citet{triaud_spin-orbit_2010} also acquired HARPS data over
 three nights for Rossiter-McLaughlin observations, these data were
-reduced with a non-standard pipeline, and so are systematically offset
-from the remaining HARPS data.  In our final fit, we therefore omitted
-the three \citet{triaud_spin-orbit_2010} nights.  We did nonetheless
-experiment with using the recent \citet{trifonov_public_2020}
-re-reductions, which allowed us to include two calibration RVs taken
-as part of the \citet{triaud_spin-orbit_2010} program. The decision
+reduced with a non-standard pipeline making them ill-suited
+for our study, and we did not include them.\footnote{This problem was
+fixed in principle by \citet{trifonov_public_2020} who performed
+a homogeneous re-reduction of the entire HARPS data archive.
+We found that the decision
 regarding whether to include or omit these points did not
-noticeably affect our results.
+noticeably affect our results.}
 
@@
 
-An initial analysis of the new HIRES observations led to our detection
-of a linear trend in the residuals after fitting out the orbit of
-WASP-4b.  This prompted us to acquire speckle images using Zorro at
-Gemini-South \citep[see][and the instrument
+Once we saw that the new HIRES observations implied a highly significant
+trend in the radial velocity,
+we sought independent evidence for a wide-orbiting companion by performing
+speckle imaging with the Zorro instrument on the 
+Gemini South 8m telescope\citep[see][and the instrument
 web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
@@
 relatively poor seeing (1.2$''$) and also on the night of September
-28-29.  Over each observation, we acquired three sets of 1000
-60$\,$msec exposures.  If a companion is present, the autocorrelation
-functions of these speckle images reveal a characteristic interference
+28-29.  On each night, we acquired three sets of $1000\times 60$\,$msec exposures.
+If a companion is present, the autocorrelation
+functions of these speckle images would reveal a characteristic interference
 pattern. This pattern in then used to determine the properties of the
-detected companion and to produce a reconstructed image.  Using the
-reconstructed speckle images,  contrast curves are produced yielding
-5-sigma detection limits  (see \citealt{howell_speckle_2011}).  No
-companions were detected, and the second night, which had better
-seeing (0.6$''$), also produced the more constraining result.  The
-832$\,$nm limits were the most useful to us given that faint
-companions are redder than the host star.  We therefore opted to use
-the 832$\,$nm September 28-29 contrast limits for the remaining
-analysis.
+detected companion and produce a reconstructed image.  Using the
+reconstructed speckle images, contrast curves are produced to determine the
+5-$\sigma$ detection limits (see \citealt{howell_speckle_2011}).  No
+companions were detected. The data from the second night, which had better
+seeing (0.6$''$), led to the most constraining limit.  The
+832$\,$nm limits were the most useful, given that any faint
+companion would likely be redder than the host star.  Therefore,
+we opted to use
+the 832$\,$nm September 28-29 contrast limits in the analysis described below.
 
@@
   The vertical axis shows the observed transit times minus the
-  calculated times assuming a constant orbital period.  More opaque
+  calculated times, assuming a constant orbital period.  More opaque
   points correspond to more precise data.  The $\pm1\sigma$
@@
   are shown in blue.
-  The binned TESS point (yellow star) is the weighted average of 18
-  TESS transits and is binned for display purposes only.  The models
-  were fitted to all of the individual transit times.
+  The yellow star represents the weighted average of 18 data points
+  obtained with TESS. The TESS data were averaged here for display purposes only;
+  our analysis used the 18 individual transit times.
   \label{fig:times}
@@
 
-We considered two models for the observed transit times.  The first
-model assumes a constant orbital period $P$ on a circular orbit:
+We fitted two simple timing models to the transit-timing data. The first
+model assumes the period $P$ to be constant:
 \begin{align}
@@
 \end{align}
-where $E$ is the integer transit number and $t_0$ is a reference
-epoch.  The second model assumes that the period changes at a steady
+where $E$ is the integer specifying the number within the sequence of orbits
+spanned by the data,
+and $t_0$ is the transit time for the event designated $E=0$.
+The second model assumes that the period changes at a steady
 rate:
@@
 \end{align}
-The free parameters are the reference epoch $t_0$, the period at the
+The free parameters are the reference time $t_0$, the period at the
 reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
 (1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
-$E=0$ is near the weighted average of the observed times.  This helps
-to reduce the covariance between $t_0$ and $P$.  A third possible
-model that we did not consider for reasons that will become apparent
-is a precessing, eccentric orbit \citep[{\it
-e.g.},][]{gimenez_revision_1995,patra_2017}.
+$E=0$ is near the weighted average of the times of observed transits.
+This choice leads to a small covariance between the uncertainties in $t_0$ and $P$.
+%A third possible
+%model that we did not consider for reasons that will become apparent
+%is a precessing, eccentric orbit \citep[{\it
+%e.g.},][]{gimenez_revision_1995,patra_2017}.
 
@@
 \citet{foreman-mackey_emcee_2013} in \texttt{emcee}.  The prior for
-the quadratic model allowed the period derivative to have any sign.
+the quadratic model allowed the period derivative to be either positive
+or negative.
 
 Figure~\ref{fig:times} shows the observed transit times, minus the
-best-fit constant period model.  The best-fitting constant-period
+best-fitting constant-period model.  The best-fitting constant-period
 model has 91 degrees of freedom, $\chi^2 = 276$, and $\chi^2_{\rm
@@
 
-From the reduced $\chi^2$ values, we can surmise that neither model
-entirely describes the transit data---there
-must be some additional source of signal or noise. In
-\citetalias{bouma_wasp4b_2019}, we found that the quadratic model for the
-(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsened
-$\chi^2_{\rm red}$ could reflect underestimated statistical
-uncertainties in any of our transit measurements.  It could also
-reflect systematic errors in the time-systems in which the transit
-measurements were recorded, though we have taken every caution against
-this latter possibility.
+From the high values of $\chi^2_{\rm red}$, we can surmise that neither model
+provides a satisfactory fit to the transit data; there
+must be some additional signal or noise.
+The earlier study by \citetalias{bouma_wasp4b_2019} found that the quadratic model for the
+(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening 
+of $\chi^2_{\rm red}$ probably reflects underestimated measurement
+uncertainties in at least some of the newly reported transit times,
+due to time correlated noise in the photometry. It might also be due
+to errors or misunderstandings of the time systems in which the
+data were recorded, despite our best efforts to interpret the literature.
+In what follows, we opted to enlarge the uncertainties in the best-fitting parameters
+of each model by a factor of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of $\approx$1.73$\times$ for the
+linear model, and $\approx$1.41$\times$ for the quadratic model.
 
-One approach to analyzing such data would be to inflate the transit
-measurement uncertainties, and lower the reduced $\chi^2$.  We do not
-think that such an approach is warranted, because it would not change
-the result that the quadratic model is strongly preferred.  Instead,
-we opt to simply inflate the uncertainties reported for each model by
-a factor of $(\chi^2_{\rm red})^{1/2}$, or $\approx$1.73$\times$ for the
-linear model, and $\approx$1.41$\times$ for the quadratic.
-
-The resulting best-fit period derivative for the quadratic model is 
+The resulting period derivative for the quadratic model is 
 \begin{equation}
@@
 \end{equation}
-This agrees to within $1\sigma$ of the value reported by
+This agrees to within 1-$\sigma$ of the value reported by
 \citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
-ms}\,{\rm yr}^{-1}$).  It is $\approx$2.3$\sigma$ larger than the rate
-of period decrease we previously reported \citepalias[$- 12.6 \pm 1.2~{\rm
-ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of the
-new data from \citeauthor{southworth_transit_2019} and
-\citeauthor{baluev_2019}  The other best-fit transit timing model
+ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate
+of period decrease reported by \citepalias[][; $- 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}$]{bouma_wasp4b_2019},
+presumably because of the new data from \citeauthor{southworth_transit_2019} and
+\citeauthor{baluev_2019}. The other best-fitting model
 parameters are reported in Table~3.

### 2020-02-09 paper/ms.tex

@@
   $-0.0422\pm 0.0028\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
-  The associated Doppler effect would cause
+  The associated Doppler effect should cause
   the apparent period to shrink at a rate of
@@
     Residuals, after subtracting the
-    maximum-likelihood orbit of WASP-4b.  The linear trend
-    inferred from the RV data is shown with a black line.
-    The trend that would be needed to
-    produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
-    1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated in purple.  The four
+    best-fitting model for the variations
+    induced by the planet WASP-4b. The black line is the linear trend
+    inferred from the RV data.
+    The purple line shows the slope that would be needed for the Doppler effect
+    to explain the entire period decrease determined from transit timing.
+    The four
     new RV measurements from this work increase the
-    significance of the linear trend from $\approx$2$\sigma$
-    to 15$\sigma$.
+    significance of the linear trend from $\approx$2-$\sigma$
+    to 15-$\sigma$.
     {\it Bottom:}
@@
 
-Our initial model for the radial velocity data was a single Keplerian
-orbit, plus instrument offsets, jitters, and a long-term linear trend
+Our initial model for the radial velocity data was a circular orbit
+plus instrument offsets, ``jitter'' values (explained below), and a long-term linear trend
 \citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
 on the orbital period and time of inferior conjunction using the
-values from Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed
-WASP-4b's eccentricity to zero
-\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
-The free parameters were the velocity semi-amplitude, the instrument
+values from Table~4 of \citetalias{bouma_wasp4b_2019}.
+We assumed the orbit to be circular based on previous studies
+that placed stringent upper limits on the eccentricity \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
+The free parameters were the orbital velocity semi-amplitude, the instrument
 zero-points, an additive ``white noise'' instrument jitter for each
-instrument, and a linear ($\dot{v_{\rm r}}$) acceleration term.  We
-also considered a model without the linear trend, and found that it
-was disfavored by $\Delta {\rm BIC}=73$.
+instrument,
+% JNW: Do you mean "added in quadrature to the formal uncertainties"? That's the usual approach.
+and a linear ($\dot{v_{\rm r}}$) acceleration term.  (Without the linear
+trend, the model is a much poorer fit with $\Delta {\rm BIC}=73$.)
 
-In the preferred ``planet+linear trend'' model, WASP-4 is accelerating
-towards our line-of-sight at high confidence,
+In the best-fitting model, WASP-4 is accelerating
+along our line-of-sight at a rate
 \begin{equation}
@@
 \end{equation}
-The remaining radial velocity model parameters are listed in Table~4.
-For comparison, before our new measurements, $\dot{\gamma}$ was
+The other model parameters are listed in Table~4.
+Based on earlier data, $\dot{\gamma}$ was
 thought to be about five times smaller, and had marginal statistical
@@
 
-The system's acceleration towards our line-of-sight causes a decrease
-in the apparent orbital period:
+Because of the Doppler effect, any line-of-sight acceleration
+should lead to an change in the observed orbital period:
 \begin{align}
@@
 \end{align}
-The majority of the period decrease seen in transits ($\dot{P}= -
-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) therefore seems to be caused by the
-acceleration of the host star.
+Therefore, most of the period derivative that was detected through transit timing ($\dot{P}= -
+8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) can be accounted for by the Doppler effect.
+Given the evidence for unmodeled noise in both the transit timing data and
+the radial-velocity data, it seems plausible that the Doppler effect can
+account for the entire observed period derivative.
 
-An important consideration is whether the measured RV trend is
-correlated with stellar activity.  We investigate this by analyzing
-WASP-4's emission in the Ca II H \& K lines, as quantified with the
-chromospheric $S$-index \citep{wright_chromospheric_2004}.  We only
-examined the HIRES velocities for this step, since they are the main
-source of signal for our analysis.  First, we subtracted the orbital
-solution from the Keck-HIRES velocities.  Then, following
+An important consideration is whether the measured RV trend is truly
+due to acceleration or whether it is due to stellar activity.
+We investigate this by analyzing
+the Ca II H \& K lines in the WASP-4 spectra, as quantified with the
+chromospheric $S$-index \citep{wright_chromospheric_2004}.  We relied only
+on the HIRES velocities, which were the crucial source of information
+in the radial-velocity analysis.  First, we subtracted the component of 
+the best-fitting model representing the orbital motion induced by the planet.
+Then, following 
 \citet{bryan_statistics_2016,bryan_excess_2019}, we calculated the
@@
 cycle.  The $S$-index values are included in Table~2.  We conclude
-that it is highly unlikely that the linear trend is caused by stellar
+that there is no evidence that the linear trend is caused by stellar
 activity.
@@
     in grayscale, and the region excluded from the speckle imaging is
-    indicated.  The expected mass to semimajor-axis degeneracy is
-    shown with a black line.
+    indicated.  The black line illustrates the expected degeneracy between mass and semimajor-axis.
 	\label{fig:mass_sma}
@@
 
-Given a linear radial velocity trend, we can place lower-limits on the
-mass and semi-major axis of additional bodies in the system.  For a
-quick estimate of the minimum mass required to explain the linear
-trend in WASP-4, we turn to \citet{feng_california_2015}.  As they
-discuss, the scenario that yields the minimum companion mass for a
+Given a linear radial velocity trend, we can place probabilistic constraints on the
+mass and semi-major axis of the additional body that is causing the trend.
+For a quick estimate of the minimum mass required to explain the linear
+trend in WASP-4, we turned to \citet{feng_california_2015}.  As they
+discussed, the scenario that yields the minimum companion mass for a
 system with a linear trend is a companion with $e\approx0.5$ and
@@
 where $\tau$ is the observing baseline.  For WASP-4, this gives
-$M_{\rm min} = 4.9 M_{\rm Jup}$.  Higher masses are allowed for
+$M_{\rm min} = 4.9\,M_{\rm Jup}$.  Higher masses are allowed for
 companions that orbit further from the star: at fixed $\dot{\gamma}$,
@@
 High-resolution images can further limit the available parameter space
-by setting an upper limit on the semimajor axis, and a maximum
-brightness (and thereby mass) of any putative companions.  The
-procedure we use to combine constraints from both radial velocities
-and high resolution imaging has been developed by
+by setting an upper limit on the companion brightness (and the corresponding
+mass) as a function of orbital distance.  The
+procedure we used to combine constraints from both radial velocities
+and high resolution imaging was developed by
 \citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
@@
 
-First, we would like to convert the contrast ratios obtained through
-the Zorro imaging (Figure~\ref{fig:zorro}) to limits on the masses of
-putative companions and their separations from the host star.
-
-To do this, we followed \citet{montet_trends_2014}, and opted to
+First, we converted the contrast curves obtained from
+speckle imaging (Figure~\ref{fig:zorro}) to upper limits on the mass of
+a companion as a function of separation. To do this, we followed \citet{montet_trends_2014}, and opted to
 employ the \citet{baraffe_evolutionary_2003} models for substellar
@@
 \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
-We assumed that the system age was 5 Gyr, so that companions would
+We assumed that the system age was 5 Gyr, at which point the companion would
 have fully contracted.
@@
 blackbody spectra. While this is a simplification, we do not
-readily have access to the planetary and stellar atmosphere models
+have ready access to the planetary and stellar atmosphere models
 needed for the consistent calculation with the \texttt{COND03} and
-\texttt{MESA} models.  We therefore adopted the effective temperatures
+\texttt{MESA} models.  We adopted the effective temperatures
 and bolometric luminosities from the \citet{baraffe_evolutionary_2003}
@@
 from the radial velocities, we mostly followed the procedure of
-\citet{bryan_excess_2019}.
-
-We began by defining a $128\times128$ grid in true planetary mass and
-semimajor axis, with even logarithmic spacing from 1 to 900$\,$$M_{\rm
+\citet{bryan_excess_2019}. We began by defining a $128\times128$ grid in true planetary mass and
+semimajor axis, over a logarithmic grid ranging from 1 to 900$\,$$M_{\rm
 Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the possibility
 that an additional companion in any particular cell could explain the
-observed linear trend.  In each cell, we simulated 512 hypothetical
-companions.
+observed linear trend.
 
-We assigned each companion a mass and semimajor axis from log-uniform
-distributions within the grid cell. We drew the inclination from a
+For each grid cell, we simulated 512 hypothetical
+companions. We assigned each companion a mass and semimajor axis from log-uniform
+distributions within the boundaries of the grid cell. We drew the inclination from a
 uniform distribution in $\cos i$.  For companion masses less than
@@
 preferentially close to zero.  The binary star distribution is
-``top-heavy'', with eccentricities closer to one.  We chose a mass
-cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
-bound observed by \citet{schlaufman_evidence_2018} between giant
-planets and brown dwarfs, though this value is also close to the
+``top-heavy'', with a broad range of eccentricities extending close to unity.
+The choice of $10\,M_{\rm Jup}$ as the dividing line between these two regimes was based on the
+empirical study of \citet{schlaufman_evidence_2018} on the distinction between giant
+planets and brown dwarfs. This value is also close to the
 $13\,M_{\rm Jup}$ deuterium-burning limit \citep[{\it
@@
 
-For each simulated companion, we then drew a sample from the converged
-chains of our initial model of WASP-4b. We subtracted the planet's
-orbital solution, leaving RV points with a linear trend.  Given
+For each simulated companion, we drew a set of model parameters from the converged
+posterior probability samples of our model for the WASP-4 radial velocities.
+We subtracted the orbital component of the model from the observed RVs,
+leaving behind the RV residuals with a linear trend.  Given
 $(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each simulated outer
-companion, and the fixed instrument offsets and jitters from the MCMC
-chains, we then performed a maximum likelihood fit for the time and
+companion, and the choice of instrument offsets and jitters,
+we performed a maximum likelihood fit for the time and
 argument of periastron of the outer simulated companion.  We converted
@@
 \subsection{Implications for WASP-4}
-Previous potential explanations for WASP-4b's decreasing orbital
-period included tidal orbital decay, orbital precession, and
-light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
+
+The previously offered explanations for WASP-4b's decreasing orbital
+period included tidal orbital decay, apsidal precession, and
+the Doppler effect \citep{bouma_wasp4b_2019}.  Our new radial
 velocity measurements strongly indicate that the least exotic
-option---light-travel time effects---is also the most likely.
-Transits show the obital period decreasing by $- 8.64 \pm 1.26~{\rm
-ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
-velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
-ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is $\approx$2$\sigma$
-discrepant, Occam's razor would suggest that most or all of the apparent
+option---the Doppler effect---is the most likely explanation.
+Long-term transit timing data show the orbital period to be decreasing by $- 8.64 \pm 1.26~{\rm
+ms}\,{\rm yr}^{-1}$.  The long-term radial velocity data show a trend that
+should lead to an apparent period derivative of $-5.94 \pm 0.39~{\rm
+ms}\,{\rm yr}^{-1}$ through the Doppler effect.
+Although these two measurements of the period derivative are
+discrepant by about 2-$\sigma$,
+Occam's razor suggests that most or all of the apparent
 decrease of WASP-4b's orbital period is caused by the line-of-sight
@@
 
-The corresponding requirements for the companion causing the
-acceleration are that it is likely either a brown-dwarf or low mass
-star, orbiting between $10$-$100\,{\rm AU}$ from the host star
-(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
-at one time have influenced the orbital evolution of the inner giant.
+Based on the available data, the companion causing the
+acceleration is probably either a brown-dwarf or low mass
+star with an orbital distance of $10$-$100\,{\rm AU}$ 
+(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion may
+have influenced the orbital evolution of the giant planet around WASP-4,
+as well as any other planets.
 The fact that most hot Jupiters have similar massive outer companions
-\citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
-evidence for certain high-eccentricity formation pathways (see
+\citep{knutson_friends_2014,bryan_statistics_2016} is compatible
+with some of the high-eccentricity formation theories for hot Jupiters (see
 \citealt{dawson_johnson_2018}).  Further radial velocity monitoring
@@
 
-\subsection{How many other hot Jupiters are accelerating towards the
-Earth?}
+\subsection{How many other hot Jupiters are accelerating towards us?}
 
@@
 \citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight
-accelerations are correspondingly common in hot Jupiter systems.
+accelerations (both positive and negative)
+should therefore be common in hot Jupiter systems.
 
@@
 \citet{knutson_friends_2014} show a non-zero radial velocity trend.
-Therefore around 1 in 3 hot Jupiters is expected to show period
-changes commensurate with WASP-4 due to acceleration by outer
-companions.  Half of these will be period decreases\footnote{ At a
-population level, one would expect half of the line-of-sight
-accelerations to be positive, and half negative.  Tidal orbital decay
-however is usually negative, and could evenutally manifest as an asymmetry in the
-$\dot{P}$ distribution of hot Jupiters.}, and will be an
-astrophysical false positive in searches for tidal orbital decay.
+Therefore, around 1 in 3 hot Jupiters is expected to show period
+changes comparable to that of WASP-4 due to acceleration by outer
+companions.  The sign of the accelerations should be random,
+with about half of them approaching and half of them receding.
+With a large enough sample of short-period hot Jupiters, one might
+be able to distinguish line-of-sight accelerations from
+tidal orbital decay by seeking evidence for a systematic tendency
+for the observed period to decrease, rather than increase.
 
@@
 msec}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
+{\bf Other citations too?}
 This is roughly commensurate with the level of signal many outer
@@
 
-It is therefore pertinent to ask at what point in time further
-detections of the light-travel time effect will become routine for hot
-Jupiters.  This question is the same as asking: at what rate does the
+At what point in time will further
+detections of the Doppler effect become routine for hot
+Jupiters?  More specifically, at what rate does the
 uncertainty in the quadratic term of Equation~\ref{eq:quadratic_model}
-scale with the observing baseline?  We take a Fisher analysis approach
-to the problem.  First, we rewrite Equation~\ref{eq:quadratic_model}
-as
+scale with the observing baseline?  This can be answered with
+a Fisher analysis of the
+model
 \begin{align}
@@
 -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
-The corresponding light-travel time effect predicts a period decrease
+The corresponding Doppler effect predicts a period decrease
 $\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$.  The
@@
 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
-acceleration of the host star~---~not tidal orbital decay, or apsidal
+acceleration of the host star~---~leaving no evidence
+for tidal orbital decay or apsidal
 precession.  The companion causing the acceleration is most likely a
@@
 which the period derivative can be measured from transits scales with
-the baseline duration to the power of 5/2 (Section~\ref{sec:fisher}), and
-so within a decade many more hot Jupiters should show orbital period
+the baseline duration to the 5/2 power (Section~\ref{sec:fisher}).
+Within a decade, many more hot Jupiters should show orbital period
 changes due to accelerations from their outer companions.  

### 2020-02-09 paper/ms.tex
fix column width

@@
 \submitjournal{AAS journals.}
-\shorttitle{WASP-4 is accelerating towards the Earth}
+\shorttitle{WASP-4 is accelerating towards the Sun}
 
@@
 
+%TODO FIXME
 \begin{abstract}
-  The orbital period of the hot Jupiter WASP-4b appears to be decreasing at
-  a rate of $-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$,
-  based on transit-timing measurements spanning 12 years.
-  Proposed explanations for the period change include
-  tidal orbital decay, apsidal precession, and acceleration of the system
-  along the line of sight.
-  To investigate further, we performed new radial velocity measurements and
-  speckle imaging of WASP-4.
-  The radial-velocity data show that the system is
-  accelerating towards the Sun at a rate of
-  $-0.0422\pm 0.0028\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
-  The associated Doppler effect should cause
-  the apparent period to shrink at a rate of
-  $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$, comparable
-  to the observed rate.
-  Thus, the observed change in the transit period is mostly
-  or entirely produced by the line-of-sight acceleration of the system.
-  This acceleration is probably caused by a wide-orbiting companion
-  of mass 10-200$\,M_{\rm Jup}$ and orbital distance 10-100$\,$AU,
-  based on the magnitude of the radial-velocity trend and the non-detection of any
-  companion in the speckle image.  We expect that 1 out of 6 hot Jupiters
-  will show apparent period changes of the same order of magnitude as WASP-4,
-  based on the hot-Jupiter companion statistics of  
-  \citet{knutson_friends_2014}. Continued radial velocity
-  monitoring of hot Jupiters is therefore essential to distinguish
-  the effects of tidal orbital decay or apsidal precession from line-of-sight acceleration.
+  The orbital period of the hot Jupiter WASP-4b appears to be
+  decreasing at a rate of $-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$,
+  based on transit-timing measurements spanning 12 years.  Proposed
+  explanations for the period change include tidal orbital decay,
+  apsidal precession, and acceleration of the system along the line of
+  sight.  To investigate further, we performed new radial velocity
+  measurements and speckle imaging of WASP-4.  The radial-velocity
+  data show that the system is accelerating towards the Sun at a rate
+  of $-0.0422\pm 0.0028\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.  The
+  associated Doppler effect should cause the apparent period to shrink
+  at a rate of $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$, comparable
+  to the observed rate.  Thus, the observed change in the transit
+  period is mostly or entirely produced by the line-of-sight
+  acceleration of the system.  This acceleration is probably caused by
+  a wide-orbiting companion of mass 10-200$\,M_{\rm Jup}$ and orbital
+  distance 10-100$\,$AU, based on the magnitude of the radial-velocity
+  trend and the non-detection of any companion in the speckle image.
+  We expect that 1 out of 6 hot Jupiters will show apparent period
+  changes of the same order of magnitude as WASP-4, based on the
+  hot-Jupiter companion statistics of  \citet{knutson_friends_2014}.
+  Continued radial velocity monitoring of hot Jupiters is therefore
+  essential to distinguish the effects of tidal orbital decay or
+  apsidal precession from line-of-sight acceleration.
 \end{abstract}
 
-\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
-(1332), Transit timing variation method (1710)}
+\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial
+velocity (1332), Transit timing variation method (1710)}
 
@@
 hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
-It is not clear, though, whether the timescale for tidal orbital decay is
-shorter or longer than the timescale for main-sequence stellar
-evolution.  This answer to this question depends on the uncertain rate at which
-friction inside the star damps the tidal oscillations, which is ultimately
-what causes the orbit to shrink
-(as reviewed by \citealt{Mazeh2008} and
-\citealt{ogilvie_tidal_2014}).
-Population studies of hot Jupiters --- based on ages, rotation rates,
-orbital distances, and Galactic kinematics --- have led to differing
-conclusions, with estimated lifetimes ranging from less than a gigayear
-to much longer than main-sequence lifetimes
-(see, {\it e.g.},
+It is not clear, though, whether the timescale for tidal orbital decay
+is shorter or longer than the timescale for main-sequence stellar
+evolution.  This answer to this question depends on the uncertain rate
+at which friction inside the star damps the tidal oscillations, which
+is ultimately what causes the orbit to shrink (as reviewed by
+\citealt{Mazeh2008} and \citealt{ogilvie_tidal_2014}).  Population
+studies of hot Jupiters --- based on ages, rotation rates, orbital
+distances, and Galactic kinematics --- have led to differing
+conclusions, with estimated lifetimes ranging from less than a
+gigayear to much longer than main-sequence lifetimes (see, {\it e.g.},
 \citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
@@
 
-An empirical resolution might be possible through long-term timing
-of transits and occultations, seeking evidence for changes in the orbital
-period.  For instance, long-term transit timing and
-radial velocity measurements for WASP-12b have revealed a secular
-decrease in the period at a rate of $\approx$30 milliseconds per year, which has
-been interpreted as the effect of tidal orbital decay
+An empirical resolution might be possible through long-term timing of
+transits and occultations, seeking evidence for changes in the orbital
+period.  For instance, long-term transit timing and radial velocity
+measurements for WASP-12b have revealed a secular decrease in the
+period at a rate of $\approx$30 milliseconds per year, which has been
+interpreted as the effect of tidal orbital decay
 \citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
 
-This study draws attention to a confounding factor that, while elementary,
-does not seem to have
-received the attention it deserves.
-The point is that observational programs
-aimed at identifying orbital decay in hot Jupiters through transit
-timing must be accompanied by concurrent long-term radial velocity
-monitoring.  The reason is that an apparent change in period
-can be produced by the Doppler effect associated with acceleration
-of the hot-Jupiter host star along the line of sight, such as the acceleration that
-might be produced by a massive, wide-orbiting companion.
-Massive outer companions to hot Jupiters are common.
-\citet{bryan_statistics_2016} calculated an
+This study draws attention to a confounding factor that, while
+elementary, does not seem to have received the attention it deserves.
+The point is that observational programs aimed at identifying orbital
+decay in hot Jupiters through transit timing must be accompanied by
+concurrent long-term radial velocity monitoring.  The reason is that
+an apparent change in period can be produced by the Doppler effect
+associated with acceleration of the hot-Jupiter host star along the
+line of sight, such as the acceleration that might be produced by a
+massive, wide-orbiting companion.  Massive outer companions to hot
+Jupiters are common.  \citet{bryan_statistics_2016} calculated an
 occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
 Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
-from 5-100$\,$AU.  Therefore, we expect that many hot Jupiters will display
-secular trends in orbital period that are unrelated to tidal orbital decay.
-This possibility can be checked by performing long-term radial-velocity monitoring
-at a level sensitive enough to detect or rule out the relevant amplitude of
-acceleration.
+from 5-100$\,$AU.  Therefore, we expect that many hot Jupiters will
+display secular trends in orbital period that are unrelated to tidal
+orbital decay.  This possibility can be checked by performing
+long-term radial-velocity monitoring at a level sensitive enough to
+detect or rule out the relevant amplitude of acceleration.
 
@@
 orbital period that appears to be decreasing by about 10 milliseconds
-per year.  The period decrease was identified by \citet{bouma_wasp4b_2019}]{bouma_wasp4b_2019},
-who combined data from
-the NASA TESS mission \citep{ricker_transiting_2015} and a decade of ground-based
-transit observations. Soon thereafter, \citet{southworth_transit_2019} reported an additional
-22 transit times and recalculated the period derivative to be $\dot{P} = -9.2 \pm 1.1$
-milliseconds per year.  A separate study by \citet{baluev_2019}
-reported on additional transit times, and pointed out
-that the period decrease was statistically significant only when
-analyzing the data with the highest precision.
+per year.  The period decrease was identified by
+\citet{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, who combined data from
+the NASA TESS mission \citep{ricker_transiting_2015} and a decade of
+ground-based transit observations. Soon thereafter,
+\citet{southworth_transit_2019} reported an additional 22 transit
+times and recalculated the period derivative to be $\dot{P} = -9.2 \pm
+1.1$ milliseconds per year.  A separate study by \citet{baluev_2019}
+reported on additional transit times, and pointed out that the period
+decrease was statistically significant only when analyzing the data
+with the highest precision.
 
@@
 additional radial velocity measurements with the Keck I 10m telescope
-and the High Resolution Echelle
-Spectrometer (HIRES; \citealt{vogt_hires_1994}).
-In doing so, we extended
-the time baseline of HIRES measurements from 3 to 9 years.
-The previously available HIRES data led to the marginal ($\approx$$2\sigma$)
+and the High Resolution Echelle Spectrometer (HIRES;
+\citealt{vogt_hires_1994}).  In doing so, we extended the time
+baseline of HIRES measurements from 3 to 9 years.  The previously
+available HIRES data led to the marginal ($\approx$$2\sigma$)
 detection of a radial-velocity trend \citep{knutson_friends_2014}.
-Our new measurements reveal a
-line-of-sight acceleration of $\dot{\gamma} =
--0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,day^{-1}}$.
-Through the Doppler effect\footnote{While the apparent period change caused by a
-line-of-sight acceleration has been referred to as the ``R{\o}mer
-effect'' \citep{yee_orbit_2020}, a simpler and probably better term is
-the Doppler effect. The R{\o}mer effect refers to the {\it delay} in
-the reception of a signal due to a change in the time required for
-light to traverse the distance between the source and observer. The Doppler
-effect refers to the change in the apparent {\it rate} of a process due
-to changes in the relative motion of the source and observer, such
-as the rate of transits.}, this translates into an expected period decrease
-$-5.9$ milliseconds per year, which is comparable to the period decrease
-that was measured from transit timing.  We undertook high-resolution
-(speckle) imaging to search for evidence of a companion that could
-be responsible for the acceleration of WASP-4.
+Our new measurements reveal a line-of-sight acceleration of
+$\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
+m\,s^{-1}\,day^{-1}}$.  Through the Doppler effect\footnote{While the
+apparent period change caused by a line-of-sight acceleration has been
+referred to as the ``R{\o}mer effect'' \citep{yee_orbit_2020}, a
+simpler and probably better term is the Doppler effect. The R{\o}mer
+effect refers to the {\it delay} in the reception of a signal due to a
+change in the time required for light to traverse the distance between
+the source and observer. The Doppler effect refers to the change in
+the apparent {\it rate} of a process due to changes in the relative
+motion of the source and observer, such as the rate of transits.},
+this translates into an expected period decrease $-5.9$ milliseconds
+per year, which is comparable to the period decrease that was measured
+from transit timing.  We undertook high-resolution (speckle) imaging
+to search for evidence of a companion that could be responsible for
+the acceleration of WASP-4.
 
 Section~\ref{sec:observations} of this paper presents all of the
-available transit data as well as the new radial velocity and
-speckle imaging observations.  Section~\ref{sec:analysis} describes
-our analysis of the data, and our interpretation
-that WASP-4 is being pulled around by a brown dwarf or low-mass star.
-Section~\ref{sec:discussion} places this
-result within the context of orbital decay searches, and points
-out that line-of-sight accelerations will be a relatively
-common type of ``false positive.''
+available transit data as well as the new radial velocity and speckle
+imaging observations.  Section~\ref{sec:analysis} describes our
+analysis of the data, and our interpretation that WASP-4 is being
+pulled around by a brown dwarf or low-mass star.
+Section~\ref{sec:discussion} places this result within the context of
+orbital decay searches, and points out that line-of-sight
+accelerations will be a relatively common type of ``false positive.''
 Section~\ref{sec:conclusions} offers concluding remarks.
@@
 Table~1 lists the transit times we collected for our analysis.  We
-included data from the peer-reviewed literature for which {\it (i)} the
-reported time was based on the data from a single transit (as opposed
-to fitting the data spanning multiple transits and assuming the
-period to be constant), {\it (ii)} the
-central transit time was allowed to be a free parameter, and {\it (iii)} the time
-system was documented clearly, in particular specifying
-whether barycentric or heliocentric corrections had been
-performed and whether leap seconds had been taken into account (TDB or UTC).
+included data from the peer-reviewed literature for which {\it (i)}
+the reported time was based on the data from a single transit (as
+opposed to fitting the data spanning multiple transits and assuming
+the period to be constant), {\it (ii)} the central transit time was
+allowed to be a free parameter, and {\it (iii)} the time system was
+documented clearly, in particular specifying whether barycentric or
+heliocentric corrections had been performed and whether leap seconds
+had been taken into account (TDB or UTC).
 
-% FIXME : verify with Michael Gillon that this is true. (P. Evans
-% done)
+% FIXME : update wording
 Most of these data are identical to the data presented by
@@
 \citet{southworth_transit_2019} are included.  These transits were
-observed from the 3.58m New Technology Telescope and the Danish 1.54m telescope at La Silla, Chile,
-and the South African Astronomical Observatory 1.0m telescope.
+observed from the 3.58m New Technology Telescope and the Danish 1.54m
+telescope at La Silla, Chile, and the South African Astronomical
+Observatory 1.0m telescope.
 % Five of the twenty-two occurred at simultaneous epochs as transits
@@
 \citet{baluev_2019}, based on a homogeneous analysis of archival
-ground-based observations.
-We included twelve of their ``high quality'' transit times from
-the TRAPPIST telescope (six transits), El~Sauce (four transits), and
-\citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
-the original observers that correct barycentric and leap-second
-corrections had been performed (M.~Gillon, P.~Evans, priv{.} comm{.}).
-We omitted the fourteen remaining \citeauthor{baluev_2019}
-ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
-whether leap-second corrections had been performed.
-We did not include in our analysis the four occultation times tabulated by
-\citetalias{bouma_wasp4b_2019}, because of the large timing
-uncertainties and negligible statistical power.
+ground-based observations.  We included twelve of their ``high
+quality'' transit times from the TRAPPIST telescope (six transits),
+El~Sauce (four transits), and \citet{petrucci_no_2013}.  For TRAPPIST
+and El~Sauce, we verified with the original observers that correct
+barycentric and leap-second corrections had been performed (M.~Gillon,
+P.~Evans, priv{.} comm{.}).  We omitted the fourteen remaining
+\citeauthor{baluev_2019} ETD\footnote{\url{http://var2.astro.cz/ETD}}
+times due to ambiguity in whether leap-second corrections had been
+performed.  We did not include in our analysis the four occultation
+times tabulated by \citetalias{bouma_wasp4b_2019}, because of the
+large timing uncertainties and negligible statistical power.
 
@@
 
-We acquired four new radial
-velocity measurements with Keck/HIRES.  Our observations
-were performed using the standard setup and reduction techniques of the
-California Planet Survey \citep{howard_cps_2010}.  Previously, the
-HIRES data-points spanned 2010 to 2013 \citep{knutson_friends_2014}.
-Our new measurements triple the HIRES observing baseline to nine
-years.
+We acquired four new radial velocity measurements with Keck/HIRES.
+Our observations were performed using the standard setup and reduction
+techniques of the California Planet Survey \citep{howard_cps_2010}.
+Previously, the HIRES data-points spanned 2010 to 2013
+\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
+observing baseline to nine years.
 
 The complete set of radial velocity observations is given in Table~2.
-Along with the 2010--2019 HIRES observations are early
-measurements with two different spectrographs.
-We included data from CORALIE ({\bf Spell it out on first use)}
-from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
-adopting the values from the homogeneous analysis of the latter authors.
-We included data from HARPS ({\bf spell it out})
-from \citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
-While \citet{triaud_spin-orbit_2010} also acquired HARPS data over
-three nights for Rossiter-McLaughlin observations, these data were
-reduced with a non-standard pipeline making them ill-suited
-for our study, and we did not include them.\footnote{This problem was
-fixed in principle by \citet{trifonov_public_2020} who performed
-a homogeneous re-reduction of the entire HARPS data archive.
-We found that the decision
-regarding whether to include or omit these points did not
+Along with the 2010--2019 HIRES observations are early measurements
+with two different spectrographs.  We included data from CORALIE ({\bf
+Spell it out on first use)} from \citet{wilson_wasp-4b_2008} and
+\citet{triaud_spin-orbit_2010}, adopting the values from the
+homogeneous analysis of the latter authors.  We included data from
+HARPS ({\bf spell it out}) from \citet{pont_determining_2011} and
+\citet{husnoo_observational_2012}.  While
+\citet{triaud_spin-orbit_2010} also acquired HARPS data over three
+nights for Rossiter-McLaughlin observations, these data were reduced
+with a non-standard pipeline making them ill-suited for our study, and
+we did not include them.\footnote{This problem was fixed in principle
+by \citet{trifonov_public_2020} who performed a homogeneous
+re-reduction of the entire HARPS data archive.  We found that the
+decision regarding whether to include or omit these points did not
 noticeably affect our results.}
@@
 relatively poor seeing (1.2$''$) and also on the night of September
-28-29.  On each night, we acquired three sets of $1000\times 60$\,$msec exposures.
-If a companion is present, the autocorrelation
-functions of these speckle images would reveal a characteristic interference
-pattern. This pattern in then used to determine the properties of the
-detected companion and produce a reconstructed image.  Using the
-reconstructed speckle images, contrast curves are produced to determine the
-5-$\sigma$ detection limits (see \citealt{howell_speckle_2011}).  No
-companions were detected. The data from the second night, which had better
-seeing (0.6$''$), led to the most constraining limit.  The
-832$\,$nm limits were the most useful, given that any faint
-companion would likely be redder than the host star.  Therefore,
-we opted to use
-the 832$\,$nm September 28-29 contrast limits in the analysis described below.
+28-29.  On each night, we acquired three sets of $1000\times
+60$$\,$msec exposures.  If a companion is present, the autocorrelation
+functions of these speckle images would reveal a characteristic
+interference pattern. This pattern in then used to determine the
+properties of the detected companion and produce a reconstructed
+image.  Using the reconstructed speckle images, contrast curves are
+produced to determine the 5-$\sigma$ detection limits (see
+\citealt{howell_speckle_2011}).  No companions were detected. The data
+from the second night, which had better seeing (0.6$''$), led to the
+most constraining limit.  The 832$\,$nm limits were the most useful,
+given that any faint companion would likely be redder than the host
+star.  Therefore, we opted to use the 832$\,$nm September 28-29
+contrast limits in the analysis described below.
 
@@
   points correspond to more precise data.  The $\pm1\sigma$
-  uncertainties of the quadratic ephemeris 
-  are shown in blue.
-  The yellow star represents the weighted average of 18 data points
-  obtained with TESS. The TESS data were averaged here for display purposes only;
-  our analysis used the 18 individual transit times.
+  uncertainties of the quadratic ephemeris are shown in blue.  The
+  yellow star represents the weighted average of 18 data points
+  obtained with TESS. The TESS data were averaged here for display
+  purposes only; our analysis used the 18 individual transit times.
   \label{fig:times}
@@
 
-We fitted two simple timing models to the transit-timing data. The first
-model assumes the period $P$ to be constant:
+We fitted two simple timing models to the transit-timing data. The
+first model assumes the period $P$ to be constant:
 \begin{align}
@@
 \end{align}
-where $E$ is the integer specifying the number within the sequence of orbits
-spanned by the data,
-and $t_0$ is the transit time for the event designated $E=0$.
-The second model assumes that the period changes at a steady
-rate:
+where $E$ is the integer specifying the number within the sequence of
+orbits spanned by the data, and $t_0$ is the transit time for the
+event designated $E=0$.  The second model assumes that the period
+changes at a steady rate:
 \begin{align}
@@
 $E=0$ is near the weighted average of the times of observed transits.
-This choice leads to a small covariance between the uncertainties in $t_0$ and $P$.
+This choice leads to a small covariance between the uncertainties in
+$t_0$ and $P$.
 %A third possible
@@
 
-We fitted each model by assuming a Gaussian likelihood and
-sampling over the posterior probability distributions.  We sampled the
+We fitted each model by assuming a Gaussian likelihood and sampling
+over the posterior probability distributions.  We sampled the
 posterior using the algorithm proposed by
@@
 \citet{foreman-mackey_emcee_2013} in \texttt{emcee}.  The prior for
-the quadratic model allowed the period derivative to be either positive
-or negative.
+the quadratic model allowed the period derivative to be either
+positive or negative.
 
@@
 model has 91 degrees of freedom, $\chi^2 = 276$, and $\chi^2_{\rm
-red}=3.0$.  The best-fitting quadratic model has 90 degrees of freedom,
-$\chi^2 = 183$, and $\chi^2_{\rm red}=2.0$.  The difference in the
-Bayesian information criteria (BIC) between the linear and quadratic
-and models is $\Delta {\rm BIC} = 89$, strongly favoring the
+red}=3.0$.  The best-fitting quadratic model has 90 degrees of
+freedom, $\chi^2 = 183$, and $\chi^2_{\rm red}=2.0$.  The difference
+in the Bayesian information criteria (BIC) between the linear and
+quadratic and models is $\Delta {\rm BIC} = 89$, strongly favoring the
 quadratic model \citep{kass_bayes_1995}.
 
-From the high values of $\chi^2_{\rm red}$, we can surmise that neither model
-provides a satisfactory fit to the transit data; there
-must be some additional signal or noise.
-The earlier study by \citetalias{bouma_wasp4b_2019} found that the quadratic model for the
-(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening 
-of $\chi^2_{\rm red}$ probably reflects underestimated measurement
+From the high values of $\chi^2_{\rm red}$, we can surmise that
+neither model provides a satisfactory fit to the transit data; there
+must be some additional signal or noise.  The earlier study by
+\citetalias{bouma_wasp4b_2019} found that the quadratic model for the
+(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening of
+$\chi^2_{\rm red}$ probably reflects underestimated measurement
 uncertainties in at least some of the newly reported transit times,
 due to time correlated noise in the photometry. It might also be due
-to errors or misunderstandings of the time systems in which the
-data were recorded, despite our best efforts to interpret the literature.
-In what follows, we opted to enlarge the uncertainties in the best-fitting parameters
-of each model by a factor of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of $\approx$1.73$\times$ for the
+to errors or misunderstandings of the time systems in which the data
+were recorded, despite our best efforts to interpret the literature.
+In what follows, we opted to enlarge the uncertainties in the
+best-fitting parameters of each model by a factor of $(\chi^2_{\rm
+red})^{1/2}$.  This was a factor of $\approx$1.73$\times$ for the
 linear model, and $\approx$1.41$\times$ for the quadratic model.
@@
 \citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
-ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate
-of period decrease reported by \citepalias[][; $- 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}$]{bouma_wasp4b_2019},
-presumably because of the new data from \citeauthor{southworth_transit_2019} and
-\citeauthor{baluev_2019}. The other best-fitting model
-parameters are reported in Table~3.
+ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate of
+period decrease reported by \citepalias[][; $- 12.6 \pm 1.2~{\rm
+ms}\,{\rm yr}^{-1}$]{bouma_wasp4b_2019}, presumably because of the new
+data from \citeauthor{southworth_transit_2019} and
+\citeauthor{baluev_2019}. The other best-fitting model parameters are
+reported in Table~3.
 
@@
     {\it Middle:}
-    Residuals, after subtracting the
-    best-fitting model for the variations
-    induced by the planet WASP-4b. The black line is the linear trend
-    inferred from the RV data.
-    The purple line shows the slope that would be needed for the Doppler effect
-    to explain the entire period decrease determined from transit timing.
-    The four
-    new RV measurements from this work increase the
-    significance of the linear trend from $\approx$2-$\sigma$
-    to 15-$\sigma$.
+    Residuals, after subtracting the best-fitting model for the
+    variations induced by the planet WASP-4b. The black line is the
+    linear trend inferred from the RV data.  The purple line shows the
+    slope that would be needed for the Doppler effect to explain the
+    entire period decrease determined from transit timing.  The four
+    new RV measurements from this work increase the significance of
+    the linear trend from $\approx$2-$\sigma$ to 15-$\sigma$.
     {\it Bottom:}
@@
 Our initial model for the radial velocity data was a circular orbit
-plus instrument offsets, ``jitter'' values (explained below), and a long-term linear trend
-\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
-on the orbital period and time of inferior conjunction using the
-values from Table~4 of \citetalias{bouma_wasp4b_2019}.
-We assumed the orbit to be circular based on previous studies
-that placed stringent upper limits on the eccentricity \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
-The free parameters were the orbital velocity semi-amplitude, the instrument
-zero-points, an additive ``white noise'' instrument jitter for each
-instrument,
-% JNW: Do you mean "added in quadrature to the formal uncertainties"? That's the usual approach.
-and a linear ($\dot{v_{\rm r}}$) acceleration term.  (Without the linear
-trend, the model is a much poorer fit with $\Delta {\rm BIC}=73$.)
+plus instrument offsets, ``jitter'' values (explained below), and a
+long-term linear trend \citep[][\texttt{radvel}]{fulton_radvel_2018}.
+We set Gaussian priors on the orbital period and time of inferior
+conjunction using the values from Table~4 of
+\citetalias{bouma_wasp4b_2019}.  We assumed the orbit to be circular
+based on previous studies that placed stringent upper limits on the
+eccentricity
+\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
+The free parameters were the orbital velocity semi-amplitude, the
+instrument zero-points, an additive ``white noise'' instrument jitter
+for each instrument, and a linear ($\dot{v_{\rm r}}$) acceleration
+term.  (Without the linear trend, the model is a much poorer fit with
+$\Delta {\rm BIC}=73$.)
 
-In the best-fitting model, WASP-4 is accelerating
-along our line-of-sight at a rate
+In the best-fitting model, WASP-4 is accelerating along our
+line-of-sight at a rate
 \begin{equation}
@@
 \end{equation}
-The other model parameters are listed in Table~4.
-Based on earlier data, $\dot{\gamma}$ was
-thought to be about five times smaller, and had marginal statistical
-significance \citep{knutson_friends_2014,bouma_wasp4b_2019}.
+The other model parameters are listed in Table~4.  Based on earlier
+data, $\dot{\gamma}$ was thought to be about five times smaller, and
+had marginal statistical significance
+\citep{knutson_friends_2014,bouma_wasp4b_2019}.
 
-Because of the Doppler effect, any line-of-sight acceleration
-should lead to an change in the observed orbital period:
+Because of the Doppler effect, any line-of-sight acceleration should
+lead to an change in the observed orbital period:
 \begin{align}
@@
 \end{align}
-Therefore, most of the period derivative that was detected through transit timing ($\dot{P}= -
-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) can be accounted for by the Doppler effect.
-Given the evidence for unmodeled noise in both the transit timing data and
-the radial-velocity data, it seems plausible that the Doppler effect can
+Therefore, most of the period derivative that was detected through
+transit timing ($\dot{P}= - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$)
+can be accounted for by the Doppler effect.  Given the evidence for
+unmodeled noise in both the transit timing data and the
+radial-velocity data, it seems plausible that the Doppler effect can
 account for the entire observed period derivative.
@@
 An important consideration is whether the measured RV trend is truly
-due to acceleration or whether it is due to stellar activity.
-We investigate this by analyzing
-the Ca II H \& K lines in the WASP-4 spectra, as quantified with the
-chromospheric $S$-index \citep{wright_chromospheric_2004}.  We relied only
-on the HIRES velocities, which were the crucial source of information
-in the radial-velocity analysis.  First, we subtracted the component of 
-the best-fitting model representing the orbital motion induced by the planet.
-Then, following 
+due to acceleration or whether it is due to stellar activity.  We
+investigate this by analyzing the Ca II H \& K lines in the WASP-4
+spectra, as quantified with the chromospheric $S$-index
+\citep{wright_chromospheric_2004}.  We relied only on the HIRES
+velocities, which were the crucial source of information in the
+radial-velocity analysis.  First, we subtracted the component of the
+best-fitting model representing the orbital motion induced by the
+planet.  Then, following
 \citet{bryan_statistics_2016,bryan_excess_2019}, we calculated the
@@
     in grayscale, and the region excluded from the speckle imaging is
-    indicated.  The black line illustrates the expected degeneracy between mass and semimajor-axis.
+    indicated.  The black line illustrates the expected degeneracy
+    between mass and semimajor-axis.
 	\label{fig:mass_sma}
@@
 
-Given a linear radial velocity trend, we can place probabilistic constraints on the
-mass and semi-major axis of the additional body that is causing the trend.
-For a quick estimate of the minimum mass required to explain the linear
-trend in WASP-4, we turned to \citet{feng_california_2015}.  As they
-discussed, the scenario that yields the minimum companion mass for a
-system with a linear trend is a companion with $e\approx0.5$ and
-$\omega=90^\circ$.  Substituting $P\approx 1.25\tau$ and $K \approx
-0.5\tau \dot{\gamma}$ into the mass function \citep[{\it
-e.g.},][]{wright_efficient_2009} yields
+Given a linear radial velocity trend, we can place probabilistic
+constraints on the mass and semi-major axis of the additional body
+that is causing the trend.  For a quick estimate of the minimum mass
+required to explain the linear trend in WASP-4, we turned to
+\citet{feng_california_2015}.  As they discussed, the scenario that
+yields the minimum companion mass for a system with a linear trend is
+a companion with $e\approx0.5$ and $\omega=90^\circ$.  Substituting
+$P\approx 1.25\tau$ and $K \approx 0.5\tau \dot{\gamma}$ into the mass
+function \citep[{\it e.g.},][]{wright_efficient_2009} yields
 \begin{equation}
@@
 High-resolution images can further limit the available parameter space
-by setting an upper limit on the companion brightness (and the corresponding
-mass) as a function of orbital distance.  The
-procedure we used to combine constraints from both radial velocities
-and high resolution imaging was developed by
-\citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
-\citet{montet_trends_2014}, \citet{knutson_friends_2014},
+by setting an upper limit on the companion brightness (and the
+corresponding mass) as a function of orbital distance.  The procedure
+we used to combine constraints from both radial velocities and high
+resolution imaging was developed by \citet{wright_linear_trends_2007},
+\citet{crepp_trends_2012}, \citet{montet_trends_2014},
+\citet{knutson_friends_2014},
 \citet{bryan_statistics_2016,bryan_excess_2019}, and others.
@@
 
-First, we converted the contrast curves obtained from
-speckle imaging (Figure~\ref{fig:zorro}) to upper limits on the mass of
-a companion as a function of separation. To do this, we followed \citet{montet_trends_2014}, and opted to
-employ the \citet{baraffe_evolutionary_2003} models for substellar
-mass objects and the MIST isochrones for stellar mass objects
+First, we converted the contrast curves obtained from speckle imaging
+(Figure~\ref{fig:zorro}) to upper limits on the mass of a companion as
+a function of separation. To do this, we followed
+\citet{montet_trends_2014}, and opted to employ the
+\citet{baraffe_evolutionary_2003} models for substellar mass objects
+and the MIST isochrones for stellar mass objects
 \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
-We assumed that the system age was 5 Gyr, at which point the companion would
-have fully contracted.
+We assumed that the system age was 5 Gyr, at which point the companion
+would have fully contracted.
 
@@
 of synthetic photometry, we further assumed that all sources had
-blackbody spectra. While this is a simplification, we do not
-have ready access to the planetary and stellar atmosphere models
-needed for the consistent calculation with the \texttt{COND03} and
-\texttt{MESA} models.  We adopted the effective temperatures
-and bolometric luminosities from the \citet{baraffe_evolutionary_2003}
-and MIST isochrones.  Using these theoretical quantities and the
+blackbody spectra. While this is a simplification, we do not have
+ready access to the planetary and stellar atmosphere models needed for
+the consistent calculation with the \texttt{COND03} and \texttt{MESA}
+models.  We adopted the effective temperatures and bolometric
+luminosities from the \citet{baraffe_evolutionary_2003} and MIST
+isochrones.  Using these theoretical quantities and the
 empirically-measured Zorro bandpasses, we calculated absolute
@@
 from the radial velocities, we mostly followed the procedure of
-\citet{bryan_excess_2019}. We began by defining a $128\times128$ grid in true planetary mass and
-semimajor axis, over a logarithmic grid ranging from 1 to 900$\,$$M_{\rm
-Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the possibility
-that an additional companion in any particular cell could explain the
-observed linear trend.
+\citet{bryan_excess_2019}. We began by defining a $128\times128$ grid
+in true planetary mass and semimajor axis, over a logarithmic grid
+ranging from 1 to 900$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We
+then considered the possibility that an additional companion in any
+particular cell could explain the observed linear trend.
 
-For each grid cell, we simulated 512 hypothetical
-companions. We assigned each companion a mass and semimajor axis from log-uniform
-distributions within the boundaries of the grid cell. We drew the inclination from a
-uniform distribution in $\cos i$.  For companion masses less than
-$10\,M_{\rm Jup}$, we drew the eccentricity from
+For each grid cell, we simulated 512 hypothetical companions. We
+assigned each companion a mass and semimajor axis from log-uniform
+distributions within the boundaries of the grid cell. We drew the
+inclination from a uniform distribution in $\cos i$.  For companion
+masses less than $10\,M_{\rm Jup}$, we drew the eccentricity from
 \citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
@@
 preferentially close to zero.  The binary star distribution is
-``top-heavy'', with a broad range of eccentricities extending close to unity.
-The choice of $10\,M_{\rm Jup}$ as the dividing line between these two regimes was based on the
-empirical study of \citet{schlaufman_evidence_2018} on the distinction between giant
-planets and brown dwarfs. This value is also close to the
-$13\,M_{\rm Jup}$ deuterium-burning limit \citep[{\it
+``top-heavy'', with a broad range of eccentricities extending close to
+unity.  The choice of $10\,M_{\rm Jup}$ as the dividing line between
+these two regimes was based on the empirical study of
+\citet{schlaufman_evidence_2018} on the distinction between giant
+planets and brown dwarfs. This value is also close to the $13\,M_{\rm
+Jup}$ deuterium-burning limit \citep[{\it
 e.g.},][]{burrows_nongray_1997}.
 
-For each simulated companion, we drew a set of model parameters from the converged
-posterior probability samples of our model for the WASP-4 radial velocities.
-We subtracted the orbital component of the model from the observed RVs,
-leaving behind the RV residuals with a linear trend.  Given
-$(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each simulated outer
-companion, and the choice of instrument offsets and jitters,
-we performed a maximum likelihood fit for the time and
+For each simulated companion, we drew a set of model parameters from
+the converged posterior probability samples of our model for the
+WASP-4 radial velocities.  We subtracted the orbital component of the
+model from the observed RVs, leaving behind the RV residuals with a
+linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each
+simulated outer companion, and the choice of instrument offsets and
+jitters, we performed a maximum likelihood fit for the time and
 argument of periastron of the outer simulated companion.  We converted
@@
 The previously offered explanations for WASP-4b's decreasing orbital
-period included tidal orbital decay, apsidal precession, and
-the Doppler effect \citep{bouma_wasp4b_2019}.  Our new radial
-velocity measurements strongly indicate that the least exotic
-option---the Doppler effect---is the most likely explanation.
-Long-term transit timing data show the orbital period to be decreasing by $- 8.64 \pm 1.26~{\rm
-ms}\,{\rm yr}^{-1}$.  The long-term radial velocity data show a trend that
-should lead to an apparent period derivative of $-5.94 \pm 0.39~{\rm
-ms}\,{\rm yr}^{-1}$ through the Doppler effect.
+period included tidal orbital decay, apsidal precession, and the
+Doppler effect \citep{bouma_wasp4b_2019}.  Our new radial velocity
+measurements strongly indicate that the least exotic option---the
+Doppler effect---is the most likely explanation.  Long-term transit
+timing data show the orbital period to be decreasing by $- 8.64 \pm
+1.26~{\rm ms}\,{\rm yr}^{-1}$.  The long-term radial velocity data
+show a trend that should lead to an apparent period derivative of
+$-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$ through the Doppler effect.
 Although these two measurements of the period derivative are
-discrepant by about 2-$\sigma$,
-Occam's razor suggests that most or all of the apparent
-decrease of WASP-4b's orbital period is caused by the line-of-sight
-acceleration.
+discrepant by about 2-$\sigma$, Occam's razor suggests that most or
+all of the apparent decrease of WASP-4b's orbital period is caused by
+the line-of-sight acceleration.
 
-Based on the available data, the companion causing the
-acceleration is probably either a brown-dwarf or low mass
-star with an orbital distance of $10$-$100\,{\rm AU}$ 
-(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion may
-have influenced the orbital evolution of the giant planet around WASP-4,
-as well as any other planets.
-The fact that most hot Jupiters have similar massive outer companions
-\citep{knutson_friends_2014,bryan_statistics_2016} is compatible
-with some of the high-eccentricity formation theories for hot Jupiters (see
+Based on the available data, the companion causing the acceleration is
+probably either a brown-dwarf or low mass star with an orbital
+distance of $10$-$100\,{\rm AU}$ (Figure~\ref{fig:mass_sma}).  Given
+such a mass, this companion may have influenced the orbital evolution
+of the giant planet around WASP-4, as well as any other planets.  The
+fact that most hot Jupiters have similar massive outer companions
+\citep{knutson_friends_2014,bryan_statistics_2016} is compatible with
+some of the high-eccentricity formation theories for hot Jupiters (see
 \citealt{dawson_johnson_2018}).  Further radial velocity monitoring
@@
   \citet{knutson_friends_2014} have shown long-term radial velocity
-  trends.  HAT-P-11 is shown, though its signal may be caused
-  by stellar activity.  Three hot Jupiters are not shown because
-  their radial velocity curves are better described as quadratic
-  trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered
-  in the $y$ dimension by the absolute value of d$P$/d$t$.
+  trends.  HAT-P-11 is shown, though its signal may be caused by
+  stellar activity.  Three hot Jupiters are not shown because their
+  radial velocity curves are better described as quadratic trends in
+  time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered in the $y$
+  dimension by the absolute value of d$P$/d$t$.
 	\label{fig:pdot_pop}
@@
 \citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight
-accelerations (both positive and negative)
-should therefore be common in hot Jupiter systems.
+accelerations (both positive and negative) should therefore be common
+in hot Jupiter systems.
 
@@
 changes comparable to that of WASP-4 due to acceleration by outer
-companions.  The sign of the accelerations should be random,
-with about half of them approaching and half of them receding.
-With a large enough sample of short-period hot Jupiters, one might
-be able to distinguish line-of-sight accelerations from
-tidal orbital decay by seeking evidence for a systematic tendency
-for the observed period to decrease, rather than increase.
+companions.  The sign of the accelerations should be random, with
+about half of them approaching and half of them receding.  With a
+large enough sample of short-period hot Jupiters, one might be able to
+distinguish line-of-sight accelerations from tidal orbital decay by
+seeking evidence for a systematic tendency for the observed period to
+decrease, rather than increase.
 
@@
 
-For hot Jupiters that have been monitored over baselines
-exceeding 10 years, secular changes in their orbital periods are
-currently being constrained to a precision of $\lesssim$$10~{\rm
-msec}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
-{\bf Other citations too?}
-This is roughly commensurate with the level of signal many outer
+For hot Jupiters that have been monitored over baselines exceeding 10
+years, secular changes in their orbital periods are currently being
+constrained to a precision of $\lesssim$$10~{\rm msec}\,{\rm yr}^{-1}$
+({\it e.g.}, K.~Patra et al{.} 2020, submitted).  {\bf Other citations
+too?} This is roughly commensurate with the level of signal many outer
 companions are expected to induce (Figure~\ref{fig:pdot_pop}).
 
-At what point in time will further
-detections of the Doppler effect become routine for hot
-Jupiters?  More specifically, at what rate does the
-uncertainty in the quadratic term of Equation~\ref{eq:quadratic_model}
-scale with the observing baseline?  This can be answered with
-a Fisher analysis of the
-model
+At what point in time will further detections of the Doppler effect
+become routine for hot Jupiters?  More specifically, at what rate does
+the uncertainty in the quadratic term of
+Equation~\ref{eq:quadratic_model} scale with the observing baseline?
+This can be answered with a Fisher analysis of the model
 \begin{align}
@@
 observations continue from ground and space-based observatories,
-period derivatives will be measured with precision exceeding
-$1~{\rm msec}\,{\rm yr}^{-1}$ within the coming decade.
+period derivatives will be measured with precision exceeding $1~{\rm
+msec}\,{\rm yr}^{-1}$ within the coming decade.
 
@@
 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
-acceleration of the host star~---~leaving no evidence
-for tidal orbital decay or apsidal
-precession.  The companion causing the acceleration is most likely a
-brown dwarf or low-mass star with semi-major axis between
-10-100$\,$AU.
+acceleration of the host star~---~leaving no evidence for tidal
+orbital decay or apsidal precession.  The companion causing the
+acceleration is most likely a brown dwarf or low-mass star with
+semi-major axis between 10-100$\,$AU.
 
@@
 Within a decade, many more hot Jupiters should show orbital period
-changes due to accelerations from their outer companions.  
-To distinguish this effect from tidal decay, further long-term radial
+changes due to accelerations from their outer companions.  To
+distinguish this effect from tidal decay, further long-term radial
 velocity measurements of hot Jupiters are strongly encouraged.

### 2020-02-09 paper/ms.tex
[WIP] tweaks to JNW merge

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: make table 4 look beter
 
@@
 
-%TODO FIXME
+%TODO FIXME: some sentences
 \begin{abstract}
@@
 The orbits of most hot Jupiters are formally unstable to tidal decay
-\citep{counselman_outcomes_1973,
-hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
+\citep{counselman_outcomes_1973,hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
 It is not clear, though, whether the timescale for tidal orbital decay
@@
 orbital period that appears to be decreasing by about 10 milliseconds
-per year.  The period decrease was identified by
-\citet{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, who combined data from
-the NASA TESS mission \citep{ricker_transiting_2015} and a decade of
-ground-based transit observations. Soon thereafter,
+per year.  The period decrease was identified by \citep[][hereafter
+\citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, who combined data
+from the NASA TESS mission \citep{ricker_transiting_2015} and a decade
+of ground-based transit observations. Soon thereafter,
 \citet{southworth_transit_2019} reported an additional 22 transit
@@
 motion of the source and observer, such as the rate of transits.},
-this translates into an expected period decrease $-5.9$ milliseconds
-per year, which is comparable to the period decrease that was measured
-from transit timing.  We undertook high-resolution (speckle) imaging
-to search for evidence of a companion that could be responsible for
-the acceleration of WASP-4.
+this translates into an expected period decrease of $-5.9$
+milliseconds per year, which is comparable to the period decrease that
+was measured from transit timing.  We undertook high-resolution
+(speckle) imaging to search for evidence of a companion that could be
+responsible for the acceleration of WASP-4.
 
@@
 
-% FIXME : update wording
 Most of these data are identical to the data presented by
@@
 telescope at La Silla, Chile, and the South African Astronomical
-Observatory 1.0m telescope.
-% Five of the twenty-two occurred at simultaneous epochs as transits
-% previously reported by \citet{hoyer_tramos_2013}.
-Additional timing measurements were also reported recently by
-\citet{baluev_2019}, based on a homogeneous analysis of archival
-ground-based observations.  We included twelve of their ``high
-quality'' transit times from the TRAPPIST telescope (six transits),
-El~Sauce (four transits), and \citet{petrucci_no_2013}.  For TRAPPIST
-and El~Sauce, we verified with the original observers that correct
-barycentric and leap-second corrections had been performed (M.~Gillon,
-P.~Evans, priv{.} comm{.}).  We omitted the fourteen remaining
-\citeauthor{baluev_2019} ETD\footnote{\url{http://var2.astro.cz/ETD}}
-times due to ambiguity in whether leap-second corrections had been
-performed.  We did not include in our analysis the four occultation
-times tabulated by \citetalias{bouma_wasp4b_2019}, because of the
-large timing uncertainties and negligible statistical power.
+Observatory 1.0m telescope.  Additional timing measurements were also
+reported recently by \citet{baluev_2019}, based on a homogeneous
+analysis of archival ground-based observations.  We included twelve of
+their ``high quality'' transit times from the TRAPPIST telescope (six
+transits), El~Sauce (four transits), and \citet{petrucci_no_2013}.
+For TRAPPIST and El~Sauce, we verified with the original observers
+that the timestamps were amenable to the appropriate barycentric and
+leap second corrections (M.~Gillon, P.~Evans, priv{.} comm{.}).  We
+omitted the fourteen remaining \citeauthor{baluev_2019}
+ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
+whether leap-second corrections had been performed.  We did not
+include in our analysis the four occultation times tabulated by
+\citetalias{bouma_wasp4b_2019}, because of the large timing
+uncertainties and negligible statistical power.
 
@@
 Along with the 2010--2019 HIRES observations are early measurements
-with two different spectrographs.  We included data from CORALIE ({\bf
-Spell it out on first use)} from \citet{wilson_wasp-4b_2008} and
-\citet{triaud_spin-orbit_2010}, adopting the values from the
-homogeneous analysis of the latter authors.  We included data from
-HARPS ({\bf spell it out}) from \citet{pont_determining_2011} and
+with two different spectrographs.  \citet{wilson_wasp-4b_2008} and
+\citet{triaud_spin-orbit_2010} observed WASP-4 with the Swiss 1.2m
+Euler Telescope and CORALIE Spectrograph; we adopted the radial
+velocity values from the homogeneous analysis of the latter authors.
+We also included data from the High Accuracy Radial Velocity Planet
+Searcher (HARPS), reported by \citet{pont_determining_2011} and
 \citet{husnoo_observational_2012}.  While
@@
 
-Once we saw that the new HIRES observations implied a highly significant
-trend in the radial velocity,
-we sought independent evidence for a wide-orbiting companion by performing
-speckle imaging with the Zorro instrument on the 
-Gemini South 8m telescope\citep[see][and the instrument
+Once we saw that the new HIRES observations implied a highly
+significant trend in the radial velocity, we sought independent
+evidence for a wide-orbiting companion by performing speckle imaging
+with the Zorro instrument on the Gemini South 8m
+telescope\citep[see][and the instrument
 web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
@@
 functions of these speckle images would reveal a characteristic
-interference pattern. This pattern in then used to determine the
+interference pattern. This pattern is then used to determine the
 properties of the detected companion and produce a reconstructed
@@
 The free parameters are the reference time $t_0$, the period at the
-reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
+reference time $P$, and the period derivative, ${\rm d}P/{\rm d}t =
 (1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
@@
 $t_0$ and $P$.
-%A third possible
-%model that we did not consider for reasons that will become apparent
-%is a precessing, eccentric orbit \citep[{\it
-%e.g.},][]{gimenez_revision_1995,patra_2017}.
 
@@
 The free parameters were the orbital velocity semi-amplitude, the
-instrument zero-points, an additive ``white noise'' instrument jitter
-for each instrument, and a linear ($\dot{v_{\rm r}}$) acceleration
-term.  (Without the linear trend, the model is a much poorer fit with
-$\Delta {\rm BIC}=73$.)
+instrument zero-points, a ``white noise'' instrument jitter for each
+instrument added in quadrature to its uncertainties, and a linear
+($\dot{v_{\rm r}}$) acceleration term.  (Without the linear trend, the
+model is a much poorer fit with $\Delta {\rm BIC}=73$.)
 
@@
 Because of the Doppler effect, any line-of-sight acceleration should
-lead to an change in the observed orbital period:
+lead to a change in the observed orbital period:
 \begin{align}
@@
 ``top-heavy'', with a broad range of eccentricities extending close to
-unity.  The choice of $10\,M_{\rm Jup}$ as the dividing line between
-these two regimes was based on the empirical study of
-\citet{schlaufman_evidence_2018} on the distinction between giant
-planets and brown dwarfs. This value is also close to the $13\,M_{\rm
-Jup}$ deuterium-burning limit \citep[{\it
-e.g.},][]{burrows_nongray_1997}.
+unity \citep{moe_mind_2017,price-whelan_close_2020}.  The choice of
+$10\,M_{\rm Jup}$ as the dividing line between these two regimes was
+based on the empirical study of \citet{schlaufman_evidence_2018} on
+the distinction between giant planets and brown dwarfs. This value is
+also close to the $13\,M_{\rm Jup}$ deuterium-burning limit
+\citep[{\it e.g.},][]{burrows_nongray_1997}.
 

### 2020-02-09 paper/ms.tex
paper improvements

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: make table 4 look beter
+% TODO: kishore reference
+% TODO: rachel comments
+% TODO: steve comments
 
@@
 \submitjournal{AAS journals.}
-\shorttitle{WASP-4 is accelerating towards the Sun}
+\shorttitle{WASP-4 is accelerating toward the Earth}
 
@@
 
-\title{WASP-4 is Accelerating Towards the Sun}
+\title{WASP-4 is Accelerating Toward the Earth}
 % ", Explaining its Decreasing Orbital Period"
@@
 %
-\author{R. A. Matson}
-\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\author[0000-0001-7233-7508]{R. A. Matson}
+\affiliation{U.S. Naval Observatory, Washington, DC 20392, USA}
 %
 
-%TODO FIXME: some sentences
 \begin{abstract}
@@
   distance 10-100$\,$AU, based on the magnitude of the radial-velocity
-  trend and the non-detection of any companion in the speckle image.
-  We expect that 1 out of 6 hot Jupiters will show apparent period
-  changes of the same order of magnitude as WASP-4, based on the
+  trend and the non-detection of any companion in the speckle images.
+  We expect that the orbital periods of 1 out of 6 hot Jupiters will
+  appear to shrink at a rate similar to WASP-4b, based on the
   hot-Jupiter companion statistics of \citet{knutson_friends_2014}.
@@
 interpreted as the effect of tidal orbital decay
-\citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.
+\citep{maciejewski_departure_2016,patra_2017,maciejewski_planet-star_2018,yee_orbit_2020}.
 
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.7\textwidth]{f1.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
+		The vertical axis shows the observed transit times minus the
+		calculated times, assuming a constant orbital period.  More opaque
+		points correspond to more precise data.  The $\pm1\sigma$
+		uncertainties of the quadratic ephemeris are shown in blue.  The
+		yellow star represents the weighted average of 18 data points
+		obtained with TESS. The TESS data were averaged here for display
+		purposes only; our analysis used the 18 individual transit times.
+		\label{fig:times}
+	}
+\end{figure*}
+
 \subsection{Transits}
@@
 leap second corrections (M.~Gillon, P.~Evans, priv{.} comm{.}).  We
-omitted the fourteen remaining \citeauthor{baluev_2019}
-ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
+omitted the fourteen remaining 
+ETD\footnote{\url{http://var2.astro.cz/ETD}} times from \citeauthor{baluev_2019} due to ambiguity in
 whether leap-second corrections had been performed.  We did not
@@
 with the Zorro instrument on the Gemini South 8m
-telescope\citep[see][and the instrument
+telescope~\citep[see][and the instrument
 web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
@@
 
-\subsection{Transits}
-\label{sec:transit_analysis}
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.7\textwidth]{f2.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Radial velocities of WASP-4.}
+		{\it Top:} RV measurements, with best-fit instrument offsets
+		added. 
+		{\it Middle:}
+		Residuals, after subtracting the best-fitting model for the
+		variations induced by the planet WASP-4b. The black line is the
+		linear trend inferred from the RV data.  The purple line shows the
+		slope that would be needed for the Doppler effect to explain the
+		entire period decrease determined from transit timing.  The four
+		new RV measurements from this work increase the significance of
+		the linear trend from $\approx$2-$\sigma$ to 15-$\sigma$.
+		{\it Bottom:}
+		Phased orbit of WASP-4b.
+		\label{fig:rvs}
+		\vspace{-0.0cm}
+	}
+\end{figure*}
+
+\begin{figure}[!t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.47\textwidth]{f3.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Zorro contrast limits derived from point-source
+			injection-recovery experiments.} Sources below the curve would
+		have been detected.  The inset shows the speckle image
+		reconstructed from 1000 60 millisecond frames in an 832$\,$nm
+		bandpass, and acquired on September 28, 2019.  The image scale
+		is $2.46''\times2.46''$.
+	}
+	\label{fig:zorro}
+\end{figure}
 
@@
 		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f1.pdf}
+		\includegraphics[width=0.8\textwidth]{f4.pdf}
 	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-  The vertical axis shows the observed transit times minus the
-  calculated times, assuming a constant orbital period.  More opaque
-  points correspond to more precise data.  The $\pm1\sigma$
-  uncertainties of the quadratic ephemeris are shown in blue.  The
-  yellow star represents the weighted average of 18 data points
-  obtained with TESS. The TESS data were averaged here for display
-  purposes only; our analysis used the 18 individual transit times.
-  \label{fig:times}
+	\vspace{-0.8cm}
+	\caption{
+		{\bf Masses and semi-major axes of companions that meet
+			requirements of both the radial velocities and the speckle
+			imaging.} The likelihood inferred from radial velocities is shown
+		in grayscale, and the region excluded from the speckle imaging is
+		indicated.  The black line illustrates the expected degeneracy
+		between mass and semimajor-axis.
+		\label{fig:mass_sma}
+		\vspace{-0cm}
 	}
 \end{figure*}
+
+\subsection{Transits}
+\label{sec:transit_analysis}
 
@@
 ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate of
-period decrease reported by \citepalias[][; $- 12.6 \pm 1.2~{\rm
-ms}\,{\rm yr}^{-1}$]{bouma_wasp4b_2019}, presumably because of the new
+period decrease reported by \citetalias{bouma_wasp4b_2019}
+($- 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}$), presumably because of the new
 data from \citeauthor{southworth_transit_2019} and
@@
 \subsection{Radial velocities}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f2.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Radial velocities of WASP-4.}
-    {\it Top:} RV measurements, with best-fit instrument offsets
-    added. 
-    {\it Middle:}
-    Residuals, after subtracting the best-fitting model for the
-    variations induced by the planet WASP-4b. The black line is the
-    linear trend inferred from the RV data.  The purple line shows the
-    slope that would be needed for the Doppler effect to explain the
-    entire period decrease determined from transit timing.  The four
-    new RV measurements from this work increase the significance of
-    the linear trend from $\approx$2-$\sigma$ to 15-$\sigma$.
-    {\it Bottom:}
-    Phased orbit of WASP-4b.
-	\label{fig:rvs}
-  \vspace{-0.0cm}
-	}
-\end{figure*}
 
@@
 \subsection{Constraints on companion masses and semi-major axes}
-
-\begin{figure}[!t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f3.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-    \caption{
-      {\bf Zorro contrast limits derived from point-source
-      injection-recovery experiments.} Sources below the curve would
-      have been detected.  The inset shows the speckle image
-      reconstructed from 1000 60 millisecond frames in an 832$\,$nm
-      bandpass, and acquired on September 28, 2019.  The image scale
-      is $2.46''\times2.46''$.
-    }
-    \label{fig:zorro}
-\end{figure}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.8cm}
-	\caption{
-    {\bf Masses and semi-major axes of companions that meet
-    requirements of both the radial velocities and the speckle
-    imaging.} The likelihood inferred from radial velocities is shown
-    in grayscale, and the region excluded from the speckle imaging is
-    indicated.  The black line illustrates the expected degeneracy
-    between mass and semimajor-axis.
-	\label{fig:mass_sma}
-  \vspace{-0cm}
-	}
-\end{figure*}
 
@@
 such a mass, this companion may have influenced the orbital evolution
-of the giant planet around WASP-4, as well as any other planets.  The
+of the hot Jupiter orbiting WASP-4, as well as any other planets.  The
 fact that most hot Jupiters have similar massive outer companions
@@
 constrained to a precision of $\lesssim$$10~{\rm msec}\,{\rm yr}^{-1}$
-({\it e.g.}, K.~Patra et al{.} 2020, submitted).  {\bf Other citations
-too?} This is roughly commensurate with the level of signal many outer
+\citep{wilkins_searching_2017,maciejewski_planet-star_2018,baluev_2019,petrucci_discarding_2020,patra_2020}.
+This is roughly commensurate with the level of signal many outer
 companions are expected to induce (Figure~\ref{fig:pdot_pop}).
@@
 acceleration of the host star~---~leaving no evidence for tidal
-orbital decay or apsidal precession.  The companion causing the
+orbital decay or apsidal precession.  
+A probabilistic analysis of the speckle imaging limits and the
+radial velocity trend showed that the companion causing the
 acceleration is most likely a brown dwarf or low-mass star with

### 2020-02-09 paper/ms.tex
add facilities section

@@
 %
+% (Some of) The data presented herein were obtained at the W. M. Keck
+% Observatory, which is operated as a scientific partnership among the
+% California Institute of Technology, the University of California and
+% the National Aeronautics and Space Administration. The Observatory was
+% made possible by the generous financial support of the W. M. Keck
+% Foundation.
+% The authors wish to recognize and acknowledge the very significant
+% cultural role and reverence that the summit of Maunakea has always had
+% within the indigenous Hawaiian community.  We are most fortunate to
+% have the opportunity to conduct observations from this mountain.
+%
 % \newline
@@
   \texttt{scipy} \citep{jones_scipy_2001}.
+}
+
+\facilities{
+	{\it Astrometry}:
+	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
+	{\it Imaging}:
+	Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
+	{\it Spectroscopy}:
+	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
+	Euler1.2m~(CORALIE),
+	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
+	{\it Photometry}:
+	CTIO:1.0m (Y4KCam),
+	Danish 1.54m Telescope,
+	El Sauce:0.356m,
+	Elizabeth 1.0m at SAAO,
+	Euler1.2m (EulerCam),
+	Magellan:Baade (MagIC),
+	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
+	NTT,
+	SOAR (SOI),
+	TESS \citep{ricker_transiting_2015},
+	TRAPPIST \citep{jehin_trappist_2011},
+	VLT:Antu (FORS2).
 }

### 2020-02-09 paper/ms.tex
minor

@@
 their ``high quality'' transit times from the TRAPPIST telescope (six
-transits), El~Sauce (four transits), and \citet{petrucci_no_2013}.
-For TRAPPIST and El~Sauce, we verified with the original observers
-that the timestamps were amenable to the appropriate barycentric and
-leap second corrections (M.~Gillon, P.~Evans, priv{.} comm{.}).  We
-omitted the fourteen remaining 
-ETD\footnote{\url{http://var2.astro.cz/ETD}} times from \citeauthor{baluev_2019} due to ambiguity in
-whether leap-second corrections had been performed.  We did not
-include in our analysis the four occultation times tabulated by
+transits), the El~Sauce 36$\,$cm (four transits), and
+\citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
+the original observers that the timestamps were amenable to the
+appropriate barycentric and leap second corrections (M.~Gillon,
+P.~Evans, priv{.} comm{.}).  We omitted the fourteen remaining
+ETD\footnote{\url{http://var2.astro.cz/ETD}} times from
+\citeauthor{baluev_2019} due to ambiguity in whether leap-second
+corrections had been performed.  We did not include in our analysis
+the four occultation times tabulated by
 \citetalias{bouma_wasp4b_2019}, because of the large timing

### 2020-02-23 paper/ms.tex
minor
text update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TODO: kishore reference
-% TODO: rachel comments
-% TODO: steve comments
 
@@
 Jupiters are common.  \citet{bryan_statistics_2016} calculated an
-occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
-Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
-from 5-100$\,$AU.  Therefore, we expect that many hot Jupiters will
-display secular trends in orbital period that are unrelated to tidal
-orbital decay.  This possibility can be checked by performing
-long-term radial-velocity monitoring at a level sensitive enough to
-detect or rule out the relevant amplitude of acceleration.
+occurrence rate of $70\pm8\%$ for outer companions to hot Jupiters
+with masses from 1-13$\,$$M_{\rm Jup}$ and semi-major axes from
+1-20$\,$AU.  Therefore, we expect that many hot Jupiters will display
+secular trends in orbital period that are unrelated to tidal orbital
+decay.  This possibility can be checked by performing long-term
+radial-velocity monitoring at a level sensitive enough to detect or
+rule out the relevant amplitude of acceleration.
 
@@
 
-Based on the available data, the companion causing the acceleration is
-probably either a brown-dwarf or low mass star with an orbital
-distance of $10$-$100\,{\rm AU}$ (Figure~\ref{fig:mass_sma}).  Given
-such a mass, this companion may have influenced the orbital evolution
-of the hot Jupiter orbiting WASP-4, as well as any other planets.  The
-fact that most hot Jupiters have similar massive outer companions
+Based on the data, the companion causing the acceleration is probably
+either a brown-dwarf or low mass star with an orbital distance of
+$10$-$100\,{\rm AU}$ (Figure~\ref{fig:mass_sma}).  Given such a mass,
+this companion may have influenced the orbital evolution of the hot
+Jupiter orbiting WASP-4, as well as any other planets in the system.
+The fact that most hot Jupiters have similar massive outer companions
 \citep{knutson_friends_2014,bryan_statistics_2016} is compatible with
@@
 should eventually reveal the orbital parameters and minimum mass of
-WASP-4's companion.
+WASP-4's massive outer companion.
+
+Based on a priori expectations, if the outer companion lives within
+100$\,$AU, then it seems more likely to be a brown-dwarf than a
+low-mass star.  The reason is that stellar companions within 100$\,$AU
+seem to be rare in systems with transiting planets, relative to field
+stars.  This issue has been reviewed by \citet{moe_impact_2019}, who
+synthesized work by \citet{wang_influence_2014,wang_influence_2015},
+\citet{kraus_impact_2016}, \citet{ngo_2015_fohj2}, \citet{ngo_friends_2016},
+\citet{matson_stellar_2018}, \citet{ziegler_soar_2020}, and others.
+A handful of systems with inner hot Jupiters and outer brown dwarfs within
+100$\,$AU are known, for instance CoRoT-20, HATS-59, WASP-53,
+and WASP-81
+\citep{triaud_peculiar_2017,rey_brown_2018,sarkis_hats-59bc_2018}.  In
+contrast, from surveys by
+\citet{knutson_friends_2014}, \citet{ngo_2015_fohj2} and
+\citet{mugrauer_search_2019} we could find only one example of a
+system with an inner hot Jupiter and an outer low-mass star within
+100$\,$AU: HAT-P-10 \citep{bakos_2009_hatp10}.  Overall, this body of
+literature indicates that the presence of a stellar-mass companion
+within 100$\,$AU of the host star could hinder the formation of
+planetary systems.  If true, then one would expect that the outer
+companion in WASP-4 would have the lowest mass allowed by the data.
 
@@
   \citet{knutson_friends_2014} have shown long-term radial velocity
-  trends.  HAT-P-11 is shown, though its signal may be caused by
-  stellar activity.  Three hot Jupiters are not shown because their
-  radial velocity curves are better described as quadratic trends in
-  time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered in the $y$
-  dimension by the absolute value of d$P$/d$t$.
+  trends.  HAT-P-11 is shown, though its signal is somewhat correlated
+  with stellar activity \citep{yee_hat-p-11_2018}.  Three hot Jupiters
+  are not shown because their radial velocity curves are better
+  described as quadratic trends in time: HAT-P-17, WASP-8, and
+  WASP-34.  Objects are ordered in the $y$ dimension by the absolute
+  value of d$P$/d$t$.
 	\label{fig:pdot_pop}

### 2020-02-28 paper/ms.tex
minor
update word

@@
 %
-\author{S. B. Howell}
+\author[0000-0002-2532-2853]{S. B. Howell}
 \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
@@
   trend and the non-detection of any companion in the speckle images.
-  We expect that the orbital periods of 1 out of 6 hot Jupiters will
-  appear to shrink at a rate similar to WASP-4b, based on the
-  hot-Jupiter companion statistics of \citet{knutson_friends_2014}.
+  We expect that the orbital periods of 1 out of 3 hot Jupiters will
+  change at rates similar to WASP-4b, based on the hot-Jupiter
+  companion statistics of \citet{knutson_friends_2014}.
   Continued radial velocity monitoring of hot Jupiters is therefore
@@
 synthesized work by \citet{wang_influence_2014,wang_influence_2015},
-\citet{kraus_impact_2016}, \citet{ngo_2015_fohj2}, \citet{ngo_friends_2016},
+\citet{ngo_2015_fohj2}, \citet{ngo_friends_2016}, \citet{kraus_impact_2016},
 \citet{matson_stellar_2018}, \citet{ziegler_soar_2020}, and others.

### 2020-03-16 second_paper_review/response_20200315.txt
[WIP] second referee report

Dear Dr. Jackson,

We thank you for organizing the review of our paper, and thank the referee for
their helpful comments.  We have considered the referee's remarks carefully and
revised our manuscript accordingly.

Below, we have reproduced the referee's report, alongside with our
point-by-point responses. A list of changes is appended to the manuscript using
the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

********************************************************************************
REFEREE COMMENT
----------
Section 2.1, Paragraph 2.

It would be helpful if the authors briefly describe the exact meaning of "high
quality" in the context of the transit times used in the analysis.

RESPONSE
----------
> Thank you for this prompt. We have removed the phrase "high quality", as its
> precise definition is not of particular relevance for this study. It is
> defined in Section 3.4 of Baluev+ 2019, through an empirical cutoff that
> Baluev+ made based on the sampling cadence and the light-curve residual RMS.

********************************************************************************
REFEREE COMMENT
----------
Section 3.1

In this section it is stated that the assumption of "a Gaussian likelihood and
sampling over the posterior probability distributions" is made. This assumption
needs further justification given the variety of data sources included. In
particular, it is later stated in Section 3.1 that the cause of the high
reduced chi2 values "must be some additional signal or noise". Alternatively,
it could indeed mean that the Gaussian likelihood assumption is not completely
valid.

RESPONSE
----------
> TODO

----------
********************************************************************************
REFEREE COMMENT
----------
Section 3.3

The discussion and figure in this section are very good. However, it is not
sufficiently demonstrated that a stellar companion is ruled out when
considering the possible angular separations of the companion resulting from
being near inferior or superior conjunction. This is particularly important for
this system since even a small degree of coplanarity of the companion with
WASP-4b would result in an edge on orbit that makes the companion difficult to
detect when near conjunction locations. Also, the very last sentence of this
section "Figure 4 shows the result." should describe the results of the figure
in more detail.

RESPONSE
----------
> TODO
>
> The comment regarding the last sentence has been addressed.

********************************************************************************
REFEREE COMMENT
----------
Section 4.1

Paragraph 1. The paragraph implies that the Doppler effect is the sole cause of
the timing derivative. However, and given the 2 sigma discrepancy, the wording
should allow for the possibility that the timing derivative is actually a
result of a combination of the described effects.

RESPONSE
----------
> We have modified the wording in this paragraph to clarify that this is a
> possibility, but that at present the data are best explained

********************************************************************************
REFEREE COMMENT
----------
Section 4.2

Paragraph 3. "hot Jupiters is expected" should be "hot Jupiters are expected".

RESPONSE
----------
> Corrected, thank you.

### 2020-03-16 paper/ms.tex
[WIP] second referee report

@@
 
-\documentclass[12pt,twocolumn,tighten]{aastex62}
+%\documentclass[12pt,twocolumn,tighten]{aastex62}
+\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 reported recently by \citet{baluev_2019}, based on a homogeneous
-analysis of archival ground-based observations.  We included twelve of
-their ``high quality'' transit times from the TRAPPIST telescope (six
-transits), the El~Sauce 36$\,$cm (four transits), and
-\citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
-the original observers that the timestamps were amenable to the
-appropriate barycentric and leap second corrections (M.~Gillon,
-P.~Evans, priv{.} comm{.}).  We omitted the fourteen remaining
+analysis of archival ground-based observations.  We included
+\deleted{twelve of} their \deleted{``high quality''} transit times
+from the TRAPPIST telescope (six transits), the El~Sauce 36$\,$cm
+(four transits), and \citet{petrucci_no_2013} \added{(two transits)}.
+For TRAPPIST and El~Sauce, we verified with the original observers
+that the timestamps were amenable to the appropriate barycentric and
+leap second corrections (M.~Gillon, P.~Evans, priv{.} comm{.}).  We
+omitted the fourteen remaining
 ETD\footnote{\url{http://var2.astro.cz/ETD}} times from
@@
 period decrease reported by \citetalias{bouma_wasp4b_2019}
-($- 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}$), presumably because of the new
+($- 12.6 \pm 1.2~{\rm msec}\,{\rm yr}^{-1}$), presumably because of the new
 data from \citeauthor{southworth_transit_2019} and
@@
 derive a probability distribution in mass and semi-major axis.
-Figure~\ref{fig:mass_sma} shows the result.
+Figure~\ref{fig:mass_sma} shows the result\replaced{.}{: the companion
+responsible for the acceleration has a true mass of 10-200$\,M_{\rm
+Jup}$ and an orbital distance of $10$-$100\,{\rm AU}$.}
 
@@
 measurements strongly indicate that the least exotic option---the
-Doppler effect---is the most likely explanation.  Long-term transit
-timing data show the orbital period to be decreasing by $- 8.64 \pm
-1.26~{\rm ms}\,{\rm yr}^{-1}$.  The long-term radial velocity data
-show a trend that should lead to an apparent period derivative of
-$-5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}$ through the Doppler effect.
-Although these two measurements of the period derivative are
-discrepant by about 2-$\sigma$, Occam's razor suggests that most or
-all of the apparent decrease of WASP-4b's orbital period is caused by
-the line-of-sight acceleration.
+Doppler effect---is the \replaced{most likely explanation}{dominant
+physical process}.  Long-term transit timing data show the orbital
+period to be decreasing by $- 8.64 \pm 1.26~{\rm msec}\,{\rm
+yr}^{-1}$.  The long-term radial velocity data show a trend that
+should lead to an apparent period derivative of $-5.94 \pm 0.39~{\rm
+msec}\,{\rm yr}^{-1}$ through the Doppler effect.  Although these two
+measurements of the period derivative are discrepant by about
+2-$\sigma$, Occam's razor \replaced{suggests that most or all of}{would suggest that} the
+apparent decrease of WASP-4b's orbital period is caused
+\added{solely} by the line-of-sight acceleration.\added{ Detection
+of any additional second-order effects will require more precise data,
+and a more significant discrepancy.}
 
@@
 \citet{knutson_friends_2014} show a non-zero radial velocity trend.
-Therefore, around 1 in 3 hot Jupiters is expected to show period
-changes comparable to that of WASP-4 due to acceleration by outer
-companions.  The sign of the accelerations should be random, with
-about half of them approaching and half of them receding.  With a
+Therefore, around 1 in 3 hot Jupiters \replaced{is}{are} expected to
+show period changes comparable to that of WASP-4 due to acceleration
+by outer companions.  The sign of the accelerations should be random,
+with about half of them approaching and half of them receding.  With a
 large enough sample of short-period hot Jupiters, one might be able to
@@
 
+\listofchanges
+
 \end{document}

### 2020-03-19 paper/ms.tex
update paper to describe better AO constraints

@@
 relatively poor seeing (1.2$''$) and also on the night of September
-28-29.  On each night, we acquired three sets of $1000\times
-60$$\,$msec exposures.  If a companion is present, the autocorrelation
-functions of these speckle images would reveal a characteristic
-interference pattern. This pattern is then used to determine the
-properties of the detected companion and produce a reconstructed
-image.  Using the reconstructed speckle images, contrast curves are
-produced to determine the 5-$\sigma$ detection limits (see
+28-29\added{, 2019}.  On each night, we acquired three sets of
+$1000\times 60$$\,$msec exposures.  If a companion is present, the
+autocorrelation functions of these speckle images would reveal a
+characteristic interference pattern. This pattern is then used to
+determine the properties of the detected companion and produce a
+reconstructed image.  Using the reconstructed speckle images, contrast
+curves are produced to determine the 5-$\sigma$ detection limits (see
 \citealt{howell_speckle_2011}).  No companions were detected. The data
@@
 We fitted each model by assuming a Gaussian likelihood and sampling
-over the posterior probability distributions.  We sampled the
-posterior using the algorithm proposed by
+over the posterior probability distributions.  
+\added{The transit timing measurements, uncertainties, and provenances
+are given in Table~1.}
+We sampled the posterior using the algorithm proposed by
 \citet{goodman_ensemble_2010} and implemented by
@@
 (sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening of
-$\chi^2_{\rm red}$ probably reflects underestimated measurement
+$\chi^2_{\rm red}$ \replaced{probably reflects}{could reflect} underestimated \replaced{measurement}{statistical}
 uncertainties in at least some of the newly reported transit times,
-due to time correlated noise in the photometry. It might also be due
-to errors or misunderstandings of the time systems in which the data
-were recorded, despite our best efforts to interpret the literature.
-In what follows, we opted to enlarge the uncertainties in the
-best-fitting parameters of each model by a factor of $(\chi^2_{\rm
-red})^{1/2}$.  This was a factor of $\approx$1.73$\times$ for the
-linear model, and $\approx$1.41$\times$ for the quadratic model.
+\added{perhaps} due to time correlated noise in the photometry. It
+might also be due to \replaced{errors or}{systematic} misunderstandings of the time systems
+in which the data were recorded, despite our best efforts to interpret
+the literature. \added{In particular, mistaken leap-second corrections can introduce systematic errors of order one minute.}  \replaced{In what follows,}{Since we have no way of
+identifying which observations are affected by these issues,} we opted to
+\added{uniformly} enlarge the
+uncertainties in the best-fitting parameters of each model by a factor
+of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of
+$\approx$1.73$\times$ for the linear model, and $\approx$1.41$\times$
+for the quadratic model.
 
@@
 \subsection{Radial velocities}
+\label{subsec:rvs}
 
@@
 
-\paragraph{Speckle imaging constraints}
+%\paragraph{Speckle imaging constraints}
+% \replaced{\paragraph{Speckle imaging constraints}}{
+\paragraph{Speckle imaging transformations}
 
@@
 (Figure~\ref{fig:zorro}) to upper limits on the mass of a companion as
-a function of separation. To do this, we followed
+a function of \added{projected} separation. To do this, we followed
 \citet{montet_trends_2014}, and opted to employ the
@@
 \citetalias{bouma_wasp4b_2019}, we derived the transformation from
-contrast ratio to companion mass.  The resulting limits are shown in
-Figure~\ref{fig:mass_sma}.
+contrast ratio to companion mass.  The resulting
+limits \added{derived if we assume maximal projected separations} are
+\replaced{shown}{shown with the blue line} in Figure~\ref{fig:mass_sma}.
+\added{We describe our procedure for incorporating projection effects
+in the following paragraphs.}
 
-\paragraph{Radial velocity constraints}
-
+%\replaced{\paragraph{Radial velocity constraints}}
+\paragraph{Combined radial velocity and imaging constraints}
 To derive constraints on possible companion masses and separations
-from the radial velocities, we mostly followed the procedure of
+\deleted{from the radial velocities}, we mostly followed the procedure of
 \citet{bryan_excess_2019}. We began by defining a $128\times128$ grid
@@
 
-For each simulated companion, we drew a set of model parameters from
-the converged posterior probability samples of our model for the
-WASP-4 radial velocities.  We subtracted the orbital component of the
+\explain{Modified wording for clarity.}
+\replaced{For each simulated companion, we drew a set of model
+parameters from the converged posterior probability samples of our
+model for the WASP-4 radial velocities.}{The orbital properties of the
+inner hot Jupiter were assigned for each simulated system by sampling
+from from the radial velocity posterior derived in
+Section~\ref{subsec:rvs}.}  We subtracted the orbital component of the
 model from the observed RVs, leaving behind the RV residuals with a
@@
 probabilities, and averaged over the samples in each grid cell to
-derive a probability distribution in mass and semi-major axis.
-Figure~\ref{fig:mass_sma} shows the result\replaced{.}{: the companion
+derive a probability distribution in mass and semi-major axis, based
+on the radial velocities.
+
+\added{We incorporated the speckle imaging limits in each simulated
+system as follows. After fitting for the time and argument of
+periastron, all the orbital parameters needed to find the projected
+separation are known (we assumed uniform sensitivity as a function of
+position angle, and therefore fixed the longitude of the
+ascending node to zero). We calculated the projected separation using
+the parametrization given by \citet{quirrenbach_2010}, which required
+solving Kepler's equation for each simulated system at the time of the
+imaging observation (September 28, 2019).   For each cell, we then
+calculated the fraction of companions whose masses and
+projected separations would have enabled detection.  }
+
+\replaced{Figure~\ref{fig:mass_sma} shows the result.}
+{The contours in Figure~\ref{fig:mass_sma} show the result of multiplying the radial
+velocity and speckle imaging probability distributions. The companion
 responsible for the acceleration has a true mass of 10-200$\,M_{\rm
 Jup}$ and an orbital distance of $10$-$100\,{\rm AU}$.}
+%FIXME: might need to update this text!
 
@@
 \added{solely} by the line-of-sight acceleration.\added{ Detection
-of any additional second-order effects will require more precise data,
+of additional second-order effects will require more precise data,
 and a more significant discrepancy.}

### 2020-03-21 paper/ms.tex
updated mass/sma constraints in paper

@@
   acceleration of the system.  This acceleration is probably caused by
-  a wide-orbiting companion of mass 10-200$\,M_{\rm Jup}$ and orbital
+  a wide-orbiting companion of mass 10-300$\,M_{\rm Jup}$ and orbital
   distance 10-100$\,$AU, based on the magnitude of the radial-velocity
@@
 
+\explain{We learned about this after submission.}
+\subsection{Astrometry}
+
+\added{
+In review, it came to our attention that observations by the Gaia satellite...
+%FIXME
+}
+
 \section{Analysis}
@@
     requirements of both the radial velocities and the speckle
-			imaging.} The likelihood inferred from radial velocities is shown
-		in grayscale, and the region excluded from the speckle imaging is
-		indicated.  The black line illustrates the expected degeneracy
-		between mass and semimajor-axis.
-		\label{fig:mass_sma}
+    imaging.} Contours show the joint radial velocity and speckle
+    imaging probability density with 1, 2, and 3-$\sigma$
+    significance. The dotted line shows the speckle imaging limits at
+    maximal projected separation.  Relevant projection effects were
+    marginalized out when calculating the contours.  The horizontal
+    dashed line is the mass limit inferred from observations that
+    WASP-4 is single-lined, requiring any companion to contribute no
+    more than one-tenth of the observed light.  The black line shows
+    the expected degeneracy between mass and semimajor-axis.
+    \explain{Caption was modified to account for updated figure. The
+    updated upper limit on the mass is now roughly 300$M_{\rm Jup}$,
+    instead of 200$M_{\rm Jup}$. The text has been modified
+    accordingly.} \label{fig:mass_sma}
 		\vspace{-0cm}
@@
 (sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening of
-$\chi^2_{\rm red}$ \replaced{probably reflects}{could reflect} underestimated \replaced{measurement}{statistical}
-uncertainties in at least some of the newly reported transit times,
-\added{perhaps} due to time correlated noise in the photometry. It
-might also be due to \replaced{errors or}{systematic} misunderstandings of the time systems
+$\chi^2_{\rm red}$ \replaced{probably reflects}{could reflect}
+underestimated \replaced{measurement}{statistical} uncertainties in at
+least some of the newly reported transit times, \added{perhaps} due to
+time correlated noise in the photometry. It might also be due to
+\replaced{errors or}{systematic} misunderstandings of the time systems
 in which the data were recorded, despite our best efforts to interpret
-the literature. \added{In particular, mistaken leap-second corrections can introduce systematic errors of order one minute.}  \replaced{In what follows,}{Since we have no way of
-identifying which observations are affected by these issues,} we opted to
-\added{uniformly} enlarge the
-uncertainties in the best-fitting parameters of each model by a factor
-of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of
+the literature. \added{In particular, mistaken leap-second corrections
+can introduce systematic errors of order one minute.}  \replaced{In
+what follows,}{Since we have no way of identifying which observations
+are affected by these issues,} we opted to \added{uniformly} enlarge
+the uncertainties in the best-fitting parameters of each model by a
+factor of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of
 $\approx$1.73$\times$ for the linear model, and $\approx$1.41$\times$
@@
 ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate of
-period decrease reported by \citetalias{bouma_wasp4b_2019}
-($- 12.6 \pm 1.2~{\rm msec}\,{\rm yr}^{-1}$), presumably because of the new
+period decrease reported by \citetalias{bouma_wasp4b_2019} ($- 12.6
+\pm 1.2~{\rm msec}\,{\rm yr}^{-1}$), presumably because of the new
 data from \citeauthor{southworth_transit_2019} and
@@
 First, we converted the contrast curves obtained from speckle imaging
-(Figure~\ref{fig:zorro}) to upper limits on the mass of a companion as
+(Figure~\ref{fig:zorro}) to upper limits on the companion mass as
 a function of \added{projected} separation. To do this, we followed
@@
 limits \added{derived if we assume maximal projected separations} are
-\replaced{shown}{shown with the blue line} in Figure~\ref{fig:mass_sma}.
-\added{We describe our procedure for incorporating projection effects
-in the following paragraphs.}
+\replaced{shown}{shown with the dotted line} in Figure~\ref{fig:mass_sma}.
+\added{
+However, because we are observing the primary star accelerate towards
+our line of sight, the companion very well be near
+inferior conjunction.
+We incorporate the relevant projection effects
+as described in the following paragraphs.}
 
@@
 jitters, we performed a maximum likelihood fit for the time and
-argument of periastron of the outer simulated companion.  We converted
+argument of periastron of the outer simulated companion.  \deleted{We converted
 the resulting $128\times128\times512$ cube of log-likelihood values to
@@
 derive a probability distribution in mass and semi-major axis, based
-on the radial velocities.
+on the radial velocities.}
 
-\added{We incorporated the speckle imaging limits in each simulated
+\added{We then incorporated the speckle imaging limits in each simulated
 system as follows. After fitting for the time and argument of
@@
 solving Kepler's equation for each simulated system at the time of the
-imaging observation (September 28, 2019).   For each cell, we then
-calculated the fraction of companions whose masses and
-projected separations would have enabled detection.  }
-
+imaging observation (September 28, 2019).  If a simulated companion's
+mass and projected separation put it above the 5$\sigma$ contrast curve,
+we assumed it would have been detected.
+We multiplied
+the resulting $128\times128\times512$ cubes of radial velocity and speckle
+imaging probabilities, and marginalized over the systems in each grid cell to
+derive a probability distribution in mass and semi-major axis.
+}
 \replaced{Figure~\ref{fig:mass_sma} shows the result.}
-{The contours in Figure~\ref{fig:mass_sma} show the result of multiplying the radial
-velocity and speckle imaging probability distributions. The companion
-responsible for the acceleration has a true mass of 10-200$\,M_{\rm
+{The contours in Figure~\ref{fig:mass_sma} show the result: the companion
+responsible for the acceleration has a true mass of 10-300$\,M_{\rm
 Jup}$ and an orbital distance of $10$-$100\,{\rm AU}$.}
@@
 measurements of the period derivative are discrepant by about
-2-$\sigma$, Occam's razor \replaced{suggests that most or all of}{would suggest that} the
-apparent decrease of WASP-4b's orbital period is caused
-\added{solely} by the line-of-sight acceleration.\added{ Detection
-of additional second-order effects will require more precise data,
-and a more significant discrepancy.}
+2-$\sigma$, Occam's razor \replaced{suggests that most or all
+of}{would suggest that} the apparent decrease of WASP-4b's orbital
+period is caused \added{solely} by the line-of-sight
+acceleration.\added{ Detection of additional second-order effects will
+require more precise data, and a more significant discrepancy.}
 
@@
 synthesized work by \citet{wang_influence_2014,wang_influence_2015},
-\citet{ngo_2015_fohj2}, \citet{ngo_friends_2016}, \citet{kraus_impact_2016},
-\citet{matson_stellar_2018}, \citet{ziegler_soar_2020}, and others.
-A handful of systems with inner hot Jupiters and outer brown dwarfs within
-100$\,$AU are known, for instance CoRoT-20, HATS-59, WASP-53,
-and WASP-81
+\citet{ngo_2015_fohj2}, \citet{ngo_friends_2016},
+\citet{kraus_impact_2016}, \citet{matson_stellar_2018},
+\citet{ziegler_soar_2020}, and others.  A handful of systems with
+inner hot Jupiters and outer brown dwarfs within 100$\,$AU are known,
+for instance CoRoT-20, HATS-59, WASP-53, and WASP-81
 \citep{triaud_peculiar_2017,rey_brown_2018,sarkis_hats-59bc_2018}.  In
-contrast, from surveys by
-\citet{knutson_friends_2014}, \citet{ngo_2015_fohj2} and
-\citet{mugrauer_search_2019} we could find only one example of a
-system with an inner hot Jupiter and an outer low-mass star within
-100$\,$AU: HAT-P-10 \citep{bakos_2009_hatp10}.  Overall, this body of
-literature indicates that the presence of a stellar-mass companion
-within 100$\,$AU of the host star could hinder the formation of
-planetary systems.  If true, then one would expect that the outer
-companion in WASP-4 would have the lowest mass allowed by the data.
+contrast, from surveys by \citet{knutson_friends_2014},
+\citet{ngo_2015_fohj2} and \citet{mugrauer_search_2019} we could find
+only one example of a system with an inner hot Jupiter and an outer
+low-mass star within 100$\,$AU: HAT-P-10 \citep{bakos_2009_hatp10}.
+Overall, this body of literature indicates that the presence of a
+stellar-mass companion within 100$\,$AU of the host star could hinder
+the formation of planetary systems.  If true, then one would expect
+that the outer companion in WASP-4 would have the lowest mass allowed
+by the data.
 

### 2020-03-21 paper/collect.sh

@@ -1,5 +1,5 @@
 cp ../results/O_minus_C_20200127_onlytransits.pdf f1.pdf
 cp ../results/rv_fitting/LGB_20190911_fix_gammaddot_paperII/WASP4_rv_multipanel.pdf f2.pdf
 cp ../results/zorro_speckle.pdf f3.pdf
-cp ../results/mass_semimaj_constraints_with_contrast_discretecolor.pdf f4.pdf
+cp ../results/mass_semimaj_contours_with_contrast_discretecolor.pdf f4.pdf
 cp ../results/k14_pdot.pdf f5.pdf

### 2020-03-21 paper/ms.tex
oops. RUWE != UWE
The UWE is 1.4

But Fig A.1 of Lindegrend+2018 clarifies that this is not significant.
minor word changes

@@
 
-%\documentclass[12pt,twocolumn,tighten]{aastex62}
-\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
+\documentclass[12pt,twocolumn,tighten]{aastex62}
+%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 %% Add "Submitted to " argument.
-\received{\today}
-\revised{---}
+\received{February 28, 2020}
+\revised{\today}
 \accepted{---}
@@
 
-\explain{We learned about this after submission.}
-\subsection{Astrometry}
-
-\added{
-In review, it came to our attention that observations by the Gaia satellite...
-%FIXME
-}
-
 \section{Analysis}
@@
 over the posterior probability distributions.  
-\added{The transit timing measurements, uncertainties, and provenances
+\added{The timing measurements, uncertainties, and provenances
 are given in Table~1.}
@@
 \added{
-However, because we are observing the primary star accelerate towards
-our line of sight, the companion very well be near
+However, because the primary star is accelerating towards
+our line of sight, the companion could very well be near
 inferior conjunction.
-We incorporate the relevant projection effects
-as described in the following paragraphs.}
+Our approach for incorporating the relevant projection effects
+is described in the following paragraphs.}
 
@@
 jitters, we performed a maximum likelihood fit for the time and
-argument of periastron of the outer simulated companion.  \deleted{We converted
-the resulting $128\times128\times512$ cube of log-likelihood values to
-probabilities, and averaged over the samples in each grid cell to
-derive a probability distribution in mass and semi-major axis, based
-on the radial velocities.}
+argument of periastron of the outer simulated companion.  \deleted{We
+converted the resulting $128\times128\times512$ cube of log-likelihood
+values to probabilities, and averaged over the samples in each grid
+cell to derive a probability distribution in mass and semi-major axis,
+based on the radial velocities.}
 
-\added{We then incorporated the speckle imaging limits in each simulated
-system as follows. After fitting for the time and argument of
-periastron, all the orbital parameters needed to find the projected
-separation are known (we assumed uniform sensitivity as a function of
-position angle, and therefore fixed the longitude of the
-ascending node to zero). We calculated the projected separation using
-the parametrization given by \citet{quirrenbach_2010}, which required
-solving Kepler's equation for each simulated system at the time of the
-imaging observation (September 28, 2019).  If a simulated companion's
-mass and projected separation put it above the 5$\sigma$ contrast curve,
-we assumed it would have been detected.
-We multiplied
-the resulting $128\times128\times512$ cubes of radial velocity and speckle
-imaging probabilities, and marginalized over the systems in each grid cell to
+\added{We then incorporated the speckle imaging limits in each
+simulated system as follows. After fitting for the time and argument
+of periastron, all the orbital parameters needed to find the projected
+separation at the time of observation are known. We assumed uniform
+sensitivity as a function of position angle, and therefore fixed the
+longitude of the ascending node to zero. We then calculated the
+projected separation using the parametrization given by
+\citet{quirrenbach_2010}.  If a simulated companion's mass and
+projected separation put it above the 5$\sigma$ contrast curve, we
+assumed it would have been detected.  We multiplied the resulting
+$128\times128\times512$ cubes of radial velocity and speckle imaging
+probabilities, and marginalized over the systems in each grid cell to
 derive a probability distribution in mass and semi-major axis.

### 2020-03-21 second_paper_review/response_20200320.txt
add referee response

Dear Dr. Jackson,

We thank you for organizing the review of our paper, and also
thank the referee for their helpful comments.  We have
considered the referee's remarks carefully and revised our
manuscript accordingly.

Below, we have reproduced the relevant portions of the
referee's report, alongside our responses. A list of changes is
appended to the manuscript using the trackchanges AASTeX
macros.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
Section 2.1, Paragraph 2.

It would be helpful if the authors briefly describe the exact
meaning of "high quality" in the context of the transit times
used in the analysis.

RESPONSE
----------
> Thank you for this prompt. We have removed the phrase "high
> quality", as its precise definition is not of particular
> importance for this study. It is defined in Section 3.4 of
> Baluev+ 2019, through an empirical cutoff that Baluev+ made
> based on the sampling cadence and the light-curve residual
> RMS.

******************************************
REFEREE COMMENT
----------
Section 3.1

In this section it is stated that the assumption of "a Gaussian
likelihood and sampling over the posterior probability
distributions" is made. This assumption needs further
justification given the variety of data sources included. In
particular, it is later stated in Section 3.1 that the cause of
the high reduced chi2 values "must be some additional signal or
noise". Alternatively, it could indeed mean that the Gaussian
likelihood assumption is not completely valid.

RESPONSE
----------
> We expect that the high reduced chi2 values are likely due to
> either underestimated statistical uncertainties on transit
> timing measurements, or that they are caused by systematic
> errors.  We do not think that the specific choice of a
> Gaussian likelihood vs. an alternate choice of likelihood is
> necessarily at issue.
>
> We made every effort to ensure the integrity of our included
> times, for instance by rejecting times that did not clearly
> specify the nature of the barycentric/heliocentric
> correction, or the nature of the leap-second correction (UTC
> vs TDB). However we ultimately need to trust timestamps that
> were recorded by observers other than ourselves.  Despite
> these efforts, some observations might have systematic errors
> (e.g., of order 1 minute for leap-second corrections), but we
> cannot be sure which ones.
>
> Our approach for dealing with this issue was to introduce a
> free parameter that uniformly inflates the uncertainties of
> derived parameters (e.g., dP/dt) by the factor that would
> ensure reduced chi2 = 1.  Any suggestions on better
> methodologies are welcome, however it is not clear to us that
> there is any way around the fundamental issue of needing to
> trust that observers can record time correctly.
>
> We have updated the wording in the manuscript to improve the
> clarity with which we present these issues.

******************************************
REFEREE COMMENT
----------
Section 3.3

The discussion and figure in this section are very good.
However, it is not sufficiently demonstrated that a stellar
companion is ruled out when considering the possible angular
separations of the companion resulting from being near inferior
or superior conjunction. This is particularly important for
this system since even a small degree of coplanarity of the
companion with WASP-4b would result in an edge on orbit that
makes the companion difficult to detect when near conjunction
locations. Also, the very last sentence of this section "Figure
4 shows the result." should describe the results of the figure
in more detail.

RESPONSE
----------
> We agree; we erroneously neglected this issue in the
> submitted manuscript.  Thank you for bringing it to our
> attention.
>
> In the revised manuscript, we amended our methodogy to
> probabilistically include the speckle imaging limits. The
> procedure is discussed in the text, but in brief, for each
> simulated companion (which has known orbital parameters), we
> calculated its projected separation on-sky at the time of the
> imaging observation. We then multiply the RV and speckle
> imaging posteriors, yielding contours in a revised Figure 4.
>
> This revision led to slightly weaker upper limits on the
> companion mass (200Mjup -> 300Mjup), but did not otherwise
> affect the primary conclusions of the study.
>
> The comment regarding the last sentence has also been
> addressed; thank you.

******************************************
REFEREE COMMENT
----------
Section 4.1

Paragraph 1. The paragraph implies that the Doppler effect is
the sole cause of the timing derivative. However, and given the
2 sigma discrepancy, the wording should allow for the
possibility that the timing derivative is actually a result of
a combination of the described effects.

RESPONSE
----------
> We have modified the wording in this paragraph to clarify
> that this is a possibility, but that at present the dominant
> physical effect is the line-of-sight acceleration.

******************************************
REFEREE COMMENT
----------
Section 4.2

Paragraph 3. "hot Jupiters is expected" should be "hot Jupiters
are expected".

RESPONSE
----------
> Corrected, thank you.

### 2020-03-25 second_paper_review/response_20200320.txt
incorporate JNW referee report edits

@@
 ----------
-> We expect that the high reduced chi2 values are likely due to
-> either underestimated statistical uncertainties on transit
-> timing measurements, or that they are caused by systematic
-> errors.  We do not think that the specific choice of a
+> We expect that the high reduced chi2 values are due to
+> underestimated statistical uncertainties on transit
+> timing measurements, or systematic errors due to stellar
+> activity or mistakes in the conversion to the standard
+> time system.  We do not think that the specific choice of a
 > Gaussian likelihood vs. an alternate choice of likelihood is
@@
 > derived parameters (e.g., dP/dt) by the factor that would
-> ensure reduced chi2 = 1.  Any suggestions on better
-> methodologies are welcome, however it is not clear to us that
-> there is any way around the fundamental issue of needing to
-> trust that observers can record time correctly.
+> ensure reduced chi2 = 1.  This seems like the simplest and most
+> reasonable way to handle this heterogeneous dataset.
 >
@@
 >
-> In the revised manuscript, we amended our methodogy to
+> In the revised manuscript, we amended our methodology to
 > probabilistically include the speckle imaging limits. The

### 2020-03-26 paper/ms.tex
fix sentence

@@
 underestimated \replaced{measurement}{statistical} uncertainties in at
-least some of the newly reported transit times, \added{perhaps} due to
-time correlated noise in the photometry. It might also be due to
+least some of the newly reported transit times\deleted{, due to
+time correlated noise in the photometry}. It might also be due to
 \replaced{errors or}{systematic} misunderstandings of the time systems

### 2020-03-31 src/format_WASP-4b_transit_time_table.py
ApJL accepted
add MRTs
fix dates

@@ -1,0 +1,29 @@
+import pandas as pd
+
+df = pd.read_csv(
+    '../paper/MRTs/WASP-4b_literature_and_TESS_times_O-C_vs_epoch_20200127_selected.csv',
+    sep = ';'
+)
+
+scols = [
+    'sel_transit_times_BJD_TDB',
+    'err_sel_transit_times_BJD_TDB',
+    'sel_epoch',
+    'original_reference',
+    'where_I_got_time'
+]
+
+df = df[scols]
+
+df['sel_epoch'] = df['sel_epoch'].astype(int)
+
+rdict = {
+    'Baluev2019supp':'2019MNRAS.490.1294B',
+    'measured_from_SPOC_alert_LC':'2019AJ....157..217B',
+    'me':'2019AJ....157..217B',
+}
+for k,v in rdict.items():
+    for c in ['where_I_got_time','original_reference']:
+        df[c] = df[c].replace(to_replace=k, value=v)
+
+df.to_csv('../paper/MRTs/transit_data.csv', index=False, float_format='%.6f')

### 2020-03-31 src/make_WASP-4b_rv_table.py

@@ -28,15 +28,15 @@ def main():
     references = []
     for ref in original_references:
         if ref == 'Triaud+2010_vizier':
-            references.append('\citet{triaud_spin-orbit_2010}')
+            references.append('T+10')
         elif ref == 'Pont+2011_vizierMNRAS':
-            references.append('\citet{pont_determining_2011}')
+            references.append('P+11')
         elif ref == 'jump_20190911':
-            references.append('This work')
+            references.append('B+20')
     references = np.array(references)
 
     outdf = pd.DataFrame(
-        {'time': np.round(nparr(df.time),8),
+        {'time': np.round(nparr(df.time-2450000),6),
          'mnvel': np.round(nparr(df.mnvel),5),
          'errvel': np.round(nparr(df.errvel),5),
          'svalue': svalue,
@@ -47,10 +47,11 @@ def main():
 
     outpath = '../paper/WASP-4b_rv_data.tex'
     with open(outpath,'w') as tf:
-        tf.write(outdf.to_latex(index=False, escape=False))
+        tf.write(outdf.to_latex(index=False, escape=False, float_format='%.6f'))
     print('wrote {:s}'.format(outpath))
 
-    outdf.to_csv('../paper/WASP-4b_rv_data.csv', index=False)
+    outdf.to_csv('../paper/WASP-4b_rv_data.csv', index=False,
+                 float_format='%.6f')
     print('also wrote ../paper/WASP-4b_rv_data.csv')
 
 
@@ -65,6 +66,7 @@ def main():
     bottomrule_ix = [ix for ix,l in enumerate(lines)
                      if r'\bottomrule' in l][0]
     sel_lines = lines[midrule_ix+1 : bottomrule_ix]
+    sel_lines = [l.replace(' \\\\', '') for l in sel_lines]
 
     with open(outpath, 'w') as f:
         f.writelines(sel_lines)
